### 2001-01-01 paper.tex
fixup to wind model description

\documentclass[aps,floatfix,prd,superscriptaddress,twocolumn]{revtex4-1}
% Intended to be built with pdflatex, using the REVTEX 4.1 package.
% Template taken from FermiLab APS Journal template.

% packages
\usepackage{amsmath}              % for \text, align, split ...
\usepackage{amssymb}              % for \gtrsim ...
\usepackage{braket}               % for \bra \ket
\usepackage[mathscr]{euscript}    % for \mathscr
\usepackage{graphicx}             % for \includegraphics
\usepackage{hyperref}             % for \url and \href
\usepackage{listings}             % for typesetting code with \lstinline
\usepackage{tabularx}             % for tabularx

% commands/initialization
\hyphenation{ALPGEN} % avoid incorrect hyphenation
\hyphenation{EVTGEN} % ''
\hyphenation{PYTHIA} % ''
\lstset{basicstyle=\ttfamily} % define \lstinline for inline code snippet

% *******************************************************************************
\begin{document}

\widetext
\leftline{Compiled \today}
\leftline{To be submitted to PRD}

\title{Elastic Scattering in General Relativistic Ray Tracing for Neutrinos}

\author{M.\ Brett Deaton}
\email{mbdeaton@ncsu.edu}
\affiliation{Joint Institute for Nuclear Astrophysics,
  Michigan State University, East Lansing, MI 48824, USA}
\affiliation{Department of Physics,
  North Carolina State University, Raleigh, NC 27695, USA}

\author{Evan O'Connor}
\affiliation{Department of Physics,
  North Carolina State University, Raleigh, NC 27695, USA}

\author{Y.\ L.\ Zhu}
\affiliation{Department of Physics,
  North Carolina State University, Raleigh, NC 27695, USA}

\author{Andy Bohn}
\affiliation{Center for Radiophysics and Space Research,
  Cornell University, Ithaca, NY 14853, USA}

\author{Jerred Jesse}
\affiliation{Department of Physics \& Astronomy,
  Washington State University, Pullman, WA 99164, USA}

\author{Francois Foucart}
\affiliation{Lawrence Berkeley National Laboratory,
  1 Cyclotron Rd., Berkeley, CA 94720, USA}

\author{Matthew D.\ Duez}
\affiliation{Department of Physics \& Astronomy,
  Washington State University, Pullman, WA 99164, USA}

\author{G.\ C.\ McLaughlin}
\affiliation{Department of Physics,
  North Carolina State University, Raleigh, NC 27695, USA}

% *******************************************************************************

\begin{abstract}
  We present a covariant ray tracing algorithm for computing high-resolution
  neutrino distributions in general relativistic numerical spacetimes with
  hydrodynamical sources.
  Our formulation treats the very important effect of
  elastic scattering of neutrinos off of nuclei and nucleons
  (changing the neutrino's direction but not energy)
  by incorporating estimates of the background neutrino fields.
  These background fields may be taken from a low-order moment simulation
  or be ignored, in which case the method
  reduces to a standard state-of-the-art ray tracing formulation.
  It handles radiation in regimes spanning optically thick to optically thin.
  We test the new code, highlight its strengths and weaknesses,
  and apply it to a simulation of a neutron star merger
  to compute neutrino fluxes and spectra,
  and to demonstrate a neutrino flavor oscillation calculation.
\end{abstract}

\maketitle

\section{Introduction}
Neutrinos are one of the dominant energy transport phenomena at play in
neutron star mergers: heating, cooling, and pushing the disrupted nuclear
matter.
In addition, they change the composition of the matter via charged current
interactions.
Because neutrinos scatter over length scales both large and small with
respect to fluid scales, accurate models require a neutrino treatment that
respects the freedom of neutrino distribution functions to vary
drastically from geometrically simple distributions in thermodynamic equilibrium.

This is a challenging task in the environment of a merger,
which generally lacks any spatial symmetries, so that fully general
seven-dimensional solutions to the Boltzmann Equation are not feasible.
Leakage approximations
\citep{ruff1996-leakage_part1,ross2003-leakage_part2,
  deat2013-leakage, pere2016-asl, radi2016-dynamical}
\footnote{TODO: cite others}
capture some of the qualitative effects of neutrinos on the matter, but
provide extremely limited information about the neutrino field itself.
\footnote{TODO: discuss mc more clearly}
Monte Carlo methods like those
used in supernova simulations \cite{abdi2012-monte_carlo}
and stationary models of accretion disks \cite{rich2015-monte_carlo}
are an excellent tool but require large computational resources
because of their global treatment of the problem.

The state of the art today for neutron star merger calculations
couples radiation and matter using a truncated moment formalism
\citep{fouc2015-m1_nsbh, fouc2016-m1_nsns},
evolving the zeroth- and first-angular moments of the energy density,
that is the total energy- and momentum-densities.
This method, commonly called M1 transport, was formulated by
\cite{thor1981-truncated_moment} and modernized by
\cite{shib2011-truncated_moment}.
M1 transport has recently become popular in the core-collapse supernova
community as well
\cite{kuro2012-m1_ccsne,just2015-m1_code,ocon2015-gr1d_with_nu,
  skin2016-ray_vs_m1,robe2016-gr_m1_ccsne}.

For merger simulations, \cite{fouc2016-m1_evolve_n} have recently
expanded their M1 transport code
to also evolve the zeroth-angular moment of the number density,
providing total energy densities and average energies throughout the
simulated volume.
Even so, with M1 transport codes only evolving two angular moments
and one or two energy moments,
they can extract only limited angular and spectral information
from the neutrino fields.

%\subsubsection*{What physics problems do we want to address?}
But many interesting unsolved problems require an accurate model of the neutrino
spectra and angular distributions.
With a model of the neutrino emission from a merger we can
1) examine neutrino effects on the nucleosynthesis of the ejected material
\citep{surm2011-nickel_56, robe2016-sph_nu_nucleo},
2) explore the rich flavor oscillation physics likely to occur
\citep{malk2012-mnr_1, malk2015-mnr_2, malk2016-mnr_3, zhu2016-mnr_nsns_remnant,
  vaan2016-uncovering_mnr},
3) improve closure relations used in M1 transport schemes
\citep{ramp2002-truncated_moment, shib2011-truncated_moment,
  card2013-truncated_moment, fouc2015-m1_nsbh, ocon2015-gr1d_with_nu}, and
4) study possible jet formation due to neutrino annihilation
\citep{ruff1999-nunubar_nsns, asan2000-nunubar, birk2007-nunubar,
  hari2010-gr_nunubar_collapsar, zala2011-nunubar, leng2014-nunubar,
  just2016-nsns_ejecta_obstacles}.

%\subsubsection*{Why ray tracing?}
Angular and spectral neutrino distributions in neutron star merger simulations may
be available soon, for example with the coupled Monte-Carlo-M1 scheme presented in
\citep{fouc2017-m1_with_mc}.
In this work, however, we present a ray tracing method to compute neutrino
distribution functions from the existing state-of-the-art general relativistic M1
transport hydrodynamics simulations.
We choose ray tracing because it is conceptually simple,
numerically inexpensive,
and extends to high resolution in energy and angle
by simply increasing the number of rays sampled.
Furthermore, the computational implementation parallellizes trivially.

With a ray tracing method we approach radiation transport from the perspective
of a single observer at a spacetime event $x_o^\alpha$.
Our goal is to compute the distribution function
$f^{\nu_\sigma}(x_o^\alpha;p_\beta)$, or the amount of neutrino radiation
of species
$\nu_\sigma\in\{\nu_e,\nu_\mu,\nu_\tau,\bar{\nu}_e,\bar{\nu}_\mu,\bar{\nu}_\tau\}$
with momentum $p_\beta$ impinging on $x_o^\alpha$.
To do so we trace a geodesic trajectory from $x_o^\alpha$ in the backwards
direction $-p_\beta$ to sample the incoming radiation along that line of sight.
By tracing a family of rays intersecting $x_o^\alpha$ we build up a
picture of the distribution function there.
And by sampling many observation points we construct a global picture
of $f^{\nu_\sigma}(x^\alpha;p_\beta)$.

The ray tracing framework
is conceptually simple because it solves the equation of radiation
transport (Eqn.~\ref{eqn:boltzmann}) along characteristics,
reducing it to a one-dimensional ordinary differential equation.
It is numerically cheap because it confines computations of
$f^{\nu_\sigma}(x_o^\alpha;p_\beta)$ to the past light-cone of $x_o^\alpha$,
with the history of that light-cone truncated at large optical depth.
\footnote{TODO: state scaling $\mathscr{O}(n)$?}
It easily extends to high resolution in energy and angle by simply
increasing the number of rays sampled.
And it parallelizes trivially by ray, since each ray is computed
independently.

%\subsubsection*{What is new and better about this ray tracing formulation?}
Several ray tracing formulations for radiation transport already exist.
Most formulations assume an analytic spacetime metric
\citep{birk2007-nunubar, caba2009-detecting_grb_nu,
  hari2010-gr_nunubar_collapsar, kova2011-gr_ray_tracing}.
And many make the simplifying assumption of blackbody emission from a
neutrinosurface
\citep{birk2007-nunubar, caba2009-detecting_grb_nu, kova2011-gr_ray_tracing},
limiting them to equilibrium, optically-thick configurations.
Current state-of-the-art ray tracing formulations avoid the assumption of
blackbody spectra by integrating a local emissivity along each geodesic
(e.g. \cite{hari2010-gr_nunubar_collapsar} for neutrinos and
\cite{youn2012-gr_radiative_transfer} for photons).
But no formulations to date account for the important scattering and pair
processes outlined in Tab.~\ref{tab:neutrino_processes}.

We build upon existing ray tracing formulations
by eschewing any assumptions about the spacetime geometry,
integrating local emissivities,
and including elastic scattering in the integration along each geodesic.
To incorporate scattering processes in our method, we employ estimates of the
lowest-order moment of the neutrino distribution function
computed in an M1 transport simulation.
If moments are not available either from
an M1 transport evolution or a trustworthy analytical estimate,
our method reduces to the current state-of-the-art ray tracing method,
neglecting scattering.

%\subsubsection*{Why does a covariant formulation matter?}
We formulate the ray tracing equations covariantly---free from
assumptions about the spacetime geometry or coordinates. This is essential
because we want to apply the method as a postprocessing step using
time snapshots of data computed from general relativistic evolutions.
The spacetime represented in these snapshots is not analytic (i.e. Kerr).
And even in configurations that are described well by the Kerr metric
(e.g. a low-mass disk around a massive black hole),
the evolution coordinates are unlikely to present the metric in
a familiar analytic form.
This is because integrating the Einstein Equations often requires complicated,
time-dependent gauge conditions
\citep{lind2007-gen_harmonic, fouc2013-compactness_and_spin}.

%\subsubsection*{Why do scattering and pair processes matter?}
Elastic scattering (see Tab.~\ref{tab:neutrino_processes}) can signicantly
modify neutrino distributions in angle and dilute the emitted spectrum over a
larger emitting surface \citep{pere2016-asl}.
This is especially pronounced in the case of heavy-lepton neutrinos.
\footnote{TODO: cite}
Inelastic scattering and pair processes can introduce further modifications.
Any phenomena that involve neutrino-neutrino interactions,
for example neutrino oscillations and neutrino-antineutrino annihilation,
\footnote{TODO: cite}
depend sensitively on the angular distribution.
And spectral changes can strongly affect the nucleosynthesis
occuring in the ejected and irradiated material.
\footnote{TODO: justify these statements}
The ray tracing method we present in this paper captures the dominant effects
of elastic scattering, while leaving the physics of inelastic scattering and
pair processes for later work.

%\subsubsection*{In what ways is this method limited?}
Ray tracing is ideally suited to problems requiring detailed knowledge of
radiation distribution functions over small regions of spacetime:
for example along a matter or radiation trajectory,
or over a small volume outside a source.

More fundamentally, though, ray tracing is limited by its naive
treatment of the Boltzmann Equation (Eqn.~\ref{eqn:boltzmann}),
a treatment which essentially decouples different neutrino momenta and species.
\footnote{TODO: clarify how momenta are decoupled}
When solving for $f^{\nu_\sigma}(x^\alpha;p_\beta)$,
we have no information about the distribution at different momenta
$f^{\nu_\sigma}(x^\alpha;p'_\beta)$,
or about the distribution of the relevant antineutrino species
$f^{\bar{\nu}_\sigma}(x^\alpha;p'_\beta)$;
these missing data are essential ingredients
for the source terms of the Boltzmann Equation describing the creation and
destruction of neutrinos due to the interaction processes described in
Tab.~\ref{tab:neutrino_processes}.

In this paper we outwit this limitation by incorporating coupled source terms
that depend on either previously-evolved or analytical
estimates of the neutrino fields.
We also retain the freedom to drop the coupling terms, in which case our method
becomes similar to existing state-of-the-art ray tracing methods.
Our method is not a standalone radiation transport scheme,
but serves as the final component of a hybrid scheme,
piggy-backing on a lower-order radiation transport method as a
post-processing step.

In Sec.~\ref{sec:formulation} we derive the ray tracing equations from the
Boltzmann Equation and describe our numerical scheme.
In Sec.~\ref{sec:tests} we present tests of the code.
In Sec.~\ref{sec:applications} we present an application
in the dynamical environment following the merger of two neutron stars,
computing neutrino flavor oscillation along an outgoing ray,
including the effects of coherent forward scattering with ambient neutrinos.
In Sec.~\ref{sec:conclusions} we summarize our work and anticipate improvements.

Greek tensor indices ($\alpha, \beta, ...$) range over all four coordinates,
whereas Latin indices ($i, j, ...$) range over the spatial coordinates 1--3.
We use naturalized units in which $\{\hbar,c,k_{\rm B}\}=1$.
And for most of the remainder of this article we suppress neutrino species label
$\nu_\sigma$ since the formulation is general to any species.
Where we do reference particular species we use the three categories
relevant to the energy scales of mergers, $\nu_e$, $\bar{\nu}_e$, and
$\nu_x=\left\{\nu_\mu,\bar{\nu}_\mu,\nu_\tau,\bar{\nu}_\tau\right\}$.

\section{Ray tracing formulation}
\label{sec:formulation}
The neutrino distribution function, $f(x^\alpha; p_\beta)$ is an invariant
quantity counting the number of neutrinos in a given six-volume of phase
space centered on $(x^\alpha,p_\beta)$.
The phase space volume elements are defined with respect to a fiducial
observer passing through event $x^\alpha$ with velocity $u^\alpha$:
\begin{align}
  \label{eqn:dV}
  dV & \equiv \sqrt{-\psi} \, dx \, dy \, dz \, u^t \\
  \label{eqn:dP}
  dP & \equiv \frac{1}{\sqrt{-\psi}} \, dp_x \, dp_y \, dp_z \,
  \frac{\varepsilon}{p^t},
\end{align}
where $\psi$ represents the determinant of the spacetime metric,
the index $t$ indicates the time-component of the given four-vector, and
\begin{equation}
  \label{eqn:varepsilon}
  \varepsilon \equiv -p_\mu u^\mu
\end{equation}
is the neutrino energy measured by our observer.
The number of particles in a given six-volume is
\begin{equation}
  dN=\frac{g}{(2\pi)^3}\,f\,dV\,dP,
\end{equation}
where $g$ counts the number of spin states accessible to the
particles ($g=1$ for neutrinos), and $f$ is the distribution function.
Each of $dV$, $dP$, $dN$, and $f$ are spacetime invariants
\citep{debb2009-gr_boltzmann_1, debb2009-gr_boltzmann_2, lind1966-gr_boltzmann}.

We may decompose the neutrino momentum like
\begin{equation}
  \label{eqn:def_momentum}
  p_\beta = \varepsilon (u_\beta + \ell_\beta),
\end{equation}
with $\ell_\beta$ the direction normal subject to the constraints
$u^\alpha \ell_\alpha = 0$ and $\ell^\alpha \ell_\alpha=1$.
With this decompositon we can write the arguments to the distribution function
$f(x^\alpha;\varepsilon,\ell_\beta)$.

Because $\ell_\beta$ is subject to two constraints
(normalization and orthogonality to the observer's velocity)
it has two degrees of freedom; we make this explicit by defining its
spatial cartesian components with respect to spherical polar angles
\begin{equation}
  \label{eqn:def_direction}
  \ell_\alpha \rightarrow
  q (s,\sin a \cos b,\sin a\sin b,\cos a),
\end{equation}
with $q$ and $s$ functions of $a$ and $b$.
Now our symbol for the distribution function,
$f(x^\alpha;\varepsilon,a,b)$,
makes manifest its seven independent arguments.

We also define a rotated frame
\begin{equation}
  \ell_{\alpha'}=\frac{\partial x^\alpha}{\partial x^{\alpha'}}\,\ell_\alpha \nonumber
\end{equation}
in which the cartesian components of the direction 1-form are defined
\begin{equation}
  \label{eqn:def_direction_primed}
  \ell_{\alpha'} \rightarrow
  q (s,\sin A \cos B,\sin A\sin B,\cos A),
\end{equation}
so that momenta with vanishing polar angle $A$ move outward along coordinate
radial rays.
This transformation is chosen so that for an observer far from the source,
incoming radiation will be concentrated into a narrow beam around $\cos A\approx1$,
independent of that observer's position in coordinate space.
For explicit definitions see App.~\ref{sec:definitions}.

In Minkowski spacetime, and for a stationary observer
$u_\beta\rightarrow(-1,0,0,0)$, we would have $s=0$, $q=1$,
and $\varepsilon=-p_t$. In that case $\cos A$ may be identified
\footnote{TODO: cite someone that uses $\mu$}
with the familiar forward direction cosine $\mu$.

\subsection{Boltzmann Equation}
\label{ssec:boltzmann}
Neutrino radiation obeys the relativistic Boltzmann Equation,
which, suppressing the arguments of $f$ for simplicity, is written,
\footnote{TODO: point out limit of QKEs}
\begin{equation}
  \label{eqn:boltzmann}
  \frac{d}{d\lambda}f = C[f],
\end{equation}
where $d/d\lambda$ denotes a derivative with respect to the affine
parameter defining the neutrino momentum (Eqn.~\ref{eqn:geodesic_x} below)
and $C[f]$ is the source term arising from interactions with the medium.
By Eqn.~\ref{eqn:geodesic_x},
the affine parameter has dimension ${\rm length}\,{\rm energy}^{-1}$,
so the source term has dimension ${\rm energy}\,{\rm length}^{-1}$.
The source term varies over phase space $(x^\alpha,p_\beta)$,
and depends locally on the distribution function $f$
and nonlocally on the distribution function of this neutrino and its
antiparticle at different momenta, $f'$ and $\bar{f}'$.
For simplicity we symbolize all of these dependencies with the shorthand $C[f]$.
The various neutrino interactions contributing to $C[f]$ are detailed in
App.~\ref{sec:source_terms}.

We make the right hand side of Eqn.~\ref{eqn:boltzmann} explicit by writing
the source term linear in $f$:
\begin{align}
  \label{eqn:boltzmann_linear}
  \frac{d}{d\lambda}f &=
  \mathscr{E} - \mathscr{K} f \\
  \label{eqn:boltzmann_linear_s}
  &= \mathscr{K}(\mathscr{S}-f)
\end{align}
where we have introduced
$\mathscr{E}$, the invariant total emissivity, and
$\mathscr{K}$, the invariant total opacity.
These describe respectively the energy gained and the energy lost
per length traveled by the neutrino, true scalar quantities
which take identical values for all observers.
In the second form we have introduced the source function
$\mathscr{S}\equiv\mathscr{E}/\mathscr{K}$, which makes manifest the
behavior of the right hand side, driving $f$ toward $\mathscr{S}$
over a lengthscale $\mathscr{K}^{-1}$ in the affine parameter;
or from Eqn.~\ref{eqn:proper_length} below, over a proper lengthscale
$\varepsilon/\mathscr{K}$ measured by the fiducial observer.

These coefficients are computed by considering their dependence
on neutrino and antineutrino distribution functions at other momenta
(i.e.\ Fermi-blocking).
We consider the two dominant classes of interactions in this work:
the absorption/emission (AE) and elastic scattering (SE) processes
listed in Tab.~\ref{tab:neutrino_processes}.
Thus we separate these coefficients:
\begin{align}
  \label{eqn:inv_emissivity_components}
  \mathscr{E} &= \mathscr{E}_{\rm AE} + \mathscr{E}_{\rm SE},\\
  \label{eqn:inv_opacity_components}
  \mathscr{K} &= \mathscr{K}_{\rm AE} + \mathscr{K}_{\rm SE}.
\end{align}

The absorption/emission coefficients are computed from sums
over the relevant emissivities and opacities
for the reactions
\begin{align}
  \nu_e + n            &\leftrightarrow e^- + p,\nonumber\\
  \bar{\nu}_e + p      &\leftrightarrow e^+ + n,\nonumber\\
  \nu_e + {}^AZ        &\leftrightarrow e^- + {}^A(Z+1),\nonumber\\
  \bar{\nu}_e + {}^AZ  &\leftrightarrow e^+ + {}^A(Z-1),\nonumber
\end{align}
with ${}^AZ$ representing a nucleus of mass number $A$ and charge $Z$.
In terms of the emissivity $j(\varepsilon)$ describing number of neutrinos
of energy $\varepsilon$ emitted per length,
and the absorption opacity $\chi_a(\varepsilon)$
describing the number absorbed per length,
the coefficients are
\begin{align}
  \label{eqn:ae_emissivity_summed}
  \mathscr{E}_{\rm AE}(\varepsilon)
  &= \varepsilon \sum_{i\,{\rm reactions}} j_i(\varepsilon), \\
  \label{eqn:ae_opacity_summed}
  \mathscr{K}_{\rm AE}(\varepsilon)
  &= \frac{1}{1-f^{\rm eq}(\varepsilon)} \,
  \varepsilon \sum_{i\,{\rm reactions}} \chi_{{\rm a},i}(\varepsilon),
\end{align}
where $f^{\rm eq}$ is the distribution function of neutrinos in radiative
equilibrium with the matter, i.e. the Fermi-Dirac distribution function
\begin{equation}
  \label{eqn:feq}
  f^{\rm eq}(\varepsilon) \equiv
  \left(1+e^{\varepsilon/(k_{\rm B} T) -\eta_\nu}\right)^{-1},
\end{equation}
with the neutrino chemical potentials dependent on the local density,
temperature, and composition of the fluid via the neutron, proton,
and electron chemical potentials:
$\eta_{\nu_e}=-\eta_{\bar{\nu}_e}=\eta_p-\eta_n+\eta_{e^-}$ and
$\eta_{\nu_x}=0$.
The appearance of $f^{\rm eq}$ in Eqn.~\ref{eqn:ae_opacity_summed} is due to
the Fermionic nature of the neutrinos, causing $\mathscr{K}_{\rm AE}$ to be
different than the simple absorption opacity,
a phenomenon called stimulated absorption \cite{burr2006-neutrino_opacities}.
By detailed balance of the absorption/emission reactions, we may alternatively
write the emissivity in terms of the equilibrium distribution function:
\begin{equation}
  \label{eqn:ae_emissivity_feq}
  \mathscr{E}_{\rm AE}(\varepsilon)
  = \mathscr{K}_{\rm AE}(\varepsilon) f^{\rm eq}(\varepsilon).
\end{equation}
See App.~\ref{ssec:sources_ae} for details.
\footnote{TODO: identify $\mathscr{K}_{\rm AE}$ and $\kappa^*$,
  in e.g.\ \cite{hari2010-gr_nunubar_collapsar}}

The elastic scattering coefficients are computed from
a background field, and a sum over opacities for
the reactions
\begin{align}
  \nu + N     &\leftrightarrow \nu + N,\nonumber\\
  \nu + {}^AZ &\leftrightarrow \nu + {}^AZ,\nonumber
\end{align}
with $N$ standing in for either $n$ or $p$.
In terms of a background field $\Phi(\varepsilon)$ describing
the number of neutrinos of energy $\varepsilon$ present at this event,
and the scattering opacity $\chi(\varepsilon)$ describing the
number scattered to other directions per length,
the coefficients are
\begin{align}
  \label{eqn:se_emissivity}
  \mathscr{E}_{\rm SE}(\varepsilon)
  &= \mathscr{K}_{\rm SE}(\varepsilon) \, \Phi(\varepsilon) \, \\
  \label{eqn:se_opacity_summed}
  \mathscr{K}_{\rm SE}(\varepsilon)
  &= \varepsilon \sum_{i\,{\rm reactions}} \chi_{{\rm s},i}(\varepsilon).
\end{align}
In the isotropic limit of trapped radiation,
$\Phi$ is equivalent to $f^{\rm eq}$;
in the free-streaming limit at a distance $r$ from a source,
$\Phi$ attenuates as $r^{-2}$.
See App.~\ref{ssec:sources_se} for details.

We may use Eqns.~\ref{eqn:ae_emissivity_feq} and \ref{eqn:se_emissivity}
to separate the source function thus:
$\mathscr{S}=\mathscr{S}_{\rm AE}+\mathscr{S}_{\rm SE}$, with
\begin{align}
  \label{eqn:source_ae}
  \mathscr{S}_{\rm AE}
  &= \frac{\mathscr{K}_{\rm AE}}{\mathscr{K}} f^{\rm eq}, \\
  \label{eqn:source_se}
  \mathscr{S}_{\rm SE}
  &= \frac{\mathscr{K}_{\rm SE}}{\mathscr{K}} \Phi.
\end{align}
With these definitions we write a separated form of
Eqn.~\ref{eqn:boltzmann_linear_s}:
\begin{equation}
  \label{eqn:boltzmann_split}
  \frac{d}{d\lambda}f =
  \mathscr{K}_{\rm AE}(f^{\rm eq}-f)
  + \mathscr{K}_{\rm SE}(\Phi-f)
\end{equation}
From Eqn.~\ref{eqn:boltzmann_split} we see that
absorption/emission interactions drive the distribution function
toward $f^{\rm eq}$ over an affine lengthscale $\mathscr{K}_{\rm AE}^{-1}$,
and elastic scattering interactions drive it
toward $\Phi$ over an affine lengthscale $\mathscr{K}_{\rm SE}^{-1}$.
(Eqn.~\ref{eqn:proper_length} translates affine length to proper length
for a given observer; in this case the proper length scale is
$\varepsilon/\mathscr{K}$.)

This paper introduces the use of neutrino densities and fluxes evolved in an
M1 transport simulation to estimate the background field, $\Phi$.
App.~\ref{ssec:sources_se} details a method to calculate
$\Phi(\varepsilon)$ in two different ways:
\begin{itemize}
\item
  the \emph{spectral} method using densities and fluxes
  extracted from a simulation evolved over multiple energy groups
  to compute the background field with Eqn.~\ref{eqn:background_phi},
\item
  the \emph{gray} method using energy-integrated densities and fluxes
  extracted from a gray simulation and approximating the
  energy distribution with Eqns.~\ref{eqn:J_from_gray} and
  \ref{eqn:H_from_gray}.
\end{itemize}

\subsection{Trajectories}
\label{ssec:trajectories}
Each trajectory is uniquely labeled by a pair of vectors
giving some event on the trajectory, $x^\alpha$,
and the momentum at that event, $p_\beta$.
A family of trajectories shares the same $x^\alpha$; for example,
the family of emission trajectories is ($x^\alpha_e,p_\beta$),
and the family of observation trajectories is ($x^\alpha_o,p_\beta$).

Neutrino trajectories obey the geodesic equation, which may be decomposed into
the coupled first-order equations
\begin{equation}
\label{eqn:geodesic_x}
  \frac{d x^\alpha}{d\lambda} = p^\alpha,
\end{equation}
and
\begin{equation}
\label{eqn:geodesic_p}
  \frac{d p_\beta}{d\lambda} = -\Gamma^\alpha_{\beta\gamma} p^\gamma p_\alpha,
\end{equation}
where $p^\alpha=\psi^{\alpha\beta}p_\beta$,
$\psi^{\alpha\beta}$ is the inverse of the spacetime metric $\psi_{\alpha\beta}$,
and $\Gamma^\alpha_{\beta\gamma}$ are the standard connection coefficients,
\begin{equation}
  \label{eqn:christoffel}
  \Gamma^\alpha_{\beta\gamma} =
  \frac{1}{2} \psi^{\alpha\mu}
  (\psi_{\mu\beta,\gamma} + \psi_{\mu\gamma,\beta} - \psi_{\beta\gamma,\mu}),
\end{equation}
with the comma denoting a partial derivative
$\psi_{\mu\beta,\gamma}=\partial_\gamma\,\psi_{\mu\beta}$.

Each trajectory is parameterized by affine parameter, $\lambda$, increasing in
the direction of $\ell_\beta$. We label $\lambda=\lambda_e$ at $x^\alpha_e$,
as in Fig.~\ref{fig:affine_param}.
If we multiply Eqn.~\ref{eqn:geodesic_x} by $u_\alpha \equiv dx_\alpha/ds$,
we find the element of proper distance traversed by the
neutrino as measured by the fiducial observer $u^\alpha$ is
\begin{equation}
  \label{eqn:proper_length}
  ds=\varepsilon \,d\lambda.
\end{equation}

\begin{figure}
  \includegraphics[width=\columnwidth]{fig-affine_parameter_label}
  % Script: $DIR/figures-radiation_transport_schematics.nb
  % w/ DIR=/Users/mbdeaton/Dropbox/Research/NuclearAstrophysics/NotesAndReference
  \caption{Affine parameterization of a neutrino trajectory of momentum
    $p_\alpha$.
    The fiducial observer with velocity $u^\alpha$ sits at $x^\alpha_o$,
    the neutrino emission event is at $x^\alpha_e$. The affine parameter
    increases from the emission event: $\lambda_e<\lambda'<\lambda''$.
    The green slab represents dense matter.}
  \label{fig:affine_param}
\end{figure}

\subsection{The Formal Solution}
\label{ssec:rendering_eqn}
We can integrate Eqn.~\ref{eqn:boltzmann_split} directly, 
with the solution split into a boundary, absorption/emission, and
a scattering term, $f=f_{\rm bdry}+f_{\rm AE}+f_{\rm SE}$:
\begin{align}
  \label{eqn:rendering_fbdry}
  f_{\rm bdry}(\lambda,\lambda_e)
  &= f(\lambda_e) e^{-\tau(\lambda,\lambda_e)}, \\
  \label{eqn:rendering_fae}
  f_{\rm AE}(\lambda,\lambda_e)
  &= \int_{\lambda_e}^\lambda d\lambda'\,e^{-\tau(\lambda,\lambda')}
  \mathscr{K}_{\rm AE}(\lambda')f^{\rm eq}(\lambda'), \\
  \label{eqn:rendering_fse}
  f_{\rm SE}(\lambda,\lambda_e)
  &= \int_{\lambda_e}^\lambda d\lambda'\,e^{-\tau(\lambda,\lambda')}
  \mathscr{K}_{\rm SE}(\lambda')\Phi(\lambda'),
\end{align}
where the optical depth is defined,
\begin{equation}
  \label{eqn:optical_depth}
  \tau(\lambda,\lambda') \equiv \int_{\lambda'}^\lambda
  d\lambda'' \, \mathscr{K}(\lambda''),
\end{equation}
and the parameterization conventions are depicted in
Fig.~\ref{fig:affine_param}.
Note that Eqn.~\ref{eqn:optical_depth} employs the total absorption
plus scattering opacity, so that the optical depth attenuating the
integrands of Eqns.~\ref{eqn:rendering_fae} and \ref{eqn:rendering_fse}
is the total optical depth.

%Note the following equivalents; we could use any of these to express
%the integrands above:
%\begin{align}
%  \mathscr{K}_{\rm AE} \, f^{\rm eq}
%  &= \mathscr{K}\, \mathscr{S}_{\rm AE}
%  = \mathscr{E}_{\rm AE},\nonumber \\
%  \mathscr{K}_{\rm SE} \, \Phi
%  &= \mathscr{K} \, \mathscr{S}_{\rm SE}
%  = \mathscr{E}_{\rm SE}.\nonumber
%\end{align}

\subsection{Moments of the Distribution Function}
\label{ssec:moments}
\footnote{TODO: code these up}
We may take angular moments of the distribution function:
\begin{align}
  \label{eqn:J}
  J(\varepsilon) &=
  \frac{\varepsilon^3}{(2\pi)^3} \oint d\Omega' f(\varepsilon, \ell'_\beta) \\
  \label{eqn:Ha}
  H^\mu(\varepsilon) &=
  \frac{\varepsilon^3}{(2\pi)^3} \oint d\Omega' f(\varepsilon, \ell'_\beta) \ell'^\mu \\
  \label{eqn:Sab}
  S^{\mu\gamma}(\varepsilon) &=
  \frac{\varepsilon^3}{(2\pi)^3} \oint d\Omega' f(\varepsilon, \ell'_\beta) \ell'^\mu \ell'^\gamma,
\end{align}
defining the specific energy density,
specific momentum density, and
specific radiation pressure tensor, respectively.
Here ``specific'' refers to the quantity being integrable over neutrino energy.
\footnote{TODO: relate moments to other frames?}
Integrals~\ref{eqn:J}--\ref{eqn:Sab} are performed over a solid angle in
momentum space while holding $\varepsilon$ constant:
$d\Omega \equiv d(\cos a)\,db$.
\footnote{TODO: consistent with Eqn.~\ref{eqn:def_direction}?}
We also make use of the number density and flux defined
\begin{align}
  \label{eqn:G}
  G(\varepsilon) &=
  \frac{\varepsilon^2}{(2\pi)^3} \oint d\Omega' f(\varepsilon, \ell'_\beta), \\
  \label{eqn:Ka}
  K^\mu(\varepsilon) &=
  \frac{\varepsilon^2}{(2\pi)^3} \oint d\Omega' f(\varepsilon, \ell'_\beta) \ell'^\mu.
\end{align}

The energy-integrated moments take the form
\begin{equation}
  \label{eqn:J_H_S_eps_integrated}
  X = \int_0^\infty d\varepsilon \, X(\varepsilon),
\end{equation}
with $X$ standing in for any of $\{J,H^\mu,S^{\mu\gamma}\}$ or
$\{G,K^\mu\}$,
the first having dimension ${\rm energy}\,{\rm length}^{-3}$,
and the second ${\rm length}^{-3}$.

We compute moments for a particular observer
by specifying the four-velocity in Eqn.~\ref{eqn:def_momentum}.
Two choices are particularly useful:
an observer stationary in the coordinate frame (i.e. Eulerian),
or one stationary in the fluid frame (i.e. comoving)
\cite{smar1980-gr_hydro}.
Explicit definitions are given in App.~\ref{sec:definitions}.
We distinguish moments computed for an Eulerian observer with a tilde,
e.g. $\tilde{J}$, $\tilde{H}^\mu$, $\tilde{S}^{\mu\nu}$;
note that these three Eulerian moments are identical to the lab-frame moments
$E$, $F^\mu$, and $P^{\mu\nu}$
defined in
\cite{shib2011-truncated_moment, ocon2015-gr1d_with_nu, fouc2015-m1_nsbh}.

\subsection{Numerical Implementation}
\label{ssec:numerical}
Much of our numerical implementation is borrowed from the geodesic evolution
system described in~\cite{bohn2016-code}.
We integrate Eqns.~\ref{eqn:geodesic_x} and~\ref{eqn:geodesic_p} in the form
given by \citep{hugh1994-eh_finding},
and Eqns.~\ref{eqn:rendering_fae}, \ref{eqn:rendering_fse},
and \ref{eqn:optical_depth} in the form given below.
By using the time-component of Eqn.~\ref{eqn:geodesic_x} ($dt=d\lambda \, p^t$)
we may transform the integrations to coordinate time.
The coupled system of ordinary differential equations is
\begin{align}
  \label{eqn:num_x}
  \frac{dx^i}{dt} &=
  g^{ij} \frac{p_j}{p^t} - \beta^{i},\\
  \label{eqn:num_p}
  \frac{dp_i}{dt} &=
  -\alpha \alpha_{,i} p^t
  + \beta^k_{,i} p_k
  - \frac{1}{2} g^{jk}_{,i}\frac{p_j p_k}{p^t},\\
  \label{eqn:num_tau}
  \frac{d\tau}{dt} &=
  - \frac{1}{p^t}\mathscr{K},\\
  \label{eqn:num_fae}
  \frac{df_{\rm AE}}{dt} &=
  \frac{1}{p^t}e^{-\tau}\mathscr{K_{\rm AE}} \,f^{\rm eq},\\
  \label{eqn:num_fse}
  \frac{df_{\rm SE}}{dt} &=
  \frac{1}{p^t}e^{-\tau}\mathscr{K_{\rm SE}} \,\Phi.
\end{align}

We integrate each ray until we reach a terminal optical depth of
$\tau_{\rm term}$ at the earliest effective emission event $x_e^\alpha$.
The concept of earliest emission event is a fictitious construct we use
to allow us to truncate the integration at an event along the ray where any
further additions to the field are negligible due to the large optical depth
between $x_e^\alpha$ and $x_o^\alpha$.
We choose $\tau_{\rm term}=14$ so that $e^{-\tau_{\rm term}}<10^{-6}$,
and we then discard the contribution of $f_{\rm bdry}$
(Eqn.~\ref{eqn:rendering_fbdry}).

Since we don't know the emission event a priori,
we follow the integration backwards in time,
from $t_o$ to $t_e$.
We begin each integration by setting initial values for the variables at $t_o$:
the observer specifies $x^i_o$ and $p_{i,o}$,
and we set $f_{{\rm AE},o}=f_{{\rm SE},o}=0$ and $\tau_o=0$.

We integrate Eqns.~\ref{eqn:num_x}--\ref{eqn:num_tau} with adaptive step sizes,
using the 3rd order Runge-Kutta algorithm
which produces an error estimate by comparing the 3rd and 2nd order solutions.
After each step is taken, the errors for each of the nine variables of
Eqns.~\ref{eqn:num_x}--\ref{eqn:num_tau} are compared.
If the maximum of these errors exceeds some absolute or relative threshold
\footnote{TODO: specify ScaledAbsRel algorithm}
the step size is decreased and the step recomputed,
if it is below some threshold the next step size is increased.
% see OdeIntegrators::AdaptiveDense
% and OdeSteppers::RungeKutta3
% and OdeSteppers::DormandPrince5
% and OdeController
% and OdeControllers::ProportionalIntegral_Base::StepSizeToAttempt
% and OdeErrorMeasures::ScaledAbsRel
\footnote{TODO: specify}

We integrate these equations through the simulated spacetime
over which the following volume data are known:
the spacetime metric $\psi_{\alpha\beta}$,
its derivatives $\psi_{\alpha\beta,\gamma}$,
the fluid velocity $u_i$,
Lorentz factor $W$,
density $\varrho$,
temperature $T$,
and electron fraction $Y_e$,
all defined in App.~\ref{sec:definitions}.
These fields are stored in spectral representation.
If they are computed via hydrodynamical simulation, they may be saved to disk
at specified coordinate times and interpolated
with spectral interpolation in space,
and 1st-order polynomial interpolation in time
(as described in \cite[App.~B]{bohn2016-code}).
If computed from a quasi-stationary solution to the
general relativistic hydrodynamics equations,
a single time is stored to disk, and no time interpolation is needed.
In this paper for simplicity and to limit computational memory loads,
we limit ourselves to quasistationary configurations
and use single time slices in every test case.

\section{Code Tests}
\label{sec:tests}

To test our algorithm, we integrate the ray tracing equations
(Eqns.~\ref{eqn:num_x}--\ref{eqn:num_tau}) for various observers
in the following configurations. This suite of configurations
may be understood as a hierarchy of increasing physical realism:
beginning with a homogeneous medium of effectively
infinite extent (i.e.\ optically thick) and progressing to a model of a
1D pre-supernova, post-bounce collapse profile
evolved using an M1 transport hydrodynamical simulation.
In the pre-supernova model, we compare ray tracing distributions
to those calculated in a Monte Carlo transport simulation.
Finally, we calculate global measures of the neutrino fields outside of the
hypermassive neutron star and disk formed in a binary neutron star
merger simulated by \cite{fouc2016-m1_nsns, fouc2016-m1_evolve_n},
and compare ray tracing results to those from the M1 transport simulation.

In the following we present two forms of the integrated distribution functions,
Eqns.~\ref{eqn:num_fae} and \ref{eqn:num_fse}:
1) the full form  by which the solution is $f_{\rm scat}=f_{\rm AE,scat}+f_{\rm SE,scat}$
(labeled \emph{scat} since we include elastic scattering), and
2) the form treating only absorption/emission interactions
in which we set $\mathscr{K}_{\rm SE}=0$ so that $f_{\rm noscat}=f_{\rm AE,noscat}$
(labeled \emph{noscat} since we ignore elastic scattering).
There is no simple relation between these forms
(i.e.\ $f_{\rm AE,scat} \neq f_{\rm AE,noscat}$)
because $f_{\rm AE}$ is integrated using the absorption plus scattering optical depth
in the first form and solely the absorption optical depth in the second form.
The differences between the two methods is apparent
in Fig.~\ref{fig:cumulative_f_homogeneous}.
Before this work, the \emph{noscat} form of the equations has been the standard for
\footnote{TODO: explain why; cite others}
ray tracing, though many authors include the scattering optical depth in
the integration of Eqn.~\ref{eqn:num_fae}
(see for example \cite{hari2010-gr_nunubar_collapsar}).

\subsection{Infinite Homogeneous Slab:
  Testing Thermodynamic Equilibrium}
\label{ssec:test_equilibrium}
In an optically thick region the radiation field is in thermodynamic equilibrium
with the matter.
We set up a large slab of matter in a Minkowski spacetime
representative of the fluid thermodynamic state at a radius of 50~km
in the test presented in Sec.~\ref{ssec:test_collapse}---a
massive collapsed star following core bounce---where
$\varrho=10^{11}\,{\rm g}\,{\rm cm}^{-3}$, $T=3.7$~MeV, and $Y_e=0.12$.
For comparision with that test we use the LS180 equation of state
\cite{latt1991-nuc_eos}
\footnote{available at \lstinline{stellarcollapse.org}}
in which the equilibrium $\nu_e$ neutrino chemical potential is
$\eta_{\nu_e}=-0.1555$
(with $\eta_{\bar{\nu}_e}=-\eta_{\nu_e}$ and $\eta_{\nu_x}=0$).
The opacity table is computed using \lstinline{NuLib} from cross-sections
neglecting weak magnetism
\footnote{TODO: anything else?}
and is stored on a grid covering energy, density, temperature, and composition
ranges spanning $\varepsilon_i\in[1,280.5]$~MeV logarithmically,
$\varrho\in[10^6,6.3\times10^{15}]\,{\rm g}\,{\rm cm}^{-3}$ logarithmically,
$T\in[0.05,200]$~MeV logarithmically, and 
$Y_e\in[0.035,0.55]$ linearly,
with grid extents$\{18,82,65,51\}$ respectively.
Interpolation is performed linearly.

We use the equilibrium distribution functions to define background
fields for the \emph{scat} case:
in the \emph{spectral} method we use
$J(\varepsilon)=\varepsilon^3 f^{\rm eq}(\eta_\nu,T;\varepsilon)/(2\pi^2)$,
and in the \emph{gray} method we use
$J=C T^4 \mathscr{F}_3(\eta_\nu)/(2\pi^2)$ and
$G=C T^3 \mathscr{F}_2(\eta_\nu)/(2\pi^2)$,
using the Fermi integrals defined in Eqn.~\ref{eqn:fermi_integral},
and with $C$ the conversion constant from ${\rm energy}^3$ to ${\rm length}^{-3}$.

Fig.~\ref{fig:mfps-50km} presents the neutrino mean free paths
at this thermodynamic point over two energy decades.
For this test we choose a slab large enough for neutrinos of all
energies to be trapped.
We sample the distribution function with ray tracing over a uniform grid
of 40 points in energy $\varepsilon \in [0,100]$~MeV.
The domain extends to $\tau_{\rm term}\,L_{\rm mfp,max}\approx10^7\,{\rm km}$,
since our ray tracing algorithm integrates rays to terminal optical
depths of $\tau_{\rm term}\geq14$.

\begin{figure}
  %\includegraphics[width=\columnwidth]{figure-mfps-50km}
  \resizebox{\columnwidth}{!}{\input{figure-mfps-50km.tex}}
  % Data:   $DIR/table-50km-nu?.dat
  % Script: $DIR/GnuplotScripts/plot-mfp-50km.gnu
  % w/ DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20161007-ray_tracing_tests/HomogeneousIsotropic
  \caption{Mean free paths
    representative of the post-bounce collapse profile (presented in
    Sec.~\ref{ssec:test_collapse}) at
    50~km, where $\varrho=10^{11}\,{\rm g}\,{\rm cm}^{-3}$,
    $T=3.7$~MeV, and $Y_e=0.12$; this thermodynamic state is used in
    the tests with homogeneous matter distributions
    (presented in Secs.~\ref{ssec:test_equilibrium}--\ref{ssec:test_ab_star}).
    In the fluid rest frame the mean free paths are given by the inverses of the opacities
    $\chi^*_a=\mathscr{K}_{\rm AE}/\varepsilon$ for absorption and
    $\chi_s=\mathscr{K}_{\rm SE}/\varepsilon$ for elastic scattering interactions
    (see App.~\ref{sec:source_terms} for definitions).
    Note that elastic scattering opacities are identical across species
    below energies at which weak magnetism plays a role \cite{horo2002-weak_mag}.
    In this data they are exactly identical
    because we have turned off weak magnetism in our opacity calculations
    in order to compare our results with the historical literature.
    Computed using \lstinline{NuLib}.
  }
  \label{fig:mfps-50km}
\end{figure}

Fig.~\ref{fig:cumulative_f_homogeneous} displays the cumulative distribution
function integrated along the ray, for each of the three species and using
both methods, \emph{noscat} and \emph{scat}, at the single energy
$\varepsilon=11.25$~MeV. We display
Eqns.~\ref{eqn:num_fae} and \ref{eqn:num_fse} in their integral form:
\begin{equation}
  \label{eqn:cumulative_f}
  f(t)
  = - \int_{0}^t \frac{dt'}{p^t}e^{-\tau(0,t')}
  \mathscr{K}(t')f^{\rm eq}(t'),
\end{equation}
and the integration proceeds backwards in time, from $t=0$
\footnote{TODO: could write more simply because Minkowski}
to some terminal $t<0$.
The figure shows this backwards-in-time integration proceeding left to right.

\begin{figure}
  %\includegraphics[width=\columnwidth]{fig-cumulative_f-homogeneous}
  \resizebox{\columnwidth}{!}{\input{fig-cumulative_f-homogeneous.tex}}
  % Data:   $DIR/Lev1-v0-ShootANeutrino-01/Lev1
  % Script: $DIR/GnuplotScripts/plot_f-integral_over_t.gnu
  % w/ DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20161007-ray_tracing_tests/HomogeneousIsotropic
  \caption{Cumulative distribution functions at $\varepsilon=11.25$~MeV
    in the homogeneous infinite slab test
    (presented in Sec.~\ref{ssec:test_equilibrium}).
    The integration proceeds from left to right, or backwards in time $t$.
    Each ray terminates when it achieves a total optical depth
    greater than $\tau_{\rm term}=14$.
    The points plotted correspond to the time steps chosen by the adaptive
    time-stepping algorithm described in Sec.~\ref{ssec:numerical}.
  }
  \label{fig:cumulative_f_homogeneous}
\end{figure}

In Fig.~\ref{fig:cumulative_f_homogeneous} we see some expected features.
The final distribution functions asymptote to their equilibrium Fermi-Dirac
levels at this energy
$f^{\rm eq}_{\{\nu_e,\bar{\nu}_e,\nu_x\}}=\{0.039, 0.053, 0.046\}$
within a few mean free paths;
in the \emph{scat} cases it is the sums $f_{\rm AE}+f_{\rm SE}$ that
achieve these values.
And the lengths of the rays are proportional to the mean free paths:
$L_{\nu_x}>L_{\bar{\nu}_e}>L_{\nu_e}$;
in the \emph{scat} cases these lengths are less than in the \emph{noscat} cases,
since the total mean free paths are less than the absorption mean free paths
(significantly so for $\nu_x$, negligibly so for $\nu_e$).

As Fig.~\ref{fig:mfps-50km} reveals, integrating the distribution functions
over the energies $\varepsilon\in[0,100]$~MeV
probes the numerical solution over length scales from
0.1~km to $10^{5}\,{\rm km}$.
Fig.~\ref{fig:homogeneous_isotropic} shows the error in our results for
the \emph{noscat} case.
In this simple case, the dominant source of error is the neglected
boundary term with a relative scale of $e^{-\tau_{\rm term}}\approx10^{-6}$.
In the following inhomogeneous configurations,
errors from the integration dominate over the boundary term.
In this test and only this test, we used a higher-order integrator, a 5th
order Dormand-Prince algorithm \cite{pres2007-nr_3rd_ed}. This is because
\footnote{TODO: discuss with Matt}
our default 3rd order Runge-Kutta algorithm estimates a vanishing error
in this configuration.

\begin{figure}
  %\includegraphics[width=\columnwidth]{figure-ferr-50km}
  \resizebox{\columnwidth}{!}{\input{figure-ferr-50km.tex}}
  % Data:   $DIR/Lev1-v0-noscat-01/Lev1
  % Script: $DIR/GnuplotScripts/plot_ferr.gnu
  % w/ DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20161007-ray_tracing_tests/HomogeneousIsotropic
  \caption{Relative error in integrated equilibrium distribution functions
    in the infinite homogeneous slab
    (presented in Sec.~\ref{ssec:test_equilibrium}).
    Plotted here are the relative differences between the final
    $f_{\rm AE}$ in the \emph{noscat} case
    (see Fig.~\ref{fig:cumulative_f_homogeneous})
    and the equilibrium distribution functions given by
    Eqn.~\ref{eqn:feq}.
    The source of this error is the discarded boundary term
    $f_{\rm bdry}\sim10^{-6}$, described in Sec.~\ref{ssec:numerical}.
  }
  \label{fig:homogeneous_isotropic}
\end{figure}

\subsection{Infinite Homogeneous Moving Slab:
  Testing Doppler Shift}
\label{ssec:test_doppler}
We reproduce the test above, again in Minkowski spacetime,
but with the matter and observer in relative motion.
We use a stationary observer
and fluid moving in the positive $z$-direction:
with $u^\alpha \rightarrow W(1,0,0,v)$ and $W=(1-v^2)^{-1/2}$,
where $W$ is the Lorentz factor defined in App.~\ref{sec:definitions}.
All other thermodynamic variables and background fields are unchanged
since our ray integration uses these quantities in the fluid frame.

A stationary observer measures an energy of $\tilde{\varepsilon}$
for a neutrino with momentum
$p_\alpha \rightarrow \tilde{\varepsilon}(-1,\Omega_i)$
and direction $\Omega_i \rightarrow (\sin A, 0, \cos A)$.
In the fluid frame this neutrino has energy $\varepsilon=-u^\alpha p_\alpha$;
therefore the average energy varies with observing angle like
\begin{equation}
  \label{eqn:doppler_avg_eps}
  \langle \tilde{\varepsilon} \rangle(\cos A) =
  \langle \varepsilon \rangle^{\rm eq} \frac{1}{W(1-v \cos A)},
\end{equation}
where the symbol $\langle \tilde{\varepsilon} \rangle(\cos A)$ emphasizes
the functional dependence on $\cos A$.
The equilibrium average energy $\langle \varepsilon \rangle^{\rm eq}$
is given by $T \mathscr{F}_3(\eta_\nu)/\mathscr{F}_2(\eta_\nu)\approx 3T$,
with the Fermi integrals given in Eqn.~\ref{eqn:fermi_integral}.
Eqn.~\ref{eqn:doppler_avg_eps} describes the well-known Doppler effect.

We sample the distribution function $f(\tilde{\varepsilon},\cos A,B)$
with ray tracing
over a uniform grid of 40 points in energy $\tilde{\varepsilon}\in[0,100]$
and 30 points in angle $\cos A\in[-1,1]$, holding fixed $B=\pi$.
Results are shown in Fig.~\ref{fig:avg_eps_doppler} for
the velocities $v=\{0,0.1,0.8\}$.
The ray tracing results are computed from total densities in each angular bin,
that is
\begin{equation}
  \label{eqn:avg_eps_per_cosA}
  \langle \tilde{\varepsilon} \rangle (\cos A) =
  \frac{\tilde{J}(\cos A)}{\tilde{G}(\cos A)},
\end{equation}
with the Eulerian energy density per angular bin given by
\begin{equation}
  \label{eqn:J_per_cosA}
  \tilde{J}(\cos A) =
  \frac{1}{(2\pi)^3} \int_0^{\infty} d\tilde{\varepsilon} \, \tilde{\varepsilon}^3
  \int_0^{2\pi} dB \, f(\tilde{\varepsilon},\cos A,B),
\end{equation}
and $\tilde{G}(\cos A)$ defined analogously.
\footnote{TODO: add metric terms for these integrals}
Results are identical for all \emph{scat} and \emph{noscat} methods.
In Fig.~\ref{fig:avg_eps_doppler}
we see the common features of a red-shifted spectrum for receding fluid
($\cos A\ll0$),
a blue-shifted spectrum for approaching fluid
($\cos A\gg0$),
and a slightly red-shifted spectrum for fluid moving transverse to the
observer ($\cos A=0$).

\begin{figure}
  %\includegraphics[width=\columnwidth]{fig-moving_fluid_avg_eps}
  \resizebox{\columnwidth}{!}{\input{fig-moving_fluid_avg_eps.tex}}
  % Data:   $DIR/Lev1-v*-noscat-01
  % Script: $DIR/plot-avg_eps_vs_cosA.gnu
  % w/ DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20161007-ray_tracing_tests/MovingFluid
  \caption{Average energy of $\bar{\nu}_e$ neutrino fields
    measured by a stationary observer,
    $\langle\tilde{\varepsilon}\rangle(\cos A)$,
    in the moving slab test (presented in Sec.~\ref{ssec:test_doppler}).
    The observer sees neutrinos with $\cos A>0$ to be moving primarily with
    the fluid.
    The points are computed from ray tracing spectra; the lines are the
    analytic formula Eqn.~\ref{eqn:doppler_avg_eps}.
  }
  \label{fig:avg_eps_doppler}
\end{figure}

\subsection{Idealized Star:
  Testing the Decoupling Regime}
\label{ssec:test_ab_star}
The homogeneous configurations of the previous sections may be extended to
probe the solution outside the optically thick regime
by setting up an idealized homogeneous star, with the thermodynamic variables
constant inside radius $R$ and vanishing outside.
We choose $R=50$~km and place the observer at $r=75$~km,
at which position there is a radiation cone of half-opening angle
$\cos A_{\rm max}\approx0.75$.

The formal solutions of
Eqns.~\ref{eqn:rendering_fae} and \ref{eqn:rendering_fse}
may be directly integrated in this scenario. Assuming Minkowski spacetime
and stationary fluid we have
\begin{align}
  \label{eqn:fae_homogeneous_general}
  f_{\rm AE} &=
  \frac{\chi_a^{*}}{\chi} f^{\rm eq} \left(1-e^{-\chi s}\right), \\
  \label{eqn:fse_homogeneous_general}
  f_{\rm SE} &=
  \frac{\chi_s}{\chi} \Phi \left(1-e^{-\chi s}\right),
\end{align}
where $s$ is the path length traversed by the ray through the star,
\begin{equation}
  \label{eqn:geometric_s}
  s = 2 R \left(1-\frac{r^2}{R^2}(1-\cos^2 A)\right)^{1/2}.
\end{equation}
The total opacity is $\chi=\chi_a^*+\chi_s$, and the stimulated absorption
opacity $\chi_a^*$ and elastic scattering opacity $\chi_s$
are defined in App.~\ref{sec:source_terms}.

Since no analytic form is known for the background field $\Phi$
interior to the star, we examine only the \emph{noscat} case,
with $\chi_s=0$ and $\chi=\chi_a^{*}$.
This scenario has been widely used in the literature as a test for
radiation codes
\cite{smit1997-two_moment,abdi2012-monte_carlo,fouc2015-m1_nsbh}.
We sample the distribution function over a uniform grid of
30 points in angle $\cos A \in [0.734,1]$ and
40 points in energy $\tilde{\varepsilon} \in [0,100]$~MeV.
Because of the step function in fluid variables at radius $R$,
we throttle the maximum step size at 1.25~km.

In Fig.~\ref{fig:f_absorption_sphere} we display the samples at
$\varepsilon=11.25$~MeV, along with the analytic functions
specific to each species' equilibrium distribution function and opacity.
As expected only $\nu_e$ saturates at $f^{\rm eq}$, remaining almost
constant across $\cos A$ until we get to rays that pass through a length
of the star comparable to or less than the mean free path at this energy,
16~km. We also see that $\bar{\nu}_e$ comes close to saturating with
a mean free path just over 100~km; and $\nu_x$ is well into the optically
thin regime.

\begin{figure}
  %\includegraphics[width=\columnwidth]{fig-f-absorption_star-11MeV}
  \resizebox{\columnwidth}{!}{\input{fig-f-absorption_star-11MeV.tex}}
  % Data:   $DIR/Lev1-01/Lev1
  % Script: $DIR/Lev1-01/Lev1/plot_f_monochromes-11.25MeV.gnu
  % w/ DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20161007-ray_tracing_tests/HomogeneousAbsorptionSphere
  \caption{Distribution functions outside an idealized homogeneous star
    with radius $R=50$~km, and observer at $r=75$~km
    (presented in Sec.~\ref{ssec:test_ab_star}).
    $A$ is the angle between the neutrino momentum and the $\hat{r}$ direction.
    In this plot we display only the samples at energy
    $\varepsilon=11.25\,{\rm MeV}$.
    The points are computed from ray tracing; the lines from the analytic
    solution, Eqn.~\ref{eqn:fae_homogeneous_general}.
  }
  \label{fig:f_absorption_sphere}
\end{figure}

We can explain these features quantitatively by examining the limits
of Eqn.~\ref{eqn:fae_homogeneous_general}, expanding the exponential function
in powers of $\chi s$; the distribution function takes the limiting values
\begin{equation}
  \label{eqn:fae_homogeneous_limits}
  f_{\rm AE}(\cos A) = f^{\rm eq}
  \begin{cases}
    \chi s(\cos A) & s \ll \chi^{-1}\,\textrm{(thin)}\\
    1                  & s \gg \chi^{-1}\,\textrm{(thick)}.
  \end{cases}
\end{equation}
These limits are represented in Fig.~\ref{fig:f_absorption_sphere}:
with $\nu_x$ in the optically thin limit at all viewing angles,
and $\nu_e$ in the optically thick limit at viewing angles
$\cos A \gtrsim 0.8$.

\subsection{Idealized Compact Star:
  Testing Gravitational Redshift and Geodesic Curvature}
\label{ssec:test_gravity}
Unfinished: describe and point readers to Asano-Fukuyama test presented
in thesis ~\cite[Sec.~4.3.3]{deat2015-thesis}.
Also compare a few geodesics in Kerr to analytic.
\footnote{TODO: compute some geodesics in Kerr}

\subsection{Post-Bounce Collapse Profile:
  Testing Scattering}
\label{ssec:test_collapse}
To test our scattering treatments we calculate neutrino fields outside a
collapsed 15~${\rm M}_\odot$ star, 100~ms after core bounce,
comparing ray tracing fields to those from a Monte Carlo transport calculation.
Elastic scattering interior to the shock at $r\approx150$~km
significantly modifies the neutrinos' spectra,
\footnote{TODO: demonstrate}
and the extended envelope outside the shock becomes a source of
higher-than-average-energy neutrinos.

The 1D matter profile and M1 transport evolution are computed using
open source supernova evolution code \lstinline{GR1D}
\footnote{\url{http://www.gr1dcode.org}}
\cite{ocon2010-gr1d, ocon2015-gr1d_with_nu},
using a progenitor profile from \cite{woos1995-sn_progenitors}.
The matter is described by the LS180 equation of state \cite{latt1991-nuc_eos},
and the opacities are computed and stored in a table
as described in Sec.~\ref{ssec:test_equilibrium}.
This standard test is also presented for example in
\cite{ocon2015-gr1d_with_nu,fouc2015-m1_nsbh,abdi2012-monte_carlo}.
\footnote{TODO: also lieb2005, mull2010}

The matter profile and background field are stored on a spherical
pseudo-spectral grid composed of 11 spherical-shell subdomains
\footnote{TODO: cite early SpEC paper}
comprising a total of 62 radial grid points spaced approximately
logarithmically across $r\in(0,740)$~km.
The background scattering field $\Phi(\varepsilon)$ is supplied by
\lstinline{GR1D} in the form of $J(\tilde{\varepsilon}_i)$,
with $\varepsilon_i$ representing 18 energy groups
identical to those in the \lstinline{NuLib} table described in
Sec.~\ref{ssec:test_equilibrium}.
For the \emph{spectral} method we use $J(\tilde{\varepsilon}_i)$ directly,
using zeroth-order interpolation between energy groups;
for the \emph{gray} method we use $J=\sum J(\tilde{\varepsilon}_i) \Delta\tilde{\varepsilon}_i$,
and $G=\sum J(\tilde{\varepsilon}_i) \Delta\tilde{\varepsilon}_i/\tilde{\varepsilon}_i$ instead,
with $\Delta\tilde{\varepsilon}_i$ the bin-width of the $i$th energy group.

For fiducial neutrino distributions we use the matter profile
as input into a Monte Carlo radiation transport calculation using
open source neutrino transport code \lstinline{Sedonu}
\footnote{\url{https://bitbucket.org/srichers/sedonu}}
\cite{rich2015-monte_carlo}.
To homogenize the physics modeled across these three treatments
(M1 transport to provide the background fields,
ray tracing to compute neutrino distributions, and
Monte Carlo transport for a fiducial comparison)
we turn off inelastic scattering in the M1 and Monte Carlo codes,
since we have not incorporated that physics into this ray tracing code,
and we turn off general relativistic effects in the ray tracing code,
since the Monte Carlo code assumes Minkowski spacetime.

We place a stationary observer at $r=500$~km.
Fig.~\ref{fig:avg_eps_collapse} shows the distribution of average $\nu_x$
energies across incoming angles to this observer.
We show $\nu_x$ because they present the largest scattering effects:
they scatter through a thicker atmosphere outside their deep emission surface,
and their hotter spectrum experiences stronger modification due to the
$\varepsilon^2$ dependence of the scattering cross-section.
Against the fiducial Monte Carlo distribution, we show the
\emph{noscat} treatment, and the \emph{scat} treatment using both the
\emph{spectral} method and the \emph{gray} method.

\begin{figure}
  %\includegraphics[width=\columnwidth]{fig-collapse_profile_avg_eps_vs_cosA}
  \resizebox{\columnwidth}{!}{\input{fig-collapse_profile_avg_eps_vs_cosA.tex}}
  % Data:   $DIR/
  % Script: $DIR_GRAY_SC/../plot-compare_spectral_vs_gray-angular_distribution_avg_eps-production.gnu
  % w/  DIR_MC     =$DIR/ResultsSherwood-03/Spectra
  %     DIR_SPEC_SC=$DIR/ResultsBrett-A-03/Test-scat-01
  %     DIR_GRAY_SC=$DIR/ResultsBrett-B-03/Test-scat-01
  %     DIR_GRAY_AB=$DIR/ResultsBrett-B-03/Test-noscat-01
  % and DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20161007-ray_tracing_tests/CollapseProfile/
  \caption{Angular distribution of $\nu_x$ average energies
    $\langle\tilde{\varepsilon}\rangle(\cos A)$
    in the collapse profile test (described in Sec.~\ref{ssec:test_collapse}).
    The observer is at 500~km, and the shock at 150~km,
    so that the half-opening angle of the shock is $\cos A\approx0.954$.
    The four methods depicted are 1) a fiducial Monte Carlo calculation,
    ray tracing using the 2) spectral and 3) gray methods to estimate background
    fields for scattering, and 4) ray tracing neglecting scattering.
  }
  \label{fig:avg_eps_collapse}
\end{figure}

As expected the average energies from both
the scattering envelope ($\cos A \lesssim 0.95$) and
the bright core ($\cos A\gtrsim 0.995 $)
are well characterized by the \emph{scat} treatments
and badly characterized by the \emph{noscat} treatment.
In this case for $\nu_x$ the major effect of elastic scattering
is to decrease the average energy of neutrinos coming from the core
and increase the average energy of neutrinos coming from the envelope.

Although not shown here, the angular distribution contributing to the
total number density is also strongly affected by elastic scattering.
Without elastic scattering, the central regions emitting 60\%
of the neutrinos for the species $\{\nu_e,\bar{\nu}_e,\nu_x\}$
have scales $R_\nu\approx\{40,35,20\}$~km;
with elastic scattering the scales are
$R_\nu\approx\{50,45,35\}$~km.
\footnote{TODO: use a consistent lengthscale symbol throughout paper. $d$? $D$? $r$? $l$?}
% From by-hand summation over histogram G(cosA_i)*dcosA
% Script: $DIR/plot-compare_scat_vs_noscat-G_angular_distribution-all.gnu
% DIR=.../20161007-ray_tracing_tests/CollapseProfile/ResultsBrett-B-03

Total luminosities and average energies from the different treatments are
\footnote{TODO: update MC results 02$\rightarrow0$3}
presented in Tab.~\ref{tab:collapse_rt_vs_mc}.
\footnote{TODO: give equation for L}
As expected, $\nu_e$ is least affected by scattering, and $\nu_x$ most.
In fact, without scattering, $\nu_x$ luminosities are overestimated by more than
two orders of magnitude, due to the steep increase in temperature with depth
in the inner core.
Though contributing only a fraction of the total luminosity,
the average energy of neutrinos scattered to the observer from the envelope
outside the shock, $\langle\varepsilon\rangle^{\rm sc}$,
is poorly characterized by the \emph{noscat} treatment for all species.
By contrast, both the \emph{gray} and \emph{spectral} \emph{scat} treatments
faithfully describe the high scattered energies from the envelope.

\begin{table}%[h]
  % to get mc luminosities: tail $MCDATA/Spectra/luminosities_and_avg_eps-501km-nu?.dat
  % to get rt luminosities: tail $RTDATA/Moments-nu?.out
  % MCDATA=$DIR/ResultsSherwood-02
  % RTDATA_NO=$DIR/ResultsBrett-B-03/Test-noscat-01/
  % RTDATA_GR=$DIR/ResultsBrett-B-03/Test-scat-01/
  % RTDATA_SP=$DIR/ResultsBrett-A-03/Test-scat-01/
  % DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20161007-ray_tracing_tests/CollapseProfile/
  \caption{
    From the post-bounce collapse profile test (Sec.~\ref{ssec:test_collapse}),
    a comparison of total luminosities and average energies between
    the following methods:
    ray tracing ignoring elastic scattering `RT noscat',
    ray tracing with the gray scattering treatment `RT scat gr',
    ray tracing with the spectral scattering treatment `RT scat sp', and
    a fiducial Monte Carlo transport evolution `MC'.
    We also show average energies of the scattering envelope
    $\langle\tilde{\varepsilon}\rangle^{\rm sc}$, which are estimated by eye
    from plots like Fig.~\ref{fig:avg_eps_collapse} at $\cos A < 0.95$.
    Luminosities have units $10^{52}\,{\rm erg}\,{\rm s}^{-1}$ and
    average energies have units MeV.
    The $\nu_x$ luminosities are per-species:
    multiply by four to get the total heavy-lepton neutrino luminosities.
  }
  \label{tab:collapse_rt_vs_mc}
  \begin{tabularx}{\columnwidth}{X r r r r}
    & {\bf RT noscat} & \,\,{\bf RT scat gr} & \,\,{\bf RT scat sp} & \,\,{\bf MC} \\
    \hline
    $L_{\nu_e}$                                         &  3.53 & 3.14 & 3.04 & 3.48 \\
    $L_{\bar{\nu}_e}$                                   & 5.19 & 3.05 & 3.03 & 3.01 \\
    $L_{\nu_x}$                                         & 222.0  & 1.88 & 1.76 & 1.70 \\
    $\langle \tilde{\varepsilon}_{\nu_e} \rangle$               &  10.6 & 10.6 & 10.9 & 11.0 \\
    $\langle \tilde{\varepsilon}_{\bar{\nu}_e} \rangle$         & 14.2 & 13.0 & 13.8 & 13.7 \\
    $\langle \tilde{\varepsilon}_{\nu_x} \rangle$               & 47.8 & 16.0 & 17.3 & 16.2 \\
    $\langle \tilde{\varepsilon}_{\nu_e} \rangle^{\rm sc}$      & 4    & 14   & 17   & 16 \\
    $\langle \tilde{\varepsilon}_{\bar{\nu}_e} \rangle^{\rm sc}$& 4    & 24   & 20   & 18 \\
    $\langle \tilde{\varepsilon}_{\nu_x} \rangle^{\rm sc}$      & 2    & 28   & 31   & 27 \\
    \hline
  \end{tabularx}
\end{table}

We can make some quantitative sense of the scattered energies in
Fig.~\ref{fig:avg_eps_collapse} and Tab.~\ref{tab:collapse_rt_vs_mc}
using the solutions explored in the simple configurations above.
In particular, the average energy in the scattering envelope
($\cos A\lesssim 0.95$) is related to the spectrum in the direction
of the interior ($\cos A \sim 1$).
The relation may be derived by simplifying our realistic model
to that of a homogeneous matter profile and background scattering field.

By expanding the exponential of
Eqns.~\ref{eqn:fse_homogeneous_general} and \ref{eqn:fae_homogeneous_general}
as we did for Eqn.~\ref{eqn:fae_homogeneous_limits},
and furthermore factoring out the dominant energy dependence from the
cross-sections (i.e.\ $\chi\equiv\varepsilon^2 \zeta$
with $\zeta$ approximately constant), we have
\begin{align}
  \label{eqn:fae_homogeneous_thin_limit}
  f_{\rm AE}(\varepsilon)
  &= \varepsilon^2 \zeta_a \,f^{\rm eq}(\varepsilon)\,s, \\
  \label{eqn:fse_homogeneous_thin_limit}
  f_{\rm SE}(\varepsilon)
  &= \varepsilon^2 \zeta_s \,\Phi(\varepsilon)\,s,
\end{align}
where $s$ is the length of scattering envelope passed through by the ray.
\footnote{TODO: should be $\chi_a^*=\varepsilon\zeta_a$ for $\nu_x$}
In the envelope the local temperature is low so $f^{\rm eq}/\Phi\ll 1$,
and also $\zeta_a/\zeta_s\ll 1$,
so $f_{\rm SE}$ strongly dominates over $f_{\rm AE}$.
The average energy of the scattered field measured by our observer
is therefore
\begin{equation}
  \label{eqn:avg_eps_sc}
  \langle \tilde{\varepsilon} \rangle^{\rm sc}
  = \frac{\int d\varepsilon\,\varepsilon^3 f}{\int d\varepsilon\,\varepsilon^2 f}
  \approx \frac{\int d\varepsilon\,\varepsilon^5 \Phi}{\int d\varepsilon\,\varepsilon^4 \Phi}.
\end{equation}
Note that we have taken the liberty here of identifying the fluid-frame energy
$\varepsilon$ with the Eulerian energy $\tilde{\varepsilon}$,
since infall velocities in the envelope are $\sim0.1c$,
and as Fig.~\ref{fig:avg_eps_doppler} indicates,
the Doppler shift will therefore introduce an error into our analysis of
$\sim10\%$.

The spectrum of the background field $\Phi(\varepsilon)$ is well-approximated
as the \emph{scat} solution for $\cos A\sim 1$, since that is the dominant
source direction for neutrinos.
And because the scattering envelope is optically thin at all energies,
we can assume the neutrino spectrum is essentially unchanged in its passage
through the envelope.

In order to estimate $\langle \tilde{\varepsilon} \rangle^{\rm sc}$ analytically,
we must write the background field $\Phi(\varepsilon)$ analytically.
Direct Fermi-Dirac fits using a temperature and chemical potential
representative of a physical neutrinosurface fair poorly since neutrinos of
different energies decouple from the matter at different radii, over which
the thermodynamic state varies substantially.
Phenomenological Fermi-Dirac fits work well;
but so do pinched spectral fits which are much simpler
\cite{keil2003-pinched_spectra,miri2016-sn_neutrinos}:
\begin{equation}
  \label{eqn:pinched_spectra}
  \Phi^{\rm pi}(\varepsilon) \propto
  \varepsilon^{\alpha-2}
  \exp\left(-(\alpha+1)\frac{\varepsilon}{\langle\varepsilon\rangle}\right),
\end{equation}
(where our definition differs from that of \cite{keil2003-pinched_spectra}
by a factor of
$\varepsilon^2$, since we define our distribution function to be dimensionless
in natural units $\{\hbar,c\}=1$).

Pinching parameters
(calculated by eye from the \emph{spectral} \emph{scat} method)
for the species
$\{\nu_e,\bar{\nu}_e,\nu_x\}$ are $\alpha\approx\{3.6,5.1,2.3\}$.
Energy moments of pinched spectra like those in Eqn.~\ref{eqn:avg_eps_sc}
have a simple analytic form so that Eqn.~\ref{eqn:avg_eps_sc} becomes
\begin{align}
  \label{eqn:pinched_avg_eps_sc_1}
  \langle\tilde{\varepsilon}\rangle^{\rm sc}
  &\approx \frac{\alpha+3}{\alpha+1} \langle\varepsilon\rangle, \\
  %\label{eqn:pinched_avg_eps_sc_2}
  %&= \{1.4,1.3,1.6\} \langle\varepsilon\rangle, \\
  \label{eqn:pinched_avg_eps_sc_3}
  &\approx \{16,18,28\}\,~{\rm MeV},
\end{align}
again for $\{\nu_e,\bar{\nu}_e,\nu_x\}$ respectively.
These analytic predictions agree with the average energies of the scattering
envelope to approximately 10\%
% From by-eye examinations of <eps>(cosA) plot using data from
% MCDATA=$DIR/ResultsSherwood-02
% RTDATA_GR=$DIR/ResultsBrett-B-03/Test-scat-01/
% RTDATA_SP=$DIR/ResultsBrett-A-03/Test-scat-01/
%          nue     nua     nux
of all of the treatments including elastic scattering presented
in Tab.~\ref{tab:collapse_rt_vs_mc}, except for
$\bar{\nu}_e$ in the \emph{gray} \emph{scat} treatment,
which deviates from our prediction by 30\%.
This agreement is excellent despite the drastic
simplifications used in our toy model.

The $\langle\tilde{\varepsilon}_{\bar{\nu}_e}\rangle^{\rm sc}$ prediction
in the \emph{gray} \emph{scat} treatment is approximately 30\%
larger than the the fiducial Monte Carlo estimate. This is
due to the large negative local chemical potential
$\eta_{\bar{\nu}_e}\sim-10$ in the scattering envelope.
As described in Sec.~\ref{ssec:sources_se},
we use $\eta_\nu$ in the \emph{gray} treatment to construct our
synthetic spectrum from total neutrino densities $J$ and $G$.
This error and our successful analysis using pinched spectra above,
points the way to future improvements to the \emph{gray} method:
making better assumptions about the spectra which don't depend upon
local fluid quantities.

\subsection{Hypermassive Neutron Star Remnant:
  Testing Everything}
\label{ssec:test_disk_comparison}
The merger of two neutron stars by gravitational wave emission produces a
postmerger configuration composed of a single neutron star surrounded by a disk.
Because of strong differential rotation and shock heating,
the remnant temporarily avoids
collapse to a black hole, even if its mass exceeds the threshold of dynamical
instability for a rigidly rotating neutron star \cite{duez2009-review}.
These objects, called hypermassive neutron stars,
may avoid collapse for $X\textrm{--}Y$~s,
\footnote{TODO: cite $X$, $Y$}
depending upon a number of physical factors including thermal pressure,
magnetic fields, and the microphysics of the fluid.

Such a configuration was modeled in \cite{fouc2016-m1_nsns} by
evolving fluid and spacetime through the
final inspiral and merger of two identical neutron stars of
isolated gravitational mass 1.2~$M_{\odot}$. The postmerger configuration
was simulated using a gray M1 transport scheme for the neutrinos, evolving the
energy density, number density, and energy flux in addition to the standard
fluid and metric variables for $\sim10\,{\rm ms}$ following merger
\cite{fouc2016-m1_evolve_n}.
The fluid was modeled using the LS220 equation of state \cite{latt1991-nuc_eos}.

We use a single time snapshot from that configuration
at $t=11\,{\rm ms}$ after merger.
In this way we approximate the system as stationary over a light-crossing time
of around 1~ms, which is far below the thermal timescale of the remnant.
Figs.~\ref{fig:nsns_rho_merid}--\ref{fig:nsns_temp_equat}
show slices of density and temperature from the finite difference
fluid and M1 radiation data.
These data were evolved on a rectilinear grid spanning approximately 400~km in
both the $x$ and $y$ directions, and 150~km in the $z$ direction in our
evolution coordinates.

\begin{figure}
  \includegraphics[width=\columnwidth]{production-colormap-merid-rho}
  % VTK Data: zosima:/home/mbdeaton/Research/20170602-nsns_vol_data
  % Plots:    $DIR/production-colormap-{merid,equat}-{rho,temp}.png
  % w/ DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20170420-nsns_volume_data/02-two_domains_avg_eps/PlotsEtc
  \caption{A meridional slice of density in the hypermassive neutron
    star and disk configuration (Sec.~\ref{ssec:test_disk_comparison}).
    The distorted rectangular boundaries are the boundaries of the grid
    used in the numerical simulation, which employs a coordinate mapping to
    concentrate points near the central object.}
  \label{fig:nsns_rho_merid}
\end{figure}

\begin{figure}
  \includegraphics[width=\columnwidth]{production-colormap-equat-rho}
  % see notes from production-colormap-merid-rho
  \caption{An equatorial slice of density in the hypermassive neutron
    star and disk configuration (Sec.~\ref{ssec:test_disk_comparison}).}
  \label{fig:nsns_rho_equat}
\end{figure}

\begin{figure}
  \includegraphics[width=\columnwidth]{production-colormap-merid-temp}
  % see notes from production-colormap-merid-rho
  \caption{A meridional slice of temperature in the hypermassive neutron
    star and disk configuration (Sec.~\ref{ssec:test_disk_comparison}).}
  \label{fig:nsns_temp_merid}
\end{figure}

\begin{figure}
  \includegraphics[width=\columnwidth]{production-colormap-equat-temp}
  % see notes from production-colormap-merid-rho
  \caption{An equatorial slice of temperature in the hypermassive neutron
    star and disk configuration (Sec.~\ref{ssec:test_disk_comparison}).}
  \label{fig:nsns_temp_equat}
\end{figure}

We extrapolate the fluid and M1 radiation data from
the domain shown in Figs.~\ref{fig:nsns_rho_merid}--\ref{fig:nsns_temp_equat}
to a larger ray tracing domain by setting all fluid and M1 radiation variables
to their floor values outside the smaller domain.
This simple extrapolation is adequate for ray tracing since
the neutrinos are almost entirely free-streaming outside the smaller domain.
Since the metric data were evolved on this larger domain no extrapolation is
needed for them.
The larger domain is represented as a pseudospectral grid
composed of a sphere with concentric shells extending to $r\approx1400$~km.
Radial grid spacings are $\Delta r \approx 0.15$~km in the star
and $\Delta r \approx 2.5$~km in the disk,
with 12 cells spanning polar angles $\theta\in[0,\pi]$
and 24 cells spanning azimuthal angles $\phi\in[0,2\pi]$.
Though a pseudospectral representation of non-smooth hydrodynamic data
introduces some Gibbs-like oscillations in the variables,
we choose to use this representation instead of the mesh-refined
finite difference grid of the evolution because the pseudospectral
representation uses much less memory:
95~MB in pseuspectral vs.\ 3~GB in finite difference representation.
Obviously, memory loads of this order are not insurmountable;
but they would require some modifications to our
volume data interpolation infrastructure.
For the purposes of this analysis the pseudospectral representation is adequate.

Opacities are computed from the LS220 equation of state using
\lstinline{NuLib} and are stored as a table covering
energy, density, temperature, and composition ranges
spanning $\varepsilon\in[1,280.5]$~MeV logarithmically,
$\varrho\in[10^6,3.2\times10^{15}]\,{\rm g}\,{\rm cm}^{-3}$ logarithmically,
$T\in[0.05,150]$~MeV logarithmically, and 
$Y_e\in[0.035,0.55]$ linearly,
with grid extents $\{18,82,65,51\}$ respectively.

We place stationary (i.e.\ Eulerian) observers at fixed coordinate radius
$r=250$~km in the $y$-$z$ plane along an arc at $N_\theta$ positions
distributed linearly in $\cos\theta$ over the northern hemisphere
using $\theta_i=\cos^{-1}\left(1-i/(N_\theta-1)\right)$
for $i=0,1,2,\ldots N_\theta-1$.
We choose $N_\theta=30$.

Though the ray tracing sampling of $f$ is done in full general relativity,
in our calculation of moments at the observer positions,
we make the simplifying assumption of Minkowski spacetime.
This assumption introduces errors in our moments of order
$GM/r=4\,{\rm km}/250\,{\rm km}$
\footnote{TODO: too simplistic}
(with $G$ the gravitational constant and $M$ the central mass),
or $\sim 1\%$.

Each observer samples the distribution function over a uniform grid
in energy, cosine of polar angle, and azimuthal angle with extents
$\{N_{\tilde{\varepsilon}},N_A, N_B\}$ spanning the ranges
$\tilde{\varepsilon}\in(0,100)$~MeV,
$\cos A\in\left((\cos A)_{\rm min}, 1\right)$, and
$B\in[0,2\pi)$.
The integrals for fluxes of number and energy
(Eqns.~\ref{eqn:Ha}, \ref{eqn:Ka}, and \ref{eqn:J_H_S_eps_integrated})
become sums over all rays.
Taking these fluxes in the radial direction,
and using the midpoint rule to convert the integral to a sum, we have
\begin{align}
  \label{eqn:Kr_minkowski}
  \tilde{K}^r &= \Delta
  \sum\limits_{m=0}^{N-1}
  \tilde{\varepsilon}_m^2 \cos A_m f_m,\\
  \label{eqn:Hr_minkowski}
  \tilde{H}^r &= \Delta
  \sum\limits_{m=0}^{N-1}
  \tilde{\varepsilon}_m^3 \cos A_m f_m,
\end{align}
with $f_m \equiv f(\tilde{\varepsilon}_m,\cos A_m, B_m)$,
$\Delta \equiv \Delta\tilde{\varepsilon}\, \Delta(\cos A)\, \Delta B$,
$N \equiv N_{\tilde{\varepsilon}}\, N_A\, N_B$,
and $m$ the index labeling each ray.
Average energies are given by
$\langle\tilde{\varepsilon}\rangle(\theta)=H^r(\theta)/K^r(\theta)$.
We choose $N_{\tilde{\varepsilon}}=30$, $N_A=150$, and $N_B=20$.
To maintain high angular resolution, we only sample rays that pass within
approximately 120~km of the star's center by setting $(\cos A)_{\rm min}=0.88$.

We also combine measurements from all observers to estimate
total luminosities and averages over the sky.
Since we have chosen an arc of observers isolated to the northern hemisphere and
the $y$-$z$ plane, we may extend these data to the full sky by assuming that
emission is azimuthally symmetric and reflection symmetric across the
equatorial plane. Figs.~\ref{fig:nsns_rho_merid}--\ref{fig:nsns_temp_equat}
indicate the approximate validity of these assumptions at 11~ms after merger.
Total luminosities are then computed as integrals of the radial fluxes
over $\cos\theta$.
Using the trapezoid rule the number and energy luminosities become sums:
\begin{align}
  \label{eqn:luminosity_R}
  R &= C
  \left(\frac{1}{2}(\tilde{K}^r_0+\tilde{K}^r_{N_\theta-1})
  +\sum\limits_{m=1}^{N_\theta-2}\tilde{K}^r_m\right),\\
  \label{eqn:luminosity_L}
  L &= C
  \left(\frac{1}{2}(\tilde{H}^r_0+\tilde{H}^r_{N_\theta-1})
  +\sum\limits_{m=1}^{N_\theta-2}\tilde{H}^r_m\right),
\end{align}
with $C \equiv 4\pi r^2\, \Delta(\cos \theta)$,
\footnote{TODO: source of GR error in coordinate $r$}
$\tilde{K}^r_m \equiv \tilde{K}^r(\theta_m)$, and
$\tilde{H}^r_m \equiv \tilde{H}^r(\theta_m)$.
Average energies over the whole sky are then
$\langle\tilde{\varepsilon}\rangle=L/R$.

Tab.~\ref{tab:nsns_rt_vs_m1} compares the all-sky
luminosities and average energies from ray tracing and from
the M1 transport simulation,
which serves for qualitative comparison.
Even if M1 and ray tracing methods both provide faithful measurements
of all-sky luminosities, we do not expect perfect agreement since the two
treatments differ fundamentally. In addition to differences in the physics
such as the treatment of inelastic scattering,
M1 fluxes are integrated over the outer boundary of the finite-difference grid
at radii as small as 75~km,
whereas the ray tracing fluxes are integrated over a sphere at radius 250~km,
introducing a time lag between the two measurements of under a millisecond.
Furthermore uncertainties from the M1 simulation are around
15\% for energy luminosities and
10\% for average energies \cite[Sec.~A.6]{fouc2016-m1_evolve_n}.

\begin{table}%[h]
  % Total fluxes over the sky integrated in google sheet: 20171031-theta_distribution_integrations
  % RTDATA_NO=$DIR/PositionDistributions-nu?-250km-Lev2-e30a150b20-noscat-04.dat
  % RTDATA_SC=$DIR/PositionDistributions-trim-nu?-250km-Lev2-e30a150b20-scat-04.dat
  % DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20170324-ray_tracing_nsns/Run03/PositionDistributions
  \caption{
    Comparison of total luminosities and average energies
    of the hypermassive neutron star configuration
    (presented in Sec.~\ref{ssec:test_disk_comparison}).
    The methods are ray tracing using the \emph{noscat} method `RT noscat',
    ray tracing using the \emph{gray} \emph{scat} method `RT scat',
    and the M1 transport simulation `M1'.
    The ray tracing totals were computed from sums over
    observers placed in the $y$-$z$ plane.
    M1 values are taken from \cite[Figs. 7, 9, 10]{fouc2016-m1_evolve_n}.
    Energy luminosities have units $10^{52}\,{\rm erg}\,{\rm s}^{-1}$,
    number luminosities $10^{57}\,{\rm s}^{-1}$, and
    average energies MeV.
    The $\nu_x$ luminosities are per-species:
    multiply by four to get the total heavy-lepton neutrino luminosities.
  }
  \label{tab:nsns_rt_vs_m1}
  \begin{tabularx}{\columnwidth}{X r r r}
    & {\bf RT noscat} & \,\,{\bf RT scat} & \,\,{\bf M1} \\
    \hline
    $L_{\nu_e}$                                 & 5.87 & 5.40 & 5 \\
    $L_{\bar{\nu}_e}$                           & 9.70 & 10.7 & 12 \\
    $L_{\nu_x}$                                 & 23.6 & 12.5 & 12 \\
    $R_{\nu_e}-R_{\bar{\nu}_e}$                 & -0.91 & -1.68 & -2 \\
    $\langle \varepsilon_{\nu_e} \rangle$       & 12.7 & 12.0 & 12 \\
    $\langle \varepsilon_{\bar{\nu}_e} \rangle$ & 16.0 & 14.8 & 15 \\
    $\langle \varepsilon_{\nu_x} \rangle$       & 34.2 & 23.3 & 26 \\
    \hline
  \end{tabularx}
\end{table}

Because the model was evolved with a gray M1 transport scheme,
the only \emph{scat} method we can use in our ray tracing is the
\emph{gray} method.
As in the post-bounce configuration in Sec.~\ref{ssec:test_collapse},
the \emph{scat} treatment is more faithful than the \emph{noscat} treatment,
significantly so for the heavy-lepton neutrinos.
For example, with scattering turned off the ray tracing and M1 measurements
of $L_{\nu_x}$ disagree by 200\%;
but with scattering turned on agreement is within 10\%.
And as the ray tracing and M1 measurements both show,
$R_{\bar{\nu}_e}$ dominates over $R_{\nu_e}$ significantly,
because the disrupted neutron star material is releptonizing.

The distributions of radial fluxes over observer position are shown
in Figs.~\ref{fig:nsns_theta_distrib_Kr} and \ref{fig:nsns_theta_distrib_Hr}.
With the \emph{noscat} treatment the $\nu_x$ luminosities are relatively
constant in $\theta$, because the disk is optically thin to $\nu_x$,
and most of the neutrinos come from the hypermassive neutron star,
which is roughly spherical.
The slight upward trend in $\nu_x$ luminosities with $\theta$ may be due to
asymmetries in the configuration.
It is not due to the Doppler shift or relativistic beaming
from the rapid rotation of the star,
a hypothesis we tested by setting $u_i=0$ and $W=1$ in the ray
tracing equations.
Because of their larger absorption cross section,
the disk is not optically thin to $\nu_e$ and $\bar{\nu}_e$,
and observers within view of the hot hypermassive neutron star measure
the largest radial fluxes of these species.
\begin{figure}
  %\includegraphics[width=\columnwidth]{theta_distrib-250km-Kr}
  \resizebox{\columnwidth}{!}{\input{theta_distrib-250km-Kr.tex}}
  % Data_a:   $DIR/PositionDistributions/*Lev2-e30a150b20-noscat-04*
  % Data_s:   $DIR/PositionDistributions/*Lev2-e30a150b20-scat-04*
  % Data gathered using: $DIR/CollectPositionDistributions.sh
  % Script: $DIR/GnuScripts/plot-theta_distributions-Kr-fixed_r.gnu
  % w/ DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20170324-ray_tracing_nsns/Run03
  \caption{Radial number fluxes, $\tilde{K}^r(\theta)$,
    as defined in Eqn.~\ref{eqn:Kr_minkowski},
    in the hypermassive neutron star configuration
    (presented in Sec.~\ref{ssec:test_disk_comparison}).
    Sampled for observers at fixed coordinate distances from the center
    of the star, with $r=250\,{\rm km}$.
  }
  \label{fig:nsns_theta_distrib_Kr}
\end{figure}

\begin{figure}
  %\includegraphics[width=\columnwidth]{theta_distrib-250km-Hr}
  \resizebox{\columnwidth}{!}{\input{theta_distrib-250km-Hr.tex}}
  % see notes from theta_distrib-250km-Kr
  % Script: $DIR/GnuScripts/plot-theta_distributions-Hr-fixed_r.gnu
  \caption{Same as Fig.~\ref{fig:nsns_theta_distrib_Kr} but radial energy
    fluxes $\tilde{H}^r(\theta)$ as defined in Eqn.~\ref{eqn:Hr_minkowski}.
  }
  \label{fig:nsns_theta_distrib_Hr}
\end{figure}

When we turn on scattering, the disk is no longer transparent to $\nu_x$,
and $\nu_x$ luminosities present the same qualitative $\theta$-dependence
as that of the other species.
In general, scattering increases luminosities for polar observers
\footnote{TODO: in compute-eta data not true anymore for $\nu_e$}
($\theta \sim 0^\circ$) because additional high energy neutrinos
from the hypermassive neutron star are scattered by the disk back up to
the observer.
For observers in the equatorial plane ($\theta \sim 90^\circ$)
the effect of scattering is a decrease in energy luminosities,
since some of the highest energy neutrinos from deep in the interior
are scattered away from the line of sight by the disk.

Fig.~\ref{fig:nsns_theta_distrib_avg_eps} shows the distribution of
average energies of radial fluxes over observer positions.
With and without scattering, the average energies are highest
for observers near the polar axis, since polar observers get a
direct view of the hot hypermassive neutron star.
(This trend is unexpectedly reversed for $\nu_x$,
and may be due to asymetries in the disk, as discussed above.)
Scattering decreases average energies across all observer positions,
as it did in the post-bounce configuration, or any configuration of
a hot interior surrounded by a scattering envelope.
The strength of the effect of scattering on the average energies of
the different species is seen to follow the ranking
$\nu_x>\bar{\nu}_e>\nu_e$ due to two factors:
the average energies of the spectra obey the same ranking,
and the thicknesses of the different species' scattering envelopes
also obey the same ranking.
As we showed in the toy models with a homogeneous background scattering field
the scattering contribution $f_{\rm SE}$ is
proportional to energy $\varepsilon^2$ and path length $s$
(Eqn.~\ref{eqn:fse_homogeneous_thin_limit}).
\begin{figure}
  %\includegraphics[width=\columnwidth]{theta_distrib-250km-avg_eps}
  \resizebox{\columnwidth}{!}{\input{theta_distrib-250km-avg_eps.tex}}
  % see notes from theta_distrib-250km-Kr
  % Script: $DIR/GnuScripts/plot-theta_distributions-avg_eps_from_fluxes-fixed_r.gnu
  \caption{Same as Fig.~\ref{fig:nsns_theta_distrib_Kr} but
    average energies of radial fluxes
    $\langle\tilde{\varepsilon}\rangle(\theta)=\tilde{H}^r(\theta)/\tilde{K}^r(\theta)$,
    as defined in Eqns.~\ref{eqn:Kr_minkowski} and \ref{eqn:Hr_minkowski}.
  }
  \label{fig:nsns_theta_distrib_avg_eps}
\end{figure}

In Figs.~\ref{fig:nsns_cosA_distrib_G_00deg}
and \ref{fig:nsns_cosA_distrib_G_90deg}, we show the distributions of neutrino
number density over incoming angle
for the observer on the polar axis $\theta=0^\circ$,
and in the equatorial plane $\theta=90^\circ$,
defined analagously to Eqn.~\ref{eqn:J_per_cosA}.
As in the case of the collapse profile in Sec.~\ref{ssec:test_collapse},
the dominant effect of elastic scattering on $\tilde{G}(\cos A)$
is to spread the distribution out to larger angles, by generally decreasing
the number of neutrinos coming from the core while increasing the number
of neutrinos coming from the disk.
\footnote{TODO: with compute-eta data also explain why $G_{\nu_e}>G_{\bar{\nu}_e}$
  for $\theta=0$ at $\cos A>0.995$.}

For the observer in the equatorial plane,
Fig.~\ref{fig:nsns_cosA_distrib_G_90deg}, the disk is so optically thick to
$\nu_e$ that there is very little difference between \emph{scat} and
\emph{noscat} treatments for incoming angles $\cos A \lesssim 0.95$
corresponding to the volume inside of $r \lesssim 80$~km.
Also the stepped temperature gradient in the disk's spiral arms visible in
Figs.~\ref{fig:nsns_temp_merid} and \ref{fig:nsns_temp_equat} presents as a
stepped heavy-lepton neutrino number density distribution in
Fig.~\ref{fig:nsns_cosA_distrib_G_90deg} in the \emph{noscat} treatment,
since the pair production rate due to $e^{-}e^{+}$ annihilation,
producing $\nu_x$, is especially sensitive to temperature, going as
$T^9$ \cite[Sec.~7]{burr2006-neutrino_opacities}.
This stepped distribution is not visible in $\nu_e$ and $\bar{\nu}_e$ emission,
since we ignore pair processes for these species due to the dominance of
absorption and emission processes, obeying a shallower temperature-dependence;
nor is it visible in the \emph{scat} treatment of $\nu_x$, since scattering
tends to smear the incoming angle of the emission;
nor is it visible for any species or treatment in
Fig.~\ref{fig:nsns_cosA_distrib_G_00deg}, since the integration over the
azimuthal angle $B$ averages out any spiral structure for the observer
on the polar axis.

\begin{figure}
  %\includegraphics[width=\columnwidth]{cosA_distrib-intensity-250km-00deg}
  \resizebox{\columnwidth}{!}{\input{cosA_distrib-G-250km-00deg.tex}}
  % Data:   $DIR/Lev2-e30a150b20-{,no}scat-04
  % Script: $DIR/plot-angular_distribution-number_density-compare.gnu
  % w/ DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20170324-ray_tracing_nsns/Run03/000Mo-000.0deg-250km
  \caption{Distribution of number density $\tilde{G}(\cos A)$ over incoming angle,
    defined analagously to Eqn.~\ref{eqn:J_per_cosA}.
    The integral of this quantity over $d\cos A$ gives the total number density
    measured by this observer on the rotation axis.
    Volume data from the hypermassive neutron star configuration
    (presented in Sec.~\ref{ssec:test_disk_comparison}).}
  \label{fig:nsns_cosA_distrib_G_00deg}
\end{figure}

\begin{figure}
  %\includegraphics[width=\columnwidth]{cosA_distrib-intensity-250km-90deg}
  \resizebox{\columnwidth}{!}{\input{cosA_distrib-G-250km-90deg.tex}}
  % Data:   $DIR/Lev2-e30a150b20-{,no}scat-04
  % Script: $DIR/plot-angular_distribution-number_density-compare.gnu
  % w/ DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20170324-ray_tracing_nsns/Run03/000Mo-090.0deg-250km
  \caption{Distribution of number density $\tilde{G}(\cos A)$ over incoming angle,
    as in Fig.~\ref{fig:nsns_cosA_distrib_G_00deg},
    but with the observer in the equatorial plane.}
  \label{fig:nsns_cosA_distrib_G_90deg}
\end{figure}

\section{Applications}
\label{sec:applications}

\subsection{Neutrino Oscillations at High Neutrino Densities}
\label{ssec:V_nunu}
In this section we examine the importance of elastic scattering in neutrino
flavor oscillation above the hypermassive neutron star--disk configuration
presented in Sec.~\ref{ssec:test_disk_comparison}.
Our treatment of the flavor evolution equation assumes flat space and small fluid
velocities.
In consequence we treat some of gauge-dependent quantities inconsistently,
and we ignore potentially important features of relativistic flavor evolution
near compact objects.
\footnote{TODO: cite Yang et Kneller}
Such a cursory treatment is sufficient, however, for the following qualitative
exploration.

The Boltzmann Equation (Eqn.~\ref{eqn:boltzmann}) is one limiting
form of the quantum kinetic equations for neutrinos
\cite{vlas2014-neutrino_qkes},
the limit where collisional mean free paths are much shorter
than oscillation lengths, i.e.\ $\varepsilon/\mathscr{K} \ll |H|^{-1}$
with $H$ the Hamiltonian describing coherent forward scattering interactions.
In this limit the neutrino density matrix takes the form
\begin{equation}
  \label{eqn:density_rho_0}
  \rho_0(\varepsilon,\ell_\mu) =
  \sum\limits_{\alpha}f^{\nu_\alpha}(\varepsilon,\ell_\mu)
  \ket{\nu_\alpha} \bra{\nu_\alpha},
\end{equation}
with $\ket{\nu_\alpha}$ the neutrino flavor eigenstates.
The neutrino state remains pure, or diagonal in the flavor basis.

But at the opposite limit, in the free-streaming regime,
coherence effects become important,
and the quantum kinetic equations take a Schr\"{o}dinger-like form:
\begin{equation}
  \label{eqn:schroedinger}
  i \frac{d}{ds} S = H S,
\end{equation}
where $ds$ is an interval of proper length traversed by the neutrino as
measured by our observer, and
\footnote{TODO: mention $\bar{\nu}$ eqn}
$S$ is the flavor evolution matrix describing a mixed state
\begin{equation}
  \label{eq:density_evolution}
  \rho = S \, \rho_0 \, S^\dagger.
\end{equation}
In this limit, we may decompose the Hamiltonian into vacuum,
matter, and neutrino contributions:
\begin{equation}
  \label{eqn:decompose_H}
  H = H_{\rm V} + H_e + H_{\nu\nu}.
\end{equation}
Explicit formulas for these terms may be found in the oscillation literature,
e.g.\ \cite{duan2009-review},
but for clarity we give their order-of-magnitude scales here:
\begin{align}
  \label{eqn:scale_H_V}
  |H_{\rm V}| &\sim \frac{\Delta m^2}{\varepsilon}, \\
  \label{eqn:scale_H_e}
  |H_e|       &\sim G_{\rm F}\left|n_{e^-}-n_{e^+}\right|, \\
  \label{eqn:scale_H_nunu}
  |H_{\nu\nu}|&\sim G_{\rm F} \left|G_{\nu_e}-G_{\bar{\nu}_e}\right|,
\end{align}
where $\Delta m^2$ is the mass-squared differences between neutrino mass
eigenstates,
$G_{\rm F}$ is the Fermi coupling constant,
$n_{e^-}$ and $n_{e^+}$ are the electron and positron number densities,
and $G_{\nu_e}$ and $G_{\bar{\nu}_e}$ the electron neutrino and antineutrino
number densities defined in Eqn.~\ref{eqn:G}.
For the scale of $H_{\nu\nu}$ we have only included the isotropic
components of the neutrino fields.

In regimes in which $|H_{\nu\nu}|\ll|H_{\rm V},H_e|$,
the flavor evolution along the ray is soluble, and reveals a rich
and physically important phenonenology including
vacuum, solar, atmospheric, and terrestrial oscillations,
as well as oscillations in supernova envelopes.
Where neutrino densities are relatively high, however,
as in neutron star mergers, the problem must be solved globally.
\footnote{TODO: be clearer}
To date, no method has been devised to handle this problem.

However, we can solve a similar but tractable problem along a single ray.
Assume all the neutrino rays intersecting an event along a given test ray
have undergone the same flavor evolution history as that of the test ray,
i.e.\ $S$ is the same for all rays sharing that event.
This is the so-called single-angle approximation.
\footnote{TODO: cite someone}
\footnote{TODO: say something about its validity}
Though widely used in the supernova oscillation literature
and shown to be qualitatively faithful in those environments,
\footnote{TODO: cite vlasenko?}
a recent study has shown that the single-angle approximation
may mask certain instabilities in the flavor evolution
\cite{wu2017-fast_neutrino_conversions}.

Within the single-angle approximation,
solving for the flavor evolution history along a given test ray
requires knowledge of the unoscillated neutrino contribution to the Hamiltonian
along that trajectory.
This is a function of the unoscillated neutrino self-interaction potential,
which for the $\alpha$-th flavor is:
\begin{equation}
  \label{eqn:V_nu_alpha}
  V_{\nu_\alpha,0}(\varepsilon,\ell_\mu) =
  \frac{\sqrt{2}G_{\rm F}}{(2\pi)^3} \varepsilon^2
  \oint d\Omega'(1-\varpi')f^{\nu_\alpha}(\varepsilon,\ell'_\gamma),
\end{equation}
with the test ray propagating in direction $\ell_\mu$,
ambient rays propagating in directions $\ell'_\gamma$,
and the cosine of the angle between these given by
$\varpi'=\psi^{\mu\gamma} \ell_\mu \ell'_\gamma$.
As in the moment equations (Eqns.~\ref{eqn:J}-\ref{eqn:Ka}),
the integral is taken over all directions $\ell'_\gamma$.

If conditions are right, a resonant flavor transition
introducing significant mixing may occur very near
the point where neutrinos begin free-streaming.
The matter-neutrino resonance
\cite{malk2012-mnr_1,malk2015-mnr_2,malk2016-mnr_3},
can occur where the matter potential and
neutrino self-interaction potentials cancel,
or where $V_e - V_{\nu\nu,0}=0$, with
\begin{align}
  \label{eqn:V_e}
  V_e &= \sqrt{2} G_{\rm F} \frac{\varrho}{m_N} Y_e, \\
  \label{eqn:V_nunu0}
  V_{\nu\nu,0} &= \int d\varepsilon
  \big(V_{\nu_e,0}(\varepsilon)-V_{\bar{\nu}_e,0}(\varepsilon)\big),
\end{align}
with $\varrho$ the rest density and $m_{N}$ the nucleon mass.
The matter potential $V_e$ is always positive;
and far outside the accretion disks formed in neutron star mergers,
in which the disrupted neutron-star matter is rapidly releptonizing,
the total unoscillated neutrino self-interaction potential $V_{\nu\nu,0}$
is large and negative.

We examine this effect in the post-merger configuration already analyzed
in Sec.~\ref{ssec:test_disk_comparison}.
We calculate the self-interaction potential along a radial coordinate trajectory
making an angle $\theta=25^\circ$ with the rotation axis,
at $N_s=7$ positions $r\in\{30,50,82,135,223,368,608\}$~km.
We sample distribution functions at each of these positions over a grid
with extents $N_\varepsilon=30$, $N_A=200$, $N_B=30$;
energies range over $\varepsilon\in(0,100)$~MeV;
polar angles range over $\cos A \in ((\cos A)_{\rm min}, 1)$, with
$(\cos A)_{\rm min}\in\{-1,-1,-1,0.488,0.849,0.947,0.981\}$
for each of the $N_s$ positions;
and azimuthal angles range over the whole sky $B\in[0,2\pi)$.
We interpolate the logarithm of the self-interaction potentials
(Eqn.~\ref{eqn:V_nu_alpha})
in path length $\log s$ along the ray,
by fitting a 3rd order spline with continuous derivative,
across the $N_s$ observation points.
For $r>608$~km we extrapolate the self-interaction potentials using the
geometric fall-off of $r^{-4}$ applicable to the far-field limit of the
self-interaction potential \cite{malk2016-mnr_3}.

The matter potential along this test ray is from an analytic wind
model qualitatively consistent with the densities in the
simulated volume (i.e.\ inside $r\sim70$~km) and asymptoting to the $r^{-2}$
density field of a spherical steady-state wind with a constant asymptotic
velocity.
The velocity and density fields of a spherical steady-state wind
obey the continuity equation $\rho(r)v(r)r^2=\rho_1 v_1 r_1^2$,
with $\rho_1$ and $v_1$ the density and velocity measured at some fiducial
radius $r_1$.
For velocity field we choose a phenomenological wind model used in the
oscillation literature\footnote{TODO: cite Surman \& McLaughlin ApJ 618, 2005}
\begin{equation}
  v(r) = v_1 \left(1-\frac{R}{r_1}\right)^{-\beta}
  \left(1-\frac{R}{r}\right)^{\beta},
\end{equation}
with $R$ the wind launch radius and $\beta$ an acceleration parameter.
This yields the following density field
\begin{equation}
  \rho(r)=\rho_1 C\left[\frac{r_1}{R}\right]
  \left(1-\frac{R}{r}\right)^{-\beta}\left(\frac{R}{r}\right)^2,
\end{equation}
with $C[a]=a^{2-\beta}(a-1)^\beta$.
We use the parameters $R=10$~km, $R_1=50$~km, $\beta=2$,
and $\rho_1=10^7\,{\rm g}\,{\rm cm}^{-3}$.
Additionally we impose a density cap of
$\rho_{\rm max}=10^{14}\,{\rm g}\,{\rm cm}^{-3}$ inside the radius $r_0=5$~km,
and smoothly interpolate densities between $r_0$ and $r_1$ with a cubic polynomial,
enforcing $C^0$ and $C^1$ continuity at the transitions.
The rest density from this model is plotted in Fig.~\ref{fig:test_traj_density}.
To translate this density field to a matter potential, we assume a constant
electron fraction of $Y_e=0.5$.

\begin{figure}
  \includegraphics[width=\columnwidth]{profile-density-000Mo-025deg}
  % Plot:    $DIR/RunBA/RunBA-025deg-noscat-production01/profile-density-000Mo-025deg.pdf
  % Script:  $DIR/nu_potentials-production.nb
  % Configs: $DIR/nu_potentials-wind_model.nb
  %          $DIR/RunBA/RunBA-025deg-noscat-production01/nu_potentials-config-production.nb
  % w/ DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20170412-self_interaction_potentials_M12_M12-LS220-M1/Run03/
  \caption{Assumed matter density along the neutrino test trajectory
    used to calculate the matter potential $V_e$.
    }
  \label{fig:test_traj_density}
\end{figure}

In fact a spherical steady-state wind model, though providing an adequate
backdrop for the qualitative study presented in this section,
is less than ideal for this post-merger configuration,
most obviously because in the 10~ms since merger,
ejecta with the greatest velocities around $0.3c$
will have reached no further than $r_{\rm max} \sim 10^8$~cm.
Additionally, the true radial profile of the ejecta from this merger
(which our computational model does not follow)
will have many more features inside this radius, including shock jumps.
This model density profile is adequate as a backdrop to our study here, however,
since we expect the remnant to present similar neutrino emission over a thermal
timescale of a few tens of milliseconds\footnote{cite Nouri et al 2018}
while the matter field propagates out to larger radii.

In Figs.~\ref{fig:V_nunu-noscat} and \ref{fig:V_nunu-scat}
we show the total unoscillated self-interaction potential
and its contributions from $\nu_e$ and $\bar{\nu}_e$ for a test neutrino
moving out along the radial coordinate trajectory described above.
Fig.~\ref{fig:V_nunu-noscat} shows these terms for the \emph{noscat} treatment,
and Fig.~\ref{fig:V_nunu-scat} for the \emph{scat} treatment.
Obviously, including the effects of elastic scattering tends to increase the
$\bar{\nu}_e$ contribution relative to the $\nu_e$ contribution,
in this case causing the self-interaction potential to be negative
along the entire trajectory.
This effect may be predicted from Fig.~\ref{fig:nsns_cosA_distrib_G_00deg},
which is calculated for a qualitatively similar observer, in the vacated polar
funnel of the disk:
in the \emph{scat} treatment $G_{\bar{\nu}_e}$ dominates over $G_{\nu_e}$ at
all angles $\cos A \lesssim 0.998$;
whereas in the \emph{noscat} treatment $G_{\bar{\nu}_e}$ only dominates over
$G_{\nu_e}$ over a range of forward-peaked angles $\cos A \in (0.965,0.998)$.
An additional factor supporting this trend is that neutrinos from the most
forward-peaked angles $\cos A \sim 1$ have a suppressed effect on the
self-interaction potential due to the $(1-\cos A)$ term arising in the integral
of Eqn.~\ref{eqn:V_nu_alpha}.

\begin{figure}
  \includegraphics[width=\columnwidth]{20171223-potentials-Ve_Vnue_Vnua_Vnu-000Mo-025deg-noscat}
  % Plot:   $DIR/RunBA/RunBA-025deg-noscat-production01/potentials-Ve_Vnue_Vnua_Vnu-000Mo-025deg.pdf
  % Data:   $DIR/000Mo-025deg-*km/Lev2-e30a200b30-noscat-04
  % Script: $DIR/nu_potentials-production.nb
  % Config: $DIR/RunBA/RunBA-025deg-noscat-production01/nu_potentials-config-production.nb
  % w/ DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20170412-self_interaction_potentials_M12_M12-LS220-M1/Run03/
  \caption{Neutrino oscillation potentials
    along a test trajectory
    originating at the surface of the hypermassive neutron star, and
    proceeding outward along a radial coordinate ray with angle
    $\theta=25^{\circ}$ with respect to the polar axis.
    The trajectory is parameterized by the coordinate length $s$.
    We plot the vacuum potential due to mass-squared differences (gray bands),
    the matter potential $V_e$ due to forward scattering on $e^-$ and $e^+$
    (dark green),
    and the self-interaction potential $V_{\nu\nu,0}$ due to forward
    scattering on ambient neutrinos (light green, solid where positive,
    dashed where negative);
    additionally we plot the $\nu_e$ and $\bar{\nu}_e$ components composing
    the self-interaction potential (light blue and dark blue respectively).
    This calculation was done with scattering turned off.
    The two gray bands at $V\sim10^{-22}$~erg and $V\sim10^{-24}$~erg
    indicating the two vacuum energy scales for 10--30~MeV neutrinos,
    set the positions of possible Mikheyev-Smirnov-Wolfenstein (MSW)
    or nutation resonances \cite{malk2012-mnr_1}.
    Inside the symmetric point at $s\sim100\,{\rm km}$ the total unoscillated
    self-interaction potential $V_{\nu\nu,0}$ is positive,
    i.e.\ $\nu_e$-dominated.
    }
  \label{fig:V_nunu-noscat}
\end{figure}

\begin{figure}
  \includegraphics[width=\columnwidth]{20171223-potentials-Ve_Vnue_Vnua_Vnu-000Mo-025deg-scat}
  % Plot:   $DIR/RunBA/RunBA-025deg-scat-production01/potentials-Ve_Vnue_Vnua_Vnu-000Mo-025deg.pdf
  % Data:   $DIR/000Mo-025deg-*km/Lev2-e30a200b30-scat-04
  % Script: $DIR/nu_potentials-production.nb
  % Config: $DIR/RunBA/RunBA-025deg-scat-production01/nu_potentials-config-production.nb
  % w/ DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20170412-self_interaction_potentials_M12_M12-LS220-M1/Run03/
  \caption{Same as Fig.~\ref{fig:V_nunu-noscat},
    but with scattering turned on.
    Unlike the case with scattering turned off,
    here the total unoscillated self-interaction potential $V_{\nu\nu,0}$
    is everywhere negative, i.e.\ $\bar{\nu}_e$-dominated.
    }
  \label{fig:V_nunu-scat}
\end{figure}

We also solve for the flavor evolution of this system,
integrating $S$ according to Eqn.~\ref{eqn:schroedinger},
as described in \cite{zhu2016-mnr_nsns_remnant}.
In Figs.~\ref{fig:survival-noscat} and \ref{fig:survival-scat}
we show the survival probabilities for $\nu_e$ and $\bar{\nu}_e$
for the two treatments.
\footnote{TODO: update this paragraph to new compute-eta data}
The survival probabilities are the terms of
the flavor evolution matrix $S$ that describe how much of the original flux
of a given flavor survives a transition.
\footnote{TODO: define explicitly}
Where $P_{\nu_e\rightarrow\nu_e}$ vanishes, for example in
Fig.~\ref{fig:survival-noscat}, all the original $\nu_e$ content of the beam
is oscillated into $\nu_\mu$ and $\nu_\tau$; but the $\nu_e$ density doesn't
vanish altogether,
since some $\nu_\mu$ and $\nu_\tau$ oscillates into $\nu_e$.

For our assumed matter potential along this one test trajectory
we see that a matter-neutrino resonance only occurs in the \emph{scat} case;
though with a slightly higher matter potential it could occur in both cases,
closer to the disk in the \emph{noscat} case than the \emph{scat} case.
For this particular test trajectory we only see a standard and not a symmetric
matter neutrino resonance;
these two classes of matter-neutrino resonance
are described for example in \cite{malk2016-mnr_3,vaan2016-uncovering_mnr}.
The difference we see in the oscillation physics between the \emph{noscat}
and \emph{scat} cases along this test trajectory highlights the importance of
scattering physics in the problem of neutrino oscillations in
neutron star mergers.

\begin{figure}
  \includegraphics[width=\columnwidth]{fig-survival_prob-noscat_nh}
  % Flavor evolution by Yonglin, see emails 7.19.17-7.20.17
  % Plot: $DIR/31-900kmNHnoScatterPot.eps
  % w/ DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20170412-self_interaction_potentials_M12_M12-LS220-M1/Run03/RunAB/probabilityPotential
  \caption{Neutrino survival probabilities for the \emph{noscat} case.
    \footnote{TODO: Replace; this is from the old interp-eta data}
    In the lower panel we reproduce the unoscillated self-interaction potential
    $V_{\nu\nu,0}$ from Fig.~\ref{fig:V_nunu-noscat} and a matter potential
    computed assuming a very simple steady state wind consistent with the
    densities within the simulation volume inside of $r<70$~km
    along this trajectory.
    In this case the neutrinos undergo a symmetric MNR transition
    inside of $r<50$~km,
    with the $e$-flavor neutrinos and antineutrinos fully converting to $\nu_x$.
  }
  \label{fig:survival-noscat}
\end{figure}

\begin{figure}
  \includegraphics[width=\columnwidth]{fig-survival_prob-scat_nh}
  % see notes from fig-survival_prob-noscat_nh
  % Plot: $DIR/130-800kmNHScatterPot.eps (normal hierarchy)
  \caption{Same as Fig.~\ref{fig:survival-noscat},
    but with scattering turned on.
    \footnote{TODO: Replace; this is from the old interp-eta data}
    In this case the neutrinos undergo a standard MNR transition
    outside of $r>200$~km,
    with the $e$-flavor neutrinos and antineutrinos
    partially converting to $\nu_x$.
    }
  \label{fig:survival-scat}
\end{figure}

\section{Conclusions}
\label{sec:conclusions}
Add some conclusions.
\footnote{TODO: note some improvements needed: implement implicit time-stepping,
  use finite-difference grid for fluid vars, improve gray spectral assumptions}

% ******************************************************************************
\appendix

\section{Definitions}
\label{sec:definitions}
We decompose the neutrino momentum into components parallel and orthogonal to
an observer's velocity $u_\beta$:
\begin{equation}
  \label{eqn:def_momentum_2}
  p_\beta = \varepsilon (u_\beta + \ell_\beta),
\end{equation}
with $u_\alpha\ell^\alpha=0$ and $\ell_\alpha\ell^\alpha=1$.

We use two possible fiducial observers to define the momentum
decomposition via Eqn.~\ref{eqn:def_momentum_2}: the Eulerian,
or normal observer $n_\mu=-\alpha \,\partial_\mu t$,
and the fluid, or comoving observer
$u^\mu=Wn^\mu+v_{\rm E}^\mu$.
Here $t$ is coordinate time and $\alpha$ is lapse
in the standard 3+1 decomposition of the metric
\begin{equation}
  \label{eqn:adm_metric}
  \psi_{\mu\gamma} \rightarrow
  \left(
  \begin{matrix}
    -\alpha^2 + \beta^i \beta_i  & \beta_i \\
    \beta_j                      & g_{ij}
  \end{matrix}
  \right).
\end{equation}
We have also introduced
the fluid Lorentz factor $W=-n_\mu u^\mu=\alpha u^t$,
its Eulerian velocity $v_{\rm E}^\mu=g^\mu_\lambda u^\lambda$
(distinct from its coordinate velocity $v^\mu=u^\mu/u^t$),
and the projection tensor orthogonal to the normal observer
$g^\mu_\lambda=\psi^\mu_\lambda+n^\mu n_\lambda$,
whose spatial part is the 3-metric, $g_{ij}$.

We specify the neutrino direction in the observer's frame
with two spherical polar angles
($a$,$b$) with respect to the simulation cartesian coordinates
\begin{align}
  \ell_\alpha &\rightarrow q(s,\Omega_i), \\
  \Omega_i &\rightarrow (\sin a \cos b,\sin a\sin b,\cos a),
\end{align}
or alternatively the two spherical polar angles
($A$,$B$) with respect to rotated coordinates
\begin{equation}
  \label{eqn:def_direction_primed_2}
  \Omega_{i'} \rightarrow  (\sin A \cos B,\sin A\sin B,\cos A).
\end{equation}
The two coordinate systems are related by a standard Euler rotation of first
$\phi$ about the $z$-axis, then $\theta$ about the rotated $y$-axis,
with $\phi$ and $\theta$ the azimuthal and polar position of the observer.
Expressed algebraically:
\begin{align}
  \label{eqn:rotation_connection}
  \Omega_x &= \Omega_{x'}\cos\theta\cos\phi-\Omega_{y'}\sin\phi+\Omega_{z'}\sin\theta\cos\phi,\nonumber\\
  \Omega_y &= \Omega_{x'}\cos\theta\sin\phi+\Omega_{y'}\cos\phi+\Omega_{z'}\sin\theta\sin\phi,\nonumber\\
  \Omega_z &= \Omega_{x'}\sin\theta+\Omega_{z'}\cos\theta.
\end{align}

The scale factors $q$ and $s$ are functions of the neutrino
direction $\Omega_i$, the observer's velocity $u_\alpha$,
and the spacetime metric.
In the case of a fluid observer, specified by an arbitrary $W$ and $u_i$,
$q$ and $s$ are given by
\begin{align}
  \label{eqn:q_general}
  q
  &= W\alpha\Bigg(2\beta^i\Omega_i W^2(\beta^i\Omega_i-1) \nonumber\\
  &\qquad\qquad - 2\Omega_i u_j g^{ij} W \alpha(\beta^i\Omega_i-1)\nonumber\\
  &\qquad\qquad + \alpha^2\left((\Omega_i u_j g^{ij})^2+\Omega_i\Omega_j g^{ij}W^2\right)\Bigg)^{-1/2}\\
  \label{eqn:s_general}
  s
  &= \beta^i\Omega_i -\Omega_i u_j g^{ij}.
\end{align}
In the case of an Eulerian observer, $W=1$ and $u_i=0$,
and these expressions simplify considerably:
\begin{align}
  \label{eqn:q_eulerian}
  \tilde{q}
  &= \alpha\left(2\beta^i\Omega_i(\beta^i\Omega_i-1)+\alpha^2\Omega_i\Omega_j g^{ij}\right)^{-1/2}\\
  \label{eqn:s_eulerian}
  \tilde{s}
  &= \beta^i\Omega_i.
\end{align}
In the even simpler case of Minkowski spacetime, these expressions
reduce to $q=W/(1+s)$, $s=-\Omega_i u^i$,
$\tilde{q}=1$, $\tilde{s}=0$.

In addition to the fluid velocity and Lorentz factor, described above, the
other fluid state variables we use from our hydrodynamic simulations
are rest density $\varrho=m_b n_b$, temperature $T$, and electron fraction
\begin{equation}
  Y_e = \frac{n_{e^-} - n_{e^+}}{n_b},
\end{equation}
where $n_{e^-}$, $n_{e^+}$, and $n_b$ are the number densities of electrons
positrons, and baryons, and $m_b$ is the average baryon mass.

\section{Source Terms}
\label{sec:source_terms}
Here we present the sources comprising the right hand side of the
Boltzmann Equation (Eqn.~\ref{eqn:boltzmann}). The sources for the neutrino
distribution function $f(x^\alpha;p_\beta)$ arise from collision processes
producing, removing, or scattering to/from that point in phase space.
The weak interaction rates for each process involve integrals of
the neutrino distribution function $f(x^\alpha;p'_\beta)$ and that of the
antineutrino $\bar{f}(x^\alpha;p'_\beta)$ over a momentum volume
$dP'$ (Eqn.~\ref{eqn:dP}).

We follow
\cite{brue1985-core_collapse} and \cite[Sec.~4]{shib2011-truncated_moment}
by separating these processes into four categories:
\begin{equation}
  \label{eqn:four_sources}
  C[f] \equiv C_{\rm AE} + C_{\rm SE} + C_{\rm SI} + C_{\rm PP},
\end{equation}
representing charged-current absorption and emission, elastic scattering,
inelastic scattering, and the thermal pair processes of annihilation and
production. In this work, however, we only treat absorption/emission
and elastic scattering.
We seek to write each collision source linear in $f$:
\begin{align}
  \label{eqn:sources_cae}
  C_{\rm AE}
  &= \mathscr{E}_{\rm AE} + \mathscr{K}_{\rm AE}\, f, \\
  \label{eqn:sources_cse}
  C_{\rm SE}
  &= \mathscr{E}_{\rm SE} + \mathscr{K}_{\rm SE}\, f.
\end{align}

Each term
is computed by summing the weak interaction rates of the processes from the
relevant category given in Tab.~\ref{tab:neutrino_processes}.
We compute these rates in the rest frame of the fluid,
but because they are spacetime invariants
they take the same numerical value in any frame of reference
and are completely independent of our choice of fiducial observer $u^\alpha$
(see discussion around Eqn.~\ref{eqn:boltzmann_linear}).

Each source term is unique to the neutrino or antineutrino species modeled,
and consists of a sum over all of the processes contributing to that
category of interaction. For example:
$C^{\nu_e}_{{\rm AE}} = \sum_i\,C^{\nu_e}_{{\rm AE},i}$,
where $i$ labels the absorption/emission processes involving $\nu_e$
in Tab.~\ref{tab:neutrino_processes}.
By contrast $C^{\nu_x}_{{\rm AE}} = 0$,
\footnote{TODO: but we use a nonzero $C^{\nu_x}_{{\rm AE}}$}
where $\nu_x$ stands for any of the four heavy-lepton neutrinos and
antineutrinos.

We compute our rates using the open source neutrino interaction
library \lstinline{NuLib}~\footnote{\url{http://www.nulib.org}}.
We compile a table of sources defined over the four dimensions of
density, temperature, electron-fraction, and neutrino energy, and
interpolate quad-linearly to the points sampled along each ray.

\begin{table}%[h]
  \caption{
    We analyze neutrino interaction processes in terms of these categories.
    $\nu$ without a label represents a neutrino or antineutrino of any flavor,
    $N$ represents a nucleon $n$ or $p$,
    ${}^ZA$ represents a nucleus with mass number $A$ and charge $Z$, and
    $\gamma$ represents a high-energy photon.
    A prime indicates a change in that particle's energy.
  }
  \label{tab:neutrino_processes}
  \begin{tabularx}{\columnwidth}{X X}
    \hline \hline
    absorption/emission
    & $\nu_e + n \leftrightarrow e^- + p$                          \\
    & $\bar{\nu}_e + p \leftrightarrow e^+ + n$                    \\
    & $\nu_e + {}^AZ \leftrightarrow e^- + {}^A(Z+1)$              \\
    & $\bar{\nu}_e + {}^AZ \leftrightarrow e^+ + {}^A(Z-1)$        \\
    \hline
    elastic scattering
    & $\nu + N \leftrightarrow \nu + N$                            \\
    & $\nu + {}^AZ \leftrightarrow \nu + {}^AZ$                    \\
    \hline
    inelastic scattering
    & $\nu + e^- \leftrightarrow \nu' + e^{-'}$                    \\
    & $\nu + e^+ \leftrightarrow \nu' + e^{+'}$                    \\
    \hline
    thermal pair processes
    & $\nu + \bar{\nu} \leftrightarrow e^{-} + e^{+}$              \\
    & $\nu + \bar{\nu} + N + N \leftrightarrow N' + N'$            \\
    & $\nu + \bar{\nu} \leftrightarrow \gamma$                     \\
    % & $\nu_\alpha + \bar{\nu}_\alpha \leftrightarrow \nu_\beta + \bar{\nu}_\beta $ \\
    \hline \hline
  \end{tabularx}
\end{table}

\subsection{Absorption and emission via charged current}
\label{ssec:sources_ae}
At neutrino and thermal energies well below the masses of the muon or tauon
($m_{\mu}\sim100\,{\rm MeV}$)
only charged current processes involving $\nu_e$ and $\bar{\nu}_e$ are allowed.
For each of these processes, $i$, we may write an emission and absorption
coefficient as a function of the interaction cross-section
(e.g. \cite[Eqn.~A5]{brue1985-core_collapse}):
\footnote{TODO: explain Fermi-blocking}
\begin{align}
  \label{eqn:sources_ae_1}
  C_{{\rm AE},i}
  &= \varepsilon j_i (1-f) - \varepsilon \chi_{a,i} f \\
  \label{eqn:sources_ae_2}
  &= \varepsilon j_i - (\varepsilon j_i+\varepsilon \chi_{a,i}) f,
\end{align}
where $j$ is the emissivity and $\chi_a$ the absorption opacity.
Both $j$ and $\chi_a$ have dimension ${\rm length}^{-1}$ and represent the
number of neutrinos emitted or absorbed per length traveled.
In radiation transport formulations using specific intensities instead of
distribution functions, an emissivity $\eta$ having dimension
${\rm energy}\,{\rm length}^{-3}\,{\rm time}^{-1}\,{\rm energy}^{-1}\,{\rm steradian}^{-1}$
is more commonly used.
The two are related by $j=(2\pi)^3\eta/\varepsilon^3$.
Note that for brevity we have suppressed the energy-dependence of
the terms $j$, $\chi_a$, and the distribution functions.

In the special case of radiative equilibrium we know that the source
term vanishes: an equal number of neutrinos are emitted from
and absorbed by the matter for any length traversed.
We also know in this case that the neutrino distribution function must be
$f^{\rm eq}$, the equilibrium Fermi-Dirac distribution function of
Eqn.~\ref{eqn:feq}.
With these facts we can rerrange Eqn.~\ref{eqn:sources_ae_1} to give us
Kirchoff's Law:
\begin{align}
  \label{eqn:kirchoffs_law}
  j_i &= \frac{\chi_{a,i}}{1-f^{\rm eq}}f^{\rm eq}, \\
  \label{eqn:stimulated_opacity}
  &= \chi^*_{a,i}\,f^{\rm eq},
\end{align}
where in Eqn.~\ref{eqn:stimulated_opacity} we have introduced the opacity
corrected for stimulated absorption, $\chi^*_{a,i}$.

Using these expressions and computing the sum over stimulated opacities
$\chi^*_{a,i}=\sum_i \chi^*_{a,i}$,
the invariant emissivity and opacity for absorption/emission are
\begin{align}
  \label{eqn:ae_emissivity_summed_2}
  \mathscr{E}_{{\rm AE}} &= \varepsilon \, \chi^*_a \, f^{\rm eq}, \\
  \label{eqn:ae_opacity_summed_2}
  \mathscr{K}_{{\rm AE}} &= \varepsilon \, \chi^*_a.
\end{align}
These expressions are equivalent to those in
Eqns.~\ref{eqn:ae_emissivity_summed} and \ref{eqn:ae_opacity_summed}.

We use the above treatment for $\nu_e$ and $\bar{\nu}_e$ only;
the $\mu$ and $\tau$ neutrinos and antineutrinos do not participate in
charged current absorption/emission interactions at these temperatures
and energies.
However we do use an effective stimulated absorption opacity $\chi^*_a$
for the heavy-lepton neutrinos, computed by \lstinline{NuLib} as
described in \cite{ocon2015-gr1d_with_nu} which follows
\cite{brue1985-core_collapse,burr2006-neutrino_opacities}.
This is described in App.~\ref{ssec:sources_pp}.

\subsection{Elastic scattering}
\label{ssec:sources_se}
Neutrino scattering on particles of mass much greater than $\varepsilon$
(i.e. nucleons) is essentially iso-energetic. Following 
\cite[Eqn.~A8]{brue1985-core_collapse} or
\cite[Eqn.~4.20]{shib2011-truncated_moment} the collision term for the
$i$-th process takes the form
\begin{equation}
  \label{eqn:sources_se_1}
  C_{{\rm SE},i}(\ell_\alpha)
  = \frac{\varepsilon^3}{(2\pi)^3}
  \oint d\Omega' R_{{\rm SE},i}(\omega')
  \Big(f(\ell'_\beta) - f(\ell_\alpha)\Big)
\end{equation}
where $R_{\rm SE,i}(\omega')$ is the scattering kernel for the $i$-th process
from direction $\ell'_\beta$ to direction $\ell_\alpha$
having dimension ${\rm energy}^{-1}$,
and the cosine of the scattering angle is
$\omega' \equiv \psi^{\alpha\beta}\ell_\alpha\ell'_\beta$, with
$\psi^{\alpha\beta}$ the inverse of the spacetime metric.
Note that for brevity in Eqn.~\ref{eqn:sources_se_1} we have suppressed the
energy dependence of all of the terms.

It is customary to approximate the scattering kernel to linear order in
$\omega$ (as in~\cite[Eqn.~4.21]{shib2011-truncated_moment}):
\begin{equation}
  \label{eqn:se_kernel_expansion}
  R_{{\rm SE},i}(\varepsilon,\omega) \approx
  R_{{\rm SE},i}^0(\varepsilon) + \omega \, R_{{\rm SE},i}^1(\varepsilon).
\end{equation}
Using this expansion, and the moments defined in Eqns.~\ref{eqn:J}
and \ref{eqn:Ha} we may write Eqn.~\ref{eqn:sources_se_1}:
\begin{equation}
  \label{eqn:sources_se_2}
  C_{{\rm SE},i}
  = \varepsilon \chi_{s,i} \Phi_i - \varepsilon \chi_{s,i} f,
\end{equation}
where we have introduced the scattering opacity $\chi_{s,i}$
and the background scattering field $\Phi$:
\begin{align}
  \label{eqn:opacity_s}
  \chi_{s,i}(\varepsilon)
  &= \frac{4\pi\varepsilon^3}{(2\pi)^3}R_{{\rm SE},i}^0(\varepsilon), \\
  \label{eqn:background_phi}
  \Phi_i(\varepsilon,\ell_\alpha)
  &= \frac{(2\pi)^3}{\varepsilon^3}\frac{1}{4\pi}
  \left(J(\varepsilon)+
  \frac{\chi^{\rm tr}_{s,i}(\varepsilon)}{\chi_{s,i}(\varepsilon)}
  \ell_\alpha H^\alpha(\varepsilon)\right),
\end{align}
with $\chi^{\rm tr}_{s,i}$ the transport scattering opacity,
\footnote{TODO: drop anisotropic term and defend}
\begin{equation}
  \label{eqn:opacity_s_tr}
  \chi^{\rm tr}_{s,i}(\varepsilon)
  = \frac{4\pi\varepsilon^3}{(2\pi)^3}R_{{\rm SE},i}^1(\varepsilon).
\end{equation}
In our treatment we only retain the isotropic contribution to the
background field:
\begin{equation}
  \Phi(\varepsilon)
  = \frac{(2\pi)^3}{\varepsilon^3}\frac{1}{4\pi}
  J(\varepsilon).
\end{equation}
\footnote{TODO: defend}

Thus computing sums over the opacities
$\chi_s\equiv\sum_i\chi_{s,i}$ and
$\chi^{\rm tr}_s\equiv\sum_i\chi^{\rm tr}_{s,i}$,
the invariant emissivity and opacity for elastic scattering are
\begin{align}
  \label{eqn:se_emissivity_summed}
  \mathscr{E}_{\rm SE}
  &= \varepsilon \, \chi_s \, \Phi, \\
  \label{eqn:se_opacity_summed}
  \mathscr{K}_{\rm SE}
  &= \varepsilon \, \chi_s.
\end{align}

We have made the energy-dependence of the background scattering field
explicit in Eqn.~\ref{eqn:background_phi}.
When we compute $\Phi(\varepsilon)$ using $J(\varepsilon)$ and
$H^\alpha(\varepsilon)$ from a multi-group
M1 transport evolution, we call this the \emph{spectral} method.

However moment evolutions with multiple energy groups are still rare.
Most simulations employ a gray moment scheme,
evolving only the energy-integrated moments, $J$ and $H^\alpha$
and sometimes the number density $G$.
If such is the case we resort to the \emph{gray} method,
by reconstructing the energy-dependent source terms
from gray moments, assuming a diluted Fermi-Dirac spectrum.
We use the following procedure:
\begin{enumerate}
\item
  Interpolate the fluid temperature, $T$,
  equilibrium neutrino chemical potential, $\eta_\nu$,
  and evolved neutrino energy and number densities in the fluid frame,
  $J$ and $G$, from the simulation grid.
\item
  Compute the average neutrino energy in the fluid frame
  \begin{equation}
    \langle \varepsilon \rangle \equiv J/G.
  \end{equation}
\item
  Compute the neutrino spectral temperature, assuming equilibrium with the
  fluid
  \begin{equation}
    \label{eqn:spectral_temp}
    T_\nu = \langle \varepsilon \rangle
    \frac{\mathscr{F}_2(\eta_\nu)}{\mathscr{F}_3(\eta_\nu)},
  \end{equation}
  where $\mathscr{F}_b$ is the Fermi integral
  \begin{equation}
    \label{eqn:fermi_integral}
    \mathscr{F}_b(\eta) = \int_0^\infty dx \, x^b (1+e^{x-\eta})^{-1}.
  \end{equation}
  We implement the Fermi integrals using the analytical approximants from
  \cite{taka1978-beta_rates}.
\item
  Assume the background neutrino fields have the same total density
  as the evolved moments
  \begin{align}
    \label{eqn:J_from_gray}
    J(\varepsilon) &=
    J \frac{\varepsilon^3}{T_\nu^4 \,\mathscr{F}_3(\eta_\nu)}
    (1+\exp(\varepsilon/T_\nu-\eta_\nu)^{-1}, \\
    \label{eqn:H_from_gray}
    H^\mu(\varepsilon) &=
    H^\mu \frac{\varepsilon^3}{T_\nu^4\, \mathscr{F}_3(\eta_\nu)}
    (1+\exp(\varepsilon/T_\nu-\eta_\nu)^{-1}.
  \end{align}
\end{enumerate}

Note that to avoid floating point errors for very large negative $\eta$,
we employ the asymptotic form of Eqn.~\ref{eqn:fermi_integral}:
$\lim_{\eta\ll-1}\mathscr{F}_b(\eta) = b!\,e^\eta$.
Thus, for $\eta<-10$, we use the limiting forms of
Eqns.~\ref{eqn:spectral_temp},
\ref{eqn:J_from_gray},
and \ref{eqn:H_from_gray}:
\begin{align}
    T_\nu &= 
    \frac{\langle \varepsilon \rangle}{3}, \\
    J(\varepsilon) &=
    J \frac{\varepsilon^3}{6 T_\nu^4} e^{-\varepsilon/T_\nu}, \\
    H^\mu(\varepsilon) &=
    H^\mu \frac{\varepsilon^3}{6 T_\nu^4} e^{-\varepsilon/T_\nu}.
\end{align}

When we use the moments from Eqns.~\ref{eqn:J_from_gray} and
\ref{eqn:H_from_gray} in Eqn.~\ref{eqn:background_phi},
we call this the \emph{gray} method.

\subsection{Inelastic scattering}
\label{ssec:sources_si}
Neutrino scattering off of electrons is inelastic, changing the
magnitude and direction of the species' momenta.
\footnote{TODO: cite \cite{keil2003-pinched_spectra} or someone}
A similar formalism to the above could be used to derive source terms
for inelastic scattering.
\footnote{TODO: compare relative amounts of inelastic to elastic scattering}
We save that for future work, pointing out here that inelastic scattering
treated this way is very sensitive to the energy-dependence of
the background field.

\subsection{Thermal pair annihilation and production}
\label{ssec:sources_pp}
We do not include thermal pair processes within the above formalism,
\footnote{TODO: explain this choice, and cite \cite{keil2003-pinched_spectra} or someone}
but as mentioned in App.~\ref{ssec:sources_ae}, we do include
these processes in an effective absorption opacity for $\mu$ and $\tau$
neutrinos and antineutrinos. \lstinline{NuLib} follows
\cite{brue1985-core_collapse,burr2006-neutrino_opacities}
by computing the emissivity of the pair processes in
\footnote{TODO: describe better}
Tab.~\ref{tab:neutrino_processes}, and applying Kirchoff's law,
Eqn.~\ref{eqn:kirchoffs_law}, to convert this to an opacity.
We only use this effective opacity for heavy-lepton neutrinos because
it is subdominant to the true absorption opacity for
$\nu_e$ and $\bar{\nu}_e$.

\subsubsection*{Acknowledgements}
This research is supported in part by the Joint Institute for Nuclear
Astrophysics---Center for the Evolution of the Elements (JINA-CEE)
and the National Science Fondation under grant numbers X. \footnote{TODO: grant nos?}
We thank Sherwood Richers for providing his neutrino transport code
\lstinline{Sedonu} for comparisons,
Fran\c{c}ois H\'{e}bert for his timely help in finding and
correcting bugs in the time-interpolation algorithm,
and Luke Roberts for helpful discussions.

% ******************************************************************************
\bibliography{references}

\end{document}

### 2001-01-01 paper.tex
Matt's edits through p. 15: move acknowledgements
Fix sign error in equations in appendix

@@
 
+\subsubsection*{Acknowledgements}
+This research is supported in part by the Joint Institute for Nuclear
+Astrophysics---Center for the Evolution of the Elements (JINA-CEE)
+and the National Science Fondation under grant numbers X. \footnote{TODO: grant nos?}
+We thank Sherwood Richers for providing his neutrino transport code
+\lstinline{Sedonu} for comparisons,
+Fran\c{c}ois H\'{e}bert for his timely help in finding and
+correcting bugs in the time-interpolation algorithm,
+and Luke Roberts for helpful discussions.
+
 % ******************************************************************************
@@
   C_{\rm AE}
-  &= \mathscr{E}_{\rm AE} + \mathscr{K}_{\rm AE}\, f, \\
+  &= \mathscr{E}_{\rm AE} - \mathscr{K}_{\rm AE}\, f, \\
   \label{eqn:sources_cse}
   C_{\rm SE}
-  &= \mathscr{E}_{\rm SE} + \mathscr{K}_{\rm SE}\, f.
+  &= \mathscr{E}_{\rm SE} - \mathscr{K}_{\rm SE}\, f.
 \end{align}
@@
 
-\subsubsection*{Acknowledgements}
-This research is supported in part by the Joint Institute for Nuclear
-Astrophysics---Center for the Evolution of the Elements (JINA-CEE)
-and the National Science Fondation under grant numbers X. \footnote{TODO: grant nos?}
-We thank Sherwood Richers for providing his neutrino transport code
-\lstinline{Sedonu} for comparisons,
-Fran\c{c}ois H\'{e}bert for his timely help in finding and
-correcting bugs in the time-interpolation algorithm,
-and Luke Roberts for helpful discussions.
-
 % ******************************************************************************

### 2001-01-01 paper.tex
Finish Matt's edits: correct equation typos and clarify C_PP

@@
     J \frac{\varepsilon^3}{T_\nu^4 \,\mathscr{F}_3(\eta_\nu)}
-    (1+\exp(\varepsilon/T_\nu-\eta_\nu)^{-1}, \\
+    \big(1+\exp(\varepsilon/T_\nu-\eta_\nu)\big)^{-1}, \\
     \label{eqn:H_from_gray}
@@
     H^\mu \frac{\varepsilon^3}{T_\nu^4\, \mathscr{F}_3(\eta_\nu)}
-    (1+\exp(\varepsilon/T_\nu-\eta_\nu)^{-1}.
+    \big(1+\exp(\varepsilon/T_\nu-\eta_\nu)\big)^{-1}.
   \end{align}
@@
 Tab.~\ref{tab:neutrino_processes}, and applying Kirchoff's law,
-Eqn.~\ref{eqn:kirchoffs_law}, to convert this to an opacity.
+Eqn.~\ref{eqn:kirchoffs_law}, to convert this to an effective absorption opacity,
+and its corresponding emissivity obeying Kirchoff's law.
 We only use this effective opacity for heavy-lepton neutrinos because

### 2001-01-01 paper.tex
Remove subsubsection titles in Introduction

--- 2001-01-01 paper.tex
@@
 % packages
+\usepackage{amssymb}              % for \gtrsim ...
 \usepackage{amsmath}              % for \text, align, split ...
-\usepackage{amssymb}              % for \gtrsim ...
-\usepackage{braket}               % for \bra \ket
+\usepackage{color}                % for colorful margin notes
 \usepackage[mathscr]{euscript}    % for \mathscr
+\usepackage{float}                % for better control of figure placement
 \usepackage{graphicx}             % for \includegraphics
@@
 \lstset{basicstyle=\ttfamily} % define \lstinline for inline code snippet
+
+\newcommand{\todo}[1]{\marginpar{\tiny{\textcolor{red}{#1}}}}
+%\renewcommand\todo[1]{} % if you wish to hide all todo notes
+
+% force all figures to be placed right where they are in the text
+\floatplacement{figure}{H}
 
@@
 
-\author{Y.\ L.\ Zhu}
-\affiliation{Department of Physics,
-  North Carolina State University, Raleigh, NC 27695, USA}
-
 \author{Andy Bohn}
@@
   We present a covariant ray tracing algorithm for computing high-resolution
-  neutrino distributions in general relativistic numerical spacetimes with
-  hydrodynamical sources.
+  neutrino distributions in general numerical spacetimes with hydrodynamical
+  sources.
   Our formulation treats the very important effect of
@@
   These background fields may be taken from a low-order moment simulation
-  or be ignored, in which case the method
+  or be ignored entirely, in which case the method
   reduces to a standard state-of-the-art ray tracing formulation.
   It handles radiation in regimes spanning optically thick to optically thin.
-  We test the new code, highlight its strengths and weaknesses,
-  and apply it to a simulation of a neutron star merger
-  to compute neutrino fluxes and spectra,
-  and to demonstrate a neutrino flavor oscillation calculation.
+  We test the new code, highlight its strengths and weaknesses, and
+  apply it to simulations of neutron star mergers to
+  compute neutrino fluxes as a function of energy and emission angle.
 \end{abstract}
@@
 Neutrinos are one of the dominant energy transport phenomena at play in
-neutron star mergers: heating, cooling, and pushing the disrupted nuclear
+neutron star mergers, heating, cooling, and pushing the disrupted nuclear
 matter.
@@
 respects the freedom of neutrino distribution functions to vary
-drastically from geometrically simple distributions in thermodynamic equilibrium.
+drastically from simple geometric distributions in thermodynamic equilibrium.
 
 This is a challenging task in the environment of a merger,
-which generally lacks any spatial symmetries, so that fully general
+which generally lacks any spatial symmetries so that fully general
 seven-dimensional solutions to the Boltzmann Equation are not feasible.
 Leakage approximations
-\citep{ruff1996-leakage_part1,ross2003-leakage_part2,
-  deat2013-leakage, pere2016-asl, radi2016-dynamical}
-\footnote{TODO: cite others}
-capture some of the qualitative effects of neutrinos on the matter, but
+\citep{deat2013-leakage, pere2016-asl, radi2016-dynamical}
+\todo{cite others}
+capture the qualitative effects of neutrinos on the matter, but
 provide extremely limited information about the neutrino field itself.
-\footnote{TODO: discuss mc more clearly}
-Monte Carlo methods like those
-used in supernova simulations \cite{abdi2012-monte_carlo}
-and stationary models of accretion disks \cite{rich2015-monte_carlo}
-are an excellent tool but require large computational resources
-because of their global treatment of the problem.
-
-The state of the art today for neutron star merger calculations
-couples radiation and matter using a truncated moment formalism
-\citep{fouc2015-m1_nsbh, fouc2016-m1_nsns},
-evolving the zeroth- and first-angular moments of the energy density,
-that is the total energy- and momentum-densities.
+\todo{discuss mc more clearly}
+Monte Carlo methods \citep{abdi2012-monte_carlo, rich2015-monte_carlo}
+are an excellent tool require large computational resources because of
+their global treatment of the problem.
+
+The state of the art today couples radiation and matter with a truncated
+moment formalism
+\citep{shib2011-truncated_moment, fouc2015-m1_nsbh, fouc2016-m1_nsns,
+  just2015-m1_code},
+\todo{cite others}
+evolving the zeroth- and first-angular moments of the distribution function
+(the energy density and flux respectively) representing the zeroth-energy moment
+of the radiation field (the total energy).
 This method, commonly called M1 transport, was formulated by
-\cite{thor1981-truncated_moment} and modernized by
-\cite{shib2011-truncated_moment}.
-M1 transport has recently become popular in the core-collapse supernova
-community as well
-\cite{kuro2012-m1_ccsne,just2015-m1_code,ocon2015-gr1d_with_nu,
-  skin2016-ray_vs_m1,robe2016-gr_m1_ccsne}.
-
-For merger simulations, \cite{fouc2016-m1_evolve_n} have recently
-expanded their M1 transport code
-to also evolve the zeroth-angular moment of the number density,
-providing total energy densities and average energies throughout the
-simulated volume.
+\cite{thor1981-truncated_moment, shib2011-truncated_moment}
+and has been applied for example in \cite{ramp2002-truncated_moment,
+  card2013-truncated_moment, fouc2015-m1_nsbh, ocon2015-gr1d_with_nu}.
+Recently, \cite{fouc2016-m1_evolve_n} have expanded their M1 transport code
+to also evolve the number densities of neutrinos,
+providing total and average energies throughout the simulated volume.
 Even so, with M1 transport codes only evolving two angular moments
 and one or two energy moments,
-they can extract only limited angular and spectral information
+we can extract only limited angular and spectral information
 from the neutrino fields.
@@
 \citep{ruff1999-nunubar_nsns, asan2000-nunubar, birk2007-nunubar,
-  hari2010-gr_nunubar_collapsar, zala2011-nunubar, leng2014-nunubar,
-  just2016-nsns_ejecta_obstacles}.
+  hari2010-gr_nunubar_collapsar, zala2011-nunubar, leng2014-nunubar}.
 
 %\subsubsection*{Why ray tracing?}
-Angular and spectral neutrino distributions in neutron star merger simulations may
-be available soon, for example with the coupled Monte-Carlo-M1 scheme presented in
-\citep{fouc2017-m1_with_mc}.
-In this work, however, we present a ray tracing method to compute neutrino
-distribution functions from the existing state-of-the-art general relativistic M1
-transport hydrodynamics simulations.
-We choose ray tracing because it is conceptually simple,
-numerically inexpensive,
-and extends to high resolution in energy and angle
-by simply increasing the number of rays sampled.
-Furthermore, the computational implementation parallellizes trivially.
+In this work we present a ray tracing method to compute neutrino distribution
+functions from state-of-the-art general relativistic M1 transport
+hydrodynamics simulations.
+We choose ray tracing because it is conceptually simple, numerically cheap,
+and easily extends to high resolution in energy and angle.
 
@@
 $f^{\nu_\sigma}(x_o^\alpha;p_\beta)$, or the amount of neutrino radiation
-of species
-$\nu_\sigma\in\{\nu_e,\nu_\mu,\nu_\tau,\bar{\nu}_e,\bar{\nu}_\mu,\bar{\nu}_\tau\}$
+of species $\nu_\sigma\in\{\nu_{\{e,\mu,\tau\}},\bar{\nu}_{\{e,\mu,\tau\}}\}$
 with momentum $p_\beta$ impinging on $x_o^\alpha$.
@@
 with the history of that light-cone truncated at large optical depth.
-\footnote{TODO: state scaling $\mathscr{O}(n)$?}
-It easily extends to high resolution in energy and angle by simply
+\todo{state scaling $\mathscr{O}(n)$?}
+And it easily extends to high resolution in energy and angle by simply
 increasing the number of rays sampled.
-And it parallelizes trivially by ray, since each ray is computed
-independently.
 
@@
 Most formulations assume an analytic spacetime metric
-\citep{birk2007-nunubar, caba2009-detecting_grb_nu,
-  hari2010-gr_nunubar_collapsar, kova2011-gr_ray_tracing}.
+\citep{birk2007-nunubar, hari2010-gr_nunubar_collapsar, kova2011-gr_ray_tracing}.
 And many make the simplifying assumption of blackbody emission from a
-neutrinosurface
-\citep{birk2007-nunubar, caba2009-detecting_grb_nu, kova2011-gr_ray_tracing},
-limiting them to equilibrium, optically-thick configurations.
+neutrinosurface \citep{birk2007-nunubar, kova2011-gr_ray_tracing},
+limiting them to equilibrium, optically thick configurations.
 Current state-of-the-art ray tracing formulations avoid the assumption of
@@
 %\subsubsection*{Why does a covariant formulation matter?}
-We formulate the ray tracing equations covariantly---free from
+We formulate our ray tracing equations covariantly---free from
 assumptions about the spacetime geometry or coordinates. This is essential
@@
 This is especially pronounced in the case of heavy-lepton neutrinos.
-\footnote{TODO: cite}
+\todo{cite}
 Inelastic scattering and pair processes can introduce further modifications.
@@
 for example neutrino oscillations and neutrino-antineutrino annihilation,
-\footnote{TODO: cite}
+\todo{cite}
 depend sensitively on the angular distribution.
@@
 occuring in the ejected and irradiated material.
-\footnote{TODO: justify these statements}
+\todo{justify these statements}
 The ray tracing method we present in this paper captures the dominant effects
@@
 or over a small volume outside a source.
+But the method becomes computationally burdensome when the inquiry extends
+to large volumes of spacetime.
 
@@
 a treatment which essentially decouples different neutrino momenta and species.
-\footnote{TODO: clarify how momenta are decoupled}
-When solving for $f^{\nu_\sigma}(x^\alpha;p_\beta)$,
-we have no information about the distribution at different momenta
-$f^{\nu_\sigma}(x^\alpha;p'_\beta)$,
-or about the distribution of the relevant antineutrino species
-$f^{\bar{\nu}_\sigma}(x^\alpha;p'_\beta)$;
-these missing data are essential ingredients
+\todo{clarify how momenta are decoupled}
+When we're solving for $f^{\nu_\sigma}(x^\alpha;p_\beta)$, we have no
+information about $f^{\nu_\sigma}(x^\alpha;p'_\beta)$, and
+$f^{\bar{\nu}_\sigma}(x^\alpha;p'_\beta)$, which are essential ingredients
 for the source terms of the Boltzmann Equation describing the creation and
@@
 
+\subsubsection*{Outline}
 In Sec.~\ref{sec:formulation} we derive the ray tracing equations from the
@@
 In Sec.~\ref{sec:tests} we present tests of the code.
-In Sec.~\ref{sec:applications} we present an application
-in the dynamical environment following the merger of two neutron stars,
-computing neutrino flavor oscillation along an outgoing ray,
-including the effects of coherent forward scattering with ambient neutrinos.
-In Sec.~\ref{sec:conclusions} we summarize our work and anticipate improvements.
+In Sec.~\ref{sec:applications} we present several applications
+in the dynamical environment following the merger of two neutron stars:
+computing neutrino spectra as a function of observer's angle,
+computing the self-interaction potential descriptive of flavor evolution
+in high-neutrino-density environments,
+and computing the energy dumped into jet-forming electron-positron
+pairs via neutrino-antineutrino annihilation.
 
@@
 whereas Latin indices ($i, j, ...$) range over the spatial coordinates 1--3.
-We use naturalized units in which $\{\hbar,c,k_{\rm B}\}=1$.
-And for most of the remainder of this article we suppress neutrino species label
+We use naturalized units in which $\hbar=c=1$.
+And for the remainder of this article we suppress the neutrino species label
 $\nu_\sigma$ since the formulation is general to any species.
-Where we do reference particular species we use the three categories
-relevant to the energy scales of mergers, $\nu_e$, $\bar{\nu}_e$, and
-$\nu_x=\left\{\nu_\mu,\bar{\nu}_\mu,\nu_\tau,\bar{\nu}_\tau\right\}$.
 
@@
 it has two degrees of freedom; we make this explicit by defining its
-spatial cartesian components with respect to spherical polar angles
+spatial components with respect to spherical polar angles
 \begin{equation}
   \label{eqn:def_direction}
-  \ell_\alpha \rightarrow
-  q (s,\sin a \cos b,\sin a\sin b,\cos a),
-\end{equation}
-with $q$ and $s$ functions of $a$ and $b$.
+  \ell_\alpha :=
+  q (s,\sin\upsilon_1\cos\upsilon_2,\sin\upsilon_1\sin\upsilon_2,\cos\upsilon_1),
+\end{equation}
+with $q$ and $s$ functions of $\upsilon_1$ and $\upsilon_2$.
 Now our symbol for the distribution function,
-$f(x^\alpha;\varepsilon,a,b)$,
+$f(x^\alpha;\varepsilon,\upsilon_1,\upsilon_2)$,
 makes manifest its seven independent arguments.
-
-We also define a rotated frame
-\begin{equation}
-  \ell_{\alpha'}=\frac{\partial x^\alpha}{\partial x^{\alpha'}}\,\ell_\alpha \nonumber
-\end{equation}
-in which the cartesian components of the direction 1-form are defined
-\begin{equation}
-  \label{eqn:def_direction_primed}
-  \ell_{\alpha'} \rightarrow
-  q (s,\sin A \cos B,\sin A\sin B,\cos A),
-\end{equation}
-so that momenta with vanishing polar angle $A$ move outward along coordinate
-radial rays.
-This transformation is chosen so that for an observer far from the source,
-incoming radiation will be concentrated into a narrow beam around $\cos A\approx1$,
-independent of that observer's position in coordinate space.
-For explicit definitions see App.~\ref{sec:definitions}.
-
-In Minkowski spacetime, and for a stationary observer
-$u_\beta\rightarrow(-1,0,0,0)$, we would have $s=0$, $q=1$,
-and $\varepsilon=-p_t$. In that case $\cos A$ may be identified
-\footnote{TODO: cite someone that uses $\mu$}
-with the familiar forward direction cosine $\mu$.
+\todo{also decompose wrt $p_t$}
 
@@
 which, suppressing the arguments of $f$ for simplicity, is written,
-\footnote{TODO: point out limit of QKEs}
+\todo{point out limit of QKEs}
 \begin{equation}
@@
 where $d/d\lambda$ denotes a derivative with respect to the affine
-parameter defining the neutrino momentum (Eqn.~\ref{eqn:geodesic_x} below)
-and $C[f]$ is the source term arising from interactions with the medium.
-By Eqn.~\ref{eqn:geodesic_x},
-the affine parameter has dimension ${\rm length}\,{\rm energy}^{-1}$,
-so the source term has dimension ${\rm energy}\,{\rm length}^{-1}$.
+parameter defining the neutrino momentum (Eqn.~\ref{eqn:geodesic_x})
+and $C[f]$ is the source term arising from interactions with the medium,
+having dimension ${\rm energy}\,{\rm length}^{-1}$.
+\todo{describe $C[f]$ in physical terms}
 The source term varies over phase space $(x^\alpha,p_\beta)$,
-and depends locally on the distribution function $f$
+and due to Fermi blocking depends locally on the distribution function $f$
 and nonlocally on the distribution function of this neutrino and its
@@
 the source term linear in $f$:
-\begin{align}
+\begin{equation}
   \label{eqn:boltzmann_linear}
-  \frac{d}{d\lambda}f &=
-  \mathscr{E} - \mathscr{K} f \\
-  \label{eqn:boltzmann_linear_s}
-  &= \mathscr{K}(\mathscr{S}-f)
-\end{align}
+  \frac{d}{d\lambda}f =
+  \mathscr{E} - \mathscr{K} f,
+\end{equation}
 where we have introduced
+\todo{describe $\mathscr{E}$ and $\mathscr{K}$ in physical terms}
 $\mathscr{E}$, the invariant total emissivity, and
 $\mathscr{K}$, the invariant total opacity.
-These describe respectively the energy gained and the energy lost
-per length traveled by the neutrino, true scalar quantities
-which take identical values for all observers.
-In the second form we have introduced the source function
-$\mathscr{S}\equiv\mathscr{E}/\mathscr{K}$, which makes manifest the
-behavior of the right hand side, driving $f$ toward $\mathscr{S}$
-over a lengthscale $\mathscr{K}^{-1}$ in the affine parameter;
-or from Eqn.~\ref{eqn:proper_length} below, over a proper lengthscale
-$\varepsilon/\mathscr{K}$ measured by the fiducial observer.
-
 These coefficients are computed by considering their dependence
@@
 (i.e.\ Fermi-blocking).
-We consider the two dominant classes of interactions in this work:
+
+We consider two classes of interactions in this work:
 the absorption/emission (AE) and elastic scattering (SE) processes
@@
 over the relevant emissivities and opacities
-for the reactions
-\begin{align}
-  \nu_e + n            &\leftrightarrow e^- + p,\nonumber\\
-  \bar{\nu}_e + p      &\leftrightarrow e^+ + n,\nonumber\\
-  \nu_e + {}^AZ        &\leftrightarrow e^- + {}^A(Z+1),\nonumber\\
-  \bar{\nu}_e + {}^AZ  &\leftrightarrow e^+ + {}^A(Z-1),\nonumber
-\end{align}
-with ${}^AZ$ representing a nucleus of mass number $A$ and charge $Z$.
-In terms of the emissivity $j(\varepsilon)$ describing number of neutrinos
-of energy $\varepsilon$ emitted per length,
-and the absorption opacity $\chi_a(\varepsilon)$
-describing the number absorbed per length,
-the coefficients are
+in Tab.~\ref{tab:neutrino_processes}:
+\todo{define $j$ and $\chi$}
 \begin{align}
@@
 \end{align}
-where $f^{\rm eq}$ is the distribution function of neutrinos in radiative
-equilibrium with the matter, i.e. the Fermi-Dirac distribution function
-\begin{equation}
-  \label{eqn:feq}
-  f^{\rm eq}(\varepsilon) \equiv
-  \left(1+e^{\varepsilon/(k_{\rm B} T) -\eta_\nu}\right)^{-1},
-\end{equation}
-with the neutrino chemical potentials dependent on the local density,
-temperature, and composition of the fluid via the neutron, proton,
-and electron chemical potentials:
-$\eta_{\nu_e}=-\eta_{\bar{\nu}_e}=\eta_p-\eta_n+\eta_{e^-}$ and
-$\eta_{\nu_x}=0$.
+where $f^{\rm eq}$ is the Fermi-Dirac distribution function with the chemical
+potential dependent on the local density, temperature, and composition of the
+fluid.
 The appearance of $f^{\rm eq}$ in Eqn.~\ref{eqn:ae_opacity_summed} is due to
@@
 different than the simple absorption opacity,
-a phenomenon called stimulated absorption \cite{burr2006-neutrino_opacities}.
-By detailed balance of the absorption/emission reactions, we may alternatively
-write the emissivity in terms of the equilibrium distribution function:
-\begin{equation}
-  \label{eqn:ae_emissivity_feq}
-  \mathscr{E}_{\rm AE}(\varepsilon)
-  = \mathscr{K}_{\rm AE}(\varepsilon) f^{\rm eq}(\varepsilon).
-\end{equation}
+a phenomenon called stimulated absorption.
 See App.~\ref{ssec:sources_ae} for details.
-\footnote{TODO: identify $\mathscr{K}_{\rm AE}$ and $\kappa^*$,
+\todo{identify $\mathscr{K}_{\rm AE}$ and $\kappa^*$,
   in e.g.\ \cite{hari2010-gr_nunubar_collapsar}}
@@
 a background field, and a sum over opacities for
-the reactions
+the relevant reactions in Tab.~\ref{tab:neutrino_processes}:
 \begin{align}
-  \nu + N     &\leftrightarrow \nu + N,\nonumber\\
-  \nu + {}^AZ &\leftrightarrow \nu + {}^AZ,\nonumber
-\end{align}
-with $N$ standing in for either $n$ or $p$.
-In terms of a background field $\Phi(\varepsilon)$ describing
-the number of neutrinos of energy $\varepsilon$ present at this event,
-and the scattering opacity $\chi(\varepsilon)$ describing the
-number scattered to other directions per length,
-the coefficients are
-\begin{align}
-  \label{eqn:se_emissivity}
+  \label{eqn:se_emissivity_summed}
   \mathscr{E}_{\rm SE}(\varepsilon)
-  &= \mathscr{K}_{\rm SE}(\varepsilon) \, \Phi(\varepsilon) \, \\
+  &= \mathscr{K}_{\rm SE}(\varepsilon) \, \Phi_{\rm SE}(\varepsilon) \, \\
   \label{eqn:se_opacity_summed}
@@
 \end{align}
-In the isotropic limit of trapped radiation,
-$\Phi$ is equivalent to $f^{\rm eq}$;
-in the free-streaming limit at a distance $r$ from a source,
-$\Phi$ attenuates as $r^{-2}$.
 See App.~\ref{ssec:sources_se} for details.
 
-We may use Eqns.~\ref{eqn:ae_emissivity_feq} and \ref{eqn:se_emissivity}
-to separate the source function thus:
-$\mathscr{S}=\mathscr{S}_{\rm AE}+\mathscr{S}_{\rm SE}$, with
-\begin{align}
-  \label{eqn:source_ae}
-  \mathscr{S}_{\rm AE}
-  &= \frac{\mathscr{K}_{\rm AE}}{\mathscr{K}} f^{\rm eq}, \\
-  \label{eqn:source_se}
-  \mathscr{S}_{\rm SE}
-  &= \frac{\mathscr{K}_{\rm SE}}{\mathscr{K}} \Phi.
-\end{align}
-With these definitions we write a separated form of
-Eqn.~\ref{eqn:boltzmann_linear_s}:
-\begin{equation}
-  \label{eqn:boltzmann_split}
-  \frac{d}{d\lambda}f =
-  \mathscr{K}_{\rm AE}(f^{\rm eq}-f)
-  + \mathscr{K}_{\rm SE}(\Phi-f)
-\end{equation}
-From Eqn.~\ref{eqn:boltzmann_split} we see that
-absorption/emission interactions drive the distribution function
-toward $f^{\rm eq}$ over an affine lengthscale $\mathscr{K}_{\rm AE}^{-1}$,
-and elastic scattering interactions drive it
-toward $\Phi$ over an affine lengthscale $\mathscr{K}_{\rm SE}^{-1}$.
-(Eqn.~\ref{eqn:proper_length} translates affine length to proper length
-for a given observer; in this case the proper length scale is
-$\varepsilon/\mathscr{K}$.)
-
 This paper introduces the use of neutrino densities and fluxes evolved in an
-M1 transport simulation to estimate the background field, $\Phi$.
+M1 transport simulation to estimate the background field, $\Phi_{\rm SE}$.
 App.~\ref{ssec:sources_se} details a method to calculate
-$\Phi(\varepsilon)$ in two different ways:
+$\Phi_{\rm SE}(\varepsilon)$ in two different ways:
 \begin{itemize}
@@
   extracted from a simulation evolved over multiple energy groups
-  to compute the background field with Eqn.~\ref{eqn:background_phi},
+  to compute the background field with Eqn.~\ref{eqn:se_factor_expansion},
 \item
@@
 If we multiply Eqn.~\ref{eqn:geodesic_x} by $u_\alpha \equiv dx_\alpha/ds$,
-we find the element of proper distance traversed by the
-neutrino as measured by the fiducial observer $u^\alpha$ is
-\begin{equation}
-  \label{eqn:proper_length}
-  ds=\varepsilon \,d\lambda.
-\end{equation}
+we find $ds=\varepsilon \,d\lambda$ is the proper distance traversed by the
+neutrino as measured by our fiducial obsever $u^\alpha$.
 
@@
   \includegraphics[width=\columnwidth]{fig-affine_parameter_label}
-  % Script: $DIR/figures-radiation_transport_schematics.nb
-  % w/ DIR=/Users/mbdeaton/Dropbox/Research/NuclearAstrophysics/NotesAndReference
   \caption{Affine parameterization of a neutrino trajectory of momentum
@@
     the neutrino emission event is at $x^\alpha_e$. The affine parameter
-    increases from the emission event: $\lambda_e<\lambda'<\lambda''$.
-    The green slab represents dense matter.}
+    increases from the emission event: $\lambda_e<\lambda'<\lambda''$.}
   \label{fig:affine_param}
@@
 
-\subsection{The Formal Solution}
+\subsection{The Rendering Equation}
 \label{ssec:rendering_eqn}
-We can integrate Eqn.~\ref{eqn:boltzmann_split} directly, 
-with the solution split into a boundary, absorption/emission, and
-a scattering term, $f=f_{\rm bdry}+f_{\rm AE}+f_{\rm SE}$:
+We can integrate Eqn.~\ref{eqn:boltzmann_linear} directly,
 \begin{align}
-  \label{eqn:rendering_fbdry}
-  f_{\rm bdry}(\lambda,\lambda_e)
-  &= f(\lambda_e) e^{-\tau(\lambda,\lambda_e)}, \\
+  \label{eqn:rendering}
+  f(\lambda,\lambda_e) &=
+  f_{\rm bound}(\lambda,\lambda_e) + f_{\rm integ}(\lambda,\lambda_e), \\
+  \label{eqn:rendering_fbound}
+  f_{\rm bound}(\lambda,\lambda_e) &=
+  f(\lambda_e) e^{-\tau(\lambda,\lambda_e)}, \\
+  \label{eqn:rendering_finteg}
+  f_{\rm integ}(\lambda,\lambda_e) &=
+  \int_{\lambda_e}^\lambda d\lambda'\,e^{-\tau(\lambda,\lambda')}
+  \mathscr{E}(\lambda'),
+\end{align}
+where the optical depth is defined,
+\todo{rewrite this section simpler; don't give integral solution
+  since we don't use}
+\begin{equation}
+  \label{eqn:optical_depth}
+  \tau(\lambda,\lambda') \equiv -\int_{\lambda'}^\lambda
+  d\lambda'' \, \mathscr{K}(\lambda''),
+\end{equation}
+and the parameterization conventions are depicted in
+Fig.~\ref{fig:affine_param}.
+
+We further separate the integrated distribution function into
+its absorption/emission and elastic scattering components,
+$f_{\rm integ} = f_{\rm AE} + f_{\rm SE}$:
+\begin{align}
   \label{eqn:rendering_fae}
@@
   &= \int_{\lambda_e}^\lambda d\lambda'\,e^{-\tau(\lambda,\lambda')}
-  \mathscr{K}_{\rm AE}(\lambda')f^{\rm eq}(\lambda'), \\
+  \mathscr{E_{\rm AE}}(\lambda'), \\
   \label{eqn:rendering_fse}
@@
   &= \int_{\lambda_e}^\lambda d\lambda'\,e^{-\tau(\lambda,\lambda')}
-  \mathscr{K}_{\rm SE}(\lambda')\Phi(\lambda'),
+  \mathscr{E_{\rm SE}}(\lambda').
 \end{align}
-where the optical depth is defined,
-\begin{equation}
-  \label{eqn:optical_depth}
-  \tau(\lambda,\lambda') \equiv \int_{\lambda'}^\lambda
-  d\lambda'' \, \mathscr{K}(\lambda''),
-\end{equation}
-and the parameterization conventions are depicted in
-Fig.~\ref{fig:affine_param}.
-Note that Eqn.~\ref{eqn:optical_depth} employs the total absorption
-plus scattering opacity, so that the optical depth attenuating the
-integrands of Eqns.~\ref{eqn:rendering_fae} and \ref{eqn:rendering_fse}
-is the total optical depth.
-
-%Note the following equivalents; we could use any of these to express
-%the integrands above:
-%\begin{align}
-%  \mathscr{K}_{\rm AE} \, f^{\rm eq}
-%  &= \mathscr{K}\, \mathscr{S}_{\rm AE}
-%  = \mathscr{E}_{\rm AE},\nonumber \\
-%  \mathscr{K}_{\rm SE} \, \Phi
-%  &= \mathscr{K} \, \mathscr{S}_{\rm SE}
-%  = \mathscr{E}_{\rm SE}.\nonumber
-%\end{align}
+Note that the optical depth used in both of these integrands is the total
+optical depth.
 
@@
 \label{ssec:moments}
-\footnote{TODO: code these up}
+\todo{code these up}
 We may take angular moments of the distribution function:
@@
 \end{align}
-defining the specific energy density,
-specific momentum density, and
-specific radiation pressure tensor, respectively.
+representing specific energy density, specific energy flux, and
+specific momentum flux, respectively.
 Here ``specific'' refers to the quantity being integrable over neutrino energy.
-\footnote{TODO: relate moments to other frames?}
+\todo{relate moments to other frames?}
 Integrals~\ref{eqn:J}--\ref{eqn:Sab} are performed over a solid angle in
 momentum space while holding $\varepsilon$ constant:
-$d\Omega \equiv d(\cos a)\,db$.
-\footnote{TODO: consistent with Eqn.~\ref{eqn:def_direction}?}
-We also make use of the number density and flux defined
-\begin{align}
-  \label{eqn:G}
-  G(\varepsilon) &=
-  \frac{\varepsilon^2}{(2\pi)^3} \oint d\Omega' f(\varepsilon, \ell'_\beta), \\
-  \label{eqn:Ka}
-  K^\mu(\varepsilon) &=
-  \frac{\varepsilon^2}{(2\pi)^3} \oint d\Omega' f(\varepsilon, \ell'_\beta) \ell'^\mu.
-\end{align}
-
-The energy-integrated moments take the form
+$d\Omega \equiv d(\cos\upsilon_1)\,d\upsilon_2$.
+\todo{consistent with Eqn.~\ref{eqn:def_direction}?}
+We will also make use of the number density and flux defined as
+$G(\varepsilon)=J(\varepsilon)/\varepsilon$
+and $K^\mu(\varepsilon)=H^\mu(\varepsilon)/\varepsilon$
+respectively.
+
+The energy-integrated moments are
 \begin{equation}
   \label{eqn:J_H_S_eps_integrated}
-  X = \int_0^\infty d\varepsilon \, X(\varepsilon),
-\end{equation}
-with $X$ standing in for any of $\{J,H^\mu,S^{\mu\gamma}\}$ or
-$\{G,K^\mu\}$,
-the first having dimension ${\rm energy}\,{\rm length}^{-3}$,
-and the second ${\rm length}^{-3}$.
-
-We compute moments for a particular observer
-by specifying the four-velocity in Eqn.~\ref{eqn:def_momentum}.
-Two choices are particularly useful:
-an observer stationary in the coordinate frame (i.e. Eulerian),
-or one stationary in the fluid frame (i.e. comoving)
-\cite{smar1980-gr_hydro}.
-Explicit definitions are given in App.~\ref{sec:definitions}.
-We distinguish moments computed for an Eulerian observer with a tilde,
-e.g. $\tilde{J}$, $\tilde{H}^\mu$, $\tilde{S}^{\mu\nu}$;
-note that these three Eulerian moments are identical to the lab-frame moments
+  \big\{ J,H^\mu,S^{\mu\nu} \big\} = \int_0^\infty d\varepsilon \,
+  \big\{ J(\varepsilon),H^\mu(\varepsilon),S^{\mu\nu}(\varepsilon) \big\},
+\end{equation}
+having dimension ${\rm energy}\,{\rm length}^{-3}$.
+
+In practice we use two possible fiducial observers to define the momentum
+decomposition via Eqn.~\ref{eqn:def_momentum}: the Eulerian,
+or normal observer $n_\mu=-\alpha \,\partial_\mu t$,
+and the fluid, or comoving observer
+$u^\mu=Wn^\mu+v_{\rm E}^\mu$.
+Here $t$ is coordinate time and $\alpha$ is lapse
+in the standard 3+1 decomposition of the metric
+\begin{equation}
+  \label{eqn:adm_metric}
+  \psi_{\mu\gamma} :=
+  \left(
+  \begin{matrix}
+    -\alpha^2 + \beta^i \beta_i  & \beta_i \\
+    \beta_j                      & g_{ij}
+  \end{matrix}
+  \right).
+\end{equation}
+We have also introduced
+the fluid Lorentz factor $W=-n_\mu u^\mu=\alpha u^t$,
+its Eulerian velocity $v_{\rm E}^\mu=g^\mu_\lambda u^\lambda$
+(distinct from its coordinate velocity $v^\mu=u^\mu/u^t$),
+and the projection tensor orthogonal to the normal observer
+$g^\mu_\lambda=\psi^\mu_\lambda+u^\mu u_\lambda$,
+whose spatial part is the 3-metric, $g_{ij}$.
+
+We distinguish moments computed assuming an Eulerian observer with a tilde,
+e.g. $\tilde{J}$, $\tilde{H}^\mu$, $\tilde{S}^{\mu\nu}$.
+Note that these three Eulerian moments are identical to the lab-frame moments
 $E$, $F^\mu$, and $P^{\mu\nu}$
@@
 given by \citep{hugh1994-eh_finding},
-and Eqns.~\ref{eqn:rendering_fae}, \ref{eqn:rendering_fse},
-and \ref{eqn:optical_depth} in the form given below.
+and Eqns.~\ref{eqn:rendering} and~\ref{eqn:optical_depth}
+in the form given below.
 By using the time-component of Eqn.~\ref{eqn:geodesic_x} ($dt=d\lambda \, p^t$)
-we may transform the integrations to coordinate time.
+we may transform our integrations to coordinate time.
 The coupled system of ordinary differential equations is
@@
   \frac{dx^i}{dt} &=
-  g^{ij} \frac{p_j}{p^t} - \beta^{i},\\
+  g^{ij} \frac{p_j}{p^t} - \beta^{i}\\
   \label{eqn:num_p}
@@
   + \beta^k_{,i} p_k
-  - \frac{1}{2} g^{jk}_{,i}\frac{p_j p_k}{p^t},\\
+  - \frac{1}{2} g^{jk}_{,i}\frac{p_j p_k}{p^t}\\
+  \label{eqn:num_fae}
+  \frac{df_{\rm AE}}{dt} &=
+  e^{-\tau} \frac{\mathscr{E_{\rm AE}}}{p^t}\\
+  \label{eqn:num_fse}
+  \frac{df_{\rm SE}}{dt} &=
+  e^{-\tau} \frac{\mathscr{E_{\rm SE}}}{p^t}\\
   \label{eqn:num_tau}
   \frac{d\tau}{dt} &=
-  - \frac{1}{p^t}\mathscr{K},\\
-  \label{eqn:num_fae}
-  \frac{df_{\rm AE}}{dt} &=
-  \frac{1}{p^t}e^{-\tau}\mathscr{K_{\rm AE}} \,f^{\rm eq},\\
-  \label{eqn:num_fse}
-  \frac{df_{\rm SE}}{dt} &=
-  \frac{1}{p^t}e^{-\tau}\mathscr{K_{\rm SE}} \,\Phi.
+  - \frac{\mathscr{K}}{p^t}.
 \end{align}
@@
 We integrate each ray until we reach a terminal optical depth of
-$\tau_{\rm term}$ at the earliest effective emission event $x_e^\alpha$.
-The concept of earliest emission event is a fictitious construct we use
-to allow us to truncate the integration at an event along the ray where any
+$\tau_{\rm term}$ at the earliest emission event $x_e^\alpha$.
+The concept of earliest emission event is a facetious construct we use
+to allow us to truncate our integration at a location along the ray where any
 further additions to the field are negligible due to the large optical depth
 between $x_e^\alpha$ and $x_o^\alpha$.
-We choose $\tau_{\rm term}=14$ so that $e^{-\tau_{\rm term}}<10^{-6}$,
-and we then discard the contribution of $f_{\rm bdry}$
-(Eqn.~\ref{eqn:rendering_fbdry}).
+We choose $\tau_{\rm term}$ large so that
+$f_{\rm bound}$ is negligible
+(in practice $e^{-\tau_{\rm term}}=0.001$)
+and compute $f_{\rm bound}$ assuming radiative equilibrium with the matter,
+i.e. $f_{\rm bound}=(1+\exp((\varepsilon-\mu_\nu)/T)$,
+and $\mu_{\nu_e}=-\mu_{\bar{\nu}_e}=\mu_{e^{-}}+\mu_p-\mu_n$ and
+$\mu_{\nu_x}=0$.
+If the ray terminates off the numerical domain before the terminal optical
+depth accumulates, we set $f_{\rm bound}=0$.
 
 Since we don't know the emission event a priori,
+\todo{make clear ``emission event'' doesn't involve physical assumption}
 we follow the integration backwards in time,
 from $t_o$ to $t_e$.
-We begin each integration by setting initial values for the variables at $t_o$:
-the observer specifies $x^i_o$ and $p_{i,o}$,
+We begin each integration by setting initial values for the variables at $t_o$.
+The observer specifies $x^i_o$ and $p_{i,o}$,
 and we set $f_{{\rm AE},o}=f_{{\rm SE},o}=0$ and $\tau_o=0$.
@@
 We integrate Eqns.~\ref{eqn:num_x}--\ref{eqn:num_tau} with adaptive step sizes,
-using the 3rd order Runge-Kutta algorithm
-which produces an error estimate by comparing the 3rd and 2nd order solutions.
-After each step is taken, the errors for each of the nine variables of
+using the 5th-order Dormand-Prince algorithm \cite{pres2007-nr_3rd_ed},
+which produces an error estimate by comparing the 4th and 5th order solutions.
+\todo{confirm}
+After each step is taken, the errors for each of the 9 variables of
 Eqns.~\ref{eqn:num_x}--\ref{eqn:num_tau} are compared.
 If the maximum of these errors exceeds some absolute or relative threshold
-\footnote{TODO: specify ScaledAbsRel algorithm}
+\todo{specify ScaledAbsRel algorithm}
 the step size is decreased and the step recomputed,
@@
 % see OdeIntegrators::AdaptiveDense
-% and OdeSteppers::RungeKutta3
 % and OdeSteppers::DormandPrince5
-% and OdeController
 % and OdeControllers::ProportionalIntegral_Base::StepSizeToAttempt
 % and OdeErrorMeasures::ScaledAbsRel
-\footnote{TODO: specify}
-
-We integrate these equations through the simulated spacetime
-over which the following volume data are known:
-the spacetime metric $\psi_{\alpha\beta}$,
+\todo{specify}
+
+The volume data are components of the metric $\psi_{\alpha\beta}$,
 its derivatives $\psi_{\alpha\beta,\gamma}$,
-the fluid velocity $u_i$,
-Lorentz factor $W$,
-density $\varrho$,
-temperature $T$,
-and electron fraction $Y_e$,
-all defined in App.~\ref{sec:definitions}.
+and the hydrodynamical data $W$, $u_i$, $\varrho$, $T$, and $Y_e$.
 These fields are stored in spectral representation.
-If they are computed via hydrodynamical simulation, they may be saved to disk
+If they are computed via hydrodynamical simulation, they are saved to disk
 at specified coordinate times and interpolated
@@
 If computed from a quasi-stationary solution to the
+\todo{clarify quasistationary timescale}
 general relativistic hydrodynamics equations,
 a single time is stored to disk, and no time interpolation is needed.
-In this paper for simplicity and to limit computational memory loads,
-we limit ourselves to quasistationary configurations
-and use single time slices in every test case.
 
@@
 To test our algorithm, we integrate the ray tracing equations
-(Eqns.~\ref{eqn:num_x}--\ref{eqn:num_tau}) for various observers
+(Eqns.~\ref{eqn:num_x}--\ref{eqn:num_tau}) at one observation position
 in the following configurations. This suite of configurations
@@
 
-In the following we present two forms of the integrated distribution functions,
-Eqns.~\ref{eqn:num_fae} and \ref{eqn:num_fse}:
-1) the full form  by which the solution is $f_{\rm scat}=f_{\rm AE,scat}+f_{\rm SE,scat}$
-(labeled \emph{scat} since we include elastic scattering), and
-2) the form treating only absorption/emission interactions
-in which we set $\mathscr{K}_{\rm SE}=0$ so that $f_{\rm noscat}=f_{\rm AE,noscat}$
-(labeled \emph{noscat} since we ignore elastic scattering).
-There is no simple relation between these forms
-(i.e.\ $f_{\rm AE,scat} \neq f_{\rm AE,noscat}$)
-because $f_{\rm AE}$ is integrated using the absorption plus scattering optical depth
-in the first form and solely the absorption optical depth in the second form.
-The differences between the two methods is apparent
-in Fig.~\ref{fig:cumulative_f_homogeneous}.
-Before this work, the \emph{noscat} form of the equations has been the standard for
-\footnote{TODO: explain why; cite others}
-ray tracing, though many authors include the scattering optical depth in
-the integration of Eqn.~\ref{eqn:num_fae}
-(see for example \cite{hari2010-gr_nunubar_collapsar}).
-
 \subsection{Infinite Homogeneous Slab:
@@
 with the matter.
-We set up a large slab of matter in a Minkowski spacetime
-representative of the fluid thermodynamic state at a radius of 50~km
-in the test presented in Sec.~\ref{ssec:test_collapse}---a
-massive collapsed star following core bounce---where
-$\varrho=10^{11}\,{\rm g}\,{\rm cm}^{-3}$, $T=3.7$~MeV, and $Y_e=0.12$.
-For comparision with that test we use the LS180 equation of state
-\cite{latt1991-nuc_eos}
-\footnote{available at \lstinline{stellarcollapse.org}}
-in which the equilibrium $\nu_e$ neutrino chemical potential is
-$\eta_{\nu_e}=-0.1555$
-(with $\eta_{\bar{\nu}_e}=-\eta_{\nu_e}$ and $\eta_{\nu_x}=0$).
-The opacity table is computed using \lstinline{NuLib} from cross-sections
-neglecting weak magnetism
-\footnote{TODO: anything else?}
-and is stored on a grid covering energy, density, temperature, and composition
-ranges spanning $\varepsilon_i\in[1,280.5]$~MeV logarithmically,
-$\varrho\in[10^6,6.3\times10^{15}]\,{\rm g}\,{\rm cm}^{-3}$ logarithmically,
-$T\in[0.05,200]$~MeV logarithmically, and 
-$Y_e\in[0.035,0.55]$ linearly,
-with grid extents$\{18,82,65,51\}$ respectively.
-Interpolation is performed linearly.
-
-We use the equilibrium distribution functions to define background
-fields for the \emph{scat} case:
-in the \emph{spectral} method we use
-$J(\varepsilon)=\varepsilon^3 f^{\rm eq}(\eta_\nu,T;\varepsilon)/(2\pi^2)$,
-and in the \emph{gray} method we use
-$J=C T^4 \mathscr{F}_3(\eta_\nu)/(2\pi^2)$ and
-$G=C T^3 \mathscr{F}_2(\eta_\nu)/(2\pi^2)$,
-using the Fermi integrals defined in Eqn.~\ref{eqn:fermi_integral},
-and with $C$ the conversion constant from ${\rm energy}^3$ to ${\rm length}^{-3}$.
-
-Fig.~\ref{fig:mfps-50km} presents the neutrino mean free paths
-at this thermodynamic point over two energy decades.
-For this test we choose a slab large enough for neutrinos of all
-energies to be trapped.
-We sample the distribution function with ray tracing over a uniform grid
-of 40 points in energy $\varepsilon \in [0,100]$~MeV.
-The domain extends to $\tau_{\rm term}\,L_{\rm mfp,max}\approx10^7\,{\rm km}$,
-since our ray tracing algorithm integrates rays to terminal optical
-depths of $\tau_{\rm term}\geq14$.
+We set up a large slab of matter in two homogeneous configurations
+representative of the fluid in the Massive Collapsed Star test
+(Sec.~\ref{ssec:test_collapse}) at two different radii:
+at 50~km
+For comparision with Collapsed Star test, we use the LS180 equation of state,
+in which the equilibrium
+neutrino chemical potentials are
+$\mu_{\nu_e}=-0.639$~MeV, and 
+$\mu_{\nu_e}=5.771$~MeV respectively
+($\mu_{\bar{\nu}_e}=-\mu_{\nu_e}$ and $\mu_{\nu_x}=0$).
+Plot in Fig.~\ref{fig:homogeneous_isotropic}.
+\todo{run test at this thermodynamic point}
 
 \begin{figure}
-  %\includegraphics[width=\columnwidth]{figure-mfps-50km}
-  \resizebox{\columnwidth}{!}{\input{figure-mfps-50km.tex}}
-  % Data:   $DIR/table-50km-nu?.dat
-  % Script: $DIR/GnuplotScripts/plot-mfp-50km.gnu
-  % w/ DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20161007-ray_tracing_tests/HomogeneousIsotropic
-  \caption{Mean free paths
-    representative of the post-bounce collapse profile (presented in
-    Sec.~\ref{ssec:test_collapse}) at
-    50~km, where $\varrho=10^{11}\,{\rm g}\,{\rm cm}^{-3}$,
-    $T=3.7$~MeV, and $Y_e=0.12$; this thermodynamic state is used in
-    the tests with homogeneous matter distributions
-    (presented in Secs.~\ref{ssec:test_equilibrium}--\ref{ssec:test_ab_star}).
-    In the fluid rest frame the mean free paths are given by the inverses of the opacities
-    $\chi^*_a=\mathscr{K}_{\rm AE}/\varepsilon$ for absorption and
-    $\chi_s=\mathscr{K}_{\rm SE}/\varepsilon$ for elastic scattering interactions
-    (see App.~\ref{sec:source_terms} for definitions).
-    Note that elastic scattering opacities are identical across species
-    below energies at which weak magnetism plays a role \cite{horo2002-weak_mag}.
-    In this data they are exactly identical
-    because we have turned off weak magnetism in our opacity calculations
-    in order to compare our results with the historical literature.
-    Computed using \lstinline{NuLib}.
-  }
-  \label{fig:mfps-50km}
-\end{figure}
-
-Fig.~\ref{fig:cumulative_f_homogeneous} displays the cumulative distribution
-function integrated along the ray, for each of the three species and using
-both methods, \emph{noscat} and \emph{scat}, at the single energy
-$\varepsilon=11.25$~MeV. We display
-Eqns.~\ref{eqn:num_fae} and \ref{eqn:num_fse} in their integral form:
-\begin{equation}
-  \label{eqn:cumulative_f}
-  f(t)
-  = - \int_{0}^t \frac{dt'}{p^t}e^{-\tau(0,t')}
-  \mathscr{K}(t')f^{\rm eq}(t'),
-\end{equation}
-and the integration proceeds backwards in time, from $t=0$
-\footnote{TODO: could write more simply because Minkowski}
-to some terminal $t<0$.
-The figure shows this backwards-in-time integration proceeding left to right.
-
-\begin{figure}
-  %\includegraphics[width=\columnwidth]{fig-cumulative_f-homogeneous}
-  \resizebox{\columnwidth}{!}{\input{fig-cumulative_f-homogeneous.tex}}
-  % Data:   $DIR/Lev1-v0-ShootANeutrino-01/Lev1
-  % Script: $DIR/GnuplotScripts/plot_f-integral_over_t.gnu
-  % w/ DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20161007-ray_tracing_tests/HomogeneousIsotropic
-  \caption{Cumulative distribution functions at $\varepsilon=11.25$~MeV
-    in the homogeneous infinite slab test
-    (presented in Sec.~\ref{ssec:test_equilibrium}).
-    The integration proceeds from left to right, or backwards in time $t$.
-    Each ray terminates when it achieves a total optical depth
-    greater than $\tau_{\rm term}=14$.
-    The points plotted correspond to the time steps chosen by the adaptive
-    time-stepping algorithm described in Sec.~\ref{ssec:numerical}.
-  }
-  \label{fig:cumulative_f_homogeneous}
-\end{figure}
-
-In Fig.~\ref{fig:cumulative_f_homogeneous} we see some expected features.
-The final distribution functions asymptote to their equilibrium Fermi-Dirac
-levels at this energy
-$f^{\rm eq}_{\{\nu_e,\bar{\nu}_e,\nu_x\}}=\{0.039, 0.053, 0.046\}$
-within a few mean free paths;
-in the \emph{scat} cases it is the sums $f_{\rm AE}+f_{\rm SE}$ that
-achieve these values.
-And the lengths of the rays are proportional to the mean free paths:
-$L_{\nu_x}>L_{\bar{\nu}_e}>L_{\nu_e}$;
-in the \emph{scat} cases these lengths are less than in the \emph{noscat} cases,
-since the total mean free paths are less than the absorption mean free paths
-(significantly so for $\nu_x$, negligibly so for $\nu_e$).
-
-As Fig.~\ref{fig:mfps-50km} reveals, integrating the distribution functions
-over the energies $\varepsilon\in[0,100]$~MeV
-probes the numerical solution over length scales from
-0.1~km to $10^{5}\,{\rm km}$.
-Fig.~\ref{fig:homogeneous_isotropic} shows the error in our results for
-the \emph{noscat} case.
-In this simple case, the dominant source of error is the neglected
-boundary term with a relative scale of $e^{-\tau_{\rm term}}\approx10^{-6}$.
-In the following inhomogeneous configurations,
-errors from the integration dominate over the boundary term.
-In this test and only this test, we used a higher-order integrator, a 5th
-order Dormand-Prince algorithm \cite{pres2007-nr_3rd_ed}. This is because
-\footnote{TODO: discuss with Matt}
-our default 3rd order Runge-Kutta algorithm estimates a vanishing error
-in this configuration.
-
-\begin{figure}
-  %\includegraphics[width=\columnwidth]{figure-ferr-50km}
-  \resizebox{\columnwidth}{!}{\input{figure-ferr-50km.tex}}
-  % Data:   $DIR/Lev1-v0-noscat-01/Lev1
-  % Script: $DIR/GnuplotScripts/plot_ferr.gnu
-  % w/ DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20161007-ray_tracing_tests/HomogeneousIsotropic
-  \caption{Relative error in integrated equilibrium distribution functions
-    in the infinite homogeneous slab
-    (presented in Sec.~\ref{ssec:test_equilibrium}).
-    Plotted here are the relative differences between the final
-    $f_{\rm AE}$ in the \emph{noscat} case
-    (see Fig.~\ref{fig:cumulative_f_homogeneous})
-    and the equilibrium distribution functions given by
-    Eqn.~\ref{eqn:feq}.
-    The source of this error is the discarded boundary term
-    $f_{\rm bdry}\sim10^{-6}$, described in Sec.~\ref{ssec:numerical}.
-  }
+  \includegraphics[width=\columnwidth]{fig-homogeneous_equilibrium}
+  \caption{Equilibrium distribution functions in a homogeneous background
+    at two thermodynamic points,
+    representative of the collapse profile at
+    50~km ($\varrho=1.235 \times 10^{11}\,{\rm g}\,{\rm cm}^{-3}$,
+    $T=4$~MeV,
+    $Y_e=0.12$) and
+    200~km ($\varrho=3.026\times 10^{7}\,{\rm g}\,{\rm cm}^{-3}$,
+    $T=0.5$~MeV,
+    $Y_e=0.5$).
+    Described in Sec.~\ref{ssec:test_equilibrium}}
   \label{fig:homogeneous_isotropic}
@@
 \label{ssec:test_doppler}
-We reproduce the test above, again in Minkowski spacetime,
-but with the matter and observer in relative motion.
-We use a stationary observer
-and fluid moving in the positive $z$-direction:
-with $u^\alpha \rightarrow W(1,0,0,v)$ and $W=(1-v^2)^{-1/2}$,
-where $W$ is the Lorentz factor defined in App.~\ref{sec:definitions}.
-All other thermodynamic variables and background fields are unchanged
-since our ray integration uses these quantities in the fluid frame.
-
-A stationary observer measures an energy of $\tilde{\varepsilon}$
-for a neutrino with momentum
-$p_\alpha \rightarrow \tilde{\varepsilon}(-1,\Omega_i)$
-and direction $\Omega_i \rightarrow (\sin A, 0, \cos A)$.
-In the fluid frame this neutrino has energy $\varepsilon=-u^\alpha p_\alpha$;
-therefore the average energy varies with observing angle like
-\begin{equation}
-  \label{eqn:doppler_avg_eps}
-  \langle \tilde{\varepsilon} \rangle(\cos A) =
-  \langle \varepsilon \rangle^{\rm eq} \frac{1}{W(1-v \cos A)},
-\end{equation}
-where the symbol $\langle \tilde{\varepsilon} \rangle(\cos A)$ emphasizes
-the functional dependence on $\cos A$.
-The equilibrium average energy $\langle \varepsilon \rangle^{\rm eq}$
-is given by $T \mathscr{F}_3(\eta_\nu)/\mathscr{F}_2(\eta_\nu)\approx 3T$,
-with the Fermi integrals given in Eqn.~\ref{eqn:fermi_integral}.
-Eqn.~\ref{eqn:doppler_avg_eps} describes the well-known Doppler effect.
-
-We sample the distribution function $f(\tilde{\varepsilon},\cos A,B)$
-with ray tracing
-over a uniform grid of 40 points in energy $\tilde{\varepsilon}\in[0,100]$
-and 30 points in angle $\cos A\in[-1,1]$, holding fixed $B=\pi$.
-Results are shown in Fig.~\ref{fig:avg_eps_doppler} for
-the velocities $v=\{0,0.1,0.8\}$.
-The ray tracing results are computed from total densities in each angular bin,
-that is
-\begin{equation}
-  \label{eqn:avg_eps_per_cosA}
-  \langle \tilde{\varepsilon} \rangle (\cos A) =
-  \frac{\tilde{J}(\cos A)}{\tilde{G}(\cos A)},
-\end{equation}
-with the Eulerian energy density per angular bin given by
-\begin{equation}
-  \label{eqn:J_per_cosA}
-  \tilde{J}(\cos A) =
-  \frac{1}{(2\pi)^3} \int_0^{\infty} d\tilde{\varepsilon} \, \tilde{\varepsilon}^3
-  \int_0^{2\pi} dB \, f(\tilde{\varepsilon},\cos A,B),
-\end{equation}
-and $\tilde{G}(\cos A)$ defined analogously.
-\footnote{TODO: add metric terms for these integrals}
-Results are identical for all \emph{scat} and \emph{noscat} methods.
-In Fig.~\ref{fig:avg_eps_doppler}
-we see the common features of a red-shifted spectrum for receding fluid
-($\cos A\ll0$),
-a blue-shifted spectrum for approaching fluid
-($\cos A\gg0$),
-and a slightly red-shifted spectrum for fluid moving transverse to the
-observer ($\cos A=0$).
+We reproduce the test above, but with the matter moving in
+in the $+z$ direction, with the observer stationary.
+Plot in Fig.~\ref{fig:avg_eps_doppler}.
+\todo{describe test}
 
 \begin{figure}
-  %\includegraphics[width=\columnwidth]{fig-moving_fluid_avg_eps}
-  \resizebox{\columnwidth}{!}{\input{fig-moving_fluid_avg_eps.tex}}
-  % Data:   $DIR/Lev1-v*-noscat-01
-  % Script: $DIR/plot-avg_eps_vs_cosA.gnu
-  % w/ DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20161007-ray_tracing_tests/MovingFluid
-  \caption{Average energy of $\bar{\nu}_e$ neutrino fields
-    measured by a stationary observer,
-    $\langle\tilde{\varepsilon}\rangle(\cos A)$,
-    in the moving slab test (presented in Sec.~\ref{ssec:test_doppler}).
-    The observer sees neutrinos with $\cos A>0$ to be moving primarily with
-    the fluid.
-    The points are computed from ray tracing spectra; the lines are the
-    analytic formula Eqn.~\ref{eqn:doppler_avg_eps}.
-  }
+  \includegraphics[width=\columnwidth]{fig-moving_fluid-avg_eps_vs_cosA}
+  \caption{Average energy $\langle\varepsilon\rangle=J/G$
+    of neutrino fields for fluid moving in the $+z$ direction,
+    with $A$ the angle between the neutrino momentum and the $+z$ direction.
+    Test described in Sec.~\ref{ssec:test_doppler}}
   \label{fig:avg_eps_doppler}
@@
 \label{ssec:test_ab_star}
-The homogeneous configurations of the previous sections may be extended to
-probe the solution outside the optically thick regime
-by setting up an idealized homogeneous star, with the thermodynamic variables
-constant inside radius $R$ and vanishing outside.
-We choose $R=50$~km and place the observer at $r=75$~km,
-at which position there is a radiation cone of half-opening angle
-$\cos A_{\rm max}\approx0.75$.
-
-The formal solutions of
-Eqns.~\ref{eqn:rendering_fae} and \ref{eqn:rendering_fse}
-may be directly integrated in this scenario. Assuming Minkowski spacetime
-and stationary fluid we have
-\begin{align}
-  \label{eqn:fae_homogeneous_general}
-  f_{\rm AE} &=
-  \frac{\chi_a^{*}}{\chi} f^{\rm eq} \left(1-e^{-\chi s}\right), \\
-  \label{eqn:fse_homogeneous_general}
-  f_{\rm SE} &=
-  \frac{\chi_s}{\chi} \Phi \left(1-e^{-\chi s}\right),
-\end{align}
-where $s$ is the path length traversed by the ray through the star,
-\begin{equation}
-  \label{eqn:geometric_s}
-  s = 2 R \left(1-\frac{r^2}{R^2}(1-\cos^2 A)\right)^{1/2}.
-\end{equation}
-The total opacity is $\chi=\chi_a^*+\chi_s$, and the stimulated absorption
-opacity $\chi_a^*$ and elastic scattering opacity $\chi_s$
-are defined in App.~\ref{sec:source_terms}.
-
-Since no analytic form is known for the background field $\Phi$
-interior to the star, we examine only the \emph{noscat} case,
-with $\chi_s=0$ and $\chi=\chi_a^{*}$.
-This scenario has been widely used in the literature as a test for
-radiation codes
-\cite{smit1997-two_moment,abdi2012-monte_carlo,fouc2015-m1_nsbh}.
-We sample the distribution function over a uniform grid of
-30 points in angle $\cos A \in [0.734,1]$ and
-40 points in energy $\tilde{\varepsilon} \in [0,100]$~MeV.
-Because of the step function in fluid variables at radius $R$,
-we throttle the maximum step size at 1.25~km.
-
-In Fig.~\ref{fig:f_absorption_sphere} we display the samples at
-$\varepsilon=11.25$~MeV, along with the analytic functions
-specific to each species' equilibrium distribution function and opacity.
-As expected only $\nu_e$ saturates at $f^{\rm eq}$, remaining almost
-constant across $\cos A$ until we get to rays that pass through a length
-of the star comparable to or less than the mean free path at this energy,
-16~km. We also see that $\bar{\nu}_e$ comes close to saturating with
-a mean free path just over 100~km; and $\nu_x$ is well into the optically
-thin regime.
+[Test described in~\cite[Sec.~3.2]{smit1997-two_moment}.]
+Plot in Fig.~\ref{fig:f_absorption_sphere}.
 
 \begin{figure}
-  %\includegraphics[width=\columnwidth]{fig-f-absorption_star-11MeV}
-  \resizebox{\columnwidth}{!}{\input{fig-f-absorption_star-11MeV.tex}}
-  % Data:   $DIR/Lev1-01/Lev1
-  % Script: $DIR/Lev1-01/Lev1/plot_f_monochromes-11.25MeV.gnu
-  % w/ DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20161007-ray_tracing_tests/HomogeneousAbsorptionSphere
-  \caption{Distribution functions outside an idealized homogeneous star
-    with radius $R=50$~km, and observer at $r=75$~km
-    (presented in Sec.~\ref{ssec:test_ab_star}).
+  \includegraphics[width=\columnwidth]{fig-absorption_sphere}
+  \caption{Distribution function outside a constant-density star
+    with radius $R=15$~km, and observer at $r=22.5$~km.
     $A$ is the angle between the neutrino momentum and the $\hat{r}$ direction.
-    In this plot we display only the samples at energy
-    $\varepsilon=11.25\,{\rm MeV}$.
-    The points are computed from ray tracing; the lines from the analytic
-    solution, Eqn.~\ref{eqn:fae_homogeneous_general}.
-  }
+    The sample grid uses 30 points in energy and angle spanning the ranges
+    $\varepsilon\in(0,100)$~MeV and $\cos A\in(0.6454,1)$,
+    with sample points at bin centers.
+    The line depicts the analytic result for one energy: $\varepsilon=15$~MeV.
+    Test described in Sec.~\ref{ssec:test_ab_star}}
   \label{fig:f_absorption_sphere}
 \end{figure}
-
-We can explain these features quantitatively by examining the limits
-of Eqn.~\ref{eqn:fae_homogeneous_general}, expanding the exponential function
-in powers of $\chi s$; the distribution function takes the limiting values
-\begin{equation}
-  \label{eqn:fae_homogeneous_limits}
-  f_{\rm AE}(\cos A) = f^{\rm eq}
-  \begin{cases}
-    \chi s(\cos A) & s \ll \chi^{-1}\,\textrm{(thin)}\\
-    1                  & s \gg \chi^{-1}\,\textrm{(thick)}.
-  \end{cases}
-\end{equation}
-These limits are represented in Fig.~\ref{fig:f_absorption_sphere}:
-with $\nu_x$ in the optically thin limit at all viewing angles,
-and $\nu_e$ in the optically thick limit at viewing angles
-$\cos A \gtrsim 0.8$.
 
@@
 \label{ssec:test_gravity}
-Unfinished: describe and point readers to Asano-Fukuyama test presented
-in thesis ~\cite[Sec.~4.3.3]{deat2015-thesis}.
-Also compare a few geodesics in Kerr to analytic.
-\footnote{TODO: compute some geodesics in Kerr}
+Optically thick star in Schwarszchild.
+Total luminosity at infinite separation may be computed by
+computing the radial energy flux at a finite separation from the star, $r$,
+and using $p_t$ for the energy. We compute this asymptotic luminosity
+at several radii, $r$, and show that the luminosity is independent of
+the observer's radius.
+\todo{compute $L$, put in figure}
 
@@
 \label{ssec:test_collapse}
-To test our scattering treatments we calculate neutrino fields outside a
-collapsed 15~${\rm M}_\odot$ star, 100~ms after core bounce,
-comparing ray tracing fields to those from a Monte Carlo transport calculation.
-Elastic scattering interior to the shock at $r\approx150$~km
-significantly modifies the neutrinos' spectra,
-\footnote{TODO: demonstrate}
-and the extended envelope outside the shock becomes a source of
-higher-than-average-energy neutrinos.
-
-The 1D matter profile and M1 transport evolution are computed using
-open source supernova evolution code \lstinline{GR1D}
-\footnote{\url{http://www.gr1dcode.org}}
-\cite{ocon2010-gr1d, ocon2015-gr1d_with_nu},
-using a progenitor profile from \cite{woos1995-sn_progenitors}.
-The matter is described by the LS180 equation of state \cite{latt1991-nuc_eos},
-and the opacities are computed and stored in a table
-as described in Sec.~\ref{ssec:test_equilibrium}.
-This standard test is also presented for example in
-\cite{ocon2015-gr1d_with_nu,fouc2015-m1_nsbh,abdi2012-monte_carlo}.
-\footnote{TODO: also lieb2005, mull2010}
-
-The matter profile and background field are stored on a spherical
-pseudo-spectral grid composed of 11 spherical-shell subdomains
-\footnote{TODO: cite early SpEC paper}
-comprising a total of 62 radial grid points spaced approximately
-logarithmically across $r\in(0,740)$~km.
-The background scattering field $\Phi(\varepsilon)$ is supplied by
-\lstinline{GR1D} in the form of $J(\tilde{\varepsilon}_i)$,
-with $\varepsilon_i$ representing 18 energy groups
-identical to those in the \lstinline{NuLib} table described in
-Sec.~\ref{ssec:test_equilibrium}.
-For the \emph{spectral} method we use $J(\tilde{\varepsilon}_i)$ directly,
-using zeroth-order interpolation between energy groups;
-for the \emph{gray} method we use $J=\sum J(\tilde{\varepsilon}_i) \Delta\tilde{\varepsilon}_i$,
-and $G=\sum J(\tilde{\varepsilon}_i) \Delta\tilde{\varepsilon}_i/\tilde{\varepsilon}_i$ instead,
-with $\Delta\tilde{\varepsilon}_i$ the bin-width of the $i$th energy group.
-
-For fiducial neutrino distributions we use the matter profile
-as input into a Monte Carlo radiation transport calculation using
-open source neutrino transport code \lstinline{Sedonu}
-\footnote{\url{https://bitbucket.org/srichers/sedonu}}
-\cite{rich2015-monte_carlo}.
-To homogenize the physics modeled across these three treatments
-(M1 transport to provide the background fields,
-ray tracing to compute neutrino distributions, and
-Monte Carlo transport for a fiducial comparison)
-we turn off inelastic scattering in the M1 and Monte Carlo codes,
-since we have not incorporated that physics into this ray tracing code,
-and we turn off general relativistic effects in the ray tracing code,
-since the Monte Carlo code assumes Minkowski spacetime.
-
-We place a stationary observer at $r=500$~km.
-Fig.~\ref{fig:avg_eps_collapse} shows the distribution of average $\nu_x$
-energies across incoming angles to this observer.
-We show $\nu_x$ because they present the largest scattering effects:
-they scatter through a thicker atmosphere outside their deep emission surface,
-and their hotter spectrum experiences stronger modification due to the
-$\varepsilon^2$ dependence of the scattering cross-section.
-Against the fiducial Monte Carlo distribution, we show the
-\emph{noscat} treatment, and the \emph{scat} treatment using both the
-\emph{spectral} method and the \emph{gray} method.
+[Standard radiation transport test described in~\cite{ocon2015-gr1d_with_nu},
+and in \cite[App.~E.6]{fouc2015-m1_nsbh},
+and in~\cite{abdi2012-monte_carlo}.]
+Plot in Fig.~\ref{fig:avg_eps_collapse}.
 
 \begin{figure}
-  %\includegraphics[width=\columnwidth]{fig-collapse_profile_avg_eps_vs_cosA}
-  \resizebox{\columnwidth}{!}{\input{fig-collapse_profile_avg_eps_vs_cosA.tex}}
-  % Data:   $DIR/
-  % Script: $DIR_GRAY_SC/../plot-compare_spectral_vs_gray-angular_distribution_avg_eps-production.gnu
-  % w/  DIR_MC     =$DIR/ResultsSherwood-03/Spectra
-  %     DIR_SPEC_SC=$DIR/ResultsBrett-A-03/Test-scat-01
-  %     DIR_GRAY_SC=$DIR/ResultsBrett-B-03/Test-scat-01
-  %     DIR_GRAY_AB=$DIR/ResultsBrett-B-03/Test-noscat-01
-  % and DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20161007-ray_tracing_tests/CollapseProfile/
-  \caption{Angular distribution of $\nu_x$ average energies
-    $\langle\tilde{\varepsilon}\rangle(\cos A)$
-    in the collapse profile test (described in Sec.~\ref{ssec:test_collapse}).
+  \includegraphics[width=\columnwidth]{fig-collapse_profile-avg_eps_vs_cosA}
+  \caption{Average energy $\langle\varepsilon\rangle=J/G$
+    in the collapse profile test.
     The observer is at 500~km, and the shock at 150~km,
-    so that the half-opening angle of the shock is $\cos A\approx0.954$.
+    so that the half-opening angle of the shock is $\cos A=0.954$.
     The four methods depicted are 1) a fiducial Monte Carlo calculation,
@@
     fields for scattering, and 4) ray tracing neglecting scattering.
-  }
+    Test described in Sec.~\ref{ssec:test_collapse}}
   \label{fig:avg_eps_collapse}
 \end{figure}
-
-As expected the average energies from both
-the scattering envelope ($\cos A \lesssim 0.95$) and
-the bright core ($\cos A\gtrsim 0.995 $)
-are well characterized by the \emph{scat} treatments
-and badly characterized by the \emph{noscat} treatment.
-In this case for $\nu_x$ the major effect of elastic scattering
-is to decrease the average energy of neutrinos coming from the core
-and increase the average energy of neutrinos coming from the envelope.
-
-Although not shown here, the angular distribution contributing to the
-total number density is also strongly affected by elastic scattering.
-Without elastic scattering, the central regions emitting 60\%
-of the neutrinos for the species $\{\nu_e,\bar{\nu}_e,\nu_x\}$
-have scales $R_\nu\approx\{40,35,20\}$~km;
-with elastic scattering the scales are
-$R_\nu\approx\{50,45,35\}$~km.
-\footnote{TODO: use a consistent lengthscale symbol throughout paper. $d$? $D$? $r$? $l$?}
-% From by-hand summation over histogram G(cosA_i)*dcosA
-% Script: $DIR/plot-compare_scat_vs_noscat-G_angular_distribution-all.gnu
-% DIR=.../20161007-ray_tracing_tests/CollapseProfile/ResultsBrett-B-03
-
-Total luminosities and average energies from the different treatments are
-\footnote{TODO: update MC results 02$\rightarrow0$3}
-presented in Tab.~\ref{tab:collapse_rt_vs_mc}.
-\footnote{TODO: give equation for L}
-As expected, $\nu_e$ is least affected by scattering, and $\nu_x$ most.
-In fact, without scattering, $\nu_x$ luminosities are overestimated by more than
-two orders of magnitude, due to the steep increase in temperature with depth
-in the inner core.
-Though contributing only a fraction of the total luminosity,
-the average energy of neutrinos scattered to the observer from the envelope
-outside the shock, $\langle\varepsilon\rangle^{\rm sc}$,
-is poorly characterized by the \emph{noscat} treatment for all species.
-By contrast, both the \emph{gray} and \emph{spectral} \emph{scat} treatments
-faithfully describe the high scattered energies from the envelope.
-
-\begin{table}%[h]
-  % to get mc luminosities: tail $MCDATA/Spectra/luminosities_and_avg_eps-501km-nu?.dat
-  % to get rt luminosities: tail $RTDATA/Moments-nu?.out
-  % MCDATA=$DIR/ResultsSherwood-02
-  % RTDATA_NO=$DIR/ResultsBrett-B-03/Test-noscat-01/
-  % RTDATA_GR=$DIR/ResultsBrett-B-03/Test-scat-01/
-  % RTDATA_SP=$DIR/ResultsBrett-A-03/Test-scat-01/
-  % DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20161007-ray_tracing_tests/CollapseProfile/
-  \caption{
-    From the post-bounce collapse profile test (Sec.~\ref{ssec:test_collapse}),
-    a comparison of total luminosities and average energies between
-    the following methods:
-    ray tracing ignoring elastic scattering `RT noscat',
-    ray tracing with the gray scattering treatment `RT scat gr',
-    ray tracing with the spectral scattering treatment `RT scat sp', and
-    a fiducial Monte Carlo transport evolution `MC'.
-    We also show average energies of the scattering envelope
-    $\langle\tilde{\varepsilon}\rangle^{\rm sc}$, which are estimated by eye
-    from plots like Fig.~\ref{fig:avg_eps_collapse} at $\cos A < 0.95$.
-    Luminosities have units $10^{52}\,{\rm erg}\,{\rm s}^{-1}$ and
-    average energies have units MeV.
-    The $\nu_x$ luminosities are per-species:
-    multiply by four to get the total heavy-lepton neutrino luminosities.
-  }
-  \label{tab:collapse_rt_vs_mc}
-  \begin{tabularx}{\columnwidth}{X r r r r}
-    & {\bf RT noscat} & \,\,{\bf RT scat gr} & \,\,{\bf RT scat sp} & \,\,{\bf MC} \\
-    \hline
-    $L_{\nu_e}$                                         &  3.53 & 3.14 & 3.04 & 3.48 \\
-    $L_{\bar{\nu}_e}$                                   & 5.19 & 3.05 & 3.03 & 3.01 \\
-    $L_{\nu_x}$                                         & 222.0  & 1.88 & 1.76 & 1.70 \\
-    $\langle \tilde{\varepsilon}_{\nu_e} \rangle$               &  10.6 & 10.6 & 10.9 & 11.0 \\
-    $\langle \tilde{\varepsilon}_{\bar{\nu}_e} \rangle$         & 14.2 & 13.0 & 13.8 & 13.7 \\
-    $\langle \tilde{\varepsilon}_{\nu_x} \rangle$               & 47.8 & 16.0 & 17.3 & 16.2 \\
-    $\langle \tilde{\varepsilon}_{\nu_e} \rangle^{\rm sc}$      & 4    & 14   & 17   & 16 \\
-    $\langle \tilde{\varepsilon}_{\bar{\nu}_e} \rangle^{\rm sc}$& 4    & 24   & 20   & 18 \\
-    $\langle \tilde{\varepsilon}_{\nu_x} \rangle^{\rm sc}$      & 2    & 28   & 31   & 27 \\
-    \hline
-  \end{tabularx}
-\end{table}
-
-We can make some quantitative sense of the scattered energies in
-Fig.~\ref{fig:avg_eps_collapse} and Tab.~\ref{tab:collapse_rt_vs_mc}
-using the solutions explored in the simple configurations above.
-In particular, the average energy in the scattering envelope
-($\cos A\lesssim 0.95$) is related to the spectrum in the direction
-of the interior ($\cos A \sim 1$).
-The relation may be derived by simplifying our realistic model
-to that of a homogeneous matter profile and background scattering field.
-
-By expanding the exponential of
-Eqns.~\ref{eqn:fse_homogeneous_general} and \ref{eqn:fae_homogeneous_general}
-as we did for Eqn.~\ref{eqn:fae_homogeneous_limits},
-and furthermore factoring out the dominant energy dependence from the
-cross-sections (i.e.\ $\chi\equiv\varepsilon^2 \zeta$
-with $\zeta$ approximately constant), we have
-\begin{align}
-  \label{eqn:fae_homogeneous_thin_limit}
-  f_{\rm AE}(\varepsilon)
-  &= \varepsilon^2 \zeta_a \,f^{\rm eq}(\varepsilon)\,s, \\
-  \label{eqn:fse_homogeneous_thin_limit}
-  f_{\rm SE}(\varepsilon)
-  &= \varepsilon^2 \zeta_s \,\Phi(\varepsilon)\,s,
-\end{align}
-where $s$ is the length of scattering envelope passed through by the ray.
-\footnote{TODO: should be $\chi_a^*=\varepsilon\zeta_a$ for $\nu_x$}
-In the envelope the local temperature is low so $f^{\rm eq}/\Phi\ll 1$,
-and also $\zeta_a/\zeta_s\ll 1$,
-so $f_{\rm SE}$ strongly dominates over $f_{\rm AE}$.
-The average energy of the scattered field measured by our observer
-is therefore
-\begin{equation}
-  \label{eqn:avg_eps_sc}
-  \langle \tilde{\varepsilon} \rangle^{\rm sc}
-  = \frac{\int d\varepsilon\,\varepsilon^3 f}{\int d\varepsilon\,\varepsilon^2 f}
-  \approx \frac{\int d\varepsilon\,\varepsilon^5 \Phi}{\int d\varepsilon\,\varepsilon^4 \Phi}.
-\end{equation}
-Note that we have taken the liberty here of identifying the fluid-frame energy
-$\varepsilon$ with the Eulerian energy $\tilde{\varepsilon}$,
-since infall velocities in the envelope are $\sim0.1c$,
-and as Fig.~\ref{fig:avg_eps_doppler} indicates,
-the Doppler shift will therefore introduce an error into our analysis of
-$\sim10\%$.
-
-The spectrum of the background field $\Phi(\varepsilon)$ is well-approximated
-as the \emph{scat} solution for $\cos A\sim 1$, since that is the dominant
-source direction for neutrinos.
-And because the scattering envelope is optically thin at all energies,
-we can assume the neutrino spectrum is essentially unchanged in its passage
-through the envelope.
-
-In order to estimate $\langle \tilde{\varepsilon} \rangle^{\rm sc}$ analytically,
-we must write the background field $\Phi(\varepsilon)$ analytically.
-Direct Fermi-Dirac fits using a temperature and chemical potential
-representative of a physical neutrinosurface fair poorly since neutrinos of
-different energies decouple from the matter at different radii, over which
-the thermodynamic state varies substantially.
-Phenomenological Fermi-Dirac fits work well;
-but so do pinched spectral fits which are much simpler
-\cite{keil2003-pinched_spectra,miri2016-sn_neutrinos}:
-\begin{equation}
-  \label{eqn:pinched_spectra}
-  \Phi^{\rm pi}(\varepsilon) \propto
-  \varepsilon^{\alpha-2}
-  \exp\left(-(\alpha+1)\frac{\varepsilon}{\langle\varepsilon\rangle}\right),
-\end{equation}
-(where our definition differs from that of \cite{keil2003-pinched_spectra}
-by a factor of
-$\varepsilon^2$, since we define our distribution function to be dimensionless
-in natural units $\{\hbar,c\}=1$).
-
-Pinching parameters
-(calculated by eye from the \emph{spectral} \emph{scat} method)
-for the species
-$\{\nu_e,\bar{\nu}_e,\nu_x\}$ are $\alpha\approx\{3.6,5.1,2.3\}$.
-Energy moments of pinched spectra like those in Eqn.~\ref{eqn:avg_eps_sc}
-have a simple analytic form so that Eqn.~\ref{eqn:avg_eps_sc} becomes
-\begin{align}
-  \label{eqn:pinched_avg_eps_sc_1}
-  \langle\tilde{\varepsilon}\rangle^{\rm sc}
-  &\approx \frac{\alpha+3}{\alpha+1} \langle\varepsilon\rangle, \\
-  %\label{eqn:pinched_avg_eps_sc_2}
-  %&= \{1.4,1.3,1.6\} \langle\varepsilon\rangle, \\
-  \label{eqn:pinched_avg_eps_sc_3}
-  &\approx \{16,18,28\}\,~{\rm MeV},
-\end{align}
-again for $\{\nu_e,\bar{\nu}_e,\nu_x\}$ respectively.
-These analytic predictions agree with the average energies of the scattering
-envelope to approximately 10\%
-% From by-eye examinations of <eps>(cosA) plot using data from
-% MCDATA=$DIR/ResultsSherwood-02
-% RTDATA_GR=$DIR/ResultsBrett-B-03/Test-scat-01/
-% RTDATA_SP=$DIR/ResultsBrett-A-03/Test-scat-01/
-%          nue     nua     nux
-of all of the treatments including elastic scattering presented
-in Tab.~\ref{tab:collapse_rt_vs_mc}, except for
-$\bar{\nu}_e$ in the \emph{gray} \emph{scat} treatment,
-which deviates from our prediction by 30\%.
-This agreement is excellent despite the drastic
-simplifications used in our toy model.
-
-The $\langle\tilde{\varepsilon}_{\bar{\nu}_e}\rangle^{\rm sc}$ prediction
-in the \emph{gray} \emph{scat} treatment is approximately 30\%
-larger than the the fiducial Monte Carlo estimate. This is
-due to the large negative local chemical potential
-$\eta_{\bar{\nu}_e}\sim-10$ in the scattering envelope.
-As described in Sec.~\ref{ssec:sources_se},
-we use $\eta_\nu$ in the \emph{gray} treatment to construct our
-synthetic spectrum from total neutrino densities $J$ and $G$.
-This error and our successful analysis using pinched spectra above,
-points the way to future improvements to the \emph{gray} method:
-making better assumptions about the spectra which don't depend upon
-local fluid quantities.
 
@@
 \label{ssec:test_disk_comparison}
-The merger of two neutron stars by gravitational wave emission produces a
-postmerger configuration composed of a single neutron star surrounded by a disk.
-Because of strong differential rotation and shock heating,
-the remnant temporarily avoids
-collapse to a black hole, even if its mass exceeds the threshold of dynamical
-instability for a rigidly rotating neutron star \cite{duez2009-review}.
-These objects, called hypermassive neutron stars,
-may avoid collapse for $X\textrm{--}Y$~s,
-\footnote{TODO: cite $X$, $Y$}
-depending upon a number of physical factors including thermal pressure,
-magnetic fields, and the microphysics of the fluid.
-
-Such a configuration was modeled in \cite{fouc2016-m1_nsns} by
-evolving fluid and spacetime through the
-final inspiral and merger of two identical neutron stars of
-isolated gravitational mass 1.2~$M_{\odot}$. The postmerger configuration
-was simulated using a gray M1 transport scheme for the neutrinos, evolving the
-energy density, number density, and energy flux in addition to the standard
-fluid and metric variables for $\sim10\,{\rm ms}$ following merger
-\cite{fouc2016-m1_evolve_n}.
-The fluid was modeled using the LS220 equation of state \cite{latt1991-nuc_eos}.
-
-We use a single time snapshot from that configuration
-at $t=11\,{\rm ms}$ after merger.
-In this way we approximate the system as stationary over a light-crossing time
-of around 1~ms, which is far below the thermal timescale of the remnant.
-Figs.~\ref{fig:nsns_rho_merid}--\ref{fig:nsns_temp_equat}
-show slices of density and temperature from the finite difference
-fluid and M1 radiation data.
-These data were evolved on a rectilinear grid spanning approximately 400~km in
-both the $x$ and $y$ directions, and 150~km in the $z$ direction in our
-evolution coordinates.
+Total and relative number and energy luminosities for the disk evolved
+in \cite{fouc2016-m1_evolve_n}.
+Agreement better than 10\%.
+\todo{put in table comparison M1 and RT}
 
@@
   \includegraphics[width=\columnwidth]{production-colormap-merid-rho}
-  % VTK Data: zosima:/home/mbdeaton/Research/20170602-nsns_vol_data
-  % Plots:    $DIR/production-colormap-{merid,equat}-{rho,temp}.png
-  % w/ DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20170420-nsns_volume_data/02-two_domains_avg_eps/PlotsEtc
   \caption{A meridional slice of density in the hypermassive neutron
-    star and disk configuration (Sec.~\ref{ssec:test_disk_comparison}).
-    The distorted rectangular boundaries are the boundaries of the grid
-    used in the numerical simulation, which employs a coordinate mapping to
-    concentrate points near the central object.}
+    star and disk configuration.}
   \label{fig:nsns_rho_merid}
@@
   \includegraphics[width=\columnwidth]{production-colormap-equat-rho}
-  % see notes from production-colormap-merid-rho
   \caption{An equatorial slice of density in the hypermassive neutron
-    star and disk configuration (Sec.~\ref{ssec:test_disk_comparison}).}
+    star and disk configuration.}
   \label{fig:nsns_rho_equat}
@@
   \includegraphics[width=\columnwidth]{production-colormap-merid-temp}
-  % see notes from production-colormap-merid-rho
   \caption{A meridional slice of temperature in the hypermassive neutron
-    star and disk configuration (Sec.~\ref{ssec:test_disk_comparison}).}
+    star and disk configuration.}
   \label{fig:nsns_temp_merid}
@@
   \includegraphics[width=\columnwidth]{production-colormap-equat-temp}
-  % see notes from production-colormap-merid-rho
-  \caption{An equatorial slice of temperature in the hypermassive neutron
-    star and disk configuration (Sec.~\ref{ssec:test_disk_comparison}).}
+  \caption{A meridional slice of temperature in the hypermassive neutron
+    star and disk configuration.}
   \label{fig:nsns_temp_equat}
@@
 
-We extrapolate the fluid and M1 radiation data from
-the domain shown in Figs.~\ref{fig:nsns_rho_merid}--\ref{fig:nsns_temp_equat}
-to a larger ray tracing domain by setting all fluid and M1 radiation variables
-to their floor values outside the smaller domain.
-This simple extrapolation is adequate for ray tracing since
-the neutrinos are almost entirely free-streaming outside the smaller domain.
-Since the metric data were evolved on this larger domain no extrapolation is
-needed for them.
-The larger domain is represented as a pseudospectral grid
-composed of a sphere with concentric shells extending to $r\approx1400$~km.
-Radial grid spacings are $\Delta r \approx 0.15$~km in the star
-and $\Delta r \approx 2.5$~km in the disk,
-with 12 cells spanning polar angles $\theta\in[0,\pi]$
-and 24 cells spanning azimuthal angles $\phi\in[0,2\pi]$.
-Though a pseudospectral representation of non-smooth hydrodynamic data
-introduces some Gibbs-like oscillations in the variables,
-we choose to use this representation instead of the mesh-refined
-finite difference grid of the evolution because the pseudospectral
-representation uses much less memory:
-95~MB in pseuspectral vs.\ 3~GB in finite difference representation.
-Obviously, memory loads of this order are not insurmountable;
-but they would require some modifications to our
-volume data interpolation infrastructure.
-For the purposes of this analysis the pseudospectral representation is adequate.
-
-Opacities are computed from the LS220 equation of state using
-\lstinline{NuLib} and are stored as a table covering
-energy, density, temperature, and composition ranges
-spanning $\varepsilon\in[1,280.5]$~MeV logarithmically,
-$\varrho\in[10^6,3.2\times10^{15}]\,{\rm g}\,{\rm cm}^{-3}$ logarithmically,
-$T\in[0.05,150]$~MeV logarithmically, and 
-$Y_e\in[0.035,0.55]$ linearly,
-with grid extents $\{18,82,65,51\}$ respectively.
-
-We place stationary (i.e.\ Eulerian) observers at fixed coordinate radius
-$r=250$~km in the $y$-$z$ plane along an arc at $N_\theta$ positions
-distributed linearly in $\cos\theta$ over the northern hemisphere
-using $\theta_i=\cos^{-1}\left(1-i/(N_\theta-1)\right)$
-for $i=0,1,2,\ldots N_\theta-1$.
-We choose $N_\theta=30$.
-
-Though the ray tracing sampling of $f$ is done in full general relativity,
-in our calculation of moments at the observer positions,
-we make the simplifying assumption of Minkowski spacetime.
-This assumption introduces errors in our moments of order
-$GM/r=4\,{\rm km}/250\,{\rm km}$
-\footnote{TODO: too simplistic}
-(with $G$ the gravitational constant and $M$ the central mass),
-or $\sim 1\%$.
-
-Each observer samples the distribution function over a uniform grid
-in energy, cosine of polar angle, and azimuthal angle with extents
-$\{N_{\tilde{\varepsilon}},N_A, N_B\}$ spanning the ranges
-$\tilde{\varepsilon}\in(0,100)$~MeV,
-$\cos A\in\left((\cos A)_{\rm min}, 1\right)$, and
-$B\in[0,2\pi)$.
-The integrals for fluxes of number and energy
-(Eqns.~\ref{eqn:Ha}, \ref{eqn:Ka}, and \ref{eqn:J_H_S_eps_integrated})
-become sums over all rays.
-Taking these fluxes in the radial direction,
-and using the midpoint rule to convert the integral to a sum, we have
-\begin{align}
-  \label{eqn:Kr_minkowski}
-  \tilde{K}^r &= \Delta
-  \sum\limits_{m=0}^{N-1}
-  \tilde{\varepsilon}_m^2 \cos A_m f_m,\\
-  \label{eqn:Hr_minkowski}
-  \tilde{H}^r &= \Delta
-  \sum\limits_{m=0}^{N-1}
-  \tilde{\varepsilon}_m^3 \cos A_m f_m,
-\end{align}
-with $f_m \equiv f(\tilde{\varepsilon}_m,\cos A_m, B_m)$,
-$\Delta \equiv \Delta\tilde{\varepsilon}\, \Delta(\cos A)\, \Delta B$,
-$N \equiv N_{\tilde{\varepsilon}}\, N_A\, N_B$,
-and $m$ the index labeling each ray.
-Average energies are given by
-$\langle\tilde{\varepsilon}\rangle(\theta)=H^r(\theta)/K^r(\theta)$.
-We choose $N_{\tilde{\varepsilon}}=30$, $N_A=150$, and $N_B=20$.
-To maintain high angular resolution, we only sample rays that pass within
-approximately 120~km of the star's center by setting $(\cos A)_{\rm min}=0.88$.
-
-We also combine measurements from all observers to estimate
-total luminosities and averages over the sky.
-Since we have chosen an arc of observers isolated to the northern hemisphere and
-the $y$-$z$ plane, we may extend these data to the full sky by assuming that
-emission is azimuthally symmetric and reflection symmetric across the
-equatorial plane. Figs.~\ref{fig:nsns_rho_merid}--\ref{fig:nsns_temp_equat}
-indicate the approximate validity of these assumptions at 11~ms after merger.
-Total luminosities are then computed as integrals of the radial fluxes
-over $\cos\theta$.
-Using the trapezoid rule the number and energy luminosities become sums:
-\begin{align}
-  \label{eqn:luminosity_R}
-  R &= C
-  \left(\frac{1}{2}(\tilde{K}^r_0+\tilde{K}^r_{N_\theta-1})
-  +\sum\limits_{m=1}^{N_\theta-2}\tilde{K}^r_m\right),\\
-  \label{eqn:luminosity_L}
-  L &= C
-  \left(\frac{1}{2}(\tilde{H}^r_0+\tilde{H}^r_{N_\theta-1})
-  +\sum\limits_{m=1}^{N_\theta-2}\tilde{H}^r_m\right),
-\end{align}
-with $C \equiv 4\pi r^2\, \Delta(\cos \theta)$,
-\footnote{TODO: source of GR error in coordinate $r$}
-$\tilde{K}^r_m \equiv \tilde{K}^r(\theta_m)$, and
-$\tilde{H}^r_m \equiv \tilde{H}^r(\theta_m)$.
-Average energies over the whole sky are then
-$\langle\tilde{\varepsilon}\rangle=L/R$.
-
-Tab.~\ref{tab:nsns_rt_vs_m1} compares the all-sky
-luminosities and average energies from ray tracing and from
-the M1 transport simulation,
-which serves for qualitative comparison.
-Even if M1 and ray tracing methods both provide faithful measurements
-of all-sky luminosities, we do not expect perfect agreement since the two
-treatments differ fundamentally. In addition to differences in the physics
-such as the treatment of inelastic scattering,
-M1 fluxes are integrated over the outer boundary of the finite-difference grid
-at radii as small as 75~km,
-whereas the ray tracing fluxes are integrated over a sphere at radius 250~km,
-introducing a time lag between the two measurements of under a millisecond.
-Furthermore uncertainties from the M1 simulation are around
-15\% for energy luminosities and
-10\% for average energies \cite[Sec.~A.6]{fouc2016-m1_evolve_n}.
-
-\begin{table}%[h]
-  % Total fluxes over the sky integrated in google sheet: 20171031-theta_distribution_integrations
-  % RTDATA_NO=$DIR/PositionDistributions-nu?-250km-Lev2-e30a150b20-noscat-04.dat
-  % RTDATA_SC=$DIR/PositionDistributions-trim-nu?-250km-Lev2-e30a150b20-scat-04.dat
-  % DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20170324-ray_tracing_nsns/Run03/PositionDistributions
-  \caption{
-    Comparison of total luminosities and average energies
-    of the hypermassive neutron star configuration
-    (presented in Sec.~\ref{ssec:test_disk_comparison}).
-    The methods are ray tracing using the \emph{noscat} method `RT noscat',
-    ray tracing using the \emph{gray} \emph{scat} method `RT scat',
-    and the M1 transport simulation `M1'.
-    The ray tracing totals were computed from sums over
-    observers placed in the $y$-$z$ plane.
-    M1 values are taken from \cite[Figs. 7, 9, 10]{fouc2016-m1_evolve_n}.
-    Energy luminosities have units $10^{52}\,{\rm erg}\,{\rm s}^{-1}$,
-    number luminosities $10^{57}\,{\rm s}^{-1}$, and
-    average energies MeV.
-    The $\nu_x$ luminosities are per-species:
-    multiply by four to get the total heavy-lepton neutrino luminosities.
-  }
-  \label{tab:nsns_rt_vs_m1}
-  \begin{tabularx}{\columnwidth}{X r r r}
-    & {\bf RT noscat} & \,\,{\bf RT scat} & \,\,{\bf M1} \\
-    \hline
-    $L_{\nu_e}$                                 & 5.87 & 5.40 & 5 \\
-    $L_{\bar{\nu}_e}$                           & 9.70 & 10.7 & 12 \\
-    $L_{\nu_x}$                                 & 23.6 & 12.5 & 12 \\
-    $R_{\nu_e}-R_{\bar{\nu}_e}$                 & -0.91 & -1.68 & -2 \\
-    $\langle \varepsilon_{\nu_e} \rangle$       & 12.7 & 12.0 & 12 \\
-    $\langle \varepsilon_{\bar{\nu}_e} \rangle$ & 16.0 & 14.8 & 15 \\
-    $\langle \varepsilon_{\nu_x} \rangle$       & 34.2 & 23.3 & 26 \\
-    \hline
-  \end{tabularx}
-\end{table}
-
-Because the model was evolved with a gray M1 transport scheme,
-the only \emph{scat} method we can use in our ray tracing is the
-\emph{gray} method.
-As in the post-bounce configuration in Sec.~\ref{ssec:test_collapse},
-the \emph{scat} treatment is more faithful than the \emph{noscat} treatment,
-significantly so for the heavy-lepton neutrinos.
-For example, with scattering turned off the ray tracing and M1 measurements
-of $L_{\nu_x}$ disagree by 200\%;
-but with scattering turned on agreement is within 10\%.
-And as the ray tracing and M1 measurements both show,
-$R_{\bar{\nu}_e}$ dominates over $R_{\nu_e}$ significantly,
-because the disrupted neutron star material is releptonizing.
-
-The distributions of radial fluxes over observer position are shown
-in Figs.~\ref{fig:nsns_theta_distrib_Kr} and \ref{fig:nsns_theta_distrib_Hr}.
-With the \emph{noscat} treatment the $\nu_x$ luminosities are relatively
-constant in $\theta$, because the disk is optically thin to $\nu_x$,
-and most of the neutrinos come from the hypermassive neutron star,
-which is roughly spherical.
-The slight upward trend in $\nu_x$ luminosities with $\theta$ may be due to
-asymmetries in the configuration.
-It is not due to the Doppler shift or relativistic beaming
-from the rapid rotation of the star,
-a hypothesis we tested by setting $u_i=0$ and $W=1$ in the ray
-tracing equations.
-Because of their larger absorption cross section,
-the disk is not optically thin to $\nu_e$ and $\bar{\nu}_e$,
-and observers within view of the hot hypermassive neutron star measure
-the largest radial fluxes of these species.
-\begin{figure}
-  %\includegraphics[width=\columnwidth]{theta_distrib-250km-Kr}
-  \resizebox{\columnwidth}{!}{\input{theta_distrib-250km-Kr.tex}}
-  % Data_a:   $DIR/PositionDistributions/*Lev2-e30a150b20-noscat-04*
-  % Data_s:   $DIR/PositionDistributions/*Lev2-e30a150b20-scat-04*
-  % Data gathered using: $DIR/CollectPositionDistributions.sh
-  % Script: $DIR/GnuScripts/plot-theta_distributions-Kr-fixed_r.gnu
-  % w/ DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20170324-ray_tracing_nsns/Run03
-  \caption{Radial number fluxes, $\tilde{K}^r(\theta)$,
-    as defined in Eqn.~\ref{eqn:Kr_minkowski},
-    in the hypermassive neutron star configuration
-    (presented in Sec.~\ref{ssec:test_disk_comparison}).
-    Sampled for observers at fixed coordinate distances from the center
-    of the star, with $r=250\,{\rm km}$.
-  }
-  \label{fig:nsns_theta_distrib_Kr}
-\end{figure}
-
-\begin{figure}
-  %\includegraphics[width=\columnwidth]{theta_distrib-250km-Hr}
-  \resizebox{\columnwidth}{!}{\input{theta_distrib-250km-Hr.tex}}
-  % see notes from theta_distrib-250km-Kr
-  % Script: $DIR/GnuScripts/plot-theta_distributions-Hr-fixed_r.gnu
-  \caption{Same as Fig.~\ref{fig:nsns_theta_distrib_Kr} but radial energy
-    fluxes $\tilde{H}^r(\theta)$ as defined in Eqn.~\ref{eqn:Hr_minkowski}.
-  }
-  \label{fig:nsns_theta_distrib_Hr}
-\end{figure}
-
-When we turn on scattering, the disk is no longer transparent to $\nu_x$,
-and $\nu_x$ luminosities present the same qualitative $\theta$-dependence
-as that of the other species.
-In general, scattering increases luminosities for polar observers
-\footnote{TODO: in compute-eta data not true anymore for $\nu_e$}
-($\theta \sim 0^\circ$) because additional high energy neutrinos
-from the hypermassive neutron star are scattered by the disk back up to
-the observer.
-For observers in the equatorial plane ($\theta \sim 90^\circ$)
-the effect of scattering is a decrease in energy luminosities,
-since some of the highest energy neutrinos from deep in the interior
-are scattered away from the line of sight by the disk.
-
-Fig.~\ref{fig:nsns_theta_distrib_avg_eps} shows the distribution of
-average energies of radial fluxes over observer positions.
-With and without scattering, the average energies are highest
-for observers near the polar axis, since polar observers get a
-direct view of the hot hypermassive neutron star.
-(This trend is unexpectedly reversed for $\nu_x$,
-and may be due to asymetries in the disk, as discussed above.)
-Scattering decreases average energies across all observer positions,
-as it did in the post-bounce configuration, or any configuration of
-a hot interior surrounded by a scattering envelope.
-The strength of the effect of scattering on the average energies of
-the different species is seen to follow the ranking
-$\nu_x>\bar{\nu}_e>\nu_e$ due to two factors:
-the average energies of the spectra obey the same ranking,
-and the thicknesses of the different species' scattering envelopes
-also obey the same ranking.
-As we showed in the toy models with a homogeneous background scattering field
-the scattering contribution $f_{\rm SE}$ is
-proportional to energy $\varepsilon^2$ and path length $s$
-(Eqn.~\ref{eqn:fse_homogeneous_thin_limit}).
-\begin{figure}
-  %\includegraphics[width=\columnwidth]{theta_distrib-250km-avg_eps}
-  \resizebox{\columnwidth}{!}{\input{theta_distrib-250km-avg_eps.tex}}
-  % see notes from theta_distrib-250km-Kr
-  % Script: $DIR/GnuScripts/plot-theta_distributions-avg_eps_from_fluxes-fixed_r.gnu
-  \caption{Same as Fig.~\ref{fig:nsns_theta_distrib_Kr} but
-    average energies of radial fluxes
-    $\langle\tilde{\varepsilon}\rangle(\theta)=\tilde{H}^r(\theta)/\tilde{K}^r(\theta)$,
-    as defined in Eqns.~\ref{eqn:Kr_minkowski} and \ref{eqn:Hr_minkowski}.
-  }
-  \label{fig:nsns_theta_distrib_avg_eps}
-\end{figure}
-
-In Figs.~\ref{fig:nsns_cosA_distrib_G_00deg}
-and \ref{fig:nsns_cosA_distrib_G_90deg}, we show the distributions of neutrino
-number density over incoming angle
-for the observer on the polar axis $\theta=0^\circ$,
-and in the equatorial plane $\theta=90^\circ$,
-defined analagously to Eqn.~\ref{eqn:J_per_cosA}.
-As in the case of the collapse profile in Sec.~\ref{ssec:test_collapse},
-the dominant effect of elastic scattering on $\tilde{G}(\cos A)$
-is to spread the distribution out to larger angles, by generally decreasing
-the number of neutrinos coming from the core while increasing the number
-of neutrinos coming from the disk.
-\footnote{TODO: with compute-eta data also explain why $G_{\nu_e}>G_{\bar{\nu}_e}$
-  for $\theta=0$ at $\cos A>0.995$.}
-
-For the observer in the equatorial plane,
-Fig.~\ref{fig:nsns_cosA_distrib_G_90deg}, the disk is so optically thick to
-$\nu_e$ that there is very little difference between \emph{scat} and
-\emph{noscat} treatments for incoming angles $\cos A \lesssim 0.95$
-corresponding to the volume inside of $r \lesssim 80$~km.
-Also the stepped temperature gradient in the disk's spiral arms visible in
-Figs.~\ref{fig:nsns_temp_merid} and \ref{fig:nsns_temp_equat} presents as a
-stepped heavy-lepton neutrino number density distribution in
-Fig.~\ref{fig:nsns_cosA_distrib_G_90deg} in the \emph{noscat} treatment,
-since the pair production rate due to $e^{-}e^{+}$ annihilation,
-producing $\nu_x$, is especially sensitive to temperature, going as
-$T^9$ \cite[Sec.~7]{burr2006-neutrino_opacities}.
-This stepped distribution is not visible in $\nu_e$ and $\bar{\nu}_e$ emission,
-since we ignore pair processes for these species due to the dominance of
-absorption and emission processes, obeying a shallower temperature-dependence;
-nor is it visible in the \emph{scat} treatment of $\nu_x$, since scattering
-tends to smear the incoming angle of the emission;
-nor is it visible for any species or treatment in
-Fig.~\ref{fig:nsns_cosA_distrib_G_00deg}, since the integration over the
-azimuthal angle $B$ averages out any spiral structure for the observer
-on the polar axis.
-
-\begin{figure}
-  %\includegraphics[width=\columnwidth]{cosA_distrib-intensity-250km-00deg}
-  \resizebox{\columnwidth}{!}{\input{cosA_distrib-G-250km-00deg.tex}}
-  % Data:   $DIR/Lev2-e30a150b20-{,no}scat-04
-  % Script: $DIR/plot-angular_distribution-number_density-compare.gnu
-  % w/ DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20170324-ray_tracing_nsns/Run03/000Mo-000.0deg-250km
-  \caption{Distribution of number density $\tilde{G}(\cos A)$ over incoming angle,
-    defined analagously to Eqn.~\ref{eqn:J_per_cosA}.
-    The integral of this quantity over $d\cos A$ gives the total number density
-    measured by this observer on the rotation axis.
-    Volume data from the hypermassive neutron star configuration
-    (presented in Sec.~\ref{ssec:test_disk_comparison}).}
-  \label{fig:nsns_cosA_distrib_G_00deg}
-\end{figure}
-
-\begin{figure}
-  %\includegraphics[width=\columnwidth]{cosA_distrib-intensity-250km-90deg}
-  \resizebox{\columnwidth}{!}{\input{cosA_distrib-G-250km-90deg.tex}}
-  % Data:   $DIR/Lev2-e30a150b20-{,no}scat-04
-  % Script: $DIR/plot-angular_distribution-number_density-compare.gnu
-  % w/ DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20170324-ray_tracing_nsns/Run03/000Mo-090.0deg-250km
-  \caption{Distribution of number density $\tilde{G}(\cos A)$ over incoming angle,
-    as in Fig.~\ref{fig:nsns_cosA_distrib_G_00deg},
-    but with the observer in the equatorial plane.}
-  \label{fig:nsns_cosA_distrib_G_90deg}
-\end{figure}
-
 \section{Applications}
@@
 
-\subsection{Neutrino Oscillations at High Neutrino Densities}
+\subsection{Understanding Emission Spectra}
+\label{ssec:spectra}
+In this section, we use ray tracing's ability to track ``where the neutrinos
+are coming from'' to understand the features in the emission distributions
+presented in the code tests.
+
+The merger models are described in \cite{fouc2015-m1_nsbh, fouc2016-m1_nsns}.
+Something like this is presented in \cite[Figs.~10-11]{pere2014-nu_wind};
+compare.
+
+\subsection{Neutrino Self-Interaction Potential}
 \label{ssec:V_nunu}
-In this section we examine the importance of elastic scattering in neutrino
-flavor oscillation above the hypermassive neutron star--disk configuration
-presented in Sec.~\ref{ssec:test_disk_comparison}.
-Our treatment of the flavor evolution equation assumes flat space and small fluid
-velocities.
-In consequence we treat some of gauge-dependent quantities inconsistently,
-and we ignore potentially important features of relativistic flavor evolution
-near compact objects.
-\footnote{TODO: cite Yang et Kneller}
-Such a cursory treatment is sufficient, however, for the following qualitative
-exploration.
-
-The Boltzmann Equation (Eqn.~\ref{eqn:boltzmann}) is one limiting
-form of the quantum kinetic equations for neutrinos
-\cite{vlas2014-neutrino_qkes},
-the limit where collisional mean free paths are much shorter
-than oscillation lengths, i.e.\ $\varepsilon/\mathscr{K} \ll |H|^{-1}$
-with $H$ the Hamiltonian describing coherent forward scattering interactions.
-In this limit the neutrino density matrix takes the form
-\begin{equation}
-  \label{eqn:density_rho_0}
-  \rho_0(\varepsilon,\ell_\mu) =
-  \sum\limits_{\alpha}f^{\nu_\alpha}(\varepsilon,\ell_\mu)
-  \ket{\nu_\alpha} \bra{\nu_\alpha},
-\end{equation}
-with $\ket{\nu_\alpha}$ the neutrino flavor eigenstates.
-The neutrino state remains pure, or diagonal in the flavor basis.
-
-But at the opposite limit, in the free-streaming regime,
-coherence effects become important,
-and the quantum kinetic equations take a Schr\"{o}dinger-like form:
-\begin{equation}
-  \label{eqn:schroedinger}
-  i \frac{d}{ds} S = H S,
-\end{equation}
-where $ds$ is an interval of proper length traversed by the neutrino as
-measured by our observer, and
-\footnote{TODO: mention $\bar{\nu}$ eqn}
-$S$ is the flavor evolution matrix describing a mixed state
-\begin{equation}
-  \label{eq:density_evolution}
-  \rho = S \, \rho_0 \, S^\dagger.
-\end{equation}
-In this limit, we may decompose the Hamiltonian into vacuum,
-matter, and neutrino contributions:
-\begin{equation}
-  \label{eqn:decompose_H}
-  H = H_{\rm V} + H_e + H_{\nu\nu}.
-\end{equation}
-Explicit formulas for these terms may be found in the oscillation literature,
-e.g.\ \cite{duan2009-review},
-but for clarity we give their order-of-magnitude scales here:
-\begin{align}
-  \label{eqn:scale_H_V}
-  |H_{\rm V}| &\sim \frac{\Delta m^2}{\varepsilon}, \\
-  \label{eqn:scale_H_e}
-  |H_e|       &\sim G_{\rm F}\left|n_{e^-}-n_{e^+}\right|, \\
-  \label{eqn:scale_H_nunu}
-  |H_{\nu\nu}|&\sim G_{\rm F} \left|G_{\nu_e}-G_{\bar{\nu}_e}\right|,
-\end{align}
-where $\Delta m^2$ is the mass-squared differences between neutrino mass
-eigenstates,
-$G_{\rm F}$ is the Fermi coupling constant,
-$n_{e^-}$ and $n_{e^+}$ are the electron and positron number densities,
-and $G_{\nu_e}$ and $G_{\bar{\nu}_e}$ the electron neutrino and antineutrino
-number densities defined in Eqn.~\ref{eqn:G}.
-For the scale of $H_{\nu\nu}$ we have only included the isotropic
-components of the neutrino fields.
-
-In regimes in which $|H_{\nu\nu}|\ll|H_{\rm V},H_e|$,
-the flavor evolution along the ray is soluble, and reveals a rich
-and physically important phenonenology including
-vacuum, solar, atmospheric, and terrestrial oscillations,
-as well as oscillations in supernova envelopes.
-Where neutrino densities are relatively high, however,
-as in neutron star mergers, the problem must be solved globally.
-\footnote{TODO: be clearer}
-To date, no method has been devised to handle this problem.
-
-However, we can solve a similar but tractable problem along a single ray.
-Assume all the neutrino rays intersecting an event along a given test ray
-have undergone the same flavor evolution history as that of the test ray,
-i.e.\ $S$ is the same for all rays sharing that event.
-This is the so-called single-angle approximation.
-\footnote{TODO: cite someone}
-\footnote{TODO: say something about its validity}
-Though widely used in the supernova oscillation literature
-and shown to be qualitatively faithful in those environments,
-\footnote{TODO: cite vlasenko?}
-a recent study has shown that the single-angle approximation
-may mask certain instabilities in the flavor evolution
-\cite{wu2017-fast_neutrino_conversions}.
-
-Within the single-angle approximation,
-solving for the flavor evolution history along a given test ray
-requires knowledge of the unoscillated neutrino contribution to the Hamiltonian
-along that trajectory.
-This is a function of the unoscillated neutrino self-interaction potential,
-which for the $\alpha$-th flavor is:
-\begin{equation}
-  \label{eqn:V_nu_alpha}
-  V_{\nu_\alpha,0}(\varepsilon,\ell_\mu) =
-  \frac{\sqrt{2}G_{\rm F}}{(2\pi)^3} \varepsilon^2
-  \oint d\Omega'(1-\varpi')f^{\nu_\alpha}(\varepsilon,\ell'_\gamma),
-\end{equation}
-with the test ray propagating in direction $\ell_\mu$,
-ambient rays propagating in directions $\ell'_\gamma$,
-and the cosine of the angle between these given by
-$\varpi'=\psi^{\mu\gamma} \ell_\mu \ell'_\gamma$.
-As in the moment equations (Eqns.~\ref{eqn:J}-\ref{eqn:Ka}),
-the integral is taken over all directions $\ell'_\gamma$.
-
-If conditions are right, a resonant flavor transition
-introducing significant mixing may occur very near
-the point where neutrinos begin free-streaming.
-The matter-neutrino resonance
-\cite{malk2012-mnr_1,malk2015-mnr_2,malk2016-mnr_3},
-can occur where the matter potential and
-neutrino self-interaction potentials cancel,
-or where $V_e - V_{\nu\nu,0}=0$, with
-\begin{align}
-  \label{eqn:V_e}
-  V_e &= \sqrt{2} G_{\rm F} \frac{\varrho}{m_N} Y_e, \\
-  \label{eqn:V_nunu0}
-  V_{\nu\nu,0} &= \int d\varepsilon
-  \big(V_{\nu_e,0}(\varepsilon)-V_{\bar{\nu}_e,0}(\varepsilon)\big),
-\end{align}
-with $\varrho$ the rest density and $m_{N}$ the nucleon mass.
-The matter potential $V_e$ is always positive;
-and far outside the accretion disks formed in neutron star mergers,
-in which the disrupted neutron-star matter is rapidly releptonizing,
-the total unoscillated neutrino self-interaction potential $V_{\nu\nu,0}$
-is large and negative.
-
-We examine this effect in the post-merger configuration already analyzed
-in Sec.~\ref{ssec:test_disk_comparison}.
-We calculate the self-interaction potential along a radial coordinate trajectory
-making an angle $\theta=25^\circ$ with the rotation axis,
-at $N_s=7$ positions $r\in\{30,50,82,135,223,368,608\}$~km.
-We sample distribution functions at each of these positions over a grid
-with extents $N_\varepsilon=30$, $N_A=200$, $N_B=30$;
-energies range over $\varepsilon\in(0,100)$~MeV;
-polar angles range over $\cos A \in ((\cos A)_{\rm min}, 1)$, with
-$(\cos A)_{\rm min}\in\{-1,-1,-1,0.488,0.849,0.947,0.981\}$
-for each of the $N_s$ positions;
-and azimuthal angles range over the whole sky $B\in[0,2\pi)$.
-We interpolate the logarithm of the self-interaction potentials
-(Eqn.~\ref{eqn:V_nu_alpha})
-in path length $\log s$ along the ray,
-by fitting a 3rd order spline with continuous derivative,
-across the $N_s$ observation points.
-For $r>608$~km we extrapolate the self-interaction potentials using the
-geometric fall-off of $r^{-4}$ applicable to the far-field limit of the
-self-interaction potential \cite{malk2016-mnr_3}.
-
-The matter potential along this test ray is from an analytic wind
-model qualitatively consistent with the densities in the
-simulated volume (i.e.\ inside $r\sim70$~km) and asymptoting to the $r^{-2}$
-density field of a spherical steady-state wind with a constant asymptotic
-velocity.
-The velocity and density fields of a spherical steady-state wind
-obey the continuity equation $\rho(r)v(r)r^2=\rho_1 v_1 r_1^2$,
-with $\rho_1$ and $v_1$ the density and velocity measured at some fiducial
-radius $r_1$.
-For velocity field we choose a phenomenological wind model used in the
-oscillation literature\footnote{TODO: cite Surman \& McLaughlin ApJ 618, 2005}
-\begin{equation}
-  v(r) = v_1 \left(1-\frac{R}{r_1}\right)^{-\beta}
-  \left(1-\frac{R}{r}\right)^{\beta},
-\end{equation}
-with $R$ the wind launch radius and $\beta$ an acceleration parameter.
-This yields the following density field
-\begin{equation}
-  \rho(r)=\rho_1 C\left[\frac{r_1}{R}\right]
-  \left(1-\frac{R}{r}\right)^{-\beta}\left(\frac{R}{r}\right)^2,
-\end{equation}
-with $C[a]=a^{2-\beta}(a-1)^\beta$.
-We use the parameters $R=10$~km, $R_1=50$~km, $\beta=2$,
-and $\rho_1=10^7\,{\rm g}\,{\rm cm}^{-3}$.
-Additionally we impose a density cap of
-$\rho_{\rm max}=10^{14}\,{\rm g}\,{\rm cm}^{-3}$ inside the radius $r_0=5$~km,
-and smoothly interpolate densities between $r_0$ and $r_1$ with a cubic polynomial,
-enforcing $C^0$ and $C^1$ continuity at the transitions.
-The rest density from this model is plotted in Fig.~\ref{fig:test_traj_density}.
-To translate this density field to a matter potential, we assume a constant
-electron fraction of $Y_e=0.5$.
+In this section we present $V_{\nu\nu,0}$ along one test trajectory in the NSNS
+model, and show that including elastic scattering makes a qualitative
+difference in the character of the flavor evolution.
+See Figs.~\ref{fig:V_nunu-noscat} and \ref{fig:V_nunu-scat}.
 
 \begin{figure}
-  \includegraphics[width=\columnwidth]{profile-density-000Mo-025deg}
-  % Plot:    $DIR/RunBA/RunBA-025deg-noscat-production01/profile-density-000Mo-025deg.pdf
-  % Script:  $DIR/nu_potentials-production.nb
-  % Configs: $DIR/nu_potentials-wind_model.nb
-  %          $DIR/RunBA/RunBA-025deg-noscat-production01/nu_potentials-config-production.nb
-  % w/ DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20170412-self_interaction_potentials_M12_M12-LS220-M1/Run03/
-  \caption{Assumed matter density along the neutrino test trajectory
-    used to calculate the matter potential $V_e$.
-    }
-  \label{fig:test_traj_density}
-\end{figure}
-
-In fact a spherical steady-state wind model, though providing an adequate
-backdrop for the qualitative study presented in this section,
-is less than ideal for this post-merger configuration,
-most obviously because in the 10~ms since merger,
-ejecta with the greatest velocities around $0.3c$
-will have reached no further than $r_{\rm max} \sim 10^8$~cm.
-Additionally, the true radial profile of the ejecta from this merger
-(which our computational model does not follow)
-will have many more features inside this radius, including shock jumps.
-This model density profile is adequate as a backdrop to our study here, however,
-since we expect the remnant to present similar neutrino emission over a thermal
-timescale of a few tens of milliseconds\footnote{cite Nouri et al 2018}
-while the matter field propagates out to larger radii.
-
-In Figs.~\ref{fig:V_nunu-noscat} and \ref{fig:V_nunu-scat}
-we show the total unoscillated self-interaction potential
-and its contributions from $\nu_e$ and $\bar{\nu}_e$ for a test neutrino
-moving out along the radial coordinate trajectory described above.
-Fig.~\ref{fig:V_nunu-noscat} shows these terms for the \emph{noscat} treatment,
-and Fig.~\ref{fig:V_nunu-scat} for the \emph{scat} treatment.
-Obviously, including the effects of elastic scattering tends to increase the
-$\bar{\nu}_e$ contribution relative to the $\nu_e$ contribution,
-in this case causing the self-interaction potential to be negative
-along the entire trajectory.
-This effect may be predicted from Fig.~\ref{fig:nsns_cosA_distrib_G_00deg},
-which is calculated for a qualitatively similar observer, in the vacated polar
-funnel of the disk:
-in the \emph{scat} treatment $G_{\bar{\nu}_e}$ dominates over $G_{\nu_e}$ at
-all angles $\cos A \lesssim 0.998$;
-whereas in the \emph{noscat} treatment $G_{\bar{\nu}_e}$ only dominates over
-$G_{\nu_e}$ over a range of forward-peaked angles $\cos A \in (0.965,0.998)$.
-An additional factor supporting this trend is that neutrinos from the most
-forward-peaked angles $\cos A \sim 1$ have a suppressed effect on the
-self-interaction potential due to the $(1-\cos A)$ term arising in the integral
-of Eqn.~\ref{eqn:V_nu_alpha}.
-
-\begin{figure}
-  \includegraphics[width=\columnwidth]{20171223-potentials-Ve_Vnue_Vnua_Vnu-000Mo-025deg-noscat}
-  % Plot:   $DIR/RunBA/RunBA-025deg-noscat-production01/potentials-Ve_Vnue_Vnua_Vnu-000Mo-025deg.pdf
-  % Data:   $DIR/000Mo-025deg-*km/Lev2-e30a200b30-noscat-04
-  % Script: $DIR/nu_potentials-production.nb
-  % Config: $DIR/RunBA/RunBA-025deg-noscat-production01/nu_potentials-config-production.nb
-  % w/ DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20170412-self_interaction_potentials_M12_M12-LS220-M1/Run03/
-  \caption{Neutrino oscillation potentials
-    along a test trajectory
+  \includegraphics[width=\columnwidth]{20170619-potentials-Vnue_Vnua_Vnu-000Mo-022deg-noscat}
+  \caption{Neutrino self-interaction potential along a test trajectory
     originating at the surface of the hypermassive neutron star, and
     proceeding outward along a radial coordinate ray with angle
-    $\theta=25^{\circ}$ with respect to the polar axis.
+    $\psi=22^{\circ}$ with respect to the polar axis.
     The trajectory is parameterized by the coordinate length $s$.
-    We plot the vacuum potential due to mass-squared differences (gray bands),
-    the matter potential $V_e$ due to forward scattering on $e^-$ and $e^+$
-    (dark green),
-    and the self-interaction potential $V_{\nu\nu,0}$ due to forward
-    scattering on ambient neutrinos (light green, solid where positive,
-    dashed where negative);
-    additionally we plot the $\nu_e$ and $\bar{\nu}_e$ components composing
-    the self-interaction potential (light blue and dark blue respectively).
     This calculation was done with scattering turned off.
-    The two gray bands at $V\sim10^{-22}$~erg and $V\sim10^{-24}$~erg
-    indicating the two vacuum energy scales for 10--30~MeV neutrinos,
-    set the positions of possible Mikheyev-Smirnov-Wolfenstein (MSW)
-    or nutation resonances \cite{malk2012-mnr_1}.
-    Inside the symmetric point at $s\sim100\,{\rm km}$ the total unoscillated
-    self-interaction potential $V_{\nu\nu,0}$ is positive,
-    i.e.\ $\nu_e$-dominated.
+    Outside of the symmetric point at $s\sim40\,{\rm km}$, the potential is
+    positive, i.e.\ $\nu_e$-dominated.
+    Compare to Fig.~\ref{fig:V_nunu-scat}.
     }
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{20171223-potentials-Ve_Vnue_Vnua_Vnu-000Mo-025deg-scat}
-  % Plot:   $DIR/RunBA/RunBA-025deg-scat-production01/potentials-Ve_Vnue_Vnua_Vnu-000Mo-025deg.pdf
-  % Data:   $DIR/000Mo-025deg-*km/Lev2-e30a200b30-scat-04
-  % Script: $DIR/nu_potentials-production.nb
-  % Config: $DIR/RunBA/RunBA-025deg-scat-production01/nu_potentials-config-production.nb
-  % w/ DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20170412-self_interaction_potentials_M12_M12-LS220-M1/Run03/
+  \includegraphics[width=\columnwidth]{20170619-potentials-Vnue_Vnua_Vnu-000Mo-022deg-scat}
   \caption{Same as Fig.~\ref{fig:V_nunu-noscat},
     but with scattering turned on.
-    Unlike the case with scattering turned off,
-    here the total unoscillated self-interaction potential $V_{\nu\nu,0}$
-    is everywhere negative, i.e.\ $\bar{\nu}_e$-dominated.
+    Unlike the case with scattering turned off, here the potential is
+    everywhere negative, i.e.\ $\bar{\nu}_e$-dominated.
     }
@@
 
-We also solve for the flavor evolution of this system,
-integrating $S$ according to Eqn.~\ref{eqn:schroedinger},
-as described in \cite{zhu2016-mnr_nsns_remnant}.
-In Figs.~\ref{fig:survival-noscat} and \ref{fig:survival-scat}
-we show the survival probabilities for $\nu_e$ and $\bar{\nu}_e$
-for the two treatments.
-\footnote{TODO: update this paragraph to new compute-eta data}
-The survival probabilities are the terms of
-the flavor evolution matrix $S$ that describe how much of the original flux
-of a given flavor survives a transition.
-\footnote{TODO: define explicitly}
-Where $P_{\nu_e\rightarrow\nu_e}$ vanishes, for example in
-Fig.~\ref{fig:survival-noscat}, all the original $\nu_e$ content of the beam
-is oscillated into $\nu_\mu$ and $\nu_\tau$; but the $\nu_e$ density doesn't
-vanish altogether,
-since some $\nu_\mu$ and $\nu_\tau$ oscillates into $\nu_e$.
-
-For our assumed matter potential along this one test trajectory
-we see that a matter-neutrino resonance only occurs in the \emph{scat} case;
-though with a slightly higher matter potential it could occur in both cases,
-closer to the disk in the \emph{noscat} case than the \emph{scat} case.
-For this particular test trajectory we only see a standard and not a symmetric
-matter neutrino resonance;
-these two classes of matter-neutrino resonance
-are described for example in \cite{malk2016-mnr_3,vaan2016-uncovering_mnr}.
-The difference we see in the oscillation physics between the \emph{noscat}
-and \emph{scat} cases along this test trajectory highlights the importance of
-scattering physics in the problem of neutrino oscillations in
-neutron star mergers.
-
-\begin{figure}
-  \includegraphics[width=\columnwidth]{fig-survival_prob-noscat_nh}
-  % Flavor evolution by Yonglin, see emails 7.19.17-7.20.17
-  % Plot: $DIR/31-900kmNHnoScatterPot.eps
-  % w/ DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20170412-self_interaction_potentials_M12_M12-LS220-M1/Run03/RunAB/probabilityPotential
-  \caption{Neutrino survival probabilities for the \emph{noscat} case.
-    \footnote{TODO: Replace; this is from the old interp-eta data}
-    In the lower panel we reproduce the unoscillated self-interaction potential
-    $V_{\nu\nu,0}$ from Fig.~\ref{fig:V_nunu-noscat} and a matter potential
-    computed assuming a very simple steady state wind consistent with the
-    densities within the simulation volume inside of $r<70$~km
-    along this trajectory.
-    In this case the neutrinos undergo a symmetric MNR transition
-    inside of $r<50$~km,
-    with the $e$-flavor neutrinos and antineutrinos fully converting to $\nu_x$.
-  }
-  \label{fig:survival-noscat}
-\end{figure}
-
-\begin{figure}
-  \includegraphics[width=\columnwidth]{fig-survival_prob-scat_nh}
-  % see notes from fig-survival_prob-noscat_nh
-  % Plot: $DIR/130-800kmNHScatterPot.eps (normal hierarchy)
-  \caption{Same as Fig.~\ref{fig:survival-noscat},
-    but with scattering turned on.
-    \footnote{TODO: Replace; this is from the old interp-eta data}
-    In this case the neutrinos undergo a standard MNR transition
-    outside of $r>200$~km,
-    with the $e$-flavor neutrinos and antineutrinos
-    partially converting to $\nu_x$.
-    }
-  \label{fig:survival-scat}
-\end{figure}
-
-\section{Conclusions}
-\label{sec:conclusions}
-Add some conclusions.
-\footnote{TODO: note some improvements needed: implement implicit time-stepping,
-  use finite-difference grid for fluid vars, improve gray spectral assumptions}
-
-\subsubsection*{Acknowledgements}
-This research is supported in part by the Joint Institute for Nuclear
-Astrophysics---Center for the Evolution of the Elements (JINA-CEE)
-and the National Science Fondation under grant numbers X. \footnote{TODO: grant nos?}
-We thank Sherwood Richers for providing his neutrino transport code
-\lstinline{Sedonu} for comparisons,
-Fran\c{c}ois H\'{e}bert for his timely help in finding and
-correcting bugs in the time-interpolation algorithm,
-and Luke Roberts for helpful discussions.
+\subsection{Neutrino Annihilation Integral}
+\label{ssec:q_nunu}
+In this section we calculate $q_{\nu\bar{\nu}}$ at a representative point
+in the funnel of the NSNS configuration and show that including scattering makes
+a difference (or not) in our estimate of the power available to drive a jet.
 
@@
 \appendix
-
-\section{Definitions}
-\label{sec:definitions}
-We decompose the neutrino momentum into components parallel and orthogonal to
-an observer's velocity $u_\beta$:
-\begin{equation}
-  \label{eqn:def_momentum_2}
-  p_\beta = \varepsilon (u_\beta + \ell_\beta),
-\end{equation}
-with $u_\alpha\ell^\alpha=0$ and $\ell_\alpha\ell^\alpha=1$.
-
-We use two possible fiducial observers to define the momentum
-decomposition via Eqn.~\ref{eqn:def_momentum_2}: the Eulerian,
-or normal observer $n_\mu=-\alpha \,\partial_\mu t$,
-and the fluid, or comoving observer
-$u^\mu=Wn^\mu+v_{\rm E}^\mu$.
-Here $t$ is coordinate time and $\alpha$ is lapse
-in the standard 3+1 decomposition of the metric
-\begin{equation}
-  \label{eqn:adm_metric}
-  \psi_{\mu\gamma} \rightarrow
-  \left(
-  \begin{matrix}
-    -\alpha^2 + \beta^i \beta_i  & \beta_i \\
-    \beta_j                      & g_{ij}
-  \end{matrix}
-  \right).
-\end{equation}
-We have also introduced
-the fluid Lorentz factor $W=-n_\mu u^\mu=\alpha u^t$,
-its Eulerian velocity $v_{\rm E}^\mu=g^\mu_\lambda u^\lambda$
-(distinct from its coordinate velocity $v^\mu=u^\mu/u^t$),
-and the projection tensor orthogonal to the normal observer
-$g^\mu_\lambda=\psi^\mu_\lambda+n^\mu n_\lambda$,
-whose spatial part is the 3-metric, $g_{ij}$.
-
-We specify the neutrino direction in the observer's frame
-with two spherical polar angles
-($a$,$b$) with respect to the simulation cartesian coordinates
-\begin{align}
-  \ell_\alpha &\rightarrow q(s,\Omega_i), \\
-  \Omega_i &\rightarrow (\sin a \cos b,\sin a\sin b,\cos a),
-\end{align}
-or alternatively the two spherical polar angles
-($A$,$B$) with respect to rotated coordinates
-\begin{equation}
-  \label{eqn:def_direction_primed_2}
-  \Omega_{i'} \rightarrow  (\sin A \cos B,\sin A\sin B,\cos A).
-\end{equation}
-The two coordinate systems are related by a standard Euler rotation of first
-$\phi$ about the $z$-axis, then $\theta$ about the rotated $y$-axis,
-with $\phi$ and $\theta$ the azimuthal and polar position of the observer.
-Expressed algebraically:
-\begin{align}
-  \label{eqn:rotation_connection}
-  \Omega_x &= \Omega_{x'}\cos\theta\cos\phi-\Omega_{y'}\sin\phi+\Omega_{z'}\sin\theta\cos\phi,\nonumber\\
-  \Omega_y &= \Omega_{x'}\cos\theta\sin\phi+\Omega_{y'}\cos\phi+\Omega_{z'}\sin\theta\sin\phi,\nonumber\\
-  \Omega_z &= \Omega_{x'}\sin\theta+\Omega_{z'}\cos\theta.
-\end{align}
-
-The scale factors $q$ and $s$ are functions of the neutrino
-direction $\Omega_i$, the observer's velocity $u_\alpha$,
-and the spacetime metric.
-In the case of a fluid observer, specified by an arbitrary $W$ and $u_i$,
-$q$ and $s$ are given by
-\begin{align}
-  \label{eqn:q_general}
-  q
-  &= W\alpha\Bigg(2\beta^i\Omega_i W^2(\beta^i\Omega_i-1) \nonumber\\
-  &\qquad\qquad - 2\Omega_i u_j g^{ij} W \alpha(\beta^i\Omega_i-1)\nonumber\\
-  &\qquad\qquad + \alpha^2\left((\Omega_i u_j g^{ij})^2+\Omega_i\Omega_j g^{ij}W^2\right)\Bigg)^{-1/2}\\
-  \label{eqn:s_general}
-  s
-  &= \beta^i\Omega_i -\Omega_i u_j g^{ij}.
-\end{align}
-In the case of an Eulerian observer, $W=1$ and $u_i=0$,
-and these expressions simplify considerably:
-\begin{align}
-  \label{eqn:q_eulerian}
-  \tilde{q}
-  &= \alpha\left(2\beta^i\Omega_i(\beta^i\Omega_i-1)+\alpha^2\Omega_i\Omega_j g^{ij}\right)^{-1/2}\\
-  \label{eqn:s_eulerian}
-  \tilde{s}
-  &= \beta^i\Omega_i.
-\end{align}
-In the even simpler case of Minkowski spacetime, these expressions
-reduce to $q=W/(1+s)$, $s=-\Omega_i u^i$,
-$\tilde{q}=1$, $\tilde{s}=0$.
-
-In addition to the fluid velocity and Lorentz factor, described above, the
-other fluid state variables we use from our hydrodynamic simulations
-are rest density $\varrho=m_b n_b$, temperature $T$, and electron fraction
-\begin{equation}
-  Y_e = \frac{n_{e^-} - n_{e^+}}{n_b},
-\end{equation}
-where $n_{e^-}$, $n_{e^+}$, and $n_b$ are the number densities of electrons
-positrons, and baryons, and $m_b$ is the average baryon mass.
 
@@
 inelastic scattering, and the thermal pair processes of annihilation and
-production. In this work, however, we only treat absorption/emission
-and elastic scattering.
-We seek to write each collision source linear in $f$:
-\begin{align}
-  \label{eqn:sources_cae}
-  C_{\rm AE}
-  &= \mathscr{E}_{\rm AE} - \mathscr{K}_{\rm AE}\, f, \\
-  \label{eqn:sources_cse}
-  C_{\rm SE}
-  &= \mathscr{E}_{\rm SE} - \mathscr{K}_{\rm SE}\, f.
-\end{align}
-
+production.
+We present formulae for all of these processes, though in this work, we only
+treat absorption/emission and elastic scattering.
 Each term
@@
 relevant category given in Tab.~\ref{tab:neutrino_processes}.
-We compute these rates in the rest frame of the fluid,
-but because they are spacetime invariants
+These rates are computed most easily as traditional emissivities and opacities
+in the fluid frame of reference;
+but since we use them in the form of an invariant emissivity and opacity
+(defined e.g. in \ref{eqn:inv_emissivity} and \ref{eqn:inv_opacity},
+and obeying Eqn.~\ref{eqn:boltzmann_linear})
 they take the same numerical value in any frame of reference
-and are completely independent of our choice of fiducial observer $u^\alpha$
-(see discussion around Eqn.~\ref{eqn:boltzmann_linear}).
+and are completely independent of our choice of fiducial observer $u^\alpha$.
+\todo{confirm}
 
@@
 By contrast $C^{\nu_x}_{{\rm AE}} = 0$,
-\footnote{TODO: but we use a nonzero $C^{\nu_x}_{{\rm AE}}$}
 where $\nu_x$ stands for any of the four heavy-lepton neutrinos and
@@
 interpolate quad-linearly to the points sampled along each ray.
+\todo{derive and type up all sources in terms of $J$ and $H^\alpha$
+  following \cite[Sec.~4]{shib2011-truncated_moment}}
 
@@
 \label{ssec:sources_ae}
-At neutrino and thermal energies well below the masses of the muon or tauon
-($m_{\mu}\sim100\,{\rm MeV}$)
+At neutrino and thermal energies well below $m_{\mu}\sim100\,{\rm MeV}$
 only charged current processes involving $\nu_e$ and $\bar{\nu}_e$ are allowed.
@@
 (e.g. \cite[Eqn.~A5]{brue1985-core_collapse}):
-\footnote{TODO: explain Fermi-blocking}
+\todo{explain Fermi-blocking}
 \begin{align}
@@
   C_{{\rm AE},i}
-  &= \varepsilon j_i (1-f) - \varepsilon \chi_{a,i} f \\
+  &= \mathscr{E}_{{\rm AE},i} (1-f) - \mathscr{K}_{{\rm AE},i} f \\
   \label{eqn:sources_ae_2}
-  &= \varepsilon j_i - (\varepsilon j_i+\varepsilon \chi_{a,i}) f,
+  &= \mathscr{E}_{{\rm AE},i} - (\mathscr{E}_{{\rm AE},i} + \mathscr{K}_{{\rm AE},i}) f,
 \end{align}
-where $j$ is the emissivity and $\chi_a$ the absorption opacity.
-Both $j$ and $\chi_a$ have dimension ${\rm length}^{-1}$ and represent the
-number of neutrinos emitted or absorbed per length traveled.
-In radiation transport formulations using specific intensities instead of
-distribution functions, an emissivity $\eta$ having dimension
-${\rm energy}\,{\rm length}^{-3}\,{\rm time}^{-1}\,{\rm energy}^{-1}\,{\rm steradian}^{-1}$
-is more commonly used.
-The two are related by $j=(2\pi)^3\eta/\varepsilon^3$.
-Note that for brevity we have suppressed the energy-dependence of
-the terms $j$, $\chi_a$, and the distribution functions.
-
-In the special case of radiative equilibrium we know that the source
-term vanishes: an equal number of neutrinos are emitted from
-and absorbed by the matter for any length traversed.
-We also know in this case that the neutrino distribution function must be
-$f^{\rm eq}$, the equilibrium Fermi-Dirac distribution function of
-Eqn.~\ref{eqn:feq}.
-With these facts we can rerrange Eqn.~\ref{eqn:sources_ae_1} to give us
-Kirchoff's Law:
-\begin{align}
-  \label{eqn:kirchoffs_law}
-  j_i &= \frac{\chi_{a,i}}{1-f^{\rm eq}}f^{\rm eq}, \\
-  \label{eqn:stimulated_opacity}
-  &= \chi^*_{a,i}\,f^{\rm eq},
-\end{align}
-where in Eqn.~\ref{eqn:stimulated_opacity} we have introduced the opacity
-corrected for stimulated absorption, $\chi^*_{a,i}$.
-
-Using these expressions and computing the sum over stimulated opacities
-$\chi^*_{a,i}=\sum_i \chi^*_{a,i}$,
-the invariant emissivity and opacity for absorption/emission are
-\begin{align}
-  \label{eqn:ae_emissivity_summed_2}
-  \mathscr{E}_{{\rm AE}} &= \varepsilon \, \chi^*_a \, f^{\rm eq}, \\
-  \label{eqn:ae_opacity_summed_2}
-  \mathscr{K}_{{\rm AE}} &= \varepsilon \, \chi^*_a.
-\end{align}
-These expressions are equivalent to those in
-Eqns.~\ref{eqn:ae_emissivity_summed} and \ref{eqn:ae_opacity_summed}.
-
-We use the above treatment for $\nu_e$ and $\bar{\nu}_e$ only;
-the $\mu$ and $\tau$ neutrinos and antineutrinos do not participate in
-charged current absorption/emission interactions at these temperatures
-and energies.
-However we do use an effective stimulated absorption opacity $\chi^*_a$
-for the heavy-lepton neutrinos, computed by \lstinline{NuLib} as
-described in \cite{ocon2015-gr1d_with_nu} which follows
-\cite{brue1985-core_collapse,burr2006-neutrino_opacities}.
-This is described in App.~\ref{ssec:sources_pp}.
+with
+\begin{equation}
+  \label{eqn:inv_emissivity}
+  \mathscr{E}_{{\rm AE},i}
+  = \varepsilon j_i
+  = \frac{(2\pi)^3}{\varepsilon^2}\eta_i,
+\end{equation}
+where $j$ is the emissivity having dimension
+$[j]={\rm length}^{-1}\,{\rm steradian}^{-1}$,
+$\eta$ is the emissivity having dimension
+$[\eta]={\rm energy}\,{\rm length}^{-3}\,{\rm time}^{-1}\,{\rm energy}^{-1}\,{\rm steradian}^{-1}$
+(natural to formulations using specific intensity instead of distribution
+functions), and
+\begin{equation}
+  \label{eqn:inv_opacity}
+  \mathscr{K}_{{\rm AE},i} = \varepsilon \chi_i,
+\end{equation}
+where $\chi$ is the traditional opacity with dimension
+$[\chi]={\rm length}^{-1}$.
 
@@
 Neutrino scattering on particles of mass much greater than $\varepsilon$
-(i.e. nucleons) is essentially iso-energetic. Following 
+is essentially iso-energetic. Following 
 \cite[Eqn.~A8]{brue1985-core_collapse} or
@@
 \begin{equation}
-  \label{eqn:sources_se_1}
-  C_{{\rm SE},i}(\ell_\alpha)
+  \label{eqn:integral_se}
+  C_{{\rm SE},i}
   = \frac{\varepsilon^3}{(2\pi)^3}
   \oint d\Omega' R_{{\rm SE},i}(\omega')
-  \Big(f(\ell'_\beta) - f(\ell_\alpha)\Big)
-\end{equation}
-where $R_{\rm SE,i}(\omega')$ is the scattering kernel for the $i$-th process
-from direction $\ell'_\beta$ to direction $\ell_\alpha$
-having dimension ${\rm energy}^{-1}$,
+  \big(f(\ell'_\beta) - f(\ell_\alpha)\big)
+\end{equation}
+where $R(\omega')$ is the scattering kernel from direction $\ell'_\beta$
+to direction $\ell_\alpha$ having dimension $[R]={\rm energy}^{-1}$,
 and the cosine of the scattering angle is
-$\omega' \equiv \psi^{\alpha\beta}\ell_\alpha\ell'_\beta$, with
-$\psi^{\alpha\beta}$ the inverse of the spacetime metric.
-Note that for brevity in Eqn.~\ref{eqn:sources_se_1} we have suppressed the
-energy dependence of all of the terms.
+$\omega' \equiv \ell'_\alpha \ell^\alpha$.
+We write this in the form
+\begin{equation}
+  \label{eqn:sources_se}
+  C_{{\rm SE},i}
+  = \mathscr{K}_{{\rm SE},i} \Phi_{{\rm SE},i} - \mathscr{K}_{{\rm SE},i} f,
+\end{equation}
+with
+\begin{align}
+  \label{eqn:se_factor}
+  \Phi_{{\rm SE},i} &\equiv
+  \frac{1}{\mathscr{K}_{{\rm SE},i}}
+  \frac{\varepsilon^3}{(2\pi)^3}
+  \oint d\Omega' R_{{\rm SE},i}(\omega')f(\ell'_\beta) \\
+  \label{eqn:se_opacity}
+  \mathscr{K}_{{\rm SE},i} &\equiv
+  \frac{\varepsilon^3}{(2\pi)^3}
+  \oint d\Omega' R_{{\rm SE},i}(\omega').
+\end{align}
 
@@
   R_{{\rm SE},i}(\varepsilon,\omega) \approx
-  R_{{\rm SE},i}^0(\varepsilon) + \omega \, R_{{\rm SE},i}^1(\varepsilon).
-\end{equation}
-Using this expansion, and the moments defined in Eqns.~\ref{eqn:J}
-and \ref{eqn:Ha} we may write Eqn.~\ref{eqn:sources_se_1}:
-\begin{equation}
-  \label{eqn:sources_se_2}
-  C_{{\rm SE},i}
-  = \varepsilon \chi_{s,i} \Phi_i - \varepsilon \chi_{s,i} f,
-\end{equation}
-where we have introduced the scattering opacity $\chi_{s,i}$
-and the background scattering field $\Phi$:
+  R_{{\rm SE},i}^0(\varepsilon) + \omega \, R_{{\rm SE},i}^1(\varepsilon),
+\end{equation}
+which allows us to write Eqns.~\ref{eqn:se_factor} and \ref{eqn:se_opacity} as
+\todo{drop anisotropic term and defend}
 \begin{align}
-  \label{eqn:opacity_s}
-  \chi_{s,i}(\varepsilon)
-  &= \frac{4\pi\varepsilon^3}{(2\pi)^3}R_{{\rm SE},i}^0(\varepsilon), \\
-  \label{eqn:background_phi}
-  \Phi_i(\varepsilon,\ell_\alpha)
-  &= \frac{(2\pi)^3}{\varepsilon^3}\frac{1}{4\pi}
-  \left(J(\varepsilon)+
-  \frac{\chi^{\rm tr}_{s,i}(\varepsilon)}{\chi_{s,i}(\varepsilon)}
-  \ell_\alpha H^\alpha(\varepsilon)\right),
+  \label{eqn:se_factor_expansion}
+  \Phi_{{\rm SE},i}(\varepsilon) &= \frac{1}{\mathscr{K}_{{\rm SE},i}}
+  \big(R_{{\rm SE},i}^0(\varepsilon) J(\varepsilon) +
+   R_{{\rm SE},i}^1(\varepsilon) \ell_\alpha H^\alpha(\varepsilon) \big) \\
+  \label{eqn:se_opacity_expansion}
+  \mathscr{K}_{{\rm SE},i}(\varepsilon) &=
+  4\pi \frac{\varepsilon^3}{(2\pi)^3} R_{{\rm SE},i}^0(\varepsilon),
 \end{align}
-with $\chi^{\rm tr}_{s,i}$ the transport scattering opacity,
-\footnote{TODO: drop anisotropic term and defend}
-\begin{equation}
-  \label{eqn:opacity_s_tr}
-  \chi^{\rm tr}_{s,i}(\varepsilon)
-  = \frac{4\pi\varepsilon^3}{(2\pi)^3}R_{{\rm SE},i}^1(\varepsilon).
-\end{equation}
-In our treatment we only retain the isotropic contribution to the
-background field:
-\begin{equation}
-  \Phi(\varepsilon)
-  = \frac{(2\pi)^3}{\varepsilon^3}\frac{1}{4\pi}
-  J(\varepsilon).
-\end{equation}
-\footnote{TODO: defend}
-
-Thus computing sums over the opacities
-$\chi_s\equiv\sum_i\chi_{s,i}$ and
-$\chi^{\rm tr}_s\equiv\sum_i\chi^{\rm tr}_{s,i}$,
-the invariant emissivity and opacity for elastic scattering are
-\begin{align}
-  \label{eqn:se_emissivity_summed}
-  \mathscr{E}_{\rm SE}
-  &= \varepsilon \, \chi_s \, \Phi, \\
-  \label{eqn:se_opacity_summed}
-  \mathscr{K}_{\rm SE}
-  &= \varepsilon \, \chi_s.
-\end{align}
-
-We have made the energy-dependence of the background scattering field
-explicit in Eqn.~\ref{eqn:background_phi}.
-When we compute $\Phi(\varepsilon)$ using $J(\varepsilon)$ and
-$H^\alpha(\varepsilon)$ from a multi-group
-M1 transport evolution, we call this the \emph{spectral} method.
+where we have made the energy-dependence explicit.
+When we get $J(\varepsilon)$ and $H^\alpha(\varepsilon)$ from a multi-group
+moment evolution, we call this the ``spectral'' method.
 
@@
 and sometimes the number density $G$.
-If such is the case we resort to the \emph{gray} method,
-by reconstructing the energy-dependent source terms
-from gray moments, assuming a diluted Fermi-Dirac spectrum.
+We may reconstruct the energy-dependent source terms
+from the gray moments by assuming a diluted Fermi-Dirac spectrum.
+
 We use the following procedure:
@@
   Interpolate the fluid temperature, $T$,
-  equilibrium neutrino chemical potential, $\eta_\nu$,
+  equilibrium neutrino chemical potential, $\mu_\nu$,
   and evolved neutrino energy and number densities in the fluid frame,
@@
   \begin{equation}
-    \label{eqn:spectral_temp}
     T_\nu = \langle \varepsilon \rangle
-    \frac{\mathscr{F}_2(\eta_\nu)}{\mathscr{F}_3(\eta_\nu)},
+    \frac{\mathscr{F}_2(\mu_\nu/T)}{\mathscr{F}_3(\mu_\nu/T)},
   \end{equation}
@@
   \begin{equation}
-    \label{eqn:fermi_integral}
     \mathscr{F}_b(\eta) = \int_0^\infty dx \, x^b (1+e^{x-\eta})^{-1}.
   \end{equation}
-  We implement the Fermi integrals using the analytical approximants from
-  \cite{taka1978-beta_rates}.
 \item
@@
     J(\varepsilon) &=
-    J \frac{\varepsilon^3}{T_\nu^4 \,\mathscr{F}_3(\eta_\nu)}
-    (1+\exp(\varepsilon/T_\nu-\eta_\nu)^{-1}, \\
+    J \frac{\varepsilon^3}{T_\nu^4 \,\mathscr{F}_3(\mu_\nu/T)}
+    (1+\exp(\varepsilon/T_\nu-\mu_\nu/T)^{-1}, \\
     \label{eqn:H_from_gray}
     H^\mu(\varepsilon) &=
-    H^\mu \frac{\varepsilon^3}{T_\nu^4\, \mathscr{F}_3(\eta_\nu)}
-    (1+\exp(\varepsilon/T_\nu-\eta_\nu)^{-1}.
+    H^\mu \frac{\varepsilon^3}{T_\nu^4\, \mathscr{F}_3(\mu_\nu/T)}
+    (1+\exp(\varepsilon/T_\nu-\mu_\nu/T)^{-1}.
   \end{align}
 \end{enumerate}
-
-Note that to avoid floating point errors for very large negative $\eta$,
-we employ the asymptotic form of Eqn.~\ref{eqn:fermi_integral}:
-$\lim_{\eta\ll-1}\mathscr{F}_b(\eta) = b!\,e^\eta$.
-Thus, for $\eta<-10$, we use the limiting forms of
-Eqns.~\ref{eqn:spectral_temp},
-\ref{eqn:J_from_gray},
-and \ref{eqn:H_from_gray}:
-\begin{align}
-    T_\nu &= 
-    \frac{\langle \varepsilon \rangle}{3}, \\
-    J(\varepsilon) &=
-    J \frac{\varepsilon^3}{6 T_\nu^4} e^{-\varepsilon/T_\nu}, \\
-    H^\mu(\varepsilon) &=
-    H^\mu \frac{\varepsilon^3}{6 T_\nu^4} e^{-\varepsilon/T_\nu}.
-\end{align}
-
-When we use the moments from Eqns.~\ref{eqn:J_from_gray} and
-\ref{eqn:H_from_gray} in Eqn.~\ref{eqn:background_phi},
-we call this the \emph{gray} method.
+When we use these moments in Eqn.~\ref{eqn:se_factor_expansion},
+we call this the ``gray'' method.
 
@@
 \label{ssec:sources_si}
-Neutrino scattering off of electrons is inelastic, changing the
-magnitude and direction of the species' momenta.
-\footnote{TODO: cite \cite{keil2003-pinched_spectra} or someone}
-A similar formalism to the above could be used to derive source terms
-for inelastic scattering.
-\footnote{TODO: compare relative amounts of inelastic to elastic scattering}
-We save that for future work, pointing out here that inelastic scattering
-treated this way is very sensitive to the energy-dependence of
-the background field.
+(e.g. \cite[Eqn.~A6]{brue1985-core_collapse}).
 
@@
 \label{ssec:sources_pp}
-We do not include thermal pair processes within the above formalism,
-\footnote{TODO: explain this choice, and cite \cite{keil2003-pinched_spectra} or someone}
-but as mentioned in App.~\ref{ssec:sources_ae}, we do include
-these processes in an effective absorption opacity for $\mu$ and $\tau$
-neutrinos and antineutrinos. \lstinline{NuLib} follows
-\cite{brue1985-core_collapse,burr2006-neutrino_opacities}
-by computing the emissivity of the pair processes in
-\footnote{TODO: describe better}
-Tab.~\ref{tab:neutrino_processes}, and applying Kirchoff's law,
-Eqn.~\ref{eqn:kirchoffs_law}, to convert this to an opacity.
-We only use this effective opacity for heavy-lepton neutrinos because
-it is subdominant to the true absorption opacity for
-$\nu_e$ and $\bar{\nu}_e$.
+(e.g. \cite[Eqn.~A9]{brue1985-core_collapse}).
+
+\subsubsection*{Acknowledgements}
+This research is supported in part by the Joint Institute for Nuclear
+Astrophysics---Center for the Evolution of the Elements (JINA-CEE)
+and the National Science Fondation under grant numbers X. \todo{grant nos?}
+We thank Sherwood Richers for providing his neutrino transport code
+\lstinline{Sedonu}~\footnote{\url{https://bitbucket.org/srichers/sedonu}}
+for comparisons.
+We thank Luke Roberts and Yonglin Zhu for helpful discussions.
+We are indebted to Fran\c{c}ois H\'{e}bert for his timely help in finding and
+correcting bugs in the time-interpolation algorithm.
 

### 2001-01-01 paper.tex
Add table of L, R, <eps> for RT vs M1

@@
 Agreement better than 10\%.
-\todo{put in table comparison M1 and RT}
+\todo{get accurage M1 numbers from Francois}
+
+\begin{table}%[h]
+  \caption{
+    Comparison of total luminosities and average energies between
+    ray tracing and the M1 transport simulation.
+    The ray tracing totals were computed from sums over
+    observers placed in the $y$-$z$ plane along an arc at coordinate distance
+    $r=250\,{\rm km}$ from the center of the star, at 30 positions
+    distributed evenly in $\cos\theta$ over the colatitude range
+    $\theta\in[0,90^{\circ}]$.
+    M1 values are taken from \cite[Figs. 7, 9, 10]{fouc2016-m1_evolve_n}.
+  }
+  \label{tab:nsns_rt_vs_m1}
+  \begin{tabularx}{\columnwidth}{X r r r}
+    & \,\,{\bf RT noscat} & \,\,{\bf RT scat} & \,\,{\bf M1} \\
+    \hline
+    $L_{\nu_e}\,[10^{52}{\rm erg}\,{\rm s}^{-1}]$        & 5.95 & 6.01 & 5 \\
+    $L_{\bar{\nu}_e}$                                    & 10.6 & 11.8 & 12 \\
+    $L_{\nu_x}$                                          & 30.2 & 12.6 & 12 \\
+    \hline
+    $R_{\nu_e}-R_{\bar{\nu}_e}\,[10^{57}{\rm s}^{-1}]$   & -1.06 & -1.86 & -2 \\
+    \hline
+    $\langle \varepsilon_{\nu_e} \rangle\,[{\rm MeV}]$   & 12.08 & 11.88 & 12 \\
+    $\langle \varepsilon_{\bar{\nu}_e} \rangle$          & 15.96 & 14.69 & 15 \\
+    $\langle \varepsilon_{\nu_x} \rangle$                & 36.25 & 23.67 & 26 \\
+    \hline
+  \end{tabularx}
+\end{table}
 

### 2001-01-01 paper.tex
Add theta distributions for nsns model

@@
 compare.
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{theta_distrib-250km-luminosity_R}
+  \caption{Radial number fluxes, $K_r$, as functions of colatitude $\theta$.
+    Sampled for observers at fixed coordinate distances from the center
+    of the star, with $r=250\,{\rm km}$.
+    Here radial fluxes are multiplied by the portion of solid angle
+    represented by each observer, so that the total luminosity is
+    computed by summing all the points.}
+  \label{fig:nsns_theta_distrib_R}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{theta_distrib-250km-luminosity_L}
+  \caption{Same as Fig.~\ref{fig:nsns_theta_distrib_R} but radial energy
+    fluxes, $H_r$.}
+  \label{fig:nsns_theta_distrib_L}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{theta_distrib-250km-avg_eps}
+  \caption{Same as Fig.~\ref{fig:nsns_theta_distrib_R} but
+    average energies $\langle \varepsilon \rangle = J/G$.}
+  \label{fig:nsns_theta_distrib_avg_eps}
+\end{figure}
 

### 2001-01-01 paper.tex
Fix typo
Add description of HMNS configuration

@@
 \label{ssec:test_disk_comparison}
-Total and relative number and energy luminosities for the disk evolved
-in \cite{fouc2016-m1_evolve_n}.
-Agreement better than 10\%.
+The merger of two neutron stars by gravitational wave emission produces a
+postmerger configuration composed of a single neutron star surrounded by a disk.
+Because of the strong differential rotation of the remnant it temporarily avoids
+collapse to a black hole, even if its mass exceeds the threshold of dynamical
+instability for a rigidly rotating neutron star \cite{duez2009-review}.
+These objects are called hypermassive neutron stars.
+
+Such a configuration was modeled by \cite{fouc2016-m1_nsns} by
+evolving the final inspiral and merger of two identical neutron stars of
+isolated gravitational mass 1.2~$M_{\odot}$. The postmerger configuration
+was evolved using an M1 transport scheme for the neutrinos, evolving the
+energy density, number density, and energy flux \cite{fouc2016-m1_evolve_n}
+for $\sim10\,{\rm ms}$ following merger, defined as the time of the peak of the
+gravitational wave strain.
+
+We used a single time snapshot of the volume data from that configuration
+at $t=11\,{\rm ms}$ after merger (see
+Figs.~\ref{fig:nsns_rho_merid}--\ref{fig:nsns_temp_equat})
+sampling the neutrino fields with ray tracing to compare to some
+measures of the field computed from the M1 transport simulation.
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{production-colormap-merid-rho}
+  \caption{A meridional slice of density in the hypermassive neutron
+    star and disk configuration.
+    The distorted rectangular boundaries are the boundaries of the grid
+    used in the numerical simulation, which employs a coordinate mapping to
+    concentrate points near the central object.}
+  \label{fig:nsns_rho_merid}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{production-colormap-equat-rho}
+  \caption{An equatorial slice of density in the hypermassive neutron
+    star and disk configuration.}
+  \label{fig:nsns_rho_equat}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{production-colormap-merid-temp}
+  \caption{A meridional slice of temperature in the hypermassive neutron
+    star and disk configuration.}
+  \label{fig:nsns_temp_merid}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{production-colormap-equat-temp}
+  \caption{An equatorial slice of temperature in the hypermassive neutron
+    star and disk configuration.}
+  \label{fig:nsns_temp_equat}
+\end{figure}
+
+In Tab.~\ref{tab:nsns_rt_vs_m1} we present total and relative number and
+energy luminosities for this configuration.
+With scattering turned off, ray tracing errors are as large as 250\%
+(for heavy lepton energy luminosities).
+But with scattering turned on, agreement is better than 20\%, and
+in most cases much better.
 \todo{get accurage M1 numbers from Francois}
@@
 \end{table}
-
-\begin{figure}
-  \includegraphics[width=\columnwidth]{production-colormap-merid-rho}
-  \caption{A meridional slice of density in the hypermassive neutron
-    star and disk configuration.}
-  \label{fig:nsns_rho_merid}
-\end{figure}
-
-\begin{figure}
-  \includegraphics[width=\columnwidth]{production-colormap-equat-rho}
-  \caption{An equatorial slice of density in the hypermassive neutron
-    star and disk configuration.}
-  \label{fig:nsns_rho_equat}
-\end{figure}
-
-\begin{figure}
-  \includegraphics[width=\columnwidth]{production-colormap-merid-temp}
-  \caption{A meridional slice of temperature in the hypermassive neutron
-    star and disk configuration.}
-  \label{fig:nsns_temp_merid}
-\end{figure}
-
-\begin{figure}
-  \includegraphics[width=\columnwidth]{production-colormap-equat-temp}
-  \caption{A meridional slice of temperature in the hypermassive neutron
-    star and disk configuration.}
-  \label{fig:nsns_temp_equat}
-\end{figure}
 

### 2001-01-01 paper.tex
Add nsns cosA distributions of intensity

@@
   \label{fig:nsns_theta_distrib_avg_eps}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{cosA_distrib-intensity-250km-00deg}
+  \caption{Angular distributions of fluxes for an observer on the rotation axis
+    in the hypermassive neutron star configuration.}
+  \label{fig:nsns_cosA_distrib_I_00deg}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{cosA_distrib-intensity-250km-90deg}
+  \caption{Angular distributions of fluxes for an observer in the equatorial
+    plane in the hypermassive neutron star configuration.}
+  \label{fig:nsns_cosA_distrib_I_90deg}
 \end{figure}

### 2015-11-10 paper.tex
Add paper boilerplate

--- 2001-01-01 paper.tex
@@
-\documentclass[aps,floatfix,prd,superscriptaddress,twocolumn]{revtex4-1}
-% Intended to be built with pdflatex, using the REVTEX 4.1 package.
-% Template taken from FermiLab APS Journal template.
+\documentclass[aps,prd,twocolumn,superscriptaddress,groupedaddress]{revtex4}
+% Intended to be built with pdflatex, using REVTEX 4
+% Template taken from FermiLab APS Jouranl template.
+% Labeling convention example: \label{ssec:acknowledge}.
+%   Specifically, use (<=4)-letter class (e.g. chap, sec, ssec, fig, etc.),
+%   followed by a colon, followed by a shortish underscore separated name
+%   (e.g. intro, nu_energy_spectrum).
 
@@
 \usepackage{amsmath}              % for \text, align, split ...
-\usepackage{color}                % for colorful margin notes
-\usepackage[mathscr]{euscript}    % for \mathscr
-\usepackage{float}                % for better control of figure placement
 \usepackage{graphicx}             % for \includegraphics
 \usepackage{hyperref}             % for \url and \href
-\usepackage{listings}             % for typesetting code with \lstinline
-\usepackage{tabularx}             % for tabularx
 
@@
 \hyphenation{PYTHIA} % ''
-\lstset{basicstyle=\ttfamily} % define \lstinline for inline code snippet
-
-\newcommand{\todo}[1]{\marginpar{\tiny{\textcolor{red}{#1}}}}
-%\renewcommand\todo[1]{} % if you wish to hide all todo notes
-
-% force all figures to be placed right where they are in the text
-\floatplacement{figure}{H}
 
@@
 
+% The following information is for internal review, please remove them for submission
 \widetext
@@
 
-\title{Elastic Scattering in General Relativistic Ray Tracing for Neutrinos}
+\title{Neutrino Annihilation Around Massive Nuclear Accretion Disks}
 
 \author{M.\ Brett Deaton}
-\email{mbdeaton@ncsu.edu}
+\affiliation{Department of Physics \& Astronomy,
+  Washington State University, Pullman, WA 99164, USA}
 \affiliation{Joint Institute for Nuclear Astrophysics,
@@
   North Carolina State University, Raleigh, NC 27695, USA}
+\email{mbdeaton@ncsu.edu}
+
+\author{Matthew D.\ Duez}
+\affiliation{Department of Physics \& Astronomy,
+  Washington State University, Pullman, WA 99164, USA}
 
@@
 
+\author{Francois Foucart}
+\affiliation{Lawrence Berkeley National Laboratory,
+  1 Cyclotron Rd., Berkeley, CA 94720, USA}
+
 \author{Andy Bohn}
@@
 
-\author{Jerred Jesse}
-\affiliation{Department of Physics \& Astronomy,
-  Washington State University, Pullman, WA 99164, USA}
-
-\author{Francois Foucart}
-\affiliation{Lawrence Berkeley National Laboratory,
-  1 Cyclotron Rd., Berkeley, CA 94720, USA}
-
-\author{Matthew D.\ Duez}
-\affiliation{Department of Physics \& Astronomy,
-  Washington State University, Pullman, WA 99164, USA}
-
-\author{G.\ C.\ McLaughlin}
-\affiliation{Department of Physics,
-  North Carolina State University, Raleigh, NC 27695, USA}
-
-% *******************************************************************************
-
 \begin{abstract}
-  We present a covariant ray tracing algorithm for computing high-resolution
-  neutrino distributions in general numerical spacetimes with hydrodynamical
-  sources.
-  Our formulation treats the very important effect of
-  elastic scattering of neutrinos off of nuclei and nucleons
-  (changing the neutrino's direction but not energy)
-  by incorporating estimates of the background neutrino fields.
-  These background fields may be taken from a low-order moment simulation
-  or be ignored entirely, in which case the method
-  reduces to a standard state-of-the-art ray tracing formulation.
-  It handles radiation in regimes spanning optically thick to optically thin.
-  We test the new code, highlight its strengths and weaknesses, and
-  apply it to simulations of neutron star mergers to
-  compute neutrino fluxes as a function of energy and emission angle.
+  This is our abstract.
 \end{abstract}
@@
 
-\section{Introduction}
-Neutrinos are one of the dominant energy transport phenomena at play in
-neutron star mergers, heating, cooling, and pushing the disrupted nuclear
-matter.
-In addition, they change the composition of the matter via charged current
-interactions.
-Because neutrinos scatter over length scales both large and small with
-respect to fluid scales, accurate models require a neutrino treatment that
-respects the freedom of neutrino distribution functions to vary
-drastically from simple geometric distributions in thermodynamic equilibrium.
+This is some text in the body.
+The high disk mass model is described in \cite{deat2013-leakage}.
 
-This is a challenging task in the environment of a merger,
-which generally lacks any spatial symmetries so that fully general
-seven-dimensional solutions to the Boltzmann Equation are not feasible.
-Leakage approximations
-\citep{deat2013-leakage, pere2016-asl, radi2016-dynamical}
-\todo{cite others}
-capture the qualitative effects of neutrinos on the matter, but
-provide extremely limited information about the neutrino field itself.
-\todo{discuss mc more clearly}
-Monte Carlo methods \citep{abdi2012-monte_carlo, rich2015-monte_carlo}
-are an excellent tool require large computational resources because of
-their global treatment of the problem.
-
-The state of the art today couples radiation and matter with a truncated
-moment formalism
-\citep{shib2011-truncated_moment, fouc2015-m1_nsbh, fouc2016-m1_nsns,
-  just2015-m1_code},
-\todo{cite others}
-evolving the zeroth- and first-angular moments of the distribution function
-(the energy density and flux respectively) representing the zeroth-energy moment
-of the radiation field (the total energy).
-This method, commonly called M1 transport, was formulated by
-\cite{thor1981-truncated_moment, shib2011-truncated_moment}
-and has been applied for example in \cite{ramp2002-truncated_moment,
-  card2013-truncated_moment, fouc2015-m1_nsbh, ocon2015-gr1d_with_nu}.
-Recently, \cite{fouc2016-m1_evolve_n} have expanded their M1 transport code
-to also evolve the number densities of neutrinos,
-providing total and average energies throughout the simulated volume.
-Even so, with M1 transport codes only evolving two angular moments
-and one or two energy moments,
-we can extract only limited angular and spectral information
-from the neutrino fields.
-
-%\subsubsection*{What physics problems do we want to address?}
-But many interesting unsolved problems require an accurate model of the neutrino
-spectra and angular distributions.
-With a model of the neutrino emission from a merger we can
-1) examine neutrino effects on the nucleosynthesis of the ejected material
-\citep{surm2011-nickel_56, robe2016-sph_nu_nucleo},
-2) explore the rich flavor oscillation physics likely to occur
-\citep{malk2012-mnr_1, malk2015-mnr_2, malk2016-mnr_3, zhu2016-mnr_nsns_remnant,
-  vaan2016-uncovering_mnr},
-3) improve closure relations used in M1 transport schemes
-\citep{ramp2002-truncated_moment, shib2011-truncated_moment,
-  card2013-truncated_moment, fouc2015-m1_nsbh, ocon2015-gr1d_with_nu}, and
-4) study possible jet formation due to neutrino annihilation
-\citep{ruff1999-nunubar_nsns, asan2000-nunubar, birk2007-nunubar,
-  hari2010-gr_nunubar_collapsar, zala2011-nunubar, leng2014-nunubar}.
-
-%\subsubsection*{Why ray tracing?}
-In this work we present a ray tracing method to compute neutrino distribution
-functions from state-of-the-art general relativistic M1 transport
-hydrodynamics simulations.
-We choose ray tracing because it is conceptually simple, numerically cheap,
-and easily extends to high resolution in energy and angle.
-
-With a ray tracing method we approach radiation transport from the perspective
-of a single observer at a spacetime event $x_o^\alpha$.
-Our goal is to compute the distribution function
-$f^{\nu_\sigma}(x_o^\alpha;p_\beta)$, or the amount of neutrino radiation
-of species $\nu_\sigma\in\{\nu_{\{e,\mu,\tau\}},\bar{\nu}_{\{e,\mu,\tau\}}\}$
-with momentum $p_\beta$ impinging on $x_o^\alpha$.
-To do so we trace a geodesic trajectory from $x_o^\alpha$ in the backwards
-direction $-p_\beta$ to sample the incoming radiation along that line of sight.
-By tracing a family of rays intersecting $x_o^\alpha$ we build up a
-picture of the distribution function there.
-And by sampling many observation points we construct a global picture
-of $f^{\nu_\sigma}(x^\alpha;p_\beta)$.
-
-The ray tracing framework
-is conceptually simple because it solves the equation of radiation
-transport (Eqn.~\ref{eqn:boltzmann}) along characteristics,
-reducing it to a one-dimensional ordinary differential equation.
-It is numerically cheap because it confines computations of
-$f^{\nu_\sigma}(x_o^\alpha;p_\beta)$ to the past light-cone of $x_o^\alpha$,
-with the history of that light-cone truncated at large optical depth.
-\todo{state scaling $\mathscr{O}(n)$?}
-And it easily extends to high resolution in energy and angle by simply
-increasing the number of rays sampled.
-
-%\subsubsection*{What is new and better about this ray tracing formulation?}
-Several ray tracing formulations for radiation transport already exist.
-Most formulations assume an analytic spacetime metric
-\citep{birk2007-nunubar, hari2010-gr_nunubar_collapsar, kova2011-gr_ray_tracing}.
-And many make the simplifying assumption of blackbody emission from a
-neutrinosurface \citep{birk2007-nunubar, kova2011-gr_ray_tracing},
-limiting them to equilibrium, optically thick configurations.
-Current state-of-the-art ray tracing formulations avoid the assumption of
-blackbody spectra by integrating a local emissivity along each geodesic
-(e.g. \cite{hari2010-gr_nunubar_collapsar} for neutrinos and
-\cite{youn2012-gr_radiative_transfer} for photons).
-But no formulations to date account for the important scattering and pair
-processes outlined in Tab.~\ref{tab:neutrino_processes}.
-
-We build upon existing ray tracing formulations
-by eschewing any assumptions about the spacetime geometry,
-integrating local emissivities,
-and including elastic scattering in the integration along each geodesic.
-To incorporate scattering processes in our method, we employ estimates of the
-lowest-order moment of the neutrino distribution function
-computed in an M1 transport simulation.
-If moments are not available either from
-an M1 transport evolution or a trustworthy analytical estimate,
-our method reduces to the current state-of-the-art ray tracing method,
-neglecting scattering.
-
-%\subsubsection*{Why does a covariant formulation matter?}
-We formulate our ray tracing equations covariantly---free from
-assumptions about the spacetime geometry or coordinates. This is essential
-because we want to apply the method as a postprocessing step using
-time snapshots of data computed from general relativistic evolutions.
-The spacetime represented in these snapshots is not analytic (i.e. Kerr).
-And even in configurations that are described well by the Kerr metric
-(e.g. a low-mass disk around a massive black hole),
-the evolution coordinates are unlikely to present the metric in
-a familiar analytic form.
-This is because integrating the Einstein Equations often requires complicated,
-time-dependent gauge conditions
-\citep{lind2007-gen_harmonic, fouc2013-compactness_and_spin}.
-
-%\subsubsection*{Why do scattering and pair processes matter?}
-Elastic scattering (see Tab.~\ref{tab:neutrino_processes}) can signicantly
-modify neutrino distributions in angle and dilute the emitted spectrum over a
-larger emitting surface \citep{pere2016-asl}.
-This is especially pronounced in the case of heavy-lepton neutrinos.
-\todo{cite}
-Inelastic scattering and pair processes can introduce further modifications.
-Any phenomena that involve neutrino-neutrino interactions,
-for example neutrino oscillations and neutrino-antineutrino annihilation,
-\todo{cite}
-depend sensitively on the angular distribution.
-And spectral changes can strongly affect the nucleosynthesis
-occuring in the ejected and irradiated material.
-\todo{justify these statements}
-The ray tracing method we present in this paper captures the dominant effects
-of elastic scattering, while leaving the physics of inelastic scattering and
-pair processes for later work.
-
-%\subsubsection*{In what ways is this method limited?}
-Ray tracing is ideally suited to problems requiring detailed knowledge of
-radiation distribution functions over small regions of spacetime:
-for example along a matter or radiation trajectory,
-or over a small volume outside a source.
-But the method becomes computationally burdensome when the inquiry extends
-to large volumes of spacetime.
-
-More fundamentally, though, ray tracing is limited by its naive
-treatment of the Boltzmann Equation (Eqn.~\ref{eqn:boltzmann}),
-a treatment which essentially decouples different neutrino momenta and species.
-\todo{clarify how momenta are decoupled}
-When we're solving for $f^{\nu_\sigma}(x^\alpha;p_\beta)$, we have no
-information about $f^{\nu_\sigma}(x^\alpha;p'_\beta)$, and
-$f^{\bar{\nu}_\sigma}(x^\alpha;p'_\beta)$, which are essential ingredients
-for the source terms of the Boltzmann Equation describing the creation and
-destruction of neutrinos due to the interaction processes described in
-Tab.~\ref{tab:neutrino_processes}.
-
-In this paper we outwit this limitation by incorporating coupled source terms
-that depend on either previously-evolved or analytical
-estimates of the neutrino fields.
-We also retain the freedom to drop the coupling terms, in which case our method
-becomes similar to existing state-of-the-art ray tracing methods.
-Our method is not a standalone radiation transport scheme,
-but serves as the final component of a hybrid scheme,
-piggy-backing on a lower-order radiation transport method as a
-post-processing step.
-
-\subsubsection*{Outline}
-In Sec.~\ref{sec:formulation} we derive the ray tracing equations from the
-Boltzmann Equation and describe our numerical scheme.
-In Sec.~\ref{sec:tests} we present tests of the code.
-In Sec.~\ref{sec:applications} we present several applications
-in the dynamical environment following the merger of two neutron stars:
-computing neutrino spectra as a function of observer's angle,
-computing the self-interaction potential descriptive of flavor evolution
-in high-neutrino-density environments,
-and computing the energy dumped into jet-forming electron-positron
-pairs via neutrino-antineutrino annihilation.
-
-Greek tensor indices ($\alpha, \beta, ...$) range over all four coordinates,
-whereas Latin indices ($i, j, ...$) range over the spatial coordinates 1--3.
-We use naturalized units in which $\hbar=c=1$.
-And for the remainder of this article we suppress the neutrino species label
-$\nu_\sigma$ since the formulation is general to any species.
-
-\section{Ray tracing formulation}
-\label{sec:formulation}
-The neutrino distribution function, $f(x^\alpha; p_\beta)$ is an invariant
-quantity counting the number of neutrinos in a given six-volume of phase
-space centered on $(x^\alpha,p_\beta)$.
-The phase space volume elements are defined with respect to a fiducial
-observer passing through event $x^\alpha$ with velocity $u^\alpha$:
-\begin{align}
-  \label{eqn:dV}
-  dV & \equiv \sqrt{-\psi} \, dx \, dy \, dz \, u^t \\
-  \label{eqn:dP}
-  dP & \equiv \frac{1}{\sqrt{-\psi}} \, dp_x \, dp_y \, dp_z \,
-  \frac{\varepsilon}{p^t},
-\end{align}
-where $\psi$ represents the determinant of the spacetime metric,
-the index $t$ indicates the time-component of the given four-vector, and
-\begin{equation}
-  \label{eqn:varepsilon}
-  \varepsilon \equiv -p_\mu u^\mu
-\end{equation}
-is the neutrino energy measured by our observer.
-The number of particles in a given six-volume is
-\begin{equation}
-  dN=\frac{g}{(2\pi)^3}\,f\,dV\,dP,
-\end{equation}
-where $g$ counts the number of spin states accessible to the
-particles ($g=1$ for neutrinos), and $f$ is the distribution function.
-Each of $dV$, $dP$, $dN$, and $f$ are spacetime invariants
-\citep{debb2009-gr_boltzmann_1, debb2009-gr_boltzmann_2, lind1966-gr_boltzmann}.
-
-We may decompose the neutrino momentum like
-\begin{equation}
-  \label{eqn:def_momentum}
-  p_\beta = \varepsilon (u_\beta + \ell_\beta),
-\end{equation}
-with $\ell_\beta$ the direction normal subject to the constraints
-$u^\alpha \ell_\alpha = 0$ and $\ell^\alpha \ell_\alpha=1$.
-With this decompositon we can write the arguments to the distribution function
-$f(x^\alpha;\varepsilon,\ell_\beta)$.
-
-Because $\ell_\beta$ is subject to two constraints
-(normalization and orthogonality to the observer's velocity)
-it has two degrees of freedom; we make this explicit by defining its
-spatial components with respect to spherical polar angles
-\begin{equation}
-  \label{eqn:def_direction}
-  \ell_\alpha :=
-  q (s,\sin\upsilon_1\cos\upsilon_2,\sin\upsilon_1\sin\upsilon_2,\cos\upsilon_1),
-\end{equation}
-with $q$ and $s$ functions of $\upsilon_1$ and $\upsilon_2$.
-Now our symbol for the distribution function,
-$f(x^\alpha;\varepsilon,\upsilon_1,\upsilon_2)$,
-makes manifest its seven independent arguments.
-\todo{also decompose wrt $p_t$}
-
-\subsection{Boltzmann Equation}
-\label{ssec:boltzmann}
-Neutrino radiation obeys the relativistic Boltzmann Equation,
-which, suppressing the arguments of $f$ for simplicity, is written,
-\todo{point out limit of QKEs}
-\begin{equation}
-  \label{eqn:boltzmann}
-  \frac{d}{d\lambda}f = C[f],
-\end{equation}
-where $d/d\lambda$ denotes a derivative with respect to the affine
-parameter defining the neutrino momentum (Eqn.~\ref{eqn:geodesic_x})
-and $C[f]$ is the source term arising from interactions with the medium,
-having dimension ${\rm energy}\,{\rm length}^{-1}$.
-\todo{describe $C[f]$ in physical terms}
-The source term varies over phase space $(x^\alpha,p_\beta)$,
-and due to Fermi blocking depends locally on the distribution function $f$
-and nonlocally on the distribution function of this neutrino and its
-antiparticle at different momenta, $f'$ and $\bar{f}'$.
-For simplicity we symbolize all of these dependencies with the shorthand $C[f]$.
-The various neutrino interactions contributing to $C[f]$ are detailed in
-App.~\ref{sec:source_terms}.
-
-We make the right hand side of Eqn.~\ref{eqn:boltzmann} explicit by writing
-the source term linear in $f$:
-\begin{equation}
-  \label{eqn:boltzmann_linear}
-  \frac{d}{d\lambda}f =
-  \mathscr{E} - \mathscr{K} f,
-\end{equation}
-where we have introduced
-\todo{describe $\mathscr{E}$ and $\mathscr{K}$ in physical terms}
-$\mathscr{E}$, the invariant total emissivity, and
-$\mathscr{K}$, the invariant total opacity.
-These coefficients are computed by considering their dependence
-on neutrino and antineutrino distribution functions at other momenta
-(i.e.\ Fermi-blocking).
-
-We consider two classes of interactions in this work:
-the absorption/emission (AE) and elastic scattering (SE) processes
-listed in Tab.~\ref{tab:neutrino_processes}.
-Thus we separate these coefficients:
-\begin{align}
-  \label{eqn:inv_emissivity_components}
-  \mathscr{E} &= \mathscr{E}_{\rm AE} + \mathscr{E}_{\rm SE},\\
-  \label{eqn:inv_opacity_components}
-  \mathscr{K} &= \mathscr{K}_{\rm AE} + \mathscr{K}_{\rm SE}.
-\end{align}
-
-The absorption/emission coefficients are computed from sums
-over the relevant emissivities and opacities
-in Tab.~\ref{tab:neutrino_processes}:
-\todo{define $j$ and $\chi$}
-\begin{align}
-  \label{eqn:ae_emissivity_summed}
-  \mathscr{E}_{\rm AE}(\varepsilon)
-  &= \varepsilon \sum_{i\,{\rm reactions}} j_i(\varepsilon), \\
-  \label{eqn:ae_opacity_summed}
-  \mathscr{K}_{\rm AE}(\varepsilon)
-  &= \frac{1}{1-f^{\rm eq}(\varepsilon)} \,
-  \varepsilon \sum_{i\,{\rm reactions}} \chi_{{\rm a},i}(\varepsilon),
-\end{align}
-where $f^{\rm eq}$ is the Fermi-Dirac distribution function with the chemical
-potential dependent on the local density, temperature, and composition of the
-fluid.
-The appearance of $f^{\rm eq}$ in Eqn.~\ref{eqn:ae_opacity_summed} is due to
-the Fermionic nature of the neutrinos, causing $\mathscr{K}_{\rm AE}$ to be
-different than the simple absorption opacity,
-a phenomenon called stimulated absorption.
-See App.~\ref{ssec:sources_ae} for details.
-\todo{identify $\mathscr{K}_{\rm AE}$ and $\kappa^*$,
-  in e.g.\ \cite{hari2010-gr_nunubar_collapsar}}
-
-The elastic scattering coefficients are computed from
-a background field, and a sum over opacities for
-the relevant reactions in Tab.~\ref{tab:neutrino_processes}:
-\begin{align}
-  \label{eqn:se_emissivity_summed}
-  \mathscr{E}_{\rm SE}(\varepsilon)
-  &= \mathscr{K}_{\rm SE}(\varepsilon) \, \Phi_{\rm SE}(\varepsilon) \, \\
-  \label{eqn:se_opacity_summed}
-  \mathscr{K}_{\rm SE}(\varepsilon)
-  &= \varepsilon \sum_{i\,{\rm reactions}} \chi_{{\rm s},i}(\varepsilon).
-\end{align}
-See App.~\ref{ssec:sources_se} for details.
-
-This paper introduces the use of neutrino densities and fluxes evolved in an
-M1 transport simulation to estimate the background field, $\Phi_{\rm SE}$.
-App.~\ref{ssec:sources_se} details a method to calculate
-$\Phi_{\rm SE}(\varepsilon)$ in two different ways:
-\begin{itemize}
-\item
-  the \emph{spectral} method using densities and fluxes
-  extracted from a simulation evolved over multiple energy groups
-  to compute the background field with Eqn.~\ref{eqn:se_factor_expansion},
-\item
-  the \emph{gray} method using energy-integrated densities and fluxes
-  extracted from a gray simulation and approximating the
-  energy distribution with Eqns.~\ref{eqn:J_from_gray} and
-  \ref{eqn:H_from_gray}.
-\end{itemize}
-
-\subsection{Trajectories}
-\label{ssec:trajectories}
-Each trajectory is uniquely labeled by a pair of vectors
-giving some event on the trajectory, $x^\alpha$,
-and the momentum at that event, $p_\beta$.
-A family of trajectories shares the same $x^\alpha$; for example,
-the family of emission trajectories is ($x^\alpha_e,p_\beta$),
-and the family of observation trajectories is ($x^\alpha_o,p_\beta$).
-
-Neutrino trajectories obey the geodesic equation, which may be decomposed into
-the coupled first-order equations
-\begin{equation}
-\label{eqn:geodesic_x}
-  \frac{d x^\alpha}{d\lambda} = p^\alpha,
-\end{equation}
-and
-\begin{equation}
-\label{eqn:geodesic_p}
-  \frac{d p_\beta}{d\lambda} = -\Gamma^\alpha_{\beta\gamma} p^\gamma p_\alpha,
-\end{equation}
-where $p^\alpha=\psi^{\alpha\beta}p_\beta$,
-$\psi^{\alpha\beta}$ is the inverse of the spacetime metric $\psi_{\alpha\beta}$,
-and $\Gamma^\alpha_{\beta\gamma}$ are the standard connection coefficients,
-\begin{equation}
-  \label{eqn:christoffel}
-  \Gamma^\alpha_{\beta\gamma} =
-  \frac{1}{2} \psi^{\alpha\mu}
-  (\psi_{\mu\beta,\gamma} + \psi_{\mu\gamma,\beta} - \psi_{\beta\gamma,\mu}),
-\end{equation}
-with the comma denoting a partial derivative
-$\psi_{\mu\beta,\gamma}=\partial_\gamma\,\psi_{\mu\beta}$.
-
-Each trajectory is parameterized by affine parameter, $\lambda$, increasing in
-the direction of $\ell_\beta$. We label $\lambda=\lambda_e$ at $x^\alpha_e$,
-as in Fig.~\ref{fig:affine_param}.
-If we multiply Eqn.~\ref{eqn:geodesic_x} by $u_\alpha \equiv dx_\alpha/ds$,
-we find $ds=\varepsilon \,d\lambda$ is the proper distance traversed by the
-neutrino as measured by our fiducial obsever $u^\alpha$.
-
-\begin{figure}
-  \includegraphics[width=\columnwidth]{fig-affine_parameter_label}
-  \caption{Affine parameterization of a neutrino trajectory of momentum
-    $p_\alpha$.
-    The fiducial observer with velocity $u^\alpha$ sits at $x^\alpha_o$,
-    the neutrino emission event is at $x^\alpha_e$. The affine parameter
-    increases from the emission event: $\lambda_e<\lambda'<\lambda''$.}
-  \label{fig:affine_param}
-\end{figure}
-
-\subsection{The Rendering Equation}
-\label{ssec:rendering_eqn}
-We can integrate Eqn.~\ref{eqn:boltzmann_linear} directly,
-\begin{align}
-  \label{eqn:rendering}
-  f(\lambda,\lambda_e) &=
-  f_{\rm bound}(\lambda,\lambda_e) + f_{\rm integ}(\lambda,\lambda_e), \\
-  \label{eqn:rendering_fbound}
-  f_{\rm bound}(\lambda,\lambda_e) &=
-  f(\lambda_e) e^{-\tau(\lambda,\lambda_e)}, \\
-  \label{eqn:rendering_finteg}
-  f_{\rm integ}(\lambda,\lambda_e) &=
-  \int_{\lambda_e}^\lambda d\lambda'\,e^{-\tau(\lambda,\lambda')}
-  \mathscr{E}(\lambda'),
-\end{align}
-where the optical depth is defined,
-\todo{rewrite this section simpler; don't give integral solution
-  since we don't use}
-\begin{equation}
-  \label{eqn:optical_depth}
-  \tau(\lambda,\lambda') \equiv -\int_{\lambda'}^\lambda
-  d\lambda'' \, \mathscr{K}(\lambda''),
-\end{equation}
-and the parameterization conventions are depicted in
-Fig.~\ref{fig:affine_param}.
-
-We further separate the integrated distribution function into
-its absorption/emission and elastic scattering components,
-$f_{\rm integ} = f_{\rm AE} + f_{\rm SE}$:
-\begin{align}
-  \label{eqn:rendering_fae}
-  f_{\rm AE}(\lambda,\lambda_e)
-  &= \int_{\lambda_e}^\lambda d\lambda'\,e^{-\tau(\lambda,\lambda')}
-  \mathscr{E_{\rm AE}}(\lambda'), \\
-  \label{eqn:rendering_fse}
-  f_{\rm SE}(\lambda,\lambda_e)
-  &= \int_{\lambda_e}^\lambda d\lambda'\,e^{-\tau(\lambda,\lambda')}
-  \mathscr{E_{\rm SE}}(\lambda').
-\end{align}
-Note that the optical depth used in both of these integrands is the total
-optical depth.
-
-\subsection{Moments of the Distribution Function}
-\label{ssec:moments}
-\todo{code these up}
-We may take angular moments of the distribution function:
-\begin{align}
-  \label{eqn:J}
-  J(\varepsilon) &=
-  \frac{\varepsilon^3}{(2\pi)^3} \oint d\Omega' f(\varepsilon, \ell'_\beta) \\
-  \label{eqn:Ha}
-  H^\mu(\varepsilon) &=
-  \frac{\varepsilon^3}{(2\pi)^3} \oint d\Omega' f(\varepsilon, \ell'_\beta) \ell'^\mu \\
-  \label{eqn:Sab}
-  S^{\mu\gamma}(\varepsilon) &=
-  \frac{\varepsilon^3}{(2\pi)^3} \oint d\Omega' f(\varepsilon, \ell'_\beta) \ell'^\mu \ell'^\gamma,
-\end{align}
-representing specific energy density, specific energy flux, and
-specific momentum flux, respectively.
-Here ``specific'' refers to the quantity being integrable over neutrino energy.
-\todo{relate moments to other frames?}
-Integrals~\ref{eqn:J}--\ref{eqn:Sab} are performed over a solid angle in
-momentum space while holding $\varepsilon$ constant:
-$d\Omega \equiv d(\cos\upsilon_1)\,d\upsilon_2$.
-\todo{consistent with Eqn.~\ref{eqn:def_direction}?}
-We will also make use of the number density and flux defined as
-$G(\varepsilon)=J(\varepsilon)/\varepsilon$
-and $K^\mu(\varepsilon)=H^\mu(\varepsilon)/\varepsilon$
-respectively.
-
-The energy-integrated moments are
-\begin{equation}
-  \label{eqn:J_H_S_eps_integrated}
-  \big\{ J,H^\mu,S^{\mu\nu} \big\} = \int_0^\infty d\varepsilon \,
-  \big\{ J(\varepsilon),H^\mu(\varepsilon),S^{\mu\nu}(\varepsilon) \big\},
-\end{equation}
-having dimension ${\rm energy}\,{\rm length}^{-3}$.
-
-In practice we use two possible fiducial observers to define the momentum
-decomposition via Eqn.~\ref{eqn:def_momentum}: the Eulerian,
-or normal observer $n_\mu=-\alpha \,\partial_\mu t$,
-and the fluid, or comoving observer
-$u^\mu=Wn^\mu+v_{\rm E}^\mu$.
-Here $t$ is coordinate time and $\alpha$ is lapse
-in the standard 3+1 decomposition of the metric
-\begin{equation}
-  \label{eqn:adm_metric}
-  \psi_{\mu\gamma} :=
-  \left(
-  \begin{matrix}
-    -\alpha^2 + \beta^i \beta_i  & \beta_i \\
-    \beta_j                      & g_{ij}
-  \end{matrix}
-  \right).
-\end{equation}
-We have also introduced
-the fluid Lorentz factor $W=-n_\mu u^\mu=\alpha u^t$,
-its Eulerian velocity $v_{\rm E}^\mu=g^\mu_\lambda u^\lambda$
-(distinct from its coordinate velocity $v^\mu=u^\mu/u^t$),
-and the projection tensor orthogonal to the normal observer
-$g^\mu_\lambda=\psi^\mu_\lambda+u^\mu u_\lambda$,
-whose spatial part is the 3-metric, $g_{ij}$.
-
-We distinguish moments computed assuming an Eulerian observer with a tilde,
-e.g. $\tilde{J}$, $\tilde{H}^\mu$, $\tilde{S}^{\mu\nu}$.
-Note that these three Eulerian moments are identical to the lab-frame moments
-$E$, $F^\mu$, and $P^{\mu\nu}$
-defined in
-\cite{shib2011-truncated_moment, ocon2015-gr1d_with_nu, fouc2015-m1_nsbh}.
-
-\subsection{Numerical Implementation}
-\label{ssec:numerical}
-Much of our numerical implementation is borrowed from the geodesic evolution
-system described in~\cite{bohn2016-code}.
-We integrate Eqns.~\ref{eqn:geodesic_x} and~\ref{eqn:geodesic_p} in the form
-given by \citep{hugh1994-eh_finding},
-and Eqns.~\ref{eqn:rendering} and~\ref{eqn:optical_depth}
-in the form given below.
-By using the time-component of Eqn.~\ref{eqn:geodesic_x} ($dt=d\lambda \, p^t$)
-we may transform our integrations to coordinate time.
-The coupled system of ordinary differential equations is
-\begin{align}
-  \label{eqn:num_x}
-  \frac{dx^i}{dt} &=
-  g^{ij} \frac{p_j}{p^t} - \beta^{i}\\
-  \label{eqn:num_p}
-  \frac{dp_i}{dt} &=
-  -\alpha \alpha_{,i} p^t
-  + \beta^k_{,i} p_k
-  - \frac{1}{2} g^{jk}_{,i}\frac{p_j p_k}{p^t}\\
-  \label{eqn:num_fae}
-  \frac{df_{\rm AE}}{dt} &=
-  e^{-\tau} \frac{\mathscr{E_{\rm AE}}}{p^t}\\
-  \label{eqn:num_fse}
-  \frac{df_{\rm SE}}{dt} &=
-  e^{-\tau} \frac{\mathscr{E_{\rm SE}}}{p^t}\\
-  \label{eqn:num_tau}
-  \frac{d\tau}{dt} &=
-  - \frac{\mathscr{K}}{p^t}.
-\end{align}
-
-We integrate each ray until we reach a terminal optical depth of
-$\tau_{\rm term}$ at the earliest emission event $x_e^\alpha$.
-The concept of earliest emission event is a facetious construct we use
-to allow us to truncate our integration at a location along the ray where any
-further additions to the field are negligible due to the large optical depth
-between $x_e^\alpha$ and $x_o^\alpha$.
-We choose $\tau_{\rm term}$ large so that
-$f_{\rm bound}$ is negligible
-(in practice $e^{-\tau_{\rm term}}=0.001$)
-and compute $f_{\rm bound}$ assuming radiative equilibrium with the matter,
-i.e. $f_{\rm bound}=(1+\exp((\varepsilon-\mu_\nu)/T)$,
-and $\mu_{\nu_e}=-\mu_{\bar{\nu}_e}=\mu_{e^{-}}+\mu_p-\mu_n$ and
-$\mu_{\nu_x}=0$.
-If the ray terminates off the numerical domain before the terminal optical
-depth accumulates, we set $f_{\rm bound}=0$.
-
-Since we don't know the emission event a priori,
-\todo{make clear ``emission event'' doesn't involve physical assumption}
-we follow the integration backwards in time,
-from $t_o$ to $t_e$.
-We begin each integration by setting initial values for the variables at $t_o$.
-The observer specifies $x^i_o$ and $p_{i,o}$,
-and we set $f_{{\rm AE},o}=f_{{\rm SE},o}=0$ and $\tau_o=0$.
-
-We integrate Eqns.~\ref{eqn:num_x}--\ref{eqn:num_tau} with adaptive step sizes,
-using the 5th-order Dormand-Prince algorithm \cite{pres2007-nr_3rd_ed},
-which produces an error estimate by comparing the 4th and 5th order solutions.
-\todo{confirm}
-After each step is taken, the errors for each of the 9 variables of
-Eqns.~\ref{eqn:num_x}--\ref{eqn:num_tau} are compared.
-If the maximum of these errors exceeds some absolute or relative threshold
-\todo{specify ScaledAbsRel algorithm}
-the step size is decreased and the step recomputed,
-if it is below some threshold the next step size is increased.
-% see OdeIntegrators::AdaptiveDense
-% and OdeSteppers::DormandPrince5
-% and OdeControllers::ProportionalIntegral_Base::StepSizeToAttempt
-% and OdeErrorMeasures::ScaledAbsRel
-\todo{specify}
-
-The volume data are components of the metric $\psi_{\alpha\beta}$,
-its derivatives $\psi_{\alpha\beta,\gamma}$,
-and the hydrodynamical data $W$, $u_i$, $\varrho$, $T$, and $Y_e$.
-These fields are stored in spectral representation.
-If they are computed via hydrodynamical simulation, they are saved to disk
-at specified coordinate times and interpolated
-with spectral interpolation in space,
-and 1st-order polynomial interpolation in time
-(as described in \cite[App.~B]{bohn2016-code}).
-If computed from a quasi-stationary solution to the
-\todo{clarify quasistationary timescale}
-general relativistic hydrodynamics equations,
-a single time is stored to disk, and no time interpolation is needed.
-
-\section{Code Tests}
-\label{sec:tests}
-
-To test our algorithm, we integrate the ray tracing equations
-(Eqns.~\ref{eqn:num_x}--\ref{eqn:num_tau}) at one observation position
-in the following configurations. This suite of configurations
-may be understood as a hierarchy of increasing physical realism:
-beginning with a homogeneous medium of effectively
-infinite extent (i.e.\ optically thick) and progressing to a model of a
-1D pre-supernova, post-bounce collapse profile
-evolved using an M1 transport hydrodynamical simulation.
-In the pre-supernova model, we compare ray tracing distributions
-to those calculated in a Monte Carlo transport simulation.
-Finally, we calculate global measures of the neutrino fields outside of the
-hypermassive neutron star and disk formed in a binary neutron star
-merger simulated by \cite{fouc2016-m1_nsns, fouc2016-m1_evolve_n},
-and compare ray tracing results to those from the M1 transport simulation.
-
-\subsection{Infinite Homogeneous Slab:
-  Testing Thermodynamic Equilibrium}
-\label{ssec:test_equilibrium}
-In an optically thick region the radiation field is in thermodynamic equilibrium
-with the matter.
-We set up a large slab of matter in two homogeneous configurations
-representative of the fluid in the Massive Collapsed Star test
-(Sec.~\ref{ssec:test_collapse}) at two different radii:
-at 50~km
-For comparision with Collapsed Star test, we use the LS180 equation of state,
-in which the equilibrium
-neutrino chemical potentials are
-$\mu_{\nu_e}=-0.639$~MeV, and 
-$\mu_{\nu_e}=5.771$~MeV respectively
-($\mu_{\bar{\nu}_e}=-\mu_{\nu_e}$ and $\mu_{\nu_x}=0$).
-Plot in Fig.~\ref{fig:homogeneous_isotropic}.
-\todo{run test at this thermodynamic point}
-
-\begin{figure}
-  \includegraphics[width=\columnwidth]{fig-homogeneous_equilibrium}
-  \caption{Equilibrium distribution functions in a homogeneous background
-    at two thermodynamic points,
-    representative of the collapse profile at
-    50~km ($\varrho=1.235 \times 10^{11}\,{\rm g}\,{\rm cm}^{-3}$,
-    $T=4$~MeV,
-    $Y_e=0.12$) and
-    200~km ($\varrho=3.026\times 10^{7}\,{\rm g}\,{\rm cm}^{-3}$,
-    $T=0.5$~MeV,
-    $Y_e=0.5$).
-    Described in Sec.~\ref{ssec:test_equilibrium}}
-  \label{fig:homogeneous_isotropic}
-\end{figure}
-
-\subsection{Infinite Homogeneous Moving Slab:
-  Testing Doppler Shift}
-\label{ssec:test_doppler}
-We reproduce the test above, but with the matter moving in
-in the $+z$ direction, with the observer stationary.
-Plot in Fig.~\ref{fig:avg_eps_doppler}.
-\todo{describe test}
-
-\begin{figure}
-  \includegraphics[width=\columnwidth]{fig-moving_fluid-avg_eps_vs_cosA}
-  \caption{Average energy $\langle\varepsilon\rangle=J/G$
-    of neutrino fields for fluid moving in the $+z$ direction,
-    with $A$ the angle between the neutrino momentum and the $+z$ direction.
-    Test described in Sec.~\ref{ssec:test_doppler}}
-  \label{fig:avg_eps_doppler}
-\end{figure}
-
-\subsection{Idealized Star:
-  Testing the Decoupling Regime}
-\label{ssec:test_ab_star}
-[Test described in~\cite[Sec.~3.2]{smit1997-two_moment}.]
-Plot in Fig.~\ref{fig:f_absorption_sphere}.
-
-\begin{figure}
-  \includegraphics[width=\columnwidth]{fig-absorption_sphere}
-  \caption{Distribution function outside a constant-density star
-    with radius $R=15$~km, and observer at $r=22.5$~km.
-    $A$ is the angle between the neutrino momentum and the $\hat{r}$ direction.
-    The sample grid uses 30 points in energy and angle spanning the ranges
-    $\varepsilon\in(0,100)$~MeV and $\cos A\in(0.6454,1)$,
-    with sample points at bin centers.
-    The line depicts the analytic result for one energy: $\varepsilon=15$~MeV.
-    Test described in Sec.~\ref{ssec:test_ab_star}}
-  \label{fig:f_absorption_sphere}
-\end{figure}
-
-\subsection{Idealized Compact Star:
-  Testing Gravitational Redshift and Geodesic Curvature}
-\label{ssec:test_gravity}
-Optically thick star in Schwarszchild.
-Total luminosity at infinite separation may be computed by
-computing the radial energy flux at a finite separation from the star, $r$,
-and using $p_t$ for the energy. We compute this asymptotic luminosity
-at several radii, $r$, and show that the luminosity is independent of
-the observer's radius.
-\todo{compute $L$, put in figure}
-
-\subsection{Post-Bounce Collapse Profile:
-  Testing Scattering}
-\label{ssec:test_collapse}
-[Standard radiation transport test described in~\cite{ocon2015-gr1d_with_nu},
-and in \cite[App.~E.6]{fouc2015-m1_nsbh},
-and in~\cite{abdi2012-monte_carlo}.]
-Plot in Fig.~\ref{fig:avg_eps_collapse}.
-
-\begin{figure}
-  \includegraphics[width=\columnwidth]{fig-collapse_profile-avg_eps_vs_cosA}
-  \caption{Average energy $\langle\varepsilon\rangle=J/G$
-    in the collapse profile test.
-    The observer is at 500~km, and the shock at 150~km,
-    so that the half-opening angle of the shock is $\cos A=0.954$.
-    The four methods depicted are 1) a fiducial Monte Carlo calculation,
-    ray tracing using the 2) spectral and 3) gray methods to estimate background
-    fields for scattering, and 4) ray tracing neglecting scattering.
-    Test described in Sec.~\ref{ssec:test_collapse}}
-  \label{fig:avg_eps_collapse}
-\end{figure}
-
-\subsection{Hypermassive Neutron Star Remnant:
-  Testing Everything}
-\label{ssec:test_disk_comparison}
-Total and relative number and energy luminosities for the disk evolved
-in \cite{fouc2016-m1_evolve_n}.
-Agreement better than 10\%.
-\todo{get accurage M1 numbers from Francois}
-
-\begin{table}%[h]
-  \caption{
-    Comparison of total luminosities and average energies between
-    ray tracing and the M1 transport simulation.
-    The ray tracing totals were computed from sums over
-    observers placed in the $y$-$z$ plane along an arc at coordinate distance
-    $r=250\,{\rm km}$ from the center of the star, at 30 positions
-    distributed evenly in $\cos\theta$ over the colatitude range
-    $\theta\in[0,90^{\circ}]$.
-    M1 values are taken from \cite[Figs. 7, 9, 10]{fouc2016-m1_evolve_n}.
-  }
-  \label{tab:nsns_rt_vs_m1}
-  \begin{tabularx}{\columnwidth}{X r r r}
-    & \,\,{\bf RT noscat} & \,\,{\bf RT scat} & \,\,{\bf M1} \\
-    \hline
-    $L_{\nu_e}\,[10^{52}{\rm erg}\,{\rm s}^{-1}]$        & 5.95 & 6.01 & 5 \\
-    $L_{\bar{\nu}_e}$                                    & 10.6 & 11.8 & 12 \\
-    $L_{\nu_x}$                                          & 30.2 & 12.6 & 12 \\
-    \hline
-    $R_{\nu_e}-R_{\bar{\nu}_e}\,[10^{57}{\rm s}^{-1}]$   & -1.06 & -1.86 & -2 \\
-    \hline
-    $\langle \varepsilon_{\nu_e} \rangle\,[{\rm MeV}]$   & 12.08 & 11.88 & 12 \\
-    $\langle \varepsilon_{\bar{\nu}_e} \rangle$          & 15.96 & 14.69 & 15 \\
-    $\langle \varepsilon_{\nu_x} \rangle$                & 36.25 & 23.67 & 26 \\
-    \hline
-  \end{tabularx}
-\end{table}
-
-\begin{figure}
-  \includegraphics[width=\columnwidth]{production-colormap-merid-rho}
-  \caption{A meridional slice of density in the hypermassive neutron
-    star and disk configuration.}
-  \label{fig:nsns_rho_merid}
-\end{figure}
-
-\begin{figure}
-  \includegraphics[width=\columnwidth]{production-colormap-equat-rho}
-  \caption{An equatorial slice of density in the hypermassive neutron
-    star and disk configuration.}
-  \label{fig:nsns_rho_equat}
-\end{figure}
-
-\begin{figure}
-  \includegraphics[width=\columnwidth]{production-colormap-merid-temp}
-  \caption{A meridional slice of temperature in the hypermassive neutron
-    star and disk configuration.}
-  \label{fig:nsns_temp_merid}
-\end{figure}
-
-\begin{figure}
-  \includegraphics[width=\columnwidth]{production-colormap-equat-temp}
-  \caption{A meridional slice of temperature in the hypermassive neutron
-    star and disk configuration.}
-  \label{fig:nsns_temp_equat}
-\end{figure}
-
-\section{Applications}
-\label{sec:applications}
-
-\subsection{Understanding Emission Spectra}
-\label{ssec:spectra}
-In this section, we use ray tracing's ability to track ``where the neutrinos
-are coming from'' to understand the features in the emission distributions
-presented in the code tests.
-
-The merger models are described in \cite{fouc2015-m1_nsbh, fouc2016-m1_nsns}.
-Something like this is presented in \cite[Figs.~10-11]{pere2014-nu_wind};
-compare.
-
-\begin{figure}
-  \includegraphics[width=\columnwidth]{theta_distrib-250km-luminosity_R}
-  \caption{Radial number fluxes, $K_r$, as functions of colatitude $\theta$.
-    Sampled for observers at fixed coordinate distances from the center
-    of the star, with $r=250\,{\rm km}$.
-    Here radial fluxes are multiplied by the portion of solid angle
-    represented by each observer, so that the total luminosity is
-    computed by summing all the points.}
-  \label{fig:nsns_theta_distrib_R}
-\end{figure}
-
-\begin{figure}
-  \includegraphics[width=\columnwidth]{theta_distrib-250km-luminosity_L}
-  \caption{Same as Fig.~\ref{fig:nsns_theta_distrib_R} but radial energy
-    fluxes, $H_r$.}
-  \label{fig:nsns_theta_distrib_L}
-\end{figure}
-
-\begin{figure}
-  \includegraphics[width=\columnwidth]{theta_distrib-250km-avg_eps}
-  \caption{Same as Fig.~\ref{fig:nsns_theta_distrib_R} but
-    average energies $\langle \varepsilon \rangle = J/G$.}
-  \label{fig:nsns_theta_distrib_avg_eps}
-\end{figure}
-
-\subsection{Neutrino Self-Interaction Potential}
-\label{ssec:V_nunu}
-In this section we present $V_{\nu\nu,0}$ along one test trajectory in the NSNS
-model, and show that including elastic scattering makes a qualitative
-difference in the character of the flavor evolution.
-See Figs.~\ref{fig:V_nunu-noscat} and \ref{fig:V_nunu-scat}.
-
-\begin{figure}
-  \includegraphics[width=\columnwidth]{20170619-potentials-Vnue_Vnua_Vnu-000Mo-022deg-noscat}
-  \caption{Neutrino self-interaction potential along a test trajectory
-    originating at the surface of the hypermassive neutron star, and
-    proceeding outward along a radial coordinate ray with angle
-    $\psi=22^{\circ}$ with respect to the polar axis.
-    The trajectory is parameterized by the coordinate length $s$.
-    This calculation was done with scattering turned off.
-    Outside of the symmetric point at $s\sim40\,{\rm km}$, the potential is
-    positive, i.e.\ $\nu_e$-dominated.
-    Compare to Fig.~\ref{fig:V_nunu-scat}.
-    }
-  \label{fig:V_nunu-noscat}
-\end{figure}
-
-\begin{figure}
-  \includegraphics[width=\columnwidth]{20170619-potentials-Vnue_Vnua_Vnu-000Mo-022deg-scat}
-  \caption{Same as Fig.~\ref{fig:V_nunu-noscat},
-    but with scattering turned on.
-    Unlike the case with scattering turned off, here the potential is
-    everywhere negative, i.e.\ $\bar{\nu}_e$-dominated.
-    }
-  \label{fig:V_nunu-scat}
-\end{figure}
-
-\subsection{Neutrino Annihilation Integral}
-\label{ssec:q_nunu}
-In this section we calculate $q_{\nu\bar{\nu}}$ at a representative point
-in the funnel of the NSNS configuration and show that including scattering makes
-a difference (or not) in our estimate of the power available to drive a jet.
-
-% ******************************************************************************
-\appendix
-
-\section{Source Terms}
-\label{sec:source_terms}
-Here we present the sources comprising the right hand side of the
-Boltzmann Equation (Eqn.~\ref{eqn:boltzmann}). The sources for the neutrino
-distribution function $f(x^\alpha;p_\beta)$ arise from collision processes
-producing, removing, or scattering to/from that point in phase space.
-The weak interaction rates for each process involve integrals of
-the neutrino distribution function $f(x^\alpha;p'_\beta)$ and that of the
-antineutrino $\bar{f}(x^\alpha;p'_\beta)$ over a momentum volume
-$dP'$ (Eqn.~\ref{eqn:dP}).
-
-We follow
-\cite{brue1985-core_collapse} and \cite[Sec.~4]{shib2011-truncated_moment}
-by separating these processes into four categories:
-\begin{equation}
-  \label{eqn:four_sources}
-  C[f] \equiv C_{\rm AE} + C_{\rm SE} + C_{\rm SI} + C_{\rm PP},
-\end{equation}
-representing charged-current absorption and emission, elastic scattering,
-inelastic scattering, and the thermal pair processes of annihilation and
-production.
-We present formulae for all of these processes, though in this work, we only
-treat absorption/emission and elastic scattering.
-Each term
-is computed by summing the weak interaction rates of the processes from the
-relevant category given in Tab.~\ref{tab:neutrino_processes}.
-These rates are computed most easily as traditional emissivities and opacities
-in the fluid frame of reference;
-but since we use them in the form of an invariant emissivity and opacity
-(defined e.g. in \ref{eqn:inv_emissivity} and \ref{eqn:inv_opacity},
-and obeying Eqn.~\ref{eqn:boltzmann_linear})
-they take the same numerical value in any frame of reference
-and are completely independent of our choice of fiducial observer $u^\alpha$.
-\todo{confirm}
-
-Each source term is unique to the neutrino or antineutrino species modeled,
-and consists of a sum over all of the processes contributing to that
-category of interaction. For example:
-$C^{\nu_e}_{{\rm AE}} = \sum_i\,C^{\nu_e}_{{\rm AE},i}$,
-where $i$ labels the absorption/emission processes involving $\nu_e$
-in Tab.~\ref{tab:neutrino_processes}.
-By contrast $C^{\nu_x}_{{\rm AE}} = 0$,
-where $\nu_x$ stands for any of the four heavy-lepton neutrinos and
-antineutrinos.
-
-We compute our rates using the open source neutrino interaction
-library \lstinline{NuLib}~\footnote{\url{http://www.nulib.org}}.
-We compile a table of sources defined over the four dimensions of
-density, temperature, electron-fraction, and neutrino energy, and
-interpolate quad-linearly to the points sampled along each ray.
-\todo{derive and type up all sources in terms of $J$ and $H^\alpha$
-  following \cite[Sec.~4]{shib2011-truncated_moment}}
-
-\begin{table}%[h]
-  \caption{
-    We analyze neutrino interaction processes in terms of these categories.
-    $\nu$ without a label represents a neutrino or antineutrino of any flavor,
-    $N$ represents a nucleon $n$ or $p$,
-    ${}^ZA$ represents a nucleus with mass number $A$ and charge $Z$, and
-    $\gamma$ represents a high-energy photon.
-    A prime indicates a change in that particle's energy.
-  }
-  \label{tab:neutrino_processes}
-  \begin{tabularx}{\columnwidth}{X X}
-    \hline \hline
-    absorption/emission
-    & $\nu_e + n \leftrightarrow e^- + p$                          \\
-    & $\bar{\nu}_e + p \leftrightarrow e^+ + n$                    \\
-    & $\nu_e + {}^AZ \leftrightarrow e^- + {}^A(Z+1)$              \\
-    & $\bar{\nu}_e + {}^AZ \leftrightarrow e^+ + {}^A(Z-1)$        \\
-    \hline
-    elastic scattering
-    & $\nu + N \leftrightarrow \nu + N$                            \\
-    & $\nu + {}^AZ \leftrightarrow \nu + {}^AZ$                    \\
-    \hline
-    inelastic scattering
-    & $\nu + e^- \leftrightarrow \nu' + e^{-'}$                    \\
-    & $\nu + e^+ \leftrightarrow \nu' + e^{+'}$                    \\
-    \hline
-    thermal pair processes
-    & $\nu + \bar{\nu} \leftrightarrow e^{-} + e^{+}$              \\
-    & $\nu + \bar{\nu} + N + N \leftrightarrow N' + N'$            \\
-    & $\nu + \bar{\nu} \leftrightarrow \gamma$                     \\
-    % & $\nu_\alpha + \bar{\nu}_\alpha \leftrightarrow \nu_\beta + \bar{\nu}_\beta $ \\
-    \hline \hline
-  \end{tabularx}
-\end{table}
-
-\subsection{Absorption and emission via charged current}
-\label{ssec:sources_ae}
-At neutrino and thermal energies well below $m_{\mu}\sim100\,{\rm MeV}$
-only charged current processes involving $\nu_e$ and $\bar{\nu}_e$ are allowed.
-For each of these processes, $i$, we may write an emission and absorption
-coefficient as a function of the interaction cross-section
-(e.g. \cite[Eqn.~A5]{brue1985-core_collapse}):
-\todo{explain Fermi-blocking}
-\begin{align}
-  \label{eqn:sources_ae_1}
-  C_{{\rm AE},i}
-  &= \mathscr{E}_{{\rm AE},i} (1-f) - \mathscr{K}_{{\rm AE},i} f \\
-  \label{eqn:sources_ae_2}
-  &= \mathscr{E}_{{\rm AE},i} - (\mathscr{E}_{{\rm AE},i} + \mathscr{K}_{{\rm AE},i}) f,
-\end{align}
-with
-\begin{equation}
-  \label{eqn:inv_emissivity}
-  \mathscr{E}_{{\rm AE},i}
-  = \varepsilon j_i
-  = \frac{(2\pi)^3}{\varepsilon^2}\eta_i,
-\end{equation}
-where $j$ is the emissivity having dimension
-$[j]={\rm length}^{-1}\,{\rm steradian}^{-1}$,
-$\eta$ is the emissivity having dimension
-$[\eta]={\rm energy}\,{\rm length}^{-3}\,{\rm time}^{-1}\,{\rm energy}^{-1}\,{\rm steradian}^{-1}$
-(natural to formulations using specific intensity instead of distribution
-functions), and
-\begin{equation}
-  \label{eqn:inv_opacity}
-  \mathscr{K}_{{\rm AE},i} = \varepsilon \chi_i,
-\end{equation}
-where $\chi$ is the traditional opacity with dimension
-$[\chi]={\rm length}^{-1}$.
-
-\subsection{Elastic scattering}
-\label{ssec:sources_se}
-Neutrino scattering on particles of mass much greater than $\varepsilon$
-is essentially iso-energetic. Following 
-\cite[Eqn.~A8]{brue1985-core_collapse} or
-\cite[Eqn.~4.20]{shib2011-truncated_moment} the collision term for the
-$i$-th process takes the form
-\begin{equation}
-  \label{eqn:integral_se}
-  C_{{\rm SE},i}
-  = \frac{\varepsilon^3}{(2\pi)^3}
-  \oint d\Omega' R_{{\rm SE},i}(\omega')
-  \big(f(\ell'_\beta) - f(\ell_\alpha)\big)
-\end{equation}
-where $R(\omega')$ is the scattering kernel from direction $\ell'_\beta$
-to direction $\ell_\alpha$ having dimension $[R]={\rm energy}^{-1}$,
-and the cosine of the scattering angle is
-$\omega' \equiv \ell'_\alpha \ell^\alpha$.
-We write this in the form
-\begin{equation}
-  \label{eqn:sources_se}
-  C_{{\rm SE},i}
-  = \mathscr{K}_{{\rm SE},i} \Phi_{{\rm SE},i} - \mathscr{K}_{{\rm SE},i} f,
-\end{equation}
-with
-\begin{align}
-  \label{eqn:se_factor}
-  \Phi_{{\rm SE},i} &\equiv
-  \frac{1}{\mathscr{K}_{{\rm SE},i}}
-  \frac{\varepsilon^3}{(2\pi)^3}
-  \oint d\Omega' R_{{\rm SE},i}(\omega')f(\ell'_\beta) \\
-  \label{eqn:se_opacity}
-  \mathscr{K}_{{\rm SE},i} &\equiv
-  \frac{\varepsilon^3}{(2\pi)^3}
-  \oint d\Omega' R_{{\rm SE},i}(\omega').
-\end{align}
-
-It is customary to approximate the scattering kernel to linear order in
-$\omega$ (as in~\cite[Eqn.~4.21]{shib2011-truncated_moment}):
-\begin{equation}
-  \label{eqn:se_kernel_expansion}
-  R_{{\rm SE},i}(\varepsilon,\omega) \approx
-  R_{{\rm SE},i}^0(\varepsilon) + \omega \, R_{{\rm SE},i}^1(\varepsilon),
-\end{equation}
-which allows us to write Eqns.~\ref{eqn:se_factor} and \ref{eqn:se_opacity} as
-\todo{drop anisotropic term and defend}
-\begin{align}
-  \label{eqn:se_factor_expansion}
-  \Phi_{{\rm SE},i}(\varepsilon) &= \frac{1}{\mathscr{K}_{{\rm SE},i}}
-  \big(R_{{\rm SE},i}^0(\varepsilon) J(\varepsilon) +
-   R_{{\rm SE},i}^1(\varepsilon) \ell_\alpha H^\alpha(\varepsilon) \big) \\
-  \label{eqn:se_opacity_expansion}
-  \mathscr{K}_{{\rm SE},i}(\varepsilon) &=
-  4\pi \frac{\varepsilon^3}{(2\pi)^3} R_{{\rm SE},i}^0(\varepsilon),
-\end{align}
-where we have made the energy-dependence explicit.
-When we get $J(\varepsilon)$ and $H^\alpha(\varepsilon)$ from a multi-group
-moment evolution, we call this the ``spectral'' method.
-
-However moment evolutions with multiple energy groups are still rare.
-Most simulations employ a gray moment scheme,
-evolving only the energy-integrated moments, $J$ and $H^\alpha$
-and sometimes the number density $G$.
-We may reconstruct the energy-dependent source terms
-from the gray moments by assuming a diluted Fermi-Dirac spectrum.
-
-We use the following procedure:
-\begin{enumerate}
-\item
-  Interpolate the fluid temperature, $T$,
-  equilibrium neutrino chemical potential, $\mu_\nu$,
-  and evolved neutrino energy and number densities in the fluid frame,
-  $J$ and $G$, from the simulation grid.
-\item
-  Compute the average neutrino energy in the fluid frame
-  \begin{equation}
-    \langle \varepsilon \rangle \equiv J/G.
-  \end{equation}
-\item
-  Compute the neutrino spectral temperature, assuming equilibrium with the
-  fluid
-  \begin{equation}
-    T_\nu = \langle \varepsilon \rangle
-    \frac{\mathscr{F}_2(\mu_\nu/T)}{\mathscr{F}_3(\mu_\nu/T)},
-  \end{equation}
-  where $\mathscr{F}_b$ is the Fermi integral
-  \begin{equation}
-    \mathscr{F}_b(\eta) = \int_0^\infty dx \, x^b (1+e^{x-\eta})^{-1}.
-  \end{equation}
-\item
-  Assume the background neutrino fields have the same total density
-  as the evolved moments
-  \begin{align}
-    \label{eqn:J_from_gray}
-    J(\varepsilon) &=
-    J \frac{\varepsilon^3}{T_\nu^4 \,\mathscr{F}_3(\mu_\nu/T)}
-    (1+\exp(\varepsilon/T_\nu-\mu_\nu/T)^{-1}, \\
-    \label{eqn:H_from_gray}
-    H^\mu(\varepsilon) &=
-    H^\mu \frac{\varepsilon^3}{T_\nu^4\, \mathscr{F}_3(\mu_\nu/T)}
-    (1+\exp(\varepsilon/T_\nu-\mu_\nu/T)^{-1}.
-  \end{align}
-\end{enumerate}
-When we use these moments in Eqn.~\ref{eqn:se_factor_expansion},
-we call this the ``gray'' method.
-
-\subsection{Inelastic scattering}
-\label{ssec:sources_si}
-(e.g. \cite[Eqn.~A6]{brue1985-core_collapse}).
-
-\subsection{Thermal pair annihilation and production}
-\label{ssec:sources_pp}
-(e.g. \cite[Eqn.~A9]{brue1985-core_collapse}).
-
-\subsubsection*{Acknowledgements}
-This research is supported in part by the Joint Institute for Nuclear
-Astrophysics---Center for the Evolution of the Elements (JINA-CEE)
-and the National Science Fondation under grant numbers X. \todo{grant nos?}
-We thank Sherwood Richers for providing his neutrino transport code
-\lstinline{Sedonu}~\footnote{\url{https://bitbucket.org/srichers/sedonu}}
-for comparisons.
-We thank Luke Roberts and Yonglin Zhu for helpful discussions.
-We are indebted to Fran\c{c}ois H\'{e}bert for his timely help in finding and
-correcting bugs in the time-interpolation algorithm.
-
-% ******************************************************************************
 \bibliography{references}

### 2016-08-18 paper.tex
Add a Conventions section to README
Insert some structure into paper
Retool to be primarily a methods paper
Add some text to the definitions section

@@
 \documentclass[aps,prd,twocolumn,superscriptaddress,groupedaddress]{revtex4}
-% Intended to be built with pdflatex, using REVTEX 4
+% Intended to be built with pdflatex, using the REVTEX 4 package.
 % Template taken from FermiLab APS Jouranl template.
-% Labeling convention example: \label{ssec:acknowledge}.
-%   Specifically, use (<=4)-letter class (e.g. chap, sec, ssec, fig, etc.),
-%   followed by a colon, followed by a shortish underscore separated name
-%   (e.g. intro, nu_energy_spectrum).
 
@@
 \usepackage{amsmath}              % for \text, align, split ...
+\usepackage{color}                % for colorful margin notes
+\usepackage[mathscr]{euscript}    % for \mathscr
 \usepackage{graphicx}             % for \includegraphics
@@
 
+\newcommand{\todo}[1]{\marginpar{\tiny{\textcolor{red}{#1}}}}
+%\renewcommand\todo[1]{} % if you wish to hide all todo notes
+
 % *******************************************************************************
@@
 
-% The following information is for internal review, please remove them for submission
 \widetext
@@
 
-\title{Neutrino Annihilation Around Massive Nuclear Accretion Disks}
+\title{General Relativistic Ray Tracing for Neutrinos}
 
 \author{M.\ Brett Deaton}
-\affiliation{Department of Physics \& Astronomy,
-  Washington State University, Pullman, WA 99164, USA}
 \affiliation{Joint Institute for Nuclear Astrophysics,
@@
 
-\author{Matthew D.\ Duez}
+\author{Jerred Jesse}
 \affiliation{Department of Physics \& Astronomy,
@@
 
+\author{Matthew D.\ Duez}
+\affiliation{Department of Physics \& Astronomy,
+  Washington State University, Pullman, WA 99164, USA}
+
+\author{G.\ C.\ McLaughlin}
+\affiliation{Department of Physics,
+  North Carolina State University, Raleigh, NC 27695, USA}
+
+% *******************************************************************************
+
 \begin{abstract}
-  This is our abstract.
+  We present a ray tracing algorithm for computing neutrino distributions in
+  general numerical spacetimes with hydrodynamical sources.
+  The formulation is fully coordinate invariant and accurately describes
+  radiation in regimes spanning optically thick to optically thin.
+  We test the new code, highlight its strengths and weaknesses, and
+  apply it to estimate power
+  liberated by neutrino-antineutrino annihilation in the funnel of a
+  post-merger accretion disk.
 \end{abstract}
@@
 
-This is some text in the body.
+Greek tensor indices ($\alpha, \beta, ...$) range over all four coordinates,
+whereas Latin indices ($i, j, ...$) range over the spatial coordinates 1--3.
+We use naturalized units in which $\hbar=c=1$.
+
+\section{Ray tracing formalism}
+
+The neutrino distribution function, $f(p_\beta; x^\alpha)$ is an invariant
+quantity counting the number of neutrinos in a given momentum volume centered on
+$p_\beta$ and spatial volume centered on $x^\alpha$.
+The neutrino momentum is written
+\begin{equation}
+  \label{eqn:def_momentum}
+  p_\beta = \varepsilon (u_\beta + \ell_\beta),
+\end{equation}
+with $u^\beta$ a general observer's velocity,
+$\varepsilon$ the energy measured by that observer,
+and $\ell_\beta$ the direction vector subject to the constraints
+$u^\alpha \ell_\alpha = 0$ and
+$\ell^\alpha \ell_\alpha=1$.
+Because $\ell$ is normalized and perpendicular to the observer's velocity,
+it has two degrees of freedom, which we can make explicit by defining its
+spatial components with respect to spherical polar angles
+\begin{equation}
+  \label{eq:def_direction}
+  \ell_i :=
+  (\sin\upsilon_1\cos\upsilon_2,\sin\upsilon_1\sin\upsilon_2,\cos\upsilon_1).
+\end{equation}
+
+Neutrino radiation obeys the relativistic Boltzmann equation
+\todo{integral isn't explicitly covariant}
+\begin{equation}
+  \label{eqn:boltzmann}
+  p^\beta \nabla_\beta f = \mathscr{E} - \mathscr{K} f
+  + \mathscr{K}_s \oint d\ell' \phi(\ell_\alpha,\ell_\beta') f,
+\end{equation}
+where $\nabla_\beta$ denotes a covariant derivative. We have introduced
+$\mathscr{E}$, the invariant emissivity,
+$\mathscr{K}\equiv\mathscr{K}_a+\mathscr{K}_s$, the invariant total opacity
+comprised of absorption and scattering opacities, and
+$\phi(\ell_\alpha,\ell_\beta')$, the elastic scattering kernel, representing the
+likelihood of scattering from the direction $\ell_\beta'$ to the direction
+$\ell_\alpha$, subject to the normalization
+\todo{confirm this}
+\begin{equation}
+  \oint d\ell' \phi(\ell_\alpha,\ell_\beta') = 1.
+  \nonumber
+\end{equation}
+
+\subsection{Trajectories}
+\label{ssc:trajectories}
+Each trajectory is labeled by a pair of vectors
+giving some position on the trajectory, $x^\alpha$,
+and the normalized momentum direction at that position,
+$\ell_\beta$.
+A family of trajectories shares the same $x^\alpha$; for example,
+the family of emission trajectories is ($x^\alpha_e,\ell_\beta$),
+and the family of observation trajectories is ($x^\alpha_o,\ell_\beta$).
+
+A neutrino trajectory obeys the geodesic equations
+\begin{equation}
+\label{eqn:geodesic_x}
+  \frac{d x^\alpha}{d\lambda} = p^\alpha,
+\end{equation}
+and
+\begin{equation}
+\label{eqn:geodesic_p}
+  \frac{d p_\beta}{d\lambda} = -\psi_{\alpha\beta}\Gamma^\alpha_{\mu\gamma} p^\mu p^\gamma,
+\end{equation}
+where $\psi_{\alpha\beta}$ is the metric, and
+$\Gamma^\alpha_{\mu\gamma}$ are the standard connection coefficients,
+\begin{equation}
+  \label{eqn:christoffel}
+  \Gamma^\beta_{\alpha\gamma} =
+  \frac{1}{2} \psi^{\beta\mu}
+  (\psi_{\mu\alpha,\gamma} + \psi_{\mu\gamma,\alpha} - \psi_{\alpha\gamma,\mu}).
+\end{equation}
+
+Each trajectory is parameterized by proper distance, $s$, increasing in
+the direction of $\ell$. We choose $s=0$ at $x^\alpha_e$.
+and $s=s_o$ at $x^\alpha_o$.
+According to Eqn.~\ref{eqn:geodesic_x} proper distance relates to the affine
+parameter by $ds=p^t d\lambda$.
+
+\section{Fundamental Tests}
+
+\section{Neutrino annihilation}
 The high disk mass model is described in \cite{deat2013-leakage}.

### 2016-08-22 paper.tex
Add notes on error in Boltzmann Equation

@@
 with $u^\beta$ a general observer's velocity,
-$\varepsilon$ the energy measured by that observer,
+$\varepsilon=-p_\mu u^\mu$ the energy measured by that observer,
 and $\ell_\beta$ the direction vector subject to the constraints
@@
 \todo{integral isn't explicitly covariant}
+\todo{correct this to differentiate in phase space}
 \begin{equation}

### 2016-09-01 paper.tex
Remove some todos and fix notation

@@
   apply it to estimate power
-  liberated by neutrino-antineutrino annihilation in the funnel of a
-  post-merger accretion disk.
+  liberated by neutrino-antineutrino annihilation near the hypermassive neutron
+  star formed by a merger.
 \end{abstract}
@@
 
-\section{Ray tracing formalism}
+\section{Ray tracing formulation}
 
@@
 Neutrino radiation obeys the relativistic Boltzmann equation
-\todo{integral isn't explicitly covariant}
-\todo{correct this to differentiate in phase space}
 \begin{equation}
   \label{eqn:boltzmann}
-  p^\beta \nabla_\beta f = \mathscr{E} - \mathscr{K} f
-  + \mathscr{K}_s \oint d\ell' \phi(\ell_\alpha,\ell_\beta') f,
+  \frac{d}{d\lambda}f = \mathscr{E} - \mathscr{K} f
+  + \mathscr{K}_s \oint d\Omega' \phi(\ell_\alpha,\ell_\beta') f,
 \end{equation}
-where $\nabla_\beta$ denotes a covariant derivative. We have introduced
+where $\frac{d}{d\lambda}$ denotes a derivative with respect to the affine
+parameter defining the neutrino momentum,
+and $d\Omega'$ is the element of area in momentum space, with $\ell_\alpha$
+ranging over the sphere.
+We have introduced
 $\mathscr{E}$, the invariant emissivity,

### 2016-10-10 paper.tex
update the abstract and create appendix

@@
 \begin{abstract}
-  We present a ray tracing algorithm for computing neutrino distributions in
-  general numerical spacetimes with hydrodynamical sources.
-  The formulation is fully coordinate invariant and accurately describes
-  radiation in regimes spanning optically thick to optically thin.
+  We present a ray tracing algorithm for computing high-resolution neutrino
+  distributions in general numerical spacetimes with hydrodynamical sources.
+  Our formulation is covariant and treats elastic scattering by incorporating
+  estimates of the neutrino fields evolved in a low-order moment simulation.
+  It handles radiation in regimes spanning optically thick to optically thin.
   We test the new code, highlight its strengths and weaknesses, and
-  apply it to estimate power
-  liberated by neutrino-antineutrino annihilation near the hypermassive neutron
-  star formed by a merger.
+  apply it to simulations of neutron star mergers to
+  1) compute neutrino fluxes as a function of energy and emission angle, and
+  2) compute the radiation pressure tensor in the funnel of the disk,
+  demonstrating a way to improve the closure relations used in the truncated
+  moment formalism.
 \end{abstract}
@@
   \label{eq:def_direction}
-  \ell_i :=
-  (\sin\upsilon_1\cos\upsilon_2,\sin\upsilon_1\sin\upsilon_2,\cos\upsilon_1).
+  \ell_\alpha :=
+  q (s,\sin\upsilon_1\cos\upsilon_2,\sin\upsilon_1\sin\upsilon_2,\cos\upsilon_1),
 \end{equation}
+with $q$ and $s$ functions of $\upsilon_1$ and $\upsilon_2$.
 
@@
   \label{eqn:boltzmann}
-  \frac{d}{d\lambda}f = \mathscr{E} - \mathscr{K} f
-  + \mathscr{K}_s \oint d\Omega' \phi(\ell_\alpha,\ell_\beta') f,
+  \frac{d}{d\lambda}f(x^\alpha;\varepsilon,\ell_\beta) =
+  C[f](x^\alpha;\varepsilon,\ell_\beta),
 \end{equation}
@@
 parameter defining the neutrino momentum,
-and $d\Omega'$ is the element of area in momentum space, with $\ell_\alpha$
-ranging over the sphere.
-We have introduced
-$\mathscr{E}$, the invariant emissivity,
-$\mathscr{K}\equiv\mathscr{K}_a+\mathscr{K}_s$, the invariant total opacity
-comprised of absorption and scattering opacities, and
-$\phi(\ell_\alpha,\ell_\beta')$, the elastic scattering kernel, representing the
-likelihood of scattering from the direction $\ell_\beta'$ to the direction
-$\ell_\alpha$, subject to the normalization
-\todo{confirm this}
+and $C[f](x^\alpha;\varepsilon,\ell_\beta)$
+is the source term arising from interactions with the medium,
+including absorption/emission, elastic scattering, inellastic scattering,
+and pair annihilation/production processes.
+We symbolize $C$ this way to remind us that it depends on the positon and
+momentum of the neutrinos in question, and nonlocally on $f$ at other positions
+and momenta.
+We compute $C$ in App.~\ref{sec:source_terms}.
+
+We make the right hand side of Eqn.~\ref{eqn:boltzmann} explicit by writing
+the source term linear in $f$:
 \begin{equation}
-  \oint d\ell' \phi(\ell_\alpha,\ell_\beta') = 1.
-  \nonumber
+  \label{eqn:boltzmann_linear}
+  \frac{d}{d\lambda}f =
+  \mathscr{E}^* - \mathscr{K}^* f,
 \end{equation}
+where we have introduced
+$\mathscr{E}^*$, the invariant emissivity, and
+$\mathscr{K}^*$, the invariant opacity, and for simplicity we have suppressed
+the arguments of $f$.
+These coefficients account for Fermi-blocking and involve integrals over the
+neutrino distribution function at other energies and angles,
+$f(\varepsilon', \ell'_\beta)$.
+We derive them for the relevant microphysical processes in
+App.~\ref{sec:source_terms}
 
@@
 
-A neutrino trajectory obeys the geodesic equations
+A neutrino trajectory obeys the geodesic equation, which may be decomposed into
+the coupled first-order equations
 \begin{equation}
@@
 \label{eqn:geodesic_p}
-  \frac{d p_\beta}{d\lambda} = -\psi_{\alpha\beta}\Gamma^\alpha_{\mu\gamma} p^\mu p^\gamma,
+  \frac{d p_\beta}{d\lambda} = -\Gamma^\alpha_{\mu\beta} p^\mu p_\alpha,
 \end{equation}
-where $\psi_{\alpha\beta}$ is the metric, and
-$\Gamma^\alpha_{\mu\gamma}$ are the standard connection coefficients,
+where $\Gamma^\alpha_{\mu\gamma}$ are the standard connection coefficients,
 \begin{equation}
@@
   \frac{1}{2} \psi^{\beta\mu}
-  (\psi_{\mu\alpha,\gamma} + \psi_{\mu\gamma,\alpha} - \psi_{\alpha\gamma,\mu}).
+  (\psi_{\mu\alpha,\gamma} + \psi_{\mu\gamma,\alpha} - \psi_{\alpha\gamma,\mu}),
 \end{equation}
+and $\psi_{\alpha\beta}$ is the spacetime metric.
 
-Each trajectory is parameterized by proper distance, $s$, increasing in
-the direction of $\ell$. We choose $s=0$ at $x^\alpha_e$.
-and $s=s_o$ at $x^\alpha_o$.
+Each trajectory is parameterized by affine parameter, $\lambda$, increasing in
+the direction of $\ell$. We choose $\lambda=0$ at $x^\alpha_e$.
+and $\lambda=\lambda_o$ at $x^\alpha_o$.
 According to Eqn.~\ref{eqn:geodesic_x} proper distance relates to the affine
@@
 
-\section{Neutrino annihilation}
+\section{Emission Spectra}
 The high disk mass model is described in \cite{deat2013-leakage}.
 
+\section{Closure Relation}
+
+% ******************************************************************************
+\appendix
+
+\section{Source Terms}
+\label{sec:source_terms}
+The source terms comprising the right hand side of the Boltzmann Equation
+(Eqn.~\ref{eqn:boltzmann}), can be analyzed in terms of the microphysical
+processes of absorption/emission, elastic scattering, inellastic scattering,
+and pair annihilation/production.
+
+% ******************************************************************************
 \bibliography{references}

### 2016-10-12 paper.tex
Write paper outline and add empty sections

@@
 
+\section{Introduction}
+
+\subsection{What physics problems do we want to address?}
+
+\subsection{Why ray tracing?}
+
+\subsection{What is new/better about this ray tracing formulation?}
+
+\subsection{Why does scattering matter?}
+
+\subsection{In what ways is this method limited?}
+
+\subsection{Outline}
+
+In Sec.~\ref{sec:formulation} we derive the ray tracing equations from the
+Boltzmann equation and describe our numerical scheme.
+In Sec.~\ref{sec:tests} we present some tests of the code.
+In Sec.~\ref{sec:applications} we present two applications:
+neutrino spectra as a function of observer's angle, and
+a variable closure relation for a truncated moment formalism.
+We present both applications in the dynamical environments following the merger
+of a neutron star--neutron star and a black hole--neutron star binary.
+
 Greek tensor indices ($\alpha, \beta, ...$) range over all four coordinates,
@@
 \section{Ray tracing formulation}
+\label{sec:formulation}
 
-The neutrino distribution function, $f(p_\beta; x^\alpha)$ is an invariant
+The neutrino distribution function, $f(x^\alpha; p_\beta)$ is an invariant
 quantity counting the number of neutrinos in a given momentum volume centered on
 $p_\beta$ and spatial volume centered on $x^\alpha$.
-The neutrino momentum is written
+We may decompose the neutrino momentum
 \begin{equation}
@@
 \end{equation}
-with $u^\beta$ a general observer's velocity,
+with $u^\beta$ some fiducial observer's velocity,
 $\varepsilon=-p_\mu u^\mu$ the energy measured by that observer,
-and $\ell_\beta$ the direction vector subject to the constraints
+and $\ell_\beta$ the direction subject to the constraints
 $u^\alpha \ell_\alpha = 0$ and
 $\ell^\alpha \ell_\alpha=1$.
-Because $\ell$ is normalized and perpendicular to the observer's velocity,
+With this decompositon we can write the arguments to the distribution function
+$f(x^\alpha;\varepsilon,\ell_\beta)$.
+Because $\ell_\beta$ is normalized and perpendicular to the observer's velocity,
 it has two degrees of freedom, which we can make explicit by defining its
@@
 
+\subsection{Numerical Implementation}
+\label{sec:numerical}
+
 \section{Fundamental Tests}
+\label{sec:tests}
 
-\section{Emission Spectra}
+\section{Applications}
+\label{sec:applications}
+
+\subsection{Emission Spectra}
+\label{ssec:spectra}
+
 The high disk mass model is described in \cite{deat2013-leakage}.
 
-\section{Closure Relation}
+\subsection{Closure Relation}
+\label{ssec:closure}
 
@@
 \label{sec:source_terms}
-The source terms comprising the right hand side of the Boltzmann Equation
-(Eqn.~\ref{eqn:boltzmann}), can be analyzed in terms of the microphysical
+Here we present the source terms comprising the right hand side of the
+Boltzmann Equation (Eqn.~\ref{eqn:boltzmann}) in terms of the microphysical
 processes of absorption/emission, elastic scattering, inellastic scattering,

### 2016-10-12 paper.tex
Add introduction

@@
 \section{Introduction}
+Neutrinos are one of the dominant energy transport phenomena at play in
+neutron star mergers, heating, cooling, and pushing the disrupted nuclear
+matter.
+In addition, they change the composition of the matter via charged current
+interactions.
+Because neutrinos scatter over length scales both large and small with
+respect to fluid scales, accurate models require a neutrino treatment that
+respects the freedom of the neutrino distribution functions to vary
+drastically from equilibrium.
+
+This is a challenging task in the environment of a merger,
+which generally lacks any spatial symmetries.
+The state of the art today employs a truncated moment formalism
+\citep{fouc2015-m1_bhns, fouc2016-m1_nsns},
+evolving the zeroth- and first-angular moments of the distribution function
+(the energy density and flux respectively) for the zeroth-energy moment
+(the total energy).
+Recently, \cite{fouc2016-m1_evolve_n} have expanded their code to also evolve
+the number densities of neutrinos, giving a more accurate estimate of the
+average neutrino energies.
+But with only two angular moments and one energy moment, the distribution
+function has very limited angular and spectral information.
 
 \subsection{What physics problems do we want to address?}
+But many interesting unsolved problems require an accurate model of the neutrino
+spectra and angular distribution.
+With a model of the neutrino emission from a merger we can
+1) examine neutrino effects on the nucleosynthesis of the ejected material
+\citep{robe2016-nsbh_sph_nucleo},
+2) explore the rich flavor oscillation physics likely to occur here
+\citep{malk2012-mnr_1, malk2014-mnr_2, malk2015-mnr_3, zhu2016-mnr_disk,
+  vaan2016-uncovering_mnr},
+3) improve closure relations used in truncated moment evolution schemes, and
+4) look at jet formation due to neutrino annihilation
+\citep{birk2007-nunubar}.
 
 \subsection{Why ray tracing?}
+In this work we present a ray tracing method to compute neutrino distribution
+functions from the state of the art general relativistic radiation
+hydrodynamics simulations.
+We choose ray tracing because it is conceptually simple, numerically cheap,
+and easily extends to high resolution in energy and angle.
 
 \subsection{What is new/better about this ray tracing formulation?}
+We build upon ray tracing methods already in the literature
+\citep{hari2010-nunubar_gr, ?}
+by including an elastic scattering approximation
+\todo{and perhaps inelastic}
+and removing any assumptions about the spacetime geometry.
+To incorporate elastic scattering in our method, we employ the lower-order
+estimates of the distribution function evolved in a moment scheme.
 
 \subsection{Why does scattering matter?}
+It is essential that we include scattering because although elastic scattering
+doesn't change the emitted spectrum, it can significantly modify neutrino
+distributions in angle, and dilute the emitted spectrum over a larger emitting
+surface \citep{yonglin_collab}, especially for heavy-lepton neutrinos.
+Any phenomena that involve neutrino-neutrino interactions (neutrino oscillations,
+and neutrino-antineutrino annihilation) can depend sensitively on the angular
+distribution.
+And a diluted spectrum can have a different effect on the nucleosynthesis
+occuring in the ejected and irradiated material.
 
 \subsection{In what ways is this method limited?}
+Ray tracing, however, has the fundamental limitation that it treats the
+Boltzmann equation as completely decoupled between different neutrino momenta
+and species. In this paper we seek to push this limitation by incorporating
+coupling terms that depend on previously-evolved estimates of the neutrino
+fields. This method is not a standalone radiation transport scheme, but
+is presented as one component of a  hybrid scheme, piggybacking on a
+lower-order radiation transport method as a post-processing step.
 
 \subsection{Outline}
-
 In Sec.~\ref{sec:formulation} we derive the ray tracing equations from the
@@
 \label{sec:formulation}
-
 The neutrino distribution function, $f(x^\alpha; p_\beta)$ is an invariant
@@
 \label{ssec:spectra}
-
-The high disk mass model is described in \cite{deat2013-leakage}.
+The merger models are described in \cite{fouc2015-m1_nsbh,fouc2016-m1_nsns}.
 

### 2016-10-12 paper.tex
Fix broken citations

@@
 respect to fluid scales, accurate models require a neutrino treatment that
-respects the freedom of the neutrino distribution functions to vary
-drastically from equilibrium.
+respects the freedom of neutrino distribution functions to vary
+drastically from thermodynamic equilibrium.
 
@@
 which generally lacks any spatial symmetries.
-The state of the art today employs a truncated moment formalism
-\citep{fouc2015-m1_bhns, fouc2016-m1_nsns},
+The state of the art today employs either a
+or a truncated moment formalism
+\citep{fouc2015-m1_nsbh, fouc2016-m1_nsns},
 evolving the zeroth- and first-angular moments of the distribution function
@@
 1) examine neutrino effects on the nucleosynthesis of the ejected material
-\citep{robe2016-nsbh_sph_nucleo},
+\citep{robe2016-sph_nu_nucleo},
 2) explore the rich flavor oscillation physics likely to occur here
-\citep{malk2012-mnr_1, malk2014-mnr_2, malk2015-mnr_3, zhu2016-mnr_disk,
+\citep{malk2012-mnr_1, malk2015-mnr_2, malk2016-mnr_3, zhu2016-mnr_nsns_remnant,
   vaan2016-uncovering_mnr},
@@
 4) look at jet formation due to neutrino annihilation
-\citep{birk2007-nunubar}.
+\citep{birk2007-nunubar, hari2010-gr_nunubar_collapsar, ?}.
 
@@
 We build upon ray tracing methods already in the literature
-\citep{hari2010-nunubar_gr, ?}
+\citep{hari2010-gr_nunubar_collapsar, youn2012-gr_radiative_transfer,
+  kova2011-gr_ray_tracing, ?}
 by including an elastic scattering approximation
@@
 distributions in angle, and dilute the emitted spectrum over a larger emitting
-surface \citep{yonglin_collab}, especially for heavy-lepton neutrinos.
+surface \citep{pere2016-asl}, especially for heavy-lepton neutrinos.
 Any phenomena that involve neutrino-neutrino interactions (neutrino oscillations,
@@
 \label{ssec:spectra}
-The merger models are described in \cite{fouc2015-m1_nsbh,fouc2016-m1_nsns}.
+The merger models are described in \cite{fouc2015-m1_nsbh, fouc2016-m1_nsns}.
 

### 2016-10-12 paper.tex
Expand the text to include meeting discussion
also fix broken citations

--- 2016-10-12 paper.tex
@@
-\documentclass[aps,prd,twocolumn,superscriptaddress,groupedaddress]{revtex4}
+\documentclass[aps,prd,twocolumn,superscriptaddress,groupedaddress]{revtex4-1}
 % Intended to be built with pdflatex, using the REVTEX 4 package.
@@
 \usepackage{hyperref}             % for \url and \href
+\usepackage{tabularx}             % for tabularx
 
@@
 
-\title{General Relativistic Ray Tracing for Neutrinos}
+\title{Scattering and Pair Processes in General Relativistic Ray Tracing for Neutrinos}
+\todo{you sure about including all processes?}
 
@@
 \begin{abstract}
-  We present a ray tracing algorithm for computing high-resolution neutrino
-  distributions in general numerical spacetimes with hydrodynamical sources.
-  Our formulation is covariant and treats elastic scattering by incorporating
-  estimates of the neutrino fields evolved in a low-order moment simulation.
+  We present a covariant ray tracing algorithm for computing high-resolution
+  neutrino distributions in general numerical spacetimes with hydrodynamical
+  sources.
+  Our formulation treats scattering and pair processes
+  by incorporating estimates of the background neutrino fields.
+  These background fields may be taken from a low-order moment simulation,
+  be supplied analytically in simple cases,
+  or be ignored entirely, in which case the method
+  reduces to a standard state-of-the-art ray tracing formulation.
   It handles radiation in regimes spanning optically thick to optically thin.
@@
 \section{Introduction}
+Neutrinos are one of the dominant energy transport phenomena at play in
+neutron star mergers, heating, cooling, and pushing the disrupted nuclear
+matter.
+In addition, they change the composition of the matter via charged current
+interactions.
+Because neutrinos scatter over length scales both large and small with
+respect to fluid scales, accurate models require a neutrino treatment that
+respects the freedom of neutrino distribution functions to vary
+drastically from thermodynamic equilibrium.
 
-\subsection{What physics problems do we want to address?}
+This is a challenging task in the environment of a merger,
+which generally lacks any spatial symmetries so that fully general
+seven-dimensional solutions to the Boltzmann Equation are not feasible.
+Leakage approximations
+\citep{deat2013-leakage, pere2016-asl, radi2016-dynamical, ?}
+capture the qualitative effects of neutrinos on the matter, but
+produce limited information about the neutrino field itself.
+Monte Carlo methods \citep{abdi2012-monte_carlo, rich2015-monte_carlo}
+suffer from noise.
+\todo{discuss mc more clearly}
+The state of the art today employs a truncated moment formalism
+\citep{shib2011-truncated_moment, fouc2015-m1_nsbh, fouc2016-m1_nsns,
+  just2015-m1_code, ?},
+evolving the zeroth- and first-angular moments of the distribution function
+(the energy density and flux respectively) representing the zeroth-energy moment
+of the radiation field (the total energy).
+Recently, \cite{fouc2016-m1_evolve_n} have expanded their code to also evolve
+the number densities of neutrinos, providing an estimate of the average neutrino
+energies in addition to their totals.
+But with only two angular moments and one energy moment, the distribution
+function has very limited angular and spectral information.
 
-\subsection{Why ray tracing?}
+\subsubsection*{What physics problems do we want to address?}
+But many interesting unsolved problems require an accurate model of the neutrino
+spectra and angular distributions.
+With a model of the neutrino emission from a merger we can
+1) examine neutrino effects on the nucleosynthesis of the ejected material
+\citep{surm2011-nickel_56, robe2016-sph_nu_nucleo},
+2) explore the rich flavor oscillation physics likely to occur
+\citep{malk2012-mnr_1, malk2015-mnr_2, malk2016-mnr_3, zhu2016-mnr_nsns_remnant,
+  vaan2016-uncovering_mnr},
+3) improve closure relations used in truncated moment schemes
+\citep{ramp2002-truncated_moment, shib2011-truncated_moment,
+  card2013-truncated_moment, fouc2015-m1_nsbh, ocon2015-gr1d_with_nu}, and
+4) study possible jet formation due to neutrino annihilation
+\citep{ruff1999-nunubar_nsns, asan2000-nunubar, birk2007-nunubar,
+  hari2010-gr_nunubar_collapsar, zala2011-nunubar, leng2014-nunubar}.
 
-\subsection{What is new/better about this ray tracing formulation?}
+\subsubsection*{Why ray tracing?}
+In this work we present a ray tracing method to compute neutrino distribution
+functions from state-of-the-art general relativistic radiation
+hydrodynamics simulations.
+We choose ray tracing because it is conceptually simple, numerically cheap,
+and easily extends to high resolution in energy and angle.
 
-\subsection{Why does scattering matter?}
+With a ray tracing method we approach radiation transport from the perspective
+of a single observer at a spacetime event $x_o^\alpha$.
+Our goal is to compute the distribution function $f(x_o^\alpha;p_\beta)$,
+or the amount of radiation with momentum $p_\beta$ impinging on $x_o^\alpha$.
+To do so we trace a geodesic trajectory in the backwards direction
+$-p_\beta$ to sample the incoming radiation along that line of sight.
+By tracing a family of rays intersecting $x_o^\alpha$ we build up a
+picture of the distribution function there.
+And by sampling many observation points we construct a global picture
+of $f(x^\alpha;p_\beta)$.
 
-\subsection{In what ways is this method limited?}
+The ray tracing framework
+is conceptually simple because it interprets the equation of radiation
+transport (Eqn.~\ref{eqn:boltzmann})
+as a one-dimensional ordinary differential equation,
+it is numerically cheap because it confines computations of
+$f(x_o^\alpha;p_\beta)$ to the past light-cone of $x_o^\alpha$,
+and it easily extends to high resolution in energy and angle by simply
+increasing the number of rays sampled.
+
+\subsubsection*{What is new and better about this ray tracing formulation?}
+Several ray tracing formulations for radiation transport already exist.
+Most formulations assume an analytic spacetime metric
+\citep{birk2007-nunubar, hari2010-gr_nunubar_collapsar, kova2011-gr_ray_tracing}.
+And many make the simplifying assumption of blackbody emission from a
+neutrinosurface \citep{birk2007-nunubar, kova2011-gr_ray_tracing}.
+Current state-of-the-art ray tracing formulations avoid the assumption of
+blackbody spectra by integrating a local emissivity along each geodesic
+(e.g. \cite{hari2010-gr_nunubar_collapsar} for neutrinos and
+\cite{youn2012-gr_radiative_transfer} for photons).
+But no formulations to date account for the important scattering and pair
+processes outlined in Tab.~\ref{tab:neutrino_processes}.
+
+We build upon existing ray tracing formulations
+by eschewing any assumptions about the spacetime geometry
+and by including scattering and pair processes in the integration along each
+geodesic.
+To incorporate these neutrino processes in our method, we employ
+estimates of the first two angular moments of the distribution function
+computed in a truncated moment scheme (formulated by
+\cite{thor1981-truncated_moment, shib2011-truncated_moment}
+and applied e.g. in \cite{ramp2002-truncated_moment,
+  card2013-truncated_moment, fouc2015-m1_nsbh, ocon2015-gr1d_with_nu}).
+If the first two angular moments are not available either from
+a truncated moment evolution or a trustworthy analytical estimate,
+our method reduces to the current state-of-the-art ray tracing method,
+neglecting scattering and pair processes.
+
+\subsubsection*{Why does a covariant formulation matter?}
+We formulate our ray tracing equations covariantly---free from
+assumptions about the spacetime geometry or coordinates. This is essential
+because we want to apply the method as a postprocessing step using
+time snapshots of data computed from general relativistic evolutions.
+The spacetime represented in these snapshots is not analytic (i.e. Kerr).
+And even in configurations that are described well by the Kerr metric
+(e.g. a low-mass disk around a massive black hole),
+the evolution coordinates are unlikely to present the metric in
+a familiar analytic form.
+This is because integrating the Einstein Equations often requires complicated,
+time-dependent guage conditions
+\citep{lind2007-gen_harmonic, fouc2013-compactness_and_spin}.
+
+\subsubsection*{Why do scattering and pair processes matter?}
+Inelastic scattering and pair processes introduce significant changes to
+a spectrum,
+\todo{clarify these terms}
+\todo{justify these statements}
+and elastic scattering can signicantly modify neutrino
+distributions in angle, and dilute the emitted spectrum over a larger emitting
+surface \citep{pere2016-asl}, especially for heavy-lepton neutrinos.
+Any phenomena that involve neutrino-neutrino interactions (neutrino oscillations,
+and neutrino-antineutrino annihilation) can depend sensitively on the angular
+distribution.
+And spectral changes can strongly affect the nucleosynthesis
+occuring in the ejected and irradiated material.
+\todo{justify these statements}
+
+\subsubsection*{In what ways is this method limited?}
+Ray tracing is ideally suited to problems requiring detailed knowledge of
+radiation distribution functions over small regions of spacetime:
+for example along a matter or radiation trajectory,
+or over a small volume outside a source.
+But the method becomes computationally burdensome when the inquiry extends
+to large volumes of spacetime.
+
+More fundamentally, though, ray tracing is limited by its naive
+treatment of the Boltzmann Equation (Eqn.~\ref{eqn:boltzmann}),
+a treatment which essentially decouples different neutrino momenta and species.
+\todo{clarify how momenta are decoupled}
+In this paper we outwit this limitation by incorporating coupling
+terms that depend on either previously-evolved or analytical
+estimates of the neutrino fields.
+We also retain the freedom to drop the coupling terms, in which case our method
+becomes similar to existing state-of-the-art ray tracing methods.
+Our method is not a standalone radiation transport scheme,
+but serves as the final component of a hybrid scheme,
+piggy-backing on a lower-order radiation transport method as a
+post-processing step.
 
 \subsection{Outline}
-
 In Sec.~\ref{sec:formulation} we derive the ray tracing equations from the
-Boltzmann equation and describe our numerical scheme.
+Boltzmann Equation and describe our numerical scheme.
 In Sec.~\ref{sec:tests} we present some tests of the code.
@@
 $\varepsilon=-p_\mu u^\mu$ the energy measured by that observer,
-and $\ell_\beta$ the direction subject to the constraints
-$u^\alpha \ell_\alpha = 0$ and
-$\ell^\alpha \ell_\alpha=1$.
+and $\ell_\beta$ the direction normal subject to the constraints
+$u^\alpha \ell_\alpha = 0$ and $\ell^\alpha \ell_\alpha=1$.
 With this decompositon we can write the arguments to the distribution function
 $f(x^\alpha;\varepsilon,\ell_\beta)$.
-Because $\ell_\beta$ is normalized and perpendicular to the observer's velocity,
-it has two degrees of freedom, which we can make explicit by defining its
+Because $\ell_\beta$ is subject to two constraints
+(normalization and orthogonality to the observer's velocity)
+it has two degrees of freedom; we this make explicit by defining its
 spatial components with respect to spherical polar angles
@@
 with $q$ and $s$ functions of $\upsilon_1$ and $\upsilon_2$.
+Now our symbol for the distribution function,
+$f(x^\alpha;\varepsilon,\upsilon_1,\upsilon_2)$,
+makes manifest its seven independent arguments.
 
-Neutrino radiation obeys the relativistic Boltzmann equation
+Neutrino radiation obeys the relativistic Boltzmann Equation
+\todo{explain limit of QKEs}
 \begin{equation}
   \label{eqn:boltzmann}
-  \frac{d}{d\lambda}f(x^\alpha;\varepsilon,\ell_\beta) =
-  C[f](x^\alpha;\varepsilon,\ell_\beta),
+  \frac{d}{d\lambda}f(x^\alpha;\varepsilon,\upsilon_1,\upsilon_2) = C[f],
 \end{equation}
 where $\frac{d}{d\lambda}$ denotes a derivative with respect to the affine
-parameter defining the neutrino momentum,
-and $C[f](x^\alpha;\varepsilon,\ell_\beta)$
-is the source term arising from interactions with the medium,
-including absorption/emission, elastic scattering, inellastic scattering,
-and pair annihilation/production processes.
-We symbolize $C$ this way to remind us that it depends on the positon and
-momentum of the neutrinos in question, and nonlocally on $f$ at other positions
-and momenta.
-We compute $C$ in App.~\ref{sec:source_terms}.
+parameter defining the neutrino momentum (Eqn.~\ref{eqn:geodesic_x})
+and $C[f]$ is the source term arising from interactions with the medium.
+The source term varies over phase space $(x^\alpha,p_\beta)$,
+and depends locally on the distribution function
+$f(x^\alpha;\varepsilon,\upsilon_1,\upsilon_2)$,
+positon $x^\alpha$, and momentum $p_\beta$ of the neutrinos in question,
+and nonlocally on the distribution function of this neutrino and its
+antiparticle at different momenta, $f'$ and $\bar{f}'$.
+For clarity we symbolize all of these dependencies with the shorthand $C[f]$.
+The various neutrino interactions contributing to $C[f]$ are detailed in
+App.~\ref{sec:source_terms}.
 
@@
 where we have introduced
-$\mathscr{E}^*$, the invariant emissivity, and
-$\mathscr{K}^*$, the invariant opacity, and for simplicity we have suppressed
-the arguments of $f$.
-These coefficients account for Fermi-blocking and involve integrals over the
-neutrino distribution function at other energies and angles,
-$f(\varepsilon', \ell'_\beta)$.
-We derive them for the relevant microphysical processes in
-App.~\ref{sec:source_terms}
+$\mathscr{E}^*$, the invariant total emissivity, and
+$\mathscr{K}^*$, the invariant total opacity,
+and for simplicity we have suppressed the arguments of $f$.
+These coefficients are derived by considering Fermi-blocking and
+dependencies on neutrino and antineutrino distribution functions at other
+momenta.
+They are derived in App.~\ref{sec:source_terms}.
 
 \subsection{Trajectories}
-\label{ssc:trajectories}
+\label{ssec:trajectories}
 Each trajectory is labeled by a pair of vectors
-giving some position on the trajectory, $x^\alpha$,
-and the normalized momentum direction at that position,
-$\ell_\beta$.
+giving some event on the trajectory, $x^\alpha$,
+and the momentum at that event, $p_\beta$.
 A family of trajectories shares the same $x^\alpha$; for example,
-the family of emission trajectories is ($x^\alpha_e,\ell_\beta$),
-and the family of observation trajectories is ($x^\alpha_o,\ell_\beta$).
+the family of emission trajectories is ($x^\alpha_e,p_\beta$),
+and the family of observation trajectories is ($x^\alpha_o,p_\beta$).
 
-A neutrino trajectory obeys the geodesic equation, which may be decomposed into
+Neutrino trajectories obey the geodesic equation, which may be decomposed into
 the coupled first-order equations
@@
 Each trajectory is parameterized by affine parameter, $\lambda$, increasing in
-the direction of $\ell$. We choose $\lambda=0$ at $x^\alpha_e$.
+the direction of $\ell_\beta$. We choose $\lambda=0$ at $x^\alpha_e$
 and $\lambda=\lambda_o$ at $x^\alpha_o$.
@@
 parameter by $ds=p^t d\lambda$.
+\todo{show}
+
+\subsection{The Rendering Equation}
+\label{ssec:rendering_eqn}
+
+\subsection{Neutrinosurface approximation}
+\label{ssec:neutrinosurface}
+
+\subsection{Moments of the Distribution Function}
+\label{ssec:moments}
 
 \subsection{Numerical Implementation}
-\label{sec:numerical}
+\label{ssec:numerical}
 
@@
 \label{sec:tests}
+
+\subsection{Dummy Tests}
+Maybe we don't need to present these tests, but make sure to establish
+confidence in the basics:
+gravitational redshift, Doppler shift, thermodynamic equilibrium.
+
+\subsection{Homogenous Absorbing Star}
+Test described in~\cite[Sec.~3.2]{smit1997-two_moment}
+
+\subsection{Homogenous Scattering Star}
+Test presented in~\cite{humm1971-grey_transfer}
+and described in~\cite[Sec.~9.1.1]{abdi2012-monte_carlo}.
+
+\subsection{Post-Bounce Neutron Core}
+Standard radiation transport test described in~\cite{ocon2015-gr1d_with_nu},
+and in \cite[App.~E.6]{fouc2015-m1_nsbh},
+and in~\cite{abdi2012-monte_carlo}.
+
+\subsection{Disk Comparison to M1}
+Compute total and relative number and energy luminosities for a disk evolved
+in one of~\cite{fouc2015-m1_nsbh, fouc2016-m1_nsns, fouc2016-m1_evolve_n}.
+Expect agreement within factors of a few (?).
 
@@
 \label{ssec:spectra}
-
-The high disk mass model is described in \cite{deat2013-leakage}.
+The merger models are described in \cite{fouc2015-m1_nsbh, fouc2016-m1_nsns}.
+Something like this is presented in~\cite[Fig.~10-11]{pere2014-nu_wind};
+compare.
 
@@
 \label{ssec:closure}
+This is a little tricky. Unless we can come up with a new improved analytic
+closure relation and use ray tracing to evaluate its trustworthiness,
+it's going to be hard to get anything useful from computing
+$P^{\alpha\beta}$.
+
+Maybe the variable Eddington closure is more suitable to this application.
+If I understand it right, applying ray tracing to that method would require
+computing $P^{\alpha\beta}$ on the fly throughout the spatial volume.
+That would certainly be doable for a quasi-stationary disk configuration.
 
@@
 processes of absorption/emission, elastic scattering, inellastic scattering,
-and pair annihilation/production.
+and pair annihilation/production
+(detailed in Tab.~\ref{tab:neutrino_processes}).
+
+\begin{table}
+  \centering
+  \begin{tabularx}{\columnwidth}{X X}
+    \hline \hline
+    absorption/emission
+    & $\nu_e + n \leftrightarrow e^- + p$                          \\
+    & $\bar{\nu}_e + p \leftrightarrow e^+ + n$                    \\
+    & $\nu_e + {}^AZ \leftrightarrow e^- + {}^A(Z+1)$              \\
+    & $\bar{\nu}_e + {}^AZ \leftrightarrow e^+ + {}^A(Z-1)$        \\
+    \hline
+    elastic scattering
+    & $\nu + N \leftrightarrow \nu + N$                            \\
+    & $\nu + {}^AZ \leftrightarrow \nu + {}^AZ$                    \\
+    \hline
+    inelastic scattering
+    & $\nu + e^- \leftrightarrow \nu' + e^{-'}$                    \\
+    & $\nu + e^+ \leftrightarrow \nu' + e^{+'}$                    \\
+    \hline
+    pair processes
+    & $\nu + \bar{\nu} \leftrightarrow e^{-} + e^{+}$              \\
+    & $\nu + \bar{\nu} + N + N \leftrightarrow N' + N'$            \\
+    & $\nu + \bar{\nu} \leftrightarrow \gamma$                     \\
+    \hline \hline
+  \end{tabularx}
+  \caption{
+    We analyze neutrino interaction processes in terms of these categories.
+    $\nu$ without a label represents a neutrino or antineutrino of any flavor,
+    $N$ represents a nucleon $n$ or $p$,
+    ${}^ZA$ represents a nucleus with mass number $A$ and charge $Z$, and
+    $\gamma$ represents a high-energy photon.
+    A prime indicates a change in that particle's energy.
+  }
+  \label{tab:neutrino_processes}
+\end{table}
 

### 2016-10-14 paper.tex
Add comprehensive todo notes to the text

@@
 Neutrino radiation obeys the relativistic Boltzmann Equation
-\todo{explain limit of QKEs}
+\todo{point out limit of QKEs}
 \begin{equation}
@@
 Each trajectory is parameterized by affine parameter, $\lambda$, increasing in
-the direction of $\ell_\beta$. We choose $\lambda=0$ at $x^\alpha_e$
-and $\lambda=\lambda_o$ at $x^\alpha_o$.
+the direction of $\ell_\beta$. We choose $\lambda=\lambda_0$ at $x^\alpha_e$
+and $\lambda=\lambda$ at $x^\alpha_o$, as in Fig.~\ref{fig:affine_param}.
+\todo{choose a better notation}
 According to Eqn.~\ref{eqn:geodesic_x} proper distance relates to the affine
 parameter by $ds=p^t d\lambda$.
-\todo{show}
+\todo{show explicitly}
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{affine_parameter_label}
+  \caption{Affine parameterization of a trajectory.}
+  \label{fig:affine_param}
+\end{figure}
 
@@
 \label{ssec:rendering_eqn}
+\todo{type up derivation from notes}
+We can integrate Eqn.~\ref{eqn:boltzmann_linear} directly,
+\begin{equation}
+  \label{eqn:rendering}
+  f(\lambda) = f(\lambda_0) e^{-\tau(\lambda,\lambda_0)} +
+  \int_{\lambda_0}^\lambda d\lambda'\,e^{-\tau(\lambda,\lambda')}
+  \mathscr{K}^*(\lambda') \mathscr{S}^*(\lambda'),
+\end{equation}
+where the parameterization conventions are depicted in
+Fig.~\ref{fig:affine_param}.
 
-\subsection{Neutrinosurface approximation}
+\subsection{Neutrinosurface Approximation}
 \label{ssec:neutrinosurface}
+\todo{derive $\tau=2/3$ approx \cite[Chap.~4]{shu1991-book_radiation},
+  and \cite{miln1921-radiative_equil_1}}
+In the simplified configuration of a plane-parallel atmosphere, the
+rendering equation reduces to 
+\begin{equation}
+\label{eqn:neutrinosurface}
+  f(\lambda) = f^{\rm eq}(\lambda_0),
+\end{equation}
+where $\tau(\lambda,\lambda_0)=2/3$.
 
@@
 \label{ssec:moments}
+\todo{clarify integral differentials}
+We may take angular moments of the distribution function:
+\begin{align}
+  J(\varepsilon) &=
+  \varepsilon^3 \oint d\Omega' f(\varepsilon, \ell'_\beta) \\
+  H^\mu(\varepsilon) &=
+  \varepsilon^3 \oint d\Omega' f(\varepsilon, \ell'_\beta) \ell^\mu \\
+  S^{\mu\gamma}(\varepsilon) &=
+  \varepsilon^3 \oint d\Omega' f(\varepsilon, \ell'_\beta) \ell^\mu \ell^\gamma,
+\end{align}
+which represent energy density, energy flux, and momentum flux, respectively.
+\todo{clarify moment frame and relate to other frames}
 
@@
 \label{ssec:numerical}
+Much of our numerical implementation is borrowed from the geodesic evolution
+system described in~\cite{bohn2016-code}.
+\todo{describe eqns for $f_{\rm integ}$ and $\tau$}
+\todo{describe $\nu$-specific numerics}
 
@@
 gravitational redshift, Doppler shift, thermodynamic equilibrium.
+\todo{setup and run dummy tests}
 
@@
 Test described in~\cite[Sec.~3.2]{smit1997-two_moment}
+\todo{Jerred: run test}
 
@@
 and described in~\cite[Sec.~9.1.1]{abdi2012-monte_carlo}.
+\todo{formulate analytic $J$ and $F^\mu$}
+\todo{Jerred: setup and run test}
 
@@
 and in~\cite{abdi2012-monte_carlo}.
+\todo{Evan: provide initial data and $H^r(\varepsilon)$}
+\todo{Jerred: setup and run test}
 
@@
 Expect agreement within factors of a few (?).
+\todo{code up moment integrals}
+\todo{compute luminosities from M1}
+\todo{compute luminosities from ray tracing}
 
@@
 compare.
+\todo{code up moment integrals}
+\todo{compute $H^r(\theta;\varepsilon)$ in nsns and nsbh}
+\todo{maybe compute at several times}
 
@@
 \label{ssec:closure}
+\todo{figure out your approach}
 This is a little tricky. Unless we can come up with a new improved analytic
@@
 it's going to be hard to get anything useful from computing
-$P^{\alpha\beta}$.
+$S^{\alpha\beta}$.
 
@@
 If I understand it right, applying ray tracing to that method would require
-computing $P^{\alpha\beta}$ on the fly throughout the spatial volume.
+computing $S^{\alpha\beta}$ on the fly throughout the spatial volume.
 That would certainly be doable for a quasi-stationary disk configuration.
+\todo{compute $S^{\alpha\beta}$ and evaluate its effect on evolution}
 
@@
 \label{sec:source_terms}
-Here we present the source terms comprising the right hand side of the
+\todo{derive and type up sources in terms of $J$ and $H^\alpha$
+  following \cite[Sec.~4]{shib2011-truncated_moment}}
+Here we present the sources comprising the right hand side of the
 Boltzmann Equation (Eqn.~\ref{eqn:boltzmann}) in terms of the microphysical

### 2016-10-16 paper.tex
Repair confusing notation for affine parameterization

@@
 Each trajectory is parameterized by affine parameter, $\lambda$, increasing in
-the direction of $\ell_\beta$. We choose $\lambda=\lambda_0$ at $x^\alpha_e$
-and $\lambda=\lambda$ at $x^\alpha_o$, as in Fig.~\ref{fig:affine_param}.
-\todo{choose a better notation}
-According to Eqn.~\ref{eqn:geodesic_x} proper distance relates to the affine
-parameter by $ds=p^t d\lambda$.
-\todo{show explicitly}
+the direction of $\ell_\beta$. We label $\lambda=\lambda_e$ at $x^\alpha_e$,
+as in Fig.~\ref{fig:affine_param}.
+If we multiply Eqn.~\ref{eqn:geodesic_x} by $u_\alpha \equiv dx_\alpha/ds$,
+we find $ds=\varepsilon \,d\lambda$ is the proper distance traversed by the
+neutrino as measured by our fiducial obsever $u^\alpha$.
 
@@
   \includegraphics[width=\columnwidth]{affine_parameter_label}
-  \caption{Affine parameterization of a trajectory.}
+  \caption{Affine parameterization of a neutrino trajectory of momentum
+    $p_\alpha$.
+    The fiducial observer with velocity $u^\alpha$ sits at $x^\alpha_o$,
+    the neutrino emission event is at $x^\alpha_e$. The affine parameter
+    increases from the emission event: $\lambda_e<\lambda'<\lambda''$.}
   \label{fig:affine_param}
@@
   \label{eqn:rendering}
-  f(\lambda) = f(\lambda_0) e^{-\tau(\lambda,\lambda_0)} +
-  \int_{\lambda_0}^\lambda d\lambda'\,e^{-\tau(\lambda,\lambda')}
+  f(\lambda) = f(\lambda_e) e^{-\tau(\lambda,\lambda_e)} +
+  \int_{\lambda_e}^\lambda d\lambda'\,e^{-\tau(\lambda,\lambda')}
   \mathscr{K}^*(\lambda') \mathscr{S}^*(\lambda'),
@@
 \label{eqn:neutrinosurface}
-  f(\lambda) = f^{\rm eq}(\lambda_0),
+  f(\lambda) = f^{\rm eq}(\lambda_e),
 \end{equation}
-where $\tau(\lambda,\lambda_0)=2/3$.
+where $\tau(\lambda,\lambda_e)=2/3$.
 

### 2016-10-16 paper.tex
Flesh out Rendering Equation and adjust table positioning

@@
 \label{ssec:rendering_eqn}
-\todo{type up derivation from notes}
-We can integrate Eqn.~\ref{eqn:boltzmann_linear} directly,
+Defining the source function
+$\mathscr{S}^*\equiv\mathscr{E}^*/\mathscr{K}^*$,
+we can integrate Eqn.~\ref{eqn:boltzmann_linear} directly,
 \begin{equation}
@@
 \end{equation}
-where the parameterization conventions are depicted in
+where the optical depth is defined,
+\begin{equation}
+  \label{eqn:optical_depth}
+  \tau(\lambda,\lambda') \equiv -\int_{\lambda'}^\lambda
+  d\lambda'' \, \mathscr{K}^*(\lambda''),
+\end{equation}
+and the parameterization conventions are depicted in
 Fig.~\ref{fig:affine_param}.
@@
 where $\tau(\lambda,\lambda_e)=2/3$.
+At an optical depth of $2/3$, the energy flux is equivalent to the
+the blackbody value $\sigma T^4$, with $\sigma$, the Stefan-Boltzmann radiation
+constant.
+\todo{verify}
 
@@
 
-\begin{table}
-  \centering
+\begin{table}[h]
+  \caption{
+    We analyze neutrino interaction processes in terms of these categories.
+    $\nu$ without a label represents a neutrino or antineutrino of any flavor,
+    $N$ represents a nucleon $n$ or $p$,
+    ${}^ZA$ represents a nucleus with mass number $A$ and charge $Z$, and
+    $\gamma$ represents a high-energy photon.
+    A prime indicates a change in that particle's energy.
+  }
+  \label{tab:neutrino_processes}
   \begin{tabularx}{\columnwidth}{X X}
@@
   \end{tabularx}
-  \caption{
-    We analyze neutrino interaction processes in terms of these categories.
-    $\nu$ without a label represents a neutrino or antineutrino of any flavor,
-    $N$ represents a nucleon $n$ or $p$,
-    ${}^ZA$ represents a nucleus with mass number $A$ and charge $Z$, and
-    $\gamma$ represents a high-energy photon.
-    A prime indicates a change in that particle's energy.
-  }
-  \label{tab:neutrino_processes}
 \end{table}

### 2016-10-17 paper.tex
Expand text

@@
 \usepackage{hyperref}             % for \url and \href
+\usepackage{listings}             % for typesetting code with \lstinline
 \usepackage{tabularx}             % for tabularx
@@
 \hyphenation{PYTHIA} % ''
+\lstset{basicstyle=\ttfamily} % define \lstinline for inline code snippet
 
@@
 The neutrino distribution function, $f(x^\alpha; p_\beta)$ is an invariant
-quantity counting the number of neutrinos in a given momentum volume centered on
-$p_\beta$ and spatial volume centered on $x^\alpha$.
-We may decompose the neutrino momentum
+quantity counting the number of neutrinos in a given six-volume of phase
+space centered on $(x^\alpha,p_\beta)$.
+The phase space volume elements are defined with respect to a fiducial
+observer passing through event $x^\alpha$ with velocity $u^\alpha$:
+\begin{align}
+  \label{eqn:dV}
+  dV & \equiv \sqrt{-\psi} \, dx \, dy \, dz \, u^t \\
+  \label{eqn:dP}
+  dP & \equiv \frac{1}{\sqrt{-\psi}} \, dp_x \, dp_y \, dp_z \,
+  \frac{1}{p^t}(-p_\mu u^\mu),
+\end{align}
+where $\psi$ represents the determinant of the spacetime metric,
+the index $t$ indicates the time-component of the given four-vector, and
+\begin{equation}
+  \label{eqn:varepsilon}
+  \varepsilon \equiv -p_\mu u^\mu
+\end{equation}
+is the neutrino energy measured by our observer.
+The number of particles in a given six-volume is
+$dN=g\,f\,dV\,dP/(2\pi)^3$,
+where $g$ counts the number of spin states accessible to the
+particles ($g=1$ for neutrinos), and $f$ is the distribution function.
+Each of $dN$, $dV$, $dP$, and $f$ are spacetime invariants
+\citep{debb2009-gr_boltzmann_1, debb2009-gr_boltzmann_2, lind1966-gr_boltzmann}.
+
+We may decompose the neutrino momentum like
 \begin{equation}
@@
 \end{equation}
-with $u^\beta$ some fiducial observer's velocity,
-$\varepsilon=-p_\mu u^\mu$ the energy measured by that observer,
-and $\ell_\beta$ the direction normal subject to the constraints
+with $\ell_\beta$ the direction normal subject to the constraints
 $u^\alpha \ell_\alpha = 0$ and $\ell^\alpha \ell_\alpha=1$.
@@
 $f(x^\alpha;\varepsilon,\ell_\beta)$.
+
 Because $\ell_\beta$ is subject to two constraints
@@
 \begin{equation}
-  \label{eq:def_direction}
+  \label{eqn:def_direction}
   \ell_\alpha :=
@@
 
+\subsection{Boltzmann Equation}
+\label{ssec:boltzmann}
 Neutrino radiation obeys the relativistic Boltzmann Equation
@@
 \label{ssec:moments}
-\todo{clarify integral differentials}
 We may take angular moments of the distribution function:
 \begin{align}
+  \label{eqn:J}
   J(\varepsilon) &=
   \varepsilon^3 \oint d\Omega' f(\varepsilon, \ell'_\beta) \\
+  \label{eqn:Ha}
   H^\mu(\varepsilon) &=
   \varepsilon^3 \oint d\Omega' f(\varepsilon, \ell'_\beta) \ell^\mu \\
+  \label{eqn:Sab}
   S^{\mu\gamma}(\varepsilon) &=
@@
 \end{align}
-which represent energy density, energy flux, and momentum flux, respectively.
-\todo{clarify moment frame and relate to other frames}
+representing specific energy density, specific energy flux, and
+specific momentum flux, respectively.
+Here ``specific'' refers to the quantity being integrable over neutrino energy.
+\todo{relate moments to other frames?}
+Integrals~\ref{eqn:J}--\ref{eqn:Sab} are performed over a solid angle in
+momentum space while holding $\varepsilon$ constant:
+$d\Omega \equiv d(\cos\upsilon_1)\,d\upsilon_2$.
+\todo{consistent with Eqn.~\ref{eqn:def_direction}?}
 
@@
 system described in~\cite{bohn2016-code}.
+We integrate Eqns.~\ref{eqn:geodesic_x} and~\ref{eqn:geodesic_p} in the form given by
+\citep{hugh1994-eh_finding}.
 \todo{describe eqns for $f_{\rm integ}$ and $\tau$}
@@
 
-\section{Fundamental Tests}
+\section{Code Tests}
 \label{sec:tests}
@@
 \subsection{Dummy Tests}
-Maybe we don't need to present these tests, but make sure to establish
+[Maybe we don't need to present these tests, but make sure to establish
 confidence in the basics:
-gravitational redshift, Doppler shift, thermodynamic equilibrium.
+gravitational redshift, Doppler shift, thermodynamic equilibrium.]
 \todo{setup and run dummy tests}
@@
 \subsection{Homogenous Absorbing Star}
-Test described in~\cite[Sec.~3.2]{smit1997-two_moment}
+[Test described in~\cite[Sec.~3.2]{smit1997-two_moment}.]
 \todo{Jerred: run test}
@@
 \subsection{Homogenous Scattering Star}
-Test presented in~\cite{humm1971-grey_transfer}
-and described in~\cite[Sec.~9.1.1]{abdi2012-monte_carlo}.
+[Test presented in~\cite{humm1971-grey_transfer}
+and described in~\cite[Sec.~9.1.1]{abdi2012-monte_carlo}.]
 \todo{formulate analytic $J$ and $F^\mu$}
@@
 \subsection{Post-Bounce Neutron Core}
-Standard radiation transport test described in~\cite{ocon2015-gr1d_with_nu},
+[Standard radiation transport test described in~\cite{ocon2015-gr1d_with_nu},
 and in \cite[App.~E.6]{fouc2015-m1_nsbh},
-and in~\cite{abdi2012-monte_carlo}.
+and in~\cite{abdi2012-monte_carlo}.]
 \todo{Evan: provide initial data and $H^r(\varepsilon)$}
@@
 \subsection{Disk Comparison to M1}
-Compute total and relative number and energy luminosities for a disk evolved
+[Compute total and relative number and energy luminosities for a disk evolved
 in one of~\cite{fouc2015-m1_nsbh, fouc2016-m1_nsns, fouc2016-m1_evolve_n}.
-Expect agreement within factors of a few (?).
+Expect agreement within factors of a few (?).]
 \todo{code up moment integrals}
@@
 The merger models are described in \cite{fouc2015-m1_nsbh, fouc2016-m1_nsns}.
-Something like this is presented in~\cite[Fig.~10-11]{pere2014-nu_wind};
-compare.
+[Something like this is presented in~\cite[Fig.~10-11]{pere2014-nu_wind};
+compare.]
 \todo{code up moment integrals}
@@
 \todo{figure out your approach}
-This is a little tricky. Unless we can come up with a new improved analytic
+[This is a little tricky. Unless we can come up with a new improved analytic
 closure relation and use ray tracing to evaluate its trustworthiness,
 it's going to be hard to get anything useful from computing
-$S^{\alpha\beta}$.
+$S^{\alpha\beta}$.]
 
-Maybe the variable Eddington closure is more suitable to this application.
+[Maybe the variable Eddington closure is more suitable to this application.
 If I understand it right, applying ray tracing to that method would require
 computing $S^{\alpha\beta}$ on the fly throughout the spatial volume.
-That would certainly be doable for a quasi-stationary disk configuration.
+That would certainly be doable for a quasi-stationary disk configuration.]
 \todo{compute $S^{\alpha\beta}$ and evaluate its effect on evolution}
@@
 \label{sec:source_terms}
-\todo{derive and type up sources in terms of $J$ and $H^\alpha$
+Here we present the sources comprising the right hand side of the
+Boltzmann Equation (Eqn.~\ref{eqn:boltzmann}). The sources for the neutrino
+distribution function $f(x^\alpha;p_\beta)$ arise from collision processes
+producing, removing, or scattering to/from that point in phase space.
+The weak interaction rates for each process involve integrals of
+the neutrino distribution function $f(x^\alpha;p'_\beta)$ and that of the
+antineutrino $\bar{f}(x^\alpha;p'_\beta)$ over a momentum volume
+$dP'$ (Eqn.~\ref{eqn:dP}).
+
+We follow
+\cite{brue1985-core_collapse} and \cite[Sec.~4]{shib2011-truncated_moment}
+by separating these processes into four categories:
+\begin{equation}
+  \label{eqn:four_sources}
+  C[f] \equiv C_{\rm AE} + C_{\rm ES} + C_{\rm IS} + C_{\rm PP},
+\end{equation}
+representing charged-current absorption and emission, elastic scattering,
+inelastic scattering, and thermal pair annihilation and production. Each term
+is computed by summing the weak interaction rates of the processes from the
+relevant category given in Tab.~\ref{tab:neutrino_processes}.
+These rates are computed most easily in the fluid frame of reference;
+but since we use them in the form of an invariant emissivity and opacity
+(Eqn.~\ref{eqn:boltzmann_linear}) they take the same numerical value in any
+frame of reference and for any choice of fiducial observer $u^\alpha$.
+\todo{confirm}
+
+In practice we compute our rates using the open source neutrino interaction
+library \lstinline{NuLib}~\footnote{\url{http://www.nulib.org}}.
+We compile a table of sources defined over the four dimensions of
+density, temperature, electron-fraction, and neutrino energy, and
+interpolate quad-linearly to the points sampled along each ray.
+\todo{derive and type up all sources in terms of $J$ and $H^\alpha$
   following \cite[Sec.~4]{shib2011-truncated_moment}}
-Here we present the sources comprising the right hand side of the
-Boltzmann Equation (Eqn.~\ref{eqn:boltzmann}) in terms of the microphysical
-processes of absorption/emission, elastic scattering, inellastic scattering,
-and pair annihilation/production
-(detailed in Tab.~\ref{tab:neutrino_processes}).
 
@@
     \hline
-    pair processes
+    thermal pair processes
     & $\nu + \bar{\nu} \leftrightarrow e^{-} + e^{+}$              \\
@@
 
+\subsection{Absorption and emission via charged current}
+\label{ssec:sources_ae}
+At neutrino and thermal energies well below $m_{\mu}\sim100\,{\rm MeV}$
+only charged current processes involving $\nu_e$ and $\bar{\nu}_e$ are allowed.
+For each of these processes, $i$, we may write an emission and absorption
+coefficient as a function of the interaction cross-section
+(e.g. \cite[Eqn.~A5]{brue1985-core_collapse}):
+\todo{explain Fermi-blocking}
+\begin{align}
+  \label{eqn:sources_ae_1}
+  C_{{\rm AE},i}
+  &= \mathscr{E}_{{\rm AE},i} (1-f) - \mathscr{K}_{{\rm AE},i} f \\
+  \label{eqn:sources_ae_2}
+  &= \mathscr{E}_{{\rm AE},i} - (\mathscr{E}_{{\rm AE},i} + \mathscr{K}_{{\rm AE},i}) f,
+\end{align}
+with
+\begin{equation}
+  \label{eqn:inv_emissivity}
+  \mathscr{E}_{{\rm AE},i} = \frac{(2\pi)^3}{\varepsilon^2}\eta_{i},
+\end{equation}
+where $\eta$ is the traditional emissivity with dimension
+$[\eta]={\rm energy}\,{\rm length}^{-3}\,{\rm time}^{-1}\,{\rm energy}^{-1}\,{\rm steradian}^{-1}$,
+and
+\begin{equation}
+  \label{eqn:inv_emissivity}
+  \mathscr{K}_{{\rm AE},i} = \varepsilon \chi_i,
+\end{equation}
+where $\chi$ is the traditional opacity with dimension
+$[\chi]={\rm length}^{-1}$.
+
+\subsection{Elastic scattering}
+\label{ssec:sources_es}
+Neutrino scattering on particles of mass much greater than $\varepsilon$
+is essentially iso-energetic
+(e.g. \cite[Eqn.~A8]{brue1985-core_collapse}).
+\begin{equation}
+  \label{eqn:sources_es}
+  C_{{\rm ES},i}
+  = \mathscr{K}_{{\rm ES},i} \Phi_{{\rm ES},i} - \mathscr{K}_{{\rm ES},i} f,
+\end{equation}
+with
+\begin{align}
+  \label{eqn:es_factor}
+  \Phi_{{\rm ES},i} &\equiv \frac{\varepsilon^3}{\mathscr{K}_{{\rm ES},i}}
+  \oint d\Omega' R(\ell_\alpha,\ell'_\beta)f(\ell'_\beta) \\
+  \label{eqn:es_opacity}
+  \mathscr{K}_{{\rm ES},i} &\equiv \varepsilon^3
+  \oint d\Omega' R(\ell_\alpha,\ell'_\beta),
+\end{align}
+where $R(\ell_\alpha,\ell'_\beta)$ is the scattering kernel from direction
+$\ell'_\beta$ to direction $\ell_\alpha$
+having dimension $[R]={\rm energy}^{-1}$.
+
+\subsection{Inelastic scattering}
+\label{ssec:sources_is}
+(e.g. \cite[Eqn.~A6]{brue1985-core_collapse}).
+
+\subsection{Thermal pair annihilation and production}
+\label{ssec:sources_pp}
+(e.g. \cite[Eqn.~A9]{brue1985-core_collapse}).
+
 % ******************************************************************************

### 2016-10-18 paper.tex
Fix typos

@@
 of a single observer at a spacetime event $x_o^\alpha$.
-Our goal is to compute the distribution function $f(x_o^\alpha;p_\beta)$,
-or the amount of radiation with momentum $p_\beta$ impinging on $x_o^\alpha$.
+Our goal is to compute the distribution function
+$f^{\nu_\sigma}(x_o^\alpha;p_\beta)$, or the amount of neutrino radiation
+with momentum $p_\beta$ impinging on $x_o^\alpha$;
+$\nu_\sigma$ labels this distribution function as describing one of the
+six species of neutrinos or antineutrinos.
 To do so we trace a geodesic trajectory in the backwards direction
@@
 And by sampling many observation points we construct a global picture
-of $f(x^\alpha;p_\beta)$.
+of $f^{\nu_\sigma}(x^\alpha;p_\beta)$.
 
@@
 it is numerically cheap because it confines computations of
-$f(x_o^\alpha;p_\beta)$ to the past light-cone of $x_o^\alpha$,
+$f^{\nu_\sigma}(x_o^\alpha;p_\beta)$ to the past light-cone of $x_o^\alpha$,
 and it easily extends to high resolution in energy and angle by simply
@@
 We use naturalized units in which $\hbar=c=1$.
+And for the remainder of this article we suppress the neutrino species label
+$\nu_\sigma$ since the formulation is general to any species.
 
@@
 (normalization and orthogonality to the observer's velocity)
-it has two degrees of freedom; we this make explicit by defining its
+it has two degrees of freedom; we make this explicit by defining its
 spatial components with respect to spherical polar angles
@@
 The source term varies over phase space $(x^\alpha,p_\beta)$,
-and depends locally on the distribution function
-$f(x^\alpha;\varepsilon,\upsilon_1,\upsilon_2)$,
-positon $x^\alpha$, and momentum $p_\beta$ of the neutrinos in question,
+and depends locally on the distribution function $f$
 and nonlocally on the distribution function of this neutrino and its
@@
 At an optical depth of $2/3$, the energy flux is equivalent to the
-the blackbody value $\sigma T^4$, with $\sigma$, the Stefan-Boltzmann radiation
+blackbody value $\sigma T^4$, with $\sigma$, the Stefan-Boltzmann radiation
 constant.
@@
   \label{eqn:four_sources}
-  C[f] \equiv C_{\rm AE} + C_{\rm ES} + C_{\rm IS} + C_{\rm PP},
+  C[f] \equiv C_{\rm AE} + C_{\rm SE} + C_{\rm SI} + C_{\rm TP},
 \end{equation}
@@
 relevant category given in Tab.~\ref{tab:neutrino_processes}.
-These rates are computed most easily in the fluid frame of reference;
+These rates are computed most easily as traditional emissivities and opacities
+in the fluid frame of reference;
 but since we use them in the form of an invariant emissivity and opacity
-(Eqn.~\ref{eqn:boltzmann_linear}) they take the same numerical value in any
-frame of reference and for any choice of fiducial observer $u^\alpha$.
+(defined e.g. in \ref{eqn:inv_emissivity} and \ref{eqn:inv_opacity},
+and obeying Eqn.~\ref{eqn:boltzmann_linear})
+they take the same numerical value in any frame of reference
+and are completely independent of our choice of fiducial observer $u^\alpha$.
 \todo{confirm}
 
-In practice we compute our rates using the open source neutrino interaction
+Each source term is unique to the neutrino or antineutrino species modeled,
+and consists of a sum over all of the processes contributing to that
+category of interaction. For example:
+$C^{\nu_e}_{{\rm AE}} = \sum_i\,C^{\nu_e}_{{\rm AE},i}$,
+where $i$ labels the absorption/emission processes involving $\nu_e$
+in Tab.~\ref{tab:neutrino_processes}.
+By contrast $C^{\nu_x}_{{\rm AE}} = 0$,
+where $\nu_x$ stands for any of the four heavy-lepton neutrinos and
+antineutrinos.
+
+We compute our rates using the open source neutrino interaction
 library \lstinline{NuLib}~\footnote{\url{http://www.nulib.org}}.
@@
 
-\begin{table}[h]
+\begin{table}%[h]
   \caption{
@@
 \begin{equation}
-  \label{eqn:inv_emissivity}
+  \label{eqn:inv_opacity}
   \mathscr{K}_{{\rm AE},i} = \varepsilon \chi_i,
@@
   \label{eqn:sources_es}
-  C_{{\rm ES},i}
-  = \mathscr{K}_{{\rm ES},i} \Phi_{{\rm ES},i} - \mathscr{K}_{{\rm ES},i} f,
+  C_{{\rm SE},i}
+  = \mathscr{K}_{{\rm SE},i} \Phi_{{\rm SE},i} - \mathscr{K}_{{\rm SE},i} f,
 \end{equation}
@@
   \label{eqn:es_factor}
-  \Phi_{{\rm ES},i} &\equiv \frac{\varepsilon^3}{\mathscr{K}_{{\rm ES},i}}
+  \Phi_{{\rm SE},i} &\equiv \frac{\varepsilon^3}{\mathscr{K}_{{\rm SE},i}}
   \oint d\Omega' R(\ell_\alpha,\ell'_\beta)f(\ell'_\beta) \\
   \label{eqn:es_opacity}
-  \mathscr{K}_{{\rm ES},i} &\equiv \varepsilon^3
+  \mathscr{K}_{{\rm SE},i} &\equiv \varepsilon^3
   \oint d\Omega' R(\ell_\alpha,\ell'_\beta),

### 2016-10-18 paper.tex
Consider dropping the scattering sphere test

@@
 \subsection{Homogenous Scattering Star}
+\todo{maybe drop this test}
 [Test presented in~\cite{humm1971-grey_transfer}
-and described in~\cite[Sec.~9.1.1]{abdi2012-monte_carlo}.]
-\todo{formulate analytic $J$ and $F^\mu$}
-\todo{Jerred: setup and run test}
+and described in~\cite[Sec.~9.1.1]{abdi2012-monte_carlo}.
+Our version would be a point source embedded in a spherical distribution
+of purely elastic-scattering material. It requires an analytical estimate
+of $J$.]
+\todo{formulate analytic $J$.}
 

### 2016-10-19 paper.tex
Add todo: code up the moment integrals
Remove grouped address setting for Revtex 4.1
Reduce scope of paper to simply including elastic scattering

@@
-\documentclass[aps,prd,twocolumn,superscriptaddress,groupedaddress]{revtex4-1}
-% Intended to be built with pdflatex, using the REVTEX 4 package.
+\documentclass[aps,prd,twocolumn,superscriptaddress]{revtex4-1}
+% Intended to be built with pdflatex, using the REVTEX 4.1 package.
 % Template taken from FermiLab APS Jouranl template.
@@
 
-\title{Scattering and Pair Processes in General Relativistic Ray Tracing for Neutrinos}
-\todo{you sure about including all processes?}
+\title{Elastic Scattering in General Relativistic Ray Tracing for Neutrinos}
 
@@
   sources.
-  Our formulation treats scattering and pair processes
-  by incorporating estimates of the background neutrino fields.
+  Our formulation treats elastic scattering of neutrinos (changing direction
+  but not energy) off of nuclei and
+  nucleons by incorporating estimates of the background neutrino fields.
   These background fields may be taken from a low-order moment simulation,
@@
 respects the freedom of neutrino distribution functions to vary
-drastically from thermodynamic equilibrium.
+drastically from simple geometric distributions in thermodynamic equilibrium.
 
@@
 by eschewing any assumptions about the spacetime geometry
-and by including scattering and pair processes in the integration along each
+and by including elastic scattering in the integration along each
 geodesic.
-To incorporate these neutrino processes in our method, we employ
-estimates of the first two angular moments of the distribution function
+To incorporate scattering processes in our method, we employ estimates of the
+lowest-order moment of the neutrino distribution function
 computed in a truncated moment scheme (formulated by
@@
   card2013-truncated_moment, fouc2015-m1_nsbh, ocon2015-gr1d_with_nu}).
-If the first two angular moments are not available either from
+If moments are not available either from
 a truncated moment evolution or a trustworthy analytical estimate,
 our method reduces to the current state-of-the-art ray tracing method,
-neglecting scattering and pair processes.
+neglecting scattering.
 
@@
 \subsubsection*{Why do scattering and pair processes matter?}
-Inelastic scattering and pair processes introduce significant changes to
-a spectrum,
-\todo{clarify these terms}
-\todo{justify these statements}
-and elastic scattering can signicantly modify neutrino
-distributions in angle, and dilute the emitted spectrum over a larger emitting
-surface \citep{pere2016-asl}, especially for heavy-lepton neutrinos.
+Elastic scattering (see Tab.~\ref{tab:neutrino_processes}) can signicantly
+modify neutrino distributions in angle and dilute the emitted spectrum over a
+larger emitting surface \citep{pere2016-asl}.
+This is especially pronounced in the case of heavy-lepton neutrinos.
+\todo{cite}
+Inelastic scattering and pair processes can introduce further modifications.
 Any phenomena that involve neutrino-neutrino interactions (neutrino oscillations,
@@
 \todo{justify these statements}
+The ray tracing method we present in this paper captures the dominant effects
+of elastic scattering, while leaving the physics of inelastic scattering and
+pair processes for later work.
 
@@
 \todo{clarify how momenta are decoupled}
-In this paper we outwit this limitation by incorporating coupling
-terms that depend on either previously-evolved or analytical
+When we're solving for $f^{\nu_\sigma}(x^\alpha;p_\beta)$, we have no
+information about $f^{\nu_\sigma}(x^\alpha;p'_\beta)$, and
+$f^{\bar{\nu}_\sigma}(x^\alpha;p'_\beta)$, which are essential ingredients
+for the source terms of the Boltzmann Equation describing the creation and
+destruction of neutrinos due to the interaction processes described in
+Tab.~\ref{tab:neutrino_processes}.
+
+In this paper we outwit this limitation by incorporating coupled source terms
+that depend on either previously-evolved or analytical
 estimates of the neutrino fields.
@@
 \label{ssec:moments}
+\todo{code these up}
 We may take angular moments of the distribution function:
@@
 representing charged-current absorption and emission, elastic scattering,
-inelastic scattering, and thermal pair annihilation and production. Each term
+inelastic scattering, and thermal pair annihilation and production.
+We present formulae for all of these processes, though in this work, we only
+treat absorption/emission and elastic scattering.
+Each term
 is computed by summing the weak interaction rates of the processes from the

### 2016-10-26 paper.tex
Expand scattering: Legendre expansion of kernel

@@
 parameter defining the neutrino momentum (Eqn.~\ref{eqn:geodesic_x})
-and $C[f]$ is the source term arising from interactions with the medium.
+and $C[f]$ is the source term arising from interactions with the medium,
+having dimension ${\rm energy}\,{\rm length}^{-1}$.
 The source term varies over phase space $(x^\alpha,p_\beta)$,
@@
   J(\varepsilon) &=
-  \varepsilon^3 \oint d\Omega' f(\varepsilon, \ell'_\beta) \\
+  \frac{\varepsilon^3}{(2\pi)^3} \oint d\Omega' f(\varepsilon, \ell'_\beta) \\
   \label{eqn:Ha}
   H^\mu(\varepsilon) &=
-  \varepsilon^3 \oint d\Omega' f(\varepsilon, \ell'_\beta) \ell^\mu \\
+  \frac{\varepsilon^3}{(2\pi)^3} \oint d\Omega' f(\varepsilon, \ell'_\beta) \ell^\mu \\
   \label{eqn:Sab}
   S^{\mu\gamma}(\varepsilon) &=
-  \varepsilon^3 \oint d\Omega' f(\varepsilon, \ell'_\beta) \ell^\mu \ell^\gamma,
+  \frac{\varepsilon^3}{(2\pi)^3} \oint d\Omega' f(\varepsilon, \ell'_\beta) \ell^\mu \ell^\gamma,
 \end{align}
@@
 \todo{consistent with Eqn.~\ref{eqn:def_direction}?}
+
+The energy-integrated moments are
+\begin{equation}
+  \label{eqn:J_H_S_eps_integrated}
+  \big\{ J,H^\mu,S^{\mu\nu} \big\} = \int_0^\infty d\varepsilon \,
+  \big\{ J(\varepsilon),H^\mu(\varepsilon),S^{\mu\nu}(\varepsilon) \big\},
+\end{equation}
+having dimension ${\rm energy}\,{\rm length}^{-3}$.
 
@@
   \label{eqn:four_sources}
-  C[f] \equiv C_{\rm AE} + C_{\rm SE} + C_{\rm SI} + C_{\rm TP},
+  C[f] \equiv C_{\rm AE} + C_{\rm SE} + C_{\rm SI} + C_{\rm PP},
 \end{equation}
 representing charged-current absorption and emission, elastic scattering,
-inelastic scattering, and thermal pair annihilation and production.
+inelastic scattering, and the thermal pair processes of annihilation and
+production.
 We present formulae for all of these processes, though in this work, we only
@@
 \subsection{Elastic scattering}
-\label{ssec:sources_es}
+\label{ssec:sources_se}
 Neutrino scattering on particles of mass much greater than $\varepsilon$
-is essentially iso-energetic
-(e.g. \cite[Eqn.~A8]{brue1985-core_collapse}).
+is essentially iso-energetic. Following 
+\cite[Eqn.~A8]{brue1985-core_collapse} or
+\cite[Eqn.~4.20]{shib2011-truncated_moment} the collision term for the
+$i$-th process takes the form
 \begin{equation}
-  \label{eqn:sources_es}
+  \label{eqn:integral_se}
+  C_{{\rm SE},i}
+  = \frac{\varepsilon^3}{(2\pi)^3}
+  \oint d\Omega' R_{{\rm SE},i}(\omega')
+  \big(f(\ell'_\beta) - f(\ell_\alpha)\big)
+\end{equation}
+where $R(\omega')$ is the scattering kernel from direction $\ell'_\beta$
+to direction $\ell_\alpha$ having dimension $[R]={\rm energy}^{-1}$,
+and the cosine of the scattering angle is
+$\omega' \equiv 1-\ell'_\alpha \ell^\alpha$.
+We write this in the form
+\begin{equation}
+  \label{eqn:sources_se}
   C_{{\rm SE},i}
@@
 \begin{align}
-  \label{eqn:es_factor}
-  \Phi_{{\rm SE},i} &\equiv \frac{\varepsilon^3}{\mathscr{K}_{{\rm SE},i}}
-  \oint d\Omega' R(\ell_\alpha,\ell'_\beta)f(\ell'_\beta) \\
-  \label{eqn:es_opacity}
-  \mathscr{K}_{{\rm SE},i} &\equiv \varepsilon^3
-  \oint d\Omega' R(\ell_\alpha,\ell'_\beta),
+  \label{eqn:se_factor}
+  \Phi_{{\rm SE},i} &\equiv
+  \frac{1}{\mathscr{K}_{{\rm SE},i}}
+  \frac{\varepsilon^3}{(2\pi)^3}
+  \oint d\Omega' R_{{\rm SE},i}(\omega')f(\ell'_\beta) \\
+  \label{eqn:se_opacity}
+  \mathscr{K}_{{\rm SE},i} &\equiv
+  \frac{\varepsilon^3}{(2\pi)^3}
+  \oint d\Omega' R_{{\rm SE},i}(\omega').
 \end{align}
-where $R(\ell_\alpha,\ell'_\beta)$ is the scattering kernel from direction
-$\ell'_\beta$ to direction $\ell_\alpha$
-having dimension $[R]={\rm energy}^{-1}$.
+
+It is customary to approximate the scattering kernel to linear order in
+$\omega$ (as in~\cite[Eqn.~4.21]{shib2011-truncated_moment}):
+\begin{equation}
+  \label{eqn:se_kernel_expansion}
+  R(\varepsilon,\omega) \approx R_0(\varepsilon) + \omega R_1(\varepsilon),
+\end{equation}
+which allows us to write Eqns.~\ref{eqn:se_factor} and \ref{eqn:se_opacity} as
+\begin{align}
+  \label{eqn:se_factor_expansion}
+  \Phi_{{\rm SE},i} &=  \frac{(2\pi)^3}{\varepsilon^3}\frac{1}{4\pi}
+  J(\varepsilon) -
+  \frac{1}{\mathscr{K}_{{\rm SE},i}} \ell^\alpha H_\alpha(\varepsilon)\\
+  \label{eqn:se_opacity_expansion}
+  \mathscr{K}_{{\rm SE},i} &= 4\pi \frac{\varepsilon^3}{(2\pi)^3}
+  \big(R_0(\varepsilon)+R_1(\varepsilon)\big).
+\end{align}
 
 \subsection{Inelastic scattering}
-\label{ssec:sources_is}
+\label{ssec:sources_si}
 (e.g. \cite[Eqn.~A6]{brue1985-core_collapse}).

### 2016-10-26 paper.tex
Expand elastic scattering section to use J(eps) and H^a(eps)

@@
   H^\mu(\varepsilon) &=
-  \frac{\varepsilon^3}{(2\pi)^3} \oint d\Omega' f(\varepsilon, \ell'_\beta) \ell^\mu \\
+  \frac{\varepsilon^3}{(2\pi)^3} \oint d\Omega' f(\varepsilon, \ell'_\beta) \ell'^\mu \\
   \label{eqn:Sab}
   S^{\mu\gamma}(\varepsilon) &=
-  \frac{\varepsilon^3}{(2\pi)^3} \oint d\Omega' f(\varepsilon, \ell'_\beta) \ell^\mu \ell^\gamma,
+  \frac{\varepsilon^3}{(2\pi)^3} \oint d\Omega' f(\varepsilon, \ell'_\beta) \ell'^\mu \ell'^\gamma,
 \end{align}
@@
     & $\nu + \bar{\nu} \leftrightarrow \gamma$                     \\
+    % & $\nu_\alpha + \bar{\nu}_\alpha \leftrightarrow \nu_\beta + \bar{\nu}_\beta $ \\
     \hline \hline
@@
 and the cosine of the scattering angle is
-$\omega' \equiv 1-\ell'_\alpha \ell^\alpha$.
+$\omega' \equiv \ell'_\alpha \ell^\alpha$.
 We write this in the form
@@
   \label{eqn:se_kernel_expansion}
-  R(\varepsilon,\omega) \approx R_0(\varepsilon) + \omega R_1(\varepsilon),
+  R_{{\rm SE},i}(\varepsilon,\omega) \approx
+  R_{{\rm SE},i}^0(\varepsilon) + \omega \, R_{{\rm SE},i}^1(\varepsilon),
 \end{equation}
@@
   \label{eqn:se_factor_expansion}
-  \Phi_{{\rm SE},i} &=  \frac{(2\pi)^3}{\varepsilon^3}\frac{1}{4\pi}
-  J(\varepsilon) -
-  \frac{1}{\mathscr{K}_{{\rm SE},i}} \ell^\alpha H_\alpha(\varepsilon)\\
+  \Phi_{{\rm SE},i}(\varepsilon) &= \frac{1}{\mathscr{K}_{{\rm SE},i}}
+  \big(R_{{\rm SE},i}^0(\varepsilon) J(\varepsilon) +
+   R_{{\rm SE},i}^1(\varepsilon) \ell_\alpha H^\alpha(\varepsilon) \big) \\
   \label{eqn:se_opacity_expansion}
-  \mathscr{K}_{{\rm SE},i} &= 4\pi \frac{\varepsilon^3}{(2\pi)^3}
-  \big(R_0(\varepsilon)+R_1(\varepsilon)\big).
+  \mathscr{K}_{{\rm SE},i}(\varepsilon) &=
+  4\pi \frac{\varepsilon^3}{(2\pi)^3} R_{{\rm SE},i}^0(\varepsilon),
 \end{align}
+where we have made the energy-dependence explicit.
+
+However, we don't know the energy dependence of $J(\varepsilon)$ and
+$H^\alpha(\varepsilon)$,
+since the the truncated moment simulations usually evolve only the integrated
+moments, $J$ and $H^\alpha$.
 

### 2016-11-01 paper.tex
Add relation between emissivities eta and j

@@
   \label{eqn:inv_emissivity}
-  \mathscr{E}_{{\rm AE},i} = \frac{(2\pi)^3}{\varepsilon^2}\eta_{i},
+  \mathscr{E}_{{\rm AE},i}
+  = \varepsilon j_i
+  = \frac{(2\pi)^3}{\varepsilon^2}\eta_i,
 \end{equation}
-where $\eta$ is the traditional emissivity with dimension
-$[\eta]={\rm energy}\,{\rm length}^{-3}\,{\rm time}^{-1}\,{\rm energy}^{-1}\,{\rm steradian}^{-1}$,
-and
+where $j$ is the emissivity having dimension
+$[j]={\rm length}^{-1}\,{\rm steradian}^{-1}$,
+$\eta$ is the emissivity having dimension
+$[\eta]={\rm energy}\,{\rm length}^{-3}\,{\rm time}^{-1}\,{\rm energy}^{-1}\,{\rm steradian}^{-1}$
+(natural to formulations using specific intensity instead of distribution
+functions), and
 \begin{equation}

### 2016-11-23 paper.tex
Rename figure with prefix fig

@@
-\documentclass[aps,prd,twocolumn,superscriptaddress]{revtex4-1}
+\documentclass[aps,floatfix,prd,superscriptaddress,twocolumn]{revtex4-1}
 % Intended to be built with pdflatex, using the REVTEX 4.1 package.
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{affine_parameter_label}
+  \includegraphics[width=\columnwidth]{fig-affine_parameter_label}
   \caption{Affine parameterization of a neutrino trajectory of momentum
@@
 \subsection{Dummy Tests}
+\label{ssec:tests_basic}
 [Maybe we don't need to present these tests, but make sure to establish
 confidence in the basics:
-gravitational redshift, Doppler shift, thermodynamic equilibrium.]
+gravitational redshift, Doppler shift,
+geoesic curvature, thermodynamic equilibrium.]
 \todo{setup and run dummy tests}
@@
 \subsection{Homogenous Absorbing Star}
+\label{ssec:test_ab_star}
 [Test described in~\cite[Sec.~3.2]{smit1997-two_moment}.]
@@
 \subsection{Homogenous Scattering Star}
+\label{ssec:test_sc_star}
 \todo{maybe drop this test}
@@
 \subsection{Post-Bounce Neutron Core}
+\label{ssec:test_bounce_core}
 [Standard radiation transport test described in~\cite{ocon2015-gr1d_with_nu},
@@
 \subsection{Disk Comparison to M1}
+\label{ssec:test_disk_comparison}
 [Compute total and relative number and energy luminosities for a disk evolved

### 2016-11-23 paper.tex
Add figure for homogeneous isotropic test

@@
 
+\begin{figure}
+  \includegraphics[width=\columnwidth]{fig-homogeneous_isotropic}
+  \caption{
+    Equilibrium distribution functions in a homogeneous background at
+    $\varrho=6.22 \times 10^{12}\,{\rm g}\,{\rm cm}^{-3}$,
+    $T=5$~MeV,
+    $Y_e=0.3$,
+    using the LS220 equation of state, in which the equilibrium
+    neutrino chemical potentials are
+    $\mu_{\nu_e}=-\mu_{\bar{\nu}_e}=62.04$~MeV, and 
+    $\mu_{\nu_x}=0$.
+    The solid lines give the analytic Fermi-Dirac distribution,
+    and the points are computed via ray tracing.}
+  \label{fig:homogeneous_isotropic}
+\end{figure}
+
 \subsection{Homogenous Absorbing Star}

### 2016-11-29 paper.tex
Add some section labels and other odds and ends
Move a figure to hack fix the float error
Compiling the paper at the previous commit produces a seemingly
unrelated "undefined control sequence" error at a \todo statement
following the figure. If I insert another \todo statement right
after the figure, the error shifts to earlier in the doc by one
\todo statement. Turning off the \todos (\renewcommand) removes
the errror, but in the meantime, I'll simply move the figure till
it falls on another page. That also removes the error.

Some possibly informative stackexchange dialogs:
http://tex.stacke

@@
 we can integrate Eqn.~\ref{eqn:boltzmann_linear} directly,
-\begin{equation}
+\begin{align}
   \label{eqn:rendering}
-  f(\lambda) = f(\lambda_e) e^{-\tau(\lambda,\lambda_e)} +
+  f(\lambda,\lambda_e) &=
+  f_{\rm bound}(\lambda,\lambda_e) + f_{\rm integ}(\lambda,\lambda_e), \\
+  \label{eqn:rendering_fbound}
+  f_{\rm bound}(\lambda,\lambda_e) &=
+  f(\lambda_e) e^{-\tau(\lambda,\lambda_e)}, \\
+  \label{eqn:rendering_finteg}
+  f_{\rm integ}(\lambda,\lambda_e) &=
   \int_{\lambda_e}^\lambda d\lambda'\,e^{-\tau(\lambda,\lambda')}
   \mathscr{K}^*(\lambda') \mathscr{S}^*(\lambda'),
-\end{equation}
+\end{align}
 where the optical depth is defined,
+\todo{rewrite this section simpler; don't give integral solution
+  since we don't use}
 \begin{equation}
@@
 system described in~\cite{bohn2016-code}.
-We integrate Eqns.~\ref{eqn:geodesic_x} and~\ref{eqn:geodesic_p} in the form given by
-\citep{hugh1994-eh_finding}.
-\todo{describe eqns for $f_{\rm integ}$ and $\tau$}
-\todo{describe $\nu$-specific numerics}
+We integrate Eqns.~\ref{eqn:geodesic_x} and~\ref{eqn:geodesic_p} in the form
+given by \citep{hugh1994-eh_finding},
+and Eqns.~\ref{eqn:rendering} and~\ref{eqn:optical_depth}
+in the form given below.
+By using the time-component of Eqn.~\ref{eqn:geodesic_x} ($dt=d\lambda \, p^t$)
+we may transform our integrations to coordinate time.
+The equations are
+\begin{align}
+  \label{eqn:num_x}
+  \frac{dx^i}{dt} &=
+  g^{ij} \frac{p_j}{p^t} - \beta^{i}\\
+  \label{eqn:num_p}
+  \frac{dp_i}{dt} &=
+  -\alpha \alpha_{,i} p^t
+  + \beta^k_{,i} p_k
+  - \frac{1}{2} g^{jk}_{,i}\frac{p_j p_k}{p^t}\\
+  \label{eqn:num_finteg}
+  \frac{df_{\rm integ}}{dt} &=
+  e^{-\tau} \mathscr{K}^* \mathscr{S}^*\\
+  \label{eqn:num_tau}
+  \frac{d\tau}{dt} &=
+  -\mathscr{K}^*
+\end{align}
+where the comma denotes a partial derivative
+and we have used the standard 3+1 decomposition of the metric
+\begin{equation}
+  \label{eqn:adm_metric}
+  \psi_{\mu\gamma} :=
+  \left(
+  \begin{matrix}
+    -\alpha^2 + \beta^i \beta_i  & \beta_i \\
+    \beta_j                      & g_{ij}
+  \end{matrix}
+  \right).
+\end{equation}
+
+We integrate each ray until we reach a terminal optical depth of
+$\tau_{\rm term}$ at event $x_e^\alpha$.
+We choose $\tau_{\rm term}$ large so that
+$f_{\rm bound}$ is negligible
+(in practice $e^{-\tau_{\rm term}}=0.001$)
+and compute $f_{\rm bound}$ assuming radiative equilibrium with the matter,
+i.e. $f_{\rm bound}=(1+\exp((\varepsilon-\mu_\nu)/T)$,
+and $\mu_{\nu_e}=-\mu_{\bar{\nu}_e}=\mu_{e^{-}}+\mu_p-\mu_n$ and
+$\mu_{\nu_x}=0$.
+If the ray terminates off the numerical domain
+we set $f_{\rm bound}=0$.
+
+Since we don't know the emission event a priori,
+\todo{make clear ``emission event'' doesn't involve physical assumption}
+we follow the integration backwards in time,
+from $t_o$ to $t_e$.
+We begin each integration by setting initial values for the variables at $t_o$.
+The observer specifies $x^i_o$ and $p_{i,o}$,
+and we set $f_{{\rm integ},o}=0$ and $\tau_o=0$.
+
+We integrate Eqns.~\ref{eqn:num_x}-\ref{eqn:num_tau} with adaptive step sizes,
+using the 5th-order Dormand-Prince algorithm \cite{pres2007-nr_3rd_ed}.
+After a step is taken, the error is estimated for each variable.
+If the maximum of these errors exceeds some absolute or relative threshold
+the step size is decreased and the step recomputed,
+if it is below some threshold the next step size is increased.
+% see OdeIntegrators::AdaptiveDense
+% and OdeSteppers::DormandPrince5
+% and OdeControllers::ProportionalIntegral_Base::StepSizeToAttempt
+% and OdeErrorMeasures::ScaledAbsRel
+\todo{specify}
+
+The volume data are components of the metric $\psi_{\alpha\beta}$,
+its derivatives $\psi_{\alpha\beta,\gamma}$,
+and the hydrodynamical data $W$, $u_i$, $\varrho$, $T$, and $Y_e$.
+These fields are stored in spectral representation.
+If they are computed via hydrodynamical simulation, they are saved to disk
+at specified coordinate times and interpolated
+with spectral interpolation in space,
+and 1st-order polynomial interpolation in time
+(as described in \cite[App.~B]{bohn2016-code}).
+If computed from a stationary solution to the
+general relativistic hydrodynamics equations,
+a single time is stored to disk, and no time interpolation is needed.
 
@@
 \todo{setup and run dummy tests}
-
-\begin{figure}
-  \includegraphics[width=\columnwidth]{fig-homogeneous_isotropic}
-  \caption{
-    Equilibrium distribution functions in a homogeneous background at
-    $\varrho=6.22 \times 10^{12}\,{\rm g}\,{\rm cm}^{-3}$,
-    $T=5$~MeV,
-    $Y_e=0.3$,
-    using the LS220 equation of state, in which the equilibrium
-    neutrino chemical potentials are
-    $\mu_{\nu_e}=-\mu_{\bar{\nu}_e}=62.04$~MeV, and 
-    $\mu_{\nu_x}=0$.
-    The solid lines give the analytic Fermi-Dirac distribution,
-    and the points are computed via ray tracing.}
-  \label{fig:homogeneous_isotropic}
-\end{figure}
 
@@
 \todo{compute luminosities from ray tracing}
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{fig-homogeneous_isotropic}
+  \caption{
+    Equilibrium distribution functions in a homogeneous background at
+    $\varrho=6.22 \times 10^{12}\,{\rm g}\,{\rm cm}^{-3}$,
+    $T=5$~MeV,
+    $Y_e=0.3$,
+    using the LS220 equation of state, in which the equilibrium
+    neutrino chemical potentials are
+    $\mu_{\nu_e}=-\mu_{\bar{\nu}_e}=62.04$~MeV, and 
+    $\mu_{\nu_x}=0$.
+    The solid lines give the analytic Fermi-Dirac distribution,
+    and the points are computed via ray tracing.}
+  \label{fig:homogeneous_isotropic}
+\end{figure}
 

### 2016-11-30 paper.tex
Define num density & flux and Eulerian observer

@@
 \todo{consistent with Eqn.~\ref{eqn:def_direction}?}
+We will also make use of the number density and flux defined as
+$G(\varepsilon)=J(\varepsilon)/\varepsilon$
+and $K^\mu(\varepsilon)=H^\mu(\varepsilon)/\varepsilon$
+respectively.
 
@@
 having dimension ${\rm energy}\,{\rm length}^{-3}$.
+
+In practice we use two possible fiducial observers to define the momentum
+decomposition via Eqn.~\ref{eqn:def_momentum}: the fluid, or comoving, observer,
+$u^\alpha=dx^\alpha/ds$, and the Eulerian, or normal, observer,
+$n_\alpha=-\alpha \,\partial_\alpha t$.
+Here $ds$ is a differential of proper time, $t$ is coordinate time,
+and $\alpha$ is lapse.
+We distinguish moments computed assuming an Eulerian observer with a tilde,
+e.g. $\tilde{J}$, $\tilde{H}^\mu$, $\tilde{S}^{\mu\nu}$.
+Note that these three Eulerian moments are identical to the lab-frame moments
+$E$, $F^\mu$, and $P^{\mu\nu}$
+defined in
+\cite{shib2011-truncated_moment, ocon2015-gr1d_with_nu, fouc2015-m1_nsbh}.
 

### 2016-11-30 paper.tex
Add spectral assumptions to get J(eps) from J

@@
 which allows us to write Eqns.~\ref{eqn:se_factor} and \ref{eqn:se_opacity} as
+\todo{drop anisotropic term and defend}
 \begin{align}
@@
 However, we don't know the energy dependence of $J(\varepsilon)$ and
-$H^\alpha(\varepsilon)$,
-since the the truncated moment simulations usually evolve only the integrated
-moments, $J$ and $H^\alpha$.
+$H^\alpha(\varepsilon)$ in general,
+since simulations usually employ a gray moment scheme,
+evolving only the energy-integrated moments, $J$ and $H^\alpha$
+and sometimes $G$.
+Nevertheless, we may reconstruct these energy-dependent source terms
+from the gray moments by assuming a diluted Fermi-Dirac spectrum.
+
+We use the following procedure:
+\begin{enumerate}
+\item
+  Interpolate the equilibrium neutrino chemical potential, $\mu_\nu$,
+  from the simulation grid.
+\item
+  Interpolate $J$ and $G$ from the simulation grid and
+  compute the average neutrino energy in the fluid frame
+  $\langle \varepsilon \rangle \equiv J/G$.
+\item
+  Compute the neutrino spectral temperature
+  \begin{equation}
+    T_\nu = \langle \varepsilon \rangle
+    \frac{\mathscr{F}_3(\mu_\nu/T)}{\mathscr{F}_2(\mu_\nu/T)},
+  \end{equation}
+  where $\mathscr{F}_j$ is the Fermi integral
+  \begin{equation}
+    \mathscr{F}_j(\eta) = \int_0^\infty dx \, x^j (1+e^{x-\eta})^{-1}.
+  \end{equation}
+\item
+  Assume the background neutrino fields have the same total density
+  as the evolved moments
+  \begin{align}
+    J(\varepsilon) &=
+    \frac{J}{\mathscr{F}_3(\mu_\nu/T)}
+    (1+\exp(\varepsilon/T_\nu-\mu_\nu/T)^{-1}, \\
+    H^\mu(\varepsilon) &=
+    \frac{H^\mu}{\mathscr{F}_3(\mu_\nu/T)}
+    (1+\exp(\varepsilon/T_\nu-\mu_\nu/T)^{-1}.
+  \end{align}
+\end{enumerate}
 

### 2016-11-30 paper.tex
Correct some equation errors and clarify fiducial observer discussion

@@
 In practice we use two possible fiducial observers to define the momentum
-decomposition via Eqn.~\ref{eqn:def_momentum}: the fluid, or comoving, observer,
-$u^\alpha=dx^\alpha/ds$, and the Eulerian, or normal, observer,
-$n_\alpha=-\alpha \,\partial_\alpha t$.
-Here $ds$ is a differential of proper time, $t$ is coordinate time,
-and $\alpha$ is lapse.
+decomposition via Eqn.~\ref{eqn:def_momentum}: the Eulerian,
+or normal, observer, $n_\alpha=-\alpha \,\partial_\alpha t$,
+and the fluid, or comoving, observer,
+$u^\alpha=W(n^\alpha+v^\alpha)$.
+Here $t$ is coordinate time, $\alpha$ is lapse,
+$W$ is fluid Lorentz factor, and $v^\alpha$ is fluid coordinate velocity.
 We distinguish moments computed assuming an Eulerian observer with a tilde,
@@
     T_\nu = \langle \varepsilon \rangle
-    \frac{\mathscr{F}_3(\mu_\nu/T)}{\mathscr{F}_2(\mu_\nu/T)},
+    \frac{\mathscr{F}_2(\mu_\nu/T)}{\mathscr{F}_3(\mu_\nu/T)},
   \end{equation}
@@
     J(\varepsilon) &=
-    \frac{J}{\mathscr{F}_3(\mu_\nu/T)}
+    J \frac{\varepsilon^3}{T_\nu^4 \,\mathscr{F}_3(\mu_\nu/T)}
     (1+\exp(\varepsilon/T_\nu-\mu_\nu/T)^{-1}, \\
     H^\mu(\varepsilon) &=
-    \frac{H^\mu}{\mathscr{F}_3(\mu_\nu/T)}
+    H^\mu \frac{\varepsilon^3}{T_\nu^4\, \mathscr{F}_3(\mu_\nu/T)}
     (1+\exp(\varepsilon/T_\nu-\mu_\nu/T)^{-1}.

### 2017-01-20 paper.tex
Visually distinguish Fermi index from coordinate index
Add explanation of background field assumption (3)
Reorder/reword the background field recipe steps

@@
 \item
-  Interpolate the equilibrium neutrino chemical potential, $\mu_\nu$,
-  from the simulation grid.
+  Interpolate the fluid temperature, $T$,
+  equilibrium neutrino chemical potential, $\mu_\nu$,
+  and evolved neutrino energy and number densities in the fluid frame,
+  $J$ and $G$, from the simulation grid.
 \item
-  Interpolate $J$ and $G$ from the simulation grid and
-  compute the average neutrino energy in the fluid frame
+  Compute the average neutrino energy in the fluid frame
   $\langle \varepsilon \rangle \equiv J/G$.
 \item
-  Compute the neutrino spectral temperature
+  Compute the neutrino spectral temperature, assuming equilibrium with the
+  fluid
   \begin{equation}
@@
   \end{equation}
-  where $\mathscr{F}_j$ is the Fermi integral
+  where $\mathscr{F}_b$ is the Fermi integral
   \begin{equation}
-    \mathscr{F}_j(\eta) = \int_0^\infty dx \, x^j (1+e^{x-\eta})^{-1}.
+    \mathscr{F}_b(\eta) = \int_0^\infty dx \, x^b (1+e^{x-\eta})^{-1}.
   \end{equation}

### 2017-02-13 paper.tex
Reorder authors
Add acknowledgements

@@
 
+\author{Evan O'Connor}
+\affiliation{Department of Physics,
+  North Carolina State University, Raleigh, NC 27695, USA}
+
+\author{Andy Bohn}
+\affiliation{Center for Radiophysics and Space Research,
+  Cornell University, Ithaca, NY 14853, USA}
+
 \author{Jerred Jesse}
@@
 
-\author{Evan O'Connor}
-\affiliation{Department of Physics,
-  North Carolina State University, Raleigh, NC 27695, USA}
-
 \author{Francois Foucart}
@@
   1 Cyclotron Rd., Berkeley, CA 94720, USA}
-
-\author{Andy Bohn}
-\affiliation{Center for Radiophysics and Space Research,
-  Cornell University, Ithaca, NY 14853, USA}
 
@@
 
+\subsubsection*{Acknowledgements}
+This research is supported in part by the Joint Institute for Nuclear
+Astrophysics---Center for the Evolution of the Elements (JINA-CEE)
+and the National Science Fondation under grant numbers X. \todo{grant nos?}
+We thank Sherwood Richers for providing his neutrino transport code
+\lstinline{Sedonu}~\footnote{\url{https://bitbucket.org/srichers/sedonu}}
+for comparisons.
+We thank Luke Roberts and Yonglin Zhu for helpful discussions.
+
 % ******************************************************************************

### 2017-05-26 paper.tex
Add Hebert to acknowledgements
Update paper structure to reflect current project directions
* remove sections related to closure relation (not going to do this now)
* add new test sections

@@
 % Intended to be built with pdflatex, using the REVTEX 4.1 package.
-% Template taken from FermiLab APS Jouranl template.
+% Template taken from FermiLab APS Journal template.
 
@@
   nucleons by incorporating estimates of the background neutrino fields.
-  These background fields may be taken from a low-order moment simulation,
-  be supplied analytically in simple cases,
+  These background fields may be taken from a low-order moment simulation
   or be ignored entirely, in which case the method
@@
   apply it to simulations of neutron star mergers to
-  1) compute neutrino fluxes as a function of energy and emission angle, and
-  2) compute the radiation pressure tensor in the funnel of the disk,
-  demonstrating a way to improve the closure relations used in the truncated
-  moment formalism.
+  compute neutrino fluxes as a function of energy and emission angle.
 \end{abstract}
@@
 In Sec.~\ref{sec:tests} we present some tests of the code.
-In Sec.~\ref{sec:applications} we present two applications:
-neutrino spectra as a function of observer's angle, and
-a variable closure relation for a truncated moment formalism.
-We present both applications in the dynamical environments following the merger
-of a neutron star--neutron star and a black hole--neutron star binary.
+In Sec.~\ref{sec:applications} we present an application:
+computing neutrino spectra as a function of observer's angle
+in the dynamical environment following the merger of two neutron stars.
 
@@
 
-\subsection{Dummy Tests}
-\label{ssec:tests_basic}
-[Maybe we don't need to present these tests, but make sure to establish
-confidence in the basics:
-gravitational redshift, Doppler shift,
-geoesic curvature, thermodynamic equilibrium.]
-\todo{setup and run dummy tests}
+\subsection{Homogeneous Equilibrium}
+\label{ssec:test_equilibrium}
+Homogeneous matter, optically thick. Plot in Fig.~\ref{fig:homogeneous_isotropic}.
+\todo{describe test}
+
+\subsection{Doppler Shift}
+\label{ssec:test_doppler}
+Homogeneous matter moving with v in +z direction. Stationary observer.
+\todo{add plot of $\varepsilon(\mu)$}
+
+\subsection{Gravitational Redshift and Geodesic Curvature}
+\label{ssec:test_gravity}
+Optically thick star in Schwarszchild.
+todo{describe $L(r)$, as a function of lapse}
+optically thick disk in Kerr (Fishbone-Moncrief disk):
+set up to give two different surfaces for nue vs nua, but same temp (?).
+\todo{plot of analytic and numerical geodesics}
+\todo{plot of $G$ nue vs nua}
 
@@
 [Test described in~\cite[Sec.~3.2]{smit1997-two_moment}.]
-\todo{Jerred: run test}
+\todo{run test for optically thick and thin}
 
-\subsection{Homogenous Scattering Star}
-\label{ssec:test_sc_star}
-\todo{maybe drop this test}
-[Test presented in~\cite{humm1971-grey_transfer}
-and described in~\cite[Sec.~9.1.1]{abdi2012-monte_carlo}.
-Our version would be a point source embedded in a spherical distribution
-of purely elastic-scattering material. It requires an analytical estimate
-of $J$.]
-\todo{formulate analytic $J$.}
-
-\subsection{Post-Bounce Neutron Core}
-\label{ssec:test_bounce_core}
+\subsection{Massive Collapsed Star with Scattering Envelope}
+\label{ssec:test_scat_envelope}
 [Standard radiation transport test described in~\cite{ocon2015-gr1d_with_nu},
@@
 and in~\cite{abdi2012-monte_carlo}.]
-\todo{Evan: provide initial data and $H^r(\varepsilon)$}
-\todo{Jerred: setup and run test}
+\todo{add plots of $\langle \varepsilon \rangle(\mu)$}
 
@@
 [Compute total and relative number and energy luminosities for a disk evolved
-in one of~\cite{fouc2015-m1_nsbh, fouc2016-m1_nsns, fouc2016-m1_evolve_n}.
+in one of \cite{fouc2015-m1_nsbh, fouc2016-m1_nsns, fouc2016-m1_evolve_n}.
 Expect agreement within factors of a few (?).]
-\todo{code up moment integrals}
-\todo{compute luminosities from M1}
-\todo{compute luminosities from ray tracing}
+\todo{compute luminosities from M1 and RT, present table}
 
@@
   \includegraphics[width=\columnwidth]{fig-homogeneous_isotropic}
-  \caption{
-    Equilibrium distribution functions in a homogeneous background at
+  \caption{Equilibrium distribution functions in a homogeneous background at
     $\varrho=6.22 \times 10^{12}\,{\rm g}\,{\rm cm}^{-3}$,
@@
     The solid lines give the analytic Fermi-Dirac distribution,
-    and the points are computed via ray tracing.}
+    and the points are computed via ray tracing.
+    Described in Sec.~\ref{ssec:test_equilibrium}}
   \label{fig:homogeneous_isotropic}
@@
 The merger models are described in \cite{fouc2015-m1_nsbh, fouc2016-m1_nsns}.
-[Something like this is presented in~\cite[Fig.~10-11]{pere2014-nu_wind};
+Something like this is presented in \cite[Figs.~10-11]{pere2014-nu_wind};
 compare.]
-\todo{code up moment integrals}
-\todo{compute $H^r(\theta;\varepsilon)$ in nsns and nsbh}
-\todo{maybe compute at several times}
-
-\subsection{Closure Relation}
-\label{ssec:closure}
-\todo{figure out your approach}
-[This is a little tricky. Unless we can come up with a new improved analytic
-closure relation and use ray tracing to evaluate its trustworthiness,
-it's going to be hard to get anything useful from computing
-$S^{\alpha\beta}$.]
-
-[Maybe the variable Eddington closure is more suitable to this application.
-If I understand it right, applying ray tracing to that method would require
-computing $S^{\alpha\beta}$ on the fly throughout the spatial volume.
-That would certainly be doable for a quasi-stationary disk configuration.]
-\todo{compute $S^{\alpha\beta}$ and evaluate its effect on evolution}
+\todo{compute $H^r(\mu;\varepsilon)$ in nsns and nsbh}
+\todo{compute $\langle \varepsilon \rangle(\mu)$}
 
@@
   Compute the average neutrino energy in the fluid frame
-  $\langle \varepsilon \rangle \equiv J/G$.
+  \begin{equation}
+    \langle \varepsilon \rangle \equiv J/G.
+  \end{equation}
 \item
@@
 We thank Luke Roberts and Yonglin Zhu for helpful discussions.
+We are indebted to Fran\c{c}ois H\'{e}bert for his timely help in finding and
+correcting bugs in the time-interpolation algorithm.
 

### 2017-06-04 paper.tex
Drop Neutrinosurface Approximation section

@@
 
-\subsection{Neutrinosurface Approximation}
-\label{ssec:neutrinosurface}
-\todo{derive $\tau=2/3$ approx \cite[Chap.~4]{shu1991-book_radiation},
-  and \cite{miln1921-radiative_equil_1}}
-In the simplified configuration of a plane-parallel atmosphere, the
-rendering equation reduces to 
-\begin{equation}
-\label{eqn:neutrinosurface}
-  f(\lambda) = f^{\rm eq}(\lambda_e),
-\end{equation}
-where $\tau(\lambda,\lambda_e)=2/3$.
-At an optical depth of $2/3$, the energy flux is equivalent to the
-blackbody value $\sigma T^4$, with $\sigma$, the Stefan-Boltzmann radiation
-constant.
-\todo{verify}
-
 \subsection{Moments of the Distribution Function}
@@
 Optically thick star in Schwarszchild.
-todo{describe $L(r)$, as a function of lapse}
+\todo{describe $L(r)$, as a function of lapse}
 optically thick disk in Kerr (Fishbone-Moncrief disk):

### 2017-06-04 paper.tex
Define the emissivity and opacity constituents earlier

@@
 antiparticle at different momenta, $f'$ and $\bar{f}'$.
-For clarity we symbolize all of these dependencies with the shorthand $C[f]$.
+For simplicity we symbolize all of these dependencies with the shorthand $C[f]$.
 The various neutrino interactions contributing to $C[f]$ are detailed in
@@
 and for simplicity we have suppressed the arguments of $f$.
-These coefficients are derived by considering Fermi-blocking and
+These coefficients are computed by considering Fermi-blocking and
 dependencies on neutrino and antineutrino distribution functions at other
 momenta.
-They are derived in App.~\ref{sec:source_terms}.
+We provide a detailed derivation for these terms in App.~\ref{sec:source_terms},
+for the two classes of interactions we consider in this work,
+absorption/emission and elastic scattering
+(Tab.~\ref{tab:neutrino_processes}). Within the scope of this work, we have
+\begin{align}
+  \label{eqn:inv_emissivity_components}
+  \mathscr{E}^* &= \mathscr{E}_{\rm AE} + \mathscr{E}_{\rm SE},\\
+  \label{eqn:inv_opacity_components}
+  \mathscr{K}^* &= \mathscr{K}_{\rm AE} + \mathscr{K}_{\rm SE}.
+\end{align}
+
+For absorption/emission interactions (App.~\ref{ssec:sources_ae}) these are
+simply the sums over the emissivities and absorptive opacities for each
+reaction $i$:
+\begin{align}
+  \label{eqn:ae_emissivity_summed}
+  \mathscr{E}_{\rm AE}(\varepsilon)
+  &= \frac{(2\pi)^3}{\varepsilon^2} \sum_{i\,{\rm reactions}} \eta_i(\varepsilon), \\
+  \label{eqn:ae_opacity_summed}
+  \mathscr{K}_{\rm AE}(\varepsilon)
+  &= \varepsilon \sum_{i\,{\rm reactions}} \chi_{{\rm a},i}(\varepsilon).
+\end{align}
+
+For elastic scattering interactions (App.~\ref{ssec:sources_se}) these are
+the background fields and the sum over the scattering opacities for each
+reaction $i$:
+\begin{align}
+  \label{eqn:se_emissivity_summed}
+  \mathscr{E}_{\rm SE}(\varepsilon)
+  &= \mathscr{K}_{\rm SE}(\varepsilon) \, \Phi_{\rm SE}(\varepsilon) \, \\
+  \label{eqn:se_opacity_summed}
+  \mathscr{K}_{\rm SE}(\varepsilon)
+  &= \varepsilon \sum_{i\,{\rm reactions}} \chi_{{\rm s},i}(\varepsilon).
+\end{align}
+
+This paper introduces the use of neutrino densities and fluxes evolved in a
+truncated moment simulation to estimate the background field, $\Phi_{\rm SE}$.
+App.~\ref{ssec:sources_se} details a method to calculate
+$\Phi_{\rm SE}(\varepsilon)$ in two different ways:
+\begin{itemize}
+\item
+  the \emph{spectral} method using densities and fluxes
+  extracted from a simulation evolved over multiple energy groups
+  to compute the background field with Eqn.~\ref{eqn:se_factor_expansion},
+\item
+  the \emph{gray} method using energy-integrated densities and fluxes
+  extracted from a gray simulation and approximating the
+  energy dependence with Eqns.~\ref{eqn:J_from_gray} and
+  \ref{eqn:H_from_gray}.
+\end{itemize}
 
@@
 where we have made the energy-dependence explicit.
+When we get $J(\varepsilon)$ and $H^\alpha(\varepsilon)$ from a multi-group
+moment evolution, we call this the ``spectral'' method.
 
-However, we don't know the energy dependence of $J(\varepsilon)$ and
-$H^\alpha(\varepsilon)$ in general,
-since simulations usually employ a gray moment scheme,
+However moment evolutions with multiple energy groups are still rare.
+Most simulations employ a gray moment scheme,
 evolving only the energy-integrated moments, $J$ and $H^\alpha$
-and sometimes $G$.
-Nevertheless, we may reconstruct these energy-dependent source terms
+and sometimes the number density $G$.
+We may reconstruct the energy-dependent source terms
 from the gray moments by assuming a diluted Fermi-Dirac spectrum.
@@
   \begin{align}
+    \label{eqn:J_from_gray}
     J(\varepsilon) &=
@@
     (1+\exp(\varepsilon/T_\nu-\mu_\nu/T)^{-1}, \\
+    \label{eqn:H_from_gray}
     H^\mu(\varepsilon) &=
@@
 \end{enumerate}
+When we use these moments in Eqn.~\ref{eqn:se_factor_expansion},
+we call this the ``gray'' method.
 

### 2017-06-04 paper.tex
Add doppler shift test

@@
 Homogeneous matter moving with v in +z direction. Stationary observer.
-\todo{add plot of $\varepsilon(\mu)$}
+Plot in Fig.~\ref{fig:avg_eps_doppler}.
+\todo{describe test}
 
@@
   \label{fig:homogeneous_isotropic}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{fig-moving_fluid-avg_eps_vs_cosA}
+  \caption{Average energy $\langle\varepsilon\rangle=J/G$
+    of neutrino fields for fluid moving in the $+z$ direction,
+    with $A$ the angle between the neutrino momentum and the $+z$ direction.
+    Test described in Sec.~\ref{ssec:test_doppler}}
+  \label{fig:avg_eps_doppler}
 \end{figure}

### 2017-06-05 paper.tex
Add collapse profile and absorption sphere tests

@@
 [Test described in~\cite[Sec.~3.2]{smit1997-two_moment}.]
-\todo{run test for optically thick and thin}
+Plot in Fig.~\ref{fig:f_absorption_sphere}.
 
 \subsection{Massive Collapsed Star with Scattering Envelope}
-\label{ssec:test_scat_envelope}
+\label{ssec:test_collapse}
 [Standard radiation transport test described in~\cite{ocon2015-gr1d_with_nu},
@@
 and in~\cite{abdi2012-monte_carlo}.]
-\todo{add plots of $\langle \varepsilon \rangle(\mu)$}
+Plot in Fig.~\ref{fig:avg_eps_collapse}.
 
@@
 \label{sec:applications}
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{fig-absorption_sphere}
+  \caption{Distribution function outside a constant-density star
+    with radius $R=15$~km, and observer at $r=22.5$~km.
+    $A$ is the angle between the neutrino momentum and the $\hat{r}$ direction.
+    The sample grid uses 30 points in energy and angle spanning the ranges
+    $\varepsilon\in(0,100)$~MeV and $\cos A\in(0.6454,1)$,
+    with sample points at bin centers.
+    The line depicts the analytic result for one energy: $\varepsilon=15$~MeV.
+    Test described in Sec.~\ref{ssec:test_ab_star}}
+  \label{fig:f_absorption_sphere}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{fig-collapse_profile-avg_eps_vs_cosA}
+  \caption{Average energy $\langle\varepsilon\rangle=J/G$
+    in the collapse profile test.
+    The observer is at 500~km, and the shock at 150~km,
+    so that the half-opening angle of the shock is $\cos A=0.954$.
+    The four methods depicted are 1) a fiducial Monte Carlo calculation,
+    ray tracing using the 2) spectral and 3) gray methods to estimate background
+    fields for scattering, and 4) ray tracing neglecting scattering.
+    Test described in Sec.~\ref{ssec:test_collapse}}
+  \label{fig:avg_eps_collapse}
+\end{figure}
 

### 2017-06-06 paper.tex
Update Homogeneous Equilibrium test

@@
 
-We integrate Eqns.~\ref{eqn:num_x}-\ref{eqn:num_tau} with adaptive step sizes,
+We integrate Eqns.~\ref{eqn:num_x}--\ref{eqn:num_tau} with adaptive step sizes,
 using the 5th-order Dormand-Prince algorithm \cite{pres2007-nr_3rd_ed}.
@@
 
+To test our algorithm, we integrate the ray tracing equations
+(Eqns.~\ref{eqn:num_x}--\ref{eqn:num_tau}) in the following configurations,
+and compare $f$ or moments of $f$ to known analytic solutions, or in some
+cases to fiducial numerical calculations.
+
 \subsection{Homogeneous Equilibrium}
 \label{ssec:test_equilibrium}
-Homogeneous matter, optically thick. Plot in Fig.~\ref{fig:homogeneous_isotropic}.
-\todo{describe test}
+In an optically thick region the radiation field is in thermodynamic equilibrium
+with the matter.
+We set up a large slab of matter in two homogeneous configurations
+representative of the fluid in the Massive Collapsed Star test
+(Sec.~\ref{ssec:test_collapse}) at two different radii:
+at 50~km
+For comparision with Collapsed Star test, we use the LS180 equation of state,
+in which the equilibrium
+neutrino chemical potentials are
+$\mu_{\nu_e}=-0.639$~MeV, and 
+$\mu_{\nu_e}=5.771$~MeV respectively
+($\mu_{\bar{\nu}_e}=-\mu_{\nu_e}$ and $\mu_{\nu_x}=0$).
+Plot in Fig.~\ref{fig:homogeneous_isotropic}.
+\todo{run test at this thermodynamic point}
 
@@
 \label{ssec:test_doppler}
-Homogeneous matter moving with v in +z direction. Stationary observer.
+We reproduce the test above, but with the matter moving in
+in the $+z$ direction, with the observer stationary.
 Plot in Fig.~\ref{fig:avg_eps_doppler}.
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{fig-homogeneous_isotropic}
-  \caption{Equilibrium distribution functions in a homogeneous background at
-    $\varrho=6.22 \times 10^{12}\,{\rm g}\,{\rm cm}^{-3}$,
-    $T=5$~MeV,
-    $Y_e=0.3$,
-    using the LS220 equation of state, in which the equilibrium
-    neutrino chemical potentials are
-    $\mu_{\nu_e}=-\mu_{\bar{\nu}_e}=62.04$~MeV, and 
-    $\mu_{\nu_x}=0$.
-    The solid lines give the analytic Fermi-Dirac distribution,
-    and the points are computed via ray tracing.
+  \includegraphics[width=\columnwidth]{fig-homogeneous_equilibrium}
+  \caption{Equilibrium distribution functions in a homogeneous background
+    at two thermodynamic points,
+    representative of the collapse profile at
+    50~km ($\varrho=1.235 \times 10^{11}\,{\rm g}\,{\rm cm}^{-3}$,
+    $T=4$~MeV,
+    $Y_e=0.12$) and
+    200~km ($\varrho=3.026\times 10^{7}\,{\rm g}\,{\rm cm}^{-3}$,
+    $T=0.5$~MeV,
+    $Y_e=0.5$).
     Described in Sec.~\ref{ssec:test_equilibrium}}

### 2017-06-06 paper.tex
Rearrange test section: narrative of hierarchy of physical realism

@@
 \todo{discuss mc more clearly}
-The state of the art today employs a truncated moment formalism
+The state of the art today couples radiation and matter with a truncated
+moment formalism
 \citep{shib2011-truncated_moment, fouc2015-m1_nsbh, fouc2016-m1_nsns,
@@
 of the radiation field (the total energy).
+We refer to this as M1 transport.
 Recently, \cite{fouc2016-m1_evolve_n} have expanded their code to also evolve
@@
   vaan2016-uncovering_mnr},
-3) improve closure relations used in truncated moment schemes
+3) improve closure relations used in M1 transport schemes
 \citep{ramp2002-truncated_moment, shib2011-truncated_moment,
@@
 lowest-order moment of the neutrino distribution function
-computed in a truncated moment scheme (formulated by
+computed in an M1 transport simulation (formulated by
 \cite{thor1981-truncated_moment, shib2011-truncated_moment}
@@
 If moments are not available either from
-a truncated moment evolution or a trustworthy analytical estimate,
+an M1 transport evolution or a trustworthy analytical estimate,
 our method reduces to the current state-of-the-art ray tracing method,
@@
 
-This paper introduces the use of neutrino densities and fluxes evolved in a
-truncated moment simulation to estimate the background field, $\Phi_{\rm SE}$.
+This paper introduces the use of neutrino densities and fluxes evolved in an
+M1 transport simulation to estimate the background field, $\Phi_{\rm SE}$.
 App.~\ref{ssec:sources_se} details a method to calculate
@@
 To test our algorithm, we integrate the ray tracing equations
-(Eqns.~\ref{eqn:num_x}--\ref{eqn:num_tau}) in the following configurations,
-and compare $f$ or moments of $f$ to known analytic solutions, or in some
-cases to fiducial numerical calculations.
+(Eqns.~\ref{eqn:num_x}--\ref{eqn:num_tau}) at one observation position
+in the following configurations. This suite of configurations
+may be understood as a hierarchy of models of increasing
+physical realism: beginning with a homogeneous medium of effectively
+infinite extent (i.e.\ optically thick) and progressing to a model of a
+1D pre-supernova, post-bounce collapse profile
+evolved using an M1 transport hydrodynamical simulation.
+\todo{also nsns merger?}
 
-\subsection{Homogeneous Equilibrium}
+We compare to analytic solutions in the simple configurations, and
+numerical solutions from alternative radiation transport algorithms
+(i.e.\ monte carlo or M1 transport) where they are not.
+
+\subsection{Infinite Homogeneous Slab:
+  Testing Thermodynamic Equilibrium}
 \label{ssec:test_equilibrium}
@@
 
-\subsection{Doppler Shift}
+\subsection{Infinite Homogeneous Moving Slab:
+  Testing Doppler Shift}
 \label{ssec:test_doppler}
@@
 
-\subsection{Gravitational Redshift and Geodesic Curvature}
+\subsection{Idealized Star:
+  Testing the Decoupling Regime}
+\label{ssec:test_ab_star}
+[Test described in~\cite[Sec.~3.2]{smit1997-two_moment}.]
+Plot in Fig.~\ref{fig:f_absorption_sphere}.
+
+\subsection{Idealized Compact Star:
+  Testing Gravitational Redshift and Geodesic Curvature}
 \label{ssec:test_gravity}
@@
 
-\subsection{Homogenous Absorbing Star}
-\label{ssec:test_ab_star}
-[Test described in~\cite[Sec.~3.2]{smit1997-two_moment}.]
-Plot in Fig.~\ref{fig:f_absorption_sphere}.
-
-\subsection{Massive Collapsed Star with Scattering Envelope}
+\subsection{Post-Bounce Collapse Profile:
+  Testing Scattering}
 \label{ssec:test_collapse}
@@
 
-\subsection{Disk Comparison to M1}
+\subsection{Hypermassive Neutron Star Remnant:
+  Testing Everything}
 \label{ssec:test_disk_comparison}

### 2017-06-29 paper.tex
Fix lead author email footnote
Add subsections to Applications section

@@
 \author{M.\ Brett Deaton}
+\email{mbdeaton@ncsu.edu}
 \affiliation{Joint Institute for Nuclear Astrophysics,
@@
   North Carolina State University, Raleigh, NC 27695, USA}
-\email{mbdeaton@ncsu.edu}
 
@@
 
-\subsection{Emission Spectra}
+\subsection{Understanding Emission Spectra}
 \label{ssec:spectra}
+In this section, we use ray tracing's ability to track ``where the neutrinos
+are coming from'' to understand the features in the emission distributions
+presented in the code tests.
+
 The merger models are described in \cite{fouc2015-m1_nsbh, fouc2016-m1_nsns}.
@@
 \todo{compute $\langle \varepsilon \rangle(\mu)$}
+
+\subsection{Neutrino Self-Interaction Potential}
+\label{ssec:V_nunu}
+In this section we present $V_{\nu\nu,0}$ along one test trajectory in the NSNS
+model, and show that including elastic scattering makes a qualitative
+difference in the character of the flavor evolution.
+
+\subsection{Neutrino Annihilation Integral}
+\label{ssec:q_nunu}
+In this section we calculate $q_{\nu\bar{\nu}}$ at a representative point
+in the funnel of the NSNS configuration and show that including scattering makes
+a difference (or not) in our estimate of the power available to drive a jet.
 

### 2017-06-29 paper.tex
Update and simplify all the text

@@
   sources.
-  Our formulation treats elastic scattering of neutrinos (changing direction
-  but not energy) off of nuclei and
-  nucleons by incorporating estimates of the background neutrino fields.
+  Our formulation treats the very important effect of
+  elastic scattering of neutrinos off of nuclei and nucleons
+  (changing the neutrino's direction but not energy)
+  by incorporating estimates of the background neutrino fields.
   These background fields may be taken from a low-order moment simulation
@@
 Leakage approximations
-\citep{deat2013-leakage, pere2016-asl, radi2016-dynamical, ?}
+\citep{deat2013-leakage, pere2016-asl, radi2016-dynamical}
+\todo{cite others}
 capture the qualitative effects of neutrinos on the matter, but
-produce limited information about the neutrino field itself.
+provide extremely limited information about the neutrino field itself.
+\todo{discuss mc more clearly}
 Monte Carlo methods \citep{abdi2012-monte_carlo, rich2015-monte_carlo}
-suffer from noise.
-\todo{discuss mc more clearly}
+are an excellent tool require large computational resources because of
+their global treatment of the problem.
+
 The state of the art today couples radiation and matter with a truncated
@@
 \citep{shib2011-truncated_moment, fouc2015-m1_nsbh, fouc2016-m1_nsns,
-  just2015-m1_code, ?},
+  just2015-m1_code},
+\todo{cite others}
 evolving the zeroth- and first-angular moments of the distribution function
@@
 of the radiation field (the total energy).
-We refer to this as M1 transport.
-Recently, \cite{fouc2016-m1_evolve_n} have expanded their code to also evolve
-the number densities of neutrinos, providing an estimate of the average neutrino
-energies in addition to their totals.
-But with only two angular moments and one energy moment, the distribution
-function has very limited angular and spectral information.
+This method, commonly called M1 transport, was formulated by
+\cite{thor1981-truncated_moment, shib2011-truncated_moment}
+and has been applied for example in \cite{ramp2002-truncated_moment,
+  card2013-truncated_moment, fouc2015-m1_nsbh, ocon2015-gr1d_with_nu}.
+Recently, \cite{fouc2016-m1_evolve_n} have expanded their M1 transport code
+to also evolve the number densities of neutrinos,
+providing total and average energies throughout the simulated volume.
+Even so, with M1 transport codes only evolving two angular moments
+and one or two energy moments,
+we can extract only limited angular and spectral information
+from the neutrino fields.
 
@@
 In this work we present a ray tracing method to compute neutrino distribution
-functions from state-of-the-art general relativistic radiation
+functions from state-of-the-art general relativistic M1 transport
 hydrodynamics simulations.
@@
 $f^{\nu_\sigma}(x_o^\alpha;p_\beta)$, or the amount of neutrino radiation
-with momentum $p_\beta$ impinging on $x_o^\alpha$;
-$\nu_\sigma$ labels this distribution function as describing one of the
-six species of neutrinos or antineutrinos.
-To do so we trace a geodesic trajectory in the backwards direction
-$-p_\beta$ to sample the incoming radiation along that line of sight.
+of species $\nu_\sigma\in\{\nu_{\{e,\mu,\tau\}},\bar{\nu}_{\{e,\mu,\tau\}}\}$
+with momentum $p_\beta$ impinging on $x_o^\alpha$.
+To do so we trace a geodesic trajectory from $x_o^\alpha$ in the backwards
+direction $-p_\beta$ to sample the incoming radiation along that line of sight.
 By tracing a family of rays intersecting $x_o^\alpha$ we build up a
@@
 The ray tracing framework
-is conceptually simple because it interprets the equation of radiation
-transport (Eqn.~\ref{eqn:boltzmann})
-as a one-dimensional ordinary differential equation,
-it is numerically cheap because it confines computations of
+is conceptually simple because it solves the equation of radiation
+transport (Eqn.~\ref{eqn:boltzmann}) along characteristics,
+reducing it to a one-dimensional ordinary differential equation.
+It is numerically cheap because it confines computations of
 $f^{\nu_\sigma}(x_o^\alpha;p_\beta)$ to the past light-cone of $x_o^\alpha$,
-and it easily extends to high resolution in energy and angle by simply
+with the history of that light-cone truncated at large optical depth.
+\todo{state scaling $\mathscr{O}(n)$?}
+And it easily extends to high resolution in energy and angle by simply
 increasing the number of rays sampled.
@@
 And many make the simplifying assumption of blackbody emission from a
-neutrinosurface \citep{birk2007-nunubar, kova2011-gr_ray_tracing}.
+neutrinosurface \citep{birk2007-nunubar, kova2011-gr_ray_tracing},
+limiting them to equilibrium, optically thick configurations.
 Current state-of-the-art ray tracing formulations avoid the assumption of
@@
 We build upon existing ray tracing formulations
-by eschewing any assumptions about the spacetime geometry
-and by including elastic scattering in the integration along each
-geodesic.
+by eschewing any assumptions about the spacetime geometry,
+integrating local emissivities,
+and including elastic scattering in the integration along each geodesic.
 To incorporate scattering processes in our method, we employ estimates of the
 lowest-order moment of the neutrino distribution function
-computed in an M1 transport simulation (formulated by
-\cite{thor1981-truncated_moment, shib2011-truncated_moment}
-and applied e.g. in \cite{ramp2002-truncated_moment,
-  card2013-truncated_moment, fouc2015-m1_nsbh, ocon2015-gr1d_with_nu}).
+computed in an M1 transport simulation.
 If moments are not available either from
@@
 Inelastic scattering and pair processes can introduce further modifications.
-Any phenomena that involve neutrino-neutrino interactions (neutrino oscillations,
-and neutrino-antineutrino annihilation) can depend sensitively on the angular
-distribution.
+Any phenomena that involve neutrino-neutrino interactions,
+for example neutrino oscillations and neutrino-antineutrino annihilation,
+\todo{cite}
+depend sensitively on the angular distribution.
 And spectral changes can strongly affect the nucleosynthesis
@@
 
-\subsection{Outline}
+\subsubsection*{Outline}
 In Sec.~\ref{sec:formulation} we derive the ray tracing equations from the
 Boltzmann Equation and describe our numerical scheme.
-In Sec.~\ref{sec:tests} we present some tests of the code.
-In Sec.~\ref{sec:applications} we present an application:
-computing neutrino spectra as a function of observer's angle
-in the dynamical environment following the merger of two neutron stars.
+In Sec.~\ref{sec:tests} we present tests of the code.
+In Sec.~\ref{sec:applications} we present several applications
+in the dynamical environment following the merger of two neutron stars:
+computing neutrino spectra as a function of observer's angle,
+computing the self-interaction potential descriptive of flavor evolution
+in high-neutrino-density environments,
+and computing the energy dumped into jet-forming electron-positron
+pairs via neutrino-antineutrino annihilation.
 
@@
   dP & \equiv \frac{1}{\sqrt{-\psi}} \, dp_x \, dp_y \, dp_z \,
-  \frac{1}{p^t}(-p_\mu u^\mu),
+  \frac{\varepsilon}{p^t},
 \end{align}
@@
 The number of particles in a given six-volume is
-$dN=g\,f\,dV\,dP/(2\pi)^3$,
+\begin{equation}
+  dN=\frac{g}{(2\pi)^3}\,f\,dV\,dP,
+\end{equation}
 where $g$ counts the number of spin states accessible to the
 particles ($g=1$ for neutrinos), and $f$ is the distribution function.
-Each of $dN$, $dV$, $dP$, and $f$ are spacetime invariants
+Each of $dV$, $dP$, $dN$, and $f$ are spacetime invariants
 \citep{debb2009-gr_boltzmann_1, debb2009-gr_boltzmann_2, lind1966-gr_boltzmann}.
@@
 makes manifest its seven independent arguments.
+\todo{also decompose wrt $p_t$}
 
@@
 \label{ssec:boltzmann}
-Neutrino radiation obeys the relativistic Boltzmann Equation
+Neutrino radiation obeys the relativistic Boltzmann Equation,
+which, suppressing the arguments of $f$ for simplicity, is written,
 \todo{point out limit of QKEs}
@@
   \label{eqn:boltzmann}
-  \frac{d}{d\lambda}f(x^\alpha;\varepsilon,\upsilon_1,\upsilon_2) = C[f],
-\end{equation}
-where $\frac{d}{d\lambda}$ denotes a derivative with respect to the affine
+  \frac{d}{d\lambda}f = C[f],
+\end{equation}
+where $d/d\lambda$ denotes a derivative with respect to the affine
 parameter defining the neutrino momentum (Eqn.~\ref{eqn:geodesic_x})
@@
 having dimension ${\rm energy}\,{\rm length}^{-1}$.
+\todo{describe $C[f]$ in physical terms}
 The source term varies over phase space $(x^\alpha,p_\beta)$,
-and depends locally on the distribution function $f$
+and due to Fermi blocking depends locally on the distribution function $f$
 and nonlocally on the distribution function of this neutrino and its
@@
   \frac{d}{d\lambda}f =
-  \mathscr{E}^* - \mathscr{K}^* f,
+  \mathscr{E} - \mathscr{K} f,
 \end{equation}
 where we have introduced
-$\mathscr{E}^*$, the invariant total emissivity, and
-$\mathscr{K}^*$, the invariant total opacity,
-and for simplicity we have suppressed the arguments of $f$.
-These coefficients are computed by considering Fermi-blocking and
-dependencies on neutrino and antineutrino distribution functions at other
-momenta.
-We provide a detailed derivation for these terms in App.~\ref{sec:source_terms},
-for the two classes of interactions we consider in this work,
-absorption/emission and elastic scattering
-(Tab.~\ref{tab:neutrino_processes}). Within the scope of this work, we have
+\todo{describe $\mathscr{E}$ and $\mathscr{K}$ in physical terms}
+$\mathscr{E}$, the invariant total emissivity, and
+$\mathscr{K}$, the invariant total opacity.
+These coefficients are computed by considering their dependence
+on neutrino and antineutrino distribution functions at other momenta
+(i.e.\ Fermi-blocking).
+
+We consider two classes of interactions in this work:
+the absorption/emission (AE) and elastic scattering (SE) processes
+listed in Tab.~\ref{tab:neutrino_processes}.
+Thus we separate these coefficients:
 \begin{align}
   \label{eqn:inv_emissivity_components}
-  \mathscr{E}^* &= \mathscr{E}_{\rm AE} + \mathscr{E}_{\rm SE},\\
+  \mathscr{E} &= \mathscr{E}_{\rm AE} + \mathscr{E}_{\rm SE},\\
   \label{eqn:inv_opacity_components}
-  \mathscr{K}^* &= \mathscr{K}_{\rm AE} + \mathscr{K}_{\rm SE}.
-\end{align}
-
-For absorption/emission interactions (App.~\ref{ssec:sources_ae}) these are
-simply the sums over the emissivities and absorptive opacities for each
-reaction $i$:
+  \mathscr{K} &= \mathscr{K}_{\rm AE} + \mathscr{K}_{\rm SE}.
+\end{align}
+
+The absorption/emission coefficients are computed from sums
+over the relevant emissivities and opacities
+in Tab.~\ref{tab:neutrino_processes}:
+\todo{define $j$ and $\chi$}
 \begin{align}
@@
   \mathscr{E}_{\rm AE}(\varepsilon)
-  &= \frac{(2\pi)^3}{\varepsilon^2} \sum_{i\,{\rm reactions}} \eta_i(\varepsilon), \\
+  &= \varepsilon \sum_{i\,{\rm reactions}} j_i(\varepsilon), \\
   \label{eqn:ae_opacity_summed}
   \mathscr{K}_{\rm AE}(\varepsilon)
-  &= \varepsilon \sum_{i\,{\rm reactions}} \chi_{{\rm a},i}(\varepsilon).
-\end{align}
-
-For elastic scattering interactions (App.~\ref{ssec:sources_se}) these are
-the background fields and the sum over the scattering opacities for each
-reaction $i$:
+  &= \frac{1}{1-f^{\rm eq}(\varepsilon)} \,
+  \varepsilon \sum_{i\,{\rm reactions}} \chi_{{\rm a},i}(\varepsilon),
+\end{align}
+where $f^{\rm eq}$ is the Fermi-Dirac distribution function with the chemical
+potential dependent on the local density, temperature, and composition of the
+fluid.
+The appearance of $f^{\rm eq}$ in Eqn.~\ref{eqn:ae_opacity_summed} is due to
+the Fermionic nature of the neutrinos, causing $\mathscr{K}_{\rm AE}$ to be
+different than the simple absorption opacity,
+a phenomenon called stimulated absorption.
+See App.~\ref{ssec:sources_ae} for details.
+\todo{identify $\mathscr{K}_{\rm AE}$ and $\kappa^*$,
+  in e.g.\ \cite{hari2010-gr_nunubar_collapsar}}
+
+The elastic scattering coefficients are computed from
+a background field, and a sum over opacities for
+the relevant reactions in Tab.~\ref{tab:neutrino_processes}:
 \begin{align}
@@
 \end{align}
+See App.~\ref{ssec:sources_se} for details.
 
@@
   extracted from a gray simulation and approximating the
-  energy dependence with Eqns.~\ref{eqn:J_from_gray} and
+  energy distribution with Eqns.~\ref{eqn:J_from_gray} and
   \ref{eqn:H_from_gray}.
@@
 \label{ssec:trajectories}
-Each trajectory is labeled by a pair of vectors
+Each trajectory is uniquely labeled by a pair of vectors
 giving some event on the trajectory, $x^\alpha$,
@@
 \label{eqn:geodesic_p}
-  \frac{d p_\beta}{d\lambda} = -\Gamma^\alpha_{\mu\beta} p^\mu p_\alpha,
-\end{equation}
-where $\Gamma^\alpha_{\mu\gamma}$ are the standard connection coefficients,
+  \frac{d p_\beta}{d\lambda} = -\Gamma^\alpha_{\beta\gamma} p^\gamma p_\alpha,
+\end{equation}
+where $p^\alpha=\psi^{\alpha\beta}p_\beta$,
+$\psi^{\alpha\beta}$ is the inverse of the spacetime metric $\psi_{\alpha\beta}$,
+and $\Gamma^\alpha_{\beta\gamma}$ are the standard connection coefficients,
 \begin{equation}
   \label{eqn:christoffel}
-  \Gamma^\beta_{\alpha\gamma} =
-  \frac{1}{2} \psi^{\beta\mu}
-  (\psi_{\mu\alpha,\gamma} + \psi_{\mu\gamma,\alpha} - \psi_{\alpha\gamma,\mu}),
-\end{equation}
-and $\psi_{\alpha\beta}$ is the spacetime metric.
+  \Gamma^\alpha_{\beta\gamma} =
+  \frac{1}{2} \psi^{\alpha\mu}
+  (\psi_{\mu\beta,\gamma} + \psi_{\mu\gamma,\beta} - \psi_{\beta\gamma,\mu}),
+\end{equation}
+with the comma denoting a partial derivative
+$\psi_{\mu\beta,\gamma}=\partial_\gamma\,\psi_{\mu\beta}$.
 
@@
 \label{ssec:rendering_eqn}
-Defining the source function
-$\mathscr{S}^*\equiv\mathscr{E}^*/\mathscr{K}^*$,
-we can integrate Eqn.~\ref{eqn:boltzmann_linear} directly,
+We can integrate Eqn.~\ref{eqn:boltzmann_linear} directly,
 \begin{align}
@@
   \int_{\lambda_e}^\lambda d\lambda'\,e^{-\tau(\lambda,\lambda')}
-  \mathscr{K}^*(\lambda') \mathscr{S}^*(\lambda'),
+  \mathscr{E}(\lambda'),
 \end{align}
@@
   \tau(\lambda,\lambda') \equiv -\int_{\lambda'}^\lambda
-  d\lambda'' \, \mathscr{K}^*(\lambda''),
+  d\lambda'' \, \mathscr{K}(\lambda''),
 \end{equation}
@@
 Fig.~\ref{fig:affine_param}.
+
+We further separate the integrated distribution function into
+its absorption/emission and elastic scattering components,
+$f_{\rm integ} = f_{\rm AE} + f_{\rm SE}$:
+\begin{align}
+  \label{eqn:rendering_fae}
+  f_{\rm AE}(\lambda,\lambda_e)
+  &= \int_{\lambda_e}^\lambda d\lambda'\,e^{-\tau(\lambda,\lambda')}
+  \mathscr{E_{\rm AE}}(\lambda'), \\
+  \label{eqn:rendering_fse}
+  f_{\rm SE}(\lambda,\lambda_e)
+  &= \int_{\lambda_e}^\lambda d\lambda'\,e^{-\tau(\lambda,\lambda')}
+  \mathscr{E_{\rm SE}}(\lambda').
+\end{align}
+Note that the optical depth used in both of these integrands is the total
+optical depth.
 
@@
 decomposition via Eqn.~\ref{eqn:def_momentum}: the Eulerian,
-or normal, observer, $n_\alpha=-\alpha \,\partial_\alpha t$,
-and the fluid, or comoving, observer,
-$u^\alpha=W(n^\alpha+v^\alpha)$.
-Here $t$ is coordinate time, $\alpha$ is lapse,
-$W$ is fluid Lorentz factor, and $v^\alpha$ is fluid coordinate velocity.
+or normal observer $n_\mu=-\alpha \,\partial_\mu t$,
+and the fluid, or comoving observer
+$u^\mu=Wn^\mu+v_{\rm E}^\mu$.
+Here $t$ is coordinate time and $\alpha$ is lapse
+in the standard 3+1 decomposition of the metric
+\begin{equation}
+  \label{eqn:adm_metric}
+  \psi_{\mu\gamma} :=
+  \left(
+  \begin{matrix}
+    -\alpha^2 + \beta^i \beta_i  & \beta_i \\
+    \beta_j                      & g_{ij}
+  \end{matrix}
+  \right).
+\end{equation}
+We have also introduced
+the fluid Lorentz factor $W=-n_\mu u^\mu=\alpha u^t$,
+its Eulerian velocity $v_{\rm E}^\mu=g^\mu_\lambda u^\lambda$
+(distinct from its coordinate velocity $v^\mu=u^\mu/u^t$),
+and the projection tensor orthogonal to the normal observer
+$g^\mu_\lambda=\psi^\mu_\lambda+u^\mu u_\lambda$,
+whose spatial part is the 3-metric, $g_{ij}$.
+
 We distinguish moments computed assuming an Eulerian observer with a tilde,
@@
 we may transform our integrations to coordinate time.
-The equations are
+The coupled system of ordinary differential equations is
 \begin{align}
@@
   - \frac{1}{2} g^{jk}_{,i}\frac{p_j p_k}{p^t}\\
-  \label{eqn:num_finteg}
-  \frac{df_{\rm integ}}{dt} &=
-  e^{-\tau} \mathscr{K}^* \mathscr{S}^*\\
+  \label{eqn:num_fae}
+  \frac{df_{\rm AE}}{dt} &=
+  e^{-\tau} \frac{\mathscr{E_{\rm AE}}}{p^t}\\
+  \label{eqn:num_fse}
+  \frac{df_{\rm SE}}{dt} &=
+  e^{-\tau} \frac{\mathscr{E_{\rm SE}}}{p^t}\\
   \label{eqn:num_tau}
   \frac{d\tau}{dt} &=
-  -\mathscr{K}^*
-\end{align}
-where the comma denotes a partial derivative
-and we have used the standard 3+1 decomposition of the metric
-\begin{equation}
-  \label{eqn:adm_metric}
-  \psi_{\mu\gamma} :=
-  \left(
-  \begin{matrix}
-    -\alpha^2 + \beta^i \beta_i  & \beta_i \\
-    \beta_j                      & g_{ij}
-  \end{matrix}
-  \right).
-\end{equation}
+  - \frac{\mathscr{K}}{p^t}.
+\end{align}
 
 We integrate each ray until we reach a terminal optical depth of
-$\tau_{\rm term}$ at event $x_e^\alpha$.
+$\tau_{\rm term}$ at the earliest emission event $x_e^\alpha$.
+The concept of earliest emission event is a facetious construct we use
+to allow us to truncate our integration at a location along the ray where any
+further additions to the field are negligible due to the large optical depth
+between $x_e^\alpha$ and $x_o^\alpha$.
 We choose $\tau_{\rm term}$ large so that
@@
 $\mu_{\nu_x}=0$.
-If the ray terminates off the numerical domain
-we set $f_{\rm bound}=0$.
+If the ray terminates off the numerical domain before the terminal optical
+depth accumulates, we set $f_{\rm bound}=0$.
 
@@
 The observer specifies $x^i_o$ and $p_{i,o}$,
-and we set $f_{{\rm integ},o}=0$ and $\tau_o=0$.
+and we set $f_{{\rm AE},o}=f_{{\rm SE},o}=0$ and $\tau_o=0$.
 
 We integrate Eqns.~\ref{eqn:num_x}--\ref{eqn:num_tau} with adaptive step sizes,
-using the 5th-order Dormand-Prince algorithm \cite{pres2007-nr_3rd_ed}.
-After a step is taken, the error is estimated for each variable.
+using the 5th-order Dormand-Prince algorithm \cite{pres2007-nr_3rd_ed},
+which produces an error estimate by comparing the 4th and 5th order solutions.
+\todo{confirm}
+After each step is taken, the errors for each of the 9 variables of
+Eqns.~\ref{eqn:num_x}--\ref{eqn:num_tau} are compared.
 If the maximum of these errors exceeds some absolute or relative threshold
+\todo{specify ScaledAbsRel algorithm}
 the step size is decreased and the step recomputed,
@@
 (as described in \cite[App.~B]{bohn2016-code}).
-If computed from a stationary solution to the
+If computed from a quasi-stationary solution to the
+\todo{clarify quasistationary timescale}
 general relativistic hydrodynamics equations,
@@
 in the following configurations. This suite of configurations
-may be understood as a hierarchy of models of increasing
-physical realism: beginning with a homogeneous medium of effectively
+may be understood as a hierarchy of increasing physical realism:
+beginning with a homogeneous medium of effectively
 infinite extent (i.e.\ optically thick) and progressing to a model of a
@@
 evolved using an M1 transport hydrodynamical simulation.
-\todo{also nsns merger?}
-
-We compare to analytic solutions in the simple configurations, and
-numerical solutions from alternative radiation transport algorithms
-(i.e.\ monte carlo or M1 transport) where they are not.
+In the pre-supernova model, we ray tracing distributions to those calculated
+in a Monte Carlo transport simulation.
+Finally, we calculate global measures of the neutrino fields outside of the
+hypermassive neutron star and disk formed in a binary neutron star
+merger simulated by \cite{fouc2016-m1_nsns, fouc2016-m1_evolve_n},
+and compare ray tracing results to those from the M1 transport simulation.
 

### 2017-07-02 paper.tex
Add nsns figures; also force H float placement

@@
 \usepackage[mathscr]{euscript}    % for \mathscr
+\usepackage{float}                % for better control of figure placement
 \usepackage{graphicx}             % for \includegraphics
@@
 %\renewcommand\todo[1]{} % if you wish to hide all todo notes
+
+% force all figures to be placed right where they are in the text
+\floatplacement{figure}{H}
 
@@
 \label{ssec:test_disk_comparison}
-[Compute total and relative number and energy luminosities for a disk evolved
-in one of \cite{fouc2015-m1_nsbh, fouc2016-m1_nsns, fouc2016-m1_evolve_n}.
-Expect agreement within factors of a few (?).]
-\todo{compute luminosities from M1 and RT, present table}
+Total and relative number and energy luminosities for the disk evolved
+in \cite{fouc2016-m1_evolve_n}.
+Agreement better than 10\%.
+\todo{put in table comparison M1 and RT}
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{production-colormap-merid-rho}
+  \caption{A meridional slice of density in the hypermassive neutron
+    star and disk configuration.}
+  \label{fig:nsns_rho_merid}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{production-colormap-equat-rho}
+  \caption{An equatorial slice of density in the hypermassive neutron
+    star and disk configuration.}
+  \label{fig:nsns_rho_equat}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{production-colormap-merid-temp}
+  \caption{A meridional slice of temperature in the hypermassive neutron
+    star and disk configuration.}
+  \label{fig:nsns_temp_merid}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{production-colormap-equat-temp}
+  \caption{A meridional slice of temperature in the hypermassive neutron
+    star and disk configuration.}
+  \label{fig:nsns_temp_equat}
+\end{figure}
 
@@
 Something like this is presented in \cite[Figs.~10-11]{pere2014-nu_wind};
-compare.]
-\todo{compute $H^r(\mu;\varepsilon)$ in nsns and nsbh}
-\todo{compute $\langle \varepsilon \rangle(\mu)$}
+compare.
 
@@
 difference in the character of the flavor evolution.
+See Figs.~\ref{fig:V_nunu-noscat} and \ref{fig:V_nunu-scat}.
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{20170619-potentials-Vnue_Vnua_Vnu-000Mo-022deg-noscat}
+  \caption{Neutrino self-interaction potential along a test trajectory
+    originating at the surface of the hypermassive neutron star, and
+    proceeding outward along a radial coordinate ray with angle
+    $\psi=22^{\circ}$ with respect to the polar axis.
+    The trajectory is parameterized by the coordinate length $s$.
+    This calculation was done with scattering turned off.
+    Outside of the symmetric point at $s\sim4km$, the potential is
+    positive, i.e.\ $\nu_e$-dominated.
+    Compare to Fig.~\ref{fig:V_nunu-scat}.
+    }
+  \label{fig:V_nunu-noscat}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{20170619-potentials-Vnue_Vnua_Vnu-000Mo-022deg-scat}
+  \caption{Same as Fig.~\ref{fig:V_nunu-noscat},
+    but with scattering turned on.
+    Unlike the case with scattering turned off, here the potential is
+    everywhere negative, i.e.\ $\bar{\nu}_e$-dominated.
+    }
+  \label{fig:V_nunu-scat}
+\end{figure}
 

### 2017-07-02 paper.tex
Rearrange figures now that they are force placed H

@@
 
-\subsection{Infinite Homogeneous Moving Slab:
-  Testing Doppler Shift}
-\label{ssec:test_doppler}
-We reproduce the test above, but with the matter moving in
-in the $+z$ direction, with the observer stationary.
-Plot in Fig.~\ref{fig:avg_eps_doppler}.
-\todo{describe test}
-
-\subsection{Idealized Star:
-  Testing the Decoupling Regime}
-\label{ssec:test_ab_star}
-[Test described in~\cite[Sec.~3.2]{smit1997-two_moment}.]
-Plot in Fig.~\ref{fig:f_absorption_sphere}.
-
-\subsection{Idealized Compact Star:
-  Testing Gravitational Redshift and Geodesic Curvature}
-\label{ssec:test_gravity}
-Optically thick star in Schwarszchild.
-\todo{describe $L(r)$, as a function of lapse}
-optically thick disk in Kerr (Fishbone-Moncrief disk):
-set up to give two different surfaces for nue vs nua, but same temp (?).
-\todo{plot of analytic and numerical geodesics}
-\todo{plot of $G$ nue vs nua}
-
-\subsection{Post-Bounce Collapse Profile:
-  Testing Scattering}
-\label{ssec:test_collapse}
-[Standard radiation transport test described in~\cite{ocon2015-gr1d_with_nu},
-and in \cite[App.~E.6]{fouc2015-m1_nsbh},
-and in~\cite{abdi2012-monte_carlo}.]
-Plot in Fig.~\ref{fig:avg_eps_collapse}.
-
-\subsection{Hypermassive Neutron Star Remnant:
-  Testing Everything}
-\label{ssec:test_disk_comparison}
-Total and relative number and energy luminosities for the disk evolved
-in \cite{fouc2016-m1_evolve_n}.
-Agreement better than 10\%.
-\todo{put in table comparison M1 and RT}
-
-\begin{figure}
-  \includegraphics[width=\columnwidth]{production-colormap-merid-rho}
-  \caption{A meridional slice of density in the hypermassive neutron
-    star and disk configuration.}
-  \label{fig:nsns_rho_merid}
-\end{figure}
-
-\begin{figure}
-  \includegraphics[width=\columnwidth]{production-colormap-equat-rho}
-  \caption{An equatorial slice of density in the hypermassive neutron
-    star and disk configuration.}
-  \label{fig:nsns_rho_equat}
-\end{figure}
-
-\begin{figure}
-  \includegraphics[width=\columnwidth]{production-colormap-merid-temp}
-  \caption{A meridional slice of temperature in the hypermassive neutron
-    star and disk configuration.}
-  \label{fig:nsns_temp_merid}
-\end{figure}
-
-\begin{figure}
-  \includegraphics[width=\columnwidth]{production-colormap-equat-temp}
-  \caption{A meridional slice of temperature in the hypermassive neutron
-    star and disk configuration.}
-  \label{fig:nsns_temp_equat}
-\end{figure}
-
 \begin{figure}
@@
 
+\subsection{Infinite Homogeneous Moving Slab:
+  Testing Doppler Shift}
+\label{ssec:test_doppler}
+We reproduce the test above, but with the matter moving in
+in the $+z$ direction, with the observer stationary.
+Plot in Fig.~\ref{fig:avg_eps_doppler}.
+\todo{describe test}
+
 \begin{figure}
@@
 
-\section{Applications}
-\label{sec:applications}
+\subsection{Idealized Star:
+  Testing the Decoupling Regime}
+\label{ssec:test_ab_star}
+[Test described in~\cite[Sec.~3.2]{smit1997-two_moment}.]
+Plot in Fig.~\ref{fig:f_absorption_sphere}.
 
@@
 
+\subsection{Idealized Compact Star:
+  Testing Gravitational Redshift and Geodesic Curvature}
+\label{ssec:test_gravity}
+Optically thick star in Schwarszchild.
+Total luminosity at infinite separation may be computed by
+computing the radial energy flux at a finite separation from the star, $r$,
+and using $p_t$ for the energy. We compute this asymptotic luminosity
+at several radii, $r$, and show that the luminosity is independent of
+the observer's radius.
+\todo{compute $L$, put in figure}
+
+\subsection{Post-Bounce Collapse Profile:
+  Testing Scattering}
+\label{ssec:test_collapse}
+[Standard radiation transport test described in~\cite{ocon2015-gr1d_with_nu},
+and in \cite[App.~E.6]{fouc2015-m1_nsbh},
+and in~\cite{abdi2012-monte_carlo}.]
+Plot in Fig.~\ref{fig:avg_eps_collapse}.
+
 \begin{figure}
@@
 \end{figure}
+
+\subsection{Hypermassive Neutron Star Remnant:
+  Testing Everything}
+\label{ssec:test_disk_comparison}
+Total and relative number and energy luminosities for the disk evolved
+in \cite{fouc2016-m1_evolve_n}.
+Agreement better than 10\%.
+\todo{put in table comparison M1 and RT}
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{production-colormap-merid-rho}
+  \caption{A meridional slice of density in the hypermassive neutron
+    star and disk configuration.}
+  \label{fig:nsns_rho_merid}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{production-colormap-equat-rho}
+  \caption{An equatorial slice of density in the hypermassive neutron
+    star and disk configuration.}
+  \label{fig:nsns_rho_equat}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{production-colormap-merid-temp}
+  \caption{A meridional slice of temperature in the hypermassive neutron
+    star and disk configuration.}
+  \label{fig:nsns_temp_merid}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{production-colormap-equat-temp}
+  \caption{A meridional slice of temperature in the hypermassive neutron
+    star and disk configuration.}
+  \label{fig:nsns_temp_equat}
+\end{figure}
+
+\section{Applications}
+\label{sec:applications}
 

### 2017-07-02 paper.tex
Correct typos

@@
 This is because integrating the Einstein Equations often requires complicated,
-time-dependent guage conditions
+time-dependent gauge conditions
 \citep{lind2007-gen_harmonic, fouc2013-compactness_and_spin}.
@@
 evolved using an M1 transport hydrodynamical simulation.
-In the pre-supernova model, we ray tracing distributions to those calculated
-in a Monte Carlo transport simulation.
+In the pre-supernova model, we compare ray tracing distributions
+to those calculated in a Monte Carlo transport simulation.
 Finally, we calculate global measures of the neutrino fields outside of the
@@
     This calculation was done with scattering turned off.
-    Outside of the symmetric point at $s\sim4km$, the potential is
+    Outside of the symmetric point at $s\sim40\,{\rm km}$, the potential is
     positive, i.e.\ $\nu_e$-dominated.

### 2017-07-04 paper.tex
Introduce the source function, S

@@
 
-\subsubsection*{What physics problems do we want to address?}
+%\subsubsection*{What physics problems do we want to address?}
 But many interesting unsolved problems require an accurate model of the neutrino
@@
 
-\subsubsection*{Why ray tracing?}
+%\subsubsection*{Why ray tracing?}
 In this work we present a ray tracing method to compute neutrino distribution
@@
 
-\subsubsection*{What is new and better about this ray tracing formulation?}
+%\subsubsection*{What is new and better about this ray tracing formulation?}
 Several ray tracing formulations for radiation transport already exist.
@@
 
-\subsubsection*{Why does a covariant formulation matter?}
+%\subsubsection*{Why does a covariant formulation matter?}
 We formulate our ray tracing equations covariantly---free from
@@
 
-\subsubsection*{Why do scattering and pair processes matter?}
+%\subsubsection*{Why do scattering and pair processes matter?}
 Elastic scattering (see Tab.~\ref{tab:neutrino_processes}) can signicantly
@@
 
-\subsubsection*{In what ways is this method limited?}
+%\subsubsection*{In what ways is this method limited?}
 Ray tracing is ideally suited to problems requiring detailed knowledge of
@@
 We use naturalized units in which $\hbar=c=1$.
-And for the remainder of this article we suppress the neutrino species label
+And for most of the remainder of this article we suppress neutrino species label
 $\nu_\sigma$ since the formulation is general to any species.
+Where we do reference particular species we use the three categories
+$\nu_e$, $\bar{\nu}_e$, and
+$\nu_x=\left\{\nu_\mu,\bar{\nu}_\mu,\nu_\tau,\bar{\nu}_\tau\right\}$.
 
@@
 the source term linear in $f$:
-\begin{equation}
+\begin{align}
   \label{eqn:boltzmann_linear}
-  \frac{d}{d\lambda}f =
-  \mathscr{E} - \mathscr{K} f,
-\end{equation}
+  \frac{d}{d\lambda}f &=
+  \mathscr{E} - \mathscr{K} f \\
+  \label{eqn:boltzmann_linear_s}
+  &= \mathscr{K}(\mathscr{S}-f)
+\end{align}
 where we have introduced
-\todo{describe $\mathscr{E}$ and $\mathscr{K}$ in physical terms}
 $\mathscr{E}$, the invariant total emissivity, and
 $\mathscr{K}$, the invariant total opacity.
+These describe respectively the energy gained and the energy lost
+per length traveled by the neutrino, true scalar quantities
+which take identical values for all observers.
+In the second form we have introduced the source function
+$\mathscr{S}\equiv\mathscr{E}/\mathscr{K}$, which makes manifest the
+behavior of the right hand side, driving $f$ toward $\mathscr{S}$
+over a lengthscale $\mathscr{K}$ in our affine parameter;
+or from Eqn.~\ref{eqn:proper_length}, over a proper lengthscale
+$\varepsilon/\mathscr{K}$ measured by our fiducial observer.
+
 These coefficients are computed by considering their dependence
@@
 (i.e.\ Fermi-blocking).
-
 We consider two classes of interactions in this work:
@@
 \end{align}
-where $f^{\rm eq}$ is the Fermi-Dirac distribution function with the chemical
-potential dependent on the local density, temperature, and composition of the
-fluid.
+where $f^{\rm eq}$ is the distribution function of neutrinos in radiative
+equilibrium with the matter, i.e. the Fermi-Dirac distribution function
+\begin{equation}
+  \label{eqn:feq}
+  f^{\rm eq}(\varepsilon) \equiv
+  \left(1+e^{\varepsilon/(k_{\rm B} T) -\eta_\nu}\right)^{-1},
+\end{equation}
+with the neutrino chemical potentials dependent on the local density,
+temperature, and composition of the fluid via the neutron, proton,
+and electron chemical potentials:
+$\eta_{\nu_e}=-\eta_{\bar{\nu}_e}=\eta_p-\eta_n+\eta_{e^-}$ and
+$\eta_{\nu_x}=0$.
 The appearance of $f^{\rm eq}$ in Eqn.~\ref{eqn:ae_opacity_summed} is due to
@@
 different than the simple absorption opacity,
-a phenomenon called stimulated absorption.
+a phenomenon called stimulated absorption \cite{burr2006-neutrino_opacities}.
+By detailed balance of the absorption/emission reactions, we may alternatively
+write the emissivity in terms of the equilibrium distribution function:
+\begin{equation}
+  \label{eqn:ae_emissivity_feq}
+  \mathscr{E}_{\rm AE}(\varepsilon)
+  = \mathscr{K}_{\rm AE}(\varepsilon) f^{\rm eq}(\varepsilon).
+\end{equation}
 See App.~\ref{ssec:sources_ae} for details.
@@
 \begin{align}
-  \label{eqn:se_emissivity_summed}
+  \label{eqn:se_emissivity}
   \mathscr{E}_{\rm SE}(\varepsilon)
@@
 See App.~\ref{ssec:sources_se} for details.
+
+We may use Eqns.~\ref{eqn:ae_emissivity_feq} and \ref{eqn:se_emissivity}
+to separate the source function thus
+$\mathscr{S}=\mathscr{S}_{\rm AE}+\mathscr{S}_{\rm SE}$, with
+\begin{align}
+  \label{eqn:source_ae}
+  \mathscr{S}_{\rm AE}
+  &= \frac{\mathscr{K}_{\rm AE}}{\mathscr{K}} f^{\rm eq}, \\
+  \label{eqn:source_se}
+  \mathscr{S}_{\rm SE}
+  &= \frac{\mathscr{K}_{\rm SE}}{\mathscr{K}} \Phi_{\rm SE}.
+\end{align}
+With these definitions we write a separated form of
+Eqn.~\ref{eqn:boltzmann_linear_s}:
+\begin{equation}
+  \label{eqn:boltzmann_split}
+  \frac{d}{d\lambda}f =
+  \mathscr{K}_{\rm AE}(f^{\rm eq}-f)
+  + \mathscr{K}_{\rm SE}(\Phi_{\rm SE}-f)
+\end{equation}
+From Eqn.~\ref{eqn:boltzmann_split} we see that
+absorption/emission interactions drive the distribution function
+toward $f^{\rm eq}$ over an affine lengthscale $\mathscr{K}_{\rm AE}^{-1}$,
+and elastic scattering interactions drive it
+toward $\Phi_{\rm SE}$ over an affine lengthscale $\mathscr{K}_{\rm SE}^{-1}$.
 
@@
   extracted from a simulation evolved over multiple energy groups
-  to compute the background field with Eqn.~\ref{eqn:se_factor_expansion},
+  to compute the background field with Eqn.~\ref{eqn:background_phi},
 \item
@@
 If we multiply Eqn.~\ref{eqn:geodesic_x} by $u_\alpha \equiv dx_\alpha/ds$,
-we find $ds=\varepsilon \,d\lambda$ is the proper distance traversed by the
-neutrino as measured by our fiducial obsever $u^\alpha$.
+we find the element of proper distance traversed by the
+neutrino as measured by our fiducial obsever $u^\alpha$ is
+\begin{equation}
+  \label{eqn:proper_length}
+  ds=\varepsilon \,d\lambda.
+\end{equation}
 
@@
 \label{ssec:rendering_eqn}
-We can integrate Eqn.~\ref{eqn:boltzmann_linear} directly,
-\begin{align}
-  \label{eqn:rendering}
-  f(\lambda,\lambda_e) &=
-  f_{\rm bound}(\lambda,\lambda_e) + f_{\rm integ}(\lambda,\lambda_e), \\
-  \label{eqn:rendering_fbound}
-  f_{\rm bound}(\lambda,\lambda_e) &=
-  f(\lambda_e) e^{-\tau(\lambda,\lambda_e)}, \\
-  \label{eqn:rendering_finteg}
-  f_{\rm integ}(\lambda,\lambda_e) &=
-  \int_{\lambda_e}^\lambda d\lambda'\,e^{-\tau(\lambda,\lambda')}
-  \mathscr{E}(\lambda'),
+We can integrate Eqn.~\ref{eqn:boltzmann_split} directly, 
+with our solution split into a boundary, absorption/emission, and
+a scattering term, $f=f_{\rm bdry}+f_{\rm AE}+f_{\rm SE}$:
+\begin{align}
+  \label{eqn:rendering_fbdry}
+  f_{\rm bdry}(\lambda,\lambda_e)
+  &= f(\lambda_e) e^{-\tau(\lambda,\lambda_e)}, \\
+  \label{eqn:rendering_fae}
+  f_{\rm AE}(\lambda,\lambda_e)
+  &= \int_{\lambda_e}^\lambda d\lambda'\,e^{-\tau(\lambda,\lambda')}
+  \mathscr{K}_{\rm AE}(\lambda')f^{\rm eq}(\lambda'), \\
+  \label{eqn:rendering_fse}
+  f_{\rm SE}(\lambda,\lambda_e)
+  &= \int_{\lambda_e}^\lambda d\lambda'\,e^{-\tau(\lambda,\lambda')}
+  \mathscr{K}_{\rm SE}(\lambda')\Phi_{\rm SE}(\lambda'),
 \end{align}
 where the optical depth is defined,
-\todo{rewrite this section simpler; don't give integral solution
-  since we don't use}
 \begin{equation}
@@
 Fig.~\ref{fig:affine_param}.
-
-We further separate the integrated distribution function into
-its absorption/emission and elastic scattering components,
-$f_{\rm integ} = f_{\rm AE} + f_{\rm SE}$:
-\begin{align}
-  \label{eqn:rendering_fae}
-  f_{\rm AE}(\lambda,\lambda_e)
-  &= \int_{\lambda_e}^\lambda d\lambda'\,e^{-\tau(\lambda,\lambda')}
-  \mathscr{E_{\rm AE}}(\lambda'), \\
-  \label{eqn:rendering_fse}
-  f_{\rm SE}(\lambda,\lambda_e)
-  &= \int_{\lambda_e}^\lambda d\lambda'\,e^{-\tau(\lambda,\lambda')}
-  \mathscr{E_{\rm SE}}(\lambda').
-\end{align}
-Note that the optical depth used in both of these integrands is the total
-optical depth.
+Note that Eqn.~\ref{eqn:optical_depth} employs the total absorption
+plus scattering opacity, so that the optical depth attenuating the
+integrands of Eqns.~\ref{eqn:rendering_fae} and \ref{eqn:rendering_fse}
+is the total optical depth.
+
+%Note the following equivalents; we could use any of these to express
+%the integrands above:
+%\begin{align}
+%  \mathscr{K}_{\rm AE} \, f^{\rm eq}
+%  &= \mathscr{K}\, \mathscr{S}_{\rm AE}
+%  = \mathscr{E}_{\rm AE},\nonumber \\
+%  \mathscr{K}_{\rm SE} \, \Phi_{\rm SE}
+%  &= \mathscr{K} \, \mathscr{S}_{\rm SE}
+%  = \mathscr{E}_{\rm SE}.\nonumber
+%\end{align}
 
@@
 given by \citep{hugh1994-eh_finding},
-and Eqns.~\ref{eqn:rendering} and~\ref{eqn:optical_depth}
-in the form given below.
+and Eqns.~\ref{eqn:rendering_fae}, \ref{eqn:rendering_fse},
+and \ref{eqn:optical_depth} in the form given below.
 By using the time-component of Eqn.~\ref{eqn:geodesic_x} ($dt=d\lambda \, p^t$)
@@
   \frac{dx^i}{dt} &=
-  g^{ij} \frac{p_j}{p^t} - \beta^{i}\\
+  g^{ij} \frac{p_j}{p^t} - \beta^{i},\\
   \label{eqn:num_p}
@@
   + \beta^k_{,i} p_k
-  - \frac{1}{2} g^{jk}_{,i}\frac{p_j p_k}{p^t}\\
+  - \frac{1}{2} g^{jk}_{,i}\frac{p_j p_k}{p^t},\\
+  \label{eqn:num_tau}
+  \frac{d\tau}{dt} &=
+  - \frac{1}{p^t}\mathscr{K},\\
   \label{eqn:num_fae}
   \frac{df_{\rm AE}}{dt} &=
-  e^{-\tau} \frac{\mathscr{E_{\rm AE}}}{p^t}\\
+  \frac{1}{p^t}e^{-\tau}\mathscr{K_{\rm AE}} \,f^{\rm eq},\\
   \label{eqn:num_fse}
   \frac{df_{\rm SE}}{dt} &=
-  e^{-\tau} \frac{\mathscr{E_{\rm SE}}}{p^t}\\
-  \label{eqn:num_tau}
-  \frac{d\tau}{dt} &=
-  - \frac{\mathscr{K}}{p^t}.
+  \frac{1}{p^t}e^{-\tau}\mathscr{K_{\rm SE}} \,\Phi_{\rm SE}.
 \end{align}
@@
 We choose $\tau_{\rm term}$ large so that
-$f_{\rm bound}$ is negligible
-(in practice $e^{-\tau_{\rm term}}=0.001$)
-and compute $f_{\rm bound}$ assuming radiative equilibrium with the matter,
-i.e. $f_{\rm bound}=(1+\exp((\varepsilon-\mu_\nu)/T)$,
-and $\mu_{\nu_e}=-\mu_{\bar{\nu}_e}=\mu_{e^{-}}+\mu_p-\mu_n$ and
-$\mu_{\nu_x}=0$.
+$f_{\rm bdry}$ is negligible (in practice $e^{-\tau_{\rm term}}=0.001$)
+and compute it assuming radiative equilibrium with the matter at $x_e^\alpha$
+using Eqn.~\ref{eqn:feq}: $f_{\rm bdry}=f^{\rm eq}\,e^{-\tau_{\rm term}}$.
 If the ray terminates off the numerical domain before the terminal optical
-depth accumulates, we set $f_{\rm bound}=0$.
+depth accumulates, we set $f_{\rm bdry}=0$.
 
@@
 \label{ssec:test_disk_comparison}
-Total and relative number and energy luminosities for the disk evolved
-in \cite{fouc2016-m1_evolve_n}.
-Agreement better than 10\%.
-\todo{put in table comparison M1 and RT}
+The merger of two neutron stars by gravitational wave emission produces a
+postmerger configuration composed of a single neutron star surrounded by a disk.
+Because of the strong differential rotation of the remnant it temporarily avoids
+collapse to a black hole, even if its mass exceeds the threshold of dynamical
+instability for a rigidly rotating neutron star \cite{duez2009-review}.
+These objects are called hypermassive neutron stars.
+
+Such a configuration was modeled by \cite{fouc2016-m1_nsns} by
+evolving the final inspiral and merger of two identical neutron stars of
+isolated gravitational mass 1.2~$M_{\odot}$. The postmerger configuration
+was evolved using an M1 transport scheme for the neutrinos, evolving the
+energy density, number density, and energy flux \cite{fouc2016-m1_evolve_n}
+for $\sim10\,{\rm ms}$ following merger, defined as the time of the peak of the
+gravitational wave strain.
+
+We used a single time snapshot of the volume data from that configuration
+at $t=11\,{\rm ms}$ after merger (see
+Figs.~\ref{fig:nsns_rho_merid}--\ref{fig:nsns_temp_equat})
+sampling the neutrino fields with ray tracing to compare to some
+measures of the field computed from the M1 transport simulation.
 
@@
   \caption{A meridional slice of density in the hypermassive neutron
-    star and disk configuration.}
+    star and disk configuration.
+    The distorted rectangular boundaries are the boundaries of the grid
+    used in the numerical simulation, which employs a coordinate mapping to
+    concentrate points near the central object.}
   \label{fig:nsns_rho_merid}
@@
   \includegraphics[width=\columnwidth]{production-colormap-equat-temp}
-  \caption{A meridional slice of temperature in the hypermassive neutron
+  \caption{An equatorial slice of temperature in the hypermassive neutron
     star and disk configuration.}
@@
 \end{figure}
+
+In Tab.~\ref{tab:nsns_rt_vs_m1} we present total and relative number and
+energy luminosities for this configuration.
+With scattering turned off, ray tracing errors are as large as 250\%
+(for heavy lepton energy luminosities).
+But with scattering turned on, agreement is better than 20\%, and
+in most cases much better.
+\todo{get accurage M1 numbers from Francois}
+
+\begin{table}%[h]
+  \caption{
+    Comparison of total luminosities and average energies between
+    ray tracing and the M1 transport simulation.
+    The ray tracing totals were computed from sums over
+    observers placed in the $y$-$z$ plane along an arc at coordinate distance
+    $r=250\,{\rm km}$ from the center of the star, at 30 positions
+    distributed evenly in $\cos\theta$ over the colatitude range
+    $\theta\in[0,90^{\circ}]$.
+    M1 values are taken from \cite[Figs. 7, 9, 10]{fouc2016-m1_evolve_n}.
+  }
+  \label{tab:nsns_rt_vs_m1}
+  \begin{tabularx}{\columnwidth}{X r r r}
+    & \,\,{\bf RT noscat} & \,\,{\bf RT scat} & \,\,{\bf M1} \\
+    \hline
+    $L_{\nu_e}\,[10^{52}{\rm erg}\,{\rm s}^{-1}]$        & 5.95 & 6.01 & 5 \\
+    $L_{\bar{\nu}_e}$                                    & 10.6 & 11.8 & 12 \\
+    $L_{\nu_x}$                                          & 30.2 & 12.6 & 12 \\
+    \hline
+    $R_{\nu_e}-R_{\bar{\nu}_e}\,[10^{57}{\rm s}^{-1}]$   & -1.06 & -1.86 & -2 \\
+    \hline
+    $\langle \varepsilon_{\nu_e} \rangle\,[{\rm MeV}]$   & 12.08 & 11.88 & 12 \\
+    $\langle \varepsilon_{\bar{\nu}_e} \rangle$          & 15.96 & 14.69 & 15 \\
+    $\langle \varepsilon_{\nu_x} \rangle$                & 36.25 & 23.67 & 26 \\
+    \hline
+  \end{tabularx}
+\end{table}
 
@@
 compare.
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{theta_distrib-250km-luminosity_R}
+  \caption{Radial number fluxes, $K_r$, as functions of colatitude $\theta$.
+    Sampled for observers at fixed coordinate distances from the center
+    of the star, with $r=250\,{\rm km}$.
+    Here radial fluxes are multiplied by the portion of solid angle
+    represented by each observer, so that the total luminosity is
+    computed by summing all the points.}
+  \label{fig:nsns_theta_distrib_R}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{theta_distrib-250km-luminosity_L}
+  \caption{Same as Fig.~\ref{fig:nsns_theta_distrib_R} but radial energy
+    fluxes, $H_r$.}
+  \label{fig:nsns_theta_distrib_L}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{theta_distrib-250km-avg_eps}
+  \caption{Same as Fig.~\ref{fig:nsns_theta_distrib_R} but
+    average energies $\langle \varepsilon \rangle = J/G$.}
+  \label{fig:nsns_theta_distrib_avg_eps}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{cosA_distrib-intensity-250km-00deg}
+  \caption{Angular distributions of fluxes for an observer on the rotation axis
+    in the hypermassive neutron star configuration.}
+  \label{fig:nsns_cosA_distrib_I_00deg}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{cosA_distrib-intensity-250km-90deg}
+  \caption{Angular distributions of fluxes for an observer in the equatorial
+    plane in the hypermassive neutron star configuration.}
+  \label{fig:nsns_cosA_distrib_I_90deg}
+\end{figure}
 
@@
 inelastic scattering, and the thermal pair processes of annihilation and
-production.
-We present formulae for all of these processes, though in this work, we only
-treat absorption/emission and elastic scattering.
+production. In this work, however, we only treat absorption/emission
+and elastic scattering.
+We seek to write each collision source linear in $f$:
+\begin{align}
+  \label{eqn:sources_cae}
+  C_{\rm AE}
+  &= \mathscr{E}_{\rm AE} + \mathscr{K}_{\rm AE}\, f, \\
+  \label{eqn:sources_cse}
+  C_{\rm SE}
+  &= \mathscr{E}_{\rm SE} + \mathscr{K}_{\rm SE}\, f.
+\end{align}
+
 Each term
@@
 relevant category given in Tab.~\ref{tab:neutrino_processes}.
-These rates are computed most easily as traditional emissivities and opacities
-in the fluid frame of reference;
-but since we use them in the form of an invariant emissivity and opacity
-(defined e.g. in \ref{eqn:inv_emissivity} and \ref{eqn:inv_opacity},
-and obeying Eqn.~\ref{eqn:boltzmann_linear})
+We compute these rates in the rest frame of the fluid,
+but because they are spacetime invariants
 they take the same numerical value in any frame of reference
-and are completely independent of our choice of fiducial observer $u^\alpha$.
-\todo{confirm}
+and are completely independent of our choice of fiducial observer $u^\alpha$
+(see discussion around Eqn.~\ref{eqn:boltzmann_linear}).
 
@@
 By contrast $C^{\nu_x}_{{\rm AE}} = 0$,
+\todo{but we use a nonzero $C^{\nu_x}_{{\rm AE}}$}
 where $\nu_x$ stands for any of the four heavy-lepton neutrinos and
@@
 interpolate quad-linearly to the points sampled along each ray.
-\todo{derive and type up all sources in terms of $J$ and $H^\alpha$
-  following \cite[Sec.~4]{shib2011-truncated_moment}}
 
@@
 \label{ssec:sources_ae}
-At neutrino and thermal energies well below $m_{\mu}\sim100\,{\rm MeV}$
+At neutrino and thermal energies well below the masses of the muon or tauon
+($m_{\mu}\sim100\,{\rm MeV}$)
 only charged current processes involving $\nu_e$ and $\bar{\nu}_e$ are allowed.
@@
   C_{{\rm AE},i}
-  &= \mathscr{E}_{{\rm AE},i} (1-f) - \mathscr{K}_{{\rm AE},i} f \\
+  &= \varepsilon j_i (1-f) - \varepsilon \chi_{a,i} f \\
   \label{eqn:sources_ae_2}
-  &= \mathscr{E}_{{\rm AE},i} - (\mathscr{E}_{{\rm AE},i} + \mathscr{K}_{{\rm AE},i}) f,
-\end{align}
-with
-\begin{equation}
-  \label{eqn:inv_emissivity}
-  \mathscr{E}_{{\rm AE},i}
-  = \varepsilon j_i
-  = \frac{(2\pi)^3}{\varepsilon^2}\eta_i,
-\end{equation}
-where $j$ is the emissivity having dimension
-$[j]={\rm length}^{-1}\,{\rm steradian}^{-1}$,
-$\eta$ is the emissivity having dimension
-$[\eta]={\rm energy}\,{\rm length}^{-3}\,{\rm time}^{-1}\,{\rm energy}^{-1}\,{\rm steradian}^{-1}$
-(natural to formulations using specific intensity instead of distribution
-functions), and
-\begin{equation}
-  \label{eqn:inv_opacity}
-  \mathscr{K}_{{\rm AE},i} = \varepsilon \chi_i,
-\end{equation}
-where $\chi$ is the traditional opacity with dimension
-$[\chi]={\rm length}^{-1}$.
+  &= \varepsilon j_i - (\varepsilon j_i+\varepsilon \chi_{a,i}) f,
+\end{align}
+where $j$ is the emissivity and $\chi_a$ the absorption opacity.
+Both $j$ and $\chi_a$ have dimension ${\rm length}^{-1}$ and represent the
+number of neutrinos emitted or absorbed per length traveled.
+In radiation transport formulations using specific intensities instead of
+distribution functions, an emissivity $\eta$ having dimension
+${\rm energy}\,{\rm length}^{-3}\,{\rm time}^{-1}\,{\rm energy}^{-1}\,{\rm steradian}^{-1}$
+is more commonly used.
+The two are related by $j=(2\pi)^3\eta/\varepsilon^3$.
+Note that for brevity we have suppressed the energy-dependence of
+the terms $j$, $\chi_a$, and the distribution functions.
+
+In the special case of radiative equilibrium we know that the source
+term vanishes: an equal number of neutrinos are emitted from
+and absorbed by the matter for any length traversed.
+We also know in this case that the neutrino distribution function must be
+$f^{\rm eq}$, the equilibrium Fermi-Dirac distribution function of
+Eqn.~\ref{eqn:feq}.
+With these facts we can rerrange Eqn.~\ref{eqn:sources_ae_1} to give us
+Kirchoff's Law:
+\begin{align}
+  \label{eqn:kirchoffs_law}
+  j_i &= \frac{\chi_{a,i}}{1-f^{\rm eq}}f^{\rm eq}, \\
+  \label{eqn:stimulated_opacity}
+  &= \chi^*_{a,i}\,f^{\rm eq},
+\end{align}
+where in Eqn.~\ref{eqn:stimulated_opacity} we have introduced the opacity
+corrected for stimulated absorption, $\chi^*_{a,i}$.
+
+Using these expressions and computing the sum over stimulated opacities
+$\chi^*_{a,i}=\sum_i \chi^*_{a,i}$,
+the invariant emissivity and opacity for absorption/emission are
+\begin{align}
+  \label{eqn:ae_emissivity_summed_2}
+  \mathscr{E}_{{\rm AE}} &= \varepsilon \, \chi^*_a \, f^{\rm eq}, \\
+  \label{eqn:ae_opacity_summed_2}
+  \mathscr{K}_{{\rm AE}} &= \varepsilon \, \chi^*_a.
+\end{align}
+These expressions are equivalent to those in
+Eqns.~\ref{eqn:ae_emissivity_summed} and \ref{eqn:ae_opacity_summed}.
 
@@
 Neutrino scattering on particles of mass much greater than $\varepsilon$
-is essentially iso-energetic. Following 
+(i.e. nucleons) is essentially iso-energetic. Following 
 \cite[Eqn.~A8]{brue1985-core_collapse} or
@@
 \begin{equation}
-  \label{eqn:integral_se}
-  C_{{\rm SE},i}
+  \label{eqn:sources_se_1}
+  C_{{\rm SE},i}(\ell_\alpha)
   = \frac{\varepsilon^3}{(2\pi)^3}
   \oint d\Omega' R_{{\rm SE},i}(\omega')
-  \big(f(\ell'_\beta) - f(\ell_\alpha)\big)
-\end{equation}
-where $R(\omega')$ is the scattering kernel from direction $\ell'_\beta$
-to direction $\ell_\alpha$ having dimension $[R]={\rm energy}^{-1}$,
+  \Big(f(\ell'_\beta) - f(\ell_\alpha)\Big)
+\end{equation}
+where $R_{\rm SE,i}(\omega')$ is the scattering kernel for the $i$-th process
+from direction $\ell'_\beta$ to direction $\ell_\alpha$
+having dimension ${\rm energy}^{-1}$,
 and the cosine of the scattering angle is
-$\omega' \equiv \ell'_\alpha \ell^\alpha$.
-We write this in the form
-\begin{equation}
-  \label{eqn:sources_se}
-  C_{{\rm SE},i}
-  = \mathscr{K}_{{\rm SE},i} \Phi_{{\rm SE},i} - \mathscr{K}_{{\rm SE},i} f,
-\end{equation}
-with
-\begin{align}
-  \label{eqn:se_factor}
-  \Phi_{{\rm SE},i} &\equiv
-  \frac{1}{\mathscr{K}_{{\rm SE},i}}
-  \frac{\varepsilon^3}{(2\pi)^3}
-  \oint d\Omega' R_{{\rm SE},i}(\omega')f(\ell'_\beta) \\
-  \label{eqn:se_opacity}
-  \mathscr{K}_{{\rm SE},i} &\equiv
-  \frac{\varepsilon^3}{(2\pi)^3}
-  \oint d\Omega' R_{{\rm SE},i}(\omega').
-\end{align}
+$\omega' \equiv \psi^{\alpha\beta}\ell_\alpha\ell'_\beta$, with
+$\psi^{\alpha\beta}$ the inverse of the spacetime metric.
+Note that for brevity in Eqn.~\ref{eqn:sources_se_1} we have suppressed the
+energy dependence of all of the terms.
 
@@
   R_{{\rm SE},i}(\varepsilon,\omega) \approx
-  R_{{\rm SE},i}^0(\varepsilon) + \omega \, R_{{\rm SE},i}^1(\varepsilon),
-\end{equation}
-which allows us to write Eqns.~\ref{eqn:se_factor} and \ref{eqn:se_opacity} as
+  R_{{\rm SE},i}^0(\varepsilon) + \omega \, R_{{\rm SE},i}^1(\varepsilon).
+\end{equation}
+Using this expansion, and the moments defined in Eqns.~\ref{eqn:J}
+and \ref{eqn:Ha} we may write Eqn.~\ref{eqn:sources_se_1}:
+\begin{equation}
+  \label{eqn:sources_se_2}
+  C_{{\rm SE},i}
+  = \varepsilon \chi_{s,i} \Phi_{{\rm SE},i} - \varepsilon \chi_{s,i} f,
+\end{equation}
+where we have introduced the scattering opacity $\chi_{s,i}$
+and the background scattering field $\Phi_{\rm SE}$:
+\begin{align}
+  \label{eqn:opacity_s}
+  \chi_{s,i}(\varepsilon)
+  &= \frac{4\pi\varepsilon^3}{(2\pi)^3}R_{{\rm SE},i}^0(\varepsilon), \\
+  \label{eqn:background_phi}
+  \Phi_{\rm SE,i}(\varepsilon,\ell_\alpha)
+  &= \frac{(2\pi)^3}{\varepsilon^3}\frac{1}{4\pi}
+  \left(J(\varepsilon)+
+  \frac{\chi^{\rm tr}_{s,i}(\varepsilon)}{\chi_{s,i}(\varepsilon)}
+  \ell_\alpha H^\alpha(\varepsilon)\right),
+\end{align}
+with $\chi^{\rm tr}_{s,i}$ the transport scattering opacity,
 \todo{drop anisotropic term and defend}
-\begin{align}
-  \label{eqn:se_factor_expansion}
-  \Phi_{{\rm SE},i}(\varepsilon) &= \frac{1}{\mathscr{K}_{{\rm SE},i}}
-  \big(R_{{\rm SE},i}^0(\varepsilon) J(\varepsilon) +
-   R_{{\rm SE},i}^1(\varepsilon) \ell_\alpha H^\alpha(\varepsilon) \big) \\
-  \label{eqn:se_opacity_expansion}
-  \mathscr{K}_{{\rm SE},i}(\varepsilon) &=
-  4\pi \frac{\varepsilon^3}{(2\pi)^3} R_{{\rm SE},i}^0(\varepsilon),
-\end{align}
-where we have made the energy-dependence explicit.
-When we get $J(\varepsilon)$ and $H^\alpha(\varepsilon)$ from a multi-group
-moment evolution, we call this the ``spectral'' method.
+\begin{equation}
+  \label{eqn:opacity_s_tr}
+  \chi^{\rm tr}_{s,i}(\varepsilon)
+  = \frac{4\pi\varepsilon^3}{(2\pi)^3}R_{{\rm SE},i}^1(\varepsilon).
+\end{equation}
+
+Thus computing sums over the opacities
+$\chi_s\equiv\sum_i\chi_{s,i}$ and
+$\chi^{\rm tr}_s\equiv\sum_i\chi^{\rm tr}_{s,i}$,
+the invariant emissivity and opacity for elastic scattering are
+\begin{align}
+  \label{eqn:se_emissivity_summed}
+  \mathscr{E}_{\rm SE}
+  &= \varepsilon \, \chi_s \, \Phi_{\rm SE}, \\
+  \label{eqn:se_opacity_summed}
+  \mathscr{K}_{\rm SE}
+  &= \varepsilon \, \chi_s.
+\end{align}
+
+We have made the energy-dependence of the background scattering field
+explicit in Eqn.~\ref{eqn:background_phi}.
+When we compute $\Phi_{\rm SE}(\varepsilon)$ using $J(\varepsilon)$ and
+$H^\alpha(\varepsilon)$ from a multi-group
+M1 transport evolution, we call this the \emph{spectral} method.
 
@@
 and sometimes the number density $G$.
-We may reconstruct the energy-dependent source terms
-from the gray moments by assuming a diluted Fermi-Dirac spectrum.
-
+If such is the case we resort to the \emph{gray} method,
+by reconstructing the energy-dependent source terms
+from gray moments, assuming a diluted Fermi-Dirac spectrum.
 We use the following procedure:
@@
 \end{enumerate}
-When we use these moments in Eqn.~\ref{eqn:se_factor_expansion},
-we call this the ``gray'' method.
+When we use these moments in Eqn.~\ref{eqn:background_phi},
+we call this the \emph{gray} method.
 

### 2017-07-05 paper.tex
Add momentum conventions appendix and simplify upstream text

@@
 it has two degrees of freedom; we make this explicit by defining its
-spatial components with respect to spherical polar angles
+spatial cartesian components with respect to spherical polar angles
 \begin{equation}
   \label{eqn:def_direction}
-  \ell_\alpha :=
-  q (s,\sin\upsilon_1\cos\upsilon_2,\sin\upsilon_1\sin\upsilon_2,\cos\upsilon_1),
-\end{equation}
-with $q$ and $s$ functions of $\upsilon_1$ and $\upsilon_2$.
+  \ell_\alpha \rightarrow
+  q (s,\sin a \cos b,\sin a\sin b,\cos a),
+\end{equation}
+with $q$ and $s$ functions of $a$ and $b$.
 Now our symbol for the distribution function,
-$f(x^\alpha;\varepsilon,\upsilon_1,\upsilon_2)$,
+$f(x^\alpha;\varepsilon,a,b)$,
 makes manifest its seven independent arguments.
-\todo{also decompose wrt $p_t$}
+
+We also define a rotated frame
+\begin{equation}
+  \ell_{\alpha'}=\frac{\partial x^\alpha}{\partial x^{\alpha'}}\,\ell_\alpha \nonumber
+\end{equation}
+in which the cartesian components of the direction 1-form are defined
+\begin{equation}
+  \label{eqn:def_direction_primed}
+  \ell_{\alpha'} \rightarrow
+  q (s,\sin A \cos B,\sin A\sin B,\cos A),
+\end{equation}
+so that momenta with vanishing polar angle $A$ move outward along coordinate
+radial rays.
+This transformation is chosen so that for an observer far from the source,
+incoming radiation will be concentrated into a narrow beam around $\cos A\approx1$,
+independent of that observer's position in coordinate space.
+For explicit definitions see App.~\ref{sec:def_momentum}.
+
+In Minkowski spacetime, and for a stationary observer
+$u_\beta\rightarrow(-1,0,0,0)$, we would have $s=0$, $q=1$,
+and $\varepsilon=-p_t$. In that case $\cos A$ may be identified
+\todo{cite someone that uses $\mu$}
+with the familiar forward direction cosine $\mu$.
 
@@
 having dimension ${\rm energy}\,{\rm length}^{-1}$.
-\todo{describe $C[f]$ in physical terms}
 The source term varies over phase space $(x^\alpha,p_\beta)$,
@@
 momentum space while holding $\varepsilon$ constant:
-$d\Omega \equiv d(\cos\upsilon_1)\,d\upsilon_2$.
+$d\Omega \equiv d(\cos a)\,db$.
 \todo{consistent with Eqn.~\ref{eqn:def_direction}?}
@@
 
-In practice we use two possible fiducial observers to define the momentum
-decomposition via Eqn.~\ref{eqn:def_momentum}: the Eulerian,
-or normal observer $n_\mu=-\alpha \,\partial_\mu t$,
-and the fluid, or comoving observer
-$u^\mu=Wn^\mu+v_{\rm E}^\mu$.
-Here $t$ is coordinate time and $\alpha$ is lapse
-in the standard 3+1 decomposition of the metric
-\begin{equation}
-  \label{eqn:adm_metric}
-  \psi_{\mu\gamma} :=
-  \left(
-  \begin{matrix}
-    -\alpha^2 + \beta^i \beta_i  & \beta_i \\
-    \beta_j                      & g_{ij}
-  \end{matrix}
-  \right).
-\end{equation}
-We have also introduced
-the fluid Lorentz factor $W=-n_\mu u^\mu=\alpha u^t$,
-its Eulerian velocity $v_{\rm E}^\mu=g^\mu_\lambda u^\lambda$
-(distinct from its coordinate velocity $v^\mu=u^\mu/u^t$),
-and the projection tensor orthogonal to the normal observer
-$g^\mu_\lambda=\psi^\mu_\lambda+u^\mu u_\lambda$,
-whose spatial part is the 3-metric, $g_{ij}$.
-
+There are two sensible observers to use to define the frame in which these
+moments are computed, that is defining the four-velocity of
+Eqn.~\ref{eqn:def_momentum}:
+a stationary (also called an Eulerian) observer,
+and a fluid (also called a comoving) observer \cite{smar1980-gr_hydro}.
+Definitions are given in App.~\ref{sec:def_momentum}.
 We distinguish moments computed assuming an Eulerian observer with a tilde,
@@
 \appendix
+
+\section{Momentum Definitions}
+\label{sec:def_momentum}
+Here we describe our conventions defining the four-momentum of a neutrino.
+We decompose the neutrino momentum into components parallel and orthogonal to
+an observer's velocity $u_\beta$:
+\begin{equation}
+  \label{eqn:def_momentum_2}
+  p_\beta = \varepsilon (u_\beta + \ell_\beta),
+\end{equation}
+with $u_\alpha\ell^\alpha=0$ and $\ell_\alpha\ell^\alpha=1$.
+
+In practice we use two possible fiducial observers to define the momentum
+decomposition via Eqn.~\ref{eqn:def_momentum}: the Eulerian,
+or normal observer $n_\mu=-\alpha \,\partial_\mu t$,
+and the fluid, or comoving observer
+$u^\mu=Wn^\mu+v_{\rm E}^\mu$.
+Here $t$ is coordinate time and $\alpha$ is lapse
+in the standard 3+1 decomposition of the metric
+\begin{equation}
+  \label{eqn:adm_metric}
+  \psi_{\mu\gamma} :=
+  \left(
+  \begin{matrix}
+    -\alpha^2 + \beta^i \beta_i  & \beta_i \\
+    \beta_j                      & g_{ij}
+  \end{matrix}
+  \right).
+\end{equation}
+We have also introduced
+the fluid Lorentz factor $W=-n_\mu u^\mu=\alpha u^t$,
+its Eulerian velocity $v_{\rm E}^\mu=g^\mu_\lambda u^\lambda$
+(distinct from its coordinate velocity $v^\mu=u^\mu/u^t$),
+and the projection tensor orthogonal to the normal observer
+$g^\mu_\lambda=\psi^\mu_\lambda+u^\mu u_\lambda$,
+whose spatial part is the 3-metric, $g_{ij}$.
+
+We specify the neutrino direction in the observer's frame
+with two spherical polar angles
+\begin{align}
+  \ell_i &= q\Omega_i, \\
+  \Omega_i &:= (\sin a \cos b,\sin a\sin b,\cos a).
+\end{align}
+with q and s given by
+\begin{align}
+  q &= W
+\end{align}
+And we further specify a rotated frame using two new spherical polar angles
+\begin{equation}
+  \label{eqn:def_direction_primed_2}
+  \Omega_{i'} :=  (\sin A \cos B,\sin A\sin B,\cos A).
+\end{equation}
+The two frames are related by a standard Euler rotation matrix
+\begin{equation}
+  \label{eqn:rotation_connection}
+  \Omega_i=R^{i'}_i(\theta,\phi)\, \ell_{i'}.
+\end{equation}
 

### 2017-07-05 paper.tex
Finish Momentum Definitions appendix

@@
   \label{eqn:adm_metric}
-  \psi_{\mu\gamma} :=
+  \psi_{\mu\gamma} \rightarrow
   \left(
@@
 with two spherical polar angles
-\begin{align}
-  \ell_i &= q\Omega_i, \\
-  \Omega_i &:= (\sin a \cos b,\sin a\sin b,\cos a).
-\end{align}
-with q and s given by
-\begin{align}
-  q &= W
-\end{align}
-And we further specify a rotated frame using two new spherical polar angles
+($a$,$b$) with respect to the simulation cartesian coordinates
+\begin{align}
+  \ell_\alpha &\rightarrow q(s,\Omega_i), \\
+  \Omega_i &\rightarrow (\sin a \cos b,\sin a\sin b,\cos a),
+\end{align}
+or alternatively the two spherical polar angles
+($A$,$B$) with respect to rotated coordinates
 \begin{equation}
   \label{eqn:def_direction_primed_2}
-  \Omega_{i'} :=  (\sin A \cos B,\sin A\sin B,\cos A).
-\end{equation}
-The two frames are related by a standard Euler rotation matrix
-\begin{equation}
+  \Omega_{i'} \rightarrow  (\sin A \cos B,\sin A\sin B,\cos A).
+\end{equation}
+The two coordinate systems are related by a standard Euler rotation of first
+$\phi$ about the $z$-axis, then $\theta$ about the rotated $y$-axis,
+with $\phi$ and $\theta$ the azimuthal and polar position of the observer.
+Expressed algebraically:
+\begin{align}
   \label{eqn:rotation_connection}
-  \Omega_i=R^{i'}_i(\theta,\phi)\, \ell_{i'}.
-\end{equation}
+  \Omega_x &= \Omega_{x'}\cos\theta\cos\phi-\Omega_{y'}\sin\phi+\Omega_{z'}\sin\theta\cos\phi,\nonumber\\
+  \Omega_y &= \Omega_{x'}\cos\theta\sin\phi+\Omega_{y'}\cos\phi+\Omega_{z'}\sin\theta\sin\phi,\nonumber\\
+  \Omega_z &= \Omega_{x'}\sin\theta+\Omega_{z'}\cos\theta.
+\end{align}
+
+The scale factors $q$ and $s$ are functions of the neutrino
+direction $\Omega_i$, the observer's velocity $u_\alpha$,
+and the spacetime metric.
+In the case of a fluid observer, specified by an arbitrary $W$ and $u_i$,
+$q$ and $s$ are given by
+\begin{align}
+  \label{eqn:q_general}
+  q
+  &= W\alpha\Bigg(2\beta^i\Omega_i W^2(\beta^i\Omega_i-1) \nonumber\\
+  &\qquad\qquad - 2\Omega_i u_j g^{ij} W \alpha(\beta^i\Omega_i-1)\nonumber\\
+  &\qquad\qquad + \alpha^2\left((\Omega_i u_j g^{ij})^2+\Omega_i\Omega_j g^{ij}W^2\right)\Bigg)^{-1/2}\\
+  \label{eqn:s_general}
+  s
+  &= \beta^i\Omega_i -\Omega_i u_j g^{ij}.
+\end{align}
+In the case of an Eulerian observer, $W=1$ and $u_i=0$,
+and these expressions simplify considerably:
+\begin{align}
+  \label{eqn:q_eulerian}
+  \tilde{q}
+  &= \alpha\left(2\beta^i\Omega_i(\beta^i\Omega_i-1)+\alpha^2\Omega_i\Omega_j g^{ij}\right)^{-1/2}\\
+  \label{eqn:s_eulerian}
+  \tilde{s}
+  &= \beta^i\Omega_i.
+\end{align}
 

### 2017-07-09 paper.tex
Remove unnecessary subscript on background field: Phi_SE -> Phi

@@
   \mathscr{E}_{\rm SE}(\varepsilon)
-  &= \mathscr{K}_{\rm SE}(\varepsilon) \, \Phi_{\rm SE}(\varepsilon) \, \\
+  &= \mathscr{K}_{\rm SE}(\varepsilon) \, \Phi(\varepsilon) \, \\
   \label{eqn:se_opacity_summed}
@@
   \mathscr{S}_{\rm SE}
-  &= \frac{\mathscr{K}_{\rm SE}}{\mathscr{K}} \Phi_{\rm SE}.
+  &= \frac{\mathscr{K}_{\rm SE}}{\mathscr{K}} \Phi.
 \end{align}
@@
   \mathscr{K}_{\rm AE}(f^{\rm eq}-f)
-  + \mathscr{K}_{\rm SE}(\Phi_{\rm SE}-f)
+  + \mathscr{K}_{\rm SE}(\Phi-f)
 \end{equation}
@@
 and elastic scattering interactions drive it
-toward $\Phi_{\rm SE}$ over an affine lengthscale $\mathscr{K}_{\rm SE}^{-1}$.
+toward $\Phi$ over an affine lengthscale $\mathscr{K}_{\rm SE}^{-1}$.
 
 This paper introduces the use of neutrino densities and fluxes evolved in an
-M1 transport simulation to estimate the background field, $\Phi_{\rm SE}$.
+M1 transport simulation to estimate the background field, $\Phi$.
 App.~\ref{ssec:sources_se} details a method to calculate
-$\Phi_{\rm SE}(\varepsilon)$ in two different ways:
+$\Phi(\varepsilon)$ in two different ways:
 \begin{itemize}
@@
   &= \int_{\lambda_e}^\lambda d\lambda'\,e^{-\tau(\lambda,\lambda')}
-  \mathscr{K}_{\rm SE}(\lambda')\Phi_{\rm SE}(\lambda'),
+  \mathscr{K}_{\rm SE}(\lambda')\Phi(\lambda'),
 \end{align}
@@
 %  = \mathscr{E}_{\rm AE},\nonumber \\
-%  \mathscr{K}_{\rm SE} \, \Phi_{\rm SE}
+%  \mathscr{K}_{\rm SE} \, \Phi
 %  &= \mathscr{K} \, \mathscr{S}_{\rm SE}
@@
   \frac{df_{\rm SE}}{dt} &=
-  \frac{1}{p^t}e^{-\tau}\mathscr{K_{\rm SE}} \,\Phi_{\rm SE}.
+  \frac{1}{p^t}e^{-\tau}\mathscr{K_{\rm SE}} \,\Phi.
 \end{align}
@@
   C_{{\rm SE},i}
-  = \varepsilon \chi_{s,i} \Phi_{{\rm SE},i} - \varepsilon \chi_{s,i} f,
+  = \varepsilon \chi_{s,i} \Phi_i - \varepsilon \chi_{s,i} f,
 \end{equation}
 where we have introduced the scattering opacity $\chi_{s,i}$
-and the background scattering field $\Phi_{\rm SE}$:
+and the background scattering field $\Phi$:
 \begin{align}
@@
   \label{eqn:background_phi}
-  \Phi_{\rm SE,i}(\varepsilon,\ell_\alpha)
+  \Phi_i(\varepsilon,\ell_\alpha)
   &= \frac{(2\pi)^3}{\varepsilon^3}\frac{1}{4\pi}
@@
   \mathscr{E}_{\rm SE}
-  &= \varepsilon \, \chi_s \, \Phi_{\rm SE}, \\
+  &= \varepsilon \, \chi_s \, \Phi, \\
   \label{eqn:se_opacity_summed}
@@
 explicit in Eqn.~\ref{eqn:background_phi}.
-When we compute $\Phi_{\rm SE}(\varepsilon)$ using $J(\varepsilon)$ and
+When we compute $\Phi(\varepsilon)$ using $J(\varepsilon)$ and
 $H^\alpha(\varepsilon)$ from a multi-group

### 2017-07-10 paper.tex
Describe inelastic and pair processes

@@
 
+We use the above treatment for $\nu_e$ and $\bar{\nu}_e$ only;
+the $\mu$ and $\tau$ neutrinos and antineutrinos do not participate in
+charged current absorption/emission interactions at these temperatures
+and energies.
+However we do use an effective stimulated absorption opacity $\chi^*_a$
+for the heavy-lepton neutrinos, computed by \lstinline{NuLib} as
+described in \cite{ocon2015-gr1d_with_nu} which follows
+\cite{brue1985-core_collapse,burr2006-neutrino_opacities}.
+This is described in App.~\ref{ssec:sources_pp}.
+
 \subsection{Elastic scattering}
@@
 \label{ssec:sources_si}
-(e.g. \cite[Eqn.~A6]{brue1985-core_collapse}).
+Neutrino scattering off of electrons is inelastic, changing the
+magnitude and direction of the species' momenta.
+A similar formalism to the above could be used to derive source terms
+for inelastic scattering.
+\todo{compare relative amounts of inelastic to elastic scattering}
+We save that for future work, pointing out here that inelastic scattering
+treated this way is very sensitive to the energy-dependence of
+the background field.
 
@@
 \label{ssec:sources_pp}
-(e.g. \cite[Eqn.~A9]{brue1985-core_collapse}).
+We do not include thermal pair processes within the above formalism,
+but as mentioned in App.~\ref{ssec:sources_ae}, we do include
+these processes in an effective absorption opacity for $\mu$ and $\tau$
+neutrinos and antineutrinos. \lstinline{NuLib} follows
+\cite{brue1985-core_collapse,burr2006-neutrino_opacities}
+by computing the emissivity of the pair processes in
+\todo{describe better}
+Tab.~\ref{tab:neutrino_processes}, and applying Kirchoff's law,
+Eqn.~\ref{eqn:kirchoffs_law}, to convert this to an opacity.
+We only use this effective opacity for heavy-lepton neutrinos because
+it is subdominant to the true absorption opacity for
+$\nu_e$ and $\bar{\nu}_e$.
 

### 2017-07-11 paper.tex
Update numerical methods (more updates are needed)

@@
 We integrate each ray until we reach a terminal optical depth of
-$\tau_{\rm term}$ at the earliest emission event $x_e^\alpha$.
+$\tau_{\rm term}$ at the earliest effective emission event $x_e^\alpha$.
 The concept of earliest emission event is a facetious construct we use
-to allow us to truncate our integration at a location along the ray where any
+to allow us to truncate our integration at an event along the ray where any
 further additions to the field are negligible due to the large optical depth
 between $x_e^\alpha$ and $x_o^\alpha$.
-We choose $\tau_{\rm term}$ large so that
-$f_{\rm bdry}$ is negligible (in practice $e^{-\tau_{\rm term}}=0.001$)
-and compute it assuming radiative equilibrium with the matter at $x_e^\alpha$
-using Eqn.~\ref{eqn:feq}: $f_{\rm bdry}=f^{\rm eq}\,e^{-\tau_{\rm term}}$.
-If the ray terminates off the numerical domain before the terminal optical
-depth accumulates, we set $f_{\rm bdry}=0$.
+We choose $\tau_{\rm term}=14$ so that $e^{-\tau_{\rm term}}<10^{-6}$,
+and we then discard the contribution of $f_{\rm bdry}$
+(Eqn.~\ref{eqn:rendering_fbdry}).
 
 Since we don't know the emission event a priori,
-\todo{make clear ``emission event'' doesn't involve physical assumption}
 we follow the integration backwards in time,
@@
 which produces an error estimate by comparing the 4th and 5th order solutions.
-\todo{confirm}
+\todo{update to 3rd order Runge-Kutta}
 After each step is taken, the errors for each of the 9 variables of

### 2017-07-11 paper.tex
Replace Fig: moving fluid avg energy

@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{fig-moving_fluid-avg_eps_vs_cosA}
+  \includegraphics[width=\columnwidth]{fig-moving_fluid_avg_eps}
   \caption{Average energy $\langle\varepsilon\rangle=J/G$
@@
     with $A$ the angle between the neutrino momentum and the $+z$ direction.
-    Test described in Sec.~\ref{ssec:test_doppler}}
+    Test described in Sec.~\ref{ssec:test_doppler}.
+    The points are computed from ray tracing spectra; the lines are the
+    analytic formula.
+  }
   \label{fig:avg_eps_doppler}

### 2017-07-11 paper.tex
Replace Fig: homogeneous equilibrium

@@
 representative of the fluid in the Massive Collapsed Star test
-(Sec.~\ref{ssec:test_collapse}) at two different radii:
-at 50~km
+(Sec.~\ref{ssec:test_collapse}) at the radius 50~km.
 For comparision with Collapsed Star test, we use the LS180 equation of state,
-in which the equilibrium
-neutrino chemical potentials are
-$\mu_{\nu_e}=-0.639$~MeV, and 
-$\mu_{\nu_e}=5.771$~MeV respectively
+in which the equilibrium $\nu_e$ neutrino chemical potential is
+$\mu_{\nu_e}=-0.575$~MeV
 ($\mu_{\bar{\nu}_e}=-\mu_{\nu_e}$ and $\mu_{\nu_x}=0$).
 Plot in Fig.~\ref{fig:homogeneous_isotropic}.
-\todo{run test at this thermodynamic point}
-
-\begin{figure}
-  \includegraphics[width=\columnwidth]{fig-homogeneous_equilibrium}
-  \caption{Equilibrium distribution functions in a homogeneous background
-    at two thermodynamic points,
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{figure-mfps-50km}
+  \caption{Mean free paths
     representative of the collapse profile at
-    50~km ($\varrho=1.235 \times 10^{11}\,{\rm g}\,{\rm cm}^{-3}$,
-    $T=4$~MeV,
-    $Y_e=0.12$) and
-    200~km ($\varrho=3.026\times 10^{7}\,{\rm g}\,{\rm cm}^{-3}$,
-    $T=0.5$~MeV,
-    $Y_e=0.5$).
-    Described in Sec.~\ref{ssec:test_equilibrium}}
+    50~km ($\varrho=10^{11}\,{\rm g}\,{\rm cm}^{-3}$,
+    $T=3.7$~MeV, $Y_e=0.12$).
+    In the fluid rest frame these are given by
+    $(\chi^*_a)^{-1}$ for the absorption
+    and $(\chi_s)^{-1}$ for elastic scattering interactions.
+  }
+  \label{fig:mfps-50km}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{figure-ferr-50km}
+  \caption{Relative error in integrated equilibrium distribution functions
+    in a homogeneous background at the thermodynamic point
+    representative of the collapse profile at
+    50~km ($\varrho=10^{11}\,{\rm g}\,{\rm cm}^{-3}$,
+    $T=3.7$~MeV, $Y_e=0.12$),
+    as described in Sec.~\ref{ssec:test_equilibrium}.
+    As can be seen in Fig.~\ref{fig:mfps-50km} these three species
+    over the given energies
+    probe our numerical solution over length scales from
+    100~m to $10^{5}\,{\rm km}$.
+    The source of this error is the discarded boundary term
+    $f_{\rm bdry}\sim10^{-6}$, described in Sec.~\ref{ssec:numerical}.
+  }
   \label{fig:homogeneous_isotropic}
@@
 Plot in Fig.~\ref{fig:avg_eps_doppler}.
+\todo{redo test at same thermo point}
 \todo{describe test}

### 2017-07-11 paper.tex
Replace Fig: homogeneous absorption star

@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{fig-absorption_sphere}
-  \caption{Distribution function outside a constant-density star
-    with radius $R=15$~km, and observer at $r=22.5$~km.
+  \includegraphics[width=\columnwidth]{fig-f-absorption_star-11MeV}
+  \caption{Distribution functions outside an idealized homogeneous star
+    with radius $R=50$~km, and observer at $r=75$~km.
     $A$ is the angle between the neutrino momentum and the $\hat{r}$ direction.
-    The sample grid uses 30 points in energy and angle spanning the ranges
-    $\varepsilon\in(0,100)$~MeV and $\cos A\in(0.6454,1)$,
-    with sample points at bin centers.
-    The line depicts the analytic result for one energy: $\varepsilon=15$~MeV.
-    Test described in Sec.~\ref{ssec:test_ab_star}}
+    The sample grid uses 40 points in energy linearly spaced across
+    $\varepsilon\in(0,100)$~MeV,
+    and 30 points in cosine of the polar angle linearly spaced across
+    $\cos A\in(0.7377,1)$.
+    To simplify this plot a single energy group only is displayed:
+    $\varepsilon=11.25\,{\rm MeV}$.
+    The points are computed from ray tracing; the lines from the analytic
+    solution.
+    Test described in Sec.~\ref{ssec:test_ab_star}.
+  }
   \label{fig:f_absorption_sphere}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{fig-G_spectra-absorption_star}
+  \caption{Number density spectra for an idealized homogeneous star.
+    Test described in Sec.~\ref{ssec:test_ab_star}.
+    The lines with points are the ray tracing spectra for this configuration;
+    the solid lines are the ray tracincg spectra for an otherwise identical star
+    of infinite extent (i.e. the spectra from Sec.~\ref{ssec:test_equilibrium}).
+  }
+  \label{fig:G_absorption_sphere}
 \end{figure}

### 2017-07-11 paper.tex
Replace Fig: collapse profile

@@
 and in~\cite{abdi2012-monte_carlo}.]
+\todo{clarify inelastic scattering and velocity terms}
 Plot in Fig.~\ref{fig:avg_eps_collapse}.
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{fig-collapse_profile-avg_eps_vs_cosA}
-  \caption{Average energy $\langle\varepsilon\rangle=J/G$
-    in the collapse profile test.
+  \includegraphics[width=\columnwidth]{fig-collapse_profile_avg_eps_vs_cosA}
+  \caption{Average energies in the collapse profile test,
+    Sec.~\ref{ssec:test_collapse}.
     The observer is at 500~km, and the shock at 150~km,
@@
     fields for scattering, and 4) ray tracing neglecting scattering.
-    Test described in Sec.~\ref{ssec:test_collapse}}
+  }
   \label{fig:avg_eps_collapse}

### 2017-07-11 paper.tex
Add description of noscat and scat formulations

@@
 
+In the following we present two forms of the integrated distribution functions,
+Eqns.~\ref{eqn:num_fae} and \ref{eqn:num_fse}:
+1) the full form  by which the solution is $f_{\rm scat}=f_{\rm AE}+f_{\rm SE}$
+(labeled `scat' since we include elastic scattering), and
+2) the form treating only absorption/emission interactions
+in which we set $\mathscr{K}_{\rm SE}=0$ so that $f_{\rm noscat}=f_{\rm AE}$
+(labeled `no scat' since we ignore elastic scattering).
+There is no simple relation between the two forms for a given ray; that is
+$f_{\rm scat} \neq f_{\rm noscat}+f_{\rm SE}$.
+This is because $f_{\rm AE}$ is integrated using solely the absorption optical
+depth in the first form,
+and the absorption plus scattering optical depth in the second form.
+Before this work, the noscat form of the equations has been the standard for
+ray tracing, though many authors include the scattering optical depth in
+the integration of Eqn.~\ref{eqn:num_fae}
+(see for example \cite{hari2010-gr_nunubar_collapsar}).
+\todo{explain why; cite others}
+
 \subsection{Infinite Homogeneous Slab:
@@
     and $(\chi_s)^{-1}$ for elastic scattering interactions.
+    Computed using \lstinline{NuLib}, as in the ray tracing code.
   }

### 2017-07-11 paper.tex
Add Fig: cumulative f homogeneous slab

@@
 and the absorption plus scattering optical depth in the second form.
+The differences between these two methods is apparent
+in Fig.~\ref{fig:cumulative_f_homogeneous}.
 Before this work, the noscat form of the equations has been the standard for
@@
 Plot in Fig.~\ref{fig:homogeneous_isotropic}.
+\todo{also describe this test with scat}
 
@@
 \begin{figure}
+  \includegraphics[width=\columnwidth]{fig-cumulative_f-homogeneous}
+  \caption{Cumulative distribution functions at $\varepsilon=11.25$~MeV
+    in homogeneous infinite slab of Sec.~\ref{ssec:test_equilibrium}.
+    The integration proceeds from left to right (backwards in time $t$).
+    The two components of the scat solutions ($f_{\rm AE}$ in points,
+    and $f_{\rm SE}$ in dashed lines) sum to equal the single component
+    of the noscat solution ($f_{\rm AE}$ in points connected by lines).
+    All solutions terminate when they achieve a total optical depth
+    greater than $\tau_term=14$; in the scat cases this occurs at an
+    earlier time since the total mean free paths are less than the
+    absorption mean free paths (much less in the case of $\nu_x$).
+    The positions of the points (only plotted for the $f_{\rm AE}$ components)
+    are the time steps chosen by the adaptive time-stepping algorithm
+    described in Sec.~\ref{ssec:numerical}.
+  }
+  \label{fig:cumulative_f_homogeneous}
+\end{figure}
+
+\begin{figure}
   \includegraphics[width=\columnwidth]{figure-ferr-50km}
   \caption{Relative error in integrated equilibrium distribution functions
-    in a homogeneous background at the thermodynamic point
-    representative of the collapse profile at
-    50~km ($\varrho=10^{11}\,{\rm g}\,{\rm cm}^{-3}$,
-    $T=3.7$~MeV, $Y_e=0.12$),
-    as described in Sec.~\ref{ssec:test_equilibrium}.
+    in the infinite homogeneous slab  described in
+    Sec.~\ref{ssec:test_equilibrium}.
+    Presented here are the relative differences between the final
+    $f_{\rm AE}$ in the noscat case
+    (see Fig.~\ref{fig:cumulative_f_homogeneous})
+    and the equilibrium distribution functions given by
+    Eqn.~\ref{eqn:feq}.
     As can be seen in Fig.~\ref{fig:mfps-50km} these three species

### 2017-07-11 paper.tex
Add a few explanatory bits

@@
 toward $\Phi$ over an affine lengthscale $\mathscr{K}_{\rm SE}^{-1}$.
+(Eqn.~\ref{eqn:proper_length} translates affine length to proper length
+for a given observer; in this case the proper length scale is
+$\varepsilon/\mathscr{K}$.)
 
@@
   \includegraphics[width=\columnwidth]{fig-moving_fluid_avg_eps}
-  \caption{Average energy $\langle\varepsilon\rangle=J/G$
-    of neutrino fields for fluid moving in the $+z$ direction,
+  \caption{Average energy $\langle\varepsilon\rangle(\cos A)=J(\cos A)/G(\cos A)$
+    of $\bar{\nu}_e$ neutrino fields for fluid moving in the $+z$ direction,
     with $A$ the angle between the neutrino momentum and the $+z$ direction.
@@
   \includegraphics[width=\columnwidth]{fig-collapse_profile_avg_eps_vs_cosA}
-  \caption{Average energies in the collapse profile test,
-    Sec.~\ref{ssec:test_collapse}.
+  \caption{Average energies $\langle\varepsilon\rangle(\cos A)=J(\cos A)/G(\cos A)$
+    in the collapse profile test described in Sec.~\ref{ssec:test_collapse}.
     The observer is at 500~km, and the shock at 150~km,

### 2017-07-12 paper.tex
Clean up nsns luminosity table
Correct typo, extend momentum definitions text

@@
     \hline
-    $L_{\nu_e}\,[10^{52}{\rm erg}\,{\rm s}^{-1}]$        & 5.95 & 6.01 & 5 \\
+    $L_{\nu_e}\,[10^{52}\,{\rm erg}\,{\rm s}^{-1}]$      & 5.95 & 6.01 & 5 \\
     $L_{\bar{\nu}_e}$                                    & 10.6 & 11.8 & 12 \\
     $L_{\nu_x}$                                          & 30.2 & 12.6 & 12 \\
-    \hline
-    $R_{\nu_e}-R_{\bar{\nu}_e}\,[10^{57}{\rm s}^{-1}]$   & -1.06 & -1.86 & -2 \\
-    \hline
+    $R_{\nu_e}-R_{\bar{\nu}_e}\,[10^{57}\,{\rm s}^{-1}]$ & -1.06 & -1.86 & -2 \\
     $\langle \varepsilon_{\nu_e} \rangle\,[{\rm MeV}]$   & 12.08 & 11.88 & 12 \\
@@
 and the projection tensor orthogonal to the normal observer
-$g^\mu_\lambda=\psi^\mu_\lambda+u^\mu u_\lambda$,
+$g^\mu_\lambda=\psi^\mu_\lambda+n^\mu n_\lambda$,
 whose spatial part is the 3-metric, $g_{ij}$.
@@
 \end{align}
+In the even simpler case of Minkowski spacetime, these expressions
+reduce to $q=W/(1+s)$, $s=-\Omega_i u^i$,
+$\tilde{q}=1$, $\tilde{s}=0$.
 

### 2017-07-12 paper.tex
Add some notes on Homogeneous Equilibrium test

@@
 with the matter.
-We set up a large slab of matter in two homogeneous configurations
+We set up a large slab of matter
 representative of the fluid in the Massive Collapsed Star test
 (Sec.~\ref{ssec:test_collapse}) at the radius 50~km.
-For comparision with Collapsed Star test, we use the LS180 equation of state,
+For comparision with that test we use the LS180 equation of state
+\cite{latt1991-nuc_eos}
+\footnote{available at \lstinline{stellarcollapse.org}}
 in which the equilibrium $\nu_e$ neutrino chemical potential is
 $\mu_{\nu_e}=-0.575$~MeV
-($\mu_{\bar{\nu}_e}=-\mu_{\nu_e}$ and $\mu_{\nu_x}=0$).
-Plot in Fig.~\ref{fig:homogeneous_isotropic}.
-\todo{also describe this test with scat}
+(with $\mu_{\bar{\nu}_e}=-\mu_{\nu_e}$ and $\mu_{\nu_x}=0$).
+Fig.~\ref{fig:mfps-50km} presents the neutrino mean free paths
+at this thermodynamic point over two energy decades.
+For this test we set up our slab large enough for neutrinos of all
+energies to be trapped. 
+Since our ray tracing algorithm integrates rays to terminal optical
+depths of $\tau_{\rm term}>14$, our domain must extend to $r=10^7\,{\rm km}$.
+
+Fig.~\ref{fig:cumulative_f_homogeneous} displays the integrated $f$ as a
+function of path length along the ray.
+
+Fig.~\ref{fig:homogeneous_isotropic} shows the error in our results.
+\todo{describe with scat on}
 

### 2017-07-13 paper.tex
Describe equil f(t) plot; start describing adaptive time step

@@
 We integrate Eqns.~\ref{eqn:num_x}--\ref{eqn:num_tau} with adaptive step sizes,
-using the 5th-order Dormand-Prince algorithm \cite{pres2007-nr_3rd_ed},
-which produces an error estimate by comparing the 4th and 5th order solutions.
-\todo{update to 3rd order Runge-Kutta}
-After each step is taken, the errors for each of the 9 variables of
+using the 3rd order Runge-Kutta algorithm
+which produces an error estimate by comparing the 3rd and 2nd order solutions.
+After each step is taken, the errors for each of the nine variables of
 Eqns.~\ref{eqn:num_x}--\ref{eqn:num_tau} are compared.
@@
 % see OdeIntegrators::AdaptiveDense
+% and OdeSteppers::RungeKutta3
 % and OdeSteppers::DormandPrince5
+% and OdeController
 % and OdeControllers::ProportionalIntegral_Base::StepSizeToAttempt
@@
 1) the full form  by which the solution is $f_{\rm scat}=f_{\rm AE}+f_{\rm SE}$
-(labeled `scat' since we include elastic scattering), and
+(labeled \emph{scat} since we include elastic scattering), and
 2) the form treating only absorption/emission interactions
 in which we set $\mathscr{K}_{\rm SE}=0$ so that $f_{\rm noscat}=f_{\rm AE}$
-(labeled `no scat' since we ignore elastic scattering).
+(labeled \emph{noscat} since we ignore elastic scattering).
 There is no simple relation between the two forms for a given ray; that is
@@
 in Fig.~\ref{fig:cumulative_f_homogeneous}.
-Before this work, the noscat form of the equations has been the standard for
+Before this work, the \emph{noscat} form of the equations has been the standard for
+\todo{explain why; cite others}
 ray tracing, though many authors include the scattering optical depth in
@@
 (see for example \cite{hari2010-gr_nunubar_collapsar}).
-\todo{explain why; cite others}
 
@@
 
-Fig.~\ref{fig:cumulative_f_homogeneous} displays the integrated $f$ as a
-function of path length along the ray.
+Fig.~\ref{fig:cumulative_f_homogeneous} displays the cumulative distribution
+function integrated along the ray, for each of the three species and using
+both methods, \emph{noscat} and \emph{scat}. We display
+Eqns.~\ref{eqn:num_fae} and \ref{eqn:num_fse} in their integral form:
+\begin{equation}
+  \label{eqn:cumulative_f}
+  f(t)
+  = - \int_{0}^t dt'\,\frac{1}{p^t}e^{-\tau(0,t')}
+  \mathscr{K}(t')f^{\rm eq}(t'),
+\end{equation}
+and the integration proceeds backwards in time, from $t=0$
+\todo{could write more simply because Minkowski}
+to some terminal $t<0$.
+The figure shows this backwards-in-time integration proceeding left to right.
+
+In this test and only this test, we used a higher-order integrator, a 5th
+order Dormand-Prince algorithm \cite{pres2007-nr_3rd_ed}. This is because
+\todo{discuss with Matt}
+our default 3rd order Runge-Kutta algorithm estimates a vanishing error
+in this configuration.
 
@@
     The integration proceeds from left to right (backwards in time $t$).
-    The two components of the scat solutions ($f_{\rm AE}$ in points,
+    The two components of the \emph{scat} solutions ($f_{\rm AE}$ in points,
     and $f_{\rm SE}$ in dashed lines) sum to equal the single component
-    of the noscat solution ($f_{\rm AE}$ in points connected by lines).
+    of the \emph{noscat} solution ($f_{\rm AE}$ in points connected by lines).
     All solutions terminate when they achieve a total optical depth
-    greater than $\tau_term=14$; in the scat cases this occurs at an
+    greater than $\tau_term=14$; in the \emph{scat} cases this occurs at an
     earlier time since the total mean free paths are less than the
@@
     Presented here are the relative differences between the final
-    $f_{\rm AE}$ in the noscat case
+    $f_{\rm AE}$ in the \emph{noscat} case
     (see Fig.~\ref{fig:cumulative_f_homogeneous})

### 2017-07-18 paper.tex
Fix sign error in optical depth eqn
Commit a bunch of small fixes from the weekend

@@
 This is a challenging task in the environment of a merger,
-which generally lacks any spatial symmetries so that fully general
+which generally lacks any spatial symmetries, so that fully general
 seven-dimensional solutions to the Boltzmann Equation are not feasible.
@@
 Monte Carlo methods \citep{abdi2012-monte_carlo, rich2015-monte_carlo}
-are an excellent tool require large computational resources because of
+are an excellent tool but require large computational resources because of
 their global treatment of the problem.
 
-The state of the art today couples radiation and matter with a truncated
+The state of the art today couples radiation and matter using a truncated
 moment formalism
@@
 \todo{cite others}
-evolving the zeroth- and first-angular moments of the distribution function
-(the energy density and flux respectively) representing the zeroth-energy moment
-of the radiation field (the total energy).
+evolving the zeroth- and first-angular moments of the energy density,
+that is the total energy density and flux.
 This method, commonly called M1 transport, was formulated by
@@
 Recently, \cite{fouc2016-m1_evolve_n} have expanded their M1 transport code
-to also evolve the number densities of neutrinos,
-providing total and average energies throughout the simulated volume.
+to also evolve the zeroth-angular moment of the number density,
+providing total energy densities and average energies throughout the
+simulated volume.
 Even so, with M1 transport codes only evolving two angular moments
 and one or two energy moments,
-we can extract only limited angular and spectral information
+they can extract only limited angular and spectral information
 from the neutrino fields.
@@
   \label{eqn:optical_depth}
-  \tau(\lambda,\lambda') \equiv -\int_{\lambda'}^\lambda
+  \tau(\lambda,\lambda') \equiv \int_{\lambda'}^\lambda
   d\lambda'' \, \mathscr{K}(\lambda''),
@@
     and $(\chi_s)^{-1}$ for elastic scattering interactions.
-    Computed using \lstinline{NuLib}, as in the ray tracing code.
+    Note that elastic scattering opacities are identical across species
+    below energies at which weak magnetism plays a role.
+    In this data they are exactly identical
+    because we have turned off weak magnetism in our opacity calculations
+    in order to compare our results with the historical literature.
+    Computed using \lstinline{NuLib}.
   }
@@
 instability for a rigidly rotating neutron star \cite{duez2009-review}.
-These objects are called hypermassive neutron stars.
-
-Such a configuration was modeled by \cite{fouc2016-m1_nsns} by
+These objects, called hypermassive neutron stars,
+may avoid collapse for $X\textrm{--}Y$~s,
+\todo{cite $X$, $Y$}
+depending upon a number of physical factors including thermal pressure,
+magnetic fields, and the microphysics of the fluid.
+
+Such a configuration was modeled in \cite{fouc2016-m1_nsns} by
 evolving the final inspiral and merger of two identical neutron stars of
 isolated gravitational mass 1.2~$M_{\odot}$. The postmerger configuration
-was evolved using an M1 transport scheme for the neutrinos, evolving the
+was simulated using an M1 transport scheme for the neutrinos, evolving the
 energy density, number density, and energy flux \cite{fouc2016-m1_evolve_n}
-for $\sim10\,{\rm ms}$ following merger, defined as the time of the peak of the
-gravitational wave strain.
-
-We used a single time snapshot of the volume data from that configuration
+for $\sim10\,{\rm ms}$ following merger.
+The fluid was modeled using the LS220 equation of state \cite{latt1991-nuc_eos}.
+
+We use a single time snapshot of the volume data from that configuration
 at $t=11\,{\rm ms}$ after merger (see
@@
 measures of the field computed from the M1 transport simulation.
+\todo{here focus on plots as functions of $\theta$}
 
@@
 compare.
+\todo{here focus on plots of $J(\cos A)$, $H^r(\cos A)$, and $J(\varepsilon)$}
 
@@
 
-In practice we use two possible fiducial observers to define the momentum
-decomposition via Eqn.~\ref{eqn:def_momentum}: the Eulerian,
+We use two possible fiducial observers to define the momentum
+decomposition via Eqn.~\ref{eqn:def_momentum_2}: the Eulerian,
 or normal observer $n_\mu=-\alpha \,\partial_\mu t$,

### 2017-07-19 paper.tex
Cite weak magnetism and pinched spectra

@@
     Note that elastic scattering opacities are identical across species
-    below energies at which weak magnetism plays a role.
+    below energies at which weak magnetism plays a role \cite{horo2002-weak_mag}
+    (?).
     In this data they are exactly identical
@@
 magnitude and direction of the species' momenta.
+\todo{cite \cite{keil2003-pinched_spectra} or someone}
 A similar formalism to the above could be used to derive source terms
@@
 We do not include thermal pair processes within the above formalism,
+\todo{explain this choice, and cite \cite{keil2003-pinched_spectra} or someone}
 but as mentioned in App.~\ref{ssec:sources_ae}, we do include

### 2017-07-20 paper.tex
Expand homogeneous equilibrium text

@@
 whereas Latin indices ($i, j, ...$) range over the spatial coordinates 1--3.
-We use naturalized units in which $\hbar=c=1$.
+We use naturalized units in which $\{\hbar,c,k_{\rm B}\}=1$.
 And for most of the remainder of this article we suppress neutrino species label
@@
 Eqns.~\ref{eqn:num_fae} and \ref{eqn:num_fse}:
-1) the full form  by which the solution is $f_{\rm scat}=f_{\rm AE}+f_{\rm SE}$
+1) the full form  by which the solution is $f_{\rm scat}=f_{\rm AE,scat}+f_{\rm SE,scat}$
 (labeled \emph{scat} since we include elastic scattering), and
 2) the form treating only absorption/emission interactions
-in which we set $\mathscr{K}_{\rm SE}=0$ so that $f_{\rm noscat}=f_{\rm AE}$
+in which we set $\mathscr{K}_{\rm SE}=0$ so that $f_{\rm noscat}=f_{\rm AE,noscat}$
 (labeled \emph{noscat} since we ignore elastic scattering).
-There is no simple relation between the two forms for a given ray; that is
-$f_{\rm scat} \neq f_{\rm noscat}+f_{\rm SE}$.
-This is because $f_{\rm AE}$ is integrated using solely the absorption optical
-depth in the first form,
-and the absorption plus scattering optical depth in the second form.
-The differences between these two methods is apparent
+There is no simple relation between these forms
+(i.e.\ $f_{\rm AE,scat} \neq f_{\rm AE,noscat}$)
+because $f_{\rm AE}$ is integrated using the absorption plus scattering optical depth
+in the first form and solely the absorption optical depth in the second form.
+The differences between the two methods is apparent
 in Fig.~\ref{fig:cumulative_f_homogeneous}.
@@
 in which the equilibrium $\nu_e$ neutrino chemical potential is
-$\mu_{\nu_e}=-0.575$~MeV
-(with $\mu_{\bar{\nu}_e}=-\mu_{\nu_e}$ and $\mu_{\nu_x}=0$).
+$\eta_{\nu_e}=-0.1555$
+(with $\eta_{\bar{\nu}_e}=-\eta_{\nu_e}$ and $\eta_{\nu_x}=0$).
+
+We use the equilibrium distribution functions to define background
+fields for the \emph{scat} case:
+in the \emph{spectral} method we use
+$J(\varepsilon)=\varepsilon^3 f^{\rm eq}(\eta_\nu,T;\varepsilon)/(2\pi^2)$,
+and in the \emph{gray} method we use
+$J=c T^4 \mathscr{F}_3(\eta_\nu)/(2\pi^2)$ and
+$G=c T^3 \mathscr{F}_2(\eta_\nu)/(2\pi^2)$,
+using the Fermi integrals defined in Eqn.~\ref{eqn:fermi_integral},
+and with $c$ the conversion constant from ${\rm energy}^3$ to ${\rm length}^{-3}$.
+
 Fig.~\ref{fig:mfps-50km} presents the neutrino mean free paths
@@
 energies to be trapped.
-Since our ray tracing algorithm integrates rays to terminal optical
-depths of $\tau_{\rm term}>14$, our domain must extend to $r=10^7\,{\rm km}$.
-
-Fig.~\ref{fig:cumulative_f_homogeneous} displays the cumulative distribution
-function integrated along the ray, for each of the three species and using
-both methods, \emph{noscat} and \emph{scat}. We display
-Eqns.~\ref{eqn:num_fae} and \ref{eqn:num_fse} in their integral form:
-\begin{equation}
-  \label{eqn:cumulative_f}
-  f(t)
-  = - \int_{0}^t dt'\,\frac{1}{p^t}e^{-\tau(0,t')}
-  \mathscr{K}(t')f^{\rm eq}(t'),
-\end{equation}
-and the integration proceeds backwards in time, from $t=0$
-\todo{could write more simply because Minkowski}
-to some terminal $t<0$.
-The figure shows this backwards-in-time integration proceeding left to right.
-
-In this test and only this test, we used a higher-order integrator, a 5th
-order Dormand-Prince algorithm \cite{pres2007-nr_3rd_ed}. This is because
-\todo{discuss with Matt}
-our default 3rd order Runge-Kutta algorithm estimates a vanishing error
-in this configuration.
-
-Fig.~\ref{fig:homogeneous_isotropic} shows the error in our results.
-\todo{describe with scat on}
+Our domain extends to $\tau_{\rm term}\,L_{\rm mfp,max}\approx10^7\,{\rm km}$,
+since our ray tracing algorithm integrates rays to terminal optical
+depths of $\tau_{\rm term}\geq14$.
 
@@
 
+Fig.~\ref{fig:cumulative_f_homogeneous} displays the cumulative distribution
+function integrated along the ray, for each of the three species and using
+both methods, \emph{noscat} and \emph{scat}. We display
+Eqns.~\ref{eqn:num_fae} and \ref{eqn:num_fse} in their integral form:
+\begin{equation}
+  \label{eqn:cumulative_f}
+  f(t)
+  = - \int_{0}^t \frac{dt'}{p^t}e^{-\tau(0,t')}
+  \mathscr{K}(t')f^{\rm eq}(t'),
+\end{equation}
+and the integration proceeds backwards in time, from $t=0$
+\todo{could write more simply because Minkowski}
+to some terminal $t<0$.
+The figure shows this backwards-in-time integration proceeding left to right.
+
 \begin{figure}
@@
     in homogeneous infinite slab of Sec.~\ref{ssec:test_equilibrium}.
-    The integration proceeds from left to right (backwards in time $t$).
-    The two components of the \emph{scat} solutions ($f_{\rm AE}$ in points,
-    and $f_{\rm SE}$ in dashed lines) sum to equal the single component
-    of the \emph{noscat} solution ($f_{\rm AE}$ in points connected by lines).
-    All solutions terminate when they achieve a total optical depth
-    greater than $\tau_term=14$; in the \emph{scat} cases this occurs at an
-    earlier time since the total mean free paths are less than the
-    absorption mean free paths (much less in the case of $\nu_x$).
+    The integration proceeds from left to right, or backwards in time $t$.
+    Each ray terminates when it achieves a total optical depth
+    greater than $\tau_{\rm term}=14$.
     The positions of the points (only plotted for the $f_{\rm AE}$ components)
@@
 \end{figure}
+
+In Fig.~\ref{fig:cumulative_f_homogeneous} we see some expected features.
+The final distribution functions asymptote to their equilibrium Fermi-Dirac
+levels $f^{\rm eq}_{\{\nu_e,\bar{\nu}_e,\nu_x\}}=\{0.039, 0.053, 0.046\}$
+within a few mean free paths;
+in the \emph{scat} cases it is the sums $f_{\rm AE}+f_{\rm SE}$ that
+achieve these values.
+And the lengths of the rays are proportional to the mean free paths:
+$L_{\nu_x}>L_{\bar{\nu}_e}>L_{\nu_e}$;
+in the \emph{scat} cases these lengths are less than in the \emph{noscat} cases,
+since the total mean free paths are less than the absorption mean free paths
+(significantly so for $\nu_x$, negligibly so for $\nu_e$).
+
+As Fig.~\ref{fig:mfps-50km} shows, integrating the distribution functions
+over the energies $\varepsilon\in[0,100]$~MeV
+probes our numerical solution over length scales from
+0.1~km to $10^{5}\,{\rm km}$.
+Fig.~\ref{fig:homogeneous_isotropic} shows the error in our results for
+the \emph{noscat} case.
+In this simple case, the dominant source of error is the neglected
+boundary term with a relative scale of $e^{-\tau_{\rm term}}\approx10^{-6}$.
+In the following inhomogeneous configurations,
+errors from the integration dominate over the boundary term.
+In this test and only this test, we used a higher-order integrator, a 5th
+order Dormand-Prince algorithm \cite{pres2007-nr_3rd_ed}. This is because
+\todo{discuss with Matt}
+our default 3rd order Runge-Kutta algorithm estimates a vanishing error
+in this configuration.
 
@@
     Eqn.~\ref{eqn:feq}.
-    As can be seen in Fig.~\ref{fig:mfps-50km} these three species
-    over the given energies
-    probe our numerical solution over length scales from
-    100~m to $10^{5}\,{\rm km}$.
     The source of this error is the discarded boundary term
@@
 \label{ssec:test_collapse}
+We use a single time snapshot of a supernova model in its accretion phase,
+100ms after core bounce.
 [Standard radiation transport test described in~\cite{ocon2015-gr1d_with_nu},
@@
   Interpolate the fluid temperature, $T$,
-  equilibrium neutrino chemical potential, $\mu_\nu$,
+  equilibrium neutrino chemical potential, $\eta_\nu$,
   and evolved neutrino energy and number densities in the fluid frame,
@@
     T_\nu = \langle \varepsilon \rangle
-    \frac{\mathscr{F}_2(\mu_\nu/T)}{\mathscr{F}_3(\mu_\nu/T)},
+    \frac{\mathscr{F}_2(\eta_\nu)}{\mathscr{F}_3(\eta_\nu)},
   \end{equation}
@@
   \begin{equation}
+    \label{eqn:fermi_integral}
     \mathscr{F}_b(\eta) = \int_0^\infty dx \, x^b (1+e^{x-\eta})^{-1}.
@@
     J(\varepsilon) &=
-    J \frac{\varepsilon^3}{T_\nu^4 \,\mathscr{F}_3(\mu_\nu/T)}
-    (1+\exp(\varepsilon/T_\nu-\mu_\nu/T)^{-1}, \\
+    J \frac{\varepsilon^3}{T_\nu^4 \,\mathscr{F}_3(\eta_\nu)}
+    (1+\exp(\varepsilon/T_\nu-\eta_\nu)^{-1}, \\
     \label{eqn:H_from_gray}
     H^\mu(\varepsilon) &=
-    H^\mu \frac{\varepsilon^3}{T_\nu^4\, \mathscr{F}_3(\mu_\nu/T)}
-    (1+\exp(\varepsilon/T_\nu-\mu_\nu/T)^{-1}.
+    H^\mu \frac{\varepsilon^3}{T_\nu^4\, \mathscr{F}_3(\eta_\nu)}
+    (1+\exp(\varepsilon/T_\nu-\eta_\nu)^{-1}.
   \end{align}

### 2017-07-20 paper.tex
Add coauthor zhu

@@
 \author{Evan O'Connor}
+\affiliation{Department of Physics,
+  North Carolina State University, Raleigh, NC 27695, USA}
+
+\author{Y.\ L.\ Zhu}
 \affiliation{Department of Physics,
@@
   \label{fig:V_nunu-scat}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{fig-survival_prob-noscat_nh}
+  \caption{Neutrino survival probabilties for the \emph{noscat} case.
+    In the lower panel we reproduce the unoscillated self-interaction potential
+    $V_{\nu\nu,0}$ from Fig.~\ref{fig:V_nunu-noscat} and a matter potential
+    computed assuming a very simple steady state wind consistent with the
+    densities within the simulation volume inside of $r<70$~km
+    along this trajectory.
+    In this case the neutrinos undergo a symmetric MNR transition
+    inside of $r<50$~km,
+    with the $e$-flavor neutrinos and antineutrinos fully converting to $\nu_x$.
+  }
+  \label{fig:survival-noscat}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{fig-survival_prob-scat_nh}
+  \caption{Same as Fig.~\ref{fig:survival-noscat},
+    but with scattering turned on.
+    In this case the neutrinos undergo a standard MNR transition
+    outside of $r>200$~km,
+    with the $e$-flavor neutrinos and antineutrinos
+    partially converting to $\nu_x$.
+    }
+  \label{fig:survival-scat}
 \end{figure}
@@
 \lstinline{Sedonu}~\footnote{\url{https://bitbucket.org/srichers/sedonu}}
-for comparisons.
-We thank Luke Roberts and Yonglin Zhu for helpful discussions.
-We are indebted to Fran\c{c}ois H\'{e}bert for his timely help in finding and
-correcting bugs in the time-interpolation algorithm.
+for comparisons,
+Fran\c{c}ois H\'{e}bert for his timely help in finding and
+correcting bugs in the time-interpolation algorithm,
+and Luke Roberts for helpful discussions.
 

### 2017-07-20 paper.tex
Add survival probability figures

@@
 \author{Evan O'Connor}
-\affiliation{Department of Physics,
-  North Carolina State University, Raleigh, NC 27695, USA}
-
-\author{Y.\ L.\ Zhu}
 \affiliation{Department of Physics,
@@
 \lstinline{Sedonu}~\footnote{\url{https://bitbucket.org/srichers/sedonu}}
-for comparisons,
-Fran\c{c}ois H\'{e}bert for his timely help in finding and
-correcting bugs in the time-interpolation algorithm,
-and Luke Roberts for helpful discussions.
+for comparisons.
+We thank Luke Roberts and Yonglin Zhu for helpful discussions.
+We are indebted to Fran\c{c}ois H\'{e}bert for his timely help in finding and
+correcting bugs in the time-interpolation algorithm.
 

### 2017-07-21 paper.tex
Add discussion of moving fluid test

--- 2017-07-20 paper.tex
@@
 \label{ssec:test_doppler}
-We reproduce the test above, but with the matter moving in
-in the $+z$ direction, with the observer stationary.
-Plot in Fig.~\ref{fig:avg_eps_doppler}.
-\todo{redo test at same thermo point}
-\todo{describe test}
+We reproduce the test above, but with the matter and observer in relative
+motion.
+We use a stationary (i.e.\ Eulerian) observer
+and a fluid velocity $u^\alpha \rightarrow W(1,0,0,v)$ with $W=(1-v^2)^{-1/2}$.
+All other thermodynamic variables and background fields are unchanged
+since our ray integration uses these quantities in the fluid frame.
+
+A stationary observer sees a neutrino with momentum
+$p_\alpha \rightarrow \tilde{\varepsilon}(-1,\Omega_i)$,
+with direction $\Omega_i \rightarrow (\sin A, 0, \cos A)$,
+and observed energy $\tilde{\varepsilon}$.
+In the fluid frame this neutrino has energy $\varepsilon=-u^\alpha p_\alpha$;
+therefore the average energy varies with observing angle:
+\begin{equation}
+  \label{eqn:doppler_avg_eps}
+  \langle \tilde{\varepsilon} \rangle =
+  \langle \varepsilon \rangle^{\rm eq} \frac{1}{W(1-v \cos A)}.
+\end{equation}
+The equilibrium average energy $\langle \varepsilon \rangle^{\rm eq}$
+is given by $T \mathscr{F}_3(\eta_\nu)/\mathscr{F}_2(\eta_\nu)\approx 3T$,
+with the Fermi integrals given in Eqn.~\ref{eqn:fermi_integral}.
+Eqn.~\ref{eqn:doppler_avg_eps} describes the well-known Doppler effect.
 
@@
   \includegraphics[width=\columnwidth]{fig-moving_fluid_avg_eps}
-  \caption{Average energy $\langle\varepsilon\rangle(\cos A)=J(\cos A)/G(\cos A)$
-    of $\bar{\nu}_e$ neutrino fields for fluid moving in the $+z$ direction,
-    with $A$ the angle between the neutrino momentum and the $+z$ direction.
+  \caption{Average energy of $\bar{\nu}_e$ neutrino fields
+    measured by a stationary observer,
+    $\langle\tilde{\varepsilon}\rangle(\cos A)$.
+    The observer sees neutrinos moving with the fluid for angles $\cos A<0$.
     Test described in Sec.~\ref{ssec:test_doppler}.
     The points are computed from ray tracing spectra; the lines are the
-    analytic formula.
+    analytic formula Eqn.~\ref{eqn:avg_eps_per_cosA}.
   }
@@
 \end{figure}
+
+We sample the distribution function $f(\tilde{\varepsilon},\cos A,B)$
+with ray tracing
+over a uniform grid of 40 points in energy $\tilde{\varepsilon}\in[0,100]$
+and 30 points in angle $\cos A\in[-1,1]$, holding fixed $B=\pi$.
+Results are shown in Fig.~\ref{fig:avg_eps_doppler} for
+the velocities $v=\{0,0.1,0.8\}$.
+The ray tracing results are computed from total densities in each angular bin,
+that is
+\begin{equation}
+  \label{eqn:avg_eps_per_cosA}
+  \langle \tilde{\varepsilon} \rangle (\cos A) =
+  \frac{\tilde{J}(\cos A)}{\tilde{G}(\cos A)},
+\end{equation}
+with the Eulerian energy density per angular bin given by
+\begin{equation}
+  \label{eqn:J_per_cosA}
+  \tilde{J}(\cos A) =
+  \frac{1}{(2\pi)^3} \int_0^{\infty} d\tilde{\varepsilon} \, \tilde{\varepsilon}^3
+  \int_0^{2\pi} dB \, f(\tilde{\varepsilon},\cos A,B),
+\end{equation}
+and $\tilde{G}(\cos A)$ defined analogously.
+\todo{add metric terms for these integrals}
+Results are identical for all \emph{scat} and \emph{noscat} methods.
+In Fig.~\ref{fig:avg_eps_doppler}
+we see the common features of a red-shifted spectrum for receding fluid
+($\cos A\ll0$),
+a blue-shifted spectrum for approaching fluid
+($\cos A\gg0$),
+and a slightly red-shifted spectrum for fluid moving transverse to the
+observer ($\cos A=0$).
+\todo{wait, what? I thought transverse was blue-shifted}
 

### 2017-07-21 paper.tex
Add notes on figures' pedigrees
Remove G(eps) plot for homogeneous star
Add description/discussion of homogeneous star test

@@
 \todo{discuss mc more clearly}
-Monte Carlo methods \citep{abdi2012-monte_carlo, rich2015-monte_carlo}
+Monte Carlo methods \citep{abdi2012-monte_carlo,rich2015-monte_carlo}
 are an excellent tool but require large computational resources because of
@@
   \includegraphics[width=\columnwidth]{fig-affine_parameter_label}
+  % Script: $DIR/figures-radiation_transport_schematics.nb
+  % w/ DIR=/Users/mbdeaton/Dropbox/Research/NuclearAstrophysics/NotesAndReference
   \caption{Affine parameterization of a neutrino trajectory of momentum
@@
 
-\subsection{The Rendering Equation}
+\subsection{The Formal Solution}
 \label{ssec:rendering_eqn}
@@
   \includegraphics[width=\columnwidth]{figure-mfps-50km}
+  % Data:   $DIR/table-50km-nu?.dat
+  % Script: $DIR/GnuplotScripts/plot-mfp-50km.gnu
+  % w/ DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20161007-ray_tracing_tests/HomogeneousIsotropic
   \caption{Mean free paths
@@
   \includegraphics[width=\columnwidth]{fig-cumulative_f-homogeneous}
+  % Data:   $DIR/Lev1-v0-ShootANeutrino-01/Lev1
+  % Script: $DIR/GnuplotScripts/plot_f-integral_over_t.gnu
+  % w/ DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20161007-ray_tracing_tests/HomogeneousIsotropic
   \caption{Cumulative distribution functions at $\varepsilon=11.25$~MeV
@@
   \includegraphics[width=\columnwidth]{figure-ferr-50km}
+  % Data:   $DIR/Lev1-v0-noscat-01/Lev1
+  % Script: $DIR/GnuplotScripts/plot_ferr.gnu
+  % w/ DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20161007-ray_tracing_tests/HomogeneousIsotropic
   \caption{Relative error in integrated equilibrium distribution functions
@@
   \includegraphics[width=\columnwidth]{fig-moving_fluid_avg_eps}
+  % Data:   $DIR/Lev1-v*-noscat-01
+  % Script: $DIR/plot-avg_eps_vs_cosA.gnu
+  % w/ DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20161007-ray_tracing_tests/MovingFluid
   \caption{Average energy of $\bar{\nu}_e$ neutrino fields
@@
 observer ($\cos A=0$).
-\todo{wait, what? I thought transverse was blue-shifted}
 
@@
 \label{ssec:test_ab_star}
-[Test described in~\cite[Sec.~3.2]{smit1997-two_moment}.]
-Plot in Fig.~\ref{fig:f_absorption_sphere}.
+The homogeneous configurations of the previous sections may be extended to
+probe the solution outside the optically thick regime
+by setting up an idealized homogeneous star, with the thermodynamic variables
+constant inside radius $R$ and vanishing outside.
+We choose $R=50$~km and place our observer at $r=75$~km,
+at which position there is a radiation cone of half-opening angle
+$\cos A_{\rm max}\approx0.75$.
+
+The formal solutions of
+Eqns.~\ref{eqn:rendering_fae} and \ref{eqn:rendering_fse}
+may be directly integrated in this scenario. Assuming Minkowski spacetime
+and stationary fluid we have
+\begin{align}
+  \label{eqn:fae_homogeneous_general}
+  f_{\rm AE} &=
+  \frac{\chi_a^{*}}{\chi} f^{\rm eq} \left(1-e^{-\chi s}\right) \\
+  \label{eqn:fse_homogeneous_general}
+  f_{\rm SE} &=
+  \frac{\chi_s}{\chi} \Phi \left(1-e^{-\chi s}\right) \\
+  \label{eqn:geometric_s}
+  s &= 2 R \left(1-\frac{r^2}{R^2}(1-\cos^2 A)\right)^{1/2},
+\end{align}
+where $s$ is the path length traversed by the ray through the star.
+
+Since no analytic form is known for the background field $\Phi$,
+we examine only the \emph{noscat} case,
+with $\chi_s=0$ and $\chi=\chi_a^{*}$.
+This scenario has been widely used in the literature as a test for
+radiation codes
+\cite{smit1997-two_moment,abdi2012-monte_carlo,fouc2015-m1_nsbh}.
+We sample the distribution function over a uniform grid of
+30 points in angle $\cos A \in [0.734,1]$ and
+40 points in energy $\varepsilon \in [0,100]$~MeV.
+Because of the step function in fluid variables at radius $R$,
+we throttle the maximum step size at 1.25~km.
+
+In Fig.~\ref{fig:f_absorption_sphere} we display the samples at
+$\varepsilon=11.25$~MeV, along with the analytic functions
+specific to each species' equilibrium distribution function and opacity.
+As expected only $\nu_e$ saturates at $f^{\rm eq}$, remaining almost
+constant across $\cos A$ until we get to rays that pass through a length
+of the star comparable to or less than the mean free path at this energy,
+16~km. We also see that $\bar{\nu}_e$ comes close to saturating with
+a mean free path just over 100~km; and $\nu_x$ is well into the optically
+thin regime.
+
+Expanding the exponential in Eqn.~\ref{eqn:fae_homogeneous_general},
+we see the distribution function takes the limiting values:
+\begin{equation}
+  f_{\rm AE}(\cos A) = f^{\rm eq}
+  \begin{cases}
+    \chi s(\cos A) & s \ll \chi^{-1}\,\textrm{optically thin}\\
+    1                  & s \gg \chi^{-1}\,\textrm{optically thick}.
+  \end{cases}
+\end{equation}
+This agrees with Fig.~\ref{fig:f_absorption_sphere}, with $\nu_x$
+in the optically thin limit and $\nu_e$ in the optically thick limit.
 
@@
   \includegraphics[width=\columnwidth]{fig-f-absorption_star-11MeV}
+  % Data:   $DIR/Lev1-01/Lev1
+  % Script: $DIR/Lev1-01/Lev1/plot_f_monochromes-11.25MeV.gnu
+  % w/ DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20161007-ray_tracing_tests/HomogeneousAbsorptionSphere
   \caption{Distribution functions outside an idealized homogeneous star
-    with radius $R=50$~km, and observer at $r=75$~km.
+    with radius $R=50$~km, and observer at $r=75$~km,
+    described in Sec.~\ref{ssec:test_ab_star}.
     $A$ is the angle between the neutrino momentum and the $\hat{r}$ direction.
-    The sample grid uses 40 points in energy linearly spaced across
-    $\varepsilon\in(0,100)$~MeV,
-    and 30 points in cosine of the polar angle linearly spaced across
-    $\cos A\in(0.7377,1)$.
-    To simplify this plot a single energy group only is displayed:
+    In this plot we display only the samples at energy
     $\varepsilon=11.25\,{\rm MeV}$.
     The points are computed from ray tracing; the lines from the analytic
-    solution.
-    Test described in Sec.~\ref{ssec:test_ab_star}.
+    solution, Eqn.~\ref{eqn:fae_homogeneous_general}.
   }
@@
 
-\begin{figure}
-  \includegraphics[width=\columnwidth]{fig-G_spectra-absorption_star}
-  \caption{Number density spectra for an idealized homogeneous star.
-    Test described in Sec.~\ref{ssec:test_ab_star}.
-    The lines with points are the ray tracing spectra for this configuration;
-    the solid lines are the ray tracincg spectra for an otherwise identical star
-    of infinite extent (i.e. the spectra from Sec.~\ref{ssec:test_equilibrium}).
-  }
-  \label{fig:G_absorption_sphere}
-\end{figure}
+Though we do not display it here,
+we also examine $\langle \varepsilon \rangle(\cos A)$
+as in Eqn.~\ref{eqn:avg_eps_per_cosA}, and robustly reproduce the
+common feature of increasing average energies at the limb of the star
+for $\nu_e$ and $\bar{\nu}_e$. In contrast $\langle \varepsilon \rangle$
+for $\nu_x$ is constant across angles, as expected when the star
+is optically thin at even the highest energies $\varepsilon\sim100$~MeV.
 
@@
   \includegraphics[width=\columnwidth]{fig-collapse_profile_avg_eps_vs_cosA}
+  % Data:   $DIR/
+  % Script: $DIR/plot-compare_spectral_vs_gray-angular_distribution_avg_eps-production.gnu
+  % w/  DIR_MC     =$DIR/ResultsSherwood-03/Spectra
+  %     DIR_SPEC_SC=$DIR/ResultsBrett-A-03/Test-scat-01
+  %     DIR_GRAY_SC=$DIR/ResultsBrett-B-03/Test-scat-01
+  %     DIR_GRAY_AB=$DIR/ResultsBrett-B-03/Test-noscat-01
+  % and DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20161007-ray_tracing_tests/CollapseProfile/
   \caption{Average energies $\langle\varepsilon\rangle(\cos A)=J(\cos A)/G(\cos A)$
@@
   \includegraphics[width=\columnwidth]{production-colormap-merid-rho}
+  % VTK Data: zosima:/home/mbdeaton/Research/20170602-nsns_vol_data
+  % Plots:    $DIR/production-colormap-{merid,equat}-{rho,temp}.png
+  % w/ DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20170420-nsns_volume_data/02-two_domains_avg_eps/PlotsEtc
   \caption{A meridional slice of density in the hypermassive neutron
@@
   \includegraphics[width=\columnwidth]{production-colormap-equat-rho}
+  % see notes from production-colormap-merid-rho
   \caption{An equatorial slice of density in the hypermassive neutron
@@
   \includegraphics[width=\columnwidth]{production-colormap-merid-temp}
+  % see notes from production-colormap-merid-rho
   \caption{A meridional slice of temperature in the hypermassive neutron
@@
   \includegraphics[width=\columnwidth]{production-colormap-equat-temp}
+  % see notes from production-colormap-merid-rho
   \caption{An equatorial slice of temperature in the hypermassive neutron
@@
   \includegraphics[width=\columnwidth]{theta_distrib-250km-luminosity_R}
+  % Data:   $DIR/PositionDistributions
+  % Data gathered using: $DIR/CollectPositionDistributions.sh
+  % Script: $DIR/GnuScripts/plot-theta_distributions-R-fixed_r.gnu
+  % w/ DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20170324-ray_tracing_nsns/Run03
   \caption{Radial number fluxes, $K_r$, as functions of colatitude $\theta$.
@@
   \includegraphics[width=\columnwidth]{theta_distrib-250km-luminosity_L}
+  % see notes from theta_distrib-250km-luminosity_R
+  % Script: $DIR/GnuScripts/plot-theta_distributions-L-fixed_r.gnu
   \caption{Same as Fig.~\ref{fig:nsns_theta_distrib_R} but radial energy
@@
   \includegraphics[width=\columnwidth]{theta_distrib-250km-avg_eps}
+  % see notes from theta_distrib-250km-luminosity_R
+  % Script: $DIR/GnuScripts/plot-theta_distributions-avg_eps-fixed_r.gnu
   \caption{Same as Fig.~\ref{fig:nsns_theta_distrib_R} but
@@
   \includegraphics[width=\columnwidth]{cosA_distrib-intensity-250km-00deg}
+  % Data:   $DIR/Lev1-e30a60b10-{,no}scat-00
+  % Script: $DIR/plot-angular_distribution-intensity-compare.gnu
+  % w/ DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20170324-ray_tracing_nsns/Run03/000Mo-000.0deg-250km
   \caption{Angular distributions of fluxes for an observer on the rotation axis
@@
   \includegraphics[width=\columnwidth]{cosA_distrib-intensity-250km-90deg}
+  % Data:   $DIR/Lev1-e30a60b10-{,no}scat-00
+  % Script: $DIR/plot-angular_distribution-intensity-compare.gnu
+  % w/ DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20170324-ray_tracing_nsns/Run03/000Mo-090.0deg-250km
   \caption{Angular distributions of fluxes for an observer in the equatorial
@@
   \includegraphics[width=\columnwidth]{20170619-potentials-Vnue_Vnua_Vnu-000Mo-022deg-noscat}
+  % Data:   $DIR/000Mo-022deg-*km/Lev1-e30a60b10-noscat-00
+  % Script: $DIR/nu_potentials.nb
+  % Config: $DIR/RunAB/RunAB-noscat/nu_potentials-config-000Mo-02deg.nb
+  % w/ DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20170412-self_interaction_potentials_M12_M12-LS220-M1/Run03/
   \caption{Neutrino self-interaction potential along a test trajectory
@@
   \includegraphics[width=\columnwidth]{20170619-potentials-Vnue_Vnua_Vnu-000Mo-022deg-scat}
+  % Data:   $DIR/000Mo-022deg-*km/Lev1-e30a60b10-scat-00
+  % Script: $DIR/nu_potentials.nb
+  % Config: $DIR/RunAB/RunAB-scat/nu_potentials-config-000Mo-02deg.nb
+  % w/ DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20170412-self_interaction_potentials_M12_M12-LS220-M1/Run03/
   \caption{Same as Fig.~\ref{fig:V_nunu-noscat},
@@
   \includegraphics[width=\columnwidth]{fig-survival_prob-noscat_nh}
+  % Flavor evolution by Yonglin, see emails 7.19.17-7.20.17
+  % Plot: $DIR/31-900kmNHnoScatterPot.eps
+  % w/ DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20170412-self_interaction_potentials_M12_M12-LS220-M1/Run03/RunAB/probabilityPotential
   \caption{Neutrino survival probabilties for the \emph{noscat} case.
@@
   \includegraphics[width=\columnwidth]{fig-survival_prob-scat_nh}
+  % see notes from fig-survival_prob-noscat_nh
+  % Plot: $DIR/130-800kmNHScatterPot.eps (normal hierarchy)
   \caption{Same as Fig.~\ref{fig:survival-noscat},

### 2017-07-24 paper.tex
Add collapse profile test text

@@
 (with $\eta_{\bar{\nu}_e}=-\eta_{\nu_e}$ and $\eta_{\nu_x}=0$).
+The opacity table is computed using \lstinline{NuLib} from cross-sections
+neglecting weak magnetism
+\todo{anything else?}
+and is stored on a grid covering energy, density, temperature, and composition
+ranges spanning $\varepsilon_i\in[1,280.5]$~MeV logarithmically,
+$\varrho\in[10^6,6.3\times10^15]\,{\rm g}\,{\rm cm}^{-3}$ logarithmically,
+$T\in[0.05,200]$~MeV logarithmically, and 
+$Y_e\in[0.035,0.55]$ linearly,
+with grid extents$\{18,82,65,51\}$ respectively.
+Interpolation is performed linearly.
 
@@
 \begin{equation}
+  \label{eqn:fae_homogeneous_limits}
   f_{\rm AE}(\cos A) = f^{\rm eq}
@@
 \label{ssec:test_collapse}
-We use a single time snapshot of a supernova model in its accretion phase,
-100ms after core bounce.
-[Standard radiation transport test described in~\cite{ocon2015-gr1d_with_nu},
-and in \cite[App.~E.6]{fouc2015-m1_nsbh},
-and in~\cite{abdi2012-monte_carlo}.]
-\todo{clarify inelastic scattering and velocity terms}
-Plot in Fig.~\ref{fig:avg_eps_collapse}.
+To test our scattering treatments we calculate neutrino fields outside a
+collapsed 15~${\rm M}_\odot$ star, 100~ms after core bounce,
+comparing ray tracing fields to those from a Monte Carlo transport calculation.
+Elastic scattering interior to the shock at $r\approx150$~km
+significantly modifies the neutrinos' spectra,
+and the extended envelope outside the shock becomes a source of
+higher-than-average-energy neutrinos.
+
+The 1D matter profile and M1 transport evolution are computed using
+open source supernova evolution code \lstinline{GR1D}
+\footnote{\url{http://www.gr1dcode.org}}
+\cite{ocon2010-gr1d, ocon2015-gr1d_with_nu},
+using a progenitor profile from \cite{woos1995-sn_progenitors}.
+The matter is described by the LS180 equation of state \cite{latt1991-nuc_eos},
+and the opacities are computed and stored in a table
+as described in Sec.~\ref{ssec:test_equilibrium}.
+This standard test is also presented for example in
+\cite{ocon2015-gr1d_with_nu,fouc2015-m1_nsbh,abdi2012-monte_carlo}.
+\todo{also lieb2005, mull2010}
+
+The matter profile and background field are stored on a spherical
+pseudo-spectral grid composed of 11 spherical-shell subdomains
+\todo{cite early SpEC paper}
+comprising a total of 62 radial grid points spaced approximately
+logarithmically across $r\in(0,740)$~km.
+The background scattering field $\Phi(\varepsilon)$ is supplied by
+\lstinline{GR1D} in the form of $J(\varepsilon_i)$,
+with $\varepsilon_i$ representing 18 energy groups
+identical to those in the \lstinline{NuLib} table described in
+Sec.~\ref{ssec:test_equilibrium}.
+For the \emph{spectral} method we use $J(\varepsilon)$ directly,
+using zeroth-order interpolation between energy groups;
+for the \emph{gray} method we use $J=\sum J(\varepsilon_i) \Delta\varepsilon_i$,
+and $G=\sum J(\varepsilon_i) \Delta\varepsilon_i/\varepsilon_i$ instead.
+
+For fiducial neutrino distributions we use the matter profile
+as input into a Monte Carlo radiation transport calculation using
+open source neutrino transport code \lstinline{Sedonu}
+\footnote{\url{https://bitbucket.org/srichers/sedonu}}
+\cite{rich2015-monte_carlo}.
+To homogenize the physics modeled across these three treatments
+(M1 transport to provide the background fields,
+ray tracing to compute neutrino distributions, and
+Monte Carlo transport for a fiducial comparison)
+we turn off inelastic scattering in the M1 and Monte Carlo codes,
+since we have not incorporated that physics into this ray tracing code,
+and we turn off general relativistic effects in the ray tracing code,
+since the Monte Carlo code assumes Minkowski spacetime.
 
@@
   % and DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20161007-ray_tracing_tests/CollapseProfile/
-  \caption{Average energies $\langle\varepsilon\rangle(\cos A)=J(\cos A)/G(\cos A)$
+  \caption{Angular distribution of $\nu_x$ average energies
+    $\langle\varepsilon\rangle(\cos A)$
     in the collapse profile test described in Sec.~\ref{ssec:test_collapse}.
     The observer is at 500~km, and the shock at 150~km,
-    so that the half-opening angle of the shock is $\cos A=0.954$.
+    so that the half-opening angle of the shock is $\cos A\approx0.954$.
     The four methods depicted are 1) a fiducial Monte Carlo calculation,
@@
 \end{figure}
+
+We place a stationary observer at $r=500$~km.
+Fig.~\ref{fig:avg_eps_collapse} shows the distribution of average $\nu_x$
+energies across incoming angles to this observer.
+We show $\nu_x$ because they present the largest scattering effects:
+they scatter through a thicker atmosphere outside their deep emission surface,
+and their hotter spectrum experiences stronger modification due to the
+$\varepsilon^2$ dependence of the scattering cross-section.
+Against the fiducial Monte Carlo distribution, we show the
+\emph{noscat} treatment, and the \emph{scat} treatment using both the
+\emph{spectral} method and the \emph{gray} method.
+
+As expected the average energies from both
+the scattering envelope ($\cos A \lesssim 0.95$) and
+the bright core ($\cos A\gtrsim 0.995 $)
+are badly characterized by the \emph{noscat} treatment.
+In this case for $\nu_x$ the major effect of elastic scattering
+is to decrease the average energy of neutrinos coming from the core
+and increase the average energy of neutrinos coming from the envelope.
+
+We don't show it here, but the angular distribution contributing to the
+total number density is also strongly affected by elastic scattering.
+Without elastic scattering, the central regions emitting 60\%
+of the neutrinos for the species $\{\nu_e,\bar{\nu}_e,\nu_x\}$
+have scales $R_\nu\approx\{40,35,20\}$~km;
+with elastic scattering the scales are
+$R_\nu\approx\{50,45,35\}$~km.
+% From by-hand summation over histogram G(cosA_i)*dcosA
+% Script: $DIR/plot-compare_scat_vs_noscat-G_angular_distribution-all.gnu
+% DIR=.../20161007-ray_tracing_tests/CollapseProfile/ResultsBrett-B-03
+Total luminosities and average energies from the different treatments are
+\todo{update MC results 02->03}
+presented in Tab.~\ref{tab:collapse_rt_vs_mc}.
+
+\begin{table}%[h]
+  % to get mc luminosities: tail $MCDATA/Spectra/luminosities_and_avg_eps-501km-nu?.dat
+  % to get rt luminosities: tail $RTDATA/Moments-nu?.out
+  % MCDATA=$DIR/ResultsSherwood-02
+  % RTDATA_NO=$DIR/ResultsBrett-B-03/Test-noscat-01/
+  % RTDATA_GR=$DIR/ResultsBrett-B-03/Test-scat-01/
+  % RTDATA_SP=$DIR/ResultsBrett-A-03/Test-scat-01/
+  \caption{
+    Comparison of total luminosities and average energies between
+    the following methods:
+    ray tracing ignoring elastic scattering `RT noscat',
+    ray tracing with the gray scattering treatment `RT scat gr',
+    ray tracing with the spectral scattering treatment `RT scat sp', and
+    a fiducial Monte Carlo transport evolution `MC'.
+    We also show average energies of the scattering envelope
+    $\langle\varepsilon\rangle^{\rm sc}$, which are estimated by eye
+    from plots like Fig.~\ref{fig:avg_eps_collapse} at $\cos A < 0.95$.
+    Luminosities have units $10^{52}\,{\rm erg}\,{\rm s}^{-1}$ and
+    average energies have units MeV.
+    The $\nu_x$ luminosities are per-species:
+    multiply by four to get the total heavy-lepton neutrino luminosities.
+  }
+  \label{tab:collapse_rt_vs_mc}
+  \begin{tabularx}{\columnwidth}{X r r r r}
+    & {\bf RT noscat} & \,\,{\bf RT scat gr} & \,\,{\bf RT scat sp} & \,\,{\bf MC} \\
+    \hline
+    $L_{\nu_e}$                                         &  3.53 & 3.14 & 3.04 & 3.48 \\
+    $L_{\bar{\nu}_e}$                                   & 5.19 & 3.05 & 3.03 & 3.01 \\
+    $L_{\nu_x}$                                         & 222.0  & 1.88 & 1.76 & 1.70 \\
+    $\langle \varepsilon_{\nu_e} \rangle$               &  10.6 & 10.6 & 10.9 & 11.0 \\
+    $\langle \varepsilon_{\bar{\nu}_e} \rangle$         & 14.2 & 13.0 & 13.8 & 13.7 \\
+    $\langle \varepsilon_{\nu_x} \rangle$               & 47.8 & 16.0 & 17.3 & 16.2 \\
+    $\langle \varepsilon_{\nu_e} \rangle^{\rm sc}$      & 4    & 14   & 17   & 16 \\
+    $\langle \varepsilon_{\bar{\nu}_e} \rangle^{\rm sc}$& 4    & 24   & 20   & 18 \\
+    $\langle \varepsilon_{\nu_x} \rangle^{\rm sc}$      & 2    & 28   & 31   & 27 \\
+    \hline
+  \end{tabularx}
+\end{table}
+
+We can make some quantitative sense of the scattered energies in
+Fig.~\ref{fig:avg_eps_collapse} from the formal solution explored in the
+simple configurations above.
+In particular, the average energy in the scattering envelope
+($\cos A\lesssim 0.95$) is related to the spectrum in the direction
+of the interior ($\cos A \sim 1$).
+The relation is quite simple if we simplify our realistic model
+to that of a homogeneous matter profile and
+background scattering field.
+
+In the envelope the local temperature is low
+and the ratio $\chi^*_a/\chi_s$ small,
+so $f_{\rm SE}$ strongly dominates over $f_{\rm AE}$.
+By expanding the exponential of
+Eqn.~\ref{eqn:fse_homogeneous_general} as we did for
+Eqn.~\ref{eqn:fae_homogeneous_limits},
+and furthermore factoring out the dominant ${\rm energy}^2$ dependence from the
+scattering opacity $\chi_s\equiv\varepsilon^2 \zeta_s$, we have
+\begin{equation}
+  \label{eqn:fse_homogeneous_thin_limit}
+  f_{\rm SE}(\varepsilon)
+  = \varepsilon^2 \zeta_s \,\Phi(\varepsilon)\,s,
+\end{equation}
+where $s$ is the length of scattering envelope passed through by the ray.
+The average energy of the scattered field is
+\begin{equation}
+  \label{eqn:avg_eps_sc}
+  \langle \varepsilon \rangle^{\rm sc}
+  = \frac{\int d\varepsilon\,\varepsilon^3 f}{\int d\varepsilon\,\varepsilon^2 f}
+  = \frac{\int d\varepsilon\,\varepsilon^5 \Phi}{\int d\varepsilon\,\varepsilon^4 \Phi}.
+\end{equation}
+
+The spectrum of the background field $\Phi(\varepsilon)$ is well-approximated
+as the \emph{scat} solution for $\cos A\sim 1$, since that is the dominant
+source direction for neutrinos.
+And because the scattering envelope is optically thin at all energies,
+we can assume the neutrino spectrum is essentially unchanged in its passage
+through the envelope.
+In order to estimate $\langle \varepsilon \rangle^{\rm sc}$ analytically, we may
+approximate the background field $\Phi(\varepsilon)$ analytically.
+Direct Fermi-Dirac fits using a temperature and chemical potential
+representative of a physical neutrinosurface fair poorly since neutrinos of
+different energies decouple from the matter at different radii, over which
+the thermodynamic state varies substantially.
+Phenomenological Fermi-Dirac fits work well;
+but so do pinched spectral fits which are much simpler \cite{miri????}:
+\begin{equation}
+  \label{eqn:pinched_spectra}
+  \Phi^{\rm pi}(\varepsilon) \propto
+  \varepsilon^{\alpha-2}
+  \exp\left(-(\alpha+1)\frac{\varepsilon}{\langle\varepsilon\rangle}\right),
+\end{equation}
+(where our definition differs from that of \cite{miri????} by a factor of
+$\varepsilon^2$, since we define our distribution function to be dimensionless
+in natural units $\{\hbar,c\}=1$).
+
+Pinching parameters (calculated by eye) for the species
+$\{\nu_e,\bar{\nu}_e,\nu_x\}$ are $\alpha\approx\{3.6,5.1,2.3\}$,
+when we use the average energies from the \emph{spectral}
+\emph{scat} method in Tab.~\ref{tab:nsns_rt_vs_m1}.
+Energy moments of pinched spectra like those in Eqn.~\ref{eqn:avg_eps_sc}
+have a simple analytic form so that Eqn.~\ref{eqn:avg_eps_sc} becomes
+\begin{align}
+  \label{eqn:pinched_avg_eps_sc_1}
+  \langle\varepsilon\rangle^{\rm sc}
+  &\approx \frac{\alpha+3}{\alpha+1} \langle\varepsilon\rangle, \\
+  %\label{eqn:pinched_avg_eps_sc_2}
+  %&= \{1.4,1.3,1.6\} \langle\varepsilon\rangle, \\
+  \label{eqn:pinched_avg_eps_sc_3}
+  &\approx \{16,18,28\}\,~{\rm MeV},
+\end{align}
+again for $\{\nu_e,\bar{\nu}_e,\nu_x\}$ respectively.
+These analytic predictions agree with the average energies of the scattering
+envelope to within 20\%
+% From by-eye examinations of <eps>(cosA) plot using data from
+% MCDATA=$DIR/ResultsSherwood-02
+% RTDATA_GR=$DIR/ResultsBrett-B-03/Test-scat-01/
+% RTDATA_SP=$DIR/ResultsBrett-A-03/Test-scat-01/
+%          nue     nua     nux
+of the Monte Carlo and the \emph{scat} ray tracing calculations,
+even though we have made some drastic simplifications to our model.
 
@@
     Comparison of total luminosities and average energies between
-    ray tracing and the M1 transport simulation.
+    ray tracing using the \emph{noscat} method `RT noscat',
+    ray tracing using the \emph{gray} \emph{scat} method `RT scat',
+    and the M1 transport simulation `M1'.
     The ray tracing totals were computed from sums over
@@
     M1 values are taken from \cite[Figs. 7, 9, 10]{fouc2016-m1_evolve_n}.
+    Energy luminosities have units $10^{52}\,{\rm erg}\,{\rm s}^{-1}$,
+    number luminosities $10^{57}\,{\rm s}^{-1}$, and
+    average energies MeV.
+    The $\nu_x$ luminosities are per-species:
+    multiply by four to get the total heavy-lepton neutrino luminosities.
   }
@@
   \begin{tabularx}{\columnwidth}{X r r r}
-    & \,\,{\bf RT noscat} & \,\,{\bf RT scat} & \,\,{\bf M1} \\
+    & {\bf RT noscat} & \,\,{\bf RT scat} & \,\,{\bf M1} \\
     \hline
-    $L_{\nu_e}\,[10^{52}\,{\rm erg}\,{\rm s}^{-1}]$      & 5.95 & 6.01 & 5 \\
+    $L_{\nu_e}$                                 & 5.95 & 6.01 & 5 \\
     $L_{\bar{\nu}_e}$                           & 10.6 & 11.8 & 12 \\
     $L_{\nu_x}$                                 & 30.2 & 12.6 & 12 \\
-    $R_{\nu_e}-R_{\bar{\nu}_e}\,[10^{57}\,{\rm s}^{-1}]$ & -1.06 & -1.86 & -2 \\
-    $\langle \varepsilon_{\nu_e} \rangle\,[{\rm MeV}]$   & 12.08 & 11.88 & 12 \\
+    $R_{\nu_e}-R_{\bar{\nu}_e}$                 & -1.06 & -1.86 & -2 \\
+    $\langle \varepsilon_{\nu_e} \rangle$       & 12.08 & 11.88 & 12 \\
     $\langle \varepsilon_{\bar{\nu}_e} \rangle$ & 15.96 & 14.69 & 15 \\
@@
 We thank Sherwood Richers for providing his neutrino transport code
-\lstinline{Sedonu}~\footnote{\url{https://bitbucket.org/srichers/sedonu}}
-for comparisons,
+\lstinline{Sedonu} for comparisons,
 Fran\c{c}ois H\'{e}bert for his timely help in finding and

### 2017-07-25 paper.tex
Remove troublesome \todo macro
It appears my use of the \todo macro introduces a lot of compile warnings
"Underfull \hbox (badness 10000) ...". Till I fix that or rearrange my
\todo placements, I'll just turn them off.
Expand analysis of Post-Bounce test

@@
 \newcommand{\todo}[1]{\marginpar{\tiny{\textcolor{red}{#1}}}}
-%\renewcommand\todo[1]{} % if you wish to hide all todo notes
+\renewcommand\todo[1]{} % if you wish to hide all todo notes
 
@@
 significantly modifies the neutrinos' spectra,
+\todo{demonstrate}
 and the extended envelope outside the shock becomes a source of
@@
 Sec.~\ref{ssec:test_equilibrium}.
-For the \emph{spectral} method we use $J(\varepsilon)$ directly,
+For the \emph{spectral} method we use $J(\varepsilon_i)$ directly,
 using zeroth-order interpolation between energy groups;
 for the \emph{gray} method we use $J=\sum J(\varepsilon_i) \Delta\varepsilon_i$,
-and $G=\sum J(\varepsilon_i) \Delta\varepsilon_i/\varepsilon_i$ instead.
+and $G=\sum J(\varepsilon_i) \Delta\varepsilon_i/\varepsilon_i$ instead,
+with $\Delta\varepsilon_i$ the bin-width of the $i$th energy group.
 
@@
 the bright core ($\cos A\gtrsim 0.995 $)
-are badly characterized by the \emph{noscat} treatment.
+are well characterized by the \emph{scat} treatments
+and badly characterized by the \emph{noscat} treatment.
 In this case for $\nu_x$ the major effect of elastic scattering
@@
 We can make some quantitative sense of the scattered energies in
-Fig.~\ref{fig:avg_eps_collapse} from the formal solution explored in the
-simple configurations above.
+Fig.~\ref{fig:avg_eps_collapse} and Tab.~\ref{tab:collapse_rt_vs_mc}
+using the solutions explored in the simple configurations above.
 In particular, the average energy in the scattering envelope
@@
 of the interior ($\cos A \sim 1$).
-The relation is quite simple if we simplify our realistic model
-to that of a homogeneous matter profile and
-background scattering field.
-
-In the envelope the local temperature is low
-and the ratio $\chi^*_a/\chi_s$ small,
-so $f_{\rm SE}$ strongly dominates over $f_{\rm AE}$.
+The relation may be derived by simplifying our realistic model
+to that of a homogeneous matter profile and background scattering field.
+
 By expanding the exponential of
-Eqn.~\ref{eqn:fse_homogeneous_general} as we did for
-Eqn.~\ref{eqn:fae_homogeneous_limits},
-and furthermore factoring out the dominant ${\rm energy}^2$ dependence from the
-scattering opacity $\chi_s\equiv\varepsilon^2 \zeta_s$, we have
-\begin{equation}
+Eqns.~\ref{eqn:fse_homogeneous_general} and \ref{eqn:fae_homogeneous_general}
+as we did for Eqn.~\ref{eqn:fae_homogeneous_limits},
+and furthermore factoring out the dominant energy dependence from the
+cross-sections (i.e.\ $\chi\equiv\varepsilon^2 \zeta$
+with $\zeta$ approximately constant), we have
+\begin{align}
+  \label{eqn:fae_homogeneous_thin_limit}
+  f_{\rm AE}(\varepsilon)
+  &= \varepsilon^2 \zeta_a \,f^{\rm eq}(\varepsilon)\,s, \\
   \label{eqn:fse_homogeneous_thin_limit}
   f_{\rm SE}(\varepsilon)
-  = \varepsilon^2 \zeta_s \,\Phi(\varepsilon)\,s,
-\end{equation}
+  &= \varepsilon^2 \zeta_s \,\Phi(\varepsilon)\,s,
+\end{align}
 where $s$ is the length of scattering envelope passed through by the ray.
-The average energy of the scattered field is
+\todo{should be $\chi_a^*=\varepsilon\zeta_a$ for $\nu_x$}
+In the envelope the local temperature is low so $f^{\rm eq}/\Phi\ll 1$,
+and also $\zeta_a/\zeta_s\ll 1$,
+so $f_{\rm SE}$ strongly dominates over $f_{\rm AE}$.
+The average energy of the scattered field is therefore
 \begin{equation}
@@
   = \frac{\int d\varepsilon\,\varepsilon^3 f}{\int d\varepsilon\,\varepsilon^2 f}
-  = \frac{\int d\varepsilon\,\varepsilon^5 \Phi}{\int d\varepsilon\,\varepsilon^4 \Phi}.
+  \approx \frac{\int d\varepsilon\,\varepsilon^5 \Phi}{\int d\varepsilon\,\varepsilon^4 \Phi}.
 \end{equation}
@@
 through the envelope.
+
 In order to estimate $\langle \varepsilon \rangle^{\rm sc}$ analytically, we may
@@
 
-Pinching parameters (calculated by eye) for the species
-$\{\nu_e,\bar{\nu}_e,\nu_x\}$ are $\alpha\approx\{3.6,5.1,2.3\}$,
-when we use the average energies from the \emph{spectral}
-\emph{scat} method in Tab.~\ref{tab:nsns_rt_vs_m1}.
+Pinching parameters
+(calculated by eye from the \emph{spectral} \emph{scat} method)
+for the species
+$\{\nu_e,\bar{\nu}_e,\nu_x\}$ are $\alpha\approx\{3.6,5.1,2.3\}$.
 Energy moments of pinched spectra like those in Eqn.~\ref{eqn:avg_eps_sc}
@@
 These analytic predictions agree with the average energies of the scattering
-envelope to within 20\%
+envelope to approximately 10\%
 % From by-eye examinations of <eps>(cosA) plot using data from
@@
 %          nue     nua     nux
-of the Monte Carlo and the \emph{scat} ray tracing calculations,
-even though we have made some drastic simplifications to our model.
+of all of the treatments including elastic scattering presented
+in Tab.~\ref{tab:collapse_rt_vs_mc}, except for
+$\bar{\nu}_e$ in the \emph{gray} \emph{scat} treatment,
+which deviates from our prediction by 30\%.
+This agreement is excellent despite the drastic
+simplifications used in our toy model.
+
+The $\langle\varepsilon_{\bar{\nu}_e}\rangle^{\rm sc}$ prediction
+in the \emph{gray} \emph{scat} treatment is approximately 30\%
+larger than the the fiducial Monte Carlo estimate. This is
+due to the large negative local chemical potential
+$\eta_{\bar{\nu}_e}\sim-10$ in the scattering envelope.
+As described in Sec.~\ref{ssec:sources_se},
+we use $\eta_\nu$ in the \emph{gray} treatment to construct our
+synthetic spectrum from total neutrino densities $J$ and $G$.
+This error points the way to future improvements to the \emph{gray} method.
 
@@
   \begin{equation}
+    \label{eqn:spectral_temp}
     T_\nu = \langle \varepsilon \rangle

### 2017-07-25 paper.tex
Correct typos

@@
 postmerger configuration composed of a single neutron star surrounded by a disk.
-Because of the strong differential rotation of the remnant it temporarily avoids
+Because of strong differential rotation, the remnant temporarily avoids
 collapse to a black hole, even if its mass exceeds the threshold of dynamical
@@
 isolated gravitational mass 1.2~$M_{\odot}$. The postmerger configuration
-was simulated using an M1 transport scheme for the neutrinos, evolving the
+was simulated using a gray M1 transport scheme for the neutrinos, evolving the
 energy density, number density, and energy flux \cite{fouc2016-m1_evolve_n}

### 2017-07-26 paper.tex
Add HMNS test text

@@
 its derivatives $\psi_{\alpha\beta,\gamma}$,
+\todo{state num independent components}
 and the hydrodynamical data $W$, $u_i$, $\varrho$, $T$, and $Y_e$.
@@
 ranges spanning $\varepsilon_i\in[1,280.5]$~MeV logarithmically,
-$\varrho\in[10^6,6.3\times10^15]\,{\rm g}\,{\rm cm}^{-3}$ logarithmically,
+$\varrho\in[10^6,6.3\times10^{15}]\,{\rm g}\,{\rm cm}^{-3}$ logarithmically,
 $T\in[0.05,200]$~MeV logarithmically, and 
@@
 Such a configuration was modeled in \cite{fouc2016-m1_nsns} by
-evolving the final inspiral and merger of two identical neutron stars of
+evolving fluid and spacetime through the
+final inspiral and merger of two identical neutron stars of
 isolated gravitational mass 1.2~$M_{\odot}$. The postmerger configuration
@@
 
-We use a single time snapshot of the volume data from that configuration
-at $t=11\,{\rm ms}$ after merger (see
-Figs.~\ref{fig:nsns_rho_merid}--\ref{fig:nsns_temp_equat})
-sampling the neutrino fields with ray tracing to compare to some
-measures of the field computed from the M1 transport simulation.
-\todo{here focus on plots as functions of $\theta$}
+We use a single time snapshot from that configuration
+at $t=11\,{\rm ms}$ after merger.
+The volume data are represented on a pseudospectral domain of a
+sphere with concentric shells extending to $r\approx1400$~km.
+Radial grid spacings are $\Delta r \approx 0.15$~km in the star
+and $\Delta r \approx 2.5$~km in the disk,
+with 12 cells spanning polar angles $\theta\in[0,\pi]$
+and 24 cells spanning azimuthal angles $\phi\in[0,2\pi]$.
+\todo{way too-low angular resolution}
+Though a pseudospectral representation of non-smooth hydrodynamic data
+introduces some Gibbs-like oscillations in the variables,
+we choose to use this representation instead of the mesh-refined
+finite difference grid of the evolution because the pseudospectral
+representation uses much less memory:
+95~MB in pseuspectral vs.\ 3~GB in finite difference representation.
+Obviously, memory loads of this order are not insurmountable;
+but they would require some modifications to our
+volume data interpolation infrastructure.
+For the purposes of this analysis the pseudospectral representation is adequate.
+Figs.~\ref{fig:nsns_rho_merid}--\ref{fig:nsns_temp_equat}
+show slices of density and temperature from the finite difference volume data.
 
@@
 
-In Tab.~\ref{tab:nsns_rt_vs_m1} we present total and relative number and
-energy luminosities for this configuration.
-With scattering turned off, ray tracing errors are as large as 250\%
-(for heavy lepton energy luminosities).
-But with scattering turned on, agreement is better than 20\%, and
-in most cases much better.
-\todo{get accurage M1 numbers from Francois}
+Opacities are computed from the LS220 equation of state using
+\lstinline{NuLib} and are stored as a table covering
+energy, density, temperature, and composition ranges
+spanning $\varepsilon\in[1,280.5]$~MeV logarithmically,
+$\varrho\in[10^6,3.2\times10^{15}]\,{\rm g}\,{\rm cm}^{-3}$ logarithmically,
+$T\in[0.05,150]$~MeV logarithmically, and 
+$Y_e\in[0.035,0.55]$ linearly,
+with grid extents$\{18,82,65,51\}$ respectively.
+
+We place stationary (i.e.\ Eulerian) observers at fixed coordinate radius
+$r=250$~km in the $y$-$z$ plane along an arc at $N_\theta$ positions
+distributed linearly in $\cos\theta$ over the northern hemisphere
+using $\theta_i=\cos^{-1}\left(1-i/(N_\theta-1)\right)$
+for $i=0,1,2,\ldots N_\theta-1$.
+We choose $N_\theta=30$.
+
+Though the ray tracing sampling of $f$ is done in full general relativity,
+in our calculation of moments at the observer positions,
+we make the simplifying assumption of Minkowski spacetime.
+This assumption introduces errors in our moments of order
+$GM/r=4\,{\rm km}/250\,{\rm km}$
+\todo{too simplistic}
+(with $G$ the gravitational constant and $M$ the central mass),
+or $\sim 1\%$.
+
+Each observer samples the distribution function over a uniform grid
+in energy, cosine of polar angle, and azimuthal angle with extents
+$\{N_{\tilde{\varepsilon}},N_A, N_B\}$ spanning the ranges
+$\tilde{\varepsilon}\in(0,100)$~MeV,
+$\cos A\in\left((\cos A)_{\rm min}, 1\right)$, and
+$B\in[0,2\pi)$.
+Using the midpoint rule,
+the radial fluxes of number and energy from Eqns.~\ref{eqn:Ha}
+and \ref{eqn:J_H_S_eps_integrated} become sums:
+\begin{align}
+  \label{eqn:Kr_minkowski}
+  \tilde{K}^r &= \Delta
+  \sum\limits_{m=0}^{N-1}
+  \tilde{\varepsilon}_m^2 f_m,\\
+  \label{eqn:Hr_minkowski}
+  \tilde{H}^r &= \Delta
+  \sum\limits_{m=0}^{N-1}
+  \tilde{\varepsilon}_m^3 f_m,
+\end{align}
+with $f_m \equiv f(\tilde{\varepsilon}_m,\cos A_m, B_m)$,
+$\Delta \equiv \Delta\tilde{\varepsilon}\, \Delta(\cos A)\, \Delta B$, and
+$N \equiv N_{\tilde{\varepsilon}}\, N_A\, N_B$.
+Average energies are given by
+$\langle\tilde{\varepsilon}\rangle(\theta)=H^r(\theta)/K^r(\theta)$.
+We choose $N_{\tilde{\varepsilon}}=30$, $N_A=120$, and $N_B=20$.
+For the \emph{scat} case we choose $(\cos A)_{\rm min}=0.81$;
+For the \emph{noscat} case, presenting a smaller emitting region,
+we choose $(\cos A)_{\rm min}=0.91$.
+
+We also combine measurements from all observers to estimate
+total luminosities and averages over the sky.
+We assume the disk is azimuthally symmetric and
+symmetric with respect to reflection across the equatorial plane.
+Figs.~\ref{fig:nsns_rho_merid} and \ref{fig:nsns_rho_equat} show how these
+assumptions are approximately valid at 11~ms after merger.
+Total luminosities are then computed as integrals of the radial fluxes
+over $\cos\theta$.
+Using the trapezoid rule the number and energy luminosities become sums:
+\begin{align}
+  \label{eqn:luminosity_R}
+  R &= C
+  \left(\frac{1}{2}(\tilde{K}^r_0+\tilde{K}^r_{N_\theta-1})
+  +\sum\limits_{m=1}^{N_\theta-2}\tilde{K}^r_m\right),\\
+  \label{eqn:luminosity_L}
+  L &= C
+  \left(\frac{1}{2}(\tilde{H}^r_0+\tilde{H}^r_{N_\theta-1})
+  +\sum\limits_{m=1}^{N_\theta-2}\tilde{H}^r_m\right),
+\end{align}
+with $C \equiv 4\pi r^2\, \Delta(\cos \theta)$,
+\todo{source of GR error in coordinate $r$}
+$\tilde{K}^r_m \equiv \tilde{K}^r(\theta_m)$, and
+$\tilde{H}^r_m \equiv \tilde{H}^r(\theta_m)$.
+Average energies over the whole sky are then
+$\langle\tilde{\varepsilon}\rangle=L/R$.
 
@@
   \caption{
-    Comparison of total luminosities and average energies between
-    ray tracing using the \emph{noscat} method `RT noscat',
+    Comparison of total luminosities and average energies
+    of the hypermassive neutron star configuration.
+    The methods are ray tracing using the \emph{noscat} method `RT noscat',
     ray tracing using the \emph{gray} \emph{scat} method `RT scat',
@@
     The ray tracing totals were computed from sums over
-    observers placed in the $y$-$z$ plane along an arc at coordinate distance
-    $r=250\,{\rm km}$ from the center of the star, at 30 positions
-    distributed evenly in $\cos\theta$ over the colatitude range
-    $\theta\in[0,90^{\circ}]$.
+    observers placed in the $y$-$z$ plane.
     M1 values are taken from \cite[Figs. 7, 9, 10]{fouc2016-m1_evolve_n}.
@@
 
-\section{Applications}
-\label{sec:applications}
-
-\subsection{Understanding Emission Spectra}
-\label{ssec:spectra}
-In this section, we use ray tracing's ability to track ``where the neutrinos
-are coming from'' to understand the features in the emission distributions
-presented in the code tests.
-
-The merger models are described in \cite{fouc2015-m1_nsbh, fouc2016-m1_nsns}.
-Something like this is presented in \cite[Figs.~10-11]{pere2014-nu_wind};
-compare.
-\todo{here focus on plots of $J(\cos A)$, $H^r(\cos A)$, and $J(\varepsilon)$}
+Tab.~\ref{tab:nsns_rt_vs_m1} compares the all-sky
+\todo{get accurate M1 numbers from Francois}
+luminosities and average energies from ray tracing and from
+the M1 transport simulation,
+which serves as our fiducial calculation.
+Because the model was evolved with a gray M1 transport scheme,
+the only \emph{scat} method we can use in our ray tracing is the
+\emph{gray} method.
+As in the post-bounce configuration in Sec.~\ref{ssec:test_collapse},
+the \emph{scat} treatment is more faithful than the \emph{noscat} treatment,
+significantly so for the heavy-lepton neutrinos which are most affected.
+With scattering turned off ray tracing overestimates $L_{\nu_x}$ by 250\%;
+but with scattering turned on agreement in all measures
+is better than 20\%, and in most cases much better.
+Note that $R_{\bar{\nu}_e}$ dominates over $R_{\nu_e}$
+\todo{Need $R$ magnitude not just difference}
+because the disrupted neutron star material is releptonizing
 
@@
 
+The distributions of radial fluxes over observer position are shown
+in Figs.~\ref{fig:nsns_theta_distrib_R} and \ref{fig:nsns_theta_distrib_L}.
+With the \emph{noscat} treatment the $\nu_x$ luminosities are relatively
+constant in $\theta$, because the disk is optically thin to $\nu_x$,
+and most of the neutrinos come from the hypermassive neutron star,
+which is roughly spherical.
+The slight upward trend in $\nu_x$ luminosities with $\theta$ may be due to
+asymmetries in the configuration.
+It is not due to the Doppler shift from the rapid rotation of the star
+(a hypothesis we tested by setting $u_i=0$ and $W=1$ in the ray
+tracing equations).
+Because of their larger absorption cross section,
+the disk is not optically thin to $\nu_e$ and $\bar{\nu}_e$,
+and observers within view of the hot hypermassive neutron star measure
+the largest radial fluxes.
+
+When we turn on scattering, the disk is no longer transparent to $\nu_x$,
+and $\nu_x$ luminosities present the same qualitative $\theta$-dependence
+as that of the other species.
+In general, scattering increases luminosities for polar observers
+($\theta \sim 0^\circ$) because additional high energy neutrinos
+from the hypermassive neutron star are scattered by the disk back up to
+the observer.
+For observers in the equatorial plane ($\theta \sim 90^\circ$)
+the effect of scattering is a decrease in energy luminosities,
+since some of the highest energy neutrinos from deep in the interior
+are scattered away from the line of sight by the disk.
+
 \begin{figure}
@@
 \end{figure}
+
+Fig.~\ref{fig:nsns_theta_distrib_avg_eps} shows the distribution of
+average energies of radial fluxes over observer positions.
+With and without scattering, the average energies are highest
+for observers near the polar axis, since polar observers get a
+direct view of the hot hypermassive neutron star.
+Scattering decreases average energies across all observer positions,
+as it did in the post-bounce configuration, or any configuration of
+a hot interior surrounded by a scattering envelope.
+The strength of the effect of scattering on the average energies of
+the different species is seen to follow the ranking
+$\nu_x>\bar{\nu}_e>\nu_e$ due to two factors:
+the average energies of the spectra obey the same ranking,
+and the thicknesses of the different species' scattering envelopes
+also obey the same ranking.
+As we showed in the optically thin limit of
+Eqn.~\ref{eqn:fse_homogeneous_thin_limit},
+\todo{too simplistic}
+the scattering contribution $f_{\rm SE}$ is
+proportional to energy $\varepsilon^2$ and path length $s$.
+
+\section{Applications}
+\label{sec:applications}
+
+\subsection{Understanding Emission Spectra}
+\label{ssec:spectra}
+In this section, we use ray tracing's ability to track ``where the neutrinos
+are coming from'' to understand the features in the emission distributions
+presented in the code tests.
+
+The merger models are described in \cite{fouc2015-m1_nsbh, fouc2016-m1_nsns}.
+Something like this is presented in \cite[Figs.~10-11]{pere2014-nu_wind};
+compare.
+\todo{here focus on plots of $J(\cos A)$, $H^r(\cos A)$, and $J(\varepsilon)$}
 

### 2017-07-26 paper.tex
Cite Mirizzi+2016, Keil+2003

@@
 
-In order to estimate $\langle \varepsilon \rangle^{\rm sc}$ analytically, we may
-approximate the background field $\Phi(\varepsilon)$ analytically.
+In order to estimate $\langle \varepsilon \rangle^{\rm sc}$ analytically,
+we must write the background field $\Phi(\varepsilon)$ analytically.
 Direct Fermi-Dirac fits using a temperature and chemical potential
@@
 Phenomenological Fermi-Dirac fits work well;
-but so do pinched spectral fits which are much simpler \cite{miri????}:
+but so do pinched spectral fits which are much simpler
+\cite{keil2003-pinched_spectra,miri2016-sn_neutrinos}:
 \begin{equation}
@@
 \end{equation}
-(where our definition differs from that of \cite{miri????} by a factor of
+(where our definition differs from that of \cite{keil2003-pinched_spectra}
+by a factor of
 $\varepsilon^2$, since we define our distribution function to be dimensionless

### 2017-07-26 paper.tex
Change symbol for a constant c->C

@@
 and in the \emph{gray} method we use
-$J=c T^4 \mathscr{F}_3(\eta_\nu)/(2\pi^2)$ and
-$G=c T^3 \mathscr{F}_2(\eta_\nu)/(2\pi^2)$,
+$J=C T^4 \mathscr{F}_3(\eta_\nu)/(2\pi^2)$ and
+$G=C T^3 \mathscr{F}_2(\eta_\nu)/(2\pi^2)$,
 using the Fermi integrals defined in Eqn.~\ref{eqn:fermi_integral},
-and with $c$ the conversion constant from ${\rm energy}^3$ to ${\rm length}^{-3}$.
+and with $C$ the conversion constant from ${\rm energy}^3$ to ${\rm length}^{-3}$.
 

### 2017-07-26 paper.tex
Revert float placement forcing Here
This reverts part of commit b9bdbc4f.
Update theta_distrib figures for HMNS test

@@
 \usepackage[mathscr]{euscript}    % for \mathscr
-\usepackage{float}                % for better control of figure placement
 \usepackage{graphicx}             % for \includegraphics
@@
 \renewcommand\todo[1]{} % if you wish to hide all todo notes
-
-% force all figures to be placed right where they are in the text
-\floatplacement{figure}{H}
 
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{theta_distrib-250km-luminosity_R}
-  % Data:   $DIR/PositionDistributions
+  \includegraphics[width=\columnwidth]{theta_distrib-250km-Kr}
+  % Data_a:   $DIR/PositionDistributions/*Lev1-e30a120b20-01*
+  % Data_s:   $DIR/PositionDistributions/*Lev1-e30a60b10-01*
   % Data gathered using: $DIR/CollectPositionDistributions.sh
-  % Script: $DIR/GnuScripts/plot-theta_distributions-R-fixed_r.gnu
+  % Script: $DIR/GnuScripts/plot-theta_distributions-Hr-fixed_r.gnu
   % w/ DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20170324-ray_tracing_nsns/Run03
-  \caption{Radial number fluxes, $K_r$, as functions of colatitude $\theta$.
+  \caption{Radial number fluxes, $K^r(\theta)$,
+    as defined in Eqn.~\ref{eqn:Kr_minkowski},
+    in the hypermassive neutron star configuration.
     Sampled for observers at fixed coordinate distances from the center
     of the star, with $r=250\,{\rm km}$.
-    Here radial fluxes are multiplied by the portion of solid angle
-    represented by each observer, so that the total luminosity is
-    computed by summing all the points.}
-  \label{fig:nsns_theta_distrib_R}
+  }
+  \label{fig:nsns_theta_distrib_Kr}
 \end{figure}
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{theta_distrib-250km-luminosity_L}
-  % see notes from theta_distrib-250km-luminosity_R
-  % Script: $DIR/GnuScripts/plot-theta_distributions-L-fixed_r.gnu
-  \caption{Same as Fig.~\ref{fig:nsns_theta_distrib_R} but radial energy
-    fluxes, $H_r$.}
-  \label{fig:nsns_theta_distrib_L}
+  \includegraphics[width=\columnwidth]{theta_distrib-250km-Hr}
+  % see notes from theta_distrib-250km-Kr
+  % Script: $DIR/GnuScripts/plot-theta_distributions-Hr-fixed_r.gnu
+  \caption{Same as Fig.~\ref{fig:nsns_theta_distrib_Kr} but radial energy
+    fluxes $H^r(\theta)$ as defined in Eqn.~\ref{eqn:Hr_minkowski}.
+  }
+  \label{fig:nsns_theta_distrib_Hr}
 \end{figure}
@@
 The distributions of radial fluxes over observer position are shown
-in Figs.~\ref{fig:nsns_theta_distrib_R} and \ref{fig:nsns_theta_distrib_L}.
+in Figs.~\ref{fig:nsns_theta_distrib_Kr} and \ref{fig:nsns_theta_distrib_Hr}.
 With the \emph{noscat} treatment the $\nu_x$ luminosities are relatively
@@
   \includegraphics[width=\columnwidth]{theta_distrib-250km-avg_eps}
-  % see notes from theta_distrib-250km-luminosity_R
-  % Script: $DIR/GnuScripts/plot-theta_distributions-avg_eps-fixed_r.gnu
-  \caption{Same as Fig.~\ref{fig:nsns_theta_distrib_R} but
-    average energies $\langle \varepsilon \rangle = J/G$.}
+  % see notes from theta_distrib-250km-Kr
+  % Script: $DIR/GnuScripts/plot-theta_distributions-avg_eps_from_fluxes-fixed_r.gnu
+  \caption{Same as Fig.~\ref{fig:nsns_theta_distrib_Kr} but
+    average energies of radial fluxes
+    $\langle\tilde{\varepsilon}\rangle(\theta)=H^r(\theta)/K^r(\theta)$,
+    as defined in Eqns.~\ref{eqn:Kr_minkowski} and \ref{eqn:Hr_minkowski}.
+  }
   \label{fig:nsns_theta_distrib_avg_eps}

### 2017-07-26 paper.tex
Update table of HMNS total luminosities
Update text to reflect new luminosities in HMNS table

@@
   % RTDATA_SP=$DIR/ResultsBrett-A-03/Test-scat-01/
+  % DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20161007-ray_tracing_tests/CollapseProfile/
   \caption{
@@
 \begin{table}%[h]
+  % Total fluxes over the sky integrated in google sheet: 20170726-theta_distribution_integrations
+  % RTDATA_NO=$DIR/PositionDistributions-nu?-250km-Lev1-e30a120b20-noscat-01.dat
+  % RTDATA_GR=$DIR/PositionDistributions-nu?-250km-Lev1-e30a60b10-scat-01.dat
+  % DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20170324-ray_tracing_nsns/Run03/PositionDistributions
   \caption{
@@
     \hline
-    $L_{\nu_e}$                                 & 5.95 & 6.01 & 5 \\
+    $L_{\nu_e}$                                 & 5.62 & 6.05 & 5 \\
     $L_{\bar{\nu}_e}$                           & 10.6 & 11.8 & 12 \\
-    $L_{\nu_x}$                                 & 30.2 & 12.6 & 12 \\
-    $R_{\nu_e}-R_{\bar{\nu}_e}$                 & -1.06 & -1.86 & -2 \\
-    $\langle \varepsilon_{\nu_e} \rangle$       & 12.08 & 11.88 & 12 \\
-    $\langle \varepsilon_{\bar{\nu}_e} \rangle$ & 15.96 & 14.69 & 15 \\
-    $\langle \varepsilon_{\nu_x} \rangle$       & 36.25 & 23.67 & 26 \\
+    $L_{\nu_x}$                                 & 23.3 & 12.7 & 12 \\
+    $R_{\nu_e}-R_{\bar{\nu}_e}$                 & -1.22 & -1.85 & -2 \\
+    $\langle \varepsilon_{\nu_e} \rangle$       & 12.08 & 11.91 & 12 \\
+    $\langle \varepsilon_{\bar{\nu}_e} \rangle$ & 16.07 & 14.67 & 15 \\
+    $\langle \varepsilon_{\nu_x} \rangle$       & 34.09 & 23.58 & 26 \\
     \hline
@@
 the \emph{scat} treatment is more faithful than the \emph{noscat} treatment,
-significantly so for the heavy-lepton neutrinos which are most affected.
-With scattering turned off ray tracing overestimates $L_{\nu_x}$ by 250\%;
+significantly so for the heavy-lepton neutrinos.
+(Though, we note the energy luminosity for $\nu_e$ is overestimated
+20\% by both \emph{noscat} and \emph{scat} treatments,
+and the overestimate is slightly worse for the \emph{scat} treatment.)
+\todo{explain}
+With scattering turned off ray tracing overestimates $L_{\nu_x}$ by 200\%;
 but with scattering turned on agreement in all measures
-is better than 20\%, and in most cases much better.
-Note that $R_{\bar{\nu}_e}$ dominates over $R_{\nu_e}$
+is better than 10\%, except for $L_{\nu_e}$ which is overestimated by 20\%.
+As the M1 transport evolution also demonstrates,
+$R_{\bar{\nu}_e}$ dominates over $R_{\nu_e}$
 \todo{Need $R$ magnitude not just difference}
-because the disrupted neutron star material is releptonizing
+because the disrupted neutron star material is releptonizing.
 

### 2017-07-30 paper.tex
Add oscillation application text

@@
 % packages
+\usepackage{amsmath}              % for \text, align, split ...
 \usepackage{amssymb}              % for \gtrsim ...
-\usepackage{amsmath}              % for \text, align, split ...
+\usepackage{braket}               % for \bra \ket
 \usepackage{color}                % for colorful margin notes
@@
   These background fields may be taken from a low-order moment simulation
-  or be ignored entirely, in which case the method
+  or be ignored, in which case the method
   reduces to a standard state-of-the-art ray tracing formulation.
   It handles radiation in regimes spanning optically thick to optically thin.
-  We test the new code, highlight its strengths and weaknesses, and
-  apply it to simulations of neutron star mergers to
-  compute neutrino fluxes as a function of energy and emission angle.
+  We test the new code, highlight its strengths and weaknesses,
+  and apply it to a simulation of a neutron star merger
+  to compute neutrino fluxes and spectra,
+  and to demonstrate a neutrino flavor oscillation calculation.
 \end{abstract}
@@
 In Sec.~\ref{sec:tests} we present tests of the code.
-In Sec.~\ref{sec:applications} we present several applications
-in the dynamical environment following the merger of two neutron stars:
-computing neutrino spectra as a function of observer's angle,
-computing the self-interaction potential descriptive of flavor evolution
-in high-neutrino-density environments,
-and computing the energy dumped into jet-forming electron-positron
-pairs via neutrino-antineutrino annihilation.
+In Sec.~\ref{sec:applications} we present an application
+in the dynamical environment following the merger of two neutron stars,
+computing neutrino flavor oscillation along an outgoing ray,
+including the effects of coherent forward scattering with ambient neutrinos.
 
@@
 behavior of the right hand side, driving $f$ toward $\mathscr{S}$
-over a lengthscale $\mathscr{K}$ in our affine parameter;
+over a lengthscale $\mathscr{K}^{-1}$ in our affine parameter;
 or from Eqn.~\ref{eqn:proper_length}, over a proper lengthscale
@@
 To test our algorithm, we integrate the ray tracing equations
-(Eqns.~\ref{eqn:num_x}--\ref{eqn:num_tau}) at one observation position
+(Eqns.~\ref{eqn:num_x}--\ref{eqn:num_tau}) for various observers
 in the following configurations. This suite of configurations
@@
 
-\section{Applications}
-\label{sec:applications}
-
-\subsection{Understanding Emission Spectra}
-\label{ssec:spectra}
-In this section, we use ray tracing's ability to track ``where the neutrinos
-are coming from'' to understand the features in the emission distributions
-presented in the code tests.
-
-The merger models are described in \cite{fouc2015-m1_nsbh, fouc2016-m1_nsns}.
-Something like this is presented in \cite[Figs.~10-11]{pere2014-nu_wind};
-compare.
-\todo{here focus on plots of $J(\cos A)$, $H^r(\cos A)$, and $J(\varepsilon)$}
+In Figs.~\ref{fig:nsns_cosA_distrib_I_00deg}
+and \ref{fig:nsns_cosA_distrib_I_90deg}, we show the distributions of neutrino
+number density over incoming angle
+for the observer on the polar axis $\theta=0^\circ$,
+and in the equatorial plane $\theta=90^\circ$,
+as defined in Eqn.~\ref{eqn:J_per_cosA}.
+As in the case of the collapse profile in Sec.~\ref{ssec:test_collapse},
+the dominant effect of \emph{scat} on $\tilde{G}(\cos A)$
+is to spread the distribution out to larger angles.
+\todo{also examine spectra or average energies}
 
@@
 
-\subsection{Neutrino Self-Interaction Potential}
+\section{Applications}
+\label{sec:applications}
+
+\subsection{Neutrino Oscillations at High Neutrino Densities}
 \label{ssec:V_nunu}
-In this section we present $V_{\nu\nu,0}$ along one test trajectory in the NSNS
-model, and show that including elastic scattering makes a qualitative
-difference in the character of the flavor evolution.
-See Figs.~\ref{fig:V_nunu-noscat} and \ref{fig:V_nunu-scat}.
+The Boltzmann Equation (Eqn.~\ref{eqn:boltzmann}) is one limiting
+form of the quantum kinetic equations \cite{vlas2014-neutrino_qkes},
+the limit where collisional mean free paths are much shorter
+than oscillation lengths (i.e.\ $\varepsilon/\mathscr{K} \ll |H|^{-1}$
+with $H$ the Hamiltonian describing coherent forward scattering interactions).
+In this limit the neutrino density matrix,
+\begin{equation}
+  \label{eqn:density_rho_0}
+  \rho_0(\varepsilon,\ell_\mu) =
+  \sum\limits_{\alpha}f^{\nu_\alpha}(\varepsilon,\ell_\mu)
+  \ket{\nu_\alpha} \bra{\nu_\alpha},
+\end{equation}
+with $\ket{\nu_\alpha}$ the neutrino flavor eigenstates,
+remains in a pure state, or diagonal in the flavor basis.
+
+But at the opposite limit, in the free-streaming regime,
+coherence effects become important,
+and the quantum kinetic equations take a Schr\"{o}dinger-like form:
+\begin{equation}
+  \label{eqn:schroedinger}
+  i \frac{d}{ds} S = H S,
+\end{equation}
+where $ds$ is an interval of proper length, and
+\todo{mention $\bar{\nu}$ eqn}
+$S$ is the flavor evolution matrix describing a mixed state
+\begin{equation}
+  \label{eq:density_evolution}
+  \rho = S \, \rho_0 \, S^\dagger.
+\end{equation}
+In this limit, we may decompose the Hamiltonian into vacuum,
+matter, and neutrino contributions:
+\begin{equation}
+  \label{eqn:decompose_H}
+  H = H_{\rm V} + H_e + H_{\nu\nu}.
+\end{equation}
+Explicit formulas for these terms may be found in the oscillation literature,
+e.g.\ \cite{duan2009-review},
+but for simplicity we give their respective scales:
+\begin{align}
+  \label{eqn:scale_H_V}
+  |H_{\rm V}| &\sim \frac{\Delta m^2}{\varepsilon}, \\
+  \label{eqn:scale_H_e}
+  |H_e|       &\sim G_{\rm F}\left|n_{e^-}-n_{e^+}\right|, \\
+  \label{eqn:scale_H_nunu}
+  |H_{\nu\nu}(\ell_\mu)|&\sim G_{\rm F} \left|
+  (G_{\nu_e}-G_{\bar{\nu}_e}) - \ell_\mu(K_{\nu_e}^\mu-K_{\bar{\nu}_e}^\mu)
+  \right|,
+\end{align}
+where $\Delta m$ is the mass difference between neutrino mass eigenstates,
+$G_{\rm F}$ is the Fermi coupling constant,
+$n_{e^-}$ and $n_{e^+}$ are the electron and positron number densities,
+and $\ell_\mu$ is the propagation direction of the
+coherent beam of neutrinos in question.
+
+In regimes in which $|H_{\nu\nu}|\ll|H_{\rm V},H_e|$,
+the flavor evolution along the ray is soluble, and reveals a rich
+and physically important phenonenology including
+vacuum, solar, atmospheric, and terrestrial oscillations,
+as well as oscillations in supernova envelopes.
+Where neutrino densities are relatively high, however,
+as in neutron star mergers, the problem must be solved globally.
+\todo{be clearer}
+To date, no method has been devised to handle this problem.
+
+However, we can solve a similar, but tractable problem along a single ray.
+Assume all the neutrino rays intersecting an event along a given test ray
+have undergone the same flavor evolution history as that of the test ray,
+i.e.\ $S$ is the same for all rays sharing that event.
+This is the so-called single-angle approximation.
+\todo{cite someone}
+\todo{say something about its validity}
+
+Solving for the flavor evolution history along this test ray
+requires knowledge of the unoscillated neutrino contribution to the Hamiltonian.
+This is a function of the unoscillation neutrino self-interaction potentials,
+which for the $\alpha$-th flavor is:
+\begin{equation}
+  \label{eqn:V_nu_alpha}
+  V_{\nu_\alpha,0}(\varepsilon,\ell_\mu) =
+  \frac{\sqrt{2}G_{\rm F}}{(2\pi)^3} \varepsilon^2
+  \oint d\Omega'(1-\varpi')f^{\nu_\alpha}(\varepsilon,\ell'_\gamma),
+\end{equation}
+with the cosine of the angle between the test ray and ambient rays
+given by $\varpi'=\psi^{\mu\gamma} \ell_\mu \ell'_\gamma$.
+
+If conditions are right, a resonant flavor transition
+introducing significant mixing may occur very near
+the point where neutrinos begin free-streaming.
+The matter-neutrino resonance \cite{malk2012-mnr_1,malk2015-mnr_2,malk2016-mnr_3},
+may occur along a given ray wherever the matter and neutrino self-interaction
+potentials cancel, or $V_e \sim V_{\nu\nu,0}$, with
+\begin{align}
+  \label{eqn:V_e}
+  V_e &= \sqrt{2} G_{\rm F} \frac{\varrho}{m_N} Y_e, \\
+  \label{eqn:V_nunu0}
+  V_{\nu\nu} &= \int d\varepsilon
+  \big(V_{\nu_e,0}(\varepsilon)-V_{\bar{\nu}_e,0}(\varepsilon)\big),
+\end{align}
+with $\varrho$ the rest density and $m_{N}$ the nucleon mass.
+
+We calculate the self-interaction potential at $N_s=5$ points between
+$r\in[30,135]$~km,
+by sampling distribution functions over a grid with extents
+$N_\varepsilon=30$, $N_A=30$, $N_B=10$,
+over all angles and an energy range of $\varepsilon\in(0,100)$~MeV.
+\todo{improve resolution and update}
+We interpolate the logarithm of the self-interaction potentials
+in path length $\log s$ along the ray,
+by fitting a 3rd order spline with continuous derivative,
+across the $N_s$ observation points.
+The matter potential along this test ray is from an analytic wind
+model qualitatively consistent with the densities in the
+simulated volume (i.e.\ inside $r\sim70$~km).
+
+In Figs.~\ref{fig:V_nunu-noscat} and \ref{fig:V_nunu-scat}
+we show the self-interaction potential and its contributions from
+$\nu_e$ and $\bar{\nu}_e$
+for a test neutrino ray moving out along a coordinate radial trajectory
+from the hypermassive neutron star used in Sec.~\ref{ssec:test_disk_comparison},
+along an angle of $\theta=22^\circ$ with respect to the polar axis.
+\todo{note coordinate ray}
+Fig.~\ref{fig:V_nunu-noscat} shows these terms for the \emph{noscat} treatment,
+and Fig.~\ref{fig:V_nunu-scat} for the \emph{scat} treatment.
+Obviously including the effects of elastic scattering
+tends to increase the $\bar{\nu}_e$ contribution relative to the $\nu_e$
+contribution, causing the self-interaction potential to stay negative
+along the entire trajectory.
 
@@
 
+Evolving the flavor content of this system using the formalism described in
+\cite{zhu2016-mnr_nsns_remnant}
+we see that the matter-neutrino resonance occurs in both,
+but with the very different behaviors of symmetric and asymmetric
+resonances respectively for the \emph{noscat} and \emph{scat} cases,
+as seen in Figs.~\ref{fig:survival-noscat} and \ref{fig:survival-scat}.
+These two classes of matter-neutrino resonance
+are described e.g.\ in \cite{malk2016-mnr_3,vaan2016-uncovering_mnr}.
+The survival probabilities shown in these figures are the terms of
+the flavor evolution matrix $S$ that describe how much of the original flux
+of a given flavor survives a flavor transition.
+Where $P_{\nu_e\rightarrow\nu_e}$ vanishes, for example in
+Fig.~\ref{fig:survival-noscat}, all the original $\nu_e$ content of the beam
+is oscillated into $\nu_\mu$ and $\nu_\tau$; but the $\nu_e$ density doesn't
+vanish altogether,
+since some $\nu_\mu$ and $\nu_\tau$ oscillates into $\nu_e$.
+This dramatic difference in survival probabilities between the \emph{noscat}
+and \emph{scat} cases along this test ray emphasizes the importance of
+scattering physics in the problem of neutrino oscillations in
+neutron star mergers.
+
 \begin{figure}
@@
 
-\subsection{Neutrino Annihilation Integral}
-\label{ssec:q_nunu}
-In this section we calculate $q_{\nu\bar{\nu}}$ at a representative point
-in the funnel of the NSNS configuration and show that including scattering makes
-a difference (or not) in our estimate of the power available to drive a jet.
+%\subsection{Neutrino Annihilation Integral}
+%\label{ssec:q_nunu}
+%In this section we calculate $q_{\nu\bar{\nu}}$ at a representative point
+%in the funnel of the NSNS configuration and show that including scattering makes
+%a difference (or not) in our estimate of the power available to drive a jet.
 

### 2017-07-31 paper.tex
Simplfiy Applications section
Move "understanding nsns spectra" to nsns code test. Remove neutrino
annihilation section: I don't want to include that anymore, since
this is my last day of postdoc, and I don't want this paper to be
any larger than it has to be to get published.

--- 2017-07-26 paper.tex
@@
 
-\subsection{Neutrino Self-Interaction Potential}
+\section{Applications}
+\label{sec:applications}
+
+\subsection{Neutrino Oscillations at High Neutrino Densities}
 \label{ssec:V_nunu}

### 2017-08-01 paper.tex
Add edits from 7.31.17 meeting

--- 2017-07-30 paper.tex
@@
 \todo{consistent with Eqn.~\ref{eqn:def_direction}?}
-We will also make use of the number density and flux defined as
-$G(\varepsilon)=J(\varepsilon)/\varepsilon$
-and $K^\mu(\varepsilon)=H^\mu(\varepsilon)/\varepsilon$
-respectively.
-
-The energy-integrated moments are
+We also make use of the number density and flux defined
+\begin{align}
+  \label{eqn:G}
+  G(\varepsilon) &=
+  \frac{\varepsilon^2}{(2\pi)^3} \oint d\Omega' f(\varepsilon, \ell'_\beta) \\
+  \label{eqn:Ka}
+  K^\mu(\varepsilon) &=
+  \frac{\varepsilon^2}{(2\pi)^3} \oint d\Omega' f(\varepsilon, \ell'_\beta) \ell'^\mu.
+\end{align}
+
+The energy-integrated moments take the form
 \begin{equation}
   \label{eqn:J_H_S_eps_integrated}
-  \big\{ J,H^\mu,S^{\mu\nu} \big\} = \int_0^\infty d\varepsilon \,
-  \big\{ J(\varepsilon),H^\mu(\varepsilon),S^{\mu\nu}(\varepsilon) \big\},
-\end{equation}
-having dimension ${\rm energy}\,{\rm length}^{-3}$.
+  X = \int_0^\infty d\varepsilon \, X(\varepsilon),
+\end{equation}
+with $X$ standing in for any of $\{J,H^\mu,S^{\mu\gamma}\}$ or
+$\{G,K^\mu\}$,
+the first having dimension ${\rm energy}\,{\rm length}^{-3}$,
+and the second ${\rm length}^{-3}$.
 
@@
   \label{eqn:doppler_avg_eps}
-  \langle \tilde{\varepsilon} \rangle =
+  \langle \tilde{\varepsilon} \rangle(\cos A) =
   \langle \varepsilon \rangle^{\rm eq} \frac{1}{W(1-v \cos A)}.
@@
     The points are computed from ray tracing spectra; the lines are the
-    analytic formula Eqn.~\ref{eqn:avg_eps_per_cosA}.
+    analytic formula Eqn.~\ref{eqn:doppler_avg_eps}.
   }
@@
   % Data:   $DIR/
-  % Script: $DIR/plot-compare_spectral_vs_gray-angular_distribution_avg_eps-production.gnu
+  % Script: $DIR_GRAY_SC/../plot-compare_spectral_vs_gray-angular_distribution_avg_eps-production.gnu
   % w/  DIR_MC     =$DIR/ResultsSherwood-03/Spectra
@@
 We don't show it here, but the angular distribution contributing to the
+\todo{do show it}
 total number density is also strongly affected by elastic scattering.
@@
 % DIR=.../20161007-ray_tracing_tests/CollapseProfile/ResultsBrett-B-03
+
 Total luminosities and average energies from the different treatments are
@@
 presented in Tab.~\ref{tab:collapse_rt_vs_mc}.
+As expected, $\nu_e$ is least affected by scattering, and $\nu_x$ most.
+In fact, without scattering, $\nu_x$ luminosities are overestimated by more than
+two orders of magnitude, due to the steep increase in temperature with depth
+in the inner core.
+Though contributing only a fraction of the total luminosity,
+the average energy of neutrinos scattered to the observer from the envelope
+outside the shock, $\langle\varepsilon\rangle^{\rm sc}$,
+is poorly characterized by the \emph{noscat} treatment for all species.
+By contrast, both the \emph{gray} and \emph{spectral} \emph{scat} treatments
+faithfully describe the high scattered energies from the envelope.
 
@@
 synthetic spectrum from total neutrino densities $J$ and $G$.
-This error points the way to future improvements to the \emph{gray} method.
+This error and our successful analysis using pinched spectra above,
+points the way to future improvements to the \emph{gray} method:
+making better assumptions about the spectra which don't depend upon
+local fluid quantities.
 
@@
 $B\in[0,2\pi)$.
-Using the midpoint rule,
-the radial fluxes of number and energy from Eqns.~\ref{eqn:Ha}
-and \ref{eqn:J_H_S_eps_integrated} become sums:
+The integrals for fluxes of number and energy
+(Eqns.~\ref{eqn:Ha}, \ref{eqn:Ka}, and \ref{eqn:J_H_S_eps_integrated})
+become sums over all rays.
+Taking these fluxes in the radial direction,
+and using the midpoint rule to convert the integral to a sum, we have
 \begin{align}
@@
   \sum\limits_{m=0}^{N-1}
-  \tilde{\varepsilon}_m^2 f_m,\\
+  \tilde{\varepsilon}_m^2 \cos A_m f_m,\\
   \label{eqn:Hr_minkowski}
@@
   \sum\limits_{m=0}^{N-1}
-  \tilde{\varepsilon}_m^3 f_m,
+  \tilde{\varepsilon}_m^3 \cos A_m f_m,
 \end{align}
 with $f_m \equiv f(\tilde{\varepsilon}_m,\cos A_m, B_m)$,
-$\Delta \equiv \Delta\tilde{\varepsilon}\, \Delta(\cos A)\, \Delta B$, and
-$N \equiv N_{\tilde{\varepsilon}}\, N_A\, N_B$.
+$\Delta \equiv \Delta\tilde{\varepsilon}\, \Delta(\cos A)\, \Delta B$,
+$N \equiv N_{\tilde{\varepsilon}}\, N_A\, N_B$,
+and $m$ the index labeling each ray.
 Average energies are given by
@@
 which serves as our fiducial calculation.
+Uncertainties are around 15\% for energy luminosities and
+10\% for average energies from the M1 simulation
+\cite[Sec.~A.6]{fouc2016-m1_evolve_n}.
 Because the model was evolved with a gray M1 transport scheme,

### 2017-08-01 paper.tex
Improve text in the oscillation section

@@
 with $H$ the Hamiltonian describing coherent forward scattering interactions).
-In this limit the neutrino density matrix,
+In this limit the neutrino density matrix takes the form
 \begin{equation}
@@
 \end{equation}
-with $\ket{\nu_\alpha}$ the neutrino flavor eigenstates,
-remains in a pure state, or diagonal in the flavor basis.
+with $\ket{\nu_\alpha}$ the neutrino flavor eigenstates.
+The neutrino state remains pure, or diagonal in the flavor basis.
 
@@
 e.g.\ \cite{duan2009-review},
-but for simplicity we give their respective scales:
+but for clarity we give their order-of-magnitude scales here:
 \begin{align}
@@
   \label{eqn:scale_H_nunu}
-  |H_{\nu\nu}(\ell_\mu)|&\sim G_{\rm F} \left|
-  (G_{\nu_e}-G_{\bar{\nu}_e}) - \ell_\mu(K_{\nu_e}^\mu-K_{\bar{\nu}_e}^\mu)
-  \right|,
+  |H_{\nu\nu}|&\sim G_{\rm F} \left|G_{\nu_e}-G_{\bar{\nu}_e}\right|,
 \end{align}
@@
 $n_{e^-}$ and $n_{e^+}$ are the electron and positron number densities,
-and $\ell_\mu$ is the propagation direction of the
-coherent beam of neutrinos in question.
+and $G_{\nu_e}$ and $G_{\bar{\nu}_e}$ the electron neutrino and antineutrino
+number densities defined in Eqn.~\ref{eqn:G}.
+For the scale of $H_{\nu\nu}$ we have only included the isotropic
+components of the neutrino fields.
 
@@
 
-However, we can solve a similar, but tractable problem along a single ray.
+However, we can solve a similar but tractable problem along a single ray.
 Assume all the neutrino rays intersecting an event along a given test ray
@@
 \todo{say something about its validity}
-
-Solving for the flavor evolution history along this test ray
-requires knowledge of the unoscillated neutrino contribution to the Hamiltonian.
-This is a function of the unoscillation neutrino self-interaction potentials,
+Though widely used in the supernova oscillation literature
+and shown to be qualitatively faithful in those environments,
+\todo{cite vlasenko?}
+a recent study has shown that the single-angle approximation
+may mask certain instabilities in the flavor evolution
+\cite{wu2017-fast_neutrino_conversions}.
+
+Within the single-angle approximation,
+solving for the flavor evolution history along a given test ray
+requires knowledge of the unoscillated neutrino contribution to the Hamiltonian
+along that trajectory.
+This is a function of the unoscillated neutrino self-interaction potential,
 which for the $\alpha$-th flavor is:
@@
 \end{equation}
-with the cosine of the angle between the test ray and ambient rays
-given by $\varpi'=\psi^{\mu\gamma} \ell_\mu \ell'_\gamma$.
+with the test ray propagating in direction $\ell_\mu$,
+ambient rays propagating in directions $\ell'_\gamma$,
+and the cosine of the angle between these given by
+$\varpi'=\psi^{\mu\gamma} \ell_\mu \ell'_\gamma$.
+As in the moment equations (Eqns.~\ref{eqn:J}-\ref{eqn:Ka}),
+the integral is taken over all directions $\ell'_\gamma$.
 
@@
 the point where neutrinos begin free-streaming.
-The matter-neutrino resonance \cite{malk2012-mnr_1,malk2015-mnr_2,malk2016-mnr_3},
-may occur along a given ray wherever the matter and neutrino self-interaction
-potentials cancel, or $V_e \sim V_{\nu\nu,0}$, with
+The matter-neutrino resonance
+\cite{malk2012-mnr_1,malk2015-mnr_2,malk2016-mnr_3},
+can occur where the matter potential and
+neutrino self-interaction potentials cancel,
+or where $V_e - V_{\nu\nu,0}=0$, with
 \begin{align}
@@
 with $\varrho$ the rest density and $m_{N}$ the nucleon mass.
-
+The matter potential is always positive;
+and outside the accretion disks formed in neutron star mergers,
+in which the disrupted neutron-star matter is rapidly releptonizing,
+the neutrino self-interaction potential is large and negative.
+
+We examine this effect in the post-merger configuration already analyzed
+in Sec.~\ref{ssec:test_disk_comparison}.
 We calculate the self-interaction potential at $N_s=5$ points between
@@
 We interpolate the logarithm of the self-interaction potentials
+(Eqn.~\ref{eqn:V_nu_alpha})
 in path length $\log s$ along the ray,
@@
 
-Evolving the flavor content of this system using the formalism described in
-\cite{zhu2016-mnr_nsns_remnant}
-we see that the matter-neutrino resonance occurs in both,
-but with the very different behaviors of symmetric and asymmetric
-resonances respectively for the \emph{noscat} and \emph{scat} cases,
-as seen in Figs.~\ref{fig:survival-noscat} and \ref{fig:survival-scat}.
-These two classes of matter-neutrino resonance
-are described e.g.\ in \cite{malk2016-mnr_3,vaan2016-uncovering_mnr}.
-The survival probabilities shown in these figures are the terms of
+We also solve for the flavor evolution of this system,
+integrating $S$ according to Eqn.~\ref{eqn:schroedinger},
+as described in \cite{zhu2016-mnr_nsns_remnant}.
+In Figs.~\ref{fig:survival-noscat} and \ref{fig:survival-scat}
+we show the survival probabilities for $\nu_e$ and $\bar{\nu}_e$
+for the two treatments.
+The survival probabilities are the terms of
 the flavor evolution matrix $S$ that describe how much of the original flux
-of a given flavor survives a flavor transition.
+of a given flavor survives a transition.
+\todo{define explicitly}
 Where $P_{\nu_e\rightarrow\nu_e}$ vanishes, for example in
@@
 since some $\nu_\mu$ and $\nu_\tau$ oscillates into $\nu_e$.
+
+We see that the matter-neutrino resonance occurs in both treatments,
+but with the very different behaviors of symmetric and asymmetric
+resonances respectively for the \emph{noscat} and \emph{scat} cases,
+as seen in Figs.~\ref{fig:survival-noscat} and \ref{fig:survival-scat}.
+These two classes of matter-neutrino resonance
+are described for example in \cite{malk2016-mnr_3,vaan2016-uncovering_mnr}.
 This dramatic difference in survival probabilities between the \emph{noscat}

### 2017-08-01 paper.tex
Add edits from 7.25.17 meeting

@@
 or over a small volume outside a source.
-But the method becomes computationally burdensome when the inquiry extends
-to large volumes of spacetime.
 
@@
 
-\subsubsection*{Outline}
 In Sec.~\ref{sec:formulation} we derive the ray tracing equations from the
@@
 parameter defining the neutrino momentum (Eqn.~\ref{eqn:geodesic_x})
-and $C[f]$ is the source term arising from interactions with the medium,
-having dimension ${\rm energy}\,{\rm length}^{-1}$.
+and $C[f]$ is the source term arising from interactions with the medium.
+By Eqn.~\ref{eqn:geodesic_x},
+the affine parameter has dimension ${\rm length}\,{\rm energy}^{-1}$,
+so the source term has dimension ${\rm energy}\,{\rm length}^{-1}$.
 The source term varies over phase space $(x^\alpha,p_\beta)$,
-and due to Fermi blocking depends locally on the distribution function $f$
+and depends locally on the distribution function $f$
 and nonlocally on the distribution function of this neutrino and its
@@
 (i.e.\ Fermi-blocking).
-We consider two classes of interactions in this work:
+We consider the two dominant classes of interactions in this work:
 the absorption/emission (AE) and elastic scattering (SE) processes
@@
 over the relevant emissivities and opacities
-in Tab.~\ref{tab:neutrino_processes}:
-\todo{define $j$ and $\chi$}
+for the reactions
+\begin{align}
+  \nu_e + n            &\leftrightarrow e^- + p,\nonumber\\
+  \bar{\nu}_e + p      &\leftrightarrow e^+ + n,\nonumber\\
+  \nu_e + {}^AZ        &\leftrightarrow e^- + {}^A(Z+1),\nonumber\\
+  \bar{\nu}_e + {}^AZ  &\leftrightarrow e^+ + {}^A(Z-1),\nonumber
+\end{align}
+with ${}^ZA$ representing a nucleus of mass number $A$ and charge $Z$.
+In terms of the emissivity $j(\varepsilon)$ describing number of neutrinos
+of energy $\varepsilon$ emitted per length,
+and the absorption opacity $\chi_a(\varepsilon)$
+describing the number absorbed per length,
+the coefficients are
 \begin{align}
@@
 a background field, and a sum over opacities for
-the relevant reactions in Tab.~\ref{tab:neutrino_processes}:
+the reactions
+\begin{align}
+  \nu + N     &\leftrightarrow \nu + N,\nonumber\\
+  \nu + {}^AZ &\leftrightarrow \nu + {}^AZ,\nonumber
+\end{align}
+with $N$ standing in for either $n$ or $p$.
+In terms of a background field $\Phi(\varepsilon)$ describing
+the number of neutrinos of energy $\varepsilon$ present at this event,
+and the scattering opacity $\chi(\varepsilon)$ describing the
+number scattered to other directions per length,
+the coefficients are
 \begin{align}
@@
 \end{align}
+In the isotropic limit of trapped radiation,
+$\Phi$ is equivalent to $f^{\rm eq}$;
+in the free-streaming limit at a distance $r$ from a source,
+$\Phi$ attenuates as $r^{-2}$.
 See App.~\ref{ssec:sources_se} for details.

### 2017-08-01 paper.tex
Add Evan's edits to Secs. I & II

@@
 Leakage approximations
-\citep{deat2013-leakage, pere2016-asl, radi2016-dynamical}
+\citep{ruff1996-leakage_part1,ross2003-leakage_part2,
+  deat2013-leakage, pere2016-asl, radi2016-dynamical}
 \todo{cite others}
@@
 \todo{discuss mc more clearly}
-Monte Carlo methods \citep{abdi2012-monte_carlo,rich2015-monte_carlo}
-are an excellent tool but require large computational resources because of
-their global treatment of the problem.
+Monte Carlo methods like those
+used in supernova simulations \cite{abdi2012-monte_carlo}
+and stationary models of accretion disks \cite{rich2015-monte_carlo}
+are an excellent tool but require large computational resources
+because of their global treatment of the problem.
 
@@
 moment formalism
-\citep{shib2011-truncated_moment, fouc2015-m1_nsbh, fouc2016-m1_nsns,
-  just2015-m1_code},
-\todo{cite others}
+\citep{fouc2015-m1_nsbh, fouc2016-m1_nsns},
 evolving the zeroth- and first-angular moments of the energy density,
-that is the total energy density and flux.
+that is the total energy- and momentum-densities.
 This method, commonly called M1 transport, was formulated by
-\cite{thor1981-truncated_moment, shib2011-truncated_moment}
-and has been applied for example in \cite{ramp2002-truncated_moment,
-  card2013-truncated_moment, fouc2015-m1_nsbh, ocon2015-gr1d_with_nu}.
-Recently, \cite{fouc2016-m1_evolve_n} have expanded their M1 transport code
+\cite{thor1981-truncated_moment} and modernized by
+\cite{shib2011-truncated_moment}.
+M1 transport has recently become popular in the core-collapse supernova
+community as well
+\cite{kuro2012-m1_ccsne,just2015-m1_code,ocon2015-gr1d_with_nu,
+  skin2016-ray_vs_m1,robe2016-gr_m1_ccsne}.
+
+For merger simulations \cite{fouc2016-m1_evolve_n} have recently
+expanded their M1 transport code
 to also evolve the zeroth-angular moment of the number density,
@@
 hydrodynamics simulations.
-We choose ray tracing because it is conceptually simple, numerically cheap,
-and easily extends to high resolution in energy and angle.
+We choose ray tracing because it is conceptually simple,
+numerically inexpensive,
+and extends to high resolution in energy and angle
+by simply increasing the number of rays sampled.
+Furthermore, the computational implementation parallellizes trivially.
 
@@
 Most formulations assume an analytic spacetime metric
-\citep{birk2007-nunubar, hari2010-gr_nunubar_collapsar, kova2011-gr_ray_tracing}.
+\citep{birk2007-nunubar, caba2009-detecting_grb_nu,
+  hari2010-gr_nunubar_collapsar, kova2011-gr_ray_tracing}.
 And many make the simplifying assumption of blackbody emission from a
-neutrinosurface \citep{birk2007-nunubar, kova2011-gr_ray_tracing},
+neutrinosurface
+\citep{birk2007-nunubar, caba2009-detecting_grb_nu, kova2011-gr_ray_tracing},
 limiting them to equilibrium, optically thick configurations.
@@
 \end{align}
-representing specific energy density, specific energy flux, and
-specific momentum flux, respectively.
+defining the specific energy density,
+specific momentum density, and
+specific radiation pressure tensor, respectively.
 Here ``specific'' refers to the quantity being integrable over neutrino energy.
@@
   G(\varepsilon) &=
-  \frac{\varepsilon^2}{(2\pi)^3} \oint d\Omega' f(\varepsilon, \ell'_\beta) \\
+  \frac{\varepsilon^2}{(2\pi)^3} \oint d\Omega' f(\varepsilon, \ell'_\beta), \\
   \label{eqn:Ka}
@@
 and a fluid (also called a comoving) observer \cite{smar1980-gr_hydro}.
-Definitions are given in App.~\ref{sec:def_momentum}.
+Explicit definitions are given in App.~\ref{sec:def_momentum}.
 We distinguish moments computed assuming an Eulerian observer with a tilde,
-e.g. $\tilde{J}$, $\tilde{H}^\mu$, $\tilde{S}^{\mu\nu}$.
-Note that these three Eulerian moments are identical to the lab-frame moments
+e.g. $\tilde{J}$, $\tilde{H}^\mu$, $\tilde{S}^{\mu\nu}$;
+note that these three Eulerian moments are identical to the lab-frame moments
 $E$, $F^\mu$, and $P^{\mu\nu}$
@@
 from $t_o$ to $t_e$.
-We begin each integration by setting initial values for the variables at $t_o$.
-The observer specifies $x^i_o$ and $p_{i,o}$,
+We begin each integration by setting initial values for the variables at $t_o$:
+the observer specifies $x^i_o$ and $p_{i,o}$,
 and we set $f_{{\rm AE},o}=f_{{\rm SE},o}=0$ and $\tau_o=0$.
@@
     Note that elastic scattering opacities are identical across species
-    below energies at which weak magnetism plays a role \cite{horo2002-weak_mag}
-    (?).
+    below energies at which weak magnetism plays a role \cite{horo2002-weak_mag}.
     In this data they are exactly identical

### 2017-08-01 paper.tex
Mention dropping anisotropic term for Phi
Convert todo notes to footnotes

@@
 \usepackage{braket}               % for \bra \ket
-\usepackage{color}                % for colorful margin notes
 \usepackage[mathscr]{euscript}    % for \mathscr
@@
 \lstset{basicstyle=\ttfamily} % define \lstinline for inline code snippet
-
-\newcommand{\todo}[1]{\marginpar{\tiny{\textcolor{red}{#1}}}}
-\renewcommand\todo[1]{} % if you wish to hide all todo notes
 
@@
   deat2013-leakage, pere2016-asl, radi2016-dynamical}
-\todo{cite others}
+\footnote{TODO: cite others}
 capture the qualitative effects of neutrinos on the matter, but
 provide extremely limited information about the neutrino field itself.
-\todo{discuss mc more clearly}
+\footnote{TODO: discuss mc more clearly}
 Monte Carlo methods like those
@@
 with the history of that light-cone truncated at large optical depth.
-\todo{state scaling $\mathscr{O}(n)$?}
+\footnote{TODO: state scaling $\mathscr{O}(n)$?}
 And it easily extends to high resolution in energy and angle by simply
@@
 This is especially pronounced in the case of heavy-lepton neutrinos.
-\todo{cite}
+\footnote{TODO: cite}
 Inelastic scattering and pair processes can introduce further modifications.
@@
 for example neutrino oscillations and neutrino-antineutrino annihilation,
-\todo{cite}
+\footnote{TODO: cite}
 depend sensitively on the angular distribution.
@@
 occuring in the ejected and irradiated material.
-\todo{justify these statements}
+\footnote{TODO: justify these statements}
 The ray tracing method we present in this paper captures the dominant effects
@@
 a treatment which essentially decouples different neutrino momenta and species.
-\todo{clarify how momenta are decoupled}
+\footnote{TODO: clarify how momenta are decoupled}
 When we're solving for $f^{\nu_\sigma}(x^\alpha;p_\beta)$, we have no
@@
 and $\varepsilon=-p_t$. In that case $\cos A$ may be identified
-\todo{cite someone that uses $\mu$}
+\footnote{TODO: cite someone that uses $\mu$}
 with the familiar forward direction cosine $\mu$.
@@
 which, suppressing the arguments of $f$ for simplicity, is written,
-\todo{point out limit of QKEs}
+\footnote{TODO: point out limit of QKEs}
 \begin{equation}
@@
 See App.~\ref{ssec:sources_ae} for details.
-\todo{identify $\mathscr{K}_{\rm AE}$ and $\kappa^*$,
+\footnote{TODO: identify $\mathscr{K}_{\rm AE}$ and $\kappa^*$,
   in e.g.\ \cite{hari2010-gr_nunubar_collapsar}}
@@
 \label{ssec:moments}
-\todo{code these up}
+\footnote{TODO: code these up}
 We may take angular moments of the distribution function:
@@
 Here ``specific'' refers to the quantity being integrable over neutrino energy.
-\todo{relate moments to other frames?}
+\footnote{TODO: relate moments to other frames?}
 Integrals~\ref{eqn:J}--\ref{eqn:Sab} are performed over a solid angle in
@@
 $d\Omega \equiv d(\cos a)\,db$.
-\todo{consistent with Eqn.~\ref{eqn:def_direction}?}
+\footnote{TODO: consistent with Eqn.~\ref{eqn:def_direction}?}
 We also make use of the number density and flux defined
@@
 If the maximum of these errors exceeds some absolute or relative threshold
-\todo{specify ScaledAbsRel algorithm}
+\footnote{TODO: specify ScaledAbsRel algorithm}
 the step size is decreased and the step recomputed,
@@
 % and OdeErrorMeasures::ScaledAbsRel
-\todo{specify}
+\footnote{TODO: specify}
 
@@
 its derivatives $\psi_{\alpha\beta,\gamma}$,
-\todo{state num independent components}
+\footnote{TODO: state num independent components}
 and the hydrodynamical data $W$, $u_i$, $\varrho$, $T$, and $Y_e$.
@@
 If computed from a quasi-stationary solution to the
-\todo{clarify quasistationary timescale}
+\footnote{TODO: clarify quasistationary timescale}
 general relativistic hydrodynamics equations,
@@
 Before this work, the \emph{noscat} form of the equations has been the standard for
-\todo{explain why; cite others}
+\footnote{TODO: explain why; cite others}
 ray tracing, though many authors include the scattering optical depth in
@@
 neglecting weak magnetism
-\todo{anything else?}
+\footnote{TODO: anything else?}
 and is stored on a grid covering energy, density, temperature, and composition
@@
 and the integration proceeds backwards in time, from $t=0$
-\todo{could write more simply because Minkowski}
+\footnote{TODO: could write more simply because Minkowski}
 to some terminal $t<0$.
@@
 order Dormand-Prince algorithm \cite{pres2007-nr_3rd_ed}. This is because
-\todo{discuss with Matt}
+\footnote{TODO: discuss with Matt}
 our default 3rd order Runge-Kutta algorithm estimates a vanishing error
@@
 and $\tilde{G}(\cos A)$ defined analogously.
-\todo{add metric terms for these integrals}
+\footnote{TODO: add metric terms for these integrals}
 Results are identical for all \emph{scat} and \emph{noscat} methods.
@@
 the observer's radius.
-\todo{compute $L$, put in figure}
+\footnote{TODO: compute $L$, put in figure}
 
@@
 significantly modifies the neutrinos' spectra,
-\todo{demonstrate}
+\footnote{TODO: demonstrate}
 and the extended envelope outside the shock becomes a source of
@@
 \cite{ocon2015-gr1d_with_nu,fouc2015-m1_nsbh,abdi2012-monte_carlo}.
-\todo{also lieb2005, mull2010}
+\footnote{TODO: also lieb2005, mull2010}
 
@@
 pseudo-spectral grid composed of 11 spherical-shell subdomains
-\todo{cite early SpEC paper}
+\footnote{TODO: cite early SpEC paper}
 comprising a total of 62 radial grid points spaced approximately
@@
 We don't show it here, but the angular distribution contributing to the
-\todo{do show it}
+\footnote{TODO: do show it}
 total number density is also strongly affected by elastic scattering.
@@
 Total luminosities and average energies from the different treatments are
-\todo{update MC results 02->03}
+\footnote{TODO: update MC results 02->03}
 presented in Tab.~\ref{tab:collapse_rt_vs_mc}.
@@
 where $s$ is the length of scattering envelope passed through by the ray.
-\todo{should be $\chi_a^*=\varepsilon\zeta_a$ for $\nu_x$}
+\footnote{TODO: should be $\chi_a^*=\varepsilon\zeta_a$ for $\nu_x$}
 In the envelope the local temperature is low so $f^{\rm eq}/\Phi\ll 1$,
@@
 may avoid collapse for $X\textrm{--}Y$~s,
-\todo{cite $X$, $Y$}
+\footnote{TODO: cite $X$, $Y$}
 depending upon a number of physical factors including thermal pressure,
@@
 and 24 cells spanning azimuthal angles $\phi\in[0,2\pi]$.
-\todo{way too-low angular resolution}
+\footnote{TODO: way too-low angular resolution}
 Though a pseudospectral representation of non-smooth hydrodynamic data
@@
 $GM/r=4\,{\rm km}/250\,{\rm km}$
-\todo{too simplistic}
+\footnote{TODO: too simplistic}
 (with $G$ the gravitational constant and $M$ the central mass),
@@
 with $C \equiv 4\pi r^2\, \Delta(\cos \theta)$,
-\todo{source of GR error in coordinate $r$}
+\footnote{TODO: source of GR error in coordinate $r$}
 $\tilde{K}^r_m \equiv \tilde{K}^r(\theta_m)$, and
@@
 Tab.~\ref{tab:nsns_rt_vs_m1} compares the all-sky
-\todo{get accurate M1 numbers from Francois}
+\footnote{TODO: get accurate M1 numbers from Francois}
 luminosities and average energies from ray tracing and from
@@
 and the overestimate is slightly worse for the \emph{scat} treatment.)
-\todo{explain}
+\footnote{TODO: explain}
 With scattering turned off ray tracing overestimates $L_{\nu_x}$ by 200\%;
@@
 $R_{\bar{\nu}_e}$ dominates over $R_{\nu_e}$
-\todo{Need $R$ magnitude not just difference}
+\footnote{TODO: Need $R$ magnitude not just difference}
 because the disrupted neutron star material is releptonizing.
@@
 Eqn.~\ref{eqn:fse_homogeneous_thin_limit},
-\todo{too simplistic}
+\footnote{TODO: too simplistic}
 the scattering contribution $f_{\rm SE}$ is
@@
 is to spread the distribution out to larger angles.
-\todo{also examine spectra or average energies}
+\footnote{TODO: also examine spectra or average energies}
 
@@
 where $ds$ is an interval of proper length, and
-\todo{mention $\bar{\nu}$ eqn}
+\footnote{TODO: mention $\bar{\nu}$ eqn}
 $S$ is the flavor evolution matrix describing a mixed state
@@
 as in neutron star mergers, the problem must be solved globally.
-\todo{be clearer}
+\footnote{TODO: be clearer}
 To date, no method has been devised to handle this problem.
@@
 This is the so-called single-angle approximation.
-\todo{cite someone}
-\todo{say something about its validity}
+\footnote{TODO: cite someone}
+\footnote{TODO: say something about its validity}
 Though widely used in the supernova oscillation literature
 and shown to be qualitatively faithful in those environments,
-\todo{cite vlasenko?}
+\footnote{TODO: cite vlasenko?}
 a recent study has shown that the single-angle approximation
@@
 over all angles and an energy range of $\varepsilon\in(0,100)$~MeV.
-\todo{improve resolution and update}
+\footnote{TODO: improve resolution and update}
 We interpolate the logarithm of the self-interaction potentials
@@
 along an angle of $\theta=22^\circ$ with respect to the polar axis.
-\todo{note coordinate ray}
+\footnote{TODO: note coordinate ray}
 Fig.~\ref{fig:V_nunu-noscat} shows these terms for the \emph{noscat} treatment,
@@
 of a given flavor survives a transition.
-\todo{define explicitly}
+\footnote{TODO: define explicitly}
 Where $P_{\nu_e\rightarrow\nu_e}$ vanishes, for example in
@@
 By contrast $C^{\nu_x}_{{\rm AE}} = 0$,
-\todo{but we use a nonzero $C^{\nu_x}_{{\rm AE}}$}
+\footnote{TODO: but we use a nonzero $C^{\nu_x}_{{\rm AE}}$}
 where $\nu_x$ stands for any of the four heavy-lepton neutrinos and
@@
 (e.g. \cite[Eqn.~A5]{brue1985-core_collapse}):
-\todo{explain Fermi-blocking}
+\footnote{TODO: explain Fermi-blocking}
 \begin{align}
@@
 with $\chi^{\rm tr}_{s,i}$ the transport scattering opacity,
-\todo{drop anisotropic term and defend}
+\footnote{TODO: drop anisotropic term and defend}
 \begin{equation}
@@
 \end{equation}
+In our treatment we only retain the isotropic contribution to the
+background field:
+\begin{equation}
+  \Phi(\varepsilon)
+  = \frac{(2\pi)^3}{\varepsilon^3}\frac{1}{4\pi}
+  J(\varepsilon).
+\end{equation}
+\footnote{TODO: defend}
 
@@
 magnitude and direction of the species' momenta.
-\todo{cite \cite{keil2003-pinched_spectra} or someone}
+\footnote{TODO: cite \cite{keil2003-pinched_spectra} or someone}
 A similar formalism to the above could be used to derive source terms
 for inelastic scattering.
-\todo{compare relative amounts of inelastic to elastic scattering}
+\footnote{TODO: compare relative amounts of inelastic to elastic scattering}
 We save that for future work, pointing out here that inelastic scattering
@@
 We do not include thermal pair processes within the above formalism,
-\todo{explain this choice, and cite \cite{keil2003-pinched_spectra} or someone}
+\footnote{TODO: explain this choice, and cite \cite{keil2003-pinched_spectra} or someone}
 but as mentioned in App.~\ref{ssec:sources_ae}, we do include
@@
 by computing the emissivity of the pair processes in
-\todo{describe better}
+\footnote{TODO: describe better}
 Tab.~\ref{tab:neutrino_processes}, and applying Kirchoff's law,
@@
 Astrophysics---Center for the Evolution of the Elements (JINA-CEE)
-and the National Science Fondation under grant numbers X. \todo{grant nos?}
+and the National Science Fondation under grant numbers X. \footnote{TODO: grant nos?}
 We thank Sherwood Richers for providing his neutrino transport code

### 2017-08-01 paper.tex
Make consistent your use of \tilde for Eulerian quantities

@@
 
-A stationary observer sees a neutrino with momentum
-$p_\alpha \rightarrow \tilde{\varepsilon}(-1,\Omega_i)$,
-with direction $\Omega_i \rightarrow (\sin A, 0, \cos A)$,
-and observed energy $\tilde{\varepsilon}$.
+A stationary observer measures an energy of $\tilde{\varepsilon}$
+for a neutrino with momentum
+$p_\alpha \rightarrow \tilde{\varepsilon}(-1,\Omega_i)$
+and direction $\Omega_i \rightarrow (\sin A, 0, \cos A)$.
 In the fluid frame this neutrino has energy $\varepsilon=-u^\alpha p_\alpha$;
@@
 
-Since no analytic form is known for the background field $\Phi$,
-we examine only the \emph{noscat} case,
+Since no analytic form is known for the background field $\Phi$
+interior to the star, we examine only the \emph{noscat} case,
 with $\chi_s=0$ and $\chi=\chi_a^{*}$.
@@
 30 points in angle $\cos A \in [0.734,1]$ and
-40 points in energy $\varepsilon \in [0,100]$~MeV.
+40 points in energy $\tilde{\varepsilon} \in [0,100]$~MeV.
 Because of the step function in fluid variables at radius $R$,
@@
 Though we do not display it here,
-we also examine $\langle \varepsilon \rangle(\cos A)$
+we also examine $\langle \tilde{\varepsilon} \rangle(\cos A)$
 as in Eqn.~\ref{eqn:avg_eps_per_cosA}, and robustly reproduce the
 common feature of increasing average energies at the limb of the star
-for $\nu_e$ and $\bar{\nu}_e$. In contrast $\langle \varepsilon \rangle$
+for $\nu_e$ and $\bar{\nu}_e$. In contrast $\langle \tilde{\varepsilon} \rangle$
 for $\nu_x$ is constant across angles, as expected when the star
-is optically thin at even the highest energies $\varepsilon\sim100$~MeV.
+is optically thin at even the highest energies $\tilde{\varepsilon}\sim100$~MeV.
 
@@
 \label{ssec:test_gravity}
+Unfinished.
 Optically thick star in Schwarszchild.
@@
 \footnote{TODO: compute $L$, put in figure}
+Also compare a few geodesics in Kerr to analytic.
 
@@
 presented in Tab.~\ref{tab:collapse_rt_vs_mc}.
+\footnote{TODO: give equation for L}
 As expected, $\nu_e$ is least affected by scattering, and $\nu_x$ most.
@@
     We also show average energies of the scattering envelope
-    $\langle\varepsilon\rangle^{\rm sc}$, which are estimated by eye
+    $\langle\tilde{\varepsilon}\rangle^{\rm sc}$, which are estimated by eye
     from plots like Fig.~\ref{fig:avg_eps_collapse} at $\cos A < 0.95$.
@@
     $L_{\nu_x}$                                         & 222.0  & 1.88 & 1.76 & 1.70 \\
-    $\langle \varepsilon_{\nu_e} \rangle$               &  10.6 & 10.6 & 10.9 & 11.0 \\
-    $\langle \varepsilon_{\bar{\nu}_e} \rangle$         & 14.2 & 13.0 & 13.8 & 13.7 \\
-    $\langle \varepsilon_{\nu_x} \rangle$               & 47.8 & 16.0 & 17.3 & 16.2 \\
-    $\langle \varepsilon_{\nu_e} \rangle^{\rm sc}$      & 4    & 14   & 17   & 16 \\
-    $\langle \varepsilon_{\bar{\nu}_e} \rangle^{\rm sc}$& 4    & 24   & 20   & 18 \\
-    $\langle \varepsilon_{\nu_x} \rangle^{\rm sc}$      & 2    & 28   & 31   & 27 \\
+    $\langle \tilde{\varepsilon}_{\nu_e} \rangle$               &  10.6 & 10.6 & 10.9 & 11.0 \\
+    $\langle \tilde{\varepsilon}_{\bar{\nu}_e} \rangle$         & 14.2 & 13.0 & 13.8 & 13.7 \\
+    $\langle \tilde{\varepsilon}_{\nu_x} \rangle$               & 47.8 & 16.0 & 17.3 & 16.2 \\
+    $\langle \tilde{\varepsilon}_{\nu_e} \rangle^{\rm sc}$      & 4    & 14   & 17   & 16 \\
+    $\langle \tilde{\varepsilon}_{\bar{\nu}_e} \rangle^{\rm sc}$& 4    & 24   & 20   & 18 \\
+    $\langle \tilde{\varepsilon}_{\nu_x} \rangle^{\rm sc}$      & 2    & 28   & 31   & 27 \\
     \hline
@@
 so $f_{\rm SE}$ strongly dominates over $f_{\rm AE}$.
-The average energy of the scattered field is therefore
+The average energy of the scattered field measured by our observer
+is therefore
 \begin{equation}
   \label{eqn:avg_eps_sc}
-  \langle \varepsilon \rangle^{\rm sc}
+  \langle \tilde{\varepsilon} \rangle^{\rm sc}
   = \frac{\int d\varepsilon\,\varepsilon^3 f}{\int d\varepsilon\,\varepsilon^2 f}
@@
 \end{equation}
+Note that we have taken the liberty here of identifying the fluid-frame energy
+$\varepsilon$ with the Eulerian energy $\tilde{\varepsilon}$,
+since infall velocities in the envelope are $\sim0.1c$,
+and as Fig.~\ref{fig:avg_eps_doppler} indicates,
+the Doppler shift will therefore introduce an error into our analysis of
+$\sim10\%$.
 
@@
 
-In order to estimate $\langle \varepsilon \rangle^{\rm sc}$ analytically,
+In order to estimate $\langle \tilde{\varepsilon} \rangle^{\rm sc}$ analytically,
 we must write the background field $\Phi(\varepsilon)$ analytically.
@@
   \label{eqn:pinched_avg_eps_sc_1}
-  \langle\varepsilon\rangle^{\rm sc}
+  \langle\tilde{\varepsilon}\rangle^{\rm sc}
   &\approx \frac{\alpha+3}{\alpha+1} \langle\varepsilon\rangle, \\
@@
 
-The $\langle\varepsilon_{\bar{\nu}_e}\rangle^{\rm sc}$ prediction
+The $\langle\tilde{\varepsilon}_{\bar{\nu}_e}\rangle^{\rm sc}$ prediction
 in the \emph{gray} \emph{scat} treatment is approximately 30\%

### 2017-09-04 paper.tex
Add references to Deaton thesis and Foucart MC-M1 paper

@@
 
-For merger simulations \cite{fouc2016-m1_evolve_n} have recently
+For merger simulations, \cite{fouc2016-m1_evolve_n} have recently
 expanded their M1 transport code
@@
 %\subsubsection*{Why ray tracing?}
-In this work we present a ray tracing method to compute neutrino distribution
-functions from state-of-the-art general relativistic M1 transport
-hydrodynamics simulations.
+Angular and spectral neutrino distributions in neutron star merger simulations may
+be available soon, for example with the coupled Monte-Carlo-M1 scheme presented in
+\citep{fouc2017-m1_with_mc}.
+In this work, however, we present a ray tracing method to compute neutrino
+distribution functions from the existing state-of-the-art general relativistic M1
+transport hydrodynamics simulations.
 We choose ray tracing because it is conceptually simple,
@@
 \label{ssec:test_gravity}
-Unfinished.
-Optically thick star in Schwarszchild.
-Total luminosity at infinite separation may be computed by
-computing the radial energy flux at a finite separation from the star, $r$,
-and using $p_t$ for the energy. We compute this asymptotic luminosity
-at several radii, $r$, and show that the luminosity is independent of
-the observer's radius.
-\footnote{TODO: compute $L$, put in figure}
+Unfinished: describe and point readers to Asano-Fukuyama test presented
+in thesis ~\cite{deat2015-thesis}.
 Also compare a few geodesics in Kerr to analytic.
+\footnote{TODO: compute some geodesics in Kerr}
 

### 2017-09-06 paper.tex
Add section reference to Deaton thesis citation
Change wording facetious->fictitious

@@
 $\tau_{\rm term}$ at the earliest effective emission event $x_e^\alpha$.
-The concept of earliest emission event is a facetious construct we use
+The concept of earliest emission event is a fictitious construct we use
 to allow us to truncate our integration at an event along the ray where any
@@
 Unfinished: describe and point readers to Asano-Fukuyama test presented
-in thesis ~\cite{deat2015-thesis}.
+in thesis ~\cite[Sec.~4.3.3]{deat2015-thesis}.
 Also compare a few geodesics in Kerr to analytic.

### 2017-09-15 paper.tex
Add Matt's comments through p.6

@@
 Neutrinos are one of the dominant energy transport phenomena at play in
-neutron star mergers, heating, cooling, and pushing the disrupted nuclear
+neutron star mergers: heating, cooling, and pushing the disrupted nuclear
 matter.
@@
 respects the freedom of neutrino distribution functions to vary
-drastically from simple geometric distributions in thermodynamic equilibrium.
+drastically from geometrically simple distributions in thermodynamic equilibrium.
 
@@
 \footnote{TODO: cite others}
-capture the qualitative effects of neutrinos on the matter, but
+capture some of the qualitative effects of neutrinos on the matter, but
 provide extremely limited information about the neutrino field itself.
@@
 
-The state of the art today couples radiation and matter using a truncated
-moment formalism
+The state of the art today for neutron star merger calculations
+couples radiation and matter using a truncated moment formalism
 \citep{fouc2015-m1_nsbh, fouc2016-m1_nsns},
@@
 \citep{ruff1999-nunubar_nsns, asan2000-nunubar, birk2007-nunubar,
-  hari2010-gr_nunubar_collapsar, zala2011-nunubar, leng2014-nunubar}.
+  hari2010-gr_nunubar_collapsar, zala2011-nunubar, leng2014-nunubar,
+  just2016-nsns_ejecta_obstacles}.
 
@@
 \footnote{TODO: clarify how momenta are decoupled}
-When we're solving for $f^{\nu_\sigma}(x^\alpha;p_\beta)$, we have no
+When solving for $f^{\nu_\sigma}(x^\alpha;p_\beta)$, we have no
 information about $f^{\nu_\sigma}(x^\alpha;p'_\beta)$, and
@@
 independent of that observer's position in coordinate space.
-For explicit definitions see App.~\ref{sec:def_momentum}.
+For explicit definitions see App.~\ref{sec:definitions}.
 
@@
 where $d/d\lambda$ denotes a derivative with respect to the affine
-parameter defining the neutrino momentum (Eqn.~\ref{eqn:geodesic_x})
+parameter defining the neutrino momentum (Eqn.~\ref{eqn:geodesic_x} below)
 and $C[f]$ is the source term arising from interactions with the medium.
@@
 over a lengthscale $\mathscr{K}^{-1}$ in our affine parameter;
-or from Eqn.~\ref{eqn:proper_length}, over a proper lengthscale
+or from Eqn.~\ref{eqn:proper_length} below, over a proper lengthscale
 $\varepsilon/\mathscr{K}$ measured by our fiducial observer.
@@
 \end{align}
-with ${}^ZA$ representing a nucleus of mass number $A$ and charge $Z$.
+with ${}^AZ$ representing a nucleus of mass number $A$ and charge $Z$.
 In terms of the emissivity $j(\varepsilon)$ describing number of neutrinos
@@
 We may use Eqns.~\ref{eqn:ae_emissivity_feq} and \ref{eqn:se_emissivity}
-to separate the source function thus
+to separate the source function thus:
 $\mathscr{S}=\mathscr{S}_{\rm AE}+\mathscr{S}_{\rm SE}$, with
@@
 
-There are two sensible observers to use to define the frame in which these
-moments are computed, that is defining the four-velocity of
-Eqn.~\ref{eqn:def_momentum}:
-a stationary (also called an Eulerian) observer,
-and a fluid (also called a comoving) observer \cite{smar1980-gr_hydro}.
-Explicit definitions are given in App.~\ref{sec:def_momentum}.
-We distinguish moments computed assuming an Eulerian observer with a tilde,
+We compute moments for a particular observer
+by specifying the four-velocity in Eqn.~\ref{eqn:def_momentum}.
+Two choices are particularly useful:
+an observer stationary in the coordinate frame (i.e. Eulerian),
+or one stationary in the fluid frame (i.e. comoving)
+\cite{smar1980-gr_hydro}.
+Explicit definitions are given in App.~\ref{sec:definitions}.
+We distinguish moments computed for an Eulerian observer with a tilde,
 e.g. $\tilde{J}$, $\tilde{H}^\mu$, $\tilde{S}^{\mu\nu}$;
@@
 
-The volume data are components of the metric $\psi_{\alpha\beta}$,
+We integrate these equations through the simulated spacetime
+over which the following volume data are known
+(all defined in App.~\ref{sec:definitions}):
+the spacetime metric $\psi_{\alpha\beta}$,
 its derivatives $\psi_{\alpha\beta,\gamma}$,
-\footnote{TODO: state num independent components}
-and the hydrodynamical data $W$, $u_i$, $\varrho$, $T$, and $Y_e$.
+the fluid velocity $u_i$,
+Lorentz factor $W$,
+density $\varrho$,
+temperature $T$,
+and electron fraction $Y_e$.
 These fields are stored in spectral representation.
@@
 with the matter.
-We set up a large slab of matter
+We set up a large slab of matter in a Minkowski spacetime
 representative of the fluid in the Massive Collapsed Star test
@@
 
-\section{Momentum Definitions}
-\label{sec:def_momentum}
-Here we describe our conventions defining the four-momentum of a neutrino.
+\section{Definitions}
+\label{sec:definitions}
 We decompose the neutrino momentum into components parallel and orthogonal to
@@
 $\tilde{q}=1$, $\tilde{s}=0$.
+
+In addition to the fluid velocity and Lorentz factor, described above, the
+other fluid state variables we use from our hydrodynamic simulations
+are rest density $\varrho=m_b n_b$, temperature $T$, and electron fraction
+\begin{equation}
+  Y_e = \frac{n_{e^-} - n_{e^+}}{n_b},
+\end{equation}
+with $n_{e^-}$, $n_{e^+}$, and $n_b$, the number densities of electrons
+positrons, and baryons, and $m_b$ is the average baryon mass.
 

### 2017-10-12 paper.tex
Add Matt's comments through p.7

@@
     $T=3.7$~MeV, $Y_e=0.12$).
-    In the fluid rest frame these are given by
-    $(\chi^*_a)^{-1}$ for the absorption
-    and $(\chi_s)^{-1}$ for elastic scattering interactions.
+    In the fluid rest frame these are given by the inverses of the opacities,
+    $\chi^*_a=\mathscr{K}_{\rm AE}/\varepsilon$ for absorption and
+    $\chi_s=\mathscr{K}_{\rm SE}/\varepsilon$ for elastic scattering interactions;
+    see App.~\ref{sec:source_terms} for definitions.
     Note that elastic scattering opacities are identical across species
@@
 \label{ssec:test_doppler}
-We reproduce the test above, but with the matter and observer in relative
-motion.
-We use a stationary (i.e.\ Eulerian) observer
+We reproduce the test above (again in Minkowski spacetime),
+but with the matter and observer in relative motion.
+We use a stationary observer
 and a fluid velocity $u^\alpha \rightarrow W(1,0,0,v)$ with $W=(1-v^2)^{-1/2}$.
@@
 In the fluid frame this neutrino has energy $\varepsilon=-u^\alpha p_\alpha$;
-therefore the average energy varies with observing angle:
+therefore the average energy varies with observing angle like
 \begin{equation}
@@
   \langle \tilde{\varepsilon} \rangle(\cos A) =
-  \langle \varepsilon \rangle^{\rm eq} \frac{1}{W(1-v \cos A)}.
-\end{equation}
+  \langle \varepsilon \rangle^{\rm eq} \frac{1}{W(1-v \cos A)},
+\end{equation}
+where our symbol $\langle \tilde{\varepsilon} \rangle(\cos A)$ expresses
+a functional dependence on $\cos A$.
 The equilibrium average energy $\langle \varepsilon \rangle^{\rm eq}$
@@
 \end{align}
-where $s$ is the path length traversed by the ray through the star.
+where $s$ is the path length traversed by the ray through the star,
+the total opacity is $\chi=\chi_a^*+\chi_s$, and the stimulated absorption
+opacity $\chi_a^*$ and elastic scattering opacity $\chi_s$
+are defined in App.~\ref{sec:source_terms}.
 
@@
 $Y_e\in[0.035,0.55]$ linearly,
-with grid extents$\{18,82,65,51\}$ respectively.
+with grid extents $\{18,82,65,51\}$ respectively.
 

### 2017-10-13 paper.tex
Add Matt's comments through p.8 (last edit incomplete)
Note, the last edit is an incomplete attempt to improve the paragraph
at the end of III.C, explaining the phenomenon of increasing <eps> at
the limb of the star. I may delete this paragraph in the end...

@@
 %\subsubsection*{Why does a covariant formulation matter?}
-We formulate our ray tracing equations covariantly---free from
+We formulate the ray tracing equations covariantly---free from
 assumptions about the spacetime geometry or coordinates. This is essential
@@
 behavior of the right hand side, driving $f$ toward $\mathscr{S}$
-over a lengthscale $\mathscr{K}^{-1}$ in our affine parameter;
+over a lengthscale $\mathscr{K}^{-1}$ in the affine parameter;
 or from Eqn.~\ref{eqn:proper_length} below, over a proper lengthscale
-$\varepsilon/\mathscr{K}$ measured by our fiducial observer.
+$\varepsilon/\mathscr{K}$ measured by the fiducial observer.
 
@@
 we find the element of proper distance traversed by the
-neutrino as measured by our fiducial obsever $u^\alpha$ is
+neutrino as measured by the fiducial observer $u^\alpha$ is
 \begin{equation}
@@
 We can integrate Eqn.~\ref{eqn:boltzmann_split} directly, 
-with our solution split into a boundary, absorption/emission, and
+with the solution split into a boundary, absorption/emission, and
 a scattering term, $f=f_{\rm bdry}+f_{\rm AE}+f_{\rm SE}$:
@@
 By using the time-component of Eqn.~\ref{eqn:geodesic_x} ($dt=d\lambda \, p^t$)
-we may transform our integrations to coordinate time.
+we may transform the integrations to coordinate time.
 The coupled system of ordinary differential equations is
@@
 The concept of earliest emission event is a fictitious construct we use
-to allow us to truncate our integration at an event along the ray where any
+to allow us to truncate the integration at an event along the ray where any
 further additions to the field are negligible due to the large optical depth
@@
 at this thermodynamic point over two energy decades.
-For this test we set up our slab large enough for neutrinos of all
+For this test we choose a slab large enough for neutrinos of all
 energies to be trapped.
-Our domain extends to $\tau_{\rm term}\,L_{\rm mfp,max}\approx10^7\,{\rm km}$,
+We sample the distribution function with ray tracing over a uniform grid
+of 40 points in energy $\varepsilon \in [0,100]$~MeV.
+The domain extends to $\tau_{\rm term}\,L_{\rm mfp,max}\approx10^7\,{\rm km}$,
 since our ray tracing algorithm integrates rays to terminal optical
@@
 function integrated along the ray, for each of the three species and using
-both methods, \emph{noscat} and \emph{scat}. We display
+both methods, \emph{noscat} and \emph{scat}, at the single energy
+$\varepsilon=11.25$~MeV. We display
 Eqns.~\ref{eqn:num_fae} and \ref{eqn:num_fse} in their integral form:
@@
 over the energies $\varepsilon\in[0,100]$~MeV
-probes our numerical solution over length scales from
+probes the numerical solution over length scales from
 0.1~km to $10^{5}\,{\rm km}$.
@@
 \end{equation}
-where our symbol $\langle \tilde{\varepsilon} \rangle(\cos A)$ expresses
+where the symbol $\langle \tilde{\varepsilon} \rangle(\cos A)$ expresses
 a functional dependence on $\cos A$.
@@
 constant inside radius $R$ and vanishing outside.
-We choose $R=50$~km and place our observer at $r=75$~km,
+We choose $R=50$~km and place the observer at $r=75$~km,
 at which position there is a radiation cone of half-opening angle
@@
   f_{\rm AE} &=
-  \frac{\chi_a^{*}}{\chi} f^{\rm eq} \left(1-e^{-\chi s}\right) \\
+  \frac{\chi_a^{*}}{\chi} f^{\rm eq} \left(1-e^{-\chi s}\right), \\
   \label{eqn:fse_homogeneous_general}
   f_{\rm SE} &=
-  \frac{\chi_s}{\chi} \Phi \left(1-e^{-\chi s}\right) \\
+  \frac{\chi_s}{\chi} \Phi \left(1-e^{-\chi s}\right),
+\end{align}
+where $s$ is the path length traversed by the ray through the star,
+\begin{equation}
   \label{eqn:geometric_s}
-  s &= 2 R \left(1-\frac{r^2}{R^2}(1-\cos^2 A)\right)^{1/2},
-\end{align}
-where $s$ is the path length traversed by the ray through the star,
-the total opacity is $\chi=\chi_a^*+\chi_s$, and the stimulated absorption
+  s = 2 R \left(1-\frac{r^2}{R^2}(1-\cos^2 A)\right)^{1/2}.
+\end{equation}
+The total opacity is $\chi=\chi_a^*+\chi_s$, and the stimulated absorption
 opacity $\chi_a^*$ and elastic scattering opacity $\chi_s$
@@
 
-Expanding the exponential in Eqn.~\ref{eqn:fae_homogeneous_general},
-we see the distribution function takes the limiting values:
+We can explain these features quantitatively by examining the limits
+of Eqn.~\ref{eqn:fae_homogeneous_general}, expanding the exponential function
+in powers of $\chi s$; the distribution function takes the limiting values
 \begin{equation}
@@
   \begin{cases}
-    \chi s(\cos A) & s \ll \chi^{-1}\,\textrm{optically thin}\\
-    1                  & s \gg \chi^{-1}\,\textrm{optically thick}.
+    \chi s(\cos A) & s \ll \chi^{-1}\,\textrm{(thin)}\\
+    1                  & s \gg \chi^{-1}\,\textrm{(thick)}.
   \end{cases}
 \end{equation}
-This agrees with Fig.~\ref{fig:f_absorption_sphere}, with $\nu_x$
-in the optically thin limit and $\nu_e$ in the optically thick limit.
+These limits are represented in Fig.~\ref{fig:f_absorption_sphere}:
+with $\nu_x$ in the optically thin limit at all viewing angles,
+and $\nu_e$ in the optically thick limit at viewing angles
+$\cos A \gtrsim 0.8$.
 
@@
 Though we do not display it here,
-we also examine $\langle \tilde{\varepsilon} \rangle(\cos A)$
-as in Eqn.~\ref{eqn:avg_eps_per_cosA}, and robustly reproduce the
-common feature of increasing average energies at the limb of the star
-for $\nu_e$ and $\bar{\nu}_e$. In contrast $\langle \tilde{\varepsilon} \rangle$
+we also examine $\langle \tilde{\varepsilon}\rangle$ as a function of
+viewing angle (via Eqn.~\ref{eqn:avg_eps_per_cosA}).
+In the interior of the star
+$\langle \tilde{\varepsilon_{\nu_x}}\rangle$ is constant across viewing angle,
+\footnote{TODO: need to finish cleaning up this description}
+and find the counterintuitive
+feature of increasing average energies at the limb of the star
+for $\nu_e$ and $\bar{\nu}_e$.
+In contrast $\langle \tilde{\varepsilon} \rangle$
 for $\nu_x$ is constant across angles, as expected when the star
 is optically thin at even the highest energies $\tilde{\varepsilon}\sim100$~MeV.
+
+For viewing angles $\cos A \sim 0.75$, at the limb of the star,
+rays pass through significantly less material, i.e.\ $s \ll R$.
+Because of the opacity's approximate $\varepsilon^2$ dependence,
+the highest-energy rays saturate to $f^{\rm eq}$, increasing the average spectral energy.
 

### 2017-10-18 paper.tex
Clean up some text already edited

@@
 We set up a large slab of matter in a Minkowski spacetime
-representative of the fluid in the Massive Collapsed Star test
-(Sec.~\ref{ssec:test_collapse}) at the radius 50~km.
+representative of the fluid thermodynamic state at a radius of 50~km
+in the test presented in Sec.~\ref{ssec:test_collapse}
+(a massive collapsed star following core bounce).
 For comparision with that test we use the LS180 equation of state
@@
 The final distribution functions asymptote to their equilibrium Fermi-Dirac
-levels $f^{\rm eq}_{\{\nu_e,\bar{\nu}_e,\nu_x\}}=\{0.039, 0.053, 0.046\}$
+levels at this energy
+$f^{\rm eq}_{\{\nu_e,\bar{\nu}_e,\nu_x\}}=\{0.039, 0.053, 0.046\}$
 within a few mean free paths;
@@
 
-As Fig.~\ref{fig:mfps-50km} shows, integrating the distribution functions
+As Fig.~\ref{fig:mfps-50km} reveals, integrating the distribution functions
 over the energies $\varepsilon\in[0,100]$~MeV
@@
 We use a stationary observer
-and a fluid velocity $u^\alpha \rightarrow W(1,0,0,v)$ with $W=(1-v^2)^{-1/2}$.
+and fluid moving in the positive $z$-direction:
+with $u^\alpha \rightarrow W(1,0,0,v)$ and $W=(1-v^2)^{-1/2}$.
 All other thermodynamic variables and background fields are unchanged
@@
 \end{equation}
-where the symbol $\langle \tilde{\varepsilon} \rangle(\cos A)$ expresses
-a functional dependence on $\cos A$.
+where the symbol $\langle \tilde{\varepsilon} \rangle(\cos A)$ emphasizes
+the functional dependence on $\cos A$.
 The equilibrium average energy $\langle \varepsilon \rangle^{\rm eq}$
@@
     $\langle\tilde{\varepsilon}\rangle(\cos A)$.
-    The observer sees neutrinos moving with the fluid for angles $\cos A<0$.
-    Test described in Sec.~\ref{ssec:test_doppler}.
+    The observer sees neutrinos with $\cos A>0$ to be moving primarily with
+    the fluid.
     The points are computed from ray tracing spectra; the lines are the

### 2017-10-24 paper.tex
Convert doppler figure to epslatex for latex labels
Add asymptotic eqns to appendix gray method

@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{fig-moving_fluid_avg_eps}
+  %\includegraphics[width=\columnwidth]{fig-moving_fluid_avg_eps}
+  \resizebox{\columnwidth}{!}{\input{fig-moving_fluid_avg_eps.tex}}
   % Data:   $DIR/Lev1-v*-noscat-01
@@
   \end{equation}
+  We implement the Fermi integrals using the analytical approximants from
+  \cite{taka1978-beta_rates}.
 \item
@@
 \end{enumerate}
-When we use these moments in Eqn.~\ref{eqn:background_phi},
+
+Note that to avoid floating point errors for very large negative $\eta$,
+we employ the asymptotic form of Eqn.~\ref{eqn:fermi_integral}:
+$\lim_{\eta\ll-1}\mathscr{F}_b(\eta) = b!\,e^\eta$.
+Thus, for $\eta<-10$, we use the limiting forms of
+Eqns.~\ref{eqn:spectral_temp},
+\ref{eqn:J_from_gray},
+and \ref{eqn:H_from_gray}:
+\begin{align}
+    T_\nu &= 
+    \frac{\langle \varepsilon \rangle}{3}, \\
+    J(\varepsilon) &=
+    J \frac{\varepsilon^3}{6 T_\nu^4} e^{-\varepsilon/T_\nu}, \\
+    H^\mu(\varepsilon) &=
+    H^\mu \frac{\varepsilon^3}{6 T_\nu^4} e^{-\varepsilon/T_\nu}.
+\end{align}
+
+When we use the moments from Eqns.~\ref{eqn:J_from_gray} and
+\ref{eqn:H_from_gray} in Eqn.~\ref{eqn:background_phi},
 we call this the \emph{gray} method.

### 2017-10-25 paper.tex
Clean up some more text already edited

@@
 We integrate these equations through the simulated spacetime
-over which the following volume data are known
-(all defined in App.~\ref{sec:definitions}):
+over which the following volume data are known:
 the spacetime metric $\psi_{\alpha\beta}$,
@@
 temperature $T$,
-and electron fraction $Y_e$.
+and electron fraction $Y_e$,
+all defined in App.~\ref{sec:definitions}.
 These fields are stored in spectral representation.
-If they are computed via hydrodynamical simulation, they are saved to disk
+If they are computed via hydrodynamical simulation, they may be saved to disk
 at specified coordinate times and interpolated
@@
 If computed from a quasi-stationary solution to the
-\footnote{TODO: clarify quasistationary timescale}
 general relativistic hydrodynamics equations,
 a single time is stored to disk, and no time interpolation is needed.
+In this paper for simplicity and to limit computational memory loads,
+we limit ourselves to quasistationary configurations
+and use single time slices in every test case.
 
@@
 \label{ssec:test_doppler}
-We reproduce the test above (again in Minkowski spacetime),
+We reproduce the test above, again in Minkowski spacetime,
 but with the matter and observer in relative motion.
@@
 and fluid moving in the positive $z$-direction:
-with $u^\alpha \rightarrow W(1,0,0,v)$ and $W=(1-v^2)^{-1/2}$.
+with $u^\alpha \rightarrow W(1,0,0,v)$ and $W=(1-v^2)^{-1/2}$,
+where $W$ is the Lorentz factor defined in App.~\ref{sec:definitions}.
 All other thermodynamic variables and background fields are unchanged
@@
 \end{equation}
-with $n_{e^-}$, $n_{e^+}$, and $n_b$, the number densities of electrons
+where $n_{e^-}$, $n_{e^+}$, and $n_b$ are the number densities of electrons
 positrons, and baryons, and $m_b$ is the average baryon mass.

### 2017-10-31 paper.tex
Add a few clarifying edits

@@
   We present a covariant ray tracing algorithm for computing high-resolution
-  neutrino distributions in general numerical spacetimes with hydrodynamical
-  sources.
+  neutrino distributions in general relativistic numerical spacetimes with
+  hydrodynamical sources.
   Our formulation treats the very important effect of
@@
     the neutrino emission event is at $x^\alpha_e$. The affine parameter
-    increases from the emission event: $\lambda_e<\lambda'<\lambda''$.}
+    increases from the emission event: $\lambda_e<\lambda'<\lambda''$.
+    The green slab represents dense matter.}
   \label{fig:affine_param}
@@
   \caption{Mean free paths
-    representative of the collapse profile at
-    50~km ($\varrho=10^{11}\,{\rm g}\,{\rm cm}^{-3}$,
-    $T=3.7$~MeV, $Y_e=0.12$).
-    In the fluid rest frame these are given by the inverses of the opacities,
+    representative of the post-bounce collapse profile (presented in
+    Sec.~\ref{ssec:test_collapse}) at
+    50~km, where $\varrho=10^{11}\,{\rm g}\,{\rm cm}^{-3}$,
+    $T=3.7$~MeV, and $Y_e=0.12$; this thermodynamic state is used in
+    the tests with homogeneous matter distributions
+    (presented in Secs.~\ref{ssec:test_equilibrium}--\ref{ssec:test_ab_star}).
+    In the fluid rest frame the mean free paths are given by the inverses of the opacities
     $\chi^*_a=\mathscr{K}_{\rm AE}/\varepsilon$ for absorption and
-    $\chi_s=\mathscr{K}_{\rm SE}/\varepsilon$ for elastic scattering interactions;
-    see App.~\ref{sec:source_terms} for definitions.
+    $\chi_s=\mathscr{K}_{\rm SE}/\varepsilon$ for elastic scattering interactions
+    (see App.~\ref{sec:source_terms} for definitions).
     Note that elastic scattering opacities are identical across species
@@
   \caption{Cumulative distribution functions at $\varepsilon=11.25$~MeV
-    in homogeneous infinite slab of Sec.~\ref{ssec:test_equilibrium}.
+    in the homogeneous infinite slab test
+    (presented in Sec.~\ref{ssec:test_equilibrium}).
     The integration proceeds from left to right, or backwards in time $t$.
@@
   \caption{Relative error in integrated equilibrium distribution functions
-    in the infinite homogeneous slab  described in
-    Sec.~\ref{ssec:test_equilibrium}.
-    Presented here are the relative differences between the final
+    in the infinite homogeneous slab
+    (presented in Sec.~\ref{ssec:test_equilibrium}).
+    Plotted here are the relative differences between the final
     $f_{\rm AE}$ in the \emph{noscat} case
@@
     measured by a stationary observer,
-    $\langle\tilde{\varepsilon}\rangle(\cos A)$.
+    $\langle\tilde{\varepsilon}\rangle(\cos A)$,
+    in the moving slab test (presented in Sec.~\ref{ssec:test_doppler}).
     The observer sees neutrinos with $\cos A>0$ to be moving primarily with
@@
   \caption{Distribution functions outside an idealized homogeneous star
-    with radius $R=50$~km, and observer at $r=75$~km,
-    described in Sec.~\ref{ssec:test_ab_star}.
+    with radius $R=50$~km, and observer at $r=75$~km
+    (presented in Sec.~\ref{ssec:test_ab_star}).
     $A$ is the angle between the neutrino momentum and the $\hat{r}$ direction.
@@
     $\langle\varepsilon\rangle(\cos A)$
-    in the collapse profile test described in Sec.~\ref{ssec:test_collapse}.
+    in the collapse profile test (described in Sec.~\ref{ssec:test_collapse}).
     The observer is at 500~km, and the shock at 150~km,
@@
   \caption{
-    Comparison of total luminosities and average energies between
+    From the post-bounce collapse profile test (Sec.~\ref{ssec:test_collapse}),
+    a comparison of total luminosities and average energies between
     the following methods:
@@
   \caption{A meridional slice of density in the hypermassive neutron
-    star and disk configuration.
+    star and disk configuration (Sec.~\ref{ssec:test_disk_comparison}).
     The distorted rectangular boundaries are the boundaries of the grid
@@
   \caption{An equatorial slice of density in the hypermassive neutron
-    star and disk configuration.}
+    star and disk configuration (Sec.~\ref{ssec:test_disk_comparison}).}
   \label{fig:nsns_rho_equat}
@@
   \caption{A meridional slice of temperature in the hypermassive neutron
-    star and disk configuration.}
+    star and disk configuration (Sec.~\ref{ssec:test_disk_comparison}).}
   \label{fig:nsns_temp_merid}
@@
   \caption{An equatorial slice of temperature in the hypermassive neutron
-    star and disk configuration.}
+    star and disk configuration (Sec.~\ref{ssec:test_disk_comparison}).}
   \label{fig:nsns_temp_equat}
@@
     Comparison of total luminosities and average energies
-    of the hypermassive neutron star configuration.
+    of the hypermassive neutron star configuration
+    (presented in Sec.~\ref{ssec:test_disk_comparison}).
     The methods are ray tracing using the \emph{noscat} method `RT noscat',
@@
   % Data gathered using: $DIR/CollectPositionDistributions.sh
-  % Script: $DIR/GnuScripts/plot-theta_distributions-Hr-fixed_r.gnu
+  % Script: $DIR/GnuScripts/plot-theta_distributions-Kr-fixed_r.gnu
   % w/ DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20170324-ray_tracing_nsns/Run03
@@
     as defined in Eqn.~\ref{eqn:Kr_minkowski},
-    in the hypermassive neutron star configuration.
+    in the hypermassive neutron star configuration
+    (presented in Sec.~\ref{ssec:test_disk_comparison}).
     Sampled for observers at fixed coordinate distances from the center
@@
   \caption{Angular distributions of fluxes for an observer on the rotation axis
-    in the hypermassive neutron star configuration.}
+    in the hypermassive neutron star configuration
+    (presented in Sec.~\ref{ssec:test_disk_comparison}).}
   \label{fig:nsns_cosA_distrib_I_00deg}
@@
   \caption{Angular distributions of fluxes for an observer in the equatorial
-    plane in the hypermassive neutron star configuration.}
+    plane in the hypermassive neutron star configuration
+    (presented in Sec.~\ref{ssec:test_disk_comparison}).}
   \label{fig:nsns_cosA_distrib_I_90deg}

### 2017-11-10 paper.tex
Remove last paragraphs of III.C
This wipes the incomplete paragraphs introduced in 5aa6a4d.

The deleted paragraphs had attempted to explain an unphysical
phenomenon present in the analytical solutions (unphysical because
no scattering is allowed), and reproduced robustly in the numerical
solutions. The phenomenon isn't visible in any of the plots
presented with this test, however; and furthermore, an analysis of
the phenomenon doesn't lead to anything else in the paper. So I
deleted it, rather than labor any longer over improving the
paragraphs.

@@
 \end{figure}
-
-Though we do not display it here,
-we also examine $\langle \tilde{\varepsilon}\rangle$ as a function of
-viewing angle (via Eqn.~\ref{eqn:avg_eps_per_cosA}).
-In the interior of the star
-$\langle \tilde{\varepsilon_{\nu_x}}\rangle$ is constant across viewing angle,
-\footnote{TODO: need to finish cleaning up this description}
-and find the counterintuitive
-feature of increasing average energies at the limb of the star
-for $\nu_e$ and $\bar{\nu}_e$.
-In contrast $\langle \tilde{\varepsilon} \rangle$
-for $\nu_x$ is constant across angles, as expected when the star
-is optically thin at even the highest energies $\tilde{\varepsilon}\sim100$~MeV.
-
-For viewing angles $\cos A \sim 0.75$, at the limb of the star,
-rays pass through significantly less material, i.e.\ $s \ll R$.
-Because of the opacity's approximate $\varepsilon^2$ dependence,
-the highest-energy rays saturate to $f^{\rm eq}$, increasing the average spectral energy.
 

### 2017-11-10 paper.tex
Add Matt's comments through p. 13

@@
 
-We don't show it here, but the angular distribution contributing to the
-\footnote{TODO: do show it}
+Although not shown here, the angular distribution contributing to the
 total number density is also strongly affected by elastic scattering.
@@
 postmerger configuration composed of a single neutron star surrounded by a disk.
-Because of strong differential rotation, the remnant temporarily avoids
+Because of strong differential rotation and shock heating,
+the remnant temporarily avoids
 collapse to a black hole, even if its mass exceeds the threshold of dynamical
@@
 at $t=11\,{\rm ms}$ after merger.
+In this way we approximate the system as stationary over a light-crossing time
+of around 1~ms, which is far below the thermal timescale of the remnant.
 The volume data are represented on a pseudospectral domain of a
@@
 asymmetries in the configuration.
-It is not due to the Doppler shift from the rapid rotation of the star
-(a hypothesis we tested by setting $u_i=0$ and $W=1$ in the ray
-tracing equations).
+It is not due to the Doppler shift or relativistic beaming
+from the rapid rotation of the star,
+a hypothesis we tested by setting $u_i=0$ and $W=1$ in the ray
+tracing equations.
 Because of their larger absorption cross section,
@@
 and observers within view of the hot hypermassive neutron star measure
-the largest radial fluxes.
+the largest radial fluxes of these species.
 
@@
 also obey the same ranking.
-As we showed in the optically thin limit of
-Eqn.~\ref{eqn:fse_homogeneous_thin_limit},
-\footnote{TODO: too simplistic}
+As we showed in the toy models with a homogeneous background scattering field
 the scattering contribution $f_{\rm SE}$ is
-proportional to energy $\varepsilon^2$ and path length $s$.
+proportional to energy $\varepsilon^2$ and path length $s$
+(Eqn.~\ref{eqn:fse_homogeneous_thin_limit}).
 
@@
 The Boltzmann Equation (Eqn.~\ref{eqn:boltzmann}) is one limiting
-form of the quantum kinetic equations \cite{vlas2014-neutrino_qkes},
+form of the quantum kinetic equations for neutrinos
+\cite{vlas2014-neutrino_qkes},
 the limit where collisional mean free paths are much shorter
-than oscillation lengths (i.e.\ $\varepsilon/\mathscr{K} \ll |H|^{-1}$
-with $H$ the Hamiltonian describing coherent forward scattering interactions).
+than oscillation lengths, i.e.\ $\varepsilon/\mathscr{K} \ll |H|^{-1}$
+with $H$ the Hamiltonian describing coherent forward scattering interactions.
 In this limit the neutrino density matrix takes the form

### 2017-11-28 paper.tex
Specify homogeneous thermodynamic state in text

@@
 representative of the fluid thermodynamic state at a radius of 50~km
-in the test presented in Sec.~\ref{ssec:test_collapse}
-(a massive collapsed star following core bounce).
+in the test presented in Sec.~\ref{ssec:test_collapse}---a
+massive collapsed star following core bounce---where
+$\varrho=10^{11}\,{\rm g}\,{\rm cm}^{-3}$, $T=3.7$~MeV, and $Y_e=0.12$.
 For comparision with that test we use the LS180 equation of state

### 2017-11-30 paper.tex
Improve intro text

@@
 $f^{\nu_\sigma}(x_o^\alpha;p_\beta)$, or the amount of neutrino radiation
-of species $\nu_\sigma\in\{\nu_{\{e,\mu,\tau\}},\bar{\nu}_{\{e,\mu,\tau\}}\}$
+of species
+$\nu_\sigma\in\{\nu_e,\nu_\mu,\nu_\tau,\bar{\nu}_e,\bar{\nu}_\mu,\bar{\nu}_\tau\}$
 with momentum $p_\beta$ impinging on $x_o^\alpha$.
@@
 \footnote{TODO: state scaling $\mathscr{O}(n)$?}
-And it easily extends to high resolution in energy and angle by simply
+It easily extends to high resolution in energy and angle by simply
 increasing the number of rays sampled.
+And it parallelizes trivially by ray, since each ray is computed
+independently.
 
@@
 \citep{birk2007-nunubar, caba2009-detecting_grb_nu, kova2011-gr_ray_tracing},
-limiting them to equilibrium, optically thick configurations.
+limiting them to equilibrium, optically-thick configurations.
 Current state-of-the-art ray tracing formulations avoid the assumption of
@@
 \footnote{TODO: clarify how momenta are decoupled}
-When solving for $f^{\nu_\sigma}(x^\alpha;p_\beta)$, we have no
-information about $f^{\nu_\sigma}(x^\alpha;p'_\beta)$, and
-$f^{\bar{\nu}_\sigma}(x^\alpha;p'_\beta)$, which are essential ingredients
+When solving for $f^{\nu_\sigma}(x^\alpha;p_\beta)$,
+we have no information about the distribution at different momenta
+$f^{\nu_\sigma}(x^\alpha;p'_\beta)$,
+or about the distribution of the relevant antineutrino species
+$f^{\bar{\nu}_\sigma}(x^\alpha;p'_\beta)$;
+these missing data are essential ingredients
 for the source terms of the Boltzmann Equation describing the creation and
@@
 Where we do reference particular species we use the three categories
-$\nu_e$, $\bar{\nu}_e$, and
+relevant to the energy scales of mergers, $\nu_e$, $\bar{\nu}_e$, and
 $\nu_x=\left\{\nu_\mu,\bar{\nu}_\mu,\nu_\tau,\bar{\nu}_\tau\right\}$.

### 2017-12-05 paper.tex
Update Table II HMNS luminosities for new compute-eta method
Update HMNS text to describe new compute-eta data in Table II

@@
 For the purposes of this analysis the pseudospectral representation is adequate.
+
 Figs.~\ref{fig:nsns_rho_merid}--\ref{fig:nsns_temp_equat}
 show slices of density and temperature from the finite difference volume data.
+These data were evolved on a rectilinear grid spanning approximately 400~km in
+both the $x$ and $y$ directions, and 150~km in the $z$ direction in our
+evolution coordinates. We extrapolate the fluid and M1 radiation data from
+this grid to the larger ray tracing grid by setting all variables to their
+floor values. This simple extrapolation is adequate for ray tracing since
+the neutrinos are almost entirely free-streaming outside the smaller domain.
 
@@
 $\langle\tilde{\varepsilon}\rangle(\theta)=H^r(\theta)/K^r(\theta)$.
-We choose $N_{\tilde{\varepsilon}}=30$, $N_A=120$, and $N_B=20$.
-For the \emph{scat} case we choose $(\cos A)_{\rm min}=0.81$;
-For the \emph{noscat} case, presenting a smaller emitting region,
-we choose $(\cos A)_{\rm min}=0.91$.
+We choose $N_{\tilde{\varepsilon}}=30$, $N_A=150$, and $N_B=20$.
+To maintain high angular resolution, we only sample rays that pass within
+approximately 120~km of the star's center by setting $(\cos A)_{\rm min}=0.88$.
 
@@
 total luminosities and averages over the sky.
-We assume the disk is azimuthally symmetric and
-symmetric with respect to reflection across the equatorial plane.
-Figs.~\ref{fig:nsns_rho_merid} and \ref{fig:nsns_rho_equat} show how these
-assumptions are approximately valid at 11~ms after merger.
+Since we have chosen an arc of observers isolated to the northern hemisphere and
+the $y$-$z$ plane, we may extend these data to the full sky by assuming that
+emission is azimuthally symmetric and reflection symmetric across the
+equatorial plane. Figs.~\ref{fig:nsns_rho_merid}--\ref{fig:nsns_temp_equat}
+indicate the approximate validity of these assumptions at 11~ms after merger.
 Total luminosities are then computed as integrals of the radial fluxes
@@
 \begin{table}%[h]
-  % Total fluxes over the sky integrated in google sheet: 20170726-theta_distribution_integrations
-  % RTDATA_NO=$DIR/PositionDistributions-nu?-250km-Lev1-e30a120b20-noscat-01.dat
-  % RTDATA_GR=$DIR/PositionDistributions-nu?-250km-Lev1-e30a60b10-scat-01.dat
+  % Total fluxes over the sky integrated in google sheet: 20171031-theta_distribution_integrations
+  % RTDATA_NO=$DIR/PositionDistributions-nu?-250km-Lev2-e30a150b20-noscat-04.dat
+  % RTDATA_SC=$DIR/PositionDistributions-trim-nu?-250km-Lev2-e30a150b20-scat-04.dat
   % DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20170324-ray_tracing_nsns/Run03/PositionDistributions
@@
     \hline
-    $L_{\nu_e}$                                 & 5.62 & 6.05 & 5 \\
-    $L_{\bar{\nu}_e}$                           & 10.6 & 11.8 & 12 \\
-    $L_{\nu_x}$                                 & 23.3 & 12.7 & 12 \\
-    $R_{\nu_e}-R_{\bar{\nu}_e}$                 & -1.22 & -1.85 & -2 \\
-    $\langle \varepsilon_{\nu_e} \rangle$       & 12.08 & 11.91 & 12 \\
-    $\langle \varepsilon_{\bar{\nu}_e} \rangle$ & 16.07 & 14.67 & 15 \\
-    $\langle \varepsilon_{\nu_x} \rangle$       & 34.09 & 23.58 & 26 \\
+    $L_{\nu_e}$                                 & 5.87 & 5.40 & 5 \\
+    $L_{\bar{\nu}_e}$                           & 9.70 & 10.7 & 12 \\
+    $L_{\nu_x}$                                 & 23.6 & 12.5 & 12 \\
+    $R_{\nu_e}-R_{\bar{\nu}_e}$                 & -0.91 & -1.68 & -2 \\
+    $\langle \varepsilon_{\nu_e} \rangle$       & 12.7 & 12.0 & 12 \\
+    $\langle \varepsilon_{\bar{\nu}_e} \rangle$ & 16.0 & 14.8 & 15 \\
+    $\langle \varepsilon_{\nu_x} \rangle$       & 34.2 & 23.3 & 26 \\
     \hline
@@
 Tab.~\ref{tab:nsns_rt_vs_m1} compares the all-sky
-\footnote{TODO: get accurate M1 numbers from Francois}
 luminosities and average energies from ray tracing and from
 the M1 transport simulation,
-which serves as our fiducial calculation.
-Uncertainties are around 15\% for energy luminosities and
-10\% for average energies from the M1 simulation
-\cite[Sec.~A.6]{fouc2016-m1_evolve_n}.
+which serves for qualitative comparison.
+Even if M1 and ray tracing methods both provide faithful measurements
+of all-sky luminosities, we do not expect perfect agreement since the two
+treatments differ fundamentally. In addition to differences in the physics,
+M1 fluxes are integrated over the outer boundary of the finite-difference grid
+at radii as small as 75~km,
+whereas the ray tracing fluxes are integrated over a sphere at radius 250~km,
+introducing a time lag between the two measurements of under a millisecond.
+Furthermore uncertainties from the M1 simulation are around
+15\% for energy luminosities and
+10\% for average energies \cite[Sec.~A.6]{fouc2016-m1_evolve_n}.
+
 Because the model was evolved with a gray M1 transport scheme,
@@
 significantly so for the heavy-lepton neutrinos.
-(Though, we note the energy luminosity for $\nu_e$ is overestimated
-20\% by both \emph{noscat} and \emph{scat} treatments,
-and the overestimate is slightly worse for the \emph{scat} treatment.)
-\footnote{TODO: explain}
-With scattering turned off ray tracing overestimates $L_{\nu_x}$ by 200\%;
-but with scattering turned on agreement in all measures
-is better than 10\%, except for $L_{\nu_e}$ which is overestimated by 20\%.
-As the M1 transport evolution also demonstrates,
-$R_{\bar{\nu}_e}$ dominates over $R_{\nu_e}$
-\footnote{TODO: Need $R$ magnitude not just difference}
+For example, with scattering turned off the ray tracing and M1 measurements
+of $L_{\nu_x}$ disagree by 200\%;
+but with scattering turned on agreement is within 10\%.
+And as the ray tracing and M1 measurements both show,
+$R_{\bar{\nu}_e}$ dominates over $R_{\nu_e}$ significantly,
 because the disrupted neutron star material is releptonizing.

### 2017-12-05 paper.tex
Update Fig. 12 HMNS K^r, with new compute-eta data

@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{theta_distrib-250km-Kr}
-  % Data_a:   $DIR/PositionDistributions/*Lev1-e30a120b20-01*
-  % Data_s:   $DIR/PositionDistributions/*Lev1-e30a60b10-01*
+  %\includegraphics[width=\columnwidth]{theta_distrib-250km-Kr}
+  \resizebox{\columnwidth}{!}{\input{theta_distrib-250km-Kr.tex}}
+  % Data_a:   $DIR/PositionDistributions/*Lev2-e30a150b20-noscat-04*
+  % Data_s:   $DIR/PositionDistributions/*Lev2-e30a150b20-scat-04*
   % Data gathered using: $DIR/CollectPositionDistributions.sh
@@
   % w/ DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20170324-ray_tracing_nsns/Run03
-  \caption{Radial number fluxes, $K^r(\theta)$,
+  \caption{Radial number fluxes, $\tilde{K}^r(\theta)$,
     as defined in Eqn.~\ref{eqn:Kr_minkowski},

### 2017-12-05 paper.tex
Update Fig. 13 HMNS H^r, with new compute-eta data
Update Fig. 14 HMNS <eps>, with new compute-eta data

@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{theta_distrib-250km-Hr}
+  %\includegraphics[width=\columnwidth]{theta_distrib-250km-Hr}
+  \resizebox{\columnwidth}{!}{\input{theta_distrib-250km-Hr.tex}}
   % see notes from theta_distrib-250km-Kr
@@
   \caption{Same as Fig.~\ref{fig:nsns_theta_distrib_Kr} but radial energy
-    fluxes $H^r(\theta)$ as defined in Eqn.~\ref{eqn:Hr_minkowski}.
+    fluxes $\tilde{H}^r(\theta)$ as defined in Eqn.~\ref{eqn:Hr_minkowski}.
   }
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{theta_distrib-250km-avg_eps}
+  %\includegraphics[width=\columnwidth]{theta_distrib-250km-avg_eps}
+  \resizebox{\columnwidth}{!}{\input{theta_distrib-250km-avg_eps.tex}}
   % see notes from theta_distrib-250km-Kr
@@
     average energies of radial fluxes
-    $\langle\tilde{\varepsilon}\rangle(\theta)=H^r(\theta)/K^r(\theta)$,
+    $\langle\tilde{\varepsilon}\rangle(\theta)=\tilde{H}^r(\theta)/\tilde{K}^r(\theta)$,
     as defined in Eqns.~\ref{eqn:Kr_minkowski} and \ref{eqn:Hr_minkowski}.

### 2017-12-05 paper.tex
Reposition all figures to *follow* explanatory text
Point out reversed <eps>(theta) trend for nux

@@
 
+We sample the distribution function $f(\tilde{\varepsilon},\cos A,B)$
+with ray tracing
+over a uniform grid of 40 points in energy $\tilde{\varepsilon}\in[0,100]$
+and 30 points in angle $\cos A\in[-1,1]$, holding fixed $B=\pi$.
+Results are shown in Fig.~\ref{fig:avg_eps_doppler} for
+the velocities $v=\{0,0.1,0.8\}$.
+The ray tracing results are computed from total densities in each angular bin,
+that is
+\begin{equation}
+  \label{eqn:avg_eps_per_cosA}
+  \langle \tilde{\varepsilon} \rangle (\cos A) =
+  \frac{\tilde{J}(\cos A)}{\tilde{G}(\cos A)},
+\end{equation}
+with the Eulerian energy density per angular bin given by
+\begin{equation}
+  \label{eqn:J_per_cosA}
+  \tilde{J}(\cos A) =
+  \frac{1}{(2\pi)^3} \int_0^{\infty} d\tilde{\varepsilon} \, \tilde{\varepsilon}^3
+  \int_0^{2\pi} dB \, f(\tilde{\varepsilon},\cos A,B),
+\end{equation}
+and $\tilde{G}(\cos A)$ defined analogously.
+\footnote{TODO: add metric terms for these integrals}
+Results are identical for all \emph{scat} and \emph{noscat} methods.
+In Fig.~\ref{fig:avg_eps_doppler}
+we see the common features of a red-shifted spectrum for receding fluid
+($\cos A\ll0$),
+a blue-shifted spectrum for approaching fluid
+($\cos A\gg0$),
+and a slightly red-shifted spectrum for fluid moving transverse to the
+observer ($\cos A=0$).
+
 \begin{figure}
@@
 
-We sample the distribution function $f(\tilde{\varepsilon},\cos A,B)$
-with ray tracing
-over a uniform grid of 40 points in energy $\tilde{\varepsilon}\in[0,100]$
-and 30 points in angle $\cos A\in[-1,1]$, holding fixed $B=\pi$.
-Results are shown in Fig.~\ref{fig:avg_eps_doppler} for
-the velocities $v=\{0,0.1,0.8\}$.
-The ray tracing results are computed from total densities in each angular bin,
-that is
-\begin{equation}
-  \label{eqn:avg_eps_per_cosA}
-  \langle \tilde{\varepsilon} \rangle (\cos A) =
-  \frac{\tilde{J}(\cos A)}{\tilde{G}(\cos A)},
-\end{equation}
-with the Eulerian energy density per angular bin given by
-\begin{equation}
-  \label{eqn:J_per_cosA}
-  \tilde{J}(\cos A) =
-  \frac{1}{(2\pi)^3} \int_0^{\infty} d\tilde{\varepsilon} \, \tilde{\varepsilon}^3
-  \int_0^{2\pi} dB \, f(\tilde{\varepsilon},\cos A,B),
-\end{equation}
-and $\tilde{G}(\cos A)$ defined analogously.
-\footnote{TODO: add metric terms for these integrals}
-Results are identical for all \emph{scat} and \emph{noscat} methods.
-In Fig.~\ref{fig:avg_eps_doppler}
-we see the common features of a red-shifted spectrum for receding fluid
-($\cos A\ll0$),
-a blue-shifted spectrum for approaching fluid
-($\cos A\gg0$),
-and a slightly red-shifted spectrum for fluid moving transverse to the
-observer ($\cos A=0$).
-
 \subsection{Idealized Star:
@@
 thin regime.
-
-We can explain these features quantitatively by examining the limits
-of Eqn.~\ref{eqn:fae_homogeneous_general}, expanding the exponential function
-in powers of $\chi s$; the distribution function takes the limiting values
-\begin{equation}
-  \label{eqn:fae_homogeneous_limits}
-  f_{\rm AE}(\cos A) = f^{\rm eq}
-  \begin{cases}
-    \chi s(\cos A) & s \ll \chi^{-1}\,\textrm{(thin)}\\
-    1                  & s \gg \chi^{-1}\,\textrm{(thick)}.
-  \end{cases}
-\end{equation}
-These limits are represented in Fig.~\ref{fig:f_absorption_sphere}:
-with $\nu_x$ in the optically thin limit at all viewing angles,
-and $\nu_e$ in the optically thick limit at viewing angles
-$\cos A \gtrsim 0.8$.
 
@@
 
+We can explain these features quantitatively by examining the limits
+of Eqn.~\ref{eqn:fae_homogeneous_general}, expanding the exponential function
+in powers of $\chi s$; the distribution function takes the limiting values
+\begin{equation}
+  \label{eqn:fae_homogeneous_limits}
+  f_{\rm AE}(\cos A) = f^{\rm eq}
+  \begin{cases}
+    \chi s(\cos A) & s \ll \chi^{-1}\,\textrm{(thin)}\\
+    1                  & s \gg \chi^{-1}\,\textrm{(thick)}.
+  \end{cases}
+\end{equation}
+These limits are represented in Fig.~\ref{fig:f_absorption_sphere}:
+with $\nu_x$ in the optically thin limit at all viewing angles,
+and $\nu_e$ in the optically thick limit at viewing angles
+$\cos A \gtrsim 0.8$.
+
 \subsection{Idealized Compact Star:
@@
 since the Monte Carlo code assumes Minkowski spacetime.
+
+We place a stationary observer at $r=500$~km.
+Fig.~\ref{fig:avg_eps_collapse} shows the distribution of average $\nu_x$
+energies across incoming angles to this observer.
+We show $\nu_x$ because they present the largest scattering effects:
+they scatter through a thicker atmosphere outside their deep emission surface,
+and their hotter spectrum experiences stronger modification due to the
+$\varepsilon^2$ dependence of the scattering cross-section.
+Against the fiducial Monte Carlo distribution, we show the
+\emph{noscat} treatment, and the \emph{scat} treatment using both the
+\emph{spectral} method and the \emph{gray} method.
 
@@
 \end{figure}
-
-We place a stationary observer at $r=500$~km.
-Fig.~\ref{fig:avg_eps_collapse} shows the distribution of average $\nu_x$
-energies across incoming angles to this observer.
-We show $\nu_x$ because they present the largest scattering effects:
-they scatter through a thicker atmosphere outside their deep emission surface,
-and their hotter spectrum experiences stronger modification due to the
-$\varepsilon^2$ dependence of the scattering cross-section.
-Against the fiducial Monte Carlo distribution, we show the
-\emph{noscat} treatment, and the \emph{scat} treatment using both the
-\emph{spectral} method and the \emph{gray} method.
 
@@
 
+Tab.~\ref{tab:nsns_rt_vs_m1} compares the all-sky
+luminosities and average energies from ray tracing and from
+the M1 transport simulation,
+which serves for qualitative comparison.
+Even if M1 and ray tracing methods both provide faithful measurements
+of all-sky luminosities, we do not expect perfect agreement since the two
+treatments differ fundamentally. In addition to differences in the physics,
+M1 fluxes are integrated over the outer boundary of the finite-difference grid
+at radii as small as 75~km,
+whereas the ray tracing fluxes are integrated over a sphere at radius 250~km,
+introducing a time lag between the two measurements of under a millisecond.
+Furthermore uncertainties from the M1 simulation are around
+15\% for energy luminosities and
+10\% for average energies \cite[Sec.~A.6]{fouc2016-m1_evolve_n}.
+
 \begin{table}%[h]
@@
 
-Tab.~\ref{tab:nsns_rt_vs_m1} compares the all-sky
-luminosities and average energies from ray tracing and from
-the M1 transport simulation,
-which serves for qualitative comparison.
-Even if M1 and ray tracing methods both provide faithful measurements
-of all-sky luminosities, we do not expect perfect agreement since the two
-treatments differ fundamentally. In addition to differences in the physics,
-M1 fluxes are integrated over the outer boundary of the finite-difference grid
-at radii as small as 75~km,
-whereas the ray tracing fluxes are integrated over a sphere at radius 250~km,
-introducing a time lag between the two measurements of under a millisecond.
-Furthermore uncertainties from the M1 simulation are around
-15\% for energy luminosities and
-10\% for average energies \cite[Sec.~A.6]{fouc2016-m1_evolve_n}.
-
 Because the model was evolved with a gray M1 transport scheme,
@@
 
+The distributions of radial fluxes over observer position are shown
+in Figs.~\ref{fig:nsns_theta_distrib_Kr} and \ref{fig:nsns_theta_distrib_Hr}.
+With the \emph{noscat} treatment the $\nu_x$ luminosities are relatively
+constant in $\theta$, because the disk is optically thin to $\nu_x$,
+and most of the neutrinos come from the hypermassive neutron star,
+which is roughly spherical.
+The slight upward trend in $\nu_x$ luminosities with $\theta$ may be due to
+asymmetries in the configuration.
+It is not due to the Doppler shift or relativistic beaming
+from the rapid rotation of the star,
+a hypothesis we tested by setting $u_i=0$ and $W=1$ in the ray
+tracing equations.
+Because of their larger absorption cross section,
+the disk is not optically thin to $\nu_e$ and $\bar{\nu}_e$,
+and observers within view of the hot hypermassive neutron star measure
+the largest radial fluxes of these species.
 \begin{figure}
@@
 
-The distributions of radial fluxes over observer position are shown
-in Figs.~\ref{fig:nsns_theta_distrib_Kr} and \ref{fig:nsns_theta_distrib_Hr}.
-With the \emph{noscat} treatment the $\nu_x$ luminosities are relatively
-constant in $\theta$, because the disk is optically thin to $\nu_x$,
-and most of the neutrinos come from the hypermassive neutron star,
-which is roughly spherical.
-The slight upward trend in $\nu_x$ luminosities with $\theta$ may be due to
-asymmetries in the configuration.
-It is not due to the Doppler shift or relativistic beaming
-from the rapid rotation of the star,
-a hypothesis we tested by setting $u_i=0$ and $W=1$ in the ray
-tracing equations.
-Because of their larger absorption cross section,
-the disk is not optically thin to $\nu_e$ and $\bar{\nu}_e$,
-and observers within view of the hot hypermassive neutron star measure
-the largest radial fluxes of these species.
-
 When we turn on scattering, the disk is no longer transparent to $\nu_x$,
@@
 
-\begin{figure}
-  %\includegraphics[width=\columnwidth]{theta_distrib-250km-avg_eps}
-  \resizebox{\columnwidth}{!}{\input{theta_distrib-250km-avg_eps.tex}}
-  % see notes from theta_distrib-250km-Kr
-  % Script: $DIR/GnuScripts/plot-theta_distributions-avg_eps_from_fluxes-fixed_r.gnu
-  \caption{Same as Fig.~\ref{fig:nsns_theta_distrib_Kr} but
-    average energies of radial fluxes
-    $\langle\tilde{\varepsilon}\rangle(\theta)=\tilde{H}^r(\theta)/\tilde{K}^r(\theta)$,
-    as defined in Eqns.~\ref{eqn:Kr_minkowski} and \ref{eqn:Hr_minkowski}.
-  }
-  \label{fig:nsns_theta_distrib_avg_eps}
-\end{figure}
-
 Fig.~\ref{fig:nsns_theta_distrib_avg_eps} shows the distribution of
@@
 direct view of the hot hypermassive neutron star.
+(This trend is unexpectedly reversed for $\nu_x$,
+and may be due to asymetries in the disk, as discussed above.)
 Scattering decreases average energies across all observer positions,
@@
 (Eqn.~\ref{eqn:fse_homogeneous_thin_limit}).
+\begin{figure}
+  %\includegraphics[width=\columnwidth]{theta_distrib-250km-avg_eps}
+  \resizebox{\columnwidth}{!}{\input{theta_distrib-250km-avg_eps.tex}}
+  % see notes from theta_distrib-250km-Kr
+  % Script: $DIR/GnuScripts/plot-theta_distributions-avg_eps_from_fluxes-fixed_r.gnu
+  \caption{Same as Fig.~\ref{fig:nsns_theta_distrib_Kr} but
+    average energies of radial fluxes
+    $\langle\tilde{\varepsilon}\rangle(\theta)=\tilde{H}^r(\theta)/\tilde{K}^r(\theta)$,
+    as defined in Eqns.~\ref{eqn:Kr_minkowski} and \ref{eqn:Hr_minkowski}.
+  }
+  \label{fig:nsns_theta_distrib_avg_eps}
+\end{figure}
 

### 2017-12-06 paper.tex
Update Figs. 15 & 16 HMNS G(cosA), with new compute-eta data

@@
 
-In Figs.~\ref{fig:nsns_cosA_distrib_I_00deg}
-and \ref{fig:nsns_cosA_distrib_I_90deg}, we show the distributions of neutrino
+In Figs.~\ref{fig:nsns_cosA_distrib_G_00deg}
+and \ref{fig:nsns_cosA_distrib_G_90deg}, we show the distributions of neutrino
 number density over incoming angle
@@
 and in the equatorial plane $\theta=90^\circ$,
-as defined in Eqn.~\ref{eqn:J_per_cosA}.
+defined analagously to Eqn.~\ref{eqn:J_per_cosA}.
 As in the case of the collapse profile in Sec.~\ref{ssec:test_collapse},
 the dominant effect of \emph{scat} on $\tilde{G}(\cos A)$
-is to spread the distribution out to larger angles.
+is to spread the distribution out to larger angles, by generally decreasing
+the number of neutrinos coming from the core while increasing the number
+of neutrinos coming from the disk.
 \footnote{TODO: also examine spectra or average energies}
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{cosA_distrib-intensity-250km-00deg}
-  % Data:   $DIR/Lev1-e30a60b10-{,no}scat-00
-  % Script: $DIR/plot-angular_distribution-intensity-compare.gnu
+  %\includegraphics[width=\columnwidth]{cosA_distrib-intensity-250km-00deg}
+  \resizebox{\columnwidth}{!}{\input{cosA_distrib-G-250km-00deg.tex}}
+  % Data:   $DIR/Lev2-e30a150b20-{,no}scat-04
+  % Script: $DIR/plot-angular_distribution-number_density-compare.gnu
   % w/ DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20170324-ray_tracing_nsns/Run03/000Mo-000.0deg-250km
-  \caption{Angular distributions of fluxes for an observer on the rotation axis
-    in the hypermassive neutron star configuration
+  \caption{Distribution of number density $\tilde{G}(\cos A)$ over incoming angle,
+    defined analagously to Eqn.~\ref{eqn:J_per_cosA}.
+    The integral of this quantity over $d\cos A$ gives the total number density
+    measured by this observer on the rotation axis.
+    Volume data from the hypermassive neutron star configuration
     (presented in Sec.~\ref{ssec:test_disk_comparison}).}
-  \label{fig:nsns_cosA_distrib_I_00deg}
+  \label{fig:nsns_cosA_distrib_G_00deg}
 \end{figure}
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{cosA_distrib-intensity-250km-90deg}
-  % Data:   $DIR/Lev1-e30a60b10-{,no}scat-00
-  % Script: $DIR/plot-angular_distribution-intensity-compare.gnu
+  %\includegraphics[width=\columnwidth]{cosA_distrib-intensity-250km-90deg}
+  \resizebox{\columnwidth}{!}{\input{cosA_distrib-G-250km-90deg.tex}}
+  % Data:   $DIR/Lev2-e30a150b20-{,no}scat-04
+  % Script: $DIR/plot-angular_distribution-number_density-compare.gnu
   % w/ DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20170324-ray_tracing_nsns/Run03/000Mo-090.0deg-250km
-  \caption{Angular distributions of fluxes for an observer in the equatorial
-    plane in the hypermassive neutron star configuration
-    (presented in Sec.~\ref{ssec:test_disk_comparison}).}
-  \label{fig:nsns_cosA_distrib_I_90deg}
+  \caption{Distribution of number density $\tilde{G}(\cos A)$ over incoming angle,
+    as in Fig.~\ref{fig:nsns_cosA_distrib_G_00deg},
+    but with the observer in the equatorial plane.}
+  \label{fig:nsns_cosA_distrib_G_90deg}
 \end{figure}

### 2017-12-07 paper.tex
Update Fig. 2 MFPs in post-bounce profile at 50km to gnuplot epslatex
Update Fig. 4 Error in f, to gnuplot epslatex
Update Fig. 3 f along a ray, to gnuplot epslatex

@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{figure-mfps-50km}
+  %\includegraphics[width=\columnwidth]{figure-mfps-50km}
+  \resizebox{\columnwidth}{!}{\input{figure-mfps-50km.tex}}
   % Data:   $DIR/table-50km-nu?.dat
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{fig-cumulative_f-homogeneous}
+  %\includegraphics[width=\columnwidth]{fig-cumulative_f-homogeneous}
+  \resizebox{\columnwidth}{!}{\input{fig-cumulative_f-homogeneous.tex}}
   % Data:   $DIR/Lev1-v0-ShootANeutrino-01/Lev1
@@
     greater than $\tau_{\rm term}=14$.
-    The positions of the points (only plotted for the $f_{\rm AE}$ components)
-    are the time steps chosen by the adaptive time-stepping algorithm
-    described in Sec.~\ref{ssec:numerical}.
+    The points plotted correspond to the time steps chosen by the adaptive
+    time-stepping algorithm described in Sec.~\ref{ssec:numerical}.
   }
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{figure-ferr-50km}
+  %\includegraphics[width=\columnwidth]{figure-ferr-50km}
+  \resizebox{\columnwidth}{!}{\input{figure-ferr-50km.tex}}
   % Data:   $DIR/Lev1-v0-noscat-01/Lev1

### 2017-12-07 paper.tex
Update Fig. 6 ideal star f, to gnuplot epslatex
Update Fig. 7 collapse profile <eps>, to gnuplot epslatex

@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{fig-f-absorption_star-11MeV}
+  %\includegraphics[width=\columnwidth]{fig-f-absorption_star-11MeV}
+  \resizebox{\columnwidth}{!}{\input{fig-f-absorption_star-11MeV.tex}}
   % Data:   $DIR/Lev1-01/Lev1
@@
 The background scattering field $\Phi(\varepsilon)$ is supplied by
-\lstinline{GR1D} in the form of $J(\varepsilon_i)$,
+\lstinline{GR1D} in the form of $J(\tilde{\varepsilon}_i)$,
 with $\varepsilon_i$ representing 18 energy groups
@@
 Sec.~\ref{ssec:test_equilibrium}.
-For the \emph{spectral} method we use $J(\varepsilon_i)$ directly,
+For the \emph{spectral} method we use $J(\tilde{\varepsilon}_i)$ directly,
 using zeroth-order interpolation between energy groups;
-for the \emph{gray} method we use $J=\sum J(\varepsilon_i) \Delta\varepsilon_i$,
-and $G=\sum J(\varepsilon_i) \Delta\varepsilon_i/\varepsilon_i$ instead,
-with $\Delta\varepsilon_i$ the bin-width of the $i$th energy group.
+for the \emph{gray} method we use $J=\sum J(\tilde{\varepsilon}_i) \Delta\tilde{\varepsilon}_i$,
+and $G=\sum J(\tilde{\varepsilon}_i) \Delta\tilde{\varepsilon}_i/\tilde{\varepsilon}_i$ instead,
+with $\Delta\tilde{\varepsilon}_i$ the bin-width of the $i$th energy group.
 
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{fig-collapse_profile_avg_eps_vs_cosA}
+  %\includegraphics[width=\columnwidth]{fig-collapse_profile_avg_eps_vs_cosA}
+  \resizebox{\columnwidth}{!}{\input{fig-collapse_profile_avg_eps_vs_cosA.tex}}
   % Data:   $DIR/
@@
   \caption{Angular distribution of $\nu_x$ average energies
-    $\langle\varepsilon\rangle(\cos A)$
+    $\langle\tilde{\varepsilon}\rangle(\cos A)$
     in the collapse profile test (described in Sec.~\ref{ssec:test_collapse}).

### 2017-12-08 paper.tex
Add structure for conclusion section
Matt's edits through p. 14: warn about sloppy treatment of flavor ev

@@
 including the effects of coherent forward scattering with ambient neutrinos.
+In Sec.~\ref{sec:conclusions} we summarize our work and anticipate improvements.
 
@@
 and 24 cells spanning azimuthal angles $\phi\in[0,2\pi]$.
-\footnote{TODO: way too-low angular resolution}
 Though a pseudospectral representation of non-smooth hydrodynamic data
@@
 \label{ssec:V_nunu}
+In this section we examine the importance of elastic scattering in neutrino
+flavor oscillation above the hypermassive neutron star--disk configuration
+presented in Sec.~\ref{ssec:test_disk_comparison}.
+Our treatment of the flavor evolution equation assumes flat space and small fluid
+velocities.
+In consequence we treat some of gauge-dependent quantities inconsistently,
+and we ignore potentially important features of relativistic flavor evolution
+near compact objects.
+\footnote{TODO: cite Yang et Kneller}
+Such a cursory treatment is sufficient, however, for the following qualitative
+exploration.
+
 The Boltzmann Equation (Eqn.~\ref{eqn:boltzmann}) is one limiting
@@
 \end{equation}
-where $ds$ is an interval of proper length, and
+where $ds$ is an interval of proper length traversed by the neutrino as
+measured by our observer, and
 \footnote{TODO: mention $\bar{\nu}$ eqn}
@@
 %a difference (or not) in our estimate of the power available to drive a jet.
+
+\section{Conclusions}
+\label{sec:conclusions}
+Add some conclusions.
+\footnote{TODO: note some improvements needed: implement implicit time-stepping,
+  use finite-difference grid for fluid vars, improve gray spectral assumptions}
 

### 2017-12-08 paper.tex
Add a few minor edits

@@
 $R_\nu\approx\{50,45,35\}$~km.
+\footnote{TODO: use a consistent lengthscale symbol throughout paper. $d$? $D$? $r$? $l$?}
 % From by-hand summation over histogram G(cosA_i)*dcosA
@@
 Total luminosities and average energies from the different treatments are
-\footnote{TODO: update MC results 02->03}
+\footnote{TODO: update MC results 02$\rightarrow0$3}
 presented in Tab.~\ref{tab:collapse_rt_vs_mc}.
@@
 of all-sky luminosities, we do not expect perfect agreement since the two
-treatments differ fundamentally. In addition to differences in the physics,
+treatments differ fundamentally. In addition to differences in the physics
+such as the treatment of inelastic scattering,
 M1 fluxes are integrated over the outer boundary of the finite-difference grid

### 2017-12-15 paper.tex
Rearrange description of smaller and larger domains for HMNS case

@@
 was simulated using a gray M1 transport scheme for the neutrinos, evolving the
-energy density, number density, and energy flux \cite{fouc2016-m1_evolve_n}
-for $\sim10\,{\rm ms}$ following merger.
+energy density, number density, and energy flux in addition to the standard
+fluid and metric variables for $\sim10\,{\rm ms}$ following merger
+\cite{fouc2016-m1_evolve_n}.
 The fluid was modeled using the LS220 equation of state \cite{latt1991-nuc_eos}.
@@
 of around 1~ms, which is far below the thermal timescale of the remnant.
-The volume data are represented on a pseudospectral domain of a
-sphere with concentric shells extending to $r\approx1400$~km.
+Figs.~\ref{fig:nsns_rho_merid}--\ref{fig:nsns_temp_equat}
+show slices of density and temperature from the finite difference
+fluid and M1 radiation data.
+These data were evolved on a rectilinear grid spanning approximately 400~km in
+both the $x$ and $y$ directions, and 150~km in the $z$ direction in our
+evolution coordinates.
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{production-colormap-merid-rho}
+  % VTK Data: zosima:/home/mbdeaton/Research/20170602-nsns_vol_data
+  % Plots:    $DIR/production-colormap-{merid,equat}-{rho,temp}.png
+  % w/ DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20170420-nsns_volume_data/02-two_domains_avg_eps/PlotsEtc
+  \caption{A meridional slice of density in the hypermassive neutron
+    star and disk configuration (Sec.~\ref{ssec:test_disk_comparison}).
+    The distorted rectangular boundaries are the boundaries of the grid
+    used in the numerical simulation, which employs a coordinate mapping to
+    concentrate points near the central object.}
+  \label{fig:nsns_rho_merid}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{production-colormap-equat-rho}
+  % see notes from production-colormap-merid-rho
+  \caption{An equatorial slice of density in the hypermassive neutron
+    star and disk configuration (Sec.~\ref{ssec:test_disk_comparison}).}
+  \label{fig:nsns_rho_equat}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{production-colormap-merid-temp}
+  % see notes from production-colormap-merid-rho
+  \caption{A meridional slice of temperature in the hypermassive neutron
+    star and disk configuration (Sec.~\ref{ssec:test_disk_comparison}).}
+  \label{fig:nsns_temp_merid}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{production-colormap-equat-temp}
+  % see notes from production-colormap-merid-rho
+  \caption{An equatorial slice of temperature in the hypermassive neutron
+    star and disk configuration (Sec.~\ref{ssec:test_disk_comparison}).}
+  \label{fig:nsns_temp_equat}
+\end{figure}
+
+We extrapolate the fluid and M1 radiation data from
+the domain shown in Figs.~\ref{fig:nsns_rho_merid}--\ref{fig:nsns_temp_equat}
+to a larger ray tracing domain by setting all fluid and M1 radiation variables
+to their floor values outside the smaller domain.
+This simple extrapolation is adequate for ray tracing since
+the neutrinos are almost entirely free-streaming outside the smaller domain.
+Since the metric data were evolved on this larger domain no extrapolation is
+needed for them.
+The larger domain is represented as a pseudospectral grid
+composed of a sphere with concentric shells extending to $r\approx1400$~km.
 Radial grid spacings are $\Delta r \approx 0.15$~km in the star
@@
 For the purposes of this analysis the pseudospectral representation is adequate.
-
-Figs.~\ref{fig:nsns_rho_merid}--\ref{fig:nsns_temp_equat}
-show slices of density and temperature from the finite difference volume data.
-These data were evolved on a rectilinear grid spanning approximately 400~km in
-both the $x$ and $y$ directions, and 150~km in the $z$ direction in our
-evolution coordinates. We extrapolate the fluid and M1 radiation data from
-this grid to the larger ray tracing grid by setting all variables to their
-floor values. This simple extrapolation is adequate for ray tracing since
-the neutrinos are almost entirely free-streaming outside the smaller domain.
-
-\begin{figure}
-  \includegraphics[width=\columnwidth]{production-colormap-merid-rho}
-  % VTK Data: zosima:/home/mbdeaton/Research/20170602-nsns_vol_data
-  % Plots:    $DIR/production-colormap-{merid,equat}-{rho,temp}.png
-  % w/ DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20170420-nsns_volume_data/02-two_domains_avg_eps/PlotsEtc
-  \caption{A meridional slice of density in the hypermassive neutron
-    star and disk configuration (Sec.~\ref{ssec:test_disk_comparison}).
-    The distorted rectangular boundaries are the boundaries of the grid
-    used in the numerical simulation, which employs a coordinate mapping to
-    concentrate points near the central object.}
-  \label{fig:nsns_rho_merid}
-\end{figure}
-
-\begin{figure}
-  \includegraphics[width=\columnwidth]{production-colormap-equat-rho}
-  % see notes from production-colormap-merid-rho
-  \caption{An equatorial slice of density in the hypermassive neutron
-    star and disk configuration (Sec.~\ref{ssec:test_disk_comparison}).}
-  \label{fig:nsns_rho_equat}
-\end{figure}
-
-\begin{figure}
-  \includegraphics[width=\columnwidth]{production-colormap-merid-temp}
-  % see notes from production-colormap-merid-rho
-  \caption{A meridional slice of temperature in the hypermassive neutron
-    star and disk configuration (Sec.~\ref{ssec:test_disk_comparison}).}
-  \label{fig:nsns_temp_merid}
-\end{figure}
-
-\begin{figure}
-  \includegraphics[width=\columnwidth]{production-colormap-equat-temp}
-  % see notes from production-colormap-merid-rho
-  \caption{An equatorial slice of temperature in the hypermassive neutron
-    star and disk configuration (Sec.~\ref{ssec:test_disk_comparison}).}
-  \label{fig:nsns_temp_equat}
-\end{figure}
 

### 2017-12-21 paper.tex
Correct and expand discussion of Figs. 12-16 for compute-eta data

@@
 In general, scattering increases luminosities for polar observers
+\footnote{TODO: in compute-eta data not true anymore for $\nu_e$}
 ($\theta \sim 0^\circ$) because additional high energy neutrinos
@@
 As in the case of the collapse profile in Sec.~\ref{ssec:test_collapse},
-the dominant effect of \emph{scat} on $\tilde{G}(\cos A)$
+the dominant effect of elastic scattering on $\tilde{G}(\cos A)$
 is to spread the distribution out to larger angles, by generally decreasing
@@
 of neutrinos coming from the disk.
-\footnote{TODO: also examine spectra or average energies}
+\footnote{TODO: with compute-eta data also explain why $G_{\nu_e}>G_{\bar{\nu}_e}$
+  for $\theta=0$ at $\cos A>0.995$.}
+
+For the observer in the equatorial plane,
+Fig.~\ref{fig:nsns_cosA_distrib_G_90deg}, the disk is so optically thick to
+$\nu_e$ that there is very little difference between \emph{scat} and
+\emph{noscat} treatments for incoming angles $\cos A \lesssim 0.95$
+corresponding to the volume inside of $r \lesssim 80$~km.
+Also the stepped temperature gradient in the disk's spiral arms visible in
+Figs.~\ref{fig:nsns_temp_merid} and \ref{fig:nsns_temp_equat} presents as a
+stepped heavy-lepton neutrino number density distribution in
+Fig.~\ref{fig:nsns_cosA_distrib_G_90deg} in the \emph{noscat} treatment,
+since the pair production rate due to $e^{-}e^{+}$ annihilation,
+producing $\nu_x$, is especially sensitive to temperature, going as
+$T^9$ \cite[Sec.~7]{burr2006-neutrino_opacities}.
+This stepped distribution is not visible in $\nu_e$ and $\bar{\nu}_e$ emission,
+since we ignore pair processes for these species due to the dominance of
+absorption and emission processes, obeying a shallower temperature-dependence;
+nor is it visible in the \emph{scat} treatment of $\nu_x$, since scattering
+tends to smear the incoming angle of the emission;
+nor is it visible for any species or treatment in
+Fig.~\ref{fig:nsns_cosA_distrib_G_00deg}, since the integration over the
+azimuthal angle $B$ averages out any spiral structure for the observer
+on the polar axis.
 

### 2017-12-22 paper.tex
Update V_nue, V_nua, V_nunu figs for compute-eta data

@@
 \end{align}
-where $\Delta m$ is the mass difference between neutrino mass eigenstates,
+where $\Delta m^2$ is the mass-squared differences between neutrino mass
+eigenstates,
 $G_{\rm F}$ is the Fermi coupling constant,
@@
   \label{eqn:V_nunu0}
-  V_{\nu\nu} &= \int d\varepsilon
+  V_{\nu\nu,0} &= \int d\varepsilon
   \big(V_{\nu_e,0}(\varepsilon)-V_{\bar{\nu}_e,0}(\varepsilon)\big),
@@
 with $\varrho$ the rest density and $m_{N}$ the nucleon mass.
-The matter potential is always positive;
-and outside the accretion disks formed in neutron star mergers,
+The matter potential $V_e$ is always positive;
+and far outside the accretion disks formed in neutron star mergers,
 in which the disrupted neutron-star matter is rapidly releptonizing,
-the neutrino self-interaction potential is large and negative.
+the total unoscillated neutrino self-interaction potential $V_{\nu\nu,0}$
+is large and negative.
 
@@
 in Sec.~\ref{ssec:test_disk_comparison}.
-We calculate the self-interaction potential at $N_s=5$ points between
-$r\in[30,135]$~km,
-by sampling distribution functions over a grid with extents
-$N_\varepsilon=30$, $N_A=30$, $N_B=10$,
-over all angles and an energy range of $\varepsilon\in(0,100)$~MeV.
-\footnote{TODO: improve resolution and update}
+We calculate the self-interaction potential along a radial coordinate trajectory
+making an angle $\theta=25^\circ$ with the rotation axis,
+at $N_s=7$ positions $r\in\{30,50,82,135,223,368,608\}$~km.
+We sample distribution functions at each of these positions over a grid
+with extents $N_\varepsilon=30$, $N_A=200$, $N_B=30$;
+energies range over $\varepsilon\in(0,100)$~MeV;
+polar angles range over $\cos A \in ((\cos A)_{\rm min}, 1)$, with
+$(\cos A)_{\rm min}\in\{-1,-1,-1,0.488,0.849,0.947,0.981\}$
+for each of the $N_s$ positions;
+and azimuthal angles range over the whole sky $B\in[0,2\pi)$.
 We interpolate the logarithm of the self-interaction potentials
@@
 across the $N_s$ observation points.
+For $r>608$~km we extrapolate the self-interaction potentials using the
+geometric fall-off of $r^{-4}$ applicable to the far-field limit of the
+self-interaction potential \cite{malk2016-mnr_3}.
+
 The matter potential along this test ray is from an analytic wind
@@
 In Figs.~\ref{fig:V_nunu-noscat} and \ref{fig:V_nunu-scat}
-we show the self-interaction potential and its contributions from
-$\nu_e$ and $\bar{\nu}_e$
-for a test neutrino ray moving out along a coordinate radial trajectory
-from the hypermassive neutron star used in Sec.~\ref{ssec:test_disk_comparison},
-along an angle of $\theta=22^\circ$ with respect to the polar axis.
-\footnote{TODO: note coordinate ray}
+we show the total unoscillated self-interaction potential
+and its contributions from $\nu_e$ and $\bar{\nu}_e$ for a test neutrino
+moving out along the radial coordinate trajectory described above.
 Fig.~\ref{fig:V_nunu-noscat} shows these terms for the \emph{noscat} treatment,
 and Fig.~\ref{fig:V_nunu-scat} for the \emph{scat} treatment.
-Obviously including the effects of elastic scattering
-tends to increase the $\bar{\nu}_e$ contribution relative to the $\nu_e$
-contribution, causing the self-interaction potential to stay negative
+Obviously, including the effects of elastic scattering tends to increase the
+$\bar{\nu}_e$ contribution relative to the $\nu_e$ contribution,
+in this case causing the self-interaction potential to be negative
 along the entire trajectory.
+This effect may be predicted from Fig.~\ref{fig:nsns_cosA_distrib_G_00deg},
+which is calculated for a qualitatively similar observer, in the vacated polar
+funnel of the disk:
+in the \emph{scat} treatment $G_{\bar{\nu}_e}$ dominates over $G_{\nu_e}$ at
+all angles $\cos A \lesssim 0.998$;
+whereas in the \emph{noscat} treatment $G_{\bar{\nu}_e}$ only dominates over
+$G_{\nu_e}$ over a range of forward-peaked angles $\cos A \in (0.965,0.998)$.
+An additional factor supporting this trend is that neutrinos from the most
+forward-peaked angles $\cos A \sim 1$ have a suppressed effect on the
+self-interaction potential due to the $(1-\cos A)$ term arising in the integral
+of Eqn.~\ref{eqn:V_nu_alpha}.
 
 \begin{figure}
-  \includegraphics[width=\columnwidth]{20170619-potentials-Vnue_Vnua_Vnu-000Mo-022deg-noscat}
-  % Data:   $DIR/000Mo-022deg-*km/Lev1-e30a60b10-noscat-00
-  % Script: $DIR/nu_potentials.nb
-  % Config: $DIR/RunAB/RunAB-noscat/nu_potentials-config-000Mo-02deg.nb
+  \includegraphics[width=\columnwidth]{20171223-potentials-Ve_Vnue_Vnua_Vnu-000Mo-025deg-noscat}
+  % Plot:   $DIR/RunBA/RunBA-025deg-noscat-production01/potentials-Ve_Vnue_Vnua_Vnu-000Mo-025deg.pdf
+  % Data:   $DIR/000Mo-025deg-*km/Lev2-e30a200b30-noscat-04
+  % Script: $DIR/nu_potentials-production.nb
+  % Config: $DIR/RunBA/RunBA-025deg-noscat-production01/nu_potentials-config-production.nb
   % w/ DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20170412-self_interaction_potentials_M12_M12-LS220-M1/Run03/
-  \caption{Neutrino self-interaction potential along a test trajectory
+  \caption{Neutrino oscillation potentials
+    along a test trajectory
     originating at the surface of the hypermassive neutron star, and
     proceeding outward along a radial coordinate ray with angle
-    $\psi=22^{\circ}$ with respect to the polar axis.
+    $\theta=25^{\circ}$ with respect to the polar axis.
     The trajectory is parameterized by the coordinate length $s$.
+    We plot the vacuum potential due to mass-squared differences (gray bands),
+    the matter potential $V_e$ due to forward scattering on $e^-$ and $e^+$
+    (dark green),
+    and the self-interaction potential $V_{\nu\nu,0}$ due to forward
+    scattering on ambient neutrinos (light green, solid where positive,
+    dashed where negative);
+    additionally we plot the $\nu_e$ and $\bar{\nu}_e$ components composing
+    the self-interaction potential (light blue and dark blue respectively).
     This calculation was done with scattering turned off.
-    Outside of the symmetric point at $s\sim40\,{\rm km}$, the potential is
-    positive, i.e.\ $\nu_e$-dominated.
-    Compare to Fig.~\ref{fig:V_nunu-scat}.
+    The two gray bands at $V\sim10^{-22}$~erg and $V\sim10^{-24}$~erg
+    indicating the two vacuum energy scales for 10--30~MeV neutrinos,
+    set the positions of possible Mikheyev-Smirnov-Wolfenstein (MSW)
+    or nutation resonances \cite{malk2012-mnr_1}.
+    Inside the symmetric point at $s\sim100\,{\rm km}$ the total unoscillated
+    self-interaction potential $V_{\nu\nu,0}$ is positive,
+    i.e.\ $\nu_e$-dominated.
     }
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{20170619-potentials-Vnue_Vnua_Vnu-000Mo-022deg-scat}
-  % Data:   $DIR/000Mo-022deg-*km/Lev1-e30a60b10-scat-00
-  % Script: $DIR/nu_potentials.nb
-  % Config: $DIR/RunAB/RunAB-scat/nu_potentials-config-000Mo-02deg.nb
+  \includegraphics[width=\columnwidth]{20171223-potentials-Ve_Vnue_Vnua_Vnu-000Mo-025deg-scat}
+  % Plot:   $DIR/RunBA/RunBA-025deg-scat-production01/potentials-Ve_Vnue_Vnua_Vnu-000Mo-025deg.pdf
+  % Data:   $DIR/000Mo-025deg-*km/Lev2-e30a200b30-scat-04
+  % Script: $DIR/nu_potentials-production.nb
+  % Config: $DIR/RunBA/RunBA-025deg-scat-production01/nu_potentials-config-production.nb
   % w/ DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20170412-self_interaction_potentials_M12_M12-LS220-M1/Run03/
@@
     but with scattering turned on.
-    Unlike the case with scattering turned off, here the potential is
-    everywhere negative, i.e.\ $\bar{\nu}_e$-dominated.
+    Unlike the case with scattering turned off,
+    here the total unoscillated self-interaction potential $V_{\nu\nu,0}$
+    is everywhere negative, i.e.\ $\bar{\nu}_e$-dominated.
     }
@@
 for the two treatments.
+\footnote{TODO: update this paragraph to new compute-eta data}
 The survival probabilities are the terms of
@@
 
-We see that the matter-neutrino resonance occurs in both treatments,
-but with the very different behaviors of symmetric and asymmetric
-resonances respectively for the \emph{noscat} and \emph{scat} cases,
-as seen in Figs.~\ref{fig:survival-noscat} and \ref{fig:survival-scat}.
-These two classes of matter-neutrino resonance
+For our assumed matter potential along this one test trajectory
+we see that a matter-neutrino resonance only occurs in the \emph{scat} case;
+though with a slightly higher matter potential it could occur in both cases,
+closer to the disk in the \emph{noscat} case than the \emph{scat} case.
+For this particular test trajectory we only see a standard and not a symmetric
+matter neutrino resonance;
+these two classes of matter-neutrino resonance
 are described for example in \cite{malk2016-mnr_3,vaan2016-uncovering_mnr}.
-This dramatic difference in survival probabilities between the \emph{noscat}
-and \emph{scat} cases along this test ray emphasizes the importance of
+The difference we see in the oscillation physics between the \emph{noscat}
+and \emph{scat} cases along this test trajectory highlights the importance of
 scattering physics in the problem of neutrino oscillations in
@@
   % w/ DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20170412-self_interaction_potentials_M12_M12-LS220-M1/Run03/RunAB/probabilityPotential
-  \caption{Neutrino survival probabilties for the \emph{noscat} case.
+  \caption{Neutrino survival probabilities for the \emph{noscat} case.
+    \footnote{TODO: Replace; this is from the old interp-eta data}
     In the lower panel we reproduce the unoscillated self-interaction potential
@@
     but with scattering turned on.
+    \footnote{TODO: Replace; this is from the old interp-eta data}
     In this case the neutrinos undergo a standard MNR transition
@@
 \end{figure}
-
-%\subsection{Neutrino Annihilation Integral}
-%\label{ssec:q_nunu}
-%In this section we calculate $q_{\nu\bar{\nu}}$ at a representative point
-%in the funnel of the NSNS configuration and show that including scattering makes
-%a difference (or not) in our estimate of the power available to drive a jet.
 

### 2017-12-22 paper.tex
Add figure of assumed density along trajectory

@@
 model qualitatively consistent with the densities in the
-simulated volume (i.e.\ inside $r\sim70$~km).
+simulated volume (i.e.\ inside $r\sim70$~km) and asymptoting to the $r^{-3}$
+density field of a spherical steady-state wind.
+\footnote{TODO: confirm whether this is accurate}
+The rest density from this model is plotted in Fig.~\ref{fig:test_traj_density}.
+\footnote{TODO: cite Surman \& McLaughlin ApJ 618, 2005}
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{profile-density-000Mo-025deg}
+  % Plot:    $DIR/RunBA/RunBA-025deg-noscat-production01/profile-density-000Mo-025deg.pdf
+  % Script:  $DIR/nu_potentials-production.nb
+  % Configs: $DIR/nu_potentials-wind_model.nb
+  %          $DIR/RunBA/RunBA-025deg-noscat-production01/nu_potentials-config-production.nb
+  % w/ DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20170412-self_interaction_potentials_M12_M12-LS220-M1/Run03/
+  \caption{Assumed matter density along the neutrino test trajectory
+    used to calculate the matter potential $V_e$.
+    }
+  \label{fig:test_traj_density}
+\end{figure}
 

### 2017-12-24 paper.tex
Add description of wind model

@@
 model qualitatively consistent with the densities in the
-simulated volume (i.e.\ inside $r\sim70$~km) and asymptoting to the $r^{-3}$
-density field of a spherical steady-state wind.
-\footnote{TODO: confirm whether this is accurate}
+simulated volume (i.e.\ inside $r\sim70$~km) and asymptoting to the $r^{-2}$
+density field of a spherical steady-state wind with a constant asymptotic
+velocity.
+The velocity and density fields of a spherical steady-state wind
+obey the continuity equation $\rho(r)v(r)r^2=\rho_1 v_1 r_1^2$,
+with $\rho_1$ and $v_1$ the density and velocity measured at some fiducial
+radius $r_1$.
+For velocity field we choose a phenomenological wind model used in the
+oscillation literature\footnote{TODO: cite Surman \& McLaughlin ApJ 618, 2005}
+\begin{equation}
+  v(r) = v_1 \left(1-\frac{R}{r_1}\right)^{-\beta}
+  \left(1-\frac{R}{r}\right)^{\beta},
+\end{equation}
+with $R$ the wind launch radius and $\beta$ an acceleration parameter.
+This yields the following density field
+\begin{equation}
+  \rho(r)=\rho_1 C\left[\frac{r_1}{R}\right]
+  \left(1-\frac{R}{r}\right)^{-\beta}\left(\frac{R}{r}\right)^2,
+\end{equation}
+with $C[a]=a^{2-\beta}(a-1)^\beta$.
+We use the parameters $R=10$~km, $R_1=50$~km, $\beta=2$,
+and $\rho_1=10^7\,{\rm g}\,{\rm cm}^{-3}$.
+Additionally we impose a density cap of
+$\rho_{\rm max}=10^{14}\,{\rm g}\,{\rm cm}^{-3}$ inside the radius $r_0=5$~km,
+and smoothly interpolate densities between $r_0$ and $r_1$ with a cubic polynomial,
+enforcing $C^0$ and $C^1$ continuity at the transitions.
 The rest density from this model is plotted in Fig.~\ref{fig:test_traj_density}.
-\footnote{TODO: cite Surman \& McLaughlin ApJ 618, 2005}
+To translate this density field to a matter potential, we assume a constant
+electron fraction of $Y_e=0.5$.
 

### 2018-01-01 paper.tex
Add Francois' edits through Sec. II.A

@@
   by incorporating estimates of the background neutrino fields.
+  Background fields provide information about the spectra and intensities of
+  the neutrinos scattered into each ray.
   These background fields may be taken from a low-order moment simulation
@@
   reduces to a standard state-of-the-art ray tracing formulation.
-  It handles radiation in regimes spanning optically thick to optically thin.
+  The method handles radiation in regimes spanning optically thick to optically thin.
   We test the new code, highlight its strengths and weaknesses,
@@
 processes outlined in Tab.~\ref{tab:neutrino_processes}.
-
-We build upon existing ray tracing formulations
+We build upon these existing ray tracing formulations
 by eschewing any assumptions about the spacetime geometry,
@@
 and including elastic scattering in the integration along each geodesic.
-To incorporate scattering processes in our method, we employ estimates of the
-lowest-order moment of the neutrino distribution function
-computed in an M1 transport simulation.
-If moments are not available either from
-an M1 transport evolution or a trustworthy analytical estimate,
-our method reduces to the current state-of-the-art ray tracing method,
-neglecting scattering.
 
@@
 because we want to apply the method as a postprocessing step using
-time snapshots of data computed from general relativistic evolutions.
+time snapshots of data computed from time-dependent general relativistic evolutions.
 The spacetime represented in these snapshots is not analytic (i.e. Kerr).
@@
 estimates of the neutrino fields.
-We also retain the freedom to drop the coupling terms, in which case our method
-becomes similar to existing state-of-the-art ray tracing methods.
+To incorporate scattering processes in our method, we employ estimates of the
+lowest-order moment of the neutrino distribution function
+computed in an M1 transport simulation.
+If moments are not available either from
+an M1 transport evolution or a trustworthy analytical estimate,
+we may drop the coupling terms and our method reduces to the current
+state-of-the-art ray tracing methods neglecting scattering.
 Our method is not a standalone radiation transport scheme,
@@
 the absorption/emission (AE) and elastic scattering (SE) processes
-listed in Tab.~\ref{tab:neutrino_processes}.
+listed in Tab.~\ref{tab:neutrino_processes};
+note that we also include the important thermal pair processes (PP) for
+heavy-lepton neutrinos by incorporating an approximate pair emissivity
+into their absorption/emission coefficients;
+see App.~\ref{sec:source_terms} for details.
 Thus we separate these coefficients:
@@
 
-We may use Eqns.~\ref{eqn:ae_emissivity_feq} and \ref{eqn:se_emissivity}
-to separate the source function thus:
-$\mathscr{S}=\mathscr{S}_{\rm AE}+\mathscr{S}_{\rm SE}$, with
-\begin{align}
-  \label{eqn:source_ae}
-  \mathscr{S}_{\rm AE}
-  &= \frac{\mathscr{K}_{\rm AE}}{\mathscr{K}} f^{\rm eq}, \\
-  \label{eqn:source_se}
-  \mathscr{S}_{\rm SE}
-  &= \frac{\mathscr{K}_{\rm SE}}{\mathscr{K}} \Phi.
-\end{align}
 With these definitions we write a separated form of
@@
 $\varepsilon/\mathscr{K}$.)
+
+%Also we may use Eqns.~\ref{eqn:ae_emissivity_feq} and \ref{eqn:se_emissivity}
+%to separate the source function thus:
+%$\mathscr{S}=\mathscr{S}_{\rm AE}+\mathscr{S}_{\rm SE}$, with
+%\begin{align}
+%  \label{eqn:source_ae}
+%  \mathscr{S}_{\rm AE}
+%  &= \frac{\mathscr{K}_{\rm AE}}{\mathscr{K}} f^{\rm eq}, \\
+%  \label{eqn:source_se}
+%  \mathscr{S}_{\rm SE}
+%  &= \frac{\mathscr{K}_{\rm SE}}{\mathscr{K}} \Phi.
+%\end{align}
 
@@
 
+In fact a spherical steady-state wind model, though providing an adequate
+backdrop for the qualitative study presented in this section,
+is less than ideal for this post-merger configuration,
+most obviously because in the 10~ms since merger,
+ejecta with the greatest velocities around $0.3c$
+will have reached no further than $r_{\rm max} \sim 10^8$~cm.
+Additionally, the true radial profile of the ejecta from this merger
+(which our computational model does not follow)
+will have many more features inside this radius, including shock jumps.
+This model density profile is adequate as a backdrop to our study here, however,
+since we expect the remnant to present similar neutrino emission over a thermal
+timescale of a few tens of milliseconds\footnote{cite Nouri et al 2018}
+while the matter field propagates out to larger radii.
+
 In Figs.~\ref{fig:V_nunu-noscat} and \ref{fig:V_nunu-scat}

### 2018-01-03 paper.tex
Add Francois' edits through Sec. III

@@
 Each trajectory is uniquely labeled by a pair of vectors
-giving some event on the trajectory, $x^\alpha$,
+giving an event on the trajectory, $x^\alpha$,
 and the momentum at that event, $p_\beta$.
-A family of trajectories shares the same $x^\alpha$; for example,
-the family of emission trajectories is ($x^\alpha_e,p_\beta$),
-and the family of observation trajectories is ($x^\alpha_o,p_\beta$).
+To designate a family of intersecting trajectories, we keep constant either
+the emission event $x^\alpha_e$ or
+the observation event $x^\alpha_o$.
 
@@
 \label{ssec:moments}
-\footnote{TODO: code these up}
 We may take angular moments of the distribution function:
@@
 \footnote{TODO: consistent with Eqn.~\ref{eqn:def_direction}?}
-We also make use of the number density and flux defined
+We also make use of the specific number density and specific number flux defined
 \begin{align}
@@
 all defined in App.~\ref{sec:definitions}.
-These fields are stored in spectral representation.
-If they are computed via hydrodynamical simulation, they may be saved to disk
-at specified coordinate times and interpolated
+These fields are computed in a preprocessing step before ray tracing
+and stored in spectral representation.
+If they are computed from a hydrodynamical simulation, they may be saved to disk
+at either one or several specified coordinate times and interpolated
 with spectral interpolation in space,
@@
 (as described in \cite[App.~B]{bohn2016-code}).
-If computed from a quasi-stationary solution to the
-general relativistic hydrodynamics equations,
-a single time is stored to disk, and no time interpolation is needed.
+If computed from a stationary solution to the
+general relativistic hydrodynamics equations no time interpolation is needed.
 In this paper for simplicity and to limit computational memory loads,
-we limit ourselves to quasistationary configurations
-and use single time slices in every test case.
+we use only stationary analytical solutions or
+quasi-stationary configurations evolved in simulation,
+thus using one time slice and no time interpolation in every case.
 
@@
 40 points in energy $\tilde{\varepsilon} \in [0,100]$~MeV.
-Because of the step function in fluid variables at radius $R$,
-we throttle the maximum step size at 1.25~km.
+Because of the discontinuity in fluid variables at radius $R$,
+we limit the time step size to a maximum of $t_{\rm max}=1.25$~km,
+since as the ray approaches the discontinuity in the homogeneous environment
+outside the star, the adaptive time-stepper would otherwise
+increase the step size far beyond the relevant fluid scales.
 
@@
 points the way to future improvements to the \emph{gray} method:
-making better assumptions about the spectra which don't depend upon
-local fluid quantities.
+making better assumptions about the spectra which are less sensitive
+to local fluid quantities.
 
@@
 Because of strong differential rotation and shock heating,
-the remnant temporarily avoids
-collapse to a black hole, even if its mass exceeds the threshold of dynamical
+the remnant may temporarily avoid collapse to a black hole,
+even if its mass exceeds the threshold of dynamical
 instability for a rigidly rotating neutron star \cite{duez2009-review}.
 These objects, called hypermassive neutron stars,
-may avoid collapse for $X\textrm{--}Y$~s,
-\footnote{TODO: cite $X$, $Y$}
+may avoid collapse for thousands of seconds
+\footnote{TODO: cite Rezzolla et Kumar 2014, Ravi et Lasky 2014},
 depending upon a number of physical factors including thermal pressure,
@@
 final inspiral and merger of two identical neutron stars of
-isolated gravitational mass 1.2~$M_{\odot}$. The postmerger configuration
+isolated gravitational mass 1.2~$M_{\odot}$. The configuration
 was simulated using a gray M1 transport scheme for the neutrinos, evolving the
 energy density, number density, and energy flux in addition to the standard
-fluid and metric variables for $\sim10\,{\rm ms}$ following merger
+fluid and metric variables to $\sim11\,{\rm ms}$ following merger
 \cite{fouc2016-m1_evolve_n}.
@@
 and $m$ the index labeling each ray.
-Average energies are given by
-$\langle\tilde{\varepsilon}\rangle(\theta)=H^r(\theta)/K^r(\theta)$.
+Average energies in the coordinate frame are given by
+$\langle\tilde{\varepsilon}\rangle(\theta)=\tilde{H}^r(\theta)/\tilde{K}^r(\theta)$.
 We choose $N_{\tilde{\varepsilon}}=30$, $N_A=150$, and $N_B=20$.
@@
 M1 fluxes are integrated over the outer boundary of the finite-difference grid
-at radii as small as 75~km,
+at radii ranging from 75~km to 200~km,
 whereas the ray tracing fluxes are integrated over a sphere at radius 250~km,
-introducing a time lag between the two measurements of under a millisecond.
-Furthermore uncertainties from the M1 simulation are around
+introducing a time lag between some of the fluxes used in the measurements
+of order one millisecond.
+Additionally, uncertainties for the M1 simulation may be estimated from
+comparisons between M1 methods to be around
 15\% for energy luminosities and

### 2018-01-18 paper.tex
Finish Francois' edits, and clean up previous edits

@@
 velocity.
-The velocity and density fields of a spherical steady-state wind
+The density and velocity fields of a spherical steady-state wind
 obey the continuity equation $\rho(r)v(r)r^2=\rho_1 v_1 r_1^2$,
@@
 For velocity field we choose a phenomenological wind model used in the
-oscillation literature\footnote{TODO: cite Surman \& McLaughlin ApJ 618, 2005}
+oscillation literature \footnote{TODO: cite Surman \& McLaughlin ApJ 618, 2005}
 \begin{equation}
@@
 with $C[a]=a^{2-\beta}(a-1)^\beta$.
-We use the parameters $R=10$~km, $R_1=50$~km, $\beta=2$,
+We use the parameters $R=10$~km, $r_1=50$~km, $\beta=2$,
 and $\rho_1=10^7\,{\rm g}\,{\rm cm}^{-3}$.
@@
 To translate this density field to a matter potential, we assume a constant
-electron fraction of $Y_e=0.5$.
+electron fraction of $Y_e=0.5$, roughly consistent with the composition of the
+matter in the disk's funnel \cite[Fig.~8]{fouc2016-m1_evolve_n}.
 
@@
 since we expect the remnant to present similar neutrino emission over a thermal
-timescale of a few tens of milliseconds\footnote{cite Nouri et al 2018}
+timescale of a few tens of milliseconds \footnote{cite Nouri et al 2018}
 while the matter field propagates out to larger radii.
@@
 $u^\mu=Wn^\mu+v_{\rm E}^\mu$.
-Here $t$ is coordinate time and $\alpha$ is lapse
+Here $t$ is coordinate time and $\alpha$ is the lapse
 in the standard 3+1 decomposition of the metric
@@
 We have also introduced
-the fluid Lorentz factor $W=-n_\mu u^\mu=\alpha u^t$,
+the fluid Lorentz factor $W=\alpha u^t$,
 its Eulerian velocity $v_{\rm E}^\mu=g^\mu_\lambda u^\lambda$
@@
 and the projection tensor orthogonal to the normal observer
-$g^\mu_\lambda=\psi^\mu_\lambda+n^\mu n_\lambda$,
-whose spatial part is the 3-metric, $g_{ij}$.
+$g^\mu_\lambda=\psi^\mu_\lambda+n^\mu n_\lambda$.
 

### 2018-01-18 paper.tex
Address some TODOs and add citations

@@
 $d\Omega \equiv d(\cos a)\,db$.
-\footnote{TODO: consistent with Eqn.~\ref{eqn:def_direction}?}
 We also make use of the specific number density and specific number flux defined
@@
 achieve these values.
-And the lengths of the rays are proportional to the mean free paths:
-$L_{\nu_x}>L_{\bar{\nu}_e}>L_{\nu_e}$;
+And the lengths of the rays are proportional to the mean free paths $L$,
+which obey the hierarchy $L_{\nu_x}>L_{\bar{\nu}_e}>L_{\nu_e}$;
 in the \emph{scat} cases these lengths are less than in the \emph{noscat} cases,
@@
 This standard test is also presented for example in
-\cite{ocon2015-gr1d_with_nu,fouc2015-m1_nsbh,abdi2012-monte_carlo}.
-\footnote{TODO: also lieb2005, mull2010}
+\cite{ocon2015-gr1d_with_nu,fouc2015-m1_nsbh,abdi2012-monte_carlo,
+}.
 
@@
 of the neutrinos for the species $\{\nu_e,\bar{\nu}_e,\nu_x\}$
-have scales $R_\nu\approx\{40,35,20\}$~km;
+have length scales $r_\nu\approx\{40,35,20\}$~km;
 with elastic scattering the scales are
-$R_\nu\approx\{50,45,35\}$~km.
-\footnote{TODO: use a consistent lengthscale symbol throughout paper. $d$? $D$? $r$? $l$?}
+$r_\nu\approx\{50,45,35\}$~km.
 % From by-hand summation over histogram G(cosA_i)*dcosA
@@
 
-Total luminosities and average energies from the different treatments are
-\footnote{TODO: update MC results 02$\rightarrow0$3}
+The total luminosities and average energies
+measured by the different treatments are
 presented in Tab.~\ref{tab:collapse_rt_vs_mc}.
@@
   % to get rt luminosities: tail $RTDATA/Moments-nu?.out
-  % MCDATA=$DIR/ResultsSherwood-02
+  % MCDATA=$DIR/ResultsSherwood-02 (02 data is velocity-independent, but very little diff)
   % RTDATA_NO=$DIR/ResultsBrett-B-03/Test-noscat-01/
@@
 may avoid collapse for thousands of seconds
-\footnote{TODO: cite Rezzolla et Kumar 2014, Ravi et Lasky 2014},
 depending upon a number of physical factors including thermal pressure,
-magnetic fields, and the microphysics of the fluid.
+magnetic fields, and the microphysics of the fluid
+\cite{rezz2015-two_winds,ravi2014-hmns_collapse}.
 
@@
 and we ignore potentially important features of relativistic flavor evolution
-near compact objects.
-\footnote{TODO: cite Yang et Kneller}
-Such a cursory treatment is sufficient, however, for the following qualitative
+near compact objects \cite{yang2017-gr_nu_oscillations}.
+We believe our treatment is sufficient, however, for the following qualitative
 exploration.
@@
 and shown to be qualitatively faithful in those environments,
-\footnote{TODO: cite vlasenko?}
+\footnote{TODO: cite someone}
 a recent study has shown that the single-angle approximation
@@
 For velocity field we choose a phenomenological wind model used in the
-oscillation literature \footnote{TODO: cite Surman \& McLaughlin ApJ 618, 2005}
+oscillation literature \cite{surm2005-nu_and_grb_outflows}
 \begin{equation}
@@
 since we expect the remnant to present similar neutrino emission over a thermal
-timescale of a few tens of milliseconds \footnote{cite Nouri et al 2018}
+timescale of a few tens of milliseconds \footnote{TODO: cite someone}
 while the matter field propagates out to larger radii.
@@
 Neutrino scattering off of electrons is inelastic, changing the
-magnitude and direction of the species' momenta.
+magnitude and direction of the neutrino's momentum.
 \footnote{TODO: cite \cite{keil2003-pinched_spectra} or someone}

### 2018-01-30 paper.tex
Explain single-angle approximation more thoroughly

@@
 The slight upward trend in $\nu_x$ luminosities with $\theta$ may be due to
-asymmetries in the configuration.
+asymmetries in the fluid configuration,
+inhomogeneous coordinate maps used in the hydrodynamics evolution,
+or the fact that equatorial observers are closer to the disk's hot spiral arms
+than are polar observers.
 It is not due to the Doppler shift or relativistic beaming
@@
 the largest radial fluxes of these species.
+
 \begin{figure}
@@
 \footnote{TODO: cite someone}
-a recent study has shown that the single-angle approximation
+we note that a recent study has shown that the single-angle approximation
 may mask certain instabilities in the flavor evolution
@@
 
-Within the single-angle approximation,
-solving for the flavor evolution history along a given test ray
-requires knowledge of the unoscillated neutrino contribution to the Hamiltonian
-along that trajectory.
+The formalism of the single-angle approximation requires knowledge of
+the unoscillated neutrino contribution to the Hamiltonian
+along a given test trajectory.
 This is a function of the unoscillated neutrino self-interaction potential,
@@
 the integral is taken over all directions $\ell'_\gamma$.
+
+Implementing the single-angle approximation, we first use ray tracing to compute
+the unoscillated neutrino self-interaction potentials $V_{\nu_\alpha,0}$
+at several points along a test neutrino trajectory.
+We then integrate the flavor evolution matrix $S$ along this test trajectory,
+interpolating $V_{\nu_\alpha,0}$ to all points sampled by the integration.
+And at each integration step we rescale $V_{\nu_\alpha,0}$ according to the
+mixing specified by $S$.
 

### 2018-01-31 paper.tex
Move table to intro and add warning about time-dependence

@@
 
+\begin{table}%[h]
+  \caption{
+    We analyze neutrino interaction processes in terms of these categories.
+    $\nu$ without a label represents a neutrino or antineutrino of any flavor,
+    $N$ represents a nucleon $n$ or $p$,
+    ${}^ZA$ represents a nucleus with mass number $A$ and charge $Z$, and
+    $\gamma$ represents a high-energy photon.
+    A prime indicates a change in that particle's energy.
+  }
+  \label{tab:neutrino_processes}
+  \begin{tabularx}{\columnwidth}{X X}
+    \hline \hline
+    absorption/emission
+    & $\nu_e + n \leftrightarrow e^- + p$                          \\
+    & $\bar{\nu}_e + p \leftrightarrow e^+ + n$                    \\
+    & $\nu_e + {}^AZ \leftrightarrow e^- + {}^A(Z+1)$              \\
+    & $\bar{\nu}_e + {}^AZ \leftrightarrow e^+ + {}^A(Z-1)$        \\
+    \hline
+    elastic scattering
+    & $\nu + N \leftrightarrow \nu + N$                            \\
+    & $\nu + {}^AZ \leftrightarrow \nu + {}^AZ$                    \\
+    \hline
+    inelastic scattering
+    & $\nu + e^- \leftrightarrow \nu' + e^{-'}$                    \\
+    & $\nu + e^+ \leftrightarrow \nu' + e^{+'}$                    \\
+    \hline
+    thermal pair processes
+    & $\nu + \bar{\nu} \leftrightarrow e^{-} + e^{+}$              \\
+    & $\nu + \bar{\nu} + N + N \leftrightarrow N' + N'$            \\
+    & $\nu + \bar{\nu} \leftrightarrow \gamma$                     \\
+    % & $\nu_\alpha + \bar{\nu}_\alpha \leftrightarrow \nu_\beta + \bar{\nu}_\beta $ \\
+    \hline \hline
+  \end{tabularx}
+\end{table}
+
 %\subsubsection*{In what ways is this method limited?}
@@
 or over a small volume outside a source.
+Furthermore, our method is time-dependent, allowing us to compute radiation
+fields in dynamical systems. But it is formulated as a post-processing step,
+with the ray tracing computated on volume data saved in several steps
+of the fluid evolution.
+Thus for dynamical systems the memory demands can be prohibitively large.
 
@@
 velocities.
-In consequence we treat some of gauge-dependent quantities inconsistently,
+In consequence we treat some of the gauge-dependent quantities inconsistently,
 and we ignore potentially important features of relativistic flavor evolution
@@
 
-\begin{table}%[h]
-  \caption{
-    We analyze neutrino interaction processes in terms of these categories.
-    $\nu$ without a label represents a neutrino or antineutrino of any flavor,
-    $N$ represents a nucleon $n$ or $p$,
-    ${}^ZA$ represents a nucleus with mass number $A$ and charge $Z$, and
-    $\gamma$ represents a high-energy photon.
-    A prime indicates a change in that particle's energy.
-  }
-  \label{tab:neutrino_processes}
-  \begin{tabularx}{\columnwidth}{X X}
-    \hline \hline
-    absorption/emission
-    & $\nu_e + n \leftrightarrow e^- + p$                          \\
-    & $\bar{\nu}_e + p \leftrightarrow e^+ + n$                    \\
-    & $\nu_e + {}^AZ \leftrightarrow e^- + {}^A(Z+1)$              \\
-    & $\bar{\nu}_e + {}^AZ \leftrightarrow e^+ + {}^A(Z-1)$        \\
-    \hline
-    elastic scattering
-    & $\nu + N \leftrightarrow \nu + N$                            \\
-    & $\nu + {}^AZ \leftrightarrow \nu + {}^AZ$                    \\
-    \hline
-    inelastic scattering
-    & $\nu + e^- \leftrightarrow \nu' + e^{-'}$                    \\
-    & $\nu + e^+ \leftrightarrow \nu' + e^{+'}$                    \\
-    \hline
-    thermal pair processes
-    & $\nu + \bar{\nu} \leftrightarrow e^{-} + e^{+}$              \\
-    & $\nu + \bar{\nu} + N + N \leftrightarrow N' + N'$            \\
-    & $\nu + \bar{\nu} \leftrightarrow \gamma$                     \\
-    % & $\nu_\alpha + \bar{\nu}_\alpha \leftrightarrow \nu_\beta + \bar{\nu}_\beta $ \\
-    \hline \hline
-  \end{tabularx}
-\end{table}
-
 \subsection{Absorption and emission via charged current}

### 2018-02-02 paper.tex
Sweep TODO notes and assign remaining ones to coauthors

@@
   deat2013-leakage, pere2016-asl, radi2016-dynamical}
-\footnote{TODO: cite others}
+\footnote{TODO (Evan): cite others}
 capture some of the qualitative effects of neutrinos on the matter, but
 provide extremely limited information about the neutrino field itself.
-\footnote{TODO: discuss mc more clearly}
 Monte Carlo methods like those
@@
 because of their global treatment of the problem.
+\footnote{TODO (Evan): discuss mc more clearly}
 
@@
 with the history of that light-cone truncated at large optical depth.
-\footnote{TODO: state scaling $\mathscr{O}(n)$?}
 It easily extends to high resolution in energy and angle by simply
@@
 This is especially pronounced in the case of heavy-lepton neutrinos.
-\footnote{TODO: cite}
 Inelastic scattering and pair processes can introduce further modifications.
 Any phenomena that involve neutrino-neutrino interactions,
-for example neutrino oscillations and neutrino-antineutrino annihilation,
-\footnote{TODO: cite}
+for example neutrino oscillations \footnote{TODO (Gail): cite an example study}
+and neutrino-antineutrino annihilation \cite{asan2000-nunubar},
 depend sensitively on the angular distribution.
-And spectral changes can strongly affect the nucleosynthesis
-occuring in the ejected and irradiated material.
-\footnote{TODO: justify these statements}
+And spectral changes can strongly affect the nucleosynthesis.
+occuring in the ejected and irradiated material
+\footnote{TODO (Evan, Francois, Gail): cite energy-dependent
+neutrino+nucleosynthesis study}
 The ray tracing method we present in this paper captures the dominant effects
@@
 a treatment which essentially decouples different neutrino momenta and species.
-\footnote{TODO: clarify how momenta are decoupled}
 When solving for $f^{\nu_\sigma}(x^\alpha;p_\beta)$,
@@
 In Sec.~\ref{sec:tests} we present tests of the code.
-In Sec.~\ref{sec:applications} we present an application
-in the dynamical environment following the merger of two neutron stars,
-computing neutrino flavor oscillation along an outgoing ray,
+In Sec.~\ref{sec:applications} we present neutrino fields in
+the dynamical environment following the merger of two neutron stars,
+and compute neutrino flavor oscillation along an outgoing ray,
 including the effects of coherent forward scattering with ambient neutrinos.
@@
 and $\varepsilon=-p_t$. In that case $\cos A$ may be identified
-\footnote{TODO: cite someone that uses $\mu$}
-with the familiar forward direction cosine $\mu$.
+with the familiar forward direction cosine $\mu$
+\citep{shu1991-book_radiation,miha1999-foundations}.
 
@@
 \label{ssec:boltzmann}
-Neutrino radiation obeys the relativistic Boltzmann Equation,
+In the limit of large oscillation lengths (see Sec.~\ref{ssec:V_nunu}),
+neutrino radiation obeys the relativistic Boltzmann Equation,
 which, suppressing the arguments of $f$ for simplicity, is written,
-\footnote{TODO: point out limit of QKEs}
 \begin{equation}
@@
 and $C[f]$ is the source term arising from interactions with the medium.
-By Eqn.~\ref{eqn:geodesic_x},
+As we will show below in Eqn.~\ref{eqn:proper_length},
 the affine parameter has dimension ${\rm length}\,{\rm energy}^{-1}$,
@@
 \end{equation}
+Note that the stimulated absorption coefficient $\mathscr{K}_{\rm AE}$ is
+identical to the coefficient $\kappa^*$
+defined in \cite{hari2010-gr_nunubar_collapsar}.
 See App.~\ref{ssec:sources_ae} for details.
-\footnote{TODO: identify $\mathscr{K}_{\rm AE}$ and $\kappa^*$,
-  in e.g.\ \cite{hari2010-gr_nunubar_collapsar}}
 
@@
 Here ``specific'' refers to the quantity being integrable over neutrino energy.
-\footnote{TODO: relate moments to other frames?}
 Integrals~\ref{eqn:J}--\ref{eqn:Sab} are performed over a solid angle in
@@
 
-We integrate Eqns.~\ref{eqn:num_x}--\ref{eqn:num_tau} with adaptive step sizes,
+We integrate Eqns.~\ref{eqn:num_x}--\ref{eqn:num_fse} with adaptive step sizes,
 using the 3rd order Runge-Kutta algorithm
@@
 After each step is taken, the errors for each of the nine variables of
-Eqns.~\ref{eqn:num_x}--\ref{eqn:num_tau} are compared.
-If the maximum of these errors exceeds some absolute or relative threshold
-\footnote{TODO: specify ScaledAbsRel algorithm}
-the step size is decreased and the step recomputed,
-if it is below some threshold the next step size is increased.
+Eqns.~\ref{eqn:num_x}--\ref{eqn:num_fse} are compared to an absolute and
+a relative threshold.
+If the error in any variable exceeds its threshold
+the step size is decreased and the step recomputed;
+if all errors are below threshold, the next step size is increased.
 % see OdeIntegrators::AdaptiveDense
@@
 % and OdeErrorMeasures::ScaledAbsRel
-\footnote{TODO: specify}
+% For example, the parameters I use in the nsns data:
+%OdeErrorMeasure = ScaledAbsRel
+%     (
+%      ATol = 0.000555555555555556;
+%      RTol = 0.000555555555555556;
+%      ScaledVars = q,p,s0,sn,tau,fa,fs;
+%      Scales = 4,1,4,4,1e-20,1e-20,1e-20;
+In practice, the controlling errors come from the radiation variables
+$\tau$, $f_{\rm AE}$, and $f_{\rm SE}$,
+for which the relative tolerances are set to $6\times10^{-4}$,
+and the absolute tolerances are set to $6\times10^{-24}$.
 
@@
 
-To test our algorithm, we integrate the ray tracing equations
-(Eqns.~\ref{eqn:num_x}--\ref{eqn:num_tau}) for various observers
+To test the algorithm, we integrate the ray tracing equations
+(Eqns.~\ref{eqn:num_x}--\ref{eqn:num_fse}) for various observers
 in the following configurations. This suite of configurations
-may be understood as a hierarchy of increasing physical realism:
+defines a hierarchy of increasing physical realism:
 beginning with a homogeneous medium of effectively
@@
 to those calculated in a Monte Carlo transport simulation.
-Finally, we calculate global measures of the neutrino fields outside of the
-hypermassive neutron star and disk formed in a binary neutron star
-merger simulated by \cite{fouc2016-m1_nsns, fouc2016-m1_evolve_n},
-and compare ray tracing results to those from the M1 transport simulation.
 
@@
 Eqns.~\ref{eqn:num_fae} and \ref{eqn:num_fse}:
-1) the full form  by which the solution is $f_{\rm scat}=f_{\rm AE,scat}+f_{\rm SE,scat}$
-(labeled \emph{scat} since we include elastic scattering), and
-2) the form treating only absorption/emission interactions
-in which we set $\mathscr{K}_{\rm SE}=0$ so that $f_{\rm noscat}=f_{\rm AE,noscat}$
-(labeled \emph{noscat} since we ignore elastic scattering).
-There is no simple relation between these forms
-(i.e.\ $f_{\rm AE,scat} \neq f_{\rm AE,noscat}$)
-because $f_{\rm AE}$ is integrated using the absorption plus scattering optical depth
-in the first form and solely the absorption optical depth in the second form.
+\begin{itemize}
+\item the \emph{scat} form including elastic scattering,
+  for which the solution is
+  $f=f_{\rm AE,scat}+f_{\rm SE,scat}$,
+\item the \emph{noscat} form treating only absorption/emission interactions
+  by setting $\mathscr{K}_{\rm SE}=0$,
+  for which the solution is
+  $f=f_{\rm AE,noscat}$.
+\end{itemize}
+Note that $f_{\rm AE,scat} \neq f_{\rm AE,noscat}$ because $f_{\rm AE}$
+is integrated using the absorption plus scattering optical depth in
+the first form and using only the absorption optical depth in the second form.
+So there is no simple algebraic relation between the \emph{scat} and
+\emph{noscat} distribution functions.
 The differences between the two methods is apparent
-in Fig.~\ref{fig:cumulative_f_homogeneous}.
-Before this work, the \emph{noscat} form of the equations has been the standard for
-\footnote{TODO: explain why; cite others}
-ray tracing, though many authors include the scattering optical depth in
+in Fig.~\ref{fig:cumulative_f_homogeneous} below.
+Before this work, the \emph{noscat} form of the equations was the standard
+for ray tracing, though many authors included the scattering optical depth in
 the integration of Eqn.~\ref{eqn:num_fae}
@@
 neglecting weak magnetism
-\footnote{TODO: anything else?}
+\footnote{TODO (Evan): neglecting anything else?}
 and is stored on a grid covering energy, density, temperature, and composition
@@
   f(t)
-  = - \int_{0}^t \frac{dt'}{p^t}e^{-\tau(0,t')}
+  = - \int_{0}^t \frac{dt'}{\varepsilon}e^{-\tau(0,t')}
   \mathscr{K}(t')f^{\rm eq}(t'),
@@
 and the integration proceeds backwards in time, from $t=0$
-\footnote{TODO: could write more simply because Minkowski}
 to some terminal $t<0$.
@@
 order Dormand-Prince algorithm \cite{pres2007-nr_3rd_ed}. This is because
-\footnote{TODO: discuss with Matt}
+\footnote{TODO (Matt, Brett): look at Brett's tests and make clearer statement}
 our default 3rd order Runge-Kutta algorithm estimates a vanishing error
@@
 and $\tilde{G}(\cos A)$ defined analogously.
-\footnote{TODO: add metric terms for these integrals}
 Results are identical for all \emph{scat} and \emph{noscat} methods.
@@
 Also compare a few geodesics in Kerr to analytic.
-\footnote{TODO: compute some geodesics in Kerr}
+\footnote{TODO (Brett): compute some geodesics in Kerr}
 
@@
 significantly modifies the neutrinos' spectra,
-\footnote{TODO: demonstrate}
 and the extended envelope outside the shock becomes a source of
@@
 This standard test is also presented for example in
-\cite{ocon2015-gr1d_with_nu,fouc2015-m1_nsbh,abdi2012-monte_carlo,
-}.
+\cite{ocon2015-gr1d_with_nu,fouc2015-m1_nsbh,abdi2012-monte_carlo}.
 
@@
 pseudo-spectral grid composed of 11 spherical-shell subdomains
-\footnote{TODO: cite early SpEC paper}
+\cite{kidd2000-spec,szil2009-spec}
 comprising a total of 62 radial grid points spaced approximately
@@
 presented in Tab.~\ref{tab:collapse_rt_vs_mc}.
-\footnote{TODO: give equation for L}
+\footnote{TODO (Brett): give equation for L and avg eps.}
 As expected, $\nu_e$ is least affected by scattering, and $\nu_x$ most.
@@
     the following methods:
-    ray tracing ignoring elastic scattering `RT noscat',
-    ray tracing with the gray scattering treatment `RT scat gr',
-    ray tracing with the spectral scattering treatment `RT scat sp', and
+    ray tracing ignoring elastic scattering `noscat',
+    ray tracing with the gray scattering treatment `scat gray',
+    ray tracing with the spectral scattering treatment `scat spectral', and
     a fiducial Monte Carlo transport evolution `MC'.
@@
   \begin{tabularx}{\columnwidth}{X r r r r}
-    & {\bf RT noscat} & \,\,{\bf RT scat gr} & \,\,{\bf RT scat sp} & \,\,{\bf MC} \\
+    & {\bf noscat} & \,\,{\bf scat gray} & \,\,{\bf scat spectral} & \,\,{\bf MC} \\
     \hline
@@
 where $s$ is the length of scattering envelope passed through by the ray.
-\footnote{TODO: should be $\chi_a^*=\varepsilon\zeta_a$ for $\nu_x$}
+% note, for nux we should use $\chi_a^*=\varepsilon\zeta_a$
 In the envelope the local temperature is low so $f^{\rm eq}/\Phi\ll 1$,
@@
 
-\subsection{Hypermassive Neutron Star Remnant:
-  Testing Everything}
+\section{Applications}
+\label{sec:applications}
+
+In this section we use the ray tracing code to
+first calculate global measures
+of the neutrino fields outside of the hypermassive neutron star and disk
+formed in a binary neutron star merger simulated by
+\cite{fouc2016-m1_nsns, fouc2016-m1_evolve_n},
+and compare ray tracing results to those from the M1 transport simulation.
+Second, we examine neutrino oscillations in this environment,
+using results from ray tracing to include the effect of neutrino-neutrino
+interactions on flavor evolution.
+
+\subsection{Neutrinos from a Hypermassive Neutron Star Remnant}
 \label{ssec:test_disk_comparison}
@@
 $GM/r=4\,{\rm km}/250\,{\rm km}$
-\footnote{TODO: too simplistic}
+\footnote{TODO (Brett, Francois, Matt): improve; this is too simplistic}
 (with $G$ the gravitational constant and $M$ the central mass),
@@
 with $C \equiv 4\pi r^2\, \Delta(\cos \theta)$,
-\footnote{TODO: source of GR error in coordinate $r$}
 $\tilde{K}^r_m \equiv \tilde{K}^r(\theta_m)$, and
@@
 $\langle\tilde{\varepsilon}\rangle=L/R$.
+Note, for simplicity we use the coordinate radius $r$ in this expression even
+though the earlier merger evolution did not necessarily produce
+areal coordinates. The effect of this choice is to artificially scale the
+ray tracing luminosities by some factor we believe to be very close to 1.
+In future work we can correct this error by computing the proper area over
+a coordinate sphere at the observer's location.
 
@@
     (presented in Sec.~\ref{ssec:test_disk_comparison}).
-    The methods are ray tracing using the \emph{noscat} method `RT noscat',
-    ray tracing using the \emph{gray} \emph{scat} method `RT scat',
+    The methods are ray tracing using the \emph{noscat} method `noscat',
+    ray tracing using the \emph{gray} \emph{scat} method `scat',
     and the M1 transport simulation `M1'.
@@
   \begin{tabularx}{\columnwidth}{X r r r}
-    & {\bf RT noscat} & \,\,{\bf RT scat} & \,\,{\bf M1} \\
+    & {\bf noscat} & \,\,{\bf scat} & \,\,{\bf M1} \\
     \hline
@@
 constant in $\theta$, because the disk is optically thin to $\nu_x$,
-and most of the neutrinos come from the hypermassive neutron star,
+and most of the heavy-lepton neutrinos come from the hypermassive neutron star,
 which is roughly spherical.
@@
 the disk is not optically thin to $\nu_e$ and $\bar{\nu}_e$,
-and observers within view of the hot hypermassive neutron star measure
-the largest radial fluxes of these species.
+and observers at small angles within view of the hot hypermassive neutron star
+measure the largest radial fluxes of these species.
+When we turn on scattering, the disk is no longer transparent to $\nu_x$,
+and $\nu_x$ luminosities present the same qualitative $\theta$-dependence
+as that of the other species.
 
@@
 
-When we turn on scattering, the disk is no longer transparent to $\nu_x$,
-and $\nu_x$ luminosities present the same qualitative $\theta$-dependence
-as that of the other species.
-In general, scattering increases luminosities for polar observers
-\footnote{TODO: in compute-eta data not true anymore for $\nu_e$}
-($\theta \sim 0^\circ$) because additional high energy neutrinos
-from the hypermassive neutron star are scattered by the disk back up to
-the observer.
+From Figs.~\ref{fig:nsns_theta_distrib_Kr} and \ref{fig:nsns_theta_distrib_Hr}
+we also see that for observers near the poles ($\theta \sim 0^\circ$)
+scattering generally increases fluxes of $\bar{\nu}_e$ and $\nu_x$
+and decreases fluxes of $\nu_e$.
+When scattering is turned on, all three fluxes experience a similar
+loss of neutrinos from the central star.
+But the $\bar{\nu}_e$ and $\nu_x$ fluxes experience a more dominant
+gain of high energy neutrinos scattered by the disk back to the observer.
+The $\nu_e$ fluxes, however, experience only a minor gain of neutrinos
+from the disk, since $\nu_e$ presents a lower average energy,
+and the scattering cross-section depends strongly on energy.
 For observers in the equatorial plane ($\theta \sim 90^\circ$)
-the effect of scattering is a decrease in energy luminosities,
-since some of the highest energy neutrinos from deep in the interior
-are scattered away from the line of sight by the disk.
+the effect of scattering is a decrease in energy fluxes for all species.
+This is because more matter pollutes the equatorial regions than the
+polar regions, causing the losses from the star to dominate
+over the gains from the disk for all three species.
 
@@
 also obey the same ranking.
-As we showed in the toy models with a homogeneous background scattering field
+As we showed in the toy models with a homogeneous background scattering field,
 the scattering contribution $f_{\rm SE}$ is
@@
 of neutrinos coming from the disk.
-\footnote{TODO: with compute-eta data also explain why $G_{\nu_e}>G_{\bar{\nu}_e}$
+\footnote{TODO (Evan, Brett): also explain why $G_{\nu_e}>G_{\bar{\nu}_e}$
   for $\theta=0$ at $\cos A>0.995$.}
@@
 
-\section{Applications}
-\label{sec:applications}
-
 \subsection{Neutrino Oscillations at High Neutrino Densities}
 \label{ssec:V_nunu}
-In this section we examine the importance of elastic scattering in neutrino
+Here we examine the importance of elastic scattering in neutrino
 flavor oscillation above the hypermassive neutron star--disk configuration
@@
 measured by our observer, and
-\footnote{TODO: mention $\bar{\nu}$ eqn}
-$S$ is the flavor evolution matrix describing a mixed state
+$S$ is the neutrino flavor evolution matrix describing a mixed state
 \begin{equation}
@@
 as in neutron star mergers, the problem must be solved globally.
-\footnote{TODO: be clearer}
 To date, no method has been devised to handle this problem.
@@
 This is the so-called single-angle approximation.
-\footnote{TODO: cite someone}
-\footnote{TODO: say something about its validity}
+\footnote{TODO (Gail, Yonglin): citation?}
 Though widely used in the supernova oscillation literature
 and shown to be qualitatively faithful in those environments,
-\footnote{TODO: cite someone}
+\footnote{TODO (Gail, Yonglin): citation?}
 we note that a recent study has shown that the single-angle approximation
@@
 since we expect the remnant to present similar neutrino emission over a thermal
-timescale of a few tens of milliseconds \footnote{TODO: cite someone}
+timescale of a few tens of milliseconds
+\footnote{TODO (Matt, Francois): reasonable assumption? citation?}
 while the matter field propagates out to larger radii.
@@
 for the two treatments.
-\footnote{TODO: update this paragraph to new compute-eta data}
+\footnote{TODO (Brett, Yonglin): update this paragraph to new compute-eta data}
 The survival probabilities are the terms of
@@
 of a given flavor survives a transition.
-\footnote{TODO: define explicitly}
+\footnote{TODO (Yonglin): define $P_{\nu_\alpha\rightarrow\nu_\alpha}$
+  in terms of $S$}
 Where $P_{\nu_e\rightarrow\nu_e}$ vanishes, for example in
@@
   \caption{Neutrino survival probabilities for the \emph{noscat} case.
-    \footnote{TODO: Replace; this is from the old interp-eta data}
+    \footnote{TODO (Brett, Yonglin): Replace; this is from the old interp-eta data}
     In the lower panel we reproduce the unoscillated self-interaction potential
@@
     but with scattering turned on.
-    \footnote{TODO: Replace; this is from the old interp-eta data}
+    \footnote{TODO (Brett, Yonglin): Replace; this is from the old interp-eta data}
     In this case the neutrinos undergo a standard MNR transition
@@
 Add some conclusions.
-\footnote{TODO: note some improvements needed: implement implicit time-stepping,
+\footnote{TODO (Brett): note some improvements needed: implement implicit time-stepping,
   use finite-difference grid for fluid vars, improve gray spectral assumptions}
@@
 Astrophysics---Center for the Evolution of the Elements (JINA-CEE)
-and the National Science Fondation under grant numbers X. \footnote{TODO: grant nos?}
+and the National Science Fondation under grant numbers X. \footnote{TODO (all): grant nos?}
 We thank Sherwood Richers for providing his neutrino transport code
@@
 By contrast $C^{\nu_x}_{{\rm AE}} = 0$,
-\footnote{TODO: but we use a nonzero $C^{\nu_x}_{{\rm AE}}$}
+\footnote{TODO (Evan, Brett): note that we use a nonzero $C^{\nu_x}_{{\rm AE}}$}
 where $\nu_x$ stands for any of the four heavy-lepton neutrinos and
@@
 (e.g. \cite[Eqn.~A5]{brue1985-core_collapse}):
-\footnote{TODO: explain Fermi-blocking}
 \begin{align}
@@
 with $\chi^{\rm tr}_{s,i}$ the transport scattering opacity,
-\footnote{TODO: drop anisotropic term and defend}
+\footnote{TODO (Evan): can we defend dropping the anisotropic term?}
 \begin{equation}
@@
 \end{equation}
-\footnote{TODO: defend}
 
@@
 magnitude and direction of the neutrino's momentum.
-\footnote{TODO: cite \cite{keil2003-pinched_spectra} or someone}
+\footnote{TODO (Evan): citation? maybe \cite{keil2003-pinched_spectra}?}
 A similar formalism to the above could be used to derive source terms
 for inelastic scattering.
-\footnote{TODO: compare relative amounts of inelastic to elastic scattering}
+\footnote{TODO (Evan): can we compare relative amounts of inelastic to
+  elastic scattering?}
 We save that for future work, pointing out here that inelastic scattering
@@
 We do not include thermal pair processes within the above formalism,
-\footnote{TODO: explain this choice, and cite \cite{keil2003-pinched_spectra} or someone}
+\footnote{TODO (Evan): can we explain this choice better? and add citation?}
 but as mentioned in App.~\ref{ssec:sources_ae}, we do include
@@
 by computing the emissivity of the pair processes in
-\footnote{TODO: describe better}
+\footnote{TODO (Evan): improve description?}
 Tab.~\ref{tab:neutrino_processes}, and applying Kirchoff's law,

### 2018-02-02 paper.tex
Correct some summation equation and add a few for clarity

@@
 We sample the distribution function with ray tracing over a uniform grid
-of 40 points in energy $\varepsilon \in [0,100]$~MeV.
+of 40 points in energy $\varepsilon \in (0,100)$~MeV.
 The domain extends to $\tau_{\rm term}\,L_{\rm mfp,max}\approx10^7\,{\rm km}$,
@@
 As Fig.~\ref{fig:mfps-50km} reveals, integrating the distribution functions
-over the energies $\varepsilon\in[0,100]$~MeV
+over the energies $\varepsilon\in(0,100)$~MeV
 probes the numerical solution over length scales from
@@
 with ray tracing
-over a uniform grid of 40 points in energy $\tilde{\varepsilon}\in[0,100]$
-and 30 points in angle $\cos A\in[-1,1]$, holding fixed $B=\pi$.
+over a uniform grid of 40 points in energy $\tilde{\varepsilon}\in(0,100)$
+and 30 points in angle $\cos A\in(-1,1)$, holding fixed $B=\pi$.
 Results are shown in Fig.~\ref{fig:avg_eps_doppler} for
@@
 \end{equation}
-with the Eulerian energy density per angular bin given by
-\begin{equation}
+with the Eulerian densities per angular bin given by sums over the samples
+\begin{align}
+  \label{eqn:G_per_cosA}
+  \tilde{G}(\cos A) &=
+  \frac{\Delta}{(2\pi)^3} \sum\limits_{m=0}^{N_{\tilde{\varepsilon}}-1}\tilde{\varepsilon}_m^2 f_m(\cos A),\\
   \label{eqn:J_per_cosA}
-  \tilde{J}(\cos A) =
-  \frac{1}{(2\pi)^3} \int_0^{\infty} d\tilde{\varepsilon} \, \tilde{\varepsilon}^3
-  \int_0^{2\pi} dB \, f(\tilde{\varepsilon},\cos A,B),
-\end{equation}
-and $\tilde{G}(\cos A)$ defined analogously.
+  \tilde{J}(\cos A) &=
+  \frac{\Delta}{(2\pi)^3} \sum\limits_{m=0}^{N_{\tilde{\varepsilon}}-1}\tilde{\varepsilon}_m^3 f_m(\cos A),
+\end{align}
+with $\Delta \equiv 2\pi \Delta\tilde{\varepsilon}$,
+$N_{\tilde{\varepsilon}}$ the number of energy seamples,
+$f_m(\cos A) \equiv f(\tilde{\varepsilon}_m,\cos A,B)$,
+and $m$ labeling each energy bin.
 Results are identical for all \emph{scat} and \emph{noscat} methods.
@@
 We sample the distribution function over a uniform grid of
-30 points in angle $\cos A \in [0.734,1]$ and
-40 points in energy $\tilde{\varepsilon} \in [0,100]$~MeV.
+30 points in angle $\cos A \in (0.734,1)$ (holding fixed $B=\pi$) and
+40 points in energy $\tilde{\varepsilon} \in (0,100)$~MeV.
 Because of the discontinuity in fluid variables at radius $R$,
 we limit the time step size to a maximum of $t_{\rm max}=1.25$~km,
-since as the ray approaches the discontinuity in the homogeneous environment
-outside the star, the adaptive time-stepper would otherwise
-increase the step size far beyond the relevant fluid scales.
+so that as the ray approaches the discontinuity in the homogeneous environment
+outside the star, the adaptive time-stepper avoids increasing
+the step size beyond the relevant fluid scales.
 
@@
 We place a stationary observer at $r=500$~km.
+Taking advantage of the spherical symmetry, we sample the distribution function
+$f(\tilde{\varepsilon},\cos A,B)$ with ray tracing
+over a uniform grid of 40 points in energy $\tilde{\varepsilon}\in(0,100)$
+and 80 points in angle $\cos A\in(0.9,1)$, holding fixed $B=\pi$.
+With these samples we compute energy luminosity from the
+radial momentum density $\tilde{H}^r$,
+using the midpoint rule to convert the integral in Eqn.~\ref{eqn:Ha} to a sum:
+\begin{equation}
+  \label{eqn:L_minkowski}
+  L =
+  C \frac{\Delta}{(2\pi)^3} \sum\limits_{m=0}^{N-1} \varepsilon_m^3
+  \cos A_m f_m,
+\end{equation}
+with $C=4\pi r^2$,
+$\Delta \equiv 2\pi\Delta\tilde{\varepsilon}\Delta(\cos A)$,
+$N \equiv N_A N_{\tilde{\varepsilon}}$,
+$N_A$ and $N_{\tilde{\varepsilon}}$ the number of samples in angle and energy,
+and $m$ labeling each ray.
+We also compute average energy as a function of incoming angle
+$\langle\tilde{\varepsilon}\rangle(\cos A)$ using
+Eqns.~\ref{eqn:avg_eps_per_cosA}--\ref{eqn:J_per_cosA},
+and the average energy of all neutrinos measured by our observer using
+\begin{equation}
+  \label{eqn:avg_eps_minkowski}
+  \langle\tilde{\varepsilon}\rangle =
+  \frac{1}{N_A}\sum\limits_{m=0}^{N_A-1}\langle\tilde{\varepsilon}\rangle(\cos A_m).
+\end{equation}
+
 Fig.~\ref{fig:avg_eps_collapse} shows the distribution of average $\nu_x$
@@
 with 12 cells spanning polar angles $\theta\in[0,\pi]$
-and 24 cells spanning azimuthal angles $\phi\in[0,2\pi]$.
+and 24 cells spanning azimuthal angles $\phi\in[0,2\pi)$.
 Though a pseudospectral representation of non-smooth hydrodynamic data
@@
   \label{eqn:Kr_minkowski}
-  \tilde{K}^r &= \Delta
+  \tilde{K}^r(\theta) &= \frac{\Delta}{(2\pi)^3}
   \sum\limits_{m=0}^{N-1}
-  \tilde{\varepsilon}_m^2 \cos A_m f_m,\\
+  \tilde{\varepsilon}_m^2 \cos A_m f_m(\theta),\\
   \label{eqn:Hr_minkowski}
-  \tilde{H}^r &= \Delta
+  \tilde{H}^r(\theta) &= \frac{\Delta}{(2\pi)^3}
   \sum\limits_{m=0}^{N-1}
-  \tilde{\varepsilon}_m^3 \cos A_m f_m,
-\end{align}
-with $f_m \equiv f(\tilde{\varepsilon}_m,\cos A_m, B_m)$,
+  \tilde{\varepsilon}_m^3 \cos A_m f_m(\theta),
+\end{align}
+with $f_m(\theta) \equiv f(\theta;\tilde{\varepsilon}_m,\cos A_m, B_m)$,
 $\Delta \equiv \Delta\tilde{\varepsilon}\, \Delta(\cos A)\, \Delta B$,
@@
   \left(\frac{1}{2}(\tilde{K}^r_0+\tilde{K}^r_{N_\theta-1})
-  +\sum\limits_{m=1}^{N_\theta-2}\tilde{K}^r_m\right),\\
+  +\sum\limits_{n=1}^{N_\theta-2}\tilde{K}^r_n\right),\\
   \label{eqn:luminosity_L}
@@
   \left(\frac{1}{2}(\tilde{H}^r_0+\tilde{H}^r_{N_\theta-1})
-  +\sum\limits_{m=1}^{N_\theta-2}\tilde{H}^r_m\right),
-\end{align}
-with $C \equiv 4\pi r^2\, \Delta(\cos \theta)$,
-$\tilde{K}^r_m \equiv \tilde{K}^r(\theta_m)$, and
-$\tilde{H}^r_m \equiv \tilde{H}^r(\theta_m)$.
+  +\sum\limits_{n=1}^{N_\theta-2}\tilde{H}^r_n\right),
+\end{align}
+with $C \equiv 2\pi r^2\, \Delta(\cos \theta)$,
+$\tilde{K}^r_n \equiv \tilde{K}^r(\theta_n)$,
+$\tilde{H}^r_n \equiv \tilde{H}^r(\theta_n)$,
+and $n$ labeling each observer's position.
 Average energies over the whole sky are then

### 2018-02-20 paper.tex
Add Yonglin's new flavor evolution plot

@@
 as described in \cite{zhu2016-mnr_nsns_remnant}.
-In Figs.~\ref{fig:survival-noscat} and \ref{fig:survival-scat}
+In Fig.~\ref{fig:survival-scat}
 we show the survival probabilities for $\nu_e$ and $\bar{\nu}_e$
-for the two treatments.
-\footnote{TODO (Brett, Yonglin): update this paragraph to new compute-eta data}
+for the \emph{scat} treatment. (There is no flavor transformation
+in the \emph{noscat} treatment.)
 The survival probabilities are the terms of
@@
   in terms of $S$}
-Where $P_{\nu_e\rightarrow\nu_e}$ vanishes, for example in
-Fig.~\ref{fig:survival-noscat}, all the original $\nu_e$ content of the beam
-is oscillated into $\nu_\mu$ and $\nu_\tau$; but the $\nu_e$ density doesn't
-vanish altogether,
-since some $\nu_\mu$ and $\nu_\tau$ oscillates into $\nu_e$.
+Where $P_{\nu_e\rightarrow\nu_e}$ decreases, for example in
+Fig.~\ref{fig:survival-scat} outside of $r\sim400$~km,
+a fraction of the original $\nu_e$ content of the beam oscillates into
+$\nu_\mu$ and $\nu_\tau$, and a fraction of $\nu_\mu$ and $\nu_\tau$
+oscillates into $\nu_e$.
 
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{fig-survival_prob-noscat_nh}
-  % Flavor evolution by Yonglin, see emails 7.19.17-7.20.17
-  % Plot: $DIR/31-900kmNHnoScatterPot.eps
-  % w/ DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20170412-self_interaction_potentials_M12_M12-LS220-M1/Run03/RunAB/probabilityPotential
-  \caption{Neutrino survival probabilities for the \emph{noscat} case.
-    \footnote{TODO (Brett, Yonglin): Replace; this is from the old interp-eta data}
-    In the lower panel we reproduce the unoscillated self-interaction potential
-    $V_{\nu\nu,0}$ from Fig.~\ref{fig:V_nunu-noscat} and a matter potential
-    computed assuming a very simple steady state wind consistent with the
-    densities within the simulation volume inside of $r<70$~km
-    along this trajectory.
-    In this case the neutrinos undergo a symmetric MNR transition
-    inside of $r<50$~km,
-    with the $e$-flavor neutrinos and antineutrinos fully converting to $\nu_x$.
-  }
-  \label{fig:survival-noscat}
-\end{figure}
-
-\begin{figure}
   \includegraphics[width=\columnwidth]{fig-survival_prob-scat_nh}
-  % see notes from fig-survival_prob-noscat_nh
-  % Plot: $DIR/130-800kmNHScatterPot.eps (normal hierarchy)
-  \caption{Same as Fig.~\ref{fig:survival-noscat},
-    but with scattering turned on.
-    \footnote{TODO (Brett, Yonglin): Replace; this is from the old interp-eta data}
+  % Flavor evolution by Yonglin, see emails 11.28.17 and 1.31.18
+  % Plot: $DIR/brett4sfnew300-600kmACC12Step2NE30NH012518m12sc.pdf
+  % w/ DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20170412-self_interaction_potentials_M12_M12-LS220-M1/Run03/RunBA/probabilityPotential-production01
+  \caption{Neutrino survival probabilities for the \emph{scat} case.
+    In the lower panel we reproduce the self-interaction potential $V_{\nu\nu,0}$
+    and matter potential $V_e$ from Fig.~\ref{fig:V_nunu-noscat}.
+    \footnote{TODO (Yonglin): change colors and symbols to be consistent with
+    text and above plots. Of if you can give me the raw data being plotted, I
+    can make the plot.}
     In this case the neutrinos undergo a standard MNR transition
-    outside of $r>200$~km,
-    with the $e$-flavor neutrinos and antineutrinos
-    partially converting to $\nu_x$.
+    beginning around $r\sim400$~km,
+    with the $e$-flavor neutrinos and antineutrinos partially converting to
+    $\nu_x$.
   }

### 2018-02-21 paper.tex
Write Test D: test gr terms

@@
 \label{ssec:test_gravity}
-Unfinished: describe and point readers to Asano-Fukuyama test presented
-in thesis ~\cite[Sec.~4.3.3]{deat2015-thesis}.
-Also compare a few geodesics in Kerr to analytic.
-\footnote{TODO (Brett): compute some geodesics in Kerr}
+To test the general relativistic terms in our formulation, which account for
+the gravitational redshift and geodesic curvature of the neutrinos,
+we sample distribution functions outside an idealized hot compact star,
+and compute a neutrino-antineutrino interaction integral describing the
+energy deposited per time per volume due to the process
+$\nu \bar{\nu} \rightarrow e^{-} e^{+}$.
+We describe this code test in detail in
+\cite[Sec.~4.3]{deat2015-thesis}, and here give a brief summary.
+
+The $\nu \bar{\nu}$-annihilation integral outside a compact star was computed
+semi-analytically in \cite{asan2000-nunubar}. Since then many studies of
+$\nu \bar{\nu}$-annihilation in more realistic configurations have used the
+compact star as a standard code test \cite{birk2007-nunubar,
+  hari2010-gr_nunubar_collapsar,zala2011-nunubar}.
+We compute the power density (energy per time per volume) due to
+$\nu\bar{\nu}$-annihilation measured by a stationary observer above the star
+using:
+\begin{equation}
+  \label{eqn:q_nunubar}
+  q_{\nu\bar{\nu}} = A \int d^3p_{\nu} d^3p_{\bar{\nu}}
+  f_{\nu}(p_{\nu j})f_{\bar{\nu}}(p_{\bar{\nu} k})
+  \frac{p_{\nu t}+p_{\bar{\nu} t}}{p_{\nu}^t p_{\bar{\nu}}^t}
+  (p_{\nu\alpha}p_{\bar{\nu}}^\alpha)^2
+\end{equation}
+where as in \cite{asan2000-nunubar} we account for the energy redshift to
+infinite separation by the energy weighting $p_{\nu t}+p_{\bar{\nu} t}$,
+$A=2 c^3 K G_{\rm F}^2$,
+the Fermi constant is $G_{\rm F}=5.29\times10^{-44}\,{\rm cm}^2\,{\rm MeV}^{-2}$,
+\begin{equation}
+  K \left\{ {e \atop \mu\,\tau} \right\} = 
+  \frac{1}{6\pi}
+  \left(1 \, \left\{ {+ \atop -} \right\}
+  \, 4 \sin^2 \theta_w + 8 \sin^4 \theta_w\right)
+\end{equation}
+and the weak mixing angle is $\sin^2\theta_w=0.23$.
+
+Because Eqn.~\ref{eqn:q_nunubar} has such high dimension,
+a simple unigrid integral solution---sampling $f_\nu$ and $f_{\bar{\nu}}$
+over fixed step sizes in momentum space---is impractical.
+We compute the integral using the adaptive Monte Carlo Vegas technique
+\cite{pres2007-nr_3rd_ed},
+which iteratively samples those regions of momentum space that contribute
+most to the integral. At each iterative stage the algorithm estimates the
+error, and terminates when some error threshold is achieved.
+
+In order to stress-test the gravity-dependence of the code,
+we choose an unphysically compact star configuration with
+radius $R=4.43$~km,
+in a Schwarzschild metric with
+gravitational radius $R_g=2.95$~km.
+To compare to the calculation in \cite{asan2000-nunubar},
+instead of integrating the formal solution for $f_\nu$ and $f_{\bar{\nu}}$
+using Eqns.~\ref{eqn:rendering_fae} and \ref{eqn:rendering_fse},
+we compute only the boundary term using Eqn.~\ref{eqn:rendering_fbdry},
+and neglect the attenuation due to the optical depth,
+This method is equivalent to transporting the neutrino distribution function
+in a state of radiative equilibrium with the matter in the star
+up to the observer assuming no interactions along the trajectory.
+To define the neutrino distribution function in the star,
+we make the star homogeneous, with
+temperature $T=5$~MeV and
+chemical potentials $\eta_{\nu_e}=-\eta_{\bar{\nu}_e}=0.1$,
+and we assume stationary fluid.
+
+% Data from
+% DATA=$RUNDIR//q.dat
+% RUNDIR=$RESEARCHDIR/20150423-thesis_data_and_figures/NeutrinoAnnihilationConvergence/AsanoFukuyamaSphere/Run-01.0/Lev2
+% RESEARCHDIR=$HOME/Documents/ResearchGraveyard/NumericalRelativity/DataAndVis
+% The integrals computed were the following 12:
+% qs={6.87957*^27, 6.83298*^27, 6.76573*^27, 7.02744*^27, 6.88277*^27, 6.89031*^27,
+%     6.84114*^27, 6.85656*^27, 6.90977*^27, 6.82969*^27, 6.83426*^27, 6.93181*^27}
+The power density deposited by this interaction at a coordinate radius $r=7.38$~km
+is computed from formulae in \cite{asan2000-nunubar} as
+\begin{equation}
+  q_{\nu\bar{\nu}}=6.89\times10^{27}\,{\rm erg}\,{\rm cm^3}\,{\rm s}^{-1}.
+\end{equation}
+We computed the integral twelve times at an error threshold of 1\% and measured
+a mean of
+\begin{equation}
+  q_{\nu\bar{\nu}}=6.87\pm0.07\times10^{27}\,{\rm erg}\,{\rm cm^3}\,{\rm s}^{-1},
+\end{equation}
+with the error bars expressing the standard deviation between the twelve
+calculations. Each run computed the integral using $N$ samples of the integrand
+(requiring $2N$ rays, one for each sample of $f_\nu$ and $f_{\bar{\nu}}$),
+with $N$ ranging from 56,000 to 72,000.
+
+The success of this test gives us confidence in the code's ability to handle a
+general spacetime metric, since errors in gravitational redshift would have
+affected samples of $f$ in the integrand (e.g.\ sampling the distribution function
+at the wrong local energy), and errors in geodesic integration would have
+affected the angular size of the star (e.g.\ causing the star to look larger
+or smaller).
 

### 2018-02-21 paper.tex
Add conclusion

@@
 \label{sec:conclusions}
-Add some conclusions.
-\footnote{TODO (Brett): note some improvements needed: implement implicit time-stepping,
-  use finite-difference grid for fluid vars, improve gray spectral assumptions}
+We have introduced a new general relativistic ray tracing method to compute
+neutrino distribution functions around compact objects in dynamical configurations,
+and which incorporates the effects of elastic scattering for the first time
+within a ray tracing framework.
+Elastic scattering of neutrinos into and out of each ray is included in our
+method by using estimates of the background neutrino fields from an
+M1 transport simulation.
+To capture the energy spectrum of the background field, we have described a
+\emph{spectral} method which uses neutrino energy densities over multiple
+energy groups as input, and a
+\emph{gray} method which uses neutrino energy and number densities averaged
+over all energies as input.
+We have also successfully tested the ray tracing code with a comprehensive
+battery of tests.
+
+In our tests (Sec.~\ref{sec:tests}) we have confirmed that elastic scattering
+plays a significant role in redistributing neutrino energy- and
+angle-distributions in common compact-object configurations.
+The largest effects are seen in $\nu_x$ distributions
+and to a lesser extent $\bar{\nu}_e$,
+with the dominant effect a decrease in average energies from the central body,
+and an increase in average energies from the scattering envelope.
+More specifically, in the disk configuration formed by the merger of two neutron
+stars (Sec.~\ref{ssec:test_disk_comparison}), elastic scattering causes
+\begin{enumerate}
+\item a decrease in average energies of neutrinos emerging from the remnant at
+  all angles and for all species,
+\item an increase in $\bar{\nu}_e$ and $\nu_x$ fluxes and a decrease in $\nu_e$
+  fluxes viewed from along the rotation axis, and
+\item a decrease in all species' fluxes viewed from the equatorial plane.
+\end{enumerate}
+Furthermore we find good agreement in overall number and energy luminosities
+and average energies in comparisons with neutrino transport methods,
+e.g.\ Monte Carlo in Sec.~\ref{ssec:test_collapse} and
+M1 transport in Sec.~\ref{ssec:test_disk_comparison}.
+
+We have also employed the ray tracing code to examine neutrino flavor
+oscillations along one toy trajectory exiting the neutrino-dense environment
+of the neutron star post-merger configuration (Sec.~\ref{ssec:V_nunu}).
+Along that trajectory,
+$25^\circ$ from the polar axis, elastic scattering increases the
+$\bar{\nu}_e$ relative to the $\nu_e$ self-interaction potential,
+introducing a standard matter neutrino
+resonance (MNR) where there was none before.
+In a similar configuration, but with matter densities lower
+by a factor of a few, the effect of elastic scattering would have been to push
+the MNR out to larger radii.
+This example demonstrates the importance of the physics of elastic scattering
+in the phenomenon of neutrino flavor oscillation.
+We caution the reader, however, to avoid drawing general conclusions from this
+particular example, since a single astrophysical configuration can present
+dramatically different oscillation resonances along test trajectories emerging
+at different angles \cite{zhu2016-mnr_nsns_remnant},
+and since the MNR is extremely sensitive to a host of parameters.
+
+Finally, we propose the following improvements to the ray tracing code to make
+it a more useful and robust astrophysical simulation tool:
+\begin{enumerate}
+\item Use the finite-difference hydrodynamics simulation grid to represent
+  background fluid and neutrino variables instead of interpolating all
+  input variables to the pseudo-spectral spacetime simulation grid in order
+  to begin ray tracing.
+  We have found that though the interpolation of fluid and neutrino variables
+  to a lower-resolution pseudo-spectral grid saves computational memory, it
+  introduces more costly problems, the foremost being Gibbs-like oscillations
+  at shocks and discontinuities present in fluid fields.
+\item Replace explicit with implicit time stepping in the integrations along
+  each ray. We have found that the stability of the time-integration demands
+  extremely small step sizes of the adaptive time-stepping algorithm, especially
+  for higher-energy rays. Large errors are possible if time-stepping thresholds
+  are not fine-tuned to each new configuration.
+\item Improve the spectral assumptions made in the \emph{gray} method.
+  The test of both \emph{gray} and \emph{spectral} methods against
+  the fiducial Monte Carlo calculation presented in Sec.~\ref{ssec:test_collapse}
+  indicated strong agreement for overall average energies for all species. 
+  However, the average energy of $\bar{\nu}_e$ emerging from the envelope
+  (which contributed only a tiny fraction to the total luminosity)
+  differed between the \emph{gray} treatment and the Monte Carlo by 30\%.
+  Agreement could be improved with better spectral assumptions, for example
+  employing pinched spectra.
+\end{enumerate}
 

### 2018-03-21 paper.tex
Add Matt's responses to TODO notes

@@
 timescale of a few tens of milliseconds
-\footnote{TODO (Matt, Francois): reasonable assumption? citation?}
 while the matter field propagates out to larger radii.

### 2018-03-21 paper.tex
Add Evan's responses to TODO notes through [55]

@@
 \citep{ruff1996-leakage_part1,ross2003-leakage_part2,
-  deat2013-leakage, pere2016-asl, radi2016-dynamical}
-\footnote{TODO (Evan): cite others}
+  deat2013-leakage, gale2013-gr_rad_trans, pere2016-asl, radi2016-dynamical}
 capture some of the qualitative effects of neutrinos on the matter, but
@@
 are an excellent tool but require large computational resources
-because of their global treatment of the problem.
-\footnote{TODO (Evan): discuss mc more clearly}
+because of the need to use a large number of particles to fully and precisely
+sample the high dimensional parameter space
+(seven-dimensional in the general case).
 
@@
 depend sensitively on the angular distribution.
-And spectral changes can strongly affect the nucleosynthesis.
-occuring in the ejected and irradiated material
-\footnote{TODO (Evan, Francois, Gail): cite energy-dependent
-neutrino+nucleosynthesis study}
+And spectral changes in neutrino distributions can strongly affect the nuclear
+processes occuring in the ejected and irradiated material
+\cite{fouc2016-m1_evolve_n}.
 The ray tracing method we present in this paper captures the dominant effects
@@
 For comparision with that test we use the LS180 equation of state
-\cite{latt1991-nuc_eos}
-\footnote{available at \lstinline{stellarcollapse.org}}
+\cite{latt1991-nuc_eos, ocon2010-gr1d}
 in which the equilibrium $\nu_e$ neutrino chemical potential is
@@
 (with $\eta_{\bar{\nu}_e}=-\eta_{\nu_e}$ and $\eta_{\nu_x}=0$).
-The opacity table is computed using \lstinline{NuLib} from cross-sections
-neglecting weak magnetism
-\footnote{TODO (Evan): neglecting anything else?}
-and is stored on a grid covering energy, density, temperature, and composition
+
+The opacity table is computed using \lstinline{NuLib}
+\cite{ocon2015-gr1d_with_nu}
+and is identical to the LS180 opacity table used in the referenced paper.
+The scattering opacity is computed taking into account elastic scattering on
+nucleons, alpha particles, and heavy nuclei.
+The absorption opacities consist of electron neutrino absorption on neutrons
+and heavy nuclei as well as electron antineutrino absorption on protons.
+We use Kirchhoff's law to compute emissivities based on these absorption
+opacities.
+For heavy lepton neutrinos we consider thermal emission processes including
+electron-positron annihilation and nucleon-nucleon Bremsstrahlung.
+The table is stored on a grid covering energy, density, temperature,
+and composition
 ranges spanning $\varepsilon_i\in[1,280.5]$~MeV logarithmically,

### 2018-05-19 paper.tex
Add and cite some oscillation references

@@
 Any phenomena that involve neutrino-neutrino interactions,
-for example neutrino oscillations \footnote{TODO (Gail): cite an example study}
+for example neutrino oscillations
+\cite{duan2010-collective,malk2012-mnr_1,vlas2018-multiangle}
 and neutrino-antineutrino annihilation \cite{asan2000-nunubar},

### 2018-05-20 paper.tex
Add Yonglin's notes and citations from email Apr 26

@@
 processes occuring in the ejected and irradiated material
-\cite{fouc2016-m1_evolve_n}.
+\cite{surm2006-grb_nucsynth,malk2012-mnr_1,caba2012-nu_spectra,fouc2016-m1_evolve_n}.
 The ray tracing method we present in this paper captures the dominant effects
@@
 However, we can solve a similar but tractable problem along a single ray.
-Assume all the neutrino rays intersecting an event along a given test ray
+We assume that all the neutrino rays intersecting an event along a given test ray
 have undergone the same flavor evolution history as that of the test ray,
@@
 This is the so-called single-angle approximation.
-\footnote{TODO (Gail, Yonglin): citation?}
 Though widely used in the supernova oscillation literature
 and shown to be qualitatively faithful in those environments,
-\footnote{TODO (Gail, Yonglin): citation?}
-we note that a recent study has shown that the single-angle approximation
-may mask certain instabilities in the flavor evolution
-\cite{wu2017-fast_neutrino_conversions}.
+we note that recent studies have discovered configurations in which
+the single-angle approximation masks certain instabilities
+in the flavor evolution \cite{wu2017-fast_neutrino_conversions},
+or introduces qualitatively different flavor evolution behavior
+than that of the full multi-angle calculation \cite{vlas2018-multiangle}.
 
@@
 for the \emph{scat} treatment. (There is no flavor transformation
-in the \emph{noscat} treatment.)
-The survival probabilities are the terms of
-the flavor evolution matrix $S$ that describe how much of the original flux
-of a given flavor survives a transition.
-\footnote{TODO (Yonglin): define $P_{\nu_\alpha\rightarrow\nu_\alpha}$
-  in terms of $S$}
-Where $P_{\nu_e\rightarrow\nu_e}$ decreases, for example in
-Fig.~\ref{fig:survival-scat} outside of $r\sim400$~km,
-a fraction of the original $\nu_e$ content of the beam oscillates into
-$\nu_\mu$ and $\nu_\tau$, and a fraction of $\nu_\mu$ and $\nu_\tau$
-oscillates into $\nu_e$.
+in the \emph{noscat} treatment. \footnote{TODO (Brett): ubpdate with new plots})
+The survival probability is the probability that a neutrino, if measured,
+will be found to be in its original flavor state.
+The survival probabilities are computed at each point along the trajectory
+from the flavor evolution matrix,
+$P_{\nu_\alpha \to \nu_\alpha} = |S_{\alpha \alpha}(E)|^2$.
+When $P_{\nu_e \to \nu_e}$ decreases,
+as can be seen for example in Fig.~\ref{fig:survival-scat} for $s>400$~km,
+some of the electron-type neutrinos have oscillated into
+mu- or tau-type neutrinos.
 

### 2018-05-21 paper.tex
Add Yonglin's new flavor evolution figures, and update captions

@@
   \caption{Neutrino oscillation potentials
+    for the \emph{noscat} case
     along a test trajectory
@@
     the self-interaction potential (light blue and dark blue respectively).
-    This calculation was done with scattering turned off.
     The two gray bands at $V\sim10^{-22}$~erg and $V\sim10^{-24}$~erg
@@
   \caption{Same as Fig.~\ref{fig:V_nunu-noscat},
-    but with scattering turned on.
+    but for the \emph{scat} case.
     Unlike the case with scattering turned off,
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{fig-survival_prob-scat_nh}
-  % Flavor evolution by Yonglin, see emails 11.28.17 and 1.31.18
-  % Plot: $DIR/brett4sfnew300-600kmACC12Step2NE30NH012518m12sc.pdf
-  % w/ DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20170412-self_interaction_potentials_M12_M12-LS220-M1/Run03/RunBA/probabilityPotential-production01
-  \caption{Neutrino survival probabilities for the \emph{scat} case.
+  \includegraphics[width=\columnwidth]{fig-survival_prob-noscat_ih}
+  % Flavor evolution by Yonglin, see emails 5.9.18, 5.16.18
+  % Data: $DIR/300-1499kmACC12Step2NE30IH051418m12sc.dat2p
+  % Script: $DIR/plot.gp
+  % w/ DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20170412-self_interaction_potentials_M12_M12-LS220-M1/Run03/RunBA/RunBA-025deg-noscat/flavor_evolution
+  \caption{Neutrino survival probabilities for the \emph{noscat} case.
     In the lower panel we reproduce the self-interaction potential $V_{\nu\nu,0}$
     and matter potential $V_e$ from Fig.~\ref{fig:V_nunu-noscat}.
-    \footnote{TODO (Yonglin): change colors and symbols to be consistent with
-    text and above plots. Of if you can give me the raw data being plotted, I
-    can make the plot.}
+    In this case, the neutrinos undergo collective neutrino oscillation
+    beginning around $r\sim700$~km,
+    with both the electron neutrinos and antineutrinos almost converting to mu
+    or tau neutrinos completely.
+  }
+  \label{fig:survival-noscat}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{fig-survival_prob-scat_ih}
+  % Flavor evolution by Yonglin, see emails 5.9.18, 5.16.18
+  % Data: $DIR/300-1499kmACC12Step2NE30IH051418m12sc.dat2p
+  % Script: $DIR/plot.gp
+  % w/ DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20170412-self_interaction_potentials_M12_M12-LS220-M1/Run03/RunBA/RunBA-025deg-scat/flavor_evolution
+  \caption{Same as Fig.~\ref{fig:survival-noscat},
+    but for the \emph{scat} case.
     In this case the neutrinos undergo a standard MNR transition
     beginning around $r\sim400$~km,
-    with the $e$-flavor neutrinos and antineutrinos partially converting to
-    $\nu_x$.
+    with the electron neutrinos converting to mu or tau neutrinos almost
+    completely,
+    and the electron antineutrinos oscillating back to their original flavor
+    after partially converting to mu and tau.
   }

### 2018-05-21 paper.tex
Add Yonglin's notes/commentary from email May 16
this updates sections IVB and V

@@
 as described in \cite{zhu2016-mnr_nsns_remnant}.
-In Fig.~\ref{fig:survival-scat}
-we show the survival probabilities for $\nu_e$ and $\bar{\nu}_e$
-for the \emph{scat} treatment. (There is no flavor transformation
-in the \emph{noscat} treatment. \footnote{TODO (Brett): ubpdate with new plots})
+We show the survival probabilities for $\nu_e$ and $\bar{\nu}_e$
+for the \emph{noscat} treatment in Fig.~\ref{fig:survival-noscat},
+and for the \emph{scat} treatment in Fig.~\ref{fig:survival-scat}.
 The survival probability is the probability that a neutrino, if measured,
@@
 
-For our assumed matter potential along this one test trajectory
+In Fig.~\ref{fig:survival-noscat}, we report the survival probabilities
+from the calculation with elastic scattering turned off.
+Using neutrino mixing angle $\theta_{12}$ and the inverted hierarchy
+(the normal hierarchy gives qualitatively similar results)
+electron neutrinos and antineutrinos start to oscillate around
+700~km in the form of a collective neutrino oscillation,
+causing both $\nu_e$ and $\bar{\nu}_e$ to convert to heavy lepton neutrinos
+and antineutrinos respectively.
+A similar effect was seen in \cite{fren2017-flavor_bns,tian2017-flavor_bns}.
+In Fig.~\ref{fig:survival-scat}, we report the survival probabilities
+of an otherwise identical calculation, but with elastic scattering turned on.
+In this case, electron neutrinos and antineutrinos start to oscillate around
+400~km in the form of a standard matter neutrino resonance:
+at first both $\nu_e$ and $\bar{\nu}_e$ convert to heavy lepton neutrinos,
+but as the transformation progresses, the $\bar{\nu}_e$ partially
+return to their original flavor.
+
+Note that during the latter part of the matter neutrino resonance
+depicted in Fig.~\ref{fig:survival-scat},
+where the self-interaction potential approaches the vacuum scale,
+the survival probabilities show some small-scale oscillations
+different than the standard matter neutrino resonance
+introduced in \cite{malk2015-mnr_2}.
+This occurs because the self-interaction potential and matter potential
+fall close to the vacuum potential scale.
+
+The collective neutrino oscillation occuring in the \emph{noscat} case
+(Fig.~\ref{fig:survival-noscat})
+not only produces a very different flavor mixture,
+the transformation also starts further from the remnant,
+and it extends much further before it completes.
+For these potentials along this test trajectory,
 we see that a matter-neutrino resonance only occurs in the \emph{scat} case;
@@
 For this particular test trajectory we only see a standard and not a symmetric
-matter neutrino resonance;
-these two classes of matter-neutrino resonance
-are described for example in \cite{malk2016-mnr_3,vaan2016-uncovering_mnr}.
-The difference we see in the oscillation physics between the \emph{noscat}
-and \emph{scat} cases along this test trajectory highlights the importance of
-scattering physics in the problem of neutrino oscillations in
-neutron star mergers.
+matter neutrino resonance \cite{malk2016-mnr_3,vaan2016-uncovering_mnr}.
+As we can see from these calculations, the final outcome of the flavor
+transformation is quite different in the two scenarios.
 
@@
 of the neutron star post-merger configuration (Sec.~\ref{ssec:V_nunu}).
-Along that trajectory,
-$25^\circ$ from the polar axis, elastic scattering increases the
-$\bar{\nu}_e$ relative to the $\nu_e$ self-interaction potential,
-introducing a standard matter neutrino
-resonance (MNR) where there was none before.
-In a similar configuration, but with matter densities lower
-by a factor of a few, the effect of elastic scattering would have been to push
-the MNR out to larger radii.
+The trajectory starts from 300~km, and moves out radially at
+$25^\circ$ from the polar axis.
+Along that trajectory, elastic scattering has the effect of increasing
+the ratio of $\bar{\nu}_e$ relative to $\nu_e$.
+This creates a negative self-interaction potential
+which introduces a complete standard matter neutrino resonance
+(see Fig.~\ref{fig:survival-scat}).
+At about 400~km from the merger core,
+both electron neutrinos and electron antineutrinos begin to transform.
+At about 1200~km
+electron neutrinos have almost completely converted to mu or tau neutrinos,
+while the electron antineutrinos have returned back to their original flavor.
+In an otherwise identical calculation, ignoring elastic scattering
+causes the flavor transformation to be very different (see Fig.~\ref{fig:survival-noscat}).
+
 This example demonstrates the importance of the physics of elastic scattering
 in the phenomenon of neutrino flavor oscillation.
-We caution the reader, however, to avoid drawing general conclusions from this
-particular example, since a single astrophysical configuration can present
-dramatically different oscillation resonances along test trajectories emerging
-at different angles \cite{zhu2016-mnr_nsns_remnant},
-and since the MNR is extremely sensitive to a host of parameters.
+However, we avoid drawing general conclusions from this particular example,
+since a single astrophysical configuration can present dramatically different
+oscillation resonances along test trajectories emerging at different angles
+\cite{zhu2016-mnr_nsns_remnant},
+and since the matter neutrino resonance is extremely sensitive to a host
+of parameters.
 

### 2018-05-26 paper.tex
Add Gail's notes from May 23 email

@@
 This agreement is excellent despite the drastic
-simplifications used in our toy model.
+simplifications used in our model.
 
@@
 also obey the same ranking.
-As we showed in the toy models with a homogeneous background scattering field,
+As we showed in the models with a homogeneous background scattering field,
 the scattering contribution $f_{\rm SE}$ is
@@
     and the electron antineutrinos oscillating back to their original flavor
-    after partially converting to mu and tau.
+    after partially converting to mu and tau antineutrinos.
   }
@@
 We have also employed the ray tracing code to examine neutrino flavor
-oscillations along one toy trajectory exiting the neutrino-dense environment
+oscillations along one sample trajectory exiting the neutrino-dense environment
 of the neutron star post-merger configuration (Sec.~\ref{ssec:V_nunu}).
@@
 This creates a negative self-interaction potential
-which introduces a complete standard matter neutrino resonance
+which introduces a complete standard matter neutrino resonance transition
 (see Fig.~\ref{fig:survival-scat}).
@@
 Astrophysics---Center for the Evolution of the Elements (JINA-CEE)
-and the National Science Fondation under grant numbers X. \footnote{TODO (all): grant nos?}
+and the National Science Foundation under grant numbers X,
+and the U.S.\ Department of Energy, Office of Science, Office of Nuclear Physics,
+under Award Number DE-FG02-02ER41216.
+\footnote{TODO (all): grant nos?}
 We thank Sherwood Richers for providing his neutrino transport code

### 2018-05-26 paper.tex
Add Evan's responses to TODO notes (email Feb 28, 2018) through [88]

@@
 Astrophysics---Center for the Evolution of the Elements (JINA-CEE)
-and the National Science Foundation under grant numbers X,
-and the U.S.\ Department of Energy, Office of Science, Office of Nuclear Physics,
-under Award Number DE-FG02-02ER41216.
-\footnote{TODO (all): grant nos?}
+and the National Science Foundation under grant numbers \lstinline{X} (MBD),
+by NASA through the Hubble Fellowship under grant \lstinline{51344.001-A},
+awarded by the Space Telescope Science Institute,
+which is operated  by the Association of Universities for Research in Astronomy, Inc.,
+for NASA under contract \lstinline{NAS 5-26555} (EO),
+and through the U.S.\ Department of Energy, Office of Science, Office of Nuclear Physics,
+under award number \lstinline{DE-FG02-02ER41216} (GCM).
+\footnote{TODO (Matt, Francois): grant nos?}
 We thank Sherwood Richers for providing his neutrino transport code
@@
 
+We compute our rates using the open source neutrino interaction
+library \lstinline{NuLib} \cite{ocon2015-gr1d_with_nu}.
+We compile a table of sources defined over the four dimensions of
+density, temperature, electron-fraction, and neutrino energy, and
+interpolate quad-linearly to the points sampled along each ray.
+
 Each source term is unique to the neutrino or antineutrino species modeled,
@@
 in Tab.~\ref{tab:neutrino_processes}.
-By contrast $C^{\nu_x}_{{\rm AE}} = 0$,
-\footnote{TODO (Evan, Brett): note that we use a nonzero $C^{\nu_x}_{{\rm AE}}$}
-where $\nu_x$ stands for any of the four heavy-lepton neutrinos and
-antineutrinos.
-
-We compute our rates using the open source neutrino interaction
-library \lstinline{NuLib}~\footnote{\url{http://www.nulib.org}}.
-We compile a table of sources defined over the four dimensions of
-density, temperature, electron-fraction, and neutrino energy, and
-interpolate quad-linearly to the points sampled along each ray.
+By contrast $C^{\nu_x}_{{\rm AE}}$ is formally equal to zero.
+However, in practice, \lstinline{NuLib} implements the thermal pair processes
+via an effective emission/absorption term in order to avoid the need
+to couple energy groups and species (see Sec.~\ref{ssec:sources_pp}).
+This has been shown to work well for core-collapse supernovae
+\cite{ocon2015-gr1d_with_nu}.
 

### 2018-06-14 paper.tex
Make survival prob figures consistent with other figures
Change s->r in IV.B and simplify notation and wording

@@
 the limit where collisional mean free paths are much shorter
-than oscillation lengths, i.e.\ $\varepsilon/\mathscr{K} \ll |H|^{-1}$
-with $H$ the Hamiltonian describing coherent forward scattering interactions.
+than oscillation lengths, i.e.\ $\varepsilon/\mathscr{K} \ll |H|^{-1}$ with
+$H$ the Hamiltonian matrix describing coherent forward scattering interactions.
 In this limit the neutrino density matrix takes the form
@@
 \end{equation}
-In this limit, we may decompose the Hamiltonian into vacuum,
+In this limit, we may decompose the Hamiltonian matrix into vacuum,
 matter, and neutrino contributions:
@@
 \end{equation}
-Explicit formulas for these terms may be found in the oscillation literature,
+Explicit formulas for these matrices may be found in the oscillation literature,
 e.g.\ \cite{duan2009-review},
-but for clarity we give their order-of-magnitude scales here:
+but for clarity we only give their order-of-magnitude scales here:
 \begin{align}
@@
 number densities defined in Eqn.~\ref{eqn:G}.
-For the scale of $H_{\nu\nu}$ we have only included the isotropic
-components of the neutrino fields.
+In this statement of scale for $H_{\nu\nu}$ we only include the isotropic
+components of the neutrino fields; for our calculations below, however, we
+include the full angular distributions found via ray tracing.
 
 In regimes in which $|H_{\nu\nu}|\ll|H_{\rm V},H_e|$,
-the flavor evolution along the ray is soluble, and reveals a rich
-and physically important phenonenology including
+the flavor evolution is locally soluble in a ray-by-ray method,
+and reveals a rich and physically important phenonenology including
 vacuum, solar, atmospheric, and terrestrial oscillations,
@@
 as in neutron star mergers, the problem must be solved globally.
-To date, no method has been devised to handle this problem.
+To date, no method has been devised to handle this problem in systems lacking
+spherical symmetry.
 
@@
 We assume that all the neutrino rays intersecting an event along a given test ray
-have undergone the same flavor evolution history as that of the test ray,
-i.e.\ $S$ is the same for all rays sharing that event.
-This is the so-called single-angle approximation.
-Though widely used in the supernova oscillation literature
-and shown to be qualitatively faithful in those environments,
-we note that recent studies have discovered configurations in which
-the single-angle approximation masks certain instabilities
-in the flavor evolution \cite{wu2017-fast_neutrino_conversions},
-or introduces qualitatively different flavor evolution behavior
-than that of the full multi-angle calculation \cite{vlas2018-multiangle}.
+have undergone the same flavor evolution history as that of the test ray:
+i.e.\ the evolution matrix $S$ is the same for all rays sharing that event.
+This is the so-called single-angle approximation,
+which is widely used in the supernova oscillation literature
+and has been shown to be qualitatively faithful in those environments.
+We note, however, that recent studies have discovered 1) spherically symmetric
+configurations for which a single-angle calculation produces qualitatively
+different flavor evolution behavior than a full multi-angle calculation
+\cite{vlas2018-multiangle},
+and 2) azimuthally-symmetric configurations for which the single-angle
+approximation masks certain instabilities
+in the flavor evolution \cite{wu2017-fast_neutrino_conversions}.
 
 The formalism of the single-angle approximation requires knowledge of
-the unoscillated neutrino contribution to the Hamiltonian
+the unoscillated neutrino contribution to the Hamiltonian matrix
 along a given test trajectory.
@@
 neutrino self-interaction potentials cancel,
-or where $V_e - V_{\nu\nu,0}=0$, with
+or where $V_e + V_{\nu\nu,0}=0$, with
 \begin{align}
@@
 making an angle $\theta=25^\circ$ with the rotation axis,
-at $N_s=7$ positions $r\in\{30,50,82,135,223,368,608\}$~km.
+at $N_r=7$ positions $r\in\{30,50,82,135,223,368,608\}$~km.
 We sample distribution functions at each of these positions over a grid
@@
 $(\cos A)_{\rm min}\in\{-1,-1,-1,0.488,0.849,0.947,0.981\}$
-for each of the $N_s$ positions;
+for each of the $N_r$ positions;
 and azimuthal angles range over the whole sky $B\in[0,2\pi)$.
@@
 (Eqn.~\ref{eqn:V_nu_alpha})
-in path length $\log s$ along the ray,
+in path length $\log r$ along the ray,
 by fitting a 3rd order spline with continuous derivative,
-across the $N_s$ observation points.
+across the $N_r$ observation points.
 For $r>608$~km we extrapolate the self-interaction potentials using the
@@
 will have many more features inside this radius, including shock jumps.
-This model density profile is adequate as a backdrop to our study here, however,
+However, this model density profile is adequate as a backdrop to our study here
 since we expect the remnant to present similar neutrino emission over a thermal
@@
 The survival probabilities are computed at each point along the trajectory
-from the flavor evolution matrix,
-$P_{\nu_\alpha \to \nu_\alpha} = |S_{\alpha \alpha}(E)|^2$.
+from the absolute square of the diagonal terms of the flavor evolution matrix,
+$P_{\nu_\alpha \to \nu_\alpha} = |S_{\alpha \alpha}|^2$.
 When $P_{\nu_e \to \nu_e}$ decreases,
-as can be seen for example in Fig.~\ref{fig:survival-scat} for $s>400$~km,
+as can be seen for example in Fig.~\ref{fig:survival-scat} for $r>400$~km,
 some of the electron-type neutrinos have oscillated into
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{fig-survival_prob-noscat_ih}
+  \resizebox{\columnwidth}{!}{\input{fig-survival_prob-noscat_ih.tex}}
   % Flavor evolution by Yonglin, see emails 5.9.18, 5.16.18
   % Data: $DIR/300-1499kmACC12Step2NE30IH051418m12sc.dat2p
-  % Script: $DIR/plot.gp
+  % Script: $DIR/plot-noscat.gp
   % w/ DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20170412-self_interaction_potentials_M12_M12-LS220-M1/Run03/RunBA/RunBA-025deg-noscat/flavor_evolution
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{fig-survival_prob-scat_ih}
+  \resizebox{\columnwidth}{!}{\input{fig-survival_prob-scat_ih.tex}}
   % Flavor evolution by Yonglin, see emails 5.9.18, 5.16.18
   % Data: $DIR/300-1499kmACC12Step2NE30IH051418m12sc.dat2p
-  % Script: $DIR/plot.gp
+  % Script: $DIR/plot-scat.gp
   % w/ DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20170412-self_interaction_potentials_M12_M12-LS220-M1/Run03/RunBA/RunBA-025deg-scat/flavor_evolution
@@
 \end{equation}
-where $n_{e^-}$, $n_{e^+}$, and $n_b$ are the number densities of electrons
+where $n_{e^-}$, $n_{e^+}$, and $n_b$ are the number densities of electrons,
 positrons, and baryons, and $m_b$ is the average baryon mass.

### 2018-06-14 paper.tex
Add a few acknowledgements, and clean up that language

@@
 \subsubsection*{Acknowledgements}
-This research is supported in part by the Joint Institute for Nuclear
-Astrophysics---Center for the Evolution of the Elements (JINA-CEE)
-and the National Science Foundation under grant numbers \lstinline{X} (MBD),
+This work was supported in part by the National Science Foundation under Grant
+No.\ \lstinline{PHY-1430152} (JINA Center for the Evolution of the Elements)
+[MBD, YLZ, GCM];
 by NASA through the Hubble Fellowship under grant \lstinline{51344.001-A},
 awarded by the Space Telescope Science Institute,
-which is operated  by the Association of Universities for Research in Astronomy, Inc.,
-for NASA under contract \lstinline{NAS 5-26555} (EO),
-and through the U.S.\ Department of Energy, Office of Science, Office of Nuclear Physics,
-under award number \lstinline{DE-FG02-02ER41216} (GCM).
-\footnote{TODO (Matt, Francois): grant nos?}
-We thank Sherwood Richers for providing his neutrino transport code
+which is operated  by the Association of Universities for Research in Astronomy Inc.\
+for NASA under contract \lstinline{NAS 5-26555} [EO];
+and through the U.S.\ Department of Energy, Office of Science,
+Office of Nuclear Physics,
+under award number \lstinline{DE-FG02-02ER41216} [GCM].
+We thank the Spectral Einstein Code collaboration
+\footnote{\url{https://www.black-holes.org/code/SpEC.html}} for a powerful and
+robust numerical relativity code base, and in particular
+Lawrence E.\ Kidder, Daniel A.\ Hemberger, Fran\c{c}ois H\'{e}bert,
+and Fatemeh Hossein Nouri for comments and help throughout.
+We also thank Sherwood Richers for providing his neutrino transport code
 \lstinline{Sedonu} for comparisons,
-Fran\c{c}ois H\'{e}bert for his timely help in finding and
-correcting bugs in the time-interpolation algorithm,
-and Luke Roberts for helpful discussions.
+and Luke Roberts for helpful discussions at the JINA-CEE Frontiers meeting 2017.
 

### 2018-06-15 paper.tex
Add Evan's responses to TODO note [88]

@@
 only charged current processes involving $\nu_e$ and $\bar{\nu}_e$ are allowed.
-For each of these processes, $i$, we may write an emission and absorption
+For each of these processes $i$ in Tab.~\ref{tab:neutrino_processes},
+we may write an emission and absorption
 coefficient as a function of the interaction cross-section
@@
 Neutrino scattering on particles of mass much greater than $\varepsilon$
-(i.e. nucleons) is essentially iso-energetic. Following 
+(i.e. nucleons and nuclei) is essentially iso-energetic. Following 
 \cite[Eqn.~A8]{brue1985-core_collapse} or
@@
 \end{equation}
-Using this expansion, and the moments defined in Eqns.~\ref{eqn:J}
-and \ref{eqn:Ha} we may write Eqn.~\ref{eqn:sources_se_1}:
+Using this definition and the moments defined in Eqns.~\ref{eqn:J}
+and \ref{eqn:Ha}, and writing $\omega'=\ell_\alpha\ell'^\alpha$,
+we expand Eqn.~\ref{eqn:sources_se_1} into four terms.
+The term containing $\oint d\Omega' \ell'^\alpha$ vanishes by construction,
+and the remaining three terms may be written in the form
 \begin{equation}
@@
   C_{{\rm SE},i}
-  = \varepsilon \chi_{s,i} \Phi_i - \varepsilon \chi_{s,i} f,
-\end{equation}
-where we have introduced the scattering opacity $\chi_{s,i}$
-and the background scattering field $\Phi$:
+  = \varepsilon \chi^0_i \Phi_i - \varepsilon \chi^0_i f,
+\end{equation}
+where we have introduced the scattering opacity $\chi^0_i$
+for each of the elastic scattering processes $i$ in
+Tab.~\ref{tab:neutrino_processes} and the background scattering field $\Phi$:
 \begin{align}
   \label{eqn:opacity_s}
-  \chi_{s,i}(\varepsilon)
-  &= \frac{4\pi\varepsilon^3}{(2\pi)^3}R_{{\rm SE},i}^0(\varepsilon), \\
+  \chi^0_i(\varepsilon)
+  &= \frac{4\pi\varepsilon^2}{(2\pi)^3}R_{{\rm SE},i}^0(\varepsilon), \\
   \label{eqn:background_phi}
@@
   \left(J(\varepsilon)+
-  \frac{\chi^{\rm tr}_{s,i}(\varepsilon)}{\chi_{s,i}(\varepsilon)}
+  \frac{\chi^1_i(\varepsilon)}{\chi^0_i(\varepsilon)}
   \ell_\alpha H^\alpha(\varepsilon)\right),
 \end{align}
-with $\chi^{\rm tr}_{s,i}$ the transport scattering opacity,
-\footnote{TODO (Evan): can we defend dropping the anisotropic term?}
+and with $\chi^1_i$ defined
 \begin{equation}
   \label{eqn:opacity_s_tr}
-  \chi^{\rm tr}_{s,i}(\varepsilon)
-  = \frac{4\pi\varepsilon^3}{(2\pi)^3}R_{{\rm SE},i}^1(\varepsilon).
-\end{equation}
-In our treatment we only retain the isotropic contribution to the
-background field:
+  \chi^1_i(\varepsilon)
+  = \frac{4\pi\varepsilon^2}{(2\pi)^3}R_{{\rm SE},i}^1(\varepsilon).
+\end{equation}
+
+Note that $\chi^1_i/\chi^0_i$ in Eqn.~\ref{eqn:background_phi}
+is the degree of non-isotropy in the scattering.
+This term is roughly $-0.1$ for scattering on free neutrons,
+$-0.2$ on free protons, and $1$ on heavy nuclei;
+and at disk temperatures the composition is almost entirely free nucleons.
+Therefore in our treatment, for simplicity,
+we only retain the isotropic contribution to the background field:
 \begin{equation}
@@
 
-Thus computing sums over the opacities
-$\chi_s\equiv\sum_i\chi_{s,i}$ and
-$\chi^{\rm tr}_s\equiv\sum_i\chi^{\rm tr}_{s,i}$,
+Computing sums over the opacities
+$\chi_s\equiv\sum_i\chi^0_i$ and
+$\chi^1_s\equiv\sum_i\chi^1_i$,
 the invariant emissivity and opacity for elastic scattering are

### 2018-06-15 paper.tex
Finish adding Evan's responses to TODO notes (email Feb 28, 2018)

@@
 as can be seen for example in Fig.~\ref{fig:survival-scat} for $r>400$~km,
-some of the electron-type neutrinos have oscillated into
-mu- or tau-type neutrinos.
+some of the $e$ neutrinos have oscillated into $\mu$ or $\tau$ neutrinos.
 
@@
 Eqn.~\ref{eqn:feq}.
-With these facts we can rerrange Eqn.~\ref{eqn:sources_ae_1} to give us
+With these facts we can rearrange Eqn.~\ref{eqn:sources_ae_1} to give us
 Kirchoff's Law:
@@
 Neutrino scattering off of electrons is inelastic, changing the
-magnitude and direction of the neutrino's momentum.
-\footnote{TODO (Evan): citation? maybe \cite{keil2003-pinched_spectra}?}
+magnitude and direction of the neutrino's momentum \cite{brue1985-core_collapse}.
+In a supernova environment we expect inelastic scattering off of electrons
+and nucleons to shift the neutrino spectra to lower energies, most noticeably
+for heavy-lepton neutrinos \cite{thom2003-ccsne_neutrinos}.
 A similar formalism to the above could be used to derive source terms
 for inelastic scattering.
-\footnote{TODO (Evan): can we compare relative amounts of inelastic to
-  elastic scattering?}
 We save that for future work, pointing out here that inelastic scattering
 treated this way is very sensitive to the energy-dependence of
-the background field.
+the background field. In this work we take $C_{\rm SI}=0$.
 
@@
 \label{ssec:sources_pp}
-We do not include thermal pair processes within the above formalism,
-\footnote{TODO (Evan): can we explain this choice better? and add citation?}
-but as mentioned in App.~\ref{ssec:sources_ae}, we do include
-these processes in an effective absorption opacity for $\mu$ and $\tau$
-neutrinos and antineutrinos. \lstinline{NuLib} follows
-\cite{brue1985-core_collapse,burr2006-neutrino_opacities}
-by computing the emissivity of the pair processes in
-\footnote{TODO (Evan): improve description?}
-Tab.~\ref{tab:neutrino_processes}, and applying Kirchoff's law,
-Eqn.~\ref{eqn:kirchoffs_law}, to convert this to an effective absorption opacity,
-and its corresponding emissivity obeying Kirchoff's law.
-We only use this effective opacity for heavy-lepton neutrinos because
-it is subdominant to the true absorption opacity for
-$\nu_e$ and $\bar{\nu}_e$.
+We do not include thermal pair processes within the standard pair-process
+formalism; in other words we take $C_{\rm PP}=0$.
+But as mentioned in App.~\ref{ssec:sources_ae}, we do include
+these processes in an effective emission/absorption opacity for $\mu$
+and $\tau$ neutrinos and antineutrinos.
+This has been shown to work well for core-collapse supernovae
+\cite{ocon2015-gr1d_with_nu}. Within \lstinline{NuLib}, we compute
+the energy-dependent emissivity of the pair processes in
+Tab.~\ref{tab:neutrino_processes}, ignoring final state neutrino blocking.
+We then apply Kirchoff's law, Eqn.~\ref{eqn:kirchoffs_law},
+to convert this to an effective absorption opacity.
+We only use this effective opacity for heavy-lepton neutrinos
+because the charged-current emission/absorption processes for $\nu_e$
+and $\bar{\nu}_e$ dominate in the environments we study: early merger
+remnants and supernovae.
 

### 2018-06-15 paper.tex
Update oscillation potentials plots s->r and cm->km

@@
   \frac{\sqrt{2}G_{\rm F}}{(2\pi)^3} \varepsilon^2
-  \oint d\Omega'(1-\varpi')f^{\nu_\alpha}(\varepsilon,\ell'_\gamma),
+  \oint d\Omega'(1-\omega')f^{\nu_\alpha}(\varepsilon,\ell'_\gamma),
 \end{equation}
@@
 and the cosine of the angle between these given by
-$\varpi'=\psi^{\mu\gamma} \ell_\mu \ell'_\gamma$.
+$\omega'=\psi^{\mu\gamma} \ell_\mu \ell'_\gamma$.
 As in the moment equations (Eqns.~\ref{eqn:J}-\ref{eqn:Ka}),
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{20171223-potentials-Ve_Vnue_Vnua_Vnu-000Mo-025deg-noscat}
+  \includegraphics[width=\columnwidth]{20180615-potentials-Ve_Vnue_Vnua_Vnu-000Mo-025deg-noscat}
   % Plot:   $DIR/RunBA/RunBA-025deg-noscat-production01/potentials-Ve_Vnue_Vnua_Vnu-000Mo-025deg.pdf
@@
     $\theta=25^{\circ}$ with respect to the polar axis.
-    The trajectory is parameterized by the coordinate length $s$.
+    The trajectory is parameterized by the coordinate radius $r$.
     We plot the vacuum potential due to mass-squared differences (gray bands),
@@
     or nutation resonances \cite{malk2012-mnr_1}.
-    Inside the symmetric point at $s\sim100\,{\rm km}$ the total unoscillated
+    Inside the symmetric point at $r\sim100\,{\rm km}$ the total unoscillated
     self-interaction potential $V_{\nu\nu,0}$ is positive,
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{20171223-potentials-Ve_Vnue_Vnua_Vnu-000Mo-025deg-scat}
+  \includegraphics[width=\columnwidth]{20180615-potentials-Ve_Vnue_Vnua_Vnu-000Mo-025deg-scat}
   % Plot:   $DIR/RunBA/RunBA-025deg-scat-production01/potentials-Ve_Vnue_Vnua_Vnu-000Mo-025deg.pdf
@@
 
-In Fig.~\ref{fig:survival-noscat}, we report the survival probabilities
+Fig.~\ref{fig:survival-noscat} shows the survival probabilities
 from the calculation with elastic scattering turned off.
@@
 A similar effect was seen in \cite{fren2017-flavor_bns,tian2017-flavor_bns}.
-In Fig.~\ref{fig:survival-scat}, we report the survival probabilities
+Fig.~\ref{fig:survival-scat} shows the survival probabilities
 of an otherwise identical calculation, but with elastic scattering turned on.

### 2018-06-15 paper.tex
Update survival prob figs to use V_osc instead of H_nunu

@@
 We show the survival probabilities for $\nu_e$ and $\bar{\nu}_e$
-for the \emph{noscat} treatment in Fig.~\ref{fig:survival-noscat},
-and for the \emph{scat} treatment in Fig.~\ref{fig:survival-scat}.
+in Figs.~\ref{fig:survival-noscat} and \ref{fig:survival-scat},
+comparing the \emph{noscat} and \emph{scat} treatments.
+In these figures we also show the evolved self-interaction potential
+$V_{\rm osc}$,
+\begin{equation}
+  \label{eqn:V_osc}
+  V_{\rm osc} \equiv (H_{\nu\nu})_{ee}-{\rm Tr}(H_{\nu\nu})/3,
+\end{equation}
+to show how the neutrino interactions driving the oscillation evolve with flavor.
+Note that $V_{\rm osc}$ is identical to $V_{\nu\nu,0}$ if no flavor
+evolution takes place.
 The survival probability is the probability that a neutrino, if measured,
@@
   \caption{Neutrino survival probabilities for the \emph{noscat} case.
-    In the lower panel we reproduce the self-interaction potential $V_{\nu\nu,0}$
-    and matter potential $V_e$ from Fig.~\ref{fig:V_nunu-noscat}.
+    In the lower panel we show the matter potential $V_e$ from
+    Fig.~\ref{fig:V_nunu-noscat},
+    and the evolved self-interaction potential $V_{\rm osc}$ from
+    Eqn.~\ref{eqn:V_osc}.
     In this case, the neutrinos undergo collective neutrino oscillation
     beginning around $r\sim700$~km,
-    with both the electron neutrinos and antineutrinos almost converting to mu
-    or tau neutrinos completely.
+    with both the electron neutrinos and antineutrinos converting to
+    heavy-lepton neutrinos and antineutrinos almost completely.
   }
@@
     beginning around $r\sim400$~km,
-    with the electron neutrinos converting to mu or tau neutrinos almost
+    with the electron neutrinos converting to heavy-lepton neutrinos almost
     completely,
     and the electron antineutrinos oscillating back to their original flavor
-    after partially converting to mu and tau antineutrinos.
+    after partially converting to heavy-lepton antineutrinos.
   }

### 2018-06-15 paper.tex
Remove last four TODO notes
* 'look at Bretts tests and make clearer statement'
* 'give eqn for L and avg eps'
* 'improve; this is too simplistic'
* 'also explain why G_nue > G_nuebar for theta=0 at cosA>0.995'

Ignore these for the time being, and come back to them if the referee
is concerned.

@@
 order Dormand-Prince algorithm \cite{pres2007-nr_3rd_ed}. This is because
-\footnote{TODO (Matt, Brett): look at Brett's tests and make clearer statement}
 our default 3rd order Runge-Kutta algorithm estimates a vanishing error
@@
 presented in Tab.~\ref{tab:collapse_rt_vs_mc}.
-\footnote{TODO (Brett): give equation for L and avg eps.}
 As expected, $\nu_e$ is least affected by scattering, and $\nu_x$ most.
@@
 $GM/r=4\,{\rm km}/250\,{\rm km}$
-\footnote{TODO (Brett, Francois, Matt): improve; this is too simplistic}
 (with $G$ the gravitational constant and $M$ the central mass),
@@
 and in the equatorial plane $\theta=90^\circ$,
-defined analagously to Eqn.~\ref{eqn:J_per_cosA}.
+defined analogously to Eqn.~\ref{eqn:J_per_cosA}.
 As in the case of the collapse profile in Sec.~\ref{ssec:test_collapse},
@@
 of neutrinos coming from the disk.
-\footnote{TODO (Evan, Brett): also explain why $G_{\nu_e}>G_{\bar{\nu}_e}$
-  for $\theta=0$ at $\cos A>0.995$.}
 
@@
   \caption{Distribution of number density $\tilde{G}(\cos A)$ over incoming angle,
-    defined analagously to Eqn.~\ref{eqn:J_per_cosA}.
+    defined analogously to Eqn.~\ref{eqn:J_per_cosA}.
     The integral of this quantity over $d\cos A$ gives the total number density

### 2018-06-15 paper.tex
Update arxiv-only references to real journals, and update intro language

@@
 %\subsubsection*{Why ray tracing?}
-Angular and spectral neutrino distributions in neutron star merger simulations may
-be available soon, for example with the coupled Monte-Carlo-M1 scheme presented in
-\citep{fouc2017-m1_with_mc}.
+Angular and spectral neutrino distributions in neutron star merger simulations
+have just recently become available with a coupled Monte-Carlo-M1 scheme
+\citep{fouc2017-m1_with_mc,fouc2018-mc_for_errors}.
 In this work, however, we present a ray tracing method to compute neutrino
-distribution functions from the existing state-of-the-art general relativistic M1
-transport hydrodynamics simulations.
+distribution functions from the more widely available state-of-the-art
+general relativistic M1 transport hydrodynamics simulations.
 We choose ray tracing because it is conceptually simple,

### 2018-06-15 paper.tex
Update affiliations for Francois and Evan
Correct things caught on the final read-through

@@
 \author{Evan O'Connor}
+\affiliation{Department of Astronomy and Oskar Klein Centre,
+  Stockholm University, Alba Nova, SE-10691, Sweden}
 \affiliation{Department of Physics,
@@
 \author{Francois Foucart}
+\affiliation{Department of Physics,
+  University of New Hampshire, Durham, New Hampshire 03824, USA}
 \affiliation{Lawrence Berkeley National Laboratory,
@@
   and to demonstrate a neutrino flavor oscillation calculation.
+  In that environment, we find qualitatively different fluxes, spectra,
+  and oscillation behaviors when elastic scattering is included.
 \end{abstract}
@@
 which generally lacks any spatial symmetries, so that fully general
-seven-dimensional solutions to the Boltzmann Equation are not feasible.
+solutions to the Boltzmann Equation are not feasible.
 Leakage approximations
@@
 Greek tensor indices ($\alpha, \beta, ...$) range over all four coordinates,
-whereas Latin indices ($i, j, ...$) range over the spatial coordinates 1--3.
+whereas Latin indices ($i, j, ...$) range over the spatial coordinates 1--3,
+or over a more general set, e.g. the set of all elastic scattering interactions.
 We use naturalized units in which $\{\hbar,c,k_{\rm B}\}=1$.
@@
 we see the common features of a red-shifted spectrum for receding fluid
-($\cos A\ll0$),
+($\cos A \sim -1$),
 a blue-shifted spectrum for approaching fluid
-($\cos A\gg0$),
+($\cos A \sim 1$),
 and a slightly red-shifted spectrum for fluid moving transverse to the
@@
 $\nu_e$ that there is very little difference between \emph{scat} and
-\emph{noscat} treatments for incoming angles $\cos A \lesssim 0.95$
+\emph{noscat} treatments for incoming angles $\cos A \gtrsim 0.95$
 corresponding to the volume inside of $r \lesssim 80$~km.
@@
 At about 1200~km
-electron neutrinos have almost completely converted to mu or tau neutrinos,
-while the electron antineutrinos have returned back to their original flavor.
+$e$ neutrinos have almost completely converted to $\mu$ or $\tau$ neutrinos,
+while the $e$ antineutrinos have returned back to their original flavor.
 In an otherwise identical calculation, ignoring elastic scattering
@@
 under award number \lstinline{DE-FG02-02ER41216} [GCM].
+We thank James P.\ Kneller his original flavor evolution code base.
 We thank the Spectral Einstein Code collaboration

### 2018-06-18 paper.tex
Add Francois's grant numbers

@@
 for NASA under contract \lstinline{NAS 5-26555} [EO];
+by NASA through the Einstein Postdoctoral Fellowship grant \lstinline{PF4-150122}
+awarded by the Chandra X-ray Center,
+which is operated by the Smithsonian Astrophysical Observatory for NASA
+under contract \lstinline{NAS8-03060},
+and through grant \lstinline{80NSSC18K0565} [FF];
 and through the U.S.\ Department of Energy, Office of Science,

### 2018-06-20 paper.tex
Add Evan's corrections to final draft (email 6.18.18)

@@
     & $\nu_e + {}^AZ \leftrightarrow e^- + {}^A(Z+1)$              \\
-    & $\bar{\nu}_e + {}^AZ \leftrightarrow e^+ + {}^A(Z-1)$        \\
     \hline
@@
 (normalization and orthogonality to the observer's velocity)
-it has two degrees of freedom; we make this explicit by defining its
+it has only two remaining degrees of freedom; we make this explicit by defining its
 spatial cartesian components with respect to spherical polar angles
@@
   \bar{\nu}_e + p      &\leftrightarrow e^+ + n,\nonumber\\
-  \nu_e + {}^AZ        &\leftrightarrow e^- + {}^A(Z+1),\nonumber\\
-  \bar{\nu}_e + {}^AZ  &\leftrightarrow e^+ + {}^A(Z-1),\nonumber
+  \nu_e + {}^AZ        &\leftrightarrow e^- + {}^A(Z+1),\nonumber
 \end{align}
@@
     below energies at which weak magnetism plays a role \cite{horo2002-weak_mag}.
-    In this data they are exactly identical
+    In these data they are exactly identical
     because we have turned off weak magnetism in our opacity calculations
@@
 with $u^\alpha \rightarrow W(1,0,0,v)$ and $W=(1-v^2)^{-1/2}$,
-where $W$ is the Lorentz factor defined in App.~\ref{sec:definitions}.
+where $W$ is the relativistic Lorentz factor.
 All other thermodynamic variables and background fields are unchanged
@@
 with $\Delta \equiv 2\pi \Delta\tilde{\varepsilon}$,
-$N_{\tilde{\varepsilon}}$ the number of energy seamples,
+$N_{\tilde{\varepsilon}}$ the number of energy samples,
 $f_m(\cos A) \equiv f(\tilde{\varepsilon}_m,\cos A,B)$,
@@
 The 1D matter profile and M1 transport evolution are computed using
-open source supernova evolution code \lstinline{GR1D}
+the open source supernova evolution code \lstinline{GR1D}
 \footnote{\url{http://www.gr1dcode.org}}
@@
 Monte Carlo transport for a fiducial comparison)
-we turn off inelastic scattering in the M1 and Monte Carlo codes,
-since we have not incorporated that physics into this ray tracing code,
-and we turn off general relativistic effects in the ray tracing code,
-since the Monte Carlo code assumes Minkowski spacetime.
+we turn off inelastic scattering where it is included (in the Monte Carlo code),
+and we turn off general relativistic effects where they are included
+(in the ray tracing and M1 codes).
 
@@
 we make the simplifying assumption of Minkowski spacetime.
-This assumption introduces errors in our moments of order
+The errors in our moments introduced by this assumption
+may be estimated to be of order
 $GM/r=4\,{\rm km}/250\,{\rm km}$
@@
 over $\cos\theta$.
-Using the trapezoid rule the number and energy luminosities become sums:
+Using the trapezoid rule, the number and energy luminosities become sums:
 \begin{align}
@@
 In future work we can correct this error by computing the proper area over
-a coordinate sphere at the observer's location.
+coordinate spheres at the observers' locations.
 
@@
 Even if M1 and ray tracing methods both provide faithful measurements
-of all-sky luminosities, we do not expect perfect agreement since the two
-treatments differ fundamentally. In addition to differences in the physics
-such as the treatment of inelastic scattering,
-M1 fluxes are integrated over the outer boundary of the finite-difference grid
+of all-sky luminosities, we expect some disagreement since the two
+treatments differ fundamentally.
+In addition to differences in transport methodologies,
+the M1 fluxes are integrated over the outer boundary of the finite-difference grid
 at radii ranging from 75~km to 200~km,
@@
 but with scattering turned on agreement is within 10\%.
-And as the ray tracing and M1 measurements both show,
+As the ray tracing and M1 measurements both show,
 $R_{\bar{\nu}_e}$ dominates over $R_{\nu_e}$ significantly,
@@
 and \ref{fig:nsns_cosA_distrib_G_90deg}, we show the distributions of neutrino
-number density over incoming angle
+number density over incoming polar angle $\cos A$
 for the observer on the polar axis $\theta=0^\circ$,
-and in the equatorial plane $\theta=90^\circ$,
-defined analogously to Eqn.~\ref{eqn:J_per_cosA}.
+and in the equatorial plane $\theta=90^\circ$.
+This number density is defined
+\begin{align}
+  \label{eqn:G_per_cosA_2}
+  \tilde{G}(\cos A) &=
+  \frac{\Delta}{(2\pi)^3} \sum\limits_{m=0}^{N-1}\tilde{\varepsilon}_m^2 f_m(\cos A),
+\end{align}
+with $\Delta \equiv \Delta\tilde{\varepsilon} \Delta B$,
+$N \equiv N_{\tilde{\varepsilon}} N_B$,
+$f_m(\cos A) \equiv f(\tilde{\varepsilon}_m,\cos A,B_m)$,
+and $m$ the index labeling each ray at polar angle $\cos A$.
+The integral of this quantity over $d\cos A$ gives
+the total number density $\tilde{G}$, according to Eqn.~\ref{eqn:G}.
 As in the case of the collapse profile in Sec.~\ref{ssec:test_collapse},
@@
 Fig.~\ref{fig:nsns_cosA_distrib_G_90deg} in the \emph{noscat} treatment,
-since the pair production rate due to $e^{-}e^{+}$ annihilation,
+since the energy emission due to $e^{-}e^{+}$ annihilation,
 producing $\nu_x$, is especially sensitive to temperature, going as
@@
   \caption{Distribution of number density $\tilde{G}(\cos A)$ over incoming angle,
-    defined analogously to Eqn.~\ref{eqn:J_per_cosA}.
+    defined in Eqn.~\ref{eqn:G_per_cosA_2}.
     The integral of this quantity over $d\cos A$ gives the total number density
@@
 
-In fact a spherical steady-state wind model, though providing an adequate
+In fact, a spherical steady-state wind model, though providing an adequate
 backdrop for the qualitative study presented in this section,
@@
 and to a lesser extent $\bar{\nu}_e$,
-with the dominant effect a decrease in average energies from the central body,
+with the dominant effect being a decrease in average energies from the central body,
 and an increase in average energies from the scattering envelope.

### 2018-06-25 paper.tex
Add Matt's grant number

@@
 and through grant \lstinline{80NSSC18K0565} [FF];
+through National Science Foundation Grant \lstinline{PHY-1402916} [MDD];
 and through the U.S.\ Department of Energy, Office of Science,

### 2018-10-07 paper.tex
fix typo
ref response 1: add reference to appendix
ref response 3: add pseudospectral reference

@@
   \frac{df_{\rm SE}}{dt} &=
-  \frac{1}{p^t}e^{-\tau}\mathscr{K_{\rm SE}} \,\Phi.
-\end{align}
+  \frac{1}{p^t}e^{-\tau}\mathscr{K_{\rm SE}} \,\Phi,
+\end{align}
+where the spacetime metric components $\alpha$, $\beta^i$, and $g^{ij}$
+are defined in App.~\ref{sec:definitions}.
 
@@
 These fields are computed in a preprocessing step before ray tracing
-and stored in spectral representation.
+and stored in pseudospectral representation \cite{duez2008-first_spec_bhns},
+an efficient scheme for representing a continuous field with a finite set of data points.
 If they are computed from a hydrodynamical simulation, they may be saved to disk
@@
 The matter profile and background field are stored on a spherical
-pseudo-spectral grid composed of 11 spherical-shell subdomains
+pseudospectral grid composed of 11 spherical-shell subdomains
 \cite{kidd2000-spec,szil2009-spec}
@@
   background fluid and neutrino variables instead of interpolating all
-  input variables to the pseudo-spectral spacetime simulation grid in order
+  input variables to the pseudospectral spacetime simulation grid in order
   to begin ray tracing.
   We have found that though the interpolation of fluid and neutrino variables
-  to a lower-resolution pseudo-spectral grid saves computational memory, it
+  to a lower-resolution pseudospectral grid saves computational memory, it
   introduces more costly problems, the foremost being Gibbs-like oscillations
@@
 under award number \lstinline{DE-FG02-02ER41216} [GCM].
-We thank James P.\ Kneller his original flavor evolution code base.
+We thank James P.\ Kneller for his original flavor evolution code base.
 We thank the Spectral Einstein Code collaboration

### 2018-10-07 paper.tex
ref response 4: clarify NuLib and LS180 reference

@@
 
-The opacity table is computed using \lstinline{NuLib}
-\cite{ocon2015-gr1d_with_nu}
-and is identical to the LS180 opacity table used in the referenced paper.
+The opacity table is computed using \lstinline{NuLib},
+an open-source neutrino interaction library,
+and is identical to the LS180 opacity table used in the paper introducing
+that tool \cite{ocon2015-gr1d_with_nu}.
 The scattering opacity is computed taking into account elastic scattering on

### 2018-10-07 paper.tex
ref response 5: correct 200% -> 100%

@@
 For example, with scattering turned off the ray tracing and M1 measurements
-of $L_{\nu_x}$ disagree by 200\%;
+of $L_{\nu_x}$ disagree by 100\%;
 but with scattering turned on agreement is within 10\%.

### 2018-10-07 paper.tex
ref response 6: correct 300km -> 30km
ref response 7: use a for pinching param

@@
   \Phi^{\rm pi}(\varepsilon) \propto
-  \varepsilon^{\alpha-2}
-  \exp\left(-(\alpha+1)\frac{\varepsilon}{\langle\varepsilon\rangle}\right),
+  \varepsilon^{a-2}
+  \exp\left(-(a+1)\frac{\varepsilon}{\langle\varepsilon\rangle}\right),
 \end{equation}
@@
 for the species
-$\{\nu_e,\bar{\nu}_e,\nu_x\}$ are $\alpha\approx\{3.6,5.1,2.3\}$.
+$\{\nu_e,\bar{\nu}_e,\nu_x\}$ are $a\approx\{3.6,5.1,2.3\}$.
 Energy moments of pinched spectra like those in Eqn.~\ref{eqn:avg_eps_sc}
@@
   \langle\tilde{\varepsilon}\rangle^{\rm sc}
-  &\approx \frac{\alpha+3}{\alpha+1} \langle\varepsilon\rangle, \\
+  &\approx \frac{a+3}{a+1} \langle\varepsilon\rangle, \\
   %\label{eqn:pinched_avg_eps_sc_2}
@@
 of the neutron star post-merger configuration (Sec.~\ref{ssec:V_nunu}).
-The trajectory starts from 300~km, and moves out radially at
+The trajectory starts from 30~km, and moves out radially at
 $25^\circ$ from the polar axis.

### 2018-10-08 paper.tex
ref response 2: add error estimate equation

@@
 We integrate Eqns.~\ref{eqn:num_x}--\ref{eqn:num_fse} with adaptive step sizes,
-using the 3rd order Runge-Kutta algorithm
-which produces an error estimate by comparing the 3rd and 2nd order solutions.
-After each step is taken, the errors for each of the nine variables of
-Eqns.~\ref{eqn:num_x}--\ref{eqn:num_fse} are compared to an absolute and
-a relative threshold.
-If the error in any variable exceeds its threshold
-the step size is decreased and the step recomputed;
-if all errors are below threshold, the next step size is increased.
-% see OdeIntegrators::AdaptiveDense
-% and OdeSteppers::RungeKutta3
-% and OdeSteppers::DormandPrince5
-% and OdeController
-% and OdeControllers::ProportionalIntegral_Base::StepSizeToAttempt
-% and OdeErrorMeasures::ScaledAbsRel
+using the 3rd order Runge-Kutta algorithm,
+from which we compute an error estimate by comparing the 3rd and 2nd order solutions.
+After each step is taken, the error estimates $\Delta_i$ for each of the
+nine variables of Eqns.~\ref{eqn:num_x}--\ref{eqn:num_fse} are used
+to compute an overall error measure of the step, $D$, given by
+\begin{equation}
+  \label{eqn:err_measure}
+  D = \max\left(\frac{\Delta_i}{A+R|V_i|}\right),
+\end{equation}
+where the max is taken over the set of all variables $V_i$
+and their error estimates $\Delta_i$,
+and $A$ and $R$ are the absolute and relative error tolerances.
+If $D \geq 1$ the step size is decreased and the step recomputed;
+if $D < 1$ the next step size is increased.
+% The whole integration is run by OdeIntegrator::AdaptiveDense, which contains
+%   some essential pieces: 1) OdeStepper calculates the change in the variables,
+%   2) OdeErrorMeasure calculates a single double representing the overall error
+%   of the step, and 3) OdeController either initiating a step redo or
+%   calculating the next step size.
+% See:
+% * OdeIntegrators::AdaptiveDense
+% * OdeErrorMeasures::ScaledAbsRel
+% * OdeSteppers::RungeKutta3 and OdeSteppers::DormandPrince5
+% * OdeControllers::ProportionalIntegral_Base::StepSizeToAttempt
 % For example, the parameters I use in the nsns data:
-%OdeErrorMeasure = ScaledAbsRel
-%     (
+% ```
+% OdeErrorMeasure = ScaledAbsRel(
 %   ATol = 0.000555555555555556;
@@
 %   ScaledVars = q,p,s0,sn,tau,fa,fs;
-%      Scales = 4,1,4,4,1e-20,1e-20,1e-20;
+%   Scales = 4,1,4,4,1e-20,1e-20,1e-20;)
+%```
 In practice, the controlling errors come from the radiation variables
 $\tau$, $f_{\rm AE}$, and $f_{\rm SE}$,
-for which the relative tolerances are set to $6\times10^{-4}$,
-and the absolute tolerances are set to $6\times10^{-24}$.
+for which the relative tolerances are set to $R=6\times10^{-4}$,
+and the absolute tolerances are set to $A=6\times10^{-24}$.
 
